### 2023-11-08 analysis/country_analysis.Rmd
Safeguard EDA

---
title: "Exploratory Data Analysis: Country"
output: word_document
date: "2023-09-11"
---

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "99%",
  fig.align = "center",
  dpi = 300
)
options(scipen = 999, digits = 4)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(prettyNum(x, big.mark = ","))
    } else{
      return(x)
    }
  }
)
```

```{r dependencies}
library(tidyverse, warn.conflicts = FALSE)
library(cowplot)
```

```{r data_prep}
library(hoaddata)

# ta_oa_inst <- readr::read_csv("https://storage.googleapis.com/hoaddata/ta_oa_inst.csv.gz")
ta_oa_inst <- readr::read_csv(here::here("data/ta_oa_inst.csv.gz"))
# add country codes
ror_country_codes <- readr::read_csv(here::here("data/ror_country_codes.csv"))
# Join
my_df <- inner_join(ta_oa_inst, ror_country_codes, by = "ror_main")

## Articles under TA
ta_df <- my_df |>
  group_by(cr_year, issn_l, country_code, cc) |>
  summarise(ta_inst_articles = n_distinct(doi),
            ta_active_articles = length(unique(doi[ta_active == TRUE])))

## Join with hoaddata affiliation data
aff_ta <- left_join(hoaddata::jn_aff, ta_df, by = c("issn_l", "cr_year", "country_code", "cc")) |>
  # Just TA 
  filter(issn_l %in% hoaddata::jct_hybrid_jns$issn_l, issn_l != "0027-8424")
```

- Proportion of inst per countries with at least one ta

```{r}
aff_country_all <- aff_ta |>
  distinct(issn_l, cr_year, country_code, articles_total) |>
  group_by(cr_year, country_code) |>
  summarise(articles = sum(articles_total, na.rm = TRUE))
aff_ta_all <- aff_ta |>
  group_by(cr_year, country_code) |>
  summarise(ta_inst_articles = sum(ta_inst_articles, na.rm = TRUE))
aff_oa_ta <- aff_ta |> 
  filter(!is.na(cc)) |>
  group_by(cr_year, country_code) |>
  summarise(cc_articles = sum(articles_under_cc_variant, na.rm = TRUE),
            ta_active_oa_articles = sum(ta_active_articles, na.rm = TRUE))
tmp <- aff_country_all |>
  left_join(aff_ta_all) |>
  left_join(aff_oa_ta) |>
  mutate(no_ta_articles = articles - ta_inst_articles, 
         no_ta_oa = cc_articles - ta_active_oa_articles)
```

### plot test

```{r}
library(ggrepel)
p <- tmp |>
  filter(cr_year == 2022, !is.na(country_code), articles > 1000) |>
  ggplot(aes(articles, cc_articles )) +
  geom_point(aes(fill = ta_active_oa_articles / cc_articles), color = "grey50", shape = 21, alpha = 0.8, size = 5) +
 scale_fill_distiller("TA OA", palette = "RdYlBu", direction = 1) +
    geom_smooth(data = filter(tmp, !country_code %in% c("CH", "US")), aes(articles, cc_articles ), method=lm) +
  ggrepel::geom_text_repel(
    data = filter(tmp, cr_year == 2022, articles > 12000),
    aes(x = articles, y = cc_articles, label = country_code),
    force = 50,
    box.padding = 0.5,
    point.padding = 0.8,
    size = 5,
    max.iter = 1000000,
    max.overlaps = Inf,
    min.segment.length = 0,
    arrow = arrow(length = unit(0.010, "npc"))) +
#    scale_y_continuous(labels =  scales::number_format(big.mark = " ")) +
    scale_x_continuous(labels =  scales::number_format(big.mark = " ")) +
  theme_minimal() +
  theme(legend.position="top", legend.justification = "center")

 qq <- tmp |>
  filter(cr_year == 2022, !is.na(country_code), articles > 1000) |>
  ggplot(aes(ta_active_oa_articles, cc_articles)) +
  geom_point(aes(fill = ta_active_oa_articles / cc_articles), color = "grey50", shape = 21, alpha = 0.8, size = 5) +
 scale_fill_distiller("TA OA", palette = "RdYlBu", direction = 1) +
 #   geom_smooth(method=lm) +
  ggrepel::geom_text_repel(
    data = filter(tmp, cr_year == 2022, articles > 12000),
    aes(x = ta_active_oa_articles, y =  cc_articles, label = country_code),
    force = 50,
    box.padding = 0.5,
    point.padding = 0.8,
    size = 5,
    max.iter = 1000000,
    max.overlaps = Inf,
    min.segment.length = 0,
    arrow = arrow(length = unit(0.010, "npc"))) +
#    scale_y_continuous(labels =  scales::number_format(big.mark = " ")) +
    scale_x_continuous(labels =  scales::number_format(big.mark = " ")) +
 #  scale_x_log10() +
  theme_minimal() +
  theme(legend.position="top", legend.justification = "center") 
```

with pie charts

```{r}
# library(ggrepel)
# library(scatterpie)
# df_22 <- tmp |>
#   filter(cr_year == 2022, !is.na(country_code), !is.na(ta_active_oa_articles)) |>
#   mutate(ta_active_oa_articles_ = ta_active_oa_articles,
#          log_articles = log(articles))
# p <- df_22 |>
#   ggplot() +
#   geom_scatterpie(data = df_22, aes(articles, cc_articles, group = country_code), cols=c("ta_active_oa_articles_", "no_ta_oa")) +
#   geom_smooth(aes(articles, cc_articles), method=lm) +
#   #coord_fixed() +
#    ggrepel::geom_text_repel(
#     data = filter(df_22, cc_articles > 2000),
#     aes(x = articles, y = cc_articles, label = country_code),
#     force = 50,
#     box.padding = 1,
#     size = 4,
#     max.iter = 1000,
#     max.overlaps = Inf,
#     min.segment.length = 0,
#     arrow = arrow(length = unit(0.010, "npc"))) 
```

### 2023-11-08 analysis/eda.Rmd
Safeguard EDA

---
title: "Exploratory Data Analysis"
output: word_document
date: "2023-09-11"
---

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "99%",
  fig.align = "center",
  dpi = 300
)
options(scipen = 999, digits = 4)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(prettyNum(x, big.mark = ","))
    } else{
      return(x)
    }
  }
)
```

```{r dependencies}
library(tidyverse, warn.conflicts = FALSE)
library(cowplot)
```

```{r data}
library(hoaddata)

# ta_oa_inst <- readr::read_csv("https://storage.googleapis.com/hoaddata/ta_oa_inst.csv.gz")
ta_oa_inst <- readr::read_csv(here::here("data/ta_oa_inst.csv.gz"))
```

What is the number and proportion of open access articles in hybrid journals published under a transformative agreement?

```{r}
ta_oa_growth <- ta_oa_inst |>
  filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424") |>
  group_by(cr_year) |>
  summarise(ta_oa = n_distinct(doi))
hybrid_oa_growth <- hoaddata::cc_articles |>
  filter(issn_l != "0027-8424") |>
  group_by(cr_year) |>
  summarise(oa_all = n_distinct(doi))
ta_df <- inner_join(ta_oa_growth, hybrid_oa_growth) |>
  mutate(ta_oa_prop = ta_oa / oa_all) |>
  mutate(no_ta = oa_all - ta_oa, 
         no_ta_prop = 1 - ta_oa_prop) |>
  mutate(cr_year = as.factor(cr_year))
# year all
year_all <-
  hoaddata::jn_ind |>
  filter(issn_l != "0027-8424") |>
  distinct(issn_l, cr_year, jn_all) |> 
  group_by(cr_year) |> 
  summarise(n = sum(jn_all))

oa_ta_plot <- ta_df |>
  select(cr_year, ta_oa, no_ta) |>
  pivot_longer(cols = c(ta_oa, no_ta), names_to = "type", values_to = "value") |>
  filter(cr_year != "2023") |>
  inner_join(year_all) |>
  # plot
  ggplot(aes(cr_year, value / n, fill = type, group = type)) +
  geom_area() +
  scale_fill_manual("Hybrid OA", values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"), labels = c(no_ta = "Other", ta_oa = "Agreement"),  guide = guide_legend(reverse = TRUE)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels = scales::percent_format(), limits = c(0, .15), breaks = c(0, .05, .1, .15)) +
  theme_minimal() +
  labs(y = "Hybrid OA Uptake", x = NULL) +
  theme(legend.position="top", legend.justification = "left", axis.text.x = element_blank())
```

```{r}
ta_share_plot <- ta_df |>
  select(cr_year, ta_oa_prop, no_ta_prop) |>
  pivot_longer(cols = c(ta_oa_prop, no_ta_prop), names_to = "type", values_to = "value") |>
  filter(cr_year != "2023", type == "ta_oa_prop") |>
  ggplot(aes(factor(cr_year), value, group = 1)) +
  geom_bar(stat = "identity", fill = "#56B4E9") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), breaks = c(0, 0.25, 0.5),
                     labels = scales::percent) +
  theme_minimal() +
  labs(y = "Percentage\nof Hybrid OA\nvia Agreement", x = "Publication Year")
```

```{r journal_boxplot}
jn_hybrid_oa_growth <- hoaddata::cc_articles |>
  filter(issn_l != "0027-8424") |>
  group_by(cr_year, issn_l) |>
  summarise(oa_all = n_distinct(doi)) |>
  mutate(cr_year = factor(cr_year))
# year all
jn_year_all <-
  hoaddata::jn_ind |>
  filter(issn_l != "0027-8424") |>
  distinct(issn_l, cr_year, jn_all) |> 
  group_by(cr_year, issn_l) |> 
  summarise(n = sum(jn_all))

oa_jn_box <- left_join(jn_hybrid_oa_growth, jn_year_all) |>
  mutate(prop = oa_all / n) |> 
  filter(cr_year != "2023") |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), prop)) +
  geom_boxplot(outlier.shape = NA) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
                     labels = scales::percent) +
  coord_cartesian(ylim = c(0,0.7)) +
  theme_minimal() +
  labs(y = "OA Uptake\nper Hybrid Journal", x = "Publication Year")
```

```{r}
library(patchwork)
design = "AC
BC"
wrap_plots(
  A = oa_ta_plot,
  B = ta_share_plot,
  C = oa_jn_box,
  design = design,
  widths = c(2, 1),
  heights = c(2, 1),
  guides = "collect"
) &
  theme(
    legend.position = "top",
    legend.justification = "left",
    legend.direction = "horizontal",
    panel.grid.minor = element_blank(),
                panel.border = element_rect(color = "grey50", fill = NA),
    axis.title=element_text(size = 10)) &
    plot_annotation(tag_levels = 'A')
```

by publisher

```{r}
jn_ind_jct <- hoaddata::jct_hybrid_jns |>
  inner_join(hoaddata::jn_ind, by = "issn_l") |>
  filter(issn_l != "0027-8424")

ta_active <- ta_oa_inst |>
  filter(issn_l != "0027-8424") |>
  filter(ta_active == TRUE, !is.na(cc)) |>
  group_by(cr_year, issn_l, esac_publisher) |>
  summarise(ta_oa = n_distinct(doi)) |>
  mutate(cr_year = as.character(cr_year))

jn_ind_jct_oa <- jn_ind_jct |>
  filter(!is.na(cc)) |>
  distinct(cr_year, issn_l, esac_publisher, cc, cc_total, jn_all) |>
  group_by(cr_year, issn_l, esac_publisher, jn_all) |>
  summarise(oa = sum(cc_total, na.rm = TRUE))

jn_ind_jct_all <- jn_ind_jct |>
  distinct(cr_year, issn_l, esac_publisher, jn_all)

publisher_df <- left_join(jn_ind_jct_all, jn_ind_jct_oa) |>
  mutate(cr_year = as.character(cr_year)) |>
  left_join(ta_active) |>
  mutate(across(everything(), .fns = ~replace_na(.,0))) |>
  mutate(no_ta = oa - ta_oa)

## plot
publisher_plot_df <- publisher_df |> 
  pivot_longer(cols = c(ta_oa, no_ta), names_to = "type", values_to = "value") |>
  mutate(
    publisher = case_when(
      esac_publisher == "Elsevier" ~ "Elsevier",
      esac_publisher == "Springer Nature" ~ "Springer Nature",
      esac_publisher == "Wiley" ~ "Wiley",
      is.character(esac_publisher) ~ "Other"
    )
  ) |>
  mutate(publisher =
           forcats::fct_relevel(
             publisher,
             c("Elsevier", "Springer Nature", "Wiley", "Other") 
           )) |>
  filter(cr_year != "2023", oa > 0)

#+ publisher ta shares
publisher_ta_prop <- publisher_plot_df |>
  filter(type == "ta_oa") |>
  group_by(cr_year, publisher) |>
  summarize(ta_oa = sum(value), oa = sum(oa), jn_all = sum(jn_all)) |>
  mutate(prop = ta_oa / oa)

oa_publisher <- publisher_plot_df |>
  group_by(cr_year, publisher, type) |>
  summarise(value = sum(value))
all_publisher <- publisher_plot_df |>
  distinct(cr_year, issn_l, publisher, jn_all) |>
  group_by(cr_year, publisher) |>
  summarise(jn_all = sum(jn_all))

publisher_summary_df <- inner_join(oa_publisher, all_publisher) |>
  mutate(prop = value / jn_all)

publisher_prop <- ggplot(publisher_summary_df, aes(gsub("^20", "'", as.character(cr_year)), prop, fill = type, group = type)) +
  geom_area() +
  facet_wrap(~publisher, nrow = 1) +
  scale_fill_manual("Hybrid OA", values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"), labels = c(no_ta = "Other", ta_oa = "Agreement"),  guide = guide_legend(reverse = TRUE)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels = scales::percent_format()) +
  theme_minimal() +
  labs(y = "Hybrid OA Uptake", x = NULL) +
  theme(legend.position="top", legend.justification = "left", axis.text.x = element_blank())

publisher_boxplot <- publisher_plot_df |>
  filter(value > 0) |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), value / jn_all, fill = type)) +
  geom_boxplot(outlier.shape = NA, show.legend = FALSE) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
                     labels = scales::percent) +
  scale_fill_manual("Hybrid OA", values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"), labels = c(no_ta = "Other", ta_oa = "Agreement")) +
  coord_cartesian(ylim = c(0,0.65)) +
  facet_wrap(~publisher, nrow = 1) +
  labs(y = "OA Uptake\nby Hybrid Journal", x = "Publication Year") +
  theme_minimal() +
    theme(strip.background = element_blank(),
    strip.text.x = element_blank()
    )

publisher_ta_prop_plot <- publisher_ta_prop |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), prop)) +
  geom_bar(stat = "identity", fill = "#56B4E9") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), breaks = c(0, 0.25,0.5,0.75), limits = c(0, .75),
                     labels = scales::percent) +
  facet_wrap(~publisher, nrow = 1) +
  theme_minimal() +
  labs(y = "Percentage\nof Hybrid OA\nvia Agreement", x = NULL) +
  theme(
    axis.text.x = element_blank(),
    strip.background = element_blank(),
    strip.text.x = element_blank())
```

```{r, fig.asp=1.1}
library(patchwork)
design = "A
B
C"
wrap_plots(
  A = publisher_prop,
  B = publisher_ta_prop_plot,
  C = publisher_boxplot,
  design = design,
#  guides = "collect",
 # widths = c(2, 1),
  heights = c(1, 0.5, 1)
) &
  theme(
    legend.position = "top",
    legend.justification = "left",
    legend.direction = "horizontal",
    panel.grid.minor = element_blank(),
                panel.border = element_rect(color = "grey50", fill = NA),
    axis.title=element_text(size = 10)) &
    plot_annotation(tag_levels = 'A') 
```

### 2023-11-08 analysis/jn_classification.Rmd
Safeguard EDA

---
title: "journal classification"
output: html_document
date: "2023-09-15"
---

```{r}
library(tidyverse)
library(hoaddata)
```

issn madness

```{r}
my_jns <- hoaddata::jct_hybrid_jns |>
 # filter(issn_l == "0002-8762") |> 
  mutate(id = issn_l) |>
  pivot_longer(cols = c(issn_l, issn)) |>
  distinct(esac_publisher, esac_id, issn_l = id, issn = value) |>
  inner_join(hoaddata::jn_ind, by = "issn_l")

my_jns_with_oa <- my_jns |>
  filter(!is.na(cc))
```

- Journals in JCT with publication activity (`jct`): `r length(unique(my_jns$issn_l))`.
- Journals in JCT with publication activity and at least one OA article in the sample (`jct_oa`): `r length(unique(my_jns_with_oa$issn_l))`

```{r}
scopus <- readxl::read_excel("~/Downloads/extlistAugust2023.xlsx", sheet = 1,
                                .name_repair = janitor::make_clean_names, col_types = "text")
```

normalize scopus jn indicators

```{r}
scopus_norm <- scopus |>
  select(1:4, all_science_journal_classification_codes_asjc) |>
  separate(print_issn, c("print_issn", "print_issn_extra")) |>
  gather(print_issn, e_issn, print_issn_extra, key = "issn_tpye", value = "issn") |> 
  filter(!is.na(issn)) |>
  # trailing zero's missing in Excel spreadsheet
  mutate(
    issn = ifelse(nchar(issn) == 5, paste0("000", issn), issn),
    issn = ifelse(nchar(issn) == 6, paste0("00", issn), issn),
    issn = ifelse(nchar(issn) == 7, paste0("0", issn), issn)) |>
  # missing hyphen
  mutate(issn = map_chr(issn, function(x) paste(c(substr(x, 1, 4), substr(x, 5, 8)), collapse = "-")))
```

Matching with hybrid journals

```{r}
matched_jns <- my_jns |>
  inner_join(scopus_norm, by = "issn")

non_matched <- my_jns |>
  filter(!issn %in% scopus_norm$issn) |>
  distinct(issn_l) |>
  filter(!issn_l %in% matched_jns$issn_l)

matched_oa_jns <- my_jns_with_oa |>
  inner_join(scopus_norm, by = "issn") 
```

- Journals (`jct`) in Scopus: `r length(unique(matched_jns$issn_l))`
- Journals with at least one OA (`jct_oa`) in Scopus: `r length(unique(matched_oa_jns$issn_l))`)

```{r final datatset}
# From https://doi.org/10.1002/asi.24549
ajsc_mapped <- readr::read_csv("https://raw.githubusercontent.com/njahn82/elsevier_hybrid_invoicing/master/data/asjc_mapped.csv", col_types = "ccc") 
subject_jn_df <- matched_jns |>
  select(-issn, -esac_id, -source_title_medline_sourced_journals_are_indicated_in_green, -issn_tpye) |>
  distinct() |>
  mutate(ajcs = strsplit(all_science_journal_classification_codes_asjc, ";")) |>
  unnest(ajcs) |>
  mutate(top_level_code = str_extract(ajcs, "\\d{2}")) |>
  inner_join(ajsc_mapped, by = "top_level_code") |>
  rename(top_level = subject_area, subject_area = description) |> 
  select(-all_science_journal_classification_codes_asjc, -ajcs) |>
  distinct()
# back up
write_csv(subject_jn_df, here::here("data", "jn_scopus_ind_subjects.csv"))
```

### Exploration

#### Journals

##### Subject Area

Number of Journals (full and fractional counting)

```{r}
subject_jn_df |> 
  group_by(issn_l, subject_area) |>
  summarise(n = n()) |>
  mutate(fractions = n / sum(n)) |> 
  ungroup() |> 
  group_by(subject_area) |>
  summarise(journals_n = n_distinct(issn_l),
            journals_frac = sum(fractions)) |>
  arrange(desc(journals_frac))
```

##### Top Level

```{r}
subject_jn_df |> 
  group_by(issn_l, top_level) |>
  summarise(n = n()) |>
  mutate(fractions = n / sum(n)) |> 
  ungroup() |> 
  group_by(top_level) |>
  summarise(journals_n = n_distinct(issn_l),
            journals_frac = sum(fractions)) |>
  arrange(desc(journals_frac))
```

#### Article Analysis

##### Top Level

- Article Volume

```{r}
subject_jn_df |>
  distinct(issn_l, jn_all, top_level) |>
  group_by(issn_l, top_level) |>
  summarise(n = sum(jn_all)) |>
  mutate(fractions_factor = n / sum(n)) |>
  mutate(fractions = n * fractions_factor) |>
  ungroup() |> 
  group_by(top_level) |>
  summarise(articles_n = sum(n),
            articles_frac = sum(fractions)) |>
  arrange(desc(articles_frac)) |>
  arrange(desc(articles_n))
```

- OA by year

```{r}
top_level_by_year_jn_oa <- subject_jn_df |>
  # With at least one cc article
  filter(!is.na(cc)) |>
  distinct(cr_year, issn_l, jn_all, top_level) |>
  group_by(cr_year, issn_l, top_level) |>
  summarise(n = sum(jn_all)) |>
  mutate(fractions_factor = n / sum(n)) |>
  mutate(fractions = n * fractions_factor) |>
  ungroup() |> 
  group_by(cr_year, top_level) |>
  summarise(articles_n = sum(n),
            articles_frac = sum(fractions)) |>
  arrange(top_level, cr_year)

top_level_by_year_oa <- subject_jn_df |>
  # With at least one cc article
  filter(!is.na(cc)) |>
  distinct(cr_year, issn_l, cc_total, top_level) |>
  group_by(cr_year, issn_l, top_level) |>
  summarise(n = sum(cc_total)) |>
  mutate(fractions_factor = n / sum(n)) |>
  mutate(fractions = n * fractions_factor) |>
  ungroup() |> 
  group_by(cr_year, top_level) |>
  summarise(oa_n = sum(n),
            oa_frac = sum(fractions)) |>
  arrange(top_level, cr_year)
```

- Join

```{r}
top_level_oa_df <- inner_join(top_level_by_year_jn_oa, top_level_by_year_oa) |>
  mutate(oa_prop = oa_n / articles_n,
         oa_frac_prop = oa_frac / articles_frac)

top_level_oa_df
```

##### Subject Area

```{r}
subject_jn_df |>
  distinct(issn_l, jn_all, subject_area) |>
  group_by(issn_l, subject_area) |>
  summarise(n = sum(jn_all)) |>
  mutate(fractions_factor = n / sum(n)) |>
  mutate(fractions = n * fractions_factor) |>
  ungroup() |> 
  group_by(subject_area) |>
  summarise(articles_n = sum(n),
            articles_frac = sum(fractions)) |>
  arrange(desc(articles_frac)) |>
  arrange(desc(articles_n))
```

- OA by year

```{r}
subject_area_by_year_jn_oa <- subject_jn_df |>
  # With at least one cc article
  filter(!is.na(cc)) |>
  distinct(cr_year, issn_l, jn_all, subject_area) |>
  group_by(cr_year, issn_l, subject_area) |>
  summarise(n = sum(jn_all)) |>
  mutate(fractions_factor = n / sum(n)) |>
  mutate(fractions = n * fractions_factor) |>
  ungroup() |> 
  group_by(cr_year, subject_area) |>
  summarise(articles_n = sum(n),
            articles_frac = sum(fractions)) |>
  arrange(subject_area, cr_year)

subject_area_by_year_oa <- subject_jn_df |>
  # With at least one cc article
  filter(!is.na(cc)) |>
  distinct(cr_year, issn_l, cc_total, subject_area) |>
  group_by(cr_year, issn_l, subject_area) |>
  summarise(n = sum(cc_total)) |>
  mutate(fractions_factor = n / sum(n)) |>
  mutate(fractions = n * fractions_factor) |>
  ungroup() |> 
  group_by(cr_year, subject_area) |>
  summarise(oa_n = sum(n),
            oa_frac = sum(fractions)) |>
  arrange(subject_area, cr_year)
```

- Join

```{r}
subject_area_oa_df <- inner_join(subject_area_by_year_jn_oa, subject_area_by_year_oa) |>
  mutate(oa_prop = oa_n / articles_n,
         oa_frac_prop = oa_frac / articles_frac)
```

- plot

```{r}
subject_area_oa_df |>
  filter(cr_year %in% c("2020", "2021", "2022")) |>
  group_by(subject_area) |>
  summarise(n = sum(articles_n),
            oa = sum(oa_n)) |>
  mutate(closed = n - oa) |>
  pivot_longer(cols = c(oa, closed)) |>
  arrange(desc(n)) |>
  ggplot(aes(reorder(subject_area, n), value, fill = name)) +
  geom_bar(stat = "identity") +
  coord_flip()
```

##### With TA info

```{r}
ta_oa_inst <- readr::read_csv(here::here("data/ta_oa_inst.csv.gz"))

jn_ind_jct <- hoaddata::jct_hybrid_jns |>
  inner_join(hoaddata::jn_ind, by = "issn_l") |>
  filter(issn_l != "0027-8424")

ta_active <- ta_oa_inst |>
  filter(issn_l != "0027-8424") |>
  filter(ta_active == TRUE, !is.na(cc)) |>
  group_by(cr_year, issn_l) |>
  summarise(ta_oa = n_distinct(doi)) |>
  mutate(cr_year = as.character(cr_year))

subject_per_year_df <- subject_jn_df |>
  distinct(issn_l, top_level) |>
  inner_join(ta_active, by = "issn_l") |>
  group_by(cr_year, top_level) |>
  summarise(ta_oa = sum(ta_oa)) |>
  # join with top level oa stats
  right_join(top_level_oa_df) |> 
  mutate(no_ta = oa_n - ta_oa) |>
  pivot_longer(
    cols = c(ta_oa, no_ta),
    names_to = "type",
    values_to = "value"
  ) |>
  # deal with periods where no ta related pub was found 
  mutate(value = ifelse(is.na(value) & type == "ta_oa", 0, value)) |>
  mutate(value = ifelse(is.na(value) & type == "no_ta", oa_n, value)) |>
  mutate(prop = value / articles_n) 
```

```{r}
plot_trend_top_level_ta <- subject_per_year_df |>
  filter(top_level != "Multidisciplinary", cr_year != "2023") |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), prop, fill = type, group = type)) +
  geom_area() +
  facet_wrap( ~ top_level, nrow = 1) +
  scale_fill_manual(
    "Hybrid OA",
    values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"),
    labels = c(no_ta = "Other", ta_oa = "Agreement"),
    guide = guide_legend(reverse = TRUE)
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    labels = scales::percent_format()
  ) +
  theme_minimal() +
  labs(y = "Hybrid OA Uptake", x = NULL) +
  theme(legend.position = "top", legend.justification = "left")
```

```{r}
# Share of TA
plot_trend_top_level_ta_share <- subject_per_year_df |>
  filter(top_level != "Multidisciplinary", cr_year != "2023") |>
  filter(type == "ta_oa") |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), value / oa_n, group = 1)) +
  geom_bar(stat = "identity", fill = "#56B4E9") +
  facet_wrap( ~ top_level, nrow = 1) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
                     labels = scales::percent) +
  theme_minimal() +
  labs(y = "Percentage\nof Hybrid OA\nvia Agreement", x = NULL)
```

```{r}
## TA boxplot
plot_trend_top_level_ta_boxplot <- subject_ta_df |>
  pivot_longer(cols = c(ta_oa, no_ta), names_to = "type", values_to = "value") |>
  filter(cr_year != "2023", oa > 0, value > 0, top_level != "Multidisciplinary") |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), value / jn_all, fill = type)) +
  geom_boxplot(outlier.shape = NA, show.legend = FALSE) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
                     labels = scales::percent) +
  scale_fill_manual("Hybrid OA", values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"), labels = c(no_ta = "Other", ta_oa = "Agreement")) +
  coord_cartesian(ylim = c(0,0.55)) +
  facet_wrap(~top_level, nrow = 1) +
  labs(y = "OA Uptake\nby Hybrid Journal", x = "Publication Year") +
  theme_minimal() +
    theme(strip.background = element_blank()
    ) 
```

```{r, fig.asp=1.1}
library(patchwork)
design = "A
B
C"
wrap_plots(
  A = plot_trend_top_level_ta,
  B = plot_trend_top_level_ta_share,
  C = plot_trend_top_level_ta_boxplot,
  design = design,
#  guides = "collect",
 # widths = c(2, 1),
  heights = c(1, 0.5, 1)
) &
  theme(
    legend.position = "top",
    legend.justification = "left",
    legend.direction = "horizontal",
    panel.grid.minor = element_blank(),
                panel.border = element_rect(color = "grey50", fill = NA),
    axis.title=element_text(size = 10)) &
    plot_annotation(tag_levels = 'A') 
```

TA analysis

```{r}
subject_ta_df |>
  pivot_longer(cols = c(ta_oa, no_ta), names_to = "type", values_to = "value") |>
  filter(cr_year != "2023", oa > 0, top_level != "Multidisciplinary") |>
   ggplot(aes(gsub("^20", "'", as.character(cr_year)), value / jn_all, fill = type)) +
  geom_boxplot(outlier.shape = NA, show.legend = FALSE) +
  coord_flip() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
                     labels = scales::percent) +
  scale_fill_manual("Hybrid OA", values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"), labels = c(no_ta = "Other", ta_oa = "Agreement")) +
  coord_cartesian(ylim = c(0,0.55)) +
  facet_wrap(~subject_area) +
  labs(y = "OA Uptake\nby Hybrid Journal", x = "Publication Year") +
  theme_minimal() +
  theme(strip.background = element_blank())
```

### 2023-11-08 data/000_data_gathering.R

@@ -1,0 +1,67 @@
+#' # OVERVIEW: Data used
+#' 
+#' ## Data
+#' 
+#' *{hoaddata} version v0.2.82* provides journal-level and article-level data used.
+#' 
+#' <https://github.com/subugoe/hoaddata/releases/tag/v0.2.82>
+#' 
+#' This R package contains: 
+#' 
+#' *Journal-level data*, retrieved from the cOAlition S public transformative
+#' agreement data on 8 May 2023
+#' 
+#' <https://github.com/njahn82/jct_data/tree/4a017bba5ea72e86840d701420afc9d922e00266/data>
+#' 
+#' The code repository also contains code used to obtain the data.
+#' 
+#' {hoaddata} also represents *article-level data* retrieved from
+#' 
+#'  - Crossref release 2023/10
+#'  - OpenAlex release 2023-10-1
+#'  
+#'  as well as the code used to compile the data from SUB Göttingen Scholarly
+#'  data warehouse.
+#'  
+#'  The main dataset, providing article-level data about publications linked to 
+#'  transformative agreements and institutions, is available as {hoaddata} 
+#'  release asset: 
+#'  <https://github.com/subugoe/hoaddata/releases/download/v0.2.82/ta_oa_inst.csv.gz>
+#'  
+#' *Raw data for reproducibility*: To improve reproducibility, 
+#' the underying raw data used to compile {hoaddata} 
+#' version v0.2.82 is available via Google BigQuery `hoa-article.hoaddata_nov23`.
+#' 
+#' *Subject classification* data was obtained from Scopus on 4 Oct 2023 and is
+#' stored in `data/jn_scopus_ind_subjects.csv`
+#' 
+#' *Country information* for institutions with transformative agreements were 
+#' obtained from OpenAlex in the following
+library(bigrquery)
+library(tidyverse)
+
+my_sql <-  "SELECT
+  DISTINCT main.ror_main,
+  country_code
+FROM
+  `hoa-article.hoaddata_nov23.ta_oa_inst` AS main
+INNER JOIN
+  `subugoe-collaborative.openalex.institutions` AS oalex
+ON
+  ror_main = oalex.ror"
+
+tb <- bq_project_query("subugoe-collaborative", my_sql)
+ror_country_codes_raw <- bq_table_download(tb)
+
+ror_country_codes <- ror_country_codes_raw |>
+  # Fix missing country codes
[…]

### 2023-11-08 analysis/001_overview.Rmd
Add ta/no ta boxplot in overview

--- 2023-11-08 analysis/eda.Rmd
@@
 ---
-title: "Exploratory Data Analysis"
-output: word_document
-date: "2023-09-11"
+title: "Overview"
+output: github_document
+date: "`r Sys.Date()`"
 ---
@@
   message = FALSE,
-  echo = FALSE,
+  echo = TRUE,
   fig.width = 6,
@@
 library(tidyverse, warn.conflicts = FALSE)
-library(cowplot)
 ```
 
-```{r data}
+```{r data, cache=TRUE}
 library(hoaddata)
 
-# ta_oa_inst <- readr::read_csv("https://storage.googleapis.com/hoaddata/ta_oa_inst.csv.gz")
-ta_oa_inst <- readr::read_csv(here::here("data/ta_oa_inst.csv.gz"))
+ta_oa_inst <- readr::read_csv("https://github.com/subugoe/hoaddata/releases/download/v0.2.82/ta_oa_inst.csv.gz")
 ```
@@
 
-```{r}
+```{r area_oa_over_the_years}
 ta_oa_growth <- ta_oa_inst |>
@@
 
-```{r}
+```{r ta_share_plot}
 ta_share_plot <- ta_df |>
@@
 ```{r}
+ta_oa_growth <- ta_oa_inst |>
+  filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424") |>
+  group_by(cr_year, issn_l) |>
+  summarise(ta_oa = n_distinct(doi))
+hybrid_oa_growth <- hoaddata::cc_articles |>
+  filter(issn_l != "0027-8424") |>
+  group_by(cr_year, issn_l) |>
+  summarise(oa_all = n_distinct(doi))
+ta_df <- left_join(hybrid_oa_growth, ta_oa_growth) |>
+  mutate(across(everything(), .fns = ~replace_na(.,0))) |>
+  mutate(no_ta = oa_all - ta_oa) |>
+  mutate(cr_year = as.factor(cr_year))
+
+summary_all <- inner_join(ta_df, jn_year_all, by = c("cr_year", "issn_l")) |>
+  mutate(prop_ta_oa = ta_oa / n,
+         prop_no_ta = no_ta / n) |>
+  ungroup() |>
+  distinct(cr_year, issn_l, prop_ta_oa, prop_no_ta) |>
+  pivot_longer(cols = c(prop_ta_oa, prop_no_ta))
+# plot
+box_ta_prop <- summary_all |>
+  filter(cr_year != "2023") |>
+  # With at least one OA article
+  filter(value > 0) |>
+  ggplot(aes(gsub("^20", "'", as.character(cr_year)), value, fill = name)) +
+  geom_boxplot(outlier.shape = NA) +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
+                     labels = scales::percent) +
+  coord_cartesian(ylim = c(0,0.5)) +
+  scale_fill_manual("Hybrid OA", values = c(prop_no_ta = "#b3b3b3a0", prop_ta_oa = "#56B4E9"), labels = c(prop_no_ta = "Other", prop_ta_oa = "Agreement")) +
+  theme_minimal() +
+  labs(y = "OA Uptake\nper Hybrid Journal", x = "Publication Year")
+
+```
+
+```{r}
 library(patchwork)
 design = "AC
-BC"
+BD"
 wrap_plots(
@@
   C = oa_jn_box,
+  D = box_ta_prop,
   design = design,

### 2023-11-21 analysis/001_overview.Rmd
Start writing Overview

@@
   message = FALSE,
-  echo = TRUE,
+  echo = FALSE,
   fig.width = 6,
@@
 )
-options(scipen = 999, digits = 4)
+options(scipen = 999, digits = 2)
 knitr::knit_hooks$set(
@@
 
-ta_oa_inst <- readr::read_csv("https://github.com/subugoe/hoaddata/releases/download/v0.2.82/ta_oa_inst.csv.gz")
+ta_oa_inst <- readr::read_csv(here::here("data", "ta_oa_inst.csv.gz"))
 ```
 
-What is the number and proportion of open access articles in hybrid journals published under a transformative agreement?
+What are the number and proportion of open access articles in hybrid journals published under a transformative agreement?
 
-```{r area_oa_over_the_years}
+```{r overview_data_prep}
 ta_oa_growth <- ta_oa_inst |>
-  filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424") |>
+  filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424", cr_year != "2023") |>
   group_by(cr_year) |>
@@
 hybrid_oa_growth <- hoaddata::cc_articles |>
-  filter(issn_l != "0027-8424") |>
+  filter(issn_l != "0027-8424", cr_year != "2023") |>
   group_by(cr_year) |>
@@
   hoaddata::jn_ind |>
-  filter(issn_l != "0027-8424") |>
+  filter(issn_l != "0027-8424", cr_year != "2023") |>
   distinct(issn_l, cr_year, jn_all) |> 
@@
 
+## Hybrid Journals
+ n_hybrid_jns_all <- hoaddata::jn_ind |>
+  filter(issn_l != "0027-8424", cr_year != "2023") |>
+  distinct(issn_l) |>
+   nrow()
+
+## Hybrid Journals with at least one OA article
+n_hybrid_jns <- hoaddata::cc_articles |>
+  filter(issn_l != "0027-8424", cr_year != "2023") |>
+  group_by(issn_l) |>
+  summarise(oa_all = n_distinct(doi)) |> 
+  nrow()
+
+## Hybrid Journals with at least one OA article
+n_ta_jns <- ta_oa_inst |>
+  filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424", cr_year != "2023") |>
+  distinct(issn_l) |>
+  nrow()
+
+## Hybrid Journals with at least one TA-OA article
+n_tas_with_oa <- ta_oa_inst |>
+  filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424", cr_year != "2023") |>
+  distinct(ta_journal_portfolio) |>
+  nrow()
+
+## Open access volume
+n_oa <-  hoaddata::cc_articles |>
+  filter(issn_l != "0027-8424", cr_year != "2023") |>
+  distinct(doi) |>
+  nrow()
+
+## Total
+n_total <- hoaddata::jn_ind |>
+  filter(issn_l != "0027-8424", cr_year != "2023", !is.na(cc)) |>
+  distinct(issn_l, cr_year, jn_all) |>
+  pull(jn_all) |>
+  sum()
+
+## TA total
+n_tas_with_oa <- ta_oa_inst |>
+  filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424", cr_year != "2023") |>
+  distinct(ta_journal_portfolio) |>
+  nrow()
+
+## TA article total
+n_ta_oa <- ta_oa_inst |>
+  filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424", cr_year != "2023") |>
+  distinct(doi) |>
+  nrow()
+```
+
+Between 2017 and 2022, a total of `r n_hybrid_jns` out of `r n_hybrid_jns_all` hybrid journals in transformative agreements provided at least one open access article under a Creative Commons license.
+During this period, `r n_oa` out of `r n_total` articles in these journals became open access, representing a six-year hybrid open access uptake rate of `r n_oa / n_total * 100`%.
+First authors affiliated with institutions that had a transformative agreement in place at the time of publication, which indicates that they could make use of such agreements, contributed `r n_ta_oa` open access articles to the total.
+
+```{r area_oa_over_the_years}
 oa_ta_plot <- ta_df |>
@@
   filter(cr_year != "2023", type == "ta_oa_prop") |>
-  ggplot(aes(factor(cr_year), value, group = 1)) +
+  ggplot(aes(gsub("^20", "'", as.character(cr_year)), value, group = 1)) +
   geom_bar(stat = "identity", fill = "#56B4E9") +
@@
                      labels = scales::percent) +
-  coord_cartesian(ylim = c(0,0.7)) +
+  coord_cartesian(ylim = c(0,0.65)) +
   theme_minimal() +
-  labs(y = "OA Uptake\nper Hybrid Journal", x = "Publication Year")
+  labs(y = "OA Uptake per Hybrid Journal", x = "Publication Year")
 ```
@@
                      labels = scales::percent) +
-  coord_cartesian(ylim = c(0,0.5)) +
-  scale_fill_manual("Hybrid OA", values = c(prop_no_ta = "#b3b3b3a0", prop_ta_oa = "#56B4E9"), labels = c(prop_no_ta = "Other", prop_ta_oa = "Agreement")) +
+  coord_cartesian(ylim = c(0,0.65)) +
+  scale_fill_manual("Hybrid OA", values = c(prop_no_ta = "#b3b3b3a0", prop_ta_oa = "#56B4E9"), labels = c(prop_no_ta = "Other", prop_ta_oa = "Agreement"), guide = FALSE) +
   theme_minimal() +
-  labs(y = "OA Uptake\nper Hybrid Journal", x = "Publication Year")
+  labs(y = NULL, x = "Publication Year")
 
@@
 library(patchwork)
-design = "AC
-BD"
+layout = "
+EEE
+ACD
+BCD"
 wrap_plots(
+  E = guide_area(),
   A = oa_ta_plot,
@@
   D = box_ta_prop,
-  design = design,
-  widths = c(2, 1),
-  heights = c(2, 1),
-  guides = "collect"
-) &
+  design = layout,
+  widths = c(0.4, 0.3, 0.3),
+  heights = c(0.2, 1, .5),
+  guides = "collect") &
   theme(
     legend.position = "top",
-    legend.justification = "left",
+    legend.justification = "right",
     legend.direction = "horizontal",
@@
 
-by publisher
+Caption:
 
@@
   design = design,
-#  guides = "collect",
+  guides = "collect",
  # widths = c(2, 1),

### 2023-11-21 data/000_data_gathering.R

@@ -18,7 +18,7 @@
 #' {hoaddata} also represents *article-level data* retrieved from
 #' 
 #'  - Crossref release 2023/10
-#'  - OpenAlex release 2023-10-1
+#'  - OpenAlex release 2023-10-18
 #'  
 #'  as well as the code used to compile the data from SUB Göttingen Scholarly
 #'  data warehouse.
@@ -28,6 +28,9 @@
 #'  release asset: 
 #'  <https://github.com/subugoe/hoaddata/releases/download/v0.2.82/ta_oa_inst.csv.gz>
 #'  
+#'  A version is provided in `data/` folder
+download.file("https://github.com/subugoe/hoaddata/releases/download/v0.2.82/ta_oa_inst.csv.gz",
+              destfile = here::here("data", "ta_oa_inst.csv.gz")) 
 #' *Raw data for reproducibility*: To improve reproducibility, 
 #' the underying raw data used to compile {hoaddata} 
 #' version v0.2.82 is available via Google BigQuery `hoa-article.hoaddata_nov23`.

### 2023-11-21 analysis/001_overview.Rmd
Analyse five-years period

@@
 ta_oa_growth <- ta_oa_inst |>
-  filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424", cr_year != "2023") |>
+  filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
   group_by(cr_year) |>
@@
 hybrid_oa_growth <- hoaddata::cc_articles |>
-  filter(issn_l != "0027-8424", cr_year != "2023") |>
+  filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
   group_by(cr_year) |>
@@
   hoaddata::jn_ind |>
-  filter(issn_l != "0027-8424", cr_year != "2023") |>
+  filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
   distinct(issn_l, cr_year, jn_all) |> 
@@
  n_hybrid_jns_all <- hoaddata::jn_ind |>
-  filter(issn_l != "0027-8424", cr_year != "2023") |>
+  filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
   distinct(issn_l) |>
@@
 n_hybrid_jns <- hoaddata::cc_articles |>
-  filter(issn_l != "0027-8424", cr_year != "2023") |>
+  filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
   group_by(issn_l) |>
@@
 n_ta_jns <- ta_oa_inst |>
-  filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424", cr_year != "2023") |>
+  filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
   distinct(issn_l) |>
@@
 n_tas_with_oa <- ta_oa_inst |>
-  filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424", cr_year != "2023") |>
+  filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
   distinct(ta_journal_portfolio) |>
@@
 n_oa <-  hoaddata::cc_articles |>
-  filter(issn_l != "0027-8424", cr_year != "2023") |>
+  filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
   distinct(doi) |>
@@
 n_total <- hoaddata::jn_ind |>
-  filter(issn_l != "0027-8424", cr_year != "2023", !is.na(cc)) |>
+  filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023"), !is.na(cc)) |>
   distinct(issn_l, cr_year, jn_all) |>
@@
 n_ta_oa <- ta_oa_inst |>
-  filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424", cr_year != "2023") |>
+  filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
   distinct(doi) |>
@@
 Between 2017 and 2022, a total of `r n_hybrid_jns` out of `r n_hybrid_jns_all` hybrid journals in transformative agreements provided at least one open access article under a Creative Commons license.
-During this period, `r n_oa` out of `r n_total` articles in these journals became open access, representing a six-year hybrid open access uptake rate of `r n_oa / n_total * 100`%.
+During this period, `r n_oa` out of `r n_total` articles in these journals became open access, representing a five-year hybrid open access uptake rate of `r n_oa / n_total * 100`%.
 First authors affiliated with institutions that had a transformative agreement in place at the time of publication, which indicates that they could make use of such agreements, contributed `r n_ta_oa` open access articles to the total.
@@
   pivot_longer(cols = c(ta_oa, no_ta), names_to = "type", values_to = "value") |>
-  filter(cr_year != "2023") |>
+  filter(!cr_year %in%  c("2017", "2023")) |>
   inner_join(year_all) |>
@@
   pivot_longer(cols = c(ta_oa_prop, no_ta_prop), names_to = "type", values_to = "value") |>
-  filter(cr_year != "2023", type == "ta_oa_prop") |>
+  filter(!cr_year %in%  c("2017", "2023"), type == "ta_oa_prop") |>
   ggplot(aes(gsub("^20", "'", as.character(cr_year)), value, group = 1)) +
@@
   mutate(prop = oa_all / n) |> 
-  filter(cr_year != "2023") |>
+  filter(!cr_year %in%  c("2017", "2023")) |>
   ggplot(aes(gsub("^20", "'", as.character(cr_year)), prop)) +
@@
 box_ta_prop <- summary_all |>
-  filter(cr_year != "2023") |>
+  filter(!cr_year %in%  c("2017", "2023")) |>
   # With at least one OA article
@@
            )) |>
-  filter(cr_year != "2023", oa > 0)
+  filter(!cr_year %in%  c("2017", "2023"), oa > 0)
 

### 2023-11-22 analysis/001_overview.Rmd
Fig caption

@@
   filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
+  #   filter(issn_l %in% hoaddata::cc_articles$issn_l) |>
+  filter(issn_l %in% hoaddata::cc_articles$issn_l) |>
   distinct(issn_l, cr_year, jn_all) |> 
@@
 Between 2017 and 2022, a total of `r n_hybrid_jns` out of `r n_hybrid_jns_all` hybrid journals in transformative agreements provided at least one open access article under a Creative Commons license.
-During this period, `r n_oa` out of `r n_total` articles in these journals became open access, representing a five-year hybrid open access uptake rate of `r n_oa / n_total * 100`%.
+During this period, these hybrid journals provided open access to `r n_oa` out of `r n_total` articles, representing a five-year hybrid open access uptake rate of `r n_oa / n_total * 100`%.
 First authors affiliated with institutions that had a transformative agreement in place at the time of publication, which indicates that they could make use of such agreements, contributed `r n_ta_oa` open access articles to the total.
@@
   geom_area() +
-  scale_fill_manual("Hybrid OA", values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"), labels = c(no_ta = "Other", ta_oa = "Agreement"),  guide = guide_legend(reverse = TRUE)) +
+  scale_fill_manual(NULL, values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"), labels = c(no_ta = "Other", ta_oa = "Agreement"),  guide = guide_legend(reverse = TRUE)) +
   scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels = scales::percent_format(), limits = c(0, .15), breaks = c(0, .05, .1, .15)) +
@@
   coord_cartesian(ylim = c(0,0.65)) +
-  scale_fill_manual("Hybrid OA", values = c(prop_no_ta = "#b3b3b3a0", prop_ta_oa = "#56B4E9"), labels = c(prop_no_ta = "Other", prop_ta_oa = "Agreement"), guide = FALSE) +
+  scale_fill_manual(NULL, values = c(prop_no_ta = "#b3b3b3a0", prop_ta_oa = "#56B4E9"), labels = c(prop_no_ta = "Other", prop_ta_oa = "Agreement"), guide = FALSE) +
   theme_minimal() +
@@
 ```{r}
+#| fig-cap: "Open access uptake in hybrid journals through transformative agreements 2018-22 per publication year.
+The blue areas represent open access through transformative agreements, as indicated by the affiliation with institutions that had a corresponding transformative agreement in place at the time of publication (according to the first author affiliation data in OpenAlex matched with journal and institutional data from cOAlition S Transformative Agreements Public Data).
+(A) Proportion of open access articles in hybrid journals per year (according to the total number of articles indexed in Crossref).
+(B) Percentage of hybrid open access via agreements per year.
+Boxplots show the proportion of open access articles by individual hybrid journals (C) and individual open access uptake rates by individual hybrid journals and open access funding (D) per publication year.
+Horizontal lines denote lower quartile, median, and upper quartile, with whiskers extending to 1.5 × IQR.
+The individual outliers are not shown.
+Note that data on transformative agreements ending before June 2021 were not available at the time of this study."
+
 library(patchwork)
@@
 ```
-
-Caption:
 

### 2023-11-27 analysis/001_overview.Rmd
Update results

@@
 
-Between 2017 and 2022, a total of `r n_hybrid_jns` out of `r n_hybrid_jns_all` hybrid journals in transformative agreements provided at least one open access article under a Creative Commons license.
+Between 2018 and 2022, a total of `r n_hybrid_jns` out of `r n_hybrid_jns_all` hybrid journals in transformative agreements published at least one open access article under a Creative Commons license.
 During this period, these hybrid journals provided open access to `r n_oa` out of `r n_total` articles, representing a five-year hybrid open access uptake rate of `r n_oa / n_total * 100`%.
@@
 ```{r area_oa_over_the_years}
-oa_ta_plot <- ta_df |>
+oa_ta_plot_df <- ta_df |>
   select(cr_year, ta_oa, no_ta) |>
@@
   filter(!cr_year %in%  c("2017", "2023")) |>
-  inner_join(year_all) |>
+  inner_join(year_all)
   # plot
-  ggplot(aes(cr_year, value / n, fill = type, group = type)) +
+oa_ta_plot <- ggplot(oa_ta_plot_df, aes(cr_year, value / n, fill = type, group = type)) +
   geom_area() +
@@
 
-oa_jn_box <- left_join(jn_hybrid_oa_growth, jn_year_all) |>
+oa_jn_box_df <- left_join(jn_hybrid_oa_growth, jn_year_all) |>
   mutate(prop = oa_all / n) |> 
-  filter(!cr_year %in%  c("2017", "2023")) |>
+  filter(!cr_year %in%  c("2017", "2023"))
+oa_jn_box <- oa_jn_box_df |>
   ggplot(aes(gsub("^20", "'", as.character(cr_year)), prop)) +
@@
   ggplot(aes(gsub("^20", "'", as.character(cr_year)), value, fill = name)) +
-  geom_boxplot(outlier.shape = NA) +
+  geom_boxplot(outlier.shape = NA, varwidth = TRUE) +
   scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
@@
 ```{r}
-#| fig-cap: "Open access uptake in hybrid journals through transformative agreements 2018-22 per publication year.
+#| fig-cap: "Growth of open access in hybrid journals through transformative agreements between 2018 and 2022 per publication year.
 The blue areas represent open access through transformative agreements, as indicated by the affiliation with institutions that had a corresponding transformative agreement in place at the time of publication (according to the first author affiliation data in OpenAlex matched with journal and institutional data from cOAlition S Transformative Agreements Public Data).
@@
 Boxplots show the proportion of open access articles by individual hybrid journals (C) and individual open access uptake rates by individual hybrid journals and open access funding (D) per publication year.
-Horizontal lines denote lower quartile, median, and upper quartile, with whiskers extending to 1.5 × IQR.
+Horizontal lines denote lower quartile, median, and upper quartile, with whiskers extending to ±1.5 of interquartile range.
 The individual outliers are not shown.
@@
 ```
+
+```{r}
+# in text data
+oa_by_year <- oa_ta_plot_df |>
+  group_by(cr_year) |>
+  summarise(oa = sum(value),
+            articles = n) |>
+  mutate(prop = oa / articles * 100) |>
+  distinct()
+```
+
+Figure 2A shows a moderate growth in the proportion of open access articles in hybrid journals, comparing the overall uptake rate and the impact of transformative agreements on this trend.
+Over the five-years period from 2018 to 2022, the overall open access uptake rate increased from `r oa_by_year |> filter(cr_year == "2018") |> pull(prop)`% (n = `r oa_by_year |> filter(cr_year == "2018") |> pull(oa)`) to `r oa_by_year |> filter(cr_year == "2022") |> pull(prop)`% (n = `r oa_by_year |> filter(cr_year == "2022") |> pull(oa)`).
+At the same time, the total publication volume of the investigated journals grew from `r oa_by_year |> filter(cr_year == "2018") |> pull(articles)` in 2018 to `r oa_by_year |> filter(cr_year == "2022") |> pull(articles)` in 2022.
+
+Figure 2B highlights that the majority of open access articles in hybrid journals were made available through transformative agreements in 2021 and 2022, contributing `r oa_ta_plot_df |> filter(cr_year == "2022", type == "ta_oa") |> pull(value) / oa_by_year |> filter(cr_year == "2022") |> pull(oa) * 100`% of the total open access volume in 2022.
+However, there was also a notable growth in open access provision through individual publication fees, which increased from `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> pull(value)`) in 2018 to  `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> pull(value)`).
+This suggests that publishers were able to gain equally from individual and institutional open access publishing options.
+
+Figure 2C depicts the substantial variations among the hybrid journals included in transformative agreements in terms of open access uptake.
+Although the median generally follows the trend shown in Figure 2A, the farther stretch of upper quartiles and whiskers over the years illustrates that an increasing number of journals published an above-average proportion of open access articles.
+In 2022, 25% of hybrid journals (n = `r oa_jn_box_df |> filter(cr_year == "2022") |> mutate(top25 = quantile(prop, .75)) |> filter(prop >= top25) |> nrow()`) had an open access uptake of `r oa_jn_box_df |> filter(cr_year == "2022") |> pull(prop) |> quantile(0.75)*100`%, and `r (oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> nrow()) / length(unique(oa_jn_box_df$issn_l)) *100`% of journals (n = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> nrow()`) provided the majority of their articles under a Creative Commons license in the same year.
+These journals were, on average, smaller (M = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> pull(n) |> mean()`, SD = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> pull(n) |> sd()`) than those with an open access share below 50% (M = `r oa_jn_box_df |> filter(cr_year == "2022", prop <= 0.5) |> pull(n) |> mean()`, SD = `r oa_jn_box_df |> filter(cr_year == "2022", prop <= 0.5) |> pull(n) |> sd()`).
+Notable exception of large journals with an above-average open access proportion were Physical Review D, a high-energy physics journal covered by the SCOAP3 consortium that provided open access to `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "2470-0010") |> pull(oa_all)` out of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "2470-0010") |> pull(n)` articles in 2022, Astronomy and Physics (`r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0004-6361") |> pull(oa_all)` out of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0004-6361") |> pull(n)` articles in 2022 were open access), which shifted to a subscribe to open business model for all accepted articles as of April 2022, the Journal of Fluid Mechanics (`r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0022-1120") |> pull(oa_all)` of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0022-1120") |> pull(n)` articles in 2022 were open access) and Bioinformatics (`r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "1367-4803") |> pull(oa_all)` out of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "1367-4803") |> pull(n)` articles in 2022 were open access), which flipped to full open access as of January 2023.
+
+When comparing the impact of open access trough transfromative agreements across journals, it shows that larger number of journals
 

### 2023-11-28 analysis/001_overview.Rmd
Finish firstdraft figure 2 interpretation

@@
 
-When comparing the impact of open access trough transfromative agreements across journals, it shows that larger number of journals
+When comparing the impact of open access trough transformative agreements across journals, it shows that for many journals these agreements have been a significant contributor to the growth of open access over the years (Figure 2D).
+Examples of such journals include those with a scope on specific countries or regions, where also transformative agreements were implemented. 
+For instance, in 2022, the Germany-based journals Zeitschrift für Erziehungswissenschaft and Zeitschrift für Politikwissenschaft, as well as the Scandinavian Political Studies adressing the Nordic countries, achieved an overall open access uptake of more than 90% just through transformative agreements.
+Despite the rise in transformative agreements, it is worth noting that other means of publishing open access in hybrid journals remained common. 
+In total, `r summary_all |> filter(cr_year == "2022", value > 0, name == "prop_no_ta") |> nrow()` journals published open access articles from authors affiliated with institutions without transformative agreements in place, while `r summary_all |> filter(cr_year == "2022", value > 0, name == "prop_ta_oa") |> nrow()` journals published at least one open access article through a transformative agreement in the same year.
 

### 2023-12-02 analysis/001_overview.Rmd
Journal investigated: hybrid journals with at least one oa article between 2018 - 2022 (check 0001-4338)

@@
 output: github_document
+always_allow_html: true
 date: "`r Sys.Date()`"
@@
 library(tidyverse, warn.conflicts = FALSE)
+library(kableExtra)
 ```
@@
   filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
-  #   filter(issn_l %in% hoaddata::cc_articles$issn_l) |>
   filter(issn_l %in% hoaddata::cc_articles$issn_l) |>
@@
 n_total <- hoaddata::jn_ind |>
-  filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023"), !is.na(cc)) |>
+  filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
+  filter(issn_l %in% hoaddata::cc_articles$issn_l) |>
   distinct(issn_l, cr_year, jn_all) |>
@@
 Between 2018 and 2022, a total of `r n_hybrid_jns` out of `r n_hybrid_jns_all` hybrid journals in transformative agreements published at least one open access article under a Creative Commons license.
-During this period, these hybrid journals provided open access to `r n_oa` out of `r n_total` articles, representing a five-year hybrid open access uptake rate of `r n_oa / n_total * 100`%.
+During this period, these hybrid journals provided open access to `r n_oa` out of `r n_total` articles, representing a five-year hybrid open access uptake of `r n_oa / n_total * 100`%.
 First authors affiliated with institutions that had a transformative agreement in place at the time of publication, which indicates that they could make use of such agreements, contributed `r n_ta_oa` open access articles to the total.
@@
   filter(issn_l != "0027-8424") |>
+  filter(issn_l %in% hoaddata::cc_articles$issn_l) |>
   distinct(issn_l, cr_year, jn_all) |> 
@@
 
-Figure 2A shows a moderate growth in the proportion of open access articles in hybrid journals, comparing the overall uptake rate and the impact of transformative agreements on this trend.
-Over the five-years period from 2018 to 2022, the overall open access uptake rate increased from `r oa_by_year |> filter(cr_year == "2018") |> pull(prop)`% (n = `r oa_by_year |> filter(cr_year == "2018") |> pull(oa)`) to `r oa_by_year |> filter(cr_year == "2022") |> pull(prop)`% (n = `r oa_by_year |> filter(cr_year == "2022") |> pull(oa)`).
-At the same time, the total publication volume of the investigated journals grew from `r oa_by_year |> filter(cr_year == "2018") |> pull(articles)` in 2018 to `r oa_by_year |> filter(cr_year == "2022") |> pull(articles)` in 2022.
+Figure 2A shows a moderate growth in the proportion of open access articles in hybrid journals, comparing the overall open access uptake and the impact of transformative agreements on this trend.
+Over the five-years period from 2018 to 2022, open access increased from `r oa_by_year |> filter(cr_year == "2018") |> pull(prop)`% (n = `r oa_by_year |> filter(cr_year == "2018") |> pull(oa)`) to `r oa_by_year |> filter(cr_year == "2022") |> pull(prop)`% (n = `r oa_by_year |> filter(cr_year == "2022") |> pull(oa)`).
+At the same time, the total article volume of the investigated journals grew from `r oa_by_year |> filter(cr_year == "2018") |> pull(articles)` in 2018 to `r oa_by_year |> filter(cr_year == "2022") |> pull(articles)` in 2022.
 
-Figure 2B highlights that the majority of open access articles in hybrid journals were made available through transformative agreements in 2021 and 2022, contributing `r oa_ta_plot_df |> filter(cr_year == "2022", type == "ta_oa") |> pull(value) / oa_by_year |> filter(cr_year == "2022") |> pull(oa) * 100`% of the total open access volume in 2022.
+Figure 2B highlights that the majority of open access articles in hybrid journals were made available through transformative agreements in 2021 and 2022, contributing `r oa_ta_plot_df |> filter(cr_year == "2022", type == "ta_oa") |> pull(value) / oa_by_year |> filter(cr_year == "2022") |> pull(oa) * 100`% of the total open access article volume in 2022.
 However, there was also a notable growth in open access provision through individual publication fees, which increased from `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> pull(value)`) in 2018 to  `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> pull(value)`).
@@
 jn_ind_jct_all <- jn_ind_jct |>
+  filter(issn_l %in% hoaddata::cc_articles$issn_l) |>
   distinct(cr_year, issn_l, esac_publisher, jn_all)
@@
            )) |>
-  filter(!cr_year %in%  c("2017", "2023"), oa > 0)
+  filter(!cr_year %in%  c("2017", "2023"))
 
@@
   ggplot(aes(gsub("^20", "'", as.character(cr_year)), value / jn_all, fill = type)) +
-  geom_boxplot(outlier.shape = NA, show.legend = FALSE) +
+  geom_boxplot(outlier.shape = NA, show.legend = FALSE, varwidth = TRUE) +
   scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
@@
 
+Analysing hybrid open access across publishers between 2018 and 2022 reveals a large concentration on a few publishers in terms of publication volume. 
+While transformative agreements could be attributed to `r length(unique(publisher_plot_df$esac_publisher))` distinct publishers, the three publishers Elsevier, Springer Nature and Wiley accounted for xxx percent of articles published.
+Together, they published xxx articles, representing a share of xx percent.
+xxx were published through transformative agreements.
+The remainder consists of
+
+```{r publisher_summary_table}
+publisher_article_total_table <- publisher_summary_df |>
+  distinct(publisher, jn_all) |>
+  group_by(publisher) |>
+  summarise(publisher_total = sum(jn_all)) |>
+  mutate(perc = publisher_total / sum(publisher_total),
+         cum = cumsum(publisher_total) / sum(publisher_total))
+
+publisher_article_oa_table <- publisher_summary_df |>
+  group_by(publisher) |>
+  summarise(oa_total = sum(value)) |>
+  mutate(oa_perc = oa_total / sum(oa_total),
+         oa_cum = cumsum(oa_total) / sum(oa_total))
+
+publisher_article_ta_oa_table <- publisher_summary_df |>
+  filter(type == "ta_oa") |>
+  group_by(publisher) |>
+  summarise(ta_oa_total = sum(value)) |>
+  mutate(ta_oa_perc = ta_oa_total / sum(ta_oa_total),
+         ta_oa_cum = cumsum(ta_oa_total) / sum(ta_oa_total))
+
+publisher_league_tbl <-
+  inner_join(publisher_article_total_table, publisher_article_oa_table) |>
+  inner_join(publisher_article_ta_oa_table)
+
+kbl(publisher_league_tbl) %>%
+  kable_classic() %>%
+  add_header_above(c("Publisher" = 1, "Group 1" = 3, "Group 2" = 3, "Group 3" = 3))
+```
+
+Figure 2 presents the growth of hybrid open access across publishers by year, highlighting the proportion of open articles published under transformative agreements.
+As can be seen from Figure 2A and 2C
+
 ```{r, fig.asp=1.1}
 library(patchwork)
-design = "A
+design = "E
+A
 B
@@
 wrap_plots(
+  E = guide_area(),
   A = publisher_prop,
@@
   guides = "collect",
- # widths = c(2, 1),
-  heights = c(1, 0.5, 1)
+  heights = c(0, 0.8, 0.3, 0.8)
 ) &
@@
     legend.position = "top",
-    legend.justification = "left",
+    legend.justification = "right",
     legend.direction = "horizontal",

### 2023-12-02 analysis/001_overview.Rmd
FInish publisher's league table

@@
   mutate(across(everything(), .fns = ~replace_na(.,0))) |>
-  mutate(no_ta = oa - ta_oa)
-
-## plot
-publisher_plot_df <- publisher_df |> 
-  pivot_longer(cols = c(ta_oa, no_ta), names_to = "type", values_to = "value") |>
+   mutate(no_ta = oa - ta_oa) |>
     mutate(
@@
            )) |>
-  filter(!cr_year %in%  c("2017", "2023"))
+  filter(!cr_year %in%  c("2017", "2023")) |>
+  # Journal transfers across publishers grouped in publisher
+  select(-esac_publisher) |>
+  distinct()
+
+## plot
+publisher_plot_df <- publisher_df |> 
+  pivot_longer(cols = c(ta_oa, no_ta), names_to = "type", values_to = "value")
 
@@
 ```{r publisher_summary_table}
-publisher_article_total_table <- publisher_summary_df |>
-  distinct(publisher, jn_all) |>
+publisher_article_total_table <- publisher_df |>
   group_by(publisher) |>
-  summarise(publisher_total = sum(jn_all)) |>
-  mutate(perc = publisher_total / sum(publisher_total),
-         cum = cumsum(publisher_total) / sum(publisher_total))
+  summarise(oa_journals = n_distinct(issn_l),
+            publisher_total = sum(jn_all)) |>
+  mutate(
+    jn_perc = oa_journals / sum(oa_journals),
+    jn_cum = cumsum(oa_journals) / sum(oa_journals),
+    perc = publisher_total / sum(publisher_total),
+    cum = cumsum(publisher_total) / sum(publisher_total)
+  ) |>
+  relocate(publisher_total, .after = jn_cum)
 
@@
   summarise(ta_oa_total = sum(value)) |>
-  mutate(ta_oa_perc = ta_oa_total / sum(ta_oa_total),
-         ta_oa_cum = cumsum(ta_oa_total) / sum(ta_oa_total))
+  mutate(
+    ta_oa_perc = ta_oa_total / sum(ta_oa_total),
+    ta_oa_cum = cumsum(ta_oa_total) / sum(ta_oa_total)
+  )
 
@@
 
-kbl(publisher_league_tbl) %>%
+publisher_league_tbl |>
+  mutate(across(contains(c("perc", "cum")), ~ paste(round(., 3) * 100))) |>
+  kbl(
+    col.names = c(
+      "Publisher",
+      "Total",
+      "%",
+      "C%",
+      "Total",
+      "%",
+      "C%",
+      "Total",
+      "%",
+      "C%",
+      "Total",
+      "%",
+      "C%"
+    ),
+    format.args = list(big.mark = ',')
+  ) %>%
   kable_classic() %>%
-  add_header_above(c("Publisher" = 1, "Group 1" = 3, "Group 2" = 3, "Group 3" = 3))
+  add_header_above(c(
+    " " = 1,
+    "Hybrid journals" = 3,
+    "Articles" = 3,
+    "OA articles" = 3,
+    "TA OA articles" = 3
+  ))
 ```

### 2023-12-02 analysis/001_overview.Rmd
Publishers league table description

@@
 
-Analysing hybrid open access across publishers between 2018 and 2022 reveals a large concentration on a few publishers in terms of publication volume. 
-While transformative agreements could be attributed to `r length(unique(publisher_plot_df$esac_publisher))` distinct publishers, the three publishers Elsevier, Springer Nature and Wiley accounted for xxx percent of articles published.
-Together, they published xxx articles, representing a share of xx percent.
-xxx were published through transformative agreements.
-The remainder consists of
-
 ```{r publisher_summary_table}
@@
   inner_join(publisher_article_ta_oa_table)
+```
 
+```{r}
 publisher_league_tbl |>
@@
   kbl(
+    caption ="Hybrid open access through transformative agreements market shares 2018-2022",
     col.names = c(
@@
 
-Figure 2 presents the growth of hybrid open access across publishers by year, highlighting the proportion of open articles published under transformative agreements.
-As can be seen from Figure 2A and 2C
+Analysing hybrid open access across publishers between 2018 and 2022 reveals a large market concentration among a small number of publishers. 
+Although `r length(unique(jn_ind_jct_all$esac_publisher))` publishers offered transformative agreements, the three major commercial publishers Elsevier, Springer Nature, and Wiley, accounted for `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(jn_cum) * 100`% of total article volume published (see Table ).
+Together, they published `r publisher_league_tbl  |> filter(publisher != "Other") |> pull(oa_total) |> sum()` open access articles, representing `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(oa_cum) * 100`% of the open access articles in hybrid journals.
+Elsevier, Springer Nature, and Wiley made `r publisher_league_tbl  |> filter(publisher != "Other") |> pull(ta_oa_total) |> sum()` articles open access in hybrid journals through transformative agreements, resulting in an even larger market share of `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(ta_oa_cum)*100`%.
+However, there are differences among the three large publishers.
+Although Elsevier published the largest volume of articles (n = `r publisher_league_tbl  |> filter(publisher == "Elsevier") |> pull(publisher_total)`, `r publisher_league_tbl  |> filter(publisher == "Elsevier") |> pull(cum)*100`%), it published a relatively low number of open access articles, including those that can be associated with transformative agreements.
+In contrast, Springer Nature and Wiley provided open access to a much larger proportion of their articles (`r (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_total)) / (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(publisher_total)) * 100`% of Springer Nature articles and `r (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(oa_total)) / (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(publisher_total)) * 100`% of Wiley articles were open acccess), leading to much higher open access market shares (`r publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_perc)*100`% Springer Nature resp.
+`r publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_perc)*100`% Wiley).
+This  difference can mainly be attributed to transformative agreements, as both publishers made the majority of their open access articles available through such deals (Springer Nature `r ((publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(ta_oa_total)) / (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_total)) * 100)`% resp.
+Wiley `r (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(ta_oa_total)) / (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(oa_total)) * 100`%).
+
+Figure 2 presents the growth of hybrid open access across publishers by year, highlighting the growth of open articles published under transformative agreements. As can be seen from Figure 2A and 2C
 

### 2023-12-03 analysis/001_overview.Rmd
tmp save publisher analysis

@@
 Between 2018 and 2022, a total of `r n_hybrid_jns` out of `r n_hybrid_jns_all` hybrid journals in transformative agreements published at least one open access article under a Creative Commons license.
-During this period, these hybrid journals provided open access to `r n_oa` out of `r n_total` articles, representing a five-year hybrid open access uptake of `r n_oa / n_total * 100`%.
-First authors affiliated with institutions that had a transformative agreement in place at the time of publication, which indicates that they could make use of such agreements, contributed `r n_ta_oa` open access articles to the total.
+During this period, these hybrid journals provided open access to `r n_oa` out of `r n_total` articles, representing a five-year open access proportion of `r n_oa / n_total * 100`%.
+Authors who could make use of transformative agreements at the time of publication contributed `r n_ta_oa` open access articles to the total.
 
@@
 
-When comparing the impact of open access trough transformative agreements across journals, it shows that for many journals these agreements have been a significant contributor to the growth of open access over the years (Figure 2D).
+When comparing the impact of open access trough transformative agreements across journals, it shows that for many journals these agreements substantially contributed to the growth of open access over the years (Figure 2D).
 Examples of such journals include those with a scope on specific countries or regions, where also transformative agreements were implemented. 
@@
 
-Analysing hybrid open access across publishers between 2018 and 2022 reveals a large market concentration among a small number of publishers. 
-Although `r length(unique(jn_ind_jct_all$esac_publisher))` publishers offered transformative agreements, the three major commercial publishers Elsevier, Springer Nature, and Wiley, accounted for `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(jn_cum) * 100`% of total article volume published (see Table ).
+Analysing hybrid open access across publishers between 2018 and 2022 reveals a large market concentration. 
+Although `r length(unique(jn_ind_jct_all$esac_publisher))` publishers offered transformative agreements, the big three commercial publishers Elsevier, Springer Nature, and Wiley, accounted for `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(jn_cum) * 100`% of total article volume published (see Table ).
 Together, they published `r publisher_league_tbl  |> filter(publisher != "Other") |> pull(oa_total) |> sum()` open access articles, representing `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(oa_cum) * 100`% of the open access articles in hybrid journals.
@@
 Although Elsevier published the largest volume of articles (n = `r publisher_league_tbl  |> filter(publisher == "Elsevier") |> pull(publisher_total)`, `r publisher_league_tbl  |> filter(publisher == "Elsevier") |> pull(cum)*100`%), it published a relatively low number of open access articles, including those that can be associated with transformative agreements.
-In contrast, Springer Nature and Wiley provided open access to a much larger proportion of their articles (`r (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_total)) / (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(publisher_total)) * 100`% of Springer Nature articles and `r (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(oa_total)) / (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(publisher_total)) * 100`% of Wiley articles were open acccess), leading to much higher open access market shares (`r publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_perc)*100`% Springer Nature resp.
+In contrast, Springer Nature and Wiley provided open access to a larger proportion of their articles (`r (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_total)) / (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(publisher_total)) * 100`% of Springer Nature articles and `r (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(oa_total)) / (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(publisher_total)) * 100`% of Wiley articles were open acccess), leading to higher open access market shares (`r publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_perc)*100`% Springer Nature resp.
 `r publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_perc)*100`% Wiley).
-This  difference can mainly be attributed to transformative agreements, as both publishers made the majority of their open access articles available through such deals (Springer Nature `r ((publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(ta_oa_total)) / (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_total)) * 100)`% resp.
+This  difference between Elsevier on the one hand and Springer Nature and Wiley on the other can  be attributed to transformative agreements, as the latter made the majority of their open access articles available through such deals (Springer Nature `r ((publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(ta_oa_total)) / (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_total)) * 100)`% resp.
 Wiley `r (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(ta_oa_total)) / (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(oa_total)) * 100`%).
 
-Figure 2 presents the growth of hybrid open access across publishers by year, highlighting the growth of open articles published under transformative agreements. As can be seen from Figure 2A and 2C
+Figure 2 takes a closer look into the growth of hybrid open access across publishers by year with a focus on open articles enabled by transformative agreements. 
+Although all publishers show a general long-term trend towards transformative agreements, Figure 2A and 2B indicate that, in particular, the open access share of Wiley's hybrid journal portfolio quadrupled from `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(prop) |> sum()*100`% in 2018 to `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(prop) |> sum()*100`% in 2022 as a result of these agreements.
+In comparison, the proportion of open access articles Elsevier hybrid journals grew from from `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Elsevier") |> pull(prop) |> sum()*100`% in 2018 to `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(prop) |> sum()*100`% in 2022.
+Unlike the other two publishers, Springer Nature started early to offer selected consortia open access agreements for its hybrid journal portfolio ("Springer Compact").
+Data about the Springer compact agreements with large consortia including Max Planck Society, the Swedish Bibsam consortium qnd the Finish	FinELib consortium that ended before start of the trnasofrmative agreement data collection in June 2021.
+ `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Springer Nature") |> pull(prop) |> sum()*100`% Springer Nature excelled in 2018, but gained a slower growth than Wiley, resulting into `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature") |> pull(prop) |> sum()*100`% of all articles being open access in Springer Nature hybrid journals in 2022.
+
+published the less than half of its `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(value) |> sum()` open access articles in 2022 through transformative agreements (`r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(value) |> sum()*100`%).
+while `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature") |> pull(value) |> sum()*100`% of Springer Nature's grew slower in the last year compared to Wiley, and  `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()*100`% of Wiley's open access articles could be linked to transformative agreements in 2022.
+Overall, Wiley published the largest number of open access articles from transformative agreements in 2022 (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()`), followed by Springer Nature (`r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()`) and Elsevier (`r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()`).
+This varying focus resulted into an different overall open access proportion in 2022 -- only 10% of all articles in Elsevier hybrid journals were open access in 2022, compared to published in springer nautre and xx in Wiley hybrid journals.
+
+The increasing impact of open access through transofmrative agreements can be also obsevered at the journal-level (Figure 2).
+While no substantial differences between open access enabled through transformative agreements can be seen across Elsevier journals, an growing number of hybrid journals published by Springer Nature and Wiley increased their open access share through these agreements.
+In 2022, 25% of Springer Nature hybrid journal portfolio comprised journals with an open access proportion of   percent, while Wiley.
+
+again a glanc eon the jorunals indicates that these ar emore smaller journals focus on a researchers affitialted with a socienty or a language springe rnautre german-speaking didactic journasl whereas in wiley british society tend to
+across all more smaller like 
+smaller and 
+of journals majority and springe rnature majority oabove 50 % . prominent example BJC
 

### 2023-12-04 analysis/001_overview.Rmd
first draft publisher analysis

@@
   scale_fill_manual("Hybrid OA", values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"), labels = c(no_ta = "Other", ta_oa = "Agreement"),  guide = guide_legend(reverse = TRUE)) +
-  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels = scales::percent_format()) +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels = scales::percent_format(),
+                     breaks = c(0.05, 0.1, 0.15, 0.2, 0.25)) +
   theme_minimal() +
@@
   coord_cartesian(ylim = c(0,0.65)) +
-  facet_wrap(~publisher, nrow = 1) +
+  facet_wrap(~publisher, nrow = 1, scales = "free") +
   labs(y = "OA Uptake\nby Hybrid Journal", x = "Publication Year") +
@@
 
-Figure 2 takes a closer look into the growth of hybrid open access across publishers by year with a focus on open articles enabled by transformative agreements. 
-Although all publishers show a general long-term trend towards transformative agreements, Figure 2A and 2B indicate that, in particular, the open access share of Wiley's hybrid journal portfolio quadrupled from `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(prop) |> sum()*100`% in 2018 to `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(prop) |> sum()*100`% in 2022 as a result of these agreements.
-In comparison, the proportion of open access articles Elsevier hybrid journals grew from from `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Elsevier") |> pull(prop) |> sum()*100`% in 2018 to `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(prop) |> sum()*100`% in 2022.
-Unlike the other two publishers, Springer Nature started early to offer selected consortia open access agreements for its hybrid journal portfolio ("Springer Compact").
-Data about the Springer compact agreements with large consortia including Max Planck Society, the Swedish Bibsam consortium qnd the Finish	FinELib consortium that ended before start of the trnasofrmative agreement data collection in June 2021.
- `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Springer Nature") |> pull(prop) |> sum()*100`% Springer Nature excelled in 2018, but gained a slower growth than Wiley, resulting into `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature") |> pull(prop) |> sum()*100`% of all articles being open access in Springer Nature hybrid journals in 2022.
-
-published the less than half of its `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(value) |> sum()` open access articles in 2022 through transformative agreements (`r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(value) |> sum()*100`%).
-while `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature") |> pull(value) |> sum()*100`% of Springer Nature's grew slower in the last year compared to Wiley, and  `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()*100`% of Wiley's open access articles could be linked to transformative agreements in 2022.
-Overall, Wiley published the largest number of open access articles from transformative agreements in 2022 (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()`), followed by Springer Nature (`r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()`) and Elsevier (`r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()`).
-This varying focus resulted into an different overall open access proportion in 2022 -- only 10% of all articles in Elsevier hybrid journals were open access in 2022, compared to published in springer nautre and xx in Wiley hybrid journals.
-
-The increasing impact of open access through transofmrative agreements can be also obsevered at the journal-level (Figure 2).
-While no substantial differences between open access enabled through transformative agreements can be seen across Elsevier journals, an growing number of hybrid journals published by Springer Nature and Wiley increased their open access share through these agreements.
-In 2022, 25% of Springer Nature hybrid journal portfolio comprised journals with an open access proportion of   percent, while Wiley.
-
-again a glanc eon the jorunals indicates that these ar emore smaller journals focus on a researchers affitialted with a socienty or a language springe rnautre german-speaking didactic journasl whereas in wiley british society tend to
-across all more smaller like 
-smaller and 
-of journals majority and springe rnature majority oabove 50 % . prominent example BJC
-
-```{r, fig.asp=1.1}
+```{r, fig.asp=0.8}
 library(patchwork)
@@
   guides = "collect",
-  heights = c(0, 0.8, 0.3, 0.8)
+  heights = c(0, 0.4, 0.3, 0.7)
 ) &
@@
 ```
+
+Figure 2 takes a closer look into the growth of hybrid open access across publishers by year with a focus on open articles enabled by transformative agreements. 
+Although all publishers show a general long-term trend towards transformative agreements, Figure 2A and 2B indicate that, in particular, Wiley's has experienced a substantial increase in its open access share from `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(value) |> sum()`) in 2018 to `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(prop) |> sum()*100`%  (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()`) in 2022, rperesenting an 4.5-fold increase.
+In contrast, Elsevier's hybrid journals deonstrated a more modest increase, from `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Elsevier") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Elsevier") |> pull(value) |> sum()`) in 2018 to `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(value) |> sum()`) in 2022, which is a relatively low open access share compared to the general trend.
+In 2018, Springer Nature had the largest open access proportion among the three publishers of `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Springer Nature") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Springer Nature") |> pull(value) |> sum()`), but experienced a relatively slower growth, resulting in `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature") |> pull(value) |> sum()`) of articles being open access in Springer Nature hybrid journals in 2022.
+
+The varying degrees of uptake of open access across the three major publishers can be attributed to distinct approaches to transformative agreements. 
+Springer Nature, for example, began in 2015 offering selected consortia, such as the Max Planck Society, the Swedish Bibsam consortium, and the Finnish FinELib consortium, open access agreements for its hybrid journal portfolio under the name Springer Compact[^springercompact].
+However, these agreements were not included in the data as they concluded prior to the start of the transformative agreement data collection in June 2021. 
+Nonetheless, the results suggest the importance of central agreements for Springer Nature's hybrid open access business over the past five years (Figure 2B).
+In 2022, `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature") |> pull(value) |> sum()*100`% (n = `r  publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature", type == "ta_oa") |> pull(value)`) of open access articles in f Springer Nature  hybrid journals were enabled through transformative agreements.
+In the same year, `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley", type == "ta_oa") |> pull(value)`) of Wiley's open access articles could be linked to transformative agreements in 2022.
+In contrast, Elsevier published fewer than half of its open access articles through transformative agreements (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier", type == "ta_oa") |> pull(value) `; `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(value) |> sum()*100`%).
+
+```{r}
+pub_75 <- publisher_plot_df |> filter(value > 0) |> mutate(prop = value /jn_all) |> group_by(publisher, type, cr_year) |> summarise(
+  top25 = quantile(prop, 0.75),
+  median = median(prop),
+  top10 = quantile(prop, 0.9))
+```
+
+The increasing trend towards transformative agreements can be also observed at the journal-level (Figure 2).
+While no substantial differences between open access enabled through transformative agreements and other revenue source could observed across Elsevier journals, the distribution of open access across Springer Nature and Wiley hybrid journals indicates that the growth is not limited to a few journals, but extends across the portfolio.
+In particular, Wiley's upper quantile, which represents the top 25% of journals in terms of the proportion of open access articles from transformative agreements, increased markedly from `r pub_75 |> filter(publisher == "Wiley", type == "ta_oa", cr_year == "2020") |> pull(top25) * 100`% in 2020 to `r pub_75 |> filter(publisher == "Wiley", type == "ta_oa", cr_year == "2022") |> pull(top25) * 100`% in 2022.
+At the same time, the median proportion grew from `r pub_75 |> filter(publisher == "Wiley", type == "ta_oa", cr_year == "2020") |> pull(median) * 100`% to  `r pub_75 |> filter(publisher == "Wiley", type == "ta_oa", cr_year == "2022") |> pull(median) * 100`%.
+It is interesting to note that a small but increasing number of journals from these two publishers are providing open access to the majority of articles through transformative agreements.
+Wiley recorded `r publisher_plot_df |>  mutate(prop = value / jn_all)  |> filter(cr_year == "2022", prop > 0.5, type == "ta_oa", publisher == "Wiley") |> nrow()` and Springer Nature `r publisher_plot_df |>  mutate(prop = value / jn_all)  |> filter(cr_year == "2022", prop > 0.5, type == "ta_oa", publisher == "Springer Nature") |> nrow()` hybrid journals with an open access share above 50% that could be solely attributed to transformative agreements.
+Upon inspection, these journals were mainly society or local language journals with a small yearly article volume.
+
+[^springercompact]: <https://web.archive.org/web/20180414062853id_/http://www.liber2015.org.uk/wp-content/uploads/2015/03/Springer-Compact.pdf>

### 2023-12-06 analysis/001_overview.Rmd
Subject figure

@@
   coord_cartesian(ylim = c(0,0.65)) +
-  facet_wrap(~publisher, nrow = 1, scales = "free") +
+  facet_wrap(~publisher, nrow = 1) +
   labs(y = "OA Uptake\nby Hybrid Journal", x = "Publication Year") +
@@
 
+Next, I focus on the subject distribution.
+
+```{r subject_data_prep}
+jn_issnl_ajsc <-
+  readr::read_csv(here::here("data", "jn_issnl_ajsc.csv")) |>
+  select(issn_l, top_level, subject_area, ajcs)
+subject_df <- left_join(jn_ind_jct_all, jn_ind_jct_oa) |>
+  mutate(cr_year = as.character(cr_year)) |>
+  left_join(ta_active) |>
+  mutate(across(everything(), .fns = ~ replace_na(., 0))) |>
+  mutate(no_ta = oa - ta_oa) |>
+  left_join(jn_issnl_ajsc, by = "issn_l")
+
+top_level_by_year_jn <- subject_df |>
+  # with at least one
+  filter(oa > 0) |>
+  distinct(cr_year, issn_l, jn_all, top_level) |>
+  group_by(cr_year, issn_l, top_level) |>
+  summarise(n = sum(jn_all)) |>
+  mutate(fractions_factor = n / sum(n)) |>
+  mutate(fractions = n * fractions_factor) |>
+  ungroup() |> 
+  group_by(cr_year, top_level) |>
+  summarise(articles_n = sum(n),
+            articles_frac = sum(fractions)) |>
+  arrange(top_level, cr_year)
+
+top_level_by_year_oa_frac <- subject_df |>
+   # with at least one
+  filter(oa > 0) |>
+  select(-ajcs, -subject_area, -esac_publisher) |>
+  distinct(cr_year, issn_l, top_level, .keep_all = TRUE) |>
+  group_by(cr_year, issn_l, top_level) |>
+  summarise(n = sum(jn_all)) |>
+  mutate(fractions_factor = n / sum(n)) |>
+  select(-n)
+
+top_level_by_year_oa <- top_level_by_year_oa_frac |>
+  inner_join(subject_df) |>
+     # with at least one
+  filter(oa > 0) |>
+  select(-ajcs, -subject_area, -esac_publisher) |>
+  distinct(cr_year, issn_l, top_level, .keep_all = TRUE) |>
+  mutate(across(c(oa, ta_oa, no_ta),  function(x) x * fractions_factor, .names = "frac_{.col}")) |>
+  ungroup() |> 
+  group_by(cr_year, top_level) |>
+  summarise(oa_n = sum(oa),
+            oa_frac = sum(frac_oa),
+            ta_oa = sum(ta_oa),
+            ta_oa_frac= sum(frac_ta_oa),
+            no_ta = sum(no_ta),
+            no_ta_fractions = sum(frac_no_ta)) |>
+  arrange(top_level, cr_year) |>
+  inner_join(top_level_by_year_jn)
+```
+
+```{r subject_plots}
+subject_trend_plot <- top_level_by_year_oa |>
+  filter(top_level != "Multidisciplinary") |>
+  filter(!cr_year %in%  c("2017", "2023")) |>
+  select(cr_year, top_level, contains("frac"),-oa_frac) |>
+  pivot_longer(
+    cols = c(ta_oa_frac, no_ta_fractions),
+    names_to = "type",
+    values_to = "value"
+  ) |>
+  mutate(prop = value / articles_frac) |>
+  ggplot(aes(
+    gsub("^20", "'", as.character(cr_year)),
+    prop,
+    fill = type,
+    group = type
+  )) +
+  geom_area() +
+  facet_wrap( ~ top_level, nrow = 1) +
+  scale_fill_manual(
+    "Hybrid OA",
+    values = c(
+      no_ta_fractions = "#b3b3b3a0",
+      ta_oa_frac = "#56B4E9"
+    ),
+    labels = c(no_ta_fractions = "Other", ta_oa_frac = "Agreement"),
+    guide = guide_legend(reverse = TRUE)
+  ) +
+  scale_y_continuous(
+    expand = expansion(mult = c(0, 0.05)),
+    labels = scales::percent_format(),
+    breaks = c(0.05, 0.1, 0.15, 0.2, 0.25)
+  ) +
+  theme_minimal() +
+  labs(y = "Hybrid OA Uptake", x = NULL) +
+  theme(
+    legend.position = "top",
+    legend.justification = "left",
+    axis.text.x = element_blank()
+  )
+
+subject_ta_trend_plot <- top_level_by_year_oa |>
+  filter(top_level != "Multidisciplinary") |>
+  filter(!cr_year %in%  c("2017", "2023")) |>
+  mutate(prop = ta_oa_frac / oa_frac) |>
+  ggplot(aes(gsub("^20", "'", as.character(cr_year)), prop)) +
+  geom_bar(stat = "identity", fill = "#56B4E9") +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), breaks = c(0, 0.25,0.5,0.75), limits = c(0, .75),
+                     labels = scales::percent) +
+  facet_wrap(~top_level, nrow = 1) +
+  theme_minimal() +
+  labs(y = "Percentage\nof Hybrid OA\nvia Agreement", x = NULL) +
+  theme(
+    axis.text.x = element_blank(),
+    strip.background = element_blank(),
+    strip.text.x = element_blank())
+subject_box_trend <- subject_df |>
+   # with at least one
+  select(-ajcs, -subject_area, -esac_publisher) |>
+  distinct(cr_year, issn_l, top_level, .keep_all = TRUE) |>
+  pivot_longer(cols = c(ta_oa, no_ta), names_to = "type", values_to = "value") |>
+  filter(value > 0) |>
+  filter(top_level != "Multidisciplinary", !cr_year %in% c("2017", "2023")) |>
+  ggplot(aes(gsub("^20", "'", as.character(cr_year)), value / jn_all, fill = type)) +
+  geom_boxplot(outlier.shape = NA, show.legend = FALSE, varwidth = TRUE) +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
+                     labels = scales::percent) +
+  scale_fill_manual("Hybrid OA", values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"), labels = c(no_ta = "Other", ta_oa = "Agreement")) +
+  coord_cartesian(ylim = c(0,0.55)) +
+  facet_wrap(~top_level, nrow = 1) +
+  labs(y = "OA Uptake\nby Hybrid Journal", x = "Publication Year") +
+  theme_minimal() +
+    theme(strip.background = element_blank(),
+    strip.text.x = element_blank()
+    )
+```
+
+```{r subject_panel, fig.asp=0.8}
+library(patchwork)
+design = "E
+A
+B
+C"
+wrap_plots(
+  E = guide_area(),
+  A = subject_trend_plot,
+  B = subject_ta_trend_plot,
+  C = subject_box_trend,
+  design = design,
+  guides = "collect",
+  heights = c(0, 0.4, 0.3, 0.7)
+) &
+  theme(
+    legend.position = "top",
+    legend.justification = "right",
+    legend.direction = "horizontal",
+    panel.grid.minor = element_blank(),
+                panel.border = element_rect(color = "grey50", fill = NA),
+    axis.title=element_text(size = 10)) &
+    plot_annotation(tag_levels = 'A') 
+```
+
 [^springercompact]: <https://web.archive.org/web/20180414062853id_/http://www.liber2015.org.uk/wp-content/uploads/2015/03/Springer-Compact.pdf>

### 2023-12-06 analysis/jn_classification.Rmd
Subject figure

@@
 ```
+
+```{r side_by_side_subject_publisher}
+# normalize publishers
+subject_short <- subject_df  |>  mutate(
+    publisher = case_when(
+      esac_publisher == "Elsevier" ~ "Elsevier",
+      esac_publisher == "Springer Nature" ~ "Springer Nature",
+      esac_publisher == "Wiley" ~ "Wiley",
+      is.character(esac_publisher) ~ "Other"
+    )
+  ) |>
+  mutate(publisher =
+           forcats::fct_relevel(
+             publisher,
+             c("Elsevier", "Springer Nature", "Wiley", "Other") 
+           )) |>
+  filter(!cr_year %in%  c("2017", "2023")) |>
+  # Journal transfers across publishers grouped in publisher
+  select(-esac_publisher) |>
+# top level
+ select(-subject_area, -ajcs) |>
+  distinct() |>
+  group_by(cr_year, top_level, publisher, issn_l) |>
+  summarise(oa = sum(oa), ta_oa = sum(ta_oa), no_ta = sum(no_ta), jn_all = sum(jn_all))
+
+subject_short |>
+  filter(top_level != "Multidisciplinary") |>
+  ggplot(aes(cr_year, oa / jn_all)) +
+  geom_boxplot(varwidth = TRUE) +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), breaks = c(0, 0.25,0.5,0.75), limits = c(0, .75),
+                     labels = scales::percent) +
+  facet_grid(top_level ~publisher)
+
+subject_short |>
+  ungroup() |>
+  filter(oa > 0) |>
+  select(-oa) |>
+  filter(top_level != "Multidisciplinary") |>
+  pivot_longer(cols = c(ta_oa, no_ta), names_to = "oa_type", values_to = "oa") |> 
+  ggplot(aes(cr_year, oa / jn_all, fill = oa_type)) +
+  geom_boxplot(varwidth = TRUE, outlier.shape = NA, notch = TRUE) +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
+                     labels = scales::percent,
+                     limits = c(0, .5)) +
+  facet_grid(top_level ~ publisher)
+
+subject_short |>
+  ungroup() |>
+  filter(oa > 0) |>
+  select(-oa) |>
+  pivot_longer(cols = c(ta_oa, no_ta), names_to = "oa_type", values_to = "oa") |> 
+  filter(top_level != "Multidisciplinary") |>
+  mutate(publisher = as.character(publisher)) |>
+  group_by(publisher, top_level, cr_year) |>
+
+subject_short |>
+  ungroup() |>
+  filter(oa > 0) |>
+  select(-oa) |>
+  pivot_longer(cols = c(ta_oa, no_ta), names_to = "oa_type", values_to = "oa") |> 
+  filter(top_level != "Multidisciplinary") |>
+  mutate(publisher = as.character(publisher)) |>
+  group_by(publisher, top_level, cr_year) |>
+  summarise(nn = wilcox.test(oa ~ oa-type)$p.value))
+
+subject_short |>
+ ungroup() |>
+filter(oa > 0) |>
+select(-oa) |>
+ pivot_longer(cols = c(ta_oa, no_ta), names_to = "oa_type", values_to = "oa") |> 
+  mutate(oa_prop = oa / jn_all) |>
+ filter(top_level != "Multidisciplinary") |>
+ mutate(publisher = as.character(publisher)) |>
+  group_by(publisher, top_level, cr_year) |>
+  summarise(p_value = wilcox.test(oa ~ oa_type, paired = TRUE)$p.value) |> View()
+
+preprints %>%
+  filter(posted_date >= analysis_start,
+         posted_date <= analysis_end) %>%
+  with(., wilcox.test(n_versions ~ covid_preprint))
+
+```

### 2023-12-06 analysis/data_prep_ajsc.R

@@ -1,0 +1,51 @@
+#' Adding AJSC codes
+library(tidyverse)
+library(hoaddata)
+library(janitor)
+#' HOAD journals from JCT
+my_jns <- hoaddata::jct_hybrid_jns |>
+  # filter(issn_l == "0002-8762") |> 
+  mutate(id = issn_l) |>
+  pivot_longer(cols = c(issn_l, issn)) |>
+  distinct(issn_l = id, issn = value)
+#' Scopus journal list from August 2023
+scopus <- readxl::read_excel("~/Downloads/extlistAugust2023.xlsx", sheet = 1,
+                             .name_repair = janitor::make_clean_names,
+                             col_types = "text")
+#' Prepare for merge
+scopus_norm <- scopus |>
+  select(1:4, all_science_journal_classification_codes_asjc) |>
+  separate(print_issn, c("print_issn", "print_issn_extra")) |>
+  gather(print_issn,
+         e_issn,
+         print_issn_extra,
+         key = "issn_tpye",
+         value = "issn") |>
+  filter(!is.na(issn)) |>
+  # trailing zero's missing in Excel spreadsheet
+  mutate(
+    issn = ifelse(nchar(issn) == 5, paste0("000", issn), issn),
+    issn = ifelse(nchar(issn) == 6, paste0("00", issn), issn),
+    issn = ifelse(nchar(issn) == 7, paste0("0", issn), issn)
+  ) |>
+  # missing hyphen
+  mutate(issn = map_chr(issn, function(x)
+    paste(c(
+      substr(x, 1, 4), substr(x, 5, 8)
+    ), collapse = "-")))
+#' Match
+matched_jns <- my_jns |>
+  inner_join(scopus_norm, by = "issn") 
+#' Add AJCS codes
+ajsc_mapped <- readr::read_csv("https://raw.githubusercontent.com/njahn82/elsevier_hybrid_invoicing/master/data/asjc_mapped.csv", col_types = "ccc") 
+subject_jn_df <- matched_jns |>
+  distinct() |>
+  mutate(ajcs = strsplit(all_science_journal_classification_codes_asjc, ";")) |>
+  unnest(ajcs) |>
+  mutate(top_level_code = str_extract(ajcs, "\\d{2}")) |>
+  inner_join(ajsc_mapped, by = "top_level_code") |>
+  rename(top_level = subject_area, subject_area = description) |> 
[…]

### 2023-12-07 analysis/001_overview.Rmd
Subject overview table

@@
 
+Here's a table. I applied fractional counting because of large overlaps.
+
+```{r subject_summary_table}
+jn_fract <- subject_df |>
+  # with at least one
+  filter(oa > 0) |>
+  distinct(issn_l, top_level) |>
+  group_by(issn_l, top_level) |>
+  summarise(n = n_distinct(issn_l)) |>
+  mutate(fractions_factor = n / sum(n)) |>
+  mutate(fractions = n * fractions_factor) |>
+  ungroup() |> 
+  group_by(top_level) |>
+  summarise(jn_frac = sum(fractions)) |>
+  arrange(top_level)
+
+subject_tbl <- top_level_by_year_oa |>
+  inner_join(jn_fract, by = "top_level") |>
+  filter(!top_level %in% c("Multidisciplinary", NA)) |>
+  group_by(top_level, jn_frac) |>
+  summarise(articles_frac = sum(articles_frac),
+            oa_frac = sum(oa_frac),
+            ta_oa_frac = sum(ta_oa_frac)) |>
+  ungroup() |>
+  mutate(
+    jn_perc = jn_frac / sum(jn_frac),
+    jn_cum = cumsum(jn_frac) / sum(jn_frac),
+    perc = articles_frac / sum(articles_frac),
+    cum = cumsum(articles_frac) / sum(articles_frac),
+    oa_perc = oa_frac / sum(oa_frac),
+    oa_cum = cumsum(oa_frac) / sum(oa_frac),
+    ta_oa_perc = ta_oa_frac / sum(ta_oa_frac),
+    ta_oa_cum = cumsum(ta_oa_frac) / sum(ta_oa_frac)
+  ) |>
+  select(top_level, jn_frac, jn_perc, jn_cum, articles_frac, perc, cum, 
+         oa_frac, oa_perc, oa_cum, ta_oa_frac,ta_oa_perc,  ta_oa_cum)
+
+subject_tbl |>
+  mutate(across(contains(c("perc", "cum")), ~ paste(round(., 3) * 100))) |>
+  kbl(
+    caption ="Hybrid open access through transformative agreements by journal subject 2018-2022",
+    col.names = c(
+      "Journal subject",
+      "Total",
+      "%",
+      "C%",
+      "Total",
+      "%",
+      "C%",
+      "Total",
+      "%",
+      "C%",
+      "Total",
+      "%",
+      "C%"
+    ),
+    format.args = list(big.mark = ',')
+  ) %>%
+  kable_classic() %>%
+  add_header_above(c(
+    " " = 1,
+    "Hybrid journals" = 3,
+    "Articles" = 3,
+    "OA articles" = 3,
+    "TA OA articles" = 3
+  ))
+```
+
 ```{r subject_plots}

### 2023-12-07 analysis/001_overview.Rmd
Country plot top 20

@@
 
+And here's my country analysis. First top 20
+
+```{r coutnry_data_prep}
+country_all <- hoaddata::jn_aff |>
+  filter(!cr_year %in% c(2017, 2023), issn_l != "0027-8424") |>
+  # only with cc
+  filter(issn_l %in% hoaddata::cc_articles$issn_l) |>
+  distinct(issn_l, cr_year, country_code, articles_total) |>
+  group_by(cr_year, country_code) |>
+  summarise(all = sum(articles_total))
+
+country_oa <- hoaddata::cc_articles |>
+  filter(!cr_year %in% c(2017, 2023), issn_l != "0027-8424") |>
+  group_by(cr_year, country_code) |>
+  summarise(oa = n_distinct(doi))
+
+country_codes <- readr::read_csv(here::here("data", "ror_country_codes.csv"))
+country_ta <-  ta_oa_inst |>
+  filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
+  left_join(country_codes) |>
+  group_by(cr_year, country_code) |>
+  summarise(ta_oa = n_distinct(doi))
+
+country_df <- country_all |>
+  left_join(country_oa, by = c("cr_year", "country_code")) |>
+  left_join(country_ta, by = c("cr_year", "country_code")) |>
+  distinct() |>
+  ungroup() |>
+  mutate(cr_year = as.character(cr_year)) |>
+  mutate(across(c(oa, ta_oa), .fns = ~replace_na(.,0))) |>
+  filter(!is.na(country_code)) |>
+  mutate(
+    country_name = countrycode::countrycode(country_code, origin = "iso2c", destination = "country.name")
+  )
+
+# Top 20 in terms of articles published
+top_20 <- country_df |>
+  filter(!is.na(country_code)) |>
+  group_by(country_code, country_name) |>
+  summarise(n = sum(all)) |>
+  arrange(desc(n)) |>
+  head(20)
+
+## Order of multiples, sort by the most productive (Top 20)
+
+country_multiple_df <- country_df |>
+  mutate(country_name_fct = forcats::fct_other(country_name, keep = top_20$country_name)) |>
+  mutate(country_name_fct = forcats::fct_relevel(country_name_fct, top_20$country_name, "Other")) |>
+  group_by(country_name_fct, cr_year) |>
+  summarise(
+    all_total = sum(all),
+    oa_total = sum(oa),
+    ta_oa_total = sum(ta_oa)
+  ) |>
+  mutate(prop = oa_total / all_total) |>
+  # remove cases where we could not retrieve a country name |>
+  filter(!is.na(country_name_fct))
+
+# Prepare facet heading
+facet_labels <- country_multiple_df |>
+  group_by(country_name_fct) |>
+  summarise(total = sum(all_total)) |>
+  mutate(prop = total / sum(total)) |>
+  ungroup() |>
+  # Highlight Germany
+  mutate(total_string = case_when(
+    total < 1000 ~ as.character(total),
+    total < 1000000 ~ paste0(format(round(total / 1e3, 1), trim = TRUE), "K"),
+    total >= 1000000 ~ paste0(format(round(total / 1e6, 2), trim = TRUE), "M")
+  )) |>
+  mutate(
+    facet_labels = glue::glue(
+      '{country_name_fct}<br>
+      {total_string} | {paste(round(prop * 100, 1), "%")}'
+    )
+  ) |>
+  distinct(country_name_fct, facet_labels)
+
+country_facet_labels <- as.character(facet_labels$facet_labels)
+names(country_facet_labels) <- facet_labels$country_name_fct
+
+top_20_plot_df <- 
+  country_multiple_df |> 
+  mutate(no_ta = oa_total - ta_oa_total) |>
+  rename(ta_oa = ta_oa_total) |>
+  pivot_longer(
+  cols = c(no_ta, ta_oa),
+  values_to = "value",
+  names_to = "type") 
+```
+
+```{r, country_plot, fig.asp=0.8}
+ggplot(top_20_plot_df, aes(
+    gsub("^20", "'", as.character(cr_year)),
+    value / all_total,
+    group = type,
+    fill = type
+  )) +
+  geom_area(stat = "identity") +
+  scale_fill_manual(
+    "Hybrid OA",
+    values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"),
+    labels = c(no_ta = "Other", ta_oa = "Agreement"),
+    guide = guide_legend(reverse = TRUE)
+  ) +
+  facet_wrap( ~ country_name_fct,
+              nrow = 3,
+              labeller = as_labeller(country_facet_labels)) +
+  scale_y_continuous(
+    labels = scales::percent_format(accuracy = 1L),
+    expand = expansion(mult = c(0, 0.05)),
+    breaks = scales::breaks_extended(4)
+  ) +
+  theme_minimal() +
+  labs(y = "OA Share", x = "2018 - 2022") +
+  theme(
+    legend.position = "top",
+    legend.justification = "right",
+    legend.direction = "horizontal",
+    panel.grid.minor = element_blank(),
+                panel.border = element_rect(color = "grey50", fill = NA),
+    axis.title=element_text(size = 10),
+    strip.text = ggtext::element_markdown(size = 6),
+    plot.subtitle = ggtext::element_markdown(hjust = 0.5),
+    panel.spacing = unit(1, "lines")
+  )
+```
+
 [^springercompact]: <https://web.archive.org/web/20180414062853id_/http://www.liber2015.org.uk/wp-content/uploads/2015/03/Springer-Compact.pdf>

### 2023-12-07 analysis/get_ror_countrycode.R

@@ -1,0 +1,28 @@
+# Get Country Codes
+library(bigrquery)
+library(tidyverse)
+
+my_sql <-  "SELECT
+  DISTINCT main.ror_main,
+  country_code
+FROM
+  `subugoe-collaborative.hoaddata.ta_oa_inst` AS main
+INNER JOIN
+  `subugoe-collaborative.openalex.institutions` AS oalex
+ON
+  ror_main = oalex.ror"
+
+tb <- bq_project_query("subugoe-collaborative", my_sql)
+ror_country_codes_raw <- bq_table_download(tb)
+
+ror_country_codes <- ror_country_codes_raw |>
+  # Fix missing country codes
+  mutate(country_code = case_when(
+    ror_main == "https://ror.org/016xje988" ~ "NA",
+    ror_main == "https://ror.org/03nnxqz81" ~ "SE",
+    ror_main == "https://ror.org/05v0p1f11" ~ "TR",
+    .default = as.character(country_code)
+  )) |>
+  distinct()
+
+readr::write_csv(ror_country_codes, here::here("data/ror_country_codes.csv"))

### 2023-12-07 analysis/001_overview.Rmd
Fisrt draft country plot and save figures in dedicated folder

@@
   fig.align = "center",
-  dpi = 300
+  dpi = 300,
+  dev = c("png", "pdf"),
+  fig.path = "fig/"
 )
@@
   theme_minimal() +
-  labs(y = "OA Share", x = "2018 - 2022") +
+  labs(y = "Hybrid OA Uptake", x = "2018 - 2022") +
   theme(
@@
     legend.direction = "horizontal",
+    axis.text.x = element_blank(),
     panel.grid.minor = element_blank(),
@@
 
+here a scatterplot and a classifciation of oecd / brics / other. first, the scatterplot 2022!
+
+```{r}
+oecd <- readr::read_csv(here::here("data", "oecd.csv")) |>
+  mutate(country_code = countrycode::countrycode(iso3, "iso3c", "iso2c"))
+
+country_scatter_df <- country_df |>
+#  filter(cr_year == "2022", !is.na(country_code)) |>
+  mutate(cat = case_when(
+    country_code %in% oecd$country_code ~ "OECD",
+    country_code %in% c("BR", "CN", "IN", "RU", "ZA") ~ "BRICS",
+    .default = "Other"
+  ))
+
+country_scatter_22 <- country_scatter_df |>
+  filter(cr_year == "2022", !is.na(country_code)) |>
+  ggplot(aes(all, oa)) +
+  geom_point(aes(color = cat, size = ta_oa), alpha = 0.8) +
+  scale_color_manual(NULL, values = c("OECD" = "#3fc1c9", "BRICS" = "#fc5185", "Other" = "#e1e1e1")) +
+  geom_smooth(aes(all, oa), method=lm) +
+  scale_size("TA OA", guide = "none") +
+  ggrepel::geom_text_repel(
+    data = filter(country_scatter_df, cr_year == 2022, cat %in% c("OECD", "BRICS"), all > 50000),
+    aes(x = all, y = oa, label = country_code),
+    force = 50,
+    box.padding = 0.5,
+    point.padding = 0.8,
+    size = 3,
+    max.iter = 1000000,
+    max.overlaps = Inf,
+    min.segment.length = 0,
+    arrow = arrow(length = unit(0.010, "npc"))) +
+  scale_x_continuous(labels =  scales::number_format(big.mark = ","), guide = guide_axis(angle = 45)) +
+  scale_y_continuous(labels =  scales::number_format(big.mark = ",")) +
+  theme_minimal() +
+  labs(x = "Total articles", y = "OA articles in hybrid journals") +
+  theme(legend.position = "top",
+    legend.justification = "left",
+    legend.direction = "horizontal")
+
+country_bar <- country_scatter_df |>
+  group_by(cr_year, cat) |>
+  summarise(all = sum(all), oa = sum(oa), ta_oa = sum(ta_oa)) |>
+  mutate(closed = all - oa) |>
+  pivot_longer(cols = c(oa, ta_oa, closed), names_to = "oa_type") |>
+  ggplot(aes(gsub("^20", "'", as.character(cr_year)), value, group = oa_type, fill = oa_type)) +
+  geom_area() +
+  facet_wrap(~ cat) +
+  scale_fill_manual(NULL, values = c(oa = "#b3b3b3a0", ta_oa = "#56B4E9", closed = "#e98b56"),
+    labels = c(oa = "OA", ta_oa = "Agreement", closed = "Closed")) +
+  theme_minimal() +
+  scale_y_continuous(labels =  scales::number_format(big.mark = ",")) +
+  theme_minimal() +
+  labs(x = "Publication year", y = "Total articles") +
+  theme(legend.position = "top",
+    legend.justification = "left",
+    legend.direction = "horizontal")
+```
+
+```{r country_patch}
+library(patchwork)
+wrap_plots(
+  A = country_scatter_22,
+  B = country_bar,
+  design = "AB",
+  widths = c(0.35, 0.65)
+) &
+  theme(
+    panel.grid.minor = element_blank(),
+                panel.border = element_rect(color = "grey50", fill = NA),
+    axis.title=element_text(size = 10)) &
+    plot_annotation(tag_levels = 'A') 
+```
+
 [^springercompact]: <https://web.archive.org/web/20180414062853id_/http://www.liber2015.org.uk/wp-content/uploads/2015/03/Springer-Compact.pdf>

### 2023-12-13 data/000_data_gathering.R
Update data to 0.2.91

@@ -29,11 +29,11 @@
 #'  <https://github.com/subugoe/hoaddata/releases/download/v0.2.82/ta_oa_inst.csv.gz>
 #'  
 #'  A version is provided in `data/` folder
-download.file("https://github.com/subugoe/hoaddata/releases/download/v0.2.82/ta_oa_inst.csv.gz",
+download.file("https://github.com/subugoe/hoaddata/releases/download/v0.2.91/ta_oa_inst.csv.gz",
               destfile = here::here("data", "ta_oa_inst.csv.gz")) 
 #' *Raw data for reproducibility*: To improve reproducibility, 
-#' the underying raw data used to compile {hoaddata} 
-#' version v0.2.82 is available via Google BigQuery `hoa-article.hoaddata_nov23`.
+#' the underlying raw data used to compile {hoaddata} 
+#' version v0.2.91 is available via Google BigQuery `hoa-article.hoaddata_nov23`.
 #' 
 #' *Subject classification* data was obtained from Scopus on 4 Oct 2023 and is
 #' stored in `data/jn_scopus_ind_subjects.csv`

### 2023-12-22 001_country_classification.R
OECD stats

@@ -1,0 +1,150 @@
+#' Apply OECD classification
+library(dplyr)
+library(bigrquery)
+library(tibble)
+library(countrycode)
+#' Prepare matching table
+oecd_countries <- readr::read_csv(here::here("data", "oecd.csv")) |>
+  mutate(cat = "OECD")
+brics <- tibble::tribble(
+  ~ country, ~ iso3,
+  "Brazil", "BRA",
+  "Russia", "RUS",
+  "India", "IND",
+  "China", "CHN",
+  "South Africa", "ZAF"
+) |>
+  mutate(cat = "BRICS")
+
+country_classification <- bind_rows(oecd_countries, brics) |>
+  mutate(iso2 = countrycode::countrycode(iso3,
+                                         origin = "iso3c",
+                                         destination = "iso2c")) |>
+  mutate(iso2 = ifelse(iso3 == "IRE", "IE", iso2))
+#' Upload to BQ
+bigrquery::bq_table_upload(
+  bq_table("hoa-article", "hoaddata_nov23", "country_classification"),
+  country_classification
+)
+#' Aggregate per group
+my_sql <- "WITH
+  all_articles AS (
+  SELECT
+    issn_l,
+    cr_year,
+    cat,
+    COUNT(DISTINCT doi) AS articles,
+  FROM
+    `hoa-article.hoaddata_nov23.cr_openalex_inst_full` AS alex
+  LEFT JOIN
+    `hoa-article.hoaddata_nov23.country_classification` AS country
+  ON
+    alex.country_code = country.iso2
+  GROUP BY
+    issn_l,
+    cr_year,
+    cat ),
+  cc_articles AS (
+  SELECT
+    issn_l,
+    cr_year,
+    cat,
+    COUNT(DISTINCT doi) AS cc_articles,
+  FROM
+    `hoa-article.hoaddata_nov23.cc_openalex_inst` AS alex
+  LEFT JOIN
+    `hoa-article.hoaddata_nov23.country_classification` AS country
+  ON
+    alex.country_code = country.iso2
+  GROUP BY
+    issn_l,
+    cr_year,
+    cat ),
+  global AS (
+  SELECT
+    all_articles.*,
+    cc_articles.cc_articles
+  FROM
+    all_articles
+  LEFT JOIN
+    cc_articles
+  ON
+    all_articles.issn_l = cc_articles.issn_l
[…]

### 2023-12-22 data/000_data_gathering.R

@@ -43,15 +43,15 @@ download.file("https://github.com/subugoe/hoaddata/releases/download/v0.2.91/ta_
 library(bigrquery)
 library(tidyverse)
 
-my_sql <-  "SELECT
-  DISTINCT main.ror_main,
-  country_code
-FROM
-  `hoa-article.hoaddata_nov23.ta_oa_inst` AS main
-INNER JOIN
-  `subugoe-collaborative.openalex.institutions` AS oalex
-ON
-  ror_main = oalex.ror"
+  my_sql <-  "SELECT
+    DISTINCT main.ror_main,
+    country_code
+  FROM
+    `hoa-article.hoaddata_nov23.ta_oa_inst` AS main
+  INNER JOIN
+    `subugoe-collaborative.openalex.institutions` AS oalex
+  ON
+    ror_main = oalex.ror"
 
 tb <- bq_project_query("subugoe-collaborative", my_sql)
 ror_country_codes_raw <- bq_table_download(tb)

### 2023-12-22 analysis/001_overview.Rmd
Update data to 0.2.91
Improve OECD ananlysi, in particular counting method!

@@
 ta_oa_inst <- readr::read_csv(here::here("data", "ta_oa_inst.csv.gz"))
+
 ```
@@
 
-country_bar <- country_scatter_df |>
+oecd_stats <- readr::read_csv(here::here("data", "oecd_stats.csv"))
+oecd_stats_df <- oecd_stats |>
+  filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
+  filter(issn_l %in% hoaddata::cc_articles$issn_l) |>
   group_by(cr_year, cat) |>
-  summarise(all = sum(all), oa = sum(oa), ta_oa = sum(ta_oa)) |>
-  mutate(closed = all - oa) |>
-  pivot_longer(cols = c(oa, ta_oa, closed), names_to = "oa_type") |>
-  ggplot(aes(gsub("^20", "'", as.character(cr_year)), value, group = oa_type, fill = oa_type)) +
+  summarise(all = sum(articles, na.rm = TRUE), oa = sum(oa,na.rm = TRUE), ta_oa = sum(ta_oa, na.rm = TRUE)) |>
+  mutate(closed = all - oa,
+         oa_other = oa - ta_oa) |>
+  pivot_longer(cols = c(ta_oa, oa_other, closed), names_to = "oa_type") |>
+  mutate(cat = ifelse(is.na(cat), "Other", cat))
+
+country_bar <- ggplot(oecd_stats_df, aes(gsub("^20", "'", as.character(cr_year)), value, group = oa_type, fill = oa_type)) +
   geom_area() +
   facet_wrap(~ cat) +
-  scale_fill_manual(NULL, values = c(oa = "#b3b3b3a0", ta_oa = "#56B4E9", closed = "#e98b56"),
-    labels = c(oa = "OA", ta_oa = "Agreement", closed = "Closed")) +
+  scale_fill_manual(NULL, values = c(oa_other = "#b3b3b3a0", ta_oa = "#56B4E9", closed = "grey50"),
+                    labels = c(oa_other = "OA Other", ta_oa = "Agreement", closed = "Closed")) +
   theme_minimal() +
@@
         legend.direction = "horizontal")
+
 ```

### 2023-12-22 001_country_classification.R

@@ -1,150 +1,0 @@
-#' Apply OECD classification
-library(dplyr)
-library(bigrquery)
-library(tibble)
-library(countrycode)
-#' Prepare matching table
-oecd_countries <- readr::read_csv(here::here("data", "oecd.csv")) |>
-  mutate(cat = "OECD")
-brics <- tibble::tribble(
-  ~ country, ~ iso3,
-  "Brazil", "BRA",
-  "Russia", "RUS",
-  "India", "IND",
-  "China", "CHN",
-  "South Africa", "ZAF"
-) |>
-  mutate(cat = "BRICS")
-
-country_classification <- bind_rows(oecd_countries, brics) |>
-  mutate(iso2 = countrycode::countrycode(iso3,
-                                         origin = "iso3c",
-                                         destination = "iso2c")) |>
-  mutate(iso2 = ifelse(iso3 == "IRE", "IE", iso2))
-#' Upload to BQ
-bigrquery::bq_table_upload(
-  bq_table("hoa-article", "hoaddata_nov23", "country_classification"),
-  country_classification
-)
-#' Aggregate per group
-my_sql <- "WITH
-  all_articles AS (
-  SELECT
-    issn_l,
-    cr_year,
-    cat,
-    COUNT(DISTINCT doi) AS articles,
-  FROM
-    `hoa-article.hoaddata_nov23.cr_openalex_inst_full` AS alex
-  LEFT JOIN
-    `hoa-article.hoaddata_nov23.country_classification` AS country
-  ON
-    alex.country_code = country.iso2
-  GROUP BY
-    issn_l,
-    cr_year,
-    cat ),
-  cc_articles AS (
-  SELECT
-    issn_l,
-    cr_year,
-    cat,
-    COUNT(DISTINCT doi) AS cc_articles,
-  FROM
-    `hoa-article.hoaddata_nov23.cc_openalex_inst` AS alex
-  LEFT JOIN
-    `hoa-article.hoaddata_nov23.country_classification` AS country
-  ON
-    alex.country_code = country.iso2
-  GROUP BY
-    issn_l,
-    cr_year,
-    cat ),
-  global AS (
-  SELECT
-    all_articles.*,
-    cc_articles.cc_articles
-  FROM
-    all_articles
-  LEFT JOIN
-    cc_articles
-  ON
-    all_articles.issn_l = cc_articles.issn_l
[…]

### 2023-12-22 data/001_country_classification.R

@@ -1,0 +1,88 @@
+#' Apply OECD classification
+library(dplyr)
+library(bigrquery)
+library(tibble)
+library(countrycode)
+#' Prepare matching table
+oecd_countries <- readr::read_csv(here::here("data", "oecd.csv")) |>
+  mutate(cat = "OECD")
+brics <- tibble::tribble(
+  ~ country, ~ iso3,
+  "Brazil", "BRA",
+  "Russia", "RUS",
+  "India", "IND",
+  "China", "CHN",
+  "South Africa", "ZAF"
+) |>
+  mutate(cat = "BRICS")
+
+country_classification <- bind_rows(oecd_countries, brics) |>
+  mutate(iso2 = countrycode::countrycode(iso3,
+                                         origin = "iso3c",
+                                         destination = "iso2c")) |>
+  mutate(iso2 = ifelse(iso3 == "IRE", "IE", iso2))
+#' Upload to BQ
+#' 
+my_bq_table <-  bq_table("hoa-article", "hoaddata_nov23", "country_classification")
+if (bigrquery::bq_table_exists(my_bq_table)) {
+  bigrquery::bq_table_delete(my_bq_table)
+}
+bigrquery::bq_table_upload(
+  my_bq_table,
+  country_classification
+)
+
+#' Aggregate per group
+my_sql <- "WITH
+  doi_raw AS (
+  SELECT
+    DISTINCT cc_md.doi,
+    cc_md.issn_l,
+    cc_md.cr_year,
+    CASE
+      WHEN cc IS NOT NULL THEN 1
+    ELSE
+    0
+  END
+    AS oa,
+    CASE
+      WHEN EXISTS ( SELECT * FROM ( SELECT doi FROM `hoa-article.hoaddata_nov23.ta_oa_inst` WHERE ta_active = TRUE AND cc IS NOT NULL ) AS t WHERE cc_md.doi = t.doi ) THEN 1
+    ELSE
+    0
+  END
+    AS ta_oa
+  FROM
+    `subugoe-collaborative.hoaddata.cc_md` AS cc_md),
+  -- Add country info
+  country_info AS (
+  SELECT
+    DISTINCT doi_raw.*,
+    country.cat
+  FROM
+    doi_raw
+  LEFT JOIN
+    `hoa-article.hoaddata_nov23.cr_openalex_inst_full` AS alex
+  ON
+    doi_raw.doi = alex.doi
+  LEFT JOIN
[…]

### 2024-01-04 analysis/001_overview.Rmd
playing around with color scales

--- 2023-12-07 analysis/001_overview.Rmd
@@
 oa_ta_plot <- ggplot(oa_ta_plot_df, aes(cr_year, value / n, fill = type, group = type)) +
-  geom_area() +
-  scale_fill_manual(NULL, values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"), labels = c(no_ta = "Other", ta_oa = "Agreement"),  guide = guide_legend(reverse = TRUE)) +
+  geom_area(color = "white",
+            alpha = 0.8,
+            linewidth = 0.3) +
+  scale_fill_manual(NULL, values = c(no_ta = "#b3b3b3", ta_oa = "#56B4E9"), labels = c(no_ta = "Other", ta_oa = "Agreement"),  guide = guide_legend(reverse = TRUE)) +
   scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels = scales::percent_format(), limits = c(0, .15), breaks = c(0, .05, .1, .15)) +
@@
 publisher_prop <- ggplot(publisher_summary_df, aes(gsub("^20", "'", as.character(cr_year)), prop, fill = type, group = type)) +
-  geom_area() +
+    geom_area(color = "white",
+            alpha = 0.8,
+            linewidth = 0.3) +
   facet_wrap(~publisher, nrow = 1) +
-  scale_fill_manual("Hybrid OA", values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"), labels = c(no_ta = "Other", ta_oa = "Agreement"),  guide = guide_legend(reverse = TRUE)) +
+  scale_fill_manual("Hybrid OA", values = c(no_ta = "#b3b3b3", ta_oa = "#56B4E9"), labels = c(no_ta = "Other", ta_oa = "Agreement"),  guide = guide_legend(reverse = TRUE)) +
   scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels = scales::percent_format(),
@@
   )) +
-  geom_area() +
+    geom_area(color = "white",
+            alpha = 0.8,
+            linewidth = 0.3) +
   facet_wrap( ~ top_level, nrow = 1) +
@@
     values = c(
-      no_ta_fractions = "#b3b3b3a0",
+      no_ta_fractions = "#b3b3b3",
       ta_oa_frac = "#56B4E9"
@@
   )) +
-  geom_area(stat = "identity") +
+  geom_area(color = "white",
+            alpha = 0.8,
+            linewidth = 0.3,
+            stat = "identity") +
 scale_fill_manual(
     "Hybrid OA",
-    values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"),
+    values = c(no_ta = "#b3b3b3", ta_oa = "#56B4E9"),
     labels = c(no_ta = "Other", ta_oa = "Agreement"),
@@
 
+here a scatterplot and a classifciation of oecd / brics / other. first, the scatterplot 2022!
+
+```{r}
+oecd <- readr::read_csv(here::here("data", "oecd.csv")) |>
+  mutate(country_code = countrycode::countrycode(iso3, "iso3c", "iso2c"))
+
+country_scatter_df <- country_df |>
+#  filter(cr_year == "2022", !is.na(country_code)) |>
+  mutate(cat = case_when(
+    country_code %in% oecd$country_code ~ "OECD",
+    country_code %in% c("BR", "CN", "IN", "RU", "ZA") ~ "BRICS",
+    .default = "Other"
+  ))
+
+```
+
+```{r country_bar_chart_top_20}
+top_20_df <- country_scatter_df |>
+  filter(cr_year == "2022") |>
+  mutate(country_name_fct = forcats::fct_other(country_name, keep = top_20$country_name)) |>
+  mutate(country_name_fct = forcats::fct_relevel(country_name_fct, top_20$country_name, "Other")) |>
+  group_by(country_name_fct, country_code, cat) |>
+  summarise(all = sum(all),
+            oa = sum(oa),
+            ta_oa = sum(ta_oa), .groups = "drop") |>
+  head(20) |>
+  mutate(cat_labels = case_when(
+   cat == "OECD" ~ glue::glue("<span style = 'color: #3fc1c9'>{country_code}</span>"),
+   cat == "BRICS" ~ glue::glue("<span style = 'color: #fc5185'>{country_code}</span>"),
+   cat == "Other" ~ glue::glue("<span style = 'color: #e1e1e1'>{country_code}</span>")
+  )) |>
+  mutate(cat_labels = fct_inorder(cat_labels))
+
+top_20_plot <- top_20_df |>
+  mutate(no_ta = oa - ta_oa) |>
+  pivot_longer(
+  cols = c(no_ta, ta_oa),
+  values_to = "value",
+  names_to = "type") |>
+  ggplot(aes(reorder(as.character(country_code),  value), value / sum(value), fill = type)) +
+    geom_bar(stat = "identity", alpha = 0.8, color = "#000000", linewidth = 0.2) +
+   scale_fill_manual(
+    NULL,
+    values = c(no_ta = "#b3b3b3", ta_oa = "#56B4E9"),
+    labels = c(no_ta = "Other", ta_oa = "Agreement"),
+    guide = FALSE) +
+  coord_flip() +
+  scale_y_continuous(
+    labels = scales::percent_format(accuracy = 1L),
+    expand = expansion(mult = c(0, 0.05)),
+    breaks = scales::breaks_extended(4)
+  ) +
+  theme_minimal() +
+#  theme(axis.text.y = ggtext::element_markdown()) +
+  labs(x = NULL, y = "% of Hybrid OA published\nin 2022")
+
+```
+
+```{r}
+country_scatter_22 <- country_scatter_df |>
+  filter(cr_year == "2022", !is.na(country_code)) |>
+  ggplot(aes(all, oa)) +
+  geom_point(aes(fill = cat, size = ta_oa), pch = 21, alpha = 0.7) +
+  scale_fill_manual(NULL, values = c("OECD" = "#3fc1c9", "BRICS" = "#fc5185", "Other" = "#e1e1e1")) +
+  geom_smooth(aes(all, oa), method=lm) +
+  scale_size("TA OA", guide = "none") +
+  ggrepel::geom_text_repel(
+    data = filter(country_scatter_df, cr_year == 2022, cat %in% c("OECD", "BRICS"), ta_oa > 5000 | country_code %in% c("CN", "IN")),
+    aes(x = all, y = oa, label = country_code),
+    force = 5,
+    segment.color = "grey30",
+    segment.size = 0.2,
+    segment.alpha = 0.8,
+    color = "grey10", 
+    size = 3,
+    box.padding = 0.3) +
+  scale_x_continuous(labels =  scales::number_format(big.mark = ","), guide = guide_axis(angle = 45)) +
+  scale_y_continuous(labels =  scales::number_format(big.mark = ",")) +
+  theme_minimal() +
+  labs(x = "Total articles in hybrid journals in 2022", y = "OA articles") +
+  theme(legend.position = "top",
+    legend.justification = "left",
+    legend.direction = "horizontal")
+
+oecd_stats <- readr::read_csv(here::here("data", "oecd_stats.csv"))
+oecd_stats_df <- oecd_stats |>
+  filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
+  filter(issn_l %in% hoaddata::cc_articles$issn_l) |>
+  group_by(cr_year, cat) |>
+  summarise(all = sum(articles, na.rm = TRUE), oa = sum(oa,na.rm = TRUE), ta_oa = sum(ta_oa, na.rm = TRUE)) |>
+  mutate(closed = all - oa,
+         oa_other = oa - ta_oa) |>
+  pivot_longer(cols = c(ta_oa, oa_other, closed), names_to = "oa_type") |>
+  mutate(cat = ifelse(is.na(cat), "Other", cat))
+
+country_bar <- ggplot(oecd_stats_df, aes(gsub("^20", "'", as.character(cr_year)), value, group = oa_type, fill = oa_type, color = oa_type)) +
+  geom_area(color = "#000000",
+            alpha = 0.8,
+            linewidth = 0.2) +
+  scale_color_manual(values = colorspace::darken(c(oa_other = "#b3b3b3", ta_oa = "#56B4E9", closed = "#e1e1e1"), 0.3)) +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels =  scales::number_format(big.mark = ",")) +
+  facet_wrap(~ cat) +
+  scale_fill_manual("Hybrid OA", values = c(oa_other = "#b3b3b3", ta_oa = "#56B4E9", closed = "#e1e1e1"),
+                    labels = c(oa_other = "Other", ta_oa = "Agreement", closed = "NA"),  guide = guide_legend(reverse = TRUE)) +
+  theme_minimal() +
+  labs(x = "Publication year", y = "Total articles")
+
+```
+
+```{r country_patch, fig.asp=0.9}
+library(patchwork)
+wrap_plots(
+  A = country_bar,
+  B = country_scatter_22,
+  C = top_20_plot,
+  design = "
+  AC
+  BC",
+#  guides = "collect",
+  widths = c(0.7, 0.3),
+  heights = c(0.5,0.5)
+) &
+  theme(
+    panel.grid.minor = element_blank(),
+                panel.border = element_rect(color = "grey50", fill = NA),
+    axis.title=element_text(size = 10),
+    legend.position = "top",
+    legend.justification = "left",
+    legend.direction = "horizontal") &
+    plot_annotation(tag_levels = 'A') 
+```
+
 [^springercompact]: <https://web.archive.org/web/20180414062853id_/http://www.liber2015.org.uk/wp-content/uploads/2015/03/Springer-Compact.pdf>

### 2024-01-05 analysis/001_overview.Rmd
Algin plot design

--- 2023-12-22 analysis/001_overview.Rmd
@@
   # plot
-oa_ta_plot <- ggplot(oa_ta_plot_df, aes(cr_year, value / n, fill = type, group = type)) +
-  geom_area() +
-  scale_fill_manual(NULL, values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"), labels = c(no_ta = "Other", ta_oa = "Agreement"),  guide = guide_legend(reverse = TRUE)) +
-  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels = scales::percent_format(), limits = c(0, .15), breaks = c(0, .05, .1, .15)) +
+oa_ta_plot <-
+  ggplot(oa_ta_plot_df, aes(cr_year, value / n, fill = type, group = type)) +
+  geom_area(color = "#000000",
+            linewidth = 0.15) +
+  scale_fill_manual(
+    NULL,
+    values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"),
+    labels = c(no_ta = "Other", ta_oa = "Agreement"),
+    guide = guide_legend(reverse = TRUE)
+  ) +
+  scale_y_continuous(
+    expand = expansion(mult = c(0, 0.05)),
+    labels = scales::percent_format(),
+    limits = c(0, .15),
+    breaks = c(0, .05, .1, .15)
+  ) +
   theme_minimal() +
   labs(y = "Hybrid OA Uptake", x = NULL) +
-  theme(legend.position="top", legend.justification = "left", axis.text.x = element_blank())
+  theme(
+    legend.position = "top",
+    legend.justification = "left",
+    axis.text.x = element_blank()
+  )
 ```
@@
   ggplot(aes(gsub("^20", "'", as.character(cr_year)), value, group = 1)) +
-  geom_bar(stat = "identity", fill = "#56B4E9") +
+  geom_bar(stat = "identity", fill = "#56B4E9", color = "#000000", linewidth = 0.15) +
   scale_y_continuous(expand = expansion(mult = c(0, 0.05)), breaks = c(0, 0.25, 0.5),
@@
   labs(y = NULL, x = "Publication Year")
-
 ```
@@
 
-publisher_prop <- ggplot(publisher_summary_df, aes(gsub("^20", "'", as.character(cr_year)), prop, fill = type, group = type)) +
-  geom_area() +
-  facet_wrap(~publisher, nrow = 1) +
-  scale_fill_manual("Hybrid OA", values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"), labels = c(no_ta = "Other", ta_oa = "Agreement"),  guide = guide_legend(reverse = TRUE)) +
-  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels = scales::percent_format(),
-                     breaks = c(0.05, 0.1, 0.15, 0.2, 0.25)) +
+publisher_prop <-
+  ggplot(publisher_summary_df, aes(
+    gsub("^20", "'", as.character(cr_year)),
+    prop,
+    fill = type,
+    group = type
+  )) +
+  geom_area(color = "#000000",
+            linewidth = 0.15) +
+  facet_wrap( ~ publisher, nrow = 1) +
+  scale_fill_manual(
+    "Hybrid OA",
+    values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"),
+    labels = c(no_ta = "Other", ta_oa = "Agreement"),
+    guide = guide_legend(reverse = TRUE)
+  ) +
+  scale_y_continuous(
+    expand = expansion(mult = c(0, 0.05)),
+    labels = scales::percent_format(),
+    breaks = c(0.05, 0.1, 0.15, 0.2, 0.25)
+  ) +
   theme_minimal() +
   labs(y = "Hybrid OA Uptake", x = NULL) +
-  theme(legend.position="top", legend.justification = "left", axis.text.x = element_blank())
+  theme(
+    legend.position = "top",
+    legend.justification = "left",
+    axis.text.x = element_blank()
+  )
 
@@
   ggplot(aes(gsub("^20", "'", as.character(cr_year)), value / jn_all, fill = type)) +
-  geom_boxplot(outlier.shape = NA, show.legend = FALSE, varwidth = TRUE) +
+  geom_boxplot(outlier.shape = NA,
+               show.legend = FALSE,
+               varwidth = TRUE) +
   scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
                      labels = scales::percent) +
-  scale_fill_manual("Hybrid OA", values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"), labels = c(no_ta = "Other", ta_oa = "Agreement")) +
-  coord_cartesian(ylim = c(0,0.65)) +
-  facet_wrap(~publisher, nrow = 1) +
+  scale_fill_manual(
+    "Hybrid OA",
+    values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"),
+    labels = c(no_ta = "Other", ta_oa = "Agreement")
+  ) +
+  coord_cartesian(ylim = c(0, 0.65)) +
+  facet_wrap( ~ publisher, nrow = 1) +
   labs(y = "OA Uptake\nby Hybrid Journal", x = "Publication Year") +
@@
   theme(strip.background = element_blank(),
-    strip.text.x = element_blank()
-    )
+        strip.text.x = element_blank())
 
@@
   ggplot(aes(gsub("^20", "'", as.character(cr_year)), prop)) +
-  geom_bar(stat = "identity", fill = "#56B4E9") +
-  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), breaks = c(0, 0.25,0.5,0.75), limits = c(0, .75),
-                     labels = scales::percent) +
-  facet_wrap(~publisher, nrow = 1) +
+  geom_bar(stat = "identity",
+           fill = "#56B4E9",
+            color = "#000000",
+           linewidth = 0.15) +
+  scale_y_continuous(
+    expand = expansion(mult = c(0, 0.05)),
+    breaks = c(0, 0.25, 0.5, 0.75),
+    limits = c(0, .75),
+    labels = scales::percent
+  ) +
+  facet_wrap( ~ publisher, nrow = 1) +
   theme_minimal() +
@@
     strip.background = element_blank(),
-    strip.text.x = element_blank())
+    strip.text.x = element_blank()
+  )
 ```
@@
   )) +
-  geom_area() +
+    geom_area(color = "#000000",
+            linewidth = 0.15) +
   facet_wrap( ~ top_level, nrow = 1) +
@@
   ggplot(aes(gsub("^20", "'", as.character(cr_year)), prop)) +
-  geom_bar(stat = "identity", fill = "#56B4E9") +
+  geom_bar(stat = "identity", fill = "#56B4E9", color = "#000000",
+            linewidth = 0.15) +
   scale_y_continuous(expand = expansion(mult = c(0, 0.05)), breaks = c(0, 0.25,0.5,0.75), limits = c(0, .75),
@@
   ungroup() |>
-  # Highlight Germany
   mutate(total_string = case_when(
@@
     facet_labels = glue::glue(
-      '{country_name_fct}<br>
+      '**{country_name_fct}**<br>
       {total_string} | {paste(round(prop * 100, 1), "%")}'
@@
   )) +
-  geom_area(stat = "identity") +
+  geom_area(color = "#000000",
+            linewidth = 0.15,
+            stat = "identity") +
 scale_fill_manual(
@@
     axis.title=element_text(size = 10),
-    strip.text = ggtext::element_markdown(size = 6),
+    strip.text = ggtext::element_markdown(size = 5.5),
     plot.subtitle = ggtext::element_markdown(hjust = 0.5),
@@
 
+```
+
+```{r country_bar_chart_top_20}
+top_20_df <- country_scatter_df |>
+  filter(cr_year == "2022") |>
+  mutate(country_name_fct = forcats::fct_other(country_name, keep = top_20$country_name)) |>
+  mutate(country_name_fct = forcats::fct_relevel(country_name_fct, top_20$country_name, "Other")) |>
+  group_by(country_name_fct, country_code, cat) |>
+  summarise(all = sum(all),
+            oa = sum(oa),
+            ta_oa = sum(ta_oa), .groups = "drop") |>
+  head(20) |>
+  mutate(cat_labels = case_when(
+   cat == "OECD" ~ glue::glue("<span style = 'color: #3fc1c9'>{country_code}</span>"),
+   cat == "BRICS" ~ glue::glue("<span style = 'color: #fc5185'>{country_code}</span>"),
+   cat == "Other" ~ glue::glue("<span style = 'color: #e1e1e1'>{country_code}</span>")
+  )) |>
+  mutate(cat_labels = fct_inorder(cat_labels))
+
+top_20_plot <- top_20_df |>
+  mutate(no_ta = oa - ta_oa) |>
+  pivot_longer(
+  cols = c(no_ta, ta_oa),
+  values_to = "value",
+  names_to = "type") |>
+  ggplot(aes(reorder(as.character(country_code),  value), value / sum(value), fill = type)) +
+    geom_bar(stat = "identity", color = "#000000", linewidth = 0.15) +
+   scale_fill_manual(
+    NULL,
+    values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"),
+    labels = c(no_ta = "Other", ta_oa = "Agreement"),
+    guide = FALSE) +
+  coord_flip() +
+  scale_y_continuous(
+    labels = scales::percent_format(accuracy = 1L),
+    expand = expansion(mult = c(0, 0.05)),
+    breaks = scales::breaks_extended(4)
+  ) +
+  theme_minimal() +
+#  theme(axis.text.y = ggtext::element_markdown()) +
+  labs(x = NULL, y = "% of Hybrid OA published\nin 2022")
+
+```
+
+```{r}
 country_scatter_22 <- country_scatter_df |>
-  filter(cr_year == "2022", !is.na(country_code)) |>
+  filter(cr_year == "2022", !is.na(country_code), all > 10000) |>
   ggplot(aes(all, oa)) +
-  geom_point(aes(color = cat, size = ta_oa), alpha = 0.8) +
-  scale_color_manual(NULL, values = c("OECD" = "#3fc1c9", "BRICS" = "#fc5185", "Other" = "#e1e1e1")) +
+  geom_point(aes(fill = cat, size = ta_oa), pch = 21, alpha = 0.7) +
+  scale_fill_manual(NULL, values = c("OECD" = "#d01c8b", "BRICS" = "#4dac26", "Other" = "#f7f7f7")) +
   geom_smooth(aes(all, oa), method=lm) +
-  scale_size("TA OA", guide = "none") +
+  scale_size("Agreement", guide = FALSE) +
   ggrepel::geom_text_repel(
-    data = filter(country_scatter_df, cr_year == 2022, cat %in% c("OECD", "BRICS"), all > 50000),
+    data = filter(country_scatter_df, cr_year == 2022, ta_oa > 5000 | country_code %in% c("CN", "IN", "BR", "RU", "IR")),
     aes(x = all, y = oa, label = country_code),
-    force = 50,
-    box.padding = 0.5,
-    point.padding = 0.8,
+            min.segment.length = .15,
+    force = 5,
+    segment.color = "grey10",
+    segment.size = 0.5,
+    segment.alpha = 0.8,
+    color = "grey10", 
     size = 3,
-    max.iter = 1000000,
-    max.overlaps = Inf,
-    min.segment.length = 0,
-    arrow = arrow(length = unit(0.010, "npc"))) +
-  scale_x_continuous(labels =  scales::number_format(big.mark = ","), guide = guide_axis(angle = 45)) +
-  scale_y_continuous(labels =  scales::number_format(big.mark = ",")) +
-  theme_minimal() +
-  labs(x = "Total articles", y = "OA articles in hybrid journals") +
-  theme(legend.position = "top",
-    legend.justification = "left",
-    legend.direction = "horizontal")
+    box.padding = 0.3) +
+  scale_y_log10(labels =  scales::number_format(big.mark = ",")) +
+  scale_x_log10(labels =  scales::number_format(big.mark = ",")) +
+  theme_minimal() +
+  labs(x = "Total articles in hybrid journals in 2022", y = "OA articles")
 
@@
 country_bar <- ggplot(oecd_stats_df, aes(gsub("^20", "'", as.character(cr_year)), value, group = oa_type, fill = oa_type)) +
-  geom_area() +
+  geom_area(color = "#000000",
+            linewidth = 0.15) +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels =  scales::number_format(big.mark = ",")) +
   facet_wrap(~ cat) +
-  scale_fill_manual(NULL, values = c(oa_other = "#b3b3b3a0", ta_oa = "#56B4E9", closed = "grey50"),
-                    labels = c(oa_other = "OA Other", ta_oa = "Agreement", closed = "Closed")) +
-  theme_minimal() +
-  scale_y_continuous(labels =  scales::number_format(big.mark = ",")) +
-  theme_minimal() +
-  labs(x = "Publication year", y = "Total articles") +
-  theme(legend.position = "top",
-        legend.justification = "left",
-        legend.direction = "horizontal")
-
-```
-
-```{r country_patch}
+  scale_fill_manual("Hybrid OA", values = c(oa_other = "#b3b3b3a0", ta_oa = "#56B4E9", closed = "#e1e1e1a0"),
+                    labels = c(oa_other = "Other", ta_oa = "Agreement", closed = "NA"),  guide = guide_legend(reverse = TRUE)) +
+  theme_minimal() +
+  labs(x = "Publication year", y = "Total articles")
+
+```
+
+```{r country_patch, fig.asp=0.9}
 library(patchwork)
 wrap_plots(
-  A = country_scatter_22,
-  B = country_bar,
-  design = "AB",
-  widths = c(0.35, 0.65)
+  A = country_bar,
+  B = country_scatter_22,
+  C = top_20_plot,
+  design = "
+  AC
+  BC",
+ # guides = "collect",
+  widths = c(0.7, 0.3),
+  heights = c(0.5,0.5)
 ) &
@@
                 panel.border = element_rect(color = "grey50", fill = NA),
-    axis.title=element_text(size = 10)) &
+    axis.title=element_text(size = 10),
+    legend.position = "top", 
+    legend.justification = "right") &
     plot_annotation(tag_levels = 'A') 

### 2024-01-16 analysis/001_overview.Rmd
Start describing subject

@@
 
-Next, I focus on the subject distribution.
-
 ```{r subject_data_prep}
@@
 
-Here's a table. I applied fractional counting because of large overlaps.
+Table presents a high-level overview of hybrid open access by AJCS subject area using fractionalised counting to account for journals belonging to more than one category. 
+Between 2018 and 2022, most hybrid journals with at least one open articles could be attributed to the social sciences including the humanities. 
+However, these journals published the fewest number of articles, whereas physical sciences journals recorded most articles (n = , 44%), followed by the health sciences and the life sciences. 
+In terms of open access,  physical sciences journals accounted for more than one third of articles published in the five-years period, followed by the health science, the social sciences and the life sciences.
 
@@
   scale_fill_manual(
-    "Hybrid OA",
+    NULL,
     values = c(
@@
 ```
+
+Figure presents the relative growth of hybrid open access by subject area between 2018-2022.
+In particular, journals in the social sciences accounted for the strongest growth by from % to %, followed by the Life Science , Health Science and Physical Sciences.
+This growth in the social sciences can be largely attributed to transformative agreements, which are prevalent across journals in this field. In 2022, 75% , (median).
+However, it is noteworthy that hybrid open access through transformative agreements plays a comparable lesser role in the life sciences and health sciences. 
+In these two subject areas, only about half of the open access articles could be linked to transformative agreements, both overall and on median average across journals.
+In contrast to the Life Science , Health Science, the majority of physical science journals, recorded an above-average open access uptake through transformative agreements.
 
@@
 ```
-
-And here's my country analysis. First top 20
 
@@
 
+### Supplements
+
+```{r country_by_publisher}
+country_jn_all <- hoaddata::jn_aff |>
+  filter(!cr_year %in% c(2017, 2023), issn_l != "0027-8424") |>
+  # only with cc
+  filter(issn_l %in% hoaddata::cc_articles$issn_l) |>
+  distinct(issn_l, cr_year, country_code, articles_total) |>
+  group_by(cr_year, country_code, issn_l) |>
+  summarise(all = sum(articles_total))
+
+country_jn_oa <- hoaddata::cc_articles |>
+  filter(!cr_year %in% c(2017, 2023), issn_l != "0027-8424") |>
+  group_by(cr_year, country_code, issn_l) |>
+  summarise(oa = n_distinct(doi))
+
+country_codes <- readr::read_csv(here::here("data", "ror_country_codes.csv"))
+country_jn_ta <-  ta_oa_inst |>
+  filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
+  left_join(country_codes) |>
+  group_by(cr_year, country_code, issn_l) |>
+  summarise(ta_oa = n_distinct(doi))
+
+country_jn_df <- country_jn_all |>
+  left_join(country_jn_oa, by = c("cr_year", "country_code", "issn_l")) |>
+  left_join(country_jn_ta, by = c("cr_year", "country_code", "issn_l")) |>
+  distinct() |>
+  ungroup() |>
+  mutate(cr_year = as.character(cr_year)) |>
+  mutate(across(c(oa, ta_oa), .fns = ~replace_na(.,0))) |>
+  filter(!is.na(country_code)) |>
+  mutate(
+    country_name = countrycode::countrycode(country_code, origin = "iso2c", destination = "country.name")
+  )
+
+## add publishers
+country_by_publisher_df <- hoaddata::jct_hybrid_jns |>
+  distinct(issn_l, esac_publisher) |>
+  inner_join(country_jn_df, by = "issn_l") |>
+  # Publisher encoding
+   mutate(
+    publisher = case_when(
+      esac_publisher == "Elsevier" ~ "Elsevier",
+      esac_publisher == "Springer Nature" ~ "Springer Nature",
+      esac_publisher == "Wiley" ~ "Wiley",
+      is.character(esac_publisher) ~ "Other"
+    )
+  ) |>
+  mutate(publisher =
+           forcats::fct_relevel(
+             publisher,
+             c("Elsevier", "Springer Nature", "Wiley", "Other") 
+           )) |>
+  # Country encoding
+   mutate(cat = case_when(
+    country_code %in% oecd$country_code ~ "OECD",
+    country_code %in% c("BR", "CN", "IN", "RU", "ZA") ~ "BRICS",
+    .default = "Other"
+  )) |>
+  # Summary
+  group_by(cr_year, publisher, cat) |>
+  summarise(all = sum(all),
+            oa = sum(oa),
+            ta_oa = sum(ta_oa), .groups = "drop") |>
+  mutate(closed = all - oa,
+         oa_other = oa - ta_oa)
+
+# Plot
+country_by_publisher_df |>
+  pivot_longer(cols = c(ta_oa, oa_other, closed), names_to = "oa_type") |>
+  ggplot(aes(gsub("^20", "'", as.character(cr_year)), value, group = oa_type, fill = oa_type)) +
+  geom_area(color = "#000000",
+            linewidth = 0.15) +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels =  scales::number_format(big.mark = ",")) +
+  facet_grid(cat ~publisher) +
+  scale_fill_manual("Hybrid OA", values = c(oa_other = "#b3b3b3a0", ta_oa = "#56B4E9", closed = "#e1e1e1a0"),
+                    labels = c(oa_other = "Other", ta_oa = "Agreement", closed = "NA"),  guide = guide_legend(reverse = TRUE)) +
+  theme_minimal() +
+  labs(x = "Publication year", y = "Total articles")
+```
+
 [^springercompact]: <https://web.archive.org/web/20180414062853id_/http://www.liber2015.org.uk/wp-content/uploads/2015/03/Springer-Compact.pdf>

### 2024-01-17 analysis/001_overview.Rmd
Subject first draft

@@
 Between 2018 and 2022, most hybrid journals with at least one open articles could be attributed to the social sciences including the humanities.
-However, these journals published the fewest number of articles, whereas physical sciences journals recorded most articles (n = , 44%), followed by the health sciences and the life sciences. 
+However, these journals published the fewest number of articles, whereas physical sciences journals recorded most articles, followed by the health sciences and the life sciences. 
 In terms of open access,  physical sciences journals accounted for more than one third of articles published in the five-years period, followed by the health science, the social sciences and the life sciences.
@@
 Figure presents the relative growth of hybrid open access by subject area between 2018-2022.
-In particular, journals in the social sciences accounted for the strongest growth by from % to %, followed by the Life Science , Health Science and Physical Sciences.
-This growth in the social sciences can be largely attributed to transformative agreements, which are prevalent across journals in this field. In 2022, 75% , (median).
-However, it is noteworthy that hybrid open access through transformative agreements plays a comparable lesser role in the life sciences and health sciences. 
-In these two subject areas, only about half of the open access articles could be linked to transformative agreements, both overall and on median average across journals.
-In contrast to the Life Science , Health Science, the majority of physical science journals, recorded an above-average open access uptake through transformative agreements.
+In particular, Social Sciences and Humanties journals accounted for the strongest growth in the five-years period from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Social Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Social Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(value) |> sum()`), followed by the Life Science from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Life Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Life Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Life Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Life Sciences") |> pull(value) |> sum()`) , Health Science from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Health Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Health Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Health Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Health Sciences") |> pull(value) |> sum()`) and Physical Sciences from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Physical Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Physical Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Physical Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Physical Sciences") |> pull(value) |> sum()`).
+<!-- In particular, journals in the subject areas recorded large high rate (see supplement)-->
+This growth in the social sciences can be largely attributed to transformative agreements.
+In 2022, two-third of open access articles (`r subject_ta_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(prop) * 100`%, n = `r subject_ta_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(ta_oa_frac)`) could be attributed to transformative agreements.
+Figure C shows that this trend is consistent across Social Sciences journals.
+In 2022, 25% of Social Science journals were able to faciliate open access to every fourth article soley through transformative agreements.
+However, it is noteworthy that hybrid open access through transformative agreements played a comparable lesser role in the Life Sciences and Health Sciences. 
+In these two subject areas, only about half of the open access articles can be linked to transformative agreements, both overall and on median average across journals.
+In contrast, the majority of Physical Science Journals, shows an increase of open access through transformative agreements compared to other options to publish open access in hybrid journals.
 

### 2024-01-18 analysis/001_overview.Rmd
start writing country comparision

@@
 
-What are the number and proportion of open access articles in hybrid journals published under a transformative agreement?
+# Results
 
@@
 ```
+
+## Overview
 
@@
 
+## Publishing market
+
 ```{r}
@@
   scale_fill_manual(
-    "Hybrid OA",
+    NULL,
     values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"),
@@
 ```
+
+## Journal subjects
 
@@
 Figure C shows that this trend is consistent across Social Sciences journals.
-In 2022, 25% of Social Science journals were able to faciliate open access to every fourth article soley through transformative agreements.
+In 2022, 25% of Social Science journals were able to faciliate open access to at least every fourth article soley through transformative agreements.
 However, it is noteworthy that hybrid open access through transformative agreements played a comparable lesser role in the Life Sciences and Health Sciences. 
@@
 ```{r, country_plot, fig.asp=0.8}
-ggplot(top_20_plot_df, aes(
+country_plot <- ggplot(top_20_plot_df, aes(
     gsub("^20", "'", as.character(cr_year)),
@@
 scale_fill_manual(
-    "Hybrid OA",
+    NULL,
     values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"),
@@
 
-here a scatterplot and a classifciation of oecd / brics / other. first, the scatterplot 2022!
+## Comparing countries
 
@@
 
+oecd_stats <- readr::read_csv(here::here("data", "oecd_stats.csv"))
+oecd_stats_global <- oecd_stats |> 
+  filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
+  filter(issn_l %in% hoaddata::cc_articles$issn_l) |>
+  group_by(cat) |> summarise(articles = sum(articles), oa = sum(oa), ta_oa = sum(ta_oa))
 ```
@@
   facet_wrap(~ cat) +
-  scale_fill_manual("Hybrid OA", values = c(oa_other = "#b3b3b3a0", ta_oa = "#56B4E9", closed = "#e1e1e1a0"),
+  scale_fill_manual(NULL, values = c(oa_other = "#b3b3b3a0", ta_oa = "#56B4E9", closed = "#e1e1e1a0"),
                     labels = c(oa_other = "Other", ta_oa = "Agreement", closed = "NA"),  guide = guide_legend(reverse = TRUE)) +
@@
 ```
+
+When analysing the country affiliations of first authors, open access publishing in hybrid journals including through transformative agreements was almost exclusively prevalent in western economies between 2018 and 2022.
+To highlight this trend, Figure 2A groups the country affiliation according to its membership in the intergovernmental organisations BRICS as of 2022, comprising the five countries Brazil, Russia, India, China and South Africa, and in the Organisation for Economic Co-operation and Development (OECD), consisting of 38 membership countries.
+The residual category "Other" comprises the remaining countries.
+Together, OECD countries recorded `r oecd_stats_global |> filter(cat == "OECD") |> pull(oa)` open access articles in hybrid journals between 2018 and 2022, representing `r oecd_stats_global |> filter(cat == "OECD") |> pull(oa) / n_oa * 100`% of the investigated open access articles in hybrid journals in the five-year period.
+The gap between OECD and the other countries is even more pronounced  when it comes to open access through transformative agreements. 
+Overall, `r oecd_stats_global |> filter(cat == "OECD") |> pull(ta_oa)` of `r n_ta_oa` open access articles were linked to an agreement, representing `r oecd_stats_global |> filter(cat == "OECD") |> pull(ta_oa) / n_ta_oa*100`%.
+Focusing on publications from first authors affiliated with institutions based in an OECD country reveals an absolute and relative growth of hybrid open access. 
+In the five-year period, the proportion of open access in hybrid journals increased from `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2018) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2018 to `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2022) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2022.
+In contrast, BRICS recorded low growth rates, from `r country_bar$data |> filter(cat == "BRICS", oa_type != "closed", cr_year == 2018) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2018 to `r country_bar$data |> filter(cat == "BRICS", oa_type != "closed", cr_year == 2022) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2022.
+
+It is worth noting that while the number of open access articles increased across OECD countries, the overall publication output decreased sharply, dropping to xxxx in 2022 from a peak of xxx articles in 2020.
+In stark contrast, the number of articles published in hybrid journals by first authors affiliated with institutions from BRICS countries increased steadily over the years. 
+Upon inspection, this trend can be observed across publishers, although the shift towards BRICS is particularly prevalent across Elsevier's hybrid journal portfolio.
+Upon closer examination, this trend can be observed across publishers, although the shift towards BRICS is particularly evident in Elsevier's hybrid journal portfolio.
+
+figure b compares oa articles with total articles in hybrid journals in the final year of your study. Trend line indicates
+it highligths that in 2022, china has been the most prodi´cutive country, followed by the usa and india. it shows taht mainly the west and european coutrnie, in particular above a
+
+in tersm of market share the uk alone contrib by .
+surpringsly, with not conevrage  ndd.
+countries with gugh rates of ar ein particular nethrrlnacd sewden and suisse.
+the former two already introduced long ihsitrory of
+
+sCountries wirh compatrabel large open access
+
+it furthermore highligths oecd countries an
 

### 2024-01-18 analysis/001_overview.Rmd
tmp save

@@
   labs(x = "Publication year", y = "Total articles")
-
+```
+
+```{r country_by_publisher}
+publisher_mapping <-  hoaddata::jct_hybrid_jns |>
+  distinct(issn_l, esac_publisher) 
+subject_mapping <- subject_df |>
+  distinct(issn_l, top_level)
+country_jn_all <- oecd_stats |>
+  filter(!cr_year %in% c(2017, 2023), issn_l != "0027-8424") |>
+  # only with cc
+  filter(issn_l %in% hoaddata::cc_articles$issn_l) |>
+  inner_join(publisher_mapping, by = "issn_l") |>
+  inner_join(subject_mapping, by = "issn_l") |>
+  # Publisher encoding
+   mutate(
+    publisher = case_when(
+      esac_publisher == "Elsevier" ~ "Elsevier",
+      esac_publisher == "Springer Nature" ~ "Springer Nature",
+      esac_publisher == "Wiley" ~ "Wiley",
+      is.character(esac_publisher) ~ "Other"
+    )
+  ) |>
+  mutate(publisher =
+           forcats::fct_relevel(
+             publisher,
+             c("Elsevier", "Springer Nature", "Wiley", "Other") 
+           ))  |> 
+  group_by(cr_year, publisher, cat, top_level) |>
+  summarise(all = sum(articles),
+            oa = sum(oa),
+            ta_oa = sum(ta_oa), .groups = "drop") |>
+  mutate(closed = all - oa,
+         oa_other = oa - ta_oa)
+
+publ_subject_output_df <- country_jn_all |>
+  distinct(cr_year, publisher, top_level, cat, all) |>
+  filter(!top_level %in% c("Multidisciplinary", NA))
+
+pub_subject_output_plot <-  ggplot(publ_subject_output_df, aes(gsub("^20", "'", as.character(cr_year)), all, fill = cat)) +
+  geom_bar(stat = "identity", position = position_dodge2(preserve = "single")) +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels =  scales::number_format(big.mark = ",")) +
+  scale_fill_manual(NULL, values = c("OECD" = "#d01c8b", "BRICS" = "#4dac26", "Other" = "#f7f7f7")) +
+  facet_grid(publisher ~ top_level) +
+  theme_minimal() +
+  labs(x = "Publication year", y = "Total articles")
+
+subject_output_plot <- publ_subject_output_df |>
+  group_by(cr_year, top_level, publisher) |>
+  summarise(all = sum(all)) |>
+  ggplot(aes(cr_year, all)) +
+  geom_bar(stat = "identity") +
+  facet_grid(top_level ~ publisher) +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels =  scales::number_format(big.mark = ",")) +
+   theme_minimal() +
+  labs(x = "Publication year", y = "Total articles")
+
+publisher_output_plot <-
+  publ_subject_output_df |>
+  group_by(cr_year, cat, publisher) |>
+  summarise(all = sum(all)) |>
+  ggplot(aes(cr_year, all)) +
+  geom_bar(stat = "identity") +
+  facet_grid(cat ~ publisher) +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels =  scales::number_format(big.mark = ",")) +
+   theme_minimal() +
+  labs(x = "Publication year", y = "Total articles")
 ```
@@
 Upon inspection, this trend can be observed across publishers, although the shift towards BRICS is particularly prevalent across Elsevier's hybrid journal portfolio.
-Upon closer examination, this trend can be observed across publishers, although the shift towards BRICS is particularly evident in Elsevier's hybrid journal portfolio.
-
-figure b compares oa articles with total articles in hybrid journals in the final year of your study. Trend line indicates
-it highligths that in 2022, china has been the most prodi´cutive country, followed by the usa and india. it shows taht mainly the west and european coutrnie, in particular above a
-
-in tersm of market share the uk alone contrib by .
-surpringsly, with not conevrage  ndd.
-countries with gugh rates of ar ein particular nethrrlnacd sewden and suisse.
-the former two already introduced long ihsitrory of
-
-sCountries wirh compatrabel large open access
-
-it furthermore highligths oecd countries an
+Upon closer examination, this trend can be observed across publishers, although the shift towards BRICS is particularly evident in Elsevier's hybrid journal portfolio, in particular with regard to articles published in Physical Sciences journals (see Supplement).
+While OECD publication output in Elsevier's Physical Sciences journals declined from `r publ_subject_output_df |> filter(cr_year == 2018, top_level == "Physical Sciences", publisher == "Elsevier", cat == "OECD") |> pull(all)` articles in 2018 to `r publ_subject_output_df |> filter(cr_year == 2022, top_level == "Physical Sciences", publisher == "Elsevier", cat == "OECD") |> pull(all)` in 2022, BRICS output increased from `r publ_subject_output_df |> filter(cr_year == 2018, top_level == "Physical Sciences", publisher == "Elsevier", cat == "BRICS") |> pull(all)` to `r publ_subject_output_df |> filter(cr_year == 2022, top_level == "Physical Sciences", publisher == "Elsevier", cat == "BRICS") |> pull(all)` in the same five-year period.
+But also the OECD publication output in Health Science Journals and Life Science journals stagnated across after a peak in 2020.
+
+Figure B and C presents a breakdown by countries for the publication year 2022.
+Figure B compares total publication output with the number of open access articles published during the final year investigated, highlighting that in 2022, china has been the most productive country, followed by the USA and India.
+However, West and Nord European countries like recorded a comparable high number of open access articles, mainly due to transformative agreements. 
+Nonetheless, China and the US recorded large open access market shares of xxx resp. xxx, although country-wide transformative agreements were not implemented.
+
+Figure highlights the development of hybrid open access and shows the top 20 most productive countries in terms of articles published in hybrid journals included in transformative agreements between 2018 and 2022.
+In particular, the UK, Netherlands, and Sweden were hybrid affin, wheras in othe rcountries like Gemrany a large grwoth could be observed from 2020 obwards where the large DEAL agreements have becpome in place.
+The high open access rates in Netherlands ,s witzerlands, and Sweden conforms the braod implementation of agreemnts across a lrge number publishers where ....
+whereas in othe rcountries, agreements were limited to a fewer number of publishers.
+Take for isntance germany, were a deal with Elsevier was not in place in 2022.
+But also publication limits or of elegible inistutions and article types might contribute to the fact taht even in countries with a broad impleemntation, 100% is not reached.
+Again, the figure points to countries with low hybrid open access, in particular  non OECD countries.
 
@@
 
-```{r country_by_publisher}
-country_jn_all <- hoaddata::jn_aff |>
-  filter(!cr_year %in% c(2017, 2023), issn_l != "0027-8424") |>
-  # only with cc
-  filter(issn_l %in% hoaddata::cc_articles$issn_l) |>
-  distinct(issn_l, cr_year, country_code, articles_total) |>
-  group_by(cr_year, country_code, issn_l) |>
-  summarise(all = sum(articles_total))
-
-country_jn_oa <- hoaddata::cc_articles |>
-  filter(!cr_year %in% c(2017, 2023), issn_l != "0027-8424") |>
-  group_by(cr_year, country_code, issn_l) |>
-  summarise(oa = n_distinct(doi))
-
-country_codes <- readr::read_csv(here::here("data", "ror_country_codes.csv"))
-country_jn_ta <-  ta_oa_inst |>
-  filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
-  left_join(country_codes) |>
-  group_by(cr_year, country_code, issn_l) |>
-  summarise(ta_oa = n_distinct(doi))
-
-country_jn_df <- country_jn_all |>
-  left_join(country_jn_oa, by = c("cr_year", "country_code", "issn_l")) |>
-  left_join(country_jn_ta, by = c("cr_year", "country_code", "issn_l")) |>
-  distinct() |>
-  ungroup() |>
-  mutate(cr_year = as.character(cr_year)) |>
-  mutate(across(c(oa, ta_oa), .fns = ~replace_na(.,0))) |>
-  filter(!is.na(country_code)) |>
-  mutate(
-    country_name = countrycode::countrycode(country_code, origin = "iso2c", destination = "country.name")
-  )
-
-## add publishers
-country_by_publisher_df <- hoaddata::jct_hybrid_jns |>
-  distinct(issn_l, esac_publisher) |>
-  inner_join(country_jn_df, by = "issn_l") |>
-  # Publisher encoding
-   mutate(
-    publisher = case_when(
-      esac_publisher == "Elsevier" ~ "Elsevier",
-      esac_publisher == "Springer Nature" ~ "Springer Nature",
-      esac_publisher == "Wiley" ~ "Wiley",
-      is.character(esac_publisher) ~ "Other"
-    )
-  ) |>
-  mutate(publisher =
-           forcats::fct_relevel(
-             publisher,
-             c("Elsevier", "Springer Nature", "Wiley", "Other") 
-           )) |>
-  # Country encoding
-   mutate(cat = case_when(
-    country_code %in% oecd$country_code ~ "OECD",
-    country_code %in% c("BR", "CN", "IN", "RU", "ZA") ~ "BRICS",
-    .default = "Other"
-  )) |>
-  # Summary
-  group_by(cr_year, publisher, cat) |>
-  summarise(all = sum(all),
-            oa = sum(oa),
-            ta_oa = sum(ta_oa), .groups = "drop") |>
-  mutate(closed = all - oa,
-         oa_other = oa - ta_oa)
-
-# Plot
-country_by_publisher_df |>
-  pivot_longer(cols = c(ta_oa, oa_other, closed), names_to = "oa_type") |>
-  ggplot(aes(gsub("^20", "'", as.character(cr_year)), value, group = oa_type, fill = oa_type)) +
-  geom_area(color = "#000000",
-            linewidth = 0.15) +
-  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels =  scales::number_format(big.mark = ",")) +
-  facet_grid(cat ~publisher) +
-  scale_fill_manual("Hybrid OA", values = c(oa_other = "#b3b3b3a0", ta_oa = "#56B4E9", closed = "#e1e1e1a0"),
-                    labels = c(oa_other = "Other", ta_oa = "Agreement", closed = "NA"),  guide = guide_legend(reverse = TRUE)) +
-  theme_minimal() +
-  labs(x = "Publication year", y = "Total articles")
+```{r}
+pub_subject_output_plot
 ```

### 2024-01-19 analysis/001_overview.Rmd
rerender

@@
 In particular, the UK, Netherlands, and Sweden were hybrid affin, wheras in othe rcountries like Gemrany a large grwoth could be observed from 2020 obwards where the large DEAL agreements have becpome in place.
-The high open access rates in Netherlands ,s witzerlands, and Sweden conforms the braod implementation of agreemnts across a lrge number publishers where ....
-whereas in othe rcountries, agreements were limited to a fewer number of publishers.
+The high open access rates in Netherlands ,s witzerlands, and Sweden conforms the braod implementation of agreemnts across a lrge number publishers where whereas in othe rcountries, agreements were limited to a fewer number of publishers.
 Take for isntance germany, were a deal with Elsevier was not in place in 2022.

### 2024-01-22 analysis/001_overview.Rmd
Results

@@
 ```{r}
-#| fig-cap: "Growth of open access in hybrid journals through transformative agreements between 2018 and 2022 per publication year.
-The blue areas represent open access through transformative agreements, as indicated by the affiliation with institutions that had a corresponding transformative agreement in place at the time of publication (according to the first author affiliation data in OpenAlex matched with journal and institutional data from cOAlition S Transformative Agreements Public Data).
-(A) Proportion of open access articles in hybrid journals per year (according to the total number of articles indexed in Crossref).
+#| fig-cap: "Relative growth of open access in hybrid journals through transformative agreements between 2018 and 2022 per publication year  according to the total number of articles indexed in Crossref.
+The blue areas represent open access through transformative agreements, as indicated by the affiliation with institutions that had a corresponding transformative agreement in place at the time of publication according to the first author affiliation data in OpenAlex matched with journal and institutional data from cOAlition S Transformative Agreements Public Data.
+Grey areas depict open access articles where no link to an agreement could be established.
+(A) Proportion of open access articles in hybrid journals per year.
 (B) Percentage of hybrid open access via agreements per year.
 Boxplots show the proportion of open access articles by individual hybrid journals (C) and individual open access uptake rates by individual hybrid journals and open access funding (D) per publication year.
-Horizontal lines denote lower quartile, median, and upper quartile, with whiskers extending to ±1.5 of interquartile range.
+Horizontal lines correspond to the lower quartile, median, and upper quartile, with whiskers extending to ±1.5 of interquartile range.
 The individual outliers are not shown.
@@
 This growth in the social sciences can be largely attributed to transformative agreements.
-In 2022, two-third of open access articles (`r subject_ta_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(prop) * 100`%, n = `r subject_ta_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(ta_oa_frac)`) could be attributed to transformative agreements.
+In 2022, two-third of open access articles (`r subject_ta_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(prop) * 100`%, n = `r subject_ta_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(ta_oa_frac)`) were published by first authors affiliated with participating institutions.
 Figure C shows that this trend is consistent across Social Sciences journals.
-In 2022, 25% of Social Science journals were able to faciliate open access to at least every fourth article soley through transformative agreements.
-However, it is noteworthy that hybrid open access through transformative agreements played a comparable lesser role in the Life Sciences and Health Sciences. 
-In these two subject areas, only about half of the open access articles can be linked to transformative agreements, both overall and on median average across journals.
+In 2022, 25% of Social Science journals provided open access to at least every fourth article exclusively through transformative agreements.
+However, hybrid open access through transformative agreements played a comparable lesser role in the Life Sciences and Health Sciences. 
+In these two subject areas, only about half of the open access articles can be linked to these agreements, both overall and on median average across journals.
 In contrast, the majority of Physical Science Journals, shows an increase of open access through transformative agreements compared to other options to publish open access in hybrid journals.
@@
 country_multiple_df <- country_df |>
-  mutate(country_name_fct = forcats::fct_other(country_name, keep = top_20$country_name)) |>
-  mutate(country_name_fct = forcats::fct_relevel(country_name_fct, top_20$country_name, "Other")) |>
+  mutate(country_name_fct = forcats::fct_other(country_code, keep = top_20$country_code)) |>
+  mutate(country_name_fct = forcats::fct_relevel(country_name_fct, top_20$country_code, "Other")) |>
   group_by(country_name_fct, cr_year) |>
@@
     axis.title=element_text(size = 10),
-    strip.text = ggtext::element_markdown(size = 5.5),
+    strip.text = ggtext::element_markdown(size = 7),
     plot.subtitle = ggtext::element_markdown(hjust = 0.5),
@@
 
-When analysing the country affiliations of first authors, open access publishing in hybrid journals including through transformative agreements was almost exclusively prevalent in western economies between 2018 and 2022.
-To highlight this trend, Figure 2A groups the country affiliation according to its membership in the intergovernmental organisations BRICS as of 2022, comprising the five countries Brazil, Russia, India, China and South Africa, and in the Organisation for Economic Co-operation and Development (OECD), consisting of 38 membership countries.
-The residual category "Other" comprises the remaining countries.
-Together, OECD countries recorded `r oecd_stats_global |> filter(cat == "OECD") |> pull(oa)` open access articles in hybrid journals between 2018 and 2022, representing `r oecd_stats_global |> filter(cat == "OECD") |> pull(oa) / n_oa * 100`% of the investigated open access articles in hybrid journals in the five-year period.
-The gap between OECD and the other countries is even more pronounced  when it comes to open access through transformative agreements. 
-Overall, `r oecd_stats_global |> filter(cat == "OECD") |> pull(ta_oa)` of `r n_ta_oa` open access articles were linked to an agreement, representing `r oecd_stats_global |> filter(cat == "OECD") |> pull(ta_oa) / n_ta_oa*100`%.
-Focusing on publications from first authors affiliated with institutions based in an OECD country reveals an absolute and relative growth of hybrid open access. 
-In the five-year period, the proportion of open access in hybrid journals increased from `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2018) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2018 to `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2022) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2022.
-In contrast, BRICS recorded low growth rates, from `r country_bar$data |> filter(cat == "BRICS", oa_type != "closed", cr_year == 2018) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2018 to `r country_bar$data |> filter(cat == "BRICS", oa_type != "closed", cr_year == 2022) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2022.
-
-It is worth noting that while the number of open access articles increased across OECD countries, the overall publication output decreased sharply, dropping to xxxx in 2022 from a peak of xxx articles in 2020.
-In stark contrast, the number of articles published in hybrid journals by first authors affiliated with institutions from BRICS countries increased steadily over the years. 
-Upon inspection, this trend can be observed across publishers, although the shift towards BRICS is particularly prevalent across Elsevier's hybrid journal portfolio.
-Upon closer examination, this trend can be observed across publishers, although the shift towards BRICS is particularly evident in Elsevier's hybrid journal portfolio, in particular with regard to articles published in Physical Sciences journals (see Supplement).
-While OECD publication output in Elsevier's Physical Sciences journals declined from `r publ_subject_output_df |> filter(cr_year == 2018, top_level == "Physical Sciences", publisher == "Elsevier", cat == "OECD") |> pull(all)` articles in 2018 to `r publ_subject_output_df |> filter(cr_year == 2022, top_level == "Physical Sciences", publisher == "Elsevier", cat == "OECD") |> pull(all)` in 2022, BRICS output increased from `r publ_subject_output_df |> filter(cr_year == 2018, top_level == "Physical Sciences", publisher == "Elsevier", cat == "BRICS") |> pull(all)` to `r publ_subject_output_df |> filter(cr_year == 2022, top_level == "Physical Sciences", publisher == "Elsevier", cat == "BRICS") |> pull(all)` in the same five-year period.
-But also the OECD publication output in Health Science Journals and Life Science journals stagnated across after a peak in 2020.
-
-Figure B and C presents a breakdown by countries for the publication year 2022.
-Figure B compares total publication output with the number of open access articles published during the final year investigated, highlighting that in 2022, china has been the most productive country, followed by the USA and India.
-However, West and Nord European countries like recorded a comparable high number of open access articles, mainly due to transformative agreements. 
-Nonetheless, China and the US recorded large open access market shares of xxx resp. xxx, although country-wide transformative agreements were not implemented.
-
-Figure highlights the development of hybrid open access and shows the top 20 most productive countries in terms of articles published in hybrid journals included in transformative agreements between 2018 and 2022.
-In particular, the UK, Netherlands, and Sweden were hybrid affin, wheras in othe rcountries like Gemrany a large grwoth could be observed from 2020 obwards where the large DEAL agreements have becpome in place.
-The high open access rates in Netherlands ,s witzerlands, and Sweden conforms the braod implementation of agreemnts across a lrge number publishers where whereas in othe rcountries, agreements were limited to a fewer number of publishers.
-Take for isntance germany, were a deal with Elsevier was not in place in 2022.
-But also publication limits or of elegible inistutions and article types might contribute to the fact taht even in countries with a broad impleemntation, 100% is not reached.
-Again, the figure points to countries with low hybrid open access, in particular  non OECD countries.
+Between 2018 and 2022, Western economies almost exclusively dominated hybrid open access publishing through transformative agreements. 
+During this period, first-authors affiliated with institutions from Organisation for Economic Co-operation and Development (OECD) member countries published `r oecd_stats_global |> filter(cat == "OECD") |> pull(oa)` open access articles in hybrid journals, representing `r oecd_stats_global |> filter(cat == "OECD") |> pull(oa) / n_oa * 100`% of the investigated open access articles.
+This disparity between OECD nations and other countries becomes even more evident when considering open access through transformative agreements, as  `r oecd_stats_global |> filter(cat == "OECD") |> pull(ta_oa)` of `r n_ta_oa`, or `r oecd_stats_global |> filter(cat == "OECD") |> pull(ta_oa) / n_ta_oa*100`% of open access articles were associated with such agreements.
 
@@
 
+Figure 2A shows the development of hybrid open access publishing by countries, comparing the OECD area with the BRICS, an intergovernmental organisation,  which comprised the countries Brazil, Russia, India, China and South Africa in 2022.
+The residual category "Other" includes the remaining countries. 
+From 2018 to 2022, the proportion of open access in hybrid journals increased from `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2018) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2018 to `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2022) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2022.
+On the other hand, BRICS recorded an low uptake, from `r country_bar$data |> filter(cat == "BRICS", oa_type != "closed", cr_year == 2018) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2018 to `r country_bar$data |> filter(cat == "BRICS", oa_type != "closed", cr_year == 2022) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2022.
+Despite rise of open access across OECD countries, the overall publication output decreased sharply, dropping to `r country_bar$data |> filter(cat == "OECD", cr_year == "2022") |> distinct(all) |> pull(all)` in 2022 after peaking `r country_bar$data |> filter(cat == "OECD", cr_year == "2020") |> distinct(all) |> pull(all)` articles in 2020.
+In stark contrast, the number of articles published in hybrid journals by first authors affiliated with institutions from BRICS countries increased steadily over the years, more than doubling from `r country_bar$data |> filter(cat == "BRICS", cr_year == "2018") |> distinct(all) |> pull(all)` in 2018 to `r country_bar$data |> filter(cat == "OECD", cr_year == "2022") |> distinct(all) |> pull(all)` in 2022.
+Upon closer examination, this trend can be observed across all big three publishers, although the shift towards BRICS is particularly evident in Elsevier's hybrid journal portfolio, in particular with regard to articles published in Physical Sciences journals (see Supplement).
+While OECD publication output in Elsevier's Physical Sciences journals declined from `r publ_subject_output_df |> filter(cr_year == 2018, top_level == "Physical Sciences", publisher == "Elsevier", cat == "OECD") |> pull(all)` articles in 2018 to `r publ_subject_output_df |> filter(cr_year == 2022, top_level == "Physical Sciences", publisher == "Elsevier", cat == "OECD") |> pull(all)` in 2022, BRICS output increased from `r publ_subject_output_df |> filter(cr_year == 2018, top_level == "Physical Sciences", publisher == "Elsevier", cat == "BRICS") |> pull(all)` to `r publ_subject_output_df |> filter(cr_year == 2022, top_level == "Physical Sciences", publisher == "Elsevier", cat == "BRICS") |> pull(all)` in the same five-year period.
+Furthermore, OECD publication output in Health Science Journals and Life Science journals stagnated across the investigated hybrid journal portfolios after a peak in 2020.
+
+To illustrate the situation in 2022, Figure B compares total publication output with the number of open access articles. 
+With `r country_scatter_df |> filter(cr_year == "2022", country_code == "CN") |> pull(all)` articles, China was the most productive country, followed by the United States (`r country_scatter_df |> filter(cr_year == "2022", country_code == "US") |> pull(all)` articles) and India (`r country_scatter_df |> filter(cr_year == "2022", country_code == "IN") |> pull(all)` articles).
+In contrast, West and Nord European countries published a considerable high number of open access articles, mainly due to transformative agreements. 
+Particularly, Germany, Great Britain, the Netherlands, Sweden, Switzerland and Spain recorded an above-average open access share as indicated by the linear trend line.
+As represented by the point size, as well as it can been seen in Figure C, transformative agreements contributed to this market position of these countries. 
+Interestingly, the United States had a notable open access market share of `r top_20_plot$data |> distinct(country_code, oa, all) |> mutate(prop = oa / sum(oa)) |> filter(country_code == "US") |> pull(prop) * 100`%, although transformative agreements contributed to a lesser extent.
+Similarly, China's open access market share of `r top_20_plot$data |> distinct(country_code, oa, all) |> mutate(prop = oa / sum(oa)) |> filter(country_code == "CN") |> pull(prop) * 100` in 2022 was comparable to that of the Netherlands, which was (`r  top_20_plot$data |> distinct(country_code, oa, all) |> mutate(prop = oa / sum(oa)) |> filter(country_code == "NL") |> pull(prop) * 100`%,).
+
+```{r country_top_20_plot, fig.asp=0.8}
+country_plot
+```
+
+Figure illustrates the development of hybrid open access from 2018 to 2022, highlighting the top 20 most productive countries in terms of articles published in hybrid journals that were included in transformative agreements over the five-year period.
+Notably, The Netherlands (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "NL") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), Sweden (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "SE") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), Poland (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "PL") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%) and Great Britain (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "GB") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%)) exhibited a relatively high level of uptake in 2018 which continued to increase in the following years.
+In 2022, Sweden had the highest proportion of open-access articles relative to its publication output (`r country_plot$data |> filter(cr_year == "2022", country_name_fct == "SE") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), followed by the Netherlands (`r country_plot$data |> filter(cr_year == "2022", country_name_fct == "NL") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%) and Switzerland (`r country_plot$data |> filter(cr_year == "2022", country_name_fct == "CH") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), with these countries benefiting from transformative agreements.
+In Germany, however, hybrid open access only began to increase from 2019 onwards after the successful negotiation of nationwide agreements with Wiley (July 2019) and Springer Nature (January 2020).
+Prior to this, only a few organisations had agreements in place, for the example the Max Planck Society with Springer Compact.
+
+Since 2021, there has been a general trend towards hybrid open access among many western countries, primarily driven by transformative agreements.
+However, proliferation of transformative agreements differe across these countries.
+For instance, Germany successfully negotiated an agreement with Elsevier not until 2023. 
+Additionally, publication limits or eligibility criteria for institutions and article types may explain why even countries with widespread agreement implementation do not achieve 100% hybrid open access.
+Interestingly, in Japan and the US other options than transformative agreements were the main driver for the increase in hybrid open access.
+Once again, the graph highlights countries with low hybrid open access, particularly non-OECD countries, where only a few or no agreements were in place.
+
 ### Supplements

### 2024-01-26 data/002_total_article_volume.R
Explore hoaddata share relative to journal articles in cr

@@ -1,0 +1,84 @@
+#' Proportion of total articles in Crossref by journal business model
+#' 
+library(bigrquery)
+library(tidyverse)
+#' 
+my_sql <- "WITH
+  hoad AS (
+  SELECT
+    doi,
+    'hoad' AS src
+  FROM
+    `hoa-article.hoaddata_nov23.cc_md` AS hoad ),
+  crossref AS (
+  SELECT
+    cr.doi,
+    EXTRACT ( YEAR
+    FROM
+      issued ) AS cr_year,
+    publisher,
+    src
+  FROM
+    `subugoe-collaborative.cr_instant.snapshot` AS cr
+  LEFT JOIN
+    hoad
+  ON
+    cr.doi = hoad.doi
+  WHERE
+    type = 'journal-article'
+    AND NOT REGEXP_CONTAINS( title, '^Author Index$|^Back Cover|^Contents$|^Contents:|^Corrigendum|^Cover Image|^Cover Picture|^Editorial Board|^Front Cover|^Frontispiece|^Inside Back Cover|^Inside Cover|^Inside Front Cover|^Issue Information|^List of contents|^Masthead|^Title page|^Correction$|^Corrections to|^Corrections$|^Withdrawn|^Frontmatter' )
+    AND ( NOT REGEXP_CONTAINS(page, '^S')
+      OR page IS NULL ) -- include online only articles, lacking page or issue
+    AND ( NOT REGEXP_CONTAINS(issue, '^S')
+      OR issue IS NULL ) )
+SELECT
+  COUNT(DISTINCT crossref.doi) AS n_articles,
+  cr_year,
+  src,
+  open_access.oa_status
+FROM
+  crossref
+LEFT JOIN
+  `subugoe-collaborative.openalex.works` AS openalex
+ON
+  crossref.doi = openalex.doi
+WHERE
+  cr_year BETWEEN 2018
+  AND 2022
+GROUP BY
+  cr_year,
+  src,
+  open_access.oa_status
+ORDER BY
+  n_articles DESC"
+
+tb <- bq_project_query("subugoe-collaborative", my_sql)
+cr_stats <- bq_table_download(tb)
+
+my_df <- cr_stats |>
+  mutate(cat = case_when(
+    is.na(src) & oa_status == "gold" ~ "gold",
+    src == "hoad" ~ "hoad",
+    .default = "other"
+  )) |> 
+  group_by(cr_year, cat) |>
[…]

### 2024-01-28 analysis/manuscript.Rmd
Writing methods

--- 2024-01-22 analysis/001_overview.Rmd
@@
 ---
-title: "Overview"
-output: github_document
-always_allow_html: true
-date: "`r Sys.Date()`"
+output:
+  pdf_document:
+    fig_caption: yes
+    keep_tex: yes
+    template: apa6.sty
+csl: apa.csl
+bibliography: references.bib
 ---
@@
 ```
+# Introduction
+
+# Methods and data
+
+This study combines data from multiple publicly available data sources as visualised in Figure 1. 
+Data gathering involved transformative agreement data retrieved from the cOAlition S Journal Checker Tool, which provided information about journal portfolios and participating institutions. 
+After identification of hybrid journals, Crossref served as the primary data source for article-level metadata including Creative Commons (CC) license information to determine open access articles on publisher websites.
+OpenAlex provided first author affiliation metadata that were subsequently linked to the transformative agreement data to identify open access articles published through these agreements. 
+In total, the compiled data set consists of `r hoaddata::jn_ind |> filter(!cr_year %in% c("2017", "2023")) |> distinct(issn_l, cr_year, jn_all) |> pull(jn_all) |> sum()` articles published in `r hoaddata::jn_ind |> filter(!cr_year %in% c("2017", "2023")) |> distinct(issn_l) |> nrow()` hybrid journals between 2018 and 2022.
+
+```{r data_workflow, echo=FALSE, fig.cap="Data collection workflow to obtain article-level transformative agreement data"}
+knitr::include_graphics(here::here("analysis", "data_collection_workflow.pdf"))
+```
+
+## Transformative agreement and hybrid journal data
+
+```{r journal_oa_exclusion}
+oa_journal_status <- readr::read_csv(here::here("data", "oa_journal_status.csv"))
+
+# OA journal total
+n_oa_journals <- oa_journal_status %>%
+  filter(openalex == TRUE | doaj == TRUE | bielefeld == TRUE) %>% 
+  distinct(issn_l) %>%
+  nrow()
+# By source
+#
+# GOLD OA
+bielefeld <- oa_journal_status %>%
+  filter(bielefeld == TRUE) %>% 
+  distinct(issn_l)  %>%
+  pull()
+# DOAJ
+doaj <- oa_journal_status %>%
+  filter(doaj == TRUE) %>% 
+  distinct(issn_l) %>%
+  pull()
+# OpenAlex venue
+openalex <- oa_journal_status %>%
+  filter(openalex == TRUE) %>% 
+  distinct(issn_l) %>%
+  pull()
+my_venn <- list(`GOLD OA` = bielefeld, `DOAJ` = doaj, `OpenAlex` = openalex)
+
+p <-
+  ggvenn::ggvenn(
+    my_venn,
+    fill_color = c("grey55",
+          "steelblue2",
+          "lightsalmon"),
+    digits = 0,
+    text_color = "black",
+    text_size = 3,
+    set_name_size = 4
+  )
+```
+
+Data gathering started with obtaining journals included in transformative agreements using the cOAlition S’ Journal Checker Tool publicly available Transformative Agreement Data dump. 
+The dump consisted of multiple online Google spreadsheets where each data file represented one agreement listed in the ESAC Transformative Agreement Registry. 
+Each file comprised information about journals and institutions involved per agreement. 
+Because the Journal Checker Tool only collected data about active transformative agreements, four different snapshots were used for this analysis:the archived version from July 2021, July 2022, May 2023 and the most current dump downloaded on 11 December 2022.
+This ensured that transformative agreements, which ended in 2021, were also included in the analysis. 
+Agreements ending before 2021 could not be obtained.
+Overall, the combined Transformative Agreement Data dumps contained `r  length(unique(hoaddata::jct_hybrid_jns$esac_id))` agreements.
+
+The Transformative Agreement Data dumps linked agreements to journals represented by journal names and ISSNs. 
+After mapping ISSN variants to the corresponding linking ISSN (ISSN-L) as provided by The ISSN International Centre.
+Journals were associated to publishers using the ESAC ID, a unique identifier for transformative agreements in the ESAC Transformative Agreement Registry.
+
+Because transformative agreements can include both full open access and hybrid journals, journal data were complemented with information about a journal's open access status using multiple sources: the Directory of Open Access Journals (DOAJ) downloaded on 4 October 2022, OpenAlex (September 2022 dump) and the the Bielefeld list of GOLD OA journals.
+In total, `r n_oa_journals` full open access journals were excluded.
+Figure shows a substantial overlap between the three data sources of
+`r round(length(unique(Reduce(intersect, my_venn))) / length(unique(Reduce(union, my_venn)))* 100,0)`%.
+However, combining different data sources considerably extended the journal matching to exclude full open access journals from transformative agreements.
+The Gold OA journals dataset alone adds `r length(unique(Reduce(setdiff, my_venn)))` journals, while the DOAJ comprised `r length(setdiff(my_venn$DOAJ, union(my_venn[["GOLD OA"]], my_venn$OpenAlex)))` full open access journals not listed in either of the other two sources.
+These full open access journals were mostly launched in 2022.
+
+## Article metadata
+
+## Data Analysis
 
@@
   kbl(
+    booktabs = TRUE, format.args = list(big.mark = ","), digits = 1, escape = TRUE, position = "H",
     caption ="Hybrid open access through transformative agreements market shares 2018-2022",
@@
       "C%"
-    ),
-    format.args = list(big.mark = ',')
-  ) %>%
-  kable_classic() %>%
+    )
+  ) |>
+  kable_styling(font_size = 10, position = "center") |>
   add_header_above(c(
@@
   kbl(
+    booktabs = TRUE, format.args = list(big.mark = ","), digits = 1, escape = TRUE, position = "H",
     caption ="Hybrid open access through transformative agreements by journal subject 2018-2022",
@@
       "C%"
-    ),
-    format.args = list(big.mark = ',')
-  ) %>%
-  kable_classic() %>%
+    )
+  ) |>
   add_header_above(c(

### 2024-01-28 analysis/manuscript.tex
Writing methods

--- 2024-01-22 analysis/001_overview.Rmd
@@
----
-title: "Overview"
-output: github_document
-always_allow_html: true
-date: "`r Sys.Date()`"
----
+\documentclass[a4paper,man,floatsintext,longtable,noextraspace,12pt]{apa6}
 
-```{r, setup, echo=FALSE}
-knitr::opts_chunk$set(
-  comment = "#>",
-  collapse = TRUE,
-  warning = FALSE,
-  message = FALSE,
-  echo = FALSE,
-  fig.width = 6,
-  fig.asp = 0.618,
-  out.width = "99%",
-  fig.align = "center",
-  dpi = 300,
-  dev = c("png", "pdf"),
-  fig.path = "fig/"
-)
-options(scipen = 999, digits = 2)
-knitr::knit_hooks$set(
-  inline = function(x) {
-    if (is.numeric(x)) {
-      return(prettyNum(x, big.mark = ","))
-    } else{
-      return(x)
+\usepackage[english]{babel}
+\usepackage[utf8x]{inputenc}
+\usepackage{amsmath}
+\usepackage{graphicx}
+\usepackage[colorinlistoftodos]{todonotes}
+\usepackage{hyperref}
+
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
+% make captions italic
+
+% bibliography
+
+% tightlist
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+\title{\textbf{How open are hybrid journals included in transformative agreements?}}
+\shorttitle{Hybrid OA}
+\author{Najko Jahn}
+\affiliation{Göttingen State and University Library, University of Göttingen\\
+Platz der Göttinger Sieben 1, 37073 Göttingen, Germany\\
+najko.jahn@sub.uni-goettingen.de
 }
+
+%\authornote{Correspondence concerning this article should be addressed to Najko Jahn}
+
+\abstract{}
+
+\begin{document}
+\maketitle
+
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
+
+\hypertarget{methods-and-data}{%
+\section{Methods and data}\label{methods-and-data}}
+
+This study combines data from multiple publicly available data sources
+as visualised in Figure 1. Data gathering involved transformative
+agreement data retrieved from the cOAlition S Journal Checker Tool,
+which provided information about journal portfolios and participating
+institutions. After identification of hybrid journals, Crossref served
+as the primary data source for article-level metadata including Creative
+Commons (CC) license information to determine open access articles on
+publisher websites. OpenAlex provided first author affiliation metadata
+that were subsequently linked to the transformative agreement data to
+identify open access articles published through these agreements. In
+total, the compiled data set consists of 8,942,669 articles published in
+12,858 hybrid journals between 2018 and 2022.
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.99\linewidth]{data_collection_workflow}
+
 }
-)
-```
 
-```{r dependencies}
-library(tidyverse, warn.conflicts = FALSE)
-library(kableExtra)
-```
+\caption{Data collection workflow to obtain article-level transformative agreement data}\label{fig:data_workflow}
+\end{figure}
 
-```{r data, cache=TRUE}
-library(hoaddata)
+\hypertarget{transformative-agreement-and-hybrid-journal-data}{%
+\subsection{Transformative agreement and hybrid journal
+data}\label{transformative-agreement-and-hybrid-journal-data}}
 
-ta_oa_inst <- readr::read_csv(here::here("data", "ta_oa_inst.csv.gz"))
+Data gathering started with obtaining journals included in
+transformative agreements using the cOAlition S' Journal Checker Tool
+publicly available Transformative Agreement Data dump. The dump
+consisted of multiple online Google spreadsheets where each data file
+represented one agreement listed in the ESAC Transformative Agreement
+Registry. Each file comprised information about journals and
+institutions involved per agreement. Because the Journal Checker Tool
+only collected data about active transformative agreements, four
+different snapshots were used for this analysis:the archived version
+from July 2021, July 2022, May 2023 and the most current dump downloaded
+on 11 December 2022. This ensured that transformative agreements, which
+ended in 2021, were also included in the analysis. Agreements ending
+before 2021 could not be obtained. Overall, the combined Transformative
+Agreement Data dumps contained 729 agreements.
 
-```
+The Transformative Agreement Data dumps linked agreements to journals
+represented by journal names and ISSNs. After mapping ISSN variants to
+the corresponding linking ISSN (ISSN-L) as provided by The ISSN
+International Centre. Journals were associated to publishers using the
+ESAC ID, a unique identifier for transformative agreements in the ESAC
+Transformative Agreement Registry.
 
-# Results
+Because transformative agreements can include both full open access and
+hybrid journals, journal data were complemented with information about a
+journal's open access status using multiple sources: the Directory of
+Open Access Journals (DOAJ) downloaded on 4 October 2022, OpenAlex
+(September 2022 dump) and the the Bielefeld list of GOLD OA journals. In
+total, 3,439 full open access journals were excluded. Figure shows a
+substantial overlap between the three data sources of 72\%. However,
+combining different data sources considerably extended the journal
+matching to exclude full open access journals from transformative
+agreements. The Gold OA journals dataset alone adds 176 journals, while
+the DOAJ comprised 10 full open access journals not listed in either of
+the other two sources. These full open access journals were mostly
+launched in 2022.
 
-```{r overview_data_prep}
-ta_oa_growth <- ta_oa_inst |>
-  filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
-  group_by(cr_year) |>
-  summarise(ta_oa = n_distinct(doi))
-hybrid_oa_growth <- hoaddata::cc_articles |>
-  filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
-  group_by(cr_year) |>
-  summarise(oa_all = n_distinct(doi))
-ta_df <- inner_join(ta_oa_growth, hybrid_oa_growth) |>
-  mutate(ta_oa_prop = ta_oa / oa_all) |>
-  mutate(no_ta = oa_all - ta_oa, 
-         no_ta_prop = 1 - ta_oa_prop) |>
-  mutate(cr_year = as.factor(cr_year))
-# year all
-year_all <-
-  hoaddata::jn_ind |>
-  filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
-  filter(issn_l %in% hoaddata::cc_articles$issn_l) |>
-  distinct(issn_l, cr_year, jn_all) |> 
-  group_by(cr_year) |> 
-  summarise(n = sum(jn_all))
+\hypertarget{article-metadata}{%
+\subsection{Article metadata}\label{article-metadata}}
 
-## Hybrid Journals
- n_hybrid_jns_all <- hoaddata::jn_ind |>
-  filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
-  distinct(issn_l) |>
-   nrow()
+\hypertarget{data-analysis}{%
+\subsection{Data Analysis}\label{data-analysis}}
 
-## Hybrid Journals with at least one OA article
-n_hybrid_jns <- hoaddata::cc_articles |>
-  filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
-  group_by(issn_l) |>
-  summarise(oa_all = n_distinct(doi)) |> 
-  nrow()
+\hypertarget{results}{%
+\section{Results}\label{results}}
 
-## Hybrid Journals with at least one OA article
-n_ta_jns <- ta_oa_inst |>
-  filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
-  distinct(issn_l) |>
-  nrow()
+\hypertarget{overview}{%
+\subsection{Overview}\label{overview}}
 
-## Hybrid Journals with at least one TA-OA article
-n_tas_with_oa <- ta_oa_inst |>
-  filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
-  distinct(ta_journal_portfolio) |>
-  nrow()
+Between 2018 and 2022, a total of 11,189 out of 12,857 hybrid journals
+in transformative agreements published at least one open access article
+under a Creative Commons license. During this period, these hybrid
+journals provided open access to 742,369 out of 8,146,958 articles,
+representing a five-year open access proportion of 9.1\%. Authors who
+could make use of transformative agreements at the time of publication
+contributed 328,957 open access articles to the total.
 
-## Open access volume
-n_oa <-  hoaddata::cc_articles |>
-  filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
-  distinct(doi) |>
-  nrow()
+\begin{figure}
 
-## Total
-n_total <- hoaddata::jn_ind |>
-  filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
-  filter(issn_l %in% hoaddata::cc_articles$issn_l) |>
-  distinct(issn_l, cr_year, jn_all) |>
-  pull(jn_all) |>
-  sum()
+{\centering \includegraphics[width=0.99\linewidth]{fig/unnamed-chunk-2-1}
 
-## TA total
-n_tas_with_oa <- ta_oa_inst |>
-  filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
-  distinct(ta_journal_portfolio) |>
-  nrow()
+}
 
-## TA article total
-n_ta_oa <- ta_oa_inst |>
-  filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
-  distinct(doi) |>
-  nrow()
-```
-
-## Overview
-
-Between 2018 and 2022, a total of `r n_hybrid_jns` out of `r n_hybrid_jns_all` hybrid journals in transformative agreements published at least one open access article under a Creative Commons license.
-During this period, these hybrid journals provided open access to `r n_oa` out of `r n_total` articles, representing a five-year open access proportion of `r n_oa / n_total * 100`%.
-Authors who could make use of transformative agreements at the time of publication contributed `r n_ta_oa` open access articles to the total.
-
-```{r area_oa_over_the_years}
-oa_ta_plot_df <- ta_df |>
-  select(cr_year, ta_oa, no_ta) |>
-  pivot_longer(cols = c(ta_oa, no_ta), names_to = "type", values_to = "value") |>
-  filter(!cr_year %in%  c("2017", "2023")) |>
-  inner_join(year_all)
-  # plot
-oa_ta_plot <-
-  ggplot(oa_ta_plot_df, aes(cr_year, value / n, fill = type, group = type)) +
-  geom_area(color = "#000000",
-            linewidth = 0.15) +
-  scale_fill_manual(
-    NULL,
-    values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"),
-    labels = c(no_ta = "Other", ta_oa = "Agreement"),
-    guide = guide_legend(reverse = TRUE)
-  ) +
-  scale_y_continuous(
-    expand = expansion(mult = c(0, 0.05)),
-    labels = scales::percent_format(),
-    limits = c(0, .15),
-    breaks = c(0, .05, .1, .15)
-  ) +
-  theme_minimal() +
-  labs(y = "Hybrid OA Uptake", x = NULL) +
-  theme(
-    legend.position = "top",
-    legend.justification = "left",
-    axis.text.x = element_blank()
-  )
-```
-
-```{r ta_share_plot}
-ta_share_plot <- ta_df |>
-  select(cr_year, ta_oa_prop, no_ta_prop) |>
-  pivot_longer(cols = c(ta_oa_prop, no_ta_prop), names_to = "type", values_to = "value") |>
-  filter(!cr_year %in%  c("2017", "2023"), type == "ta_oa_prop") |>
-  ggplot(aes(gsub("^20", "'", as.character(cr_year)), value, group = 1)) +
-  geom_bar(stat = "identity", fill = "#56B4E9", color = "#000000", linewidth = 0.15) +
-  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), breaks = c(0, 0.25, 0.5),
-                     labels = scales::percent) +
-  theme_minimal() +
-  labs(y = "Percentage\nof Hybrid OA\nvia Agreement", x = "Publication Year")
-```
-
-```{r journal_boxplot}
-jn_hybrid_oa_growth <- hoaddata::cc_articles |>
-  filter(issn_l != "0027-8424") |>
-  group_by(cr_year, issn_l) |>
-  summarise(oa_all = n_distinct(doi)) |>
-  mutate(cr_year = factor(cr_year))
-# year all
-jn_year_all <-
-  hoaddata::jn_ind |>
-  filter(issn_l != "0027-8424") |>
-  filter(issn_l %in% hoaddata::cc_articles$issn_l) |>
-  distinct(issn_l, cr_year, jn_all) |> 
-  group_by(cr_year, issn_l) |> 
-  summarise(n = sum(jn_all))
-
-oa_jn_box_df <- left_join(jn_hybrid_oa_growth, jn_year_all) |>
-  mutate(prop = oa_all / n) |> 
-  filter(!cr_year %in%  c("2017", "2023"))
-oa_jn_box <- oa_jn_box_df |>
-  ggplot(aes(gsub("^20", "'", as.character(cr_year)), prop)) +
-  geom_boxplot(outlier.shape = NA) +
-  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
-                     labels = scales::percent) +
-  coord_cartesian(ylim = c(0,0.65)) +
-  theme_minimal() +
-  labs(y = "OA Uptake per Hybrid Journal", x = "Publication Year")
-```
-
-```{r}
-ta_oa_growth <- ta_oa_inst |>
-  filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424") |>
-  group_by(cr_year, issn_l) |>
-  summarise(ta_oa = n_distinct(doi))
-hybrid_oa_growth <- hoaddata::cc_articles |>
-  filter(issn_l != "0027-8424") |>
-  group_by(cr_year, issn_l) |>
-  summarise(oa_all = n_distinct(doi))
-ta_df <- left_join(hybrid_oa_growth, ta_oa_growth) |>
-  mutate(across(everything(), .fns = ~replace_na(.,0))) |>
-  mutate(no_ta = oa_all - ta_oa) |>
-  mutate(cr_year = as.factor(cr_year))
-
-summary_all <- inner_join(ta_df, jn_year_all, by = c("cr_year", "issn_l")) |>
-  mutate(prop_ta_oa = ta_oa / n,
-         prop_no_ta = no_ta / n) |>
-  ungroup() |>
-  distinct(cr_year, issn_l, prop_ta_oa, prop_no_ta) |>
-  pivot_longer(cols = c(prop_ta_oa, prop_no_ta))
-# plot
-box_ta_prop <- summary_all |>
-  filter(!cr_year %in%  c("2017", "2023")) |>
-  # With at least one OA article
-  filter(value > 0) |>
-  ggplot(aes(gsub("^20", "'", as.character(cr_year)), value, fill = name)) +
-  geom_boxplot(outlier.shape = NA, varwidth = TRUE) +
-  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
-                     labels = scales::percent) +
-  coord_cartesian(ylim = c(0,0.65)) +
-  scale_fill_manual(NULL, values = c(prop_no_ta = "#b3b3b3a0", prop_ta_oa = "#56B4E9"), labels = c(prop_no_ta = "Other", prop_ta_oa = "Agreement"), guide = FALSE) +
-  theme_minimal() +
-  labs(y = NULL, x = "Publication Year")
-```
-
-```{r}
-#| fig-cap: "Relative growth of open access in hybrid journals through transformative agreements between 2018 and 2022 per publication year  according to the total number of articles indexed in Crossref.
+\caption{Relative growth of open access in hybrid journals through transformative agreements between 2018 and 2022 per publication year  according to the total number of articles indexed in Crossref.
 The blue areas represent open access through transformative agreements, as indicated by the affiliation with institutions that had a corresponding transformative agreement in place at the time of publication according to the first author affiliation data in OpenAlex matched with journal and institutional data from cOAlition S Transformative Agreements Public Data.
@@
 The individual outliers are not shown.
-Note that data on transformative agreements ending before June 2021 were not available at the time of this study."
+Note that data on transformative agreements ending before June 2021 were not available at the time of this study.}\label{fig:unnamed-chunk-2}
+\end{figure}
 
-library(patchwork)
-layout = "
-EEE
-ACD
-BCD"
-wrap_plots(
-  E = guide_area(),
-  A = oa_ta_plot,
-  B = ta_share_plot,
-  C = oa_jn_box,
-  D = box_ta_prop,
-  design = layout,
-  widths = c(0.4, 0.3, 0.3),
-  heights = c(0.2, 1, .5),
-  guides = "collect") &
-  theme(
-    legend.position = "top",
-    legend.justification = "right",
-    legend.direction = "horizontal",
-    panel.grid.minor = element_blank(),
-                panel.border = element_rect(color = "grey50", fill = NA),
-    axis.title=element_text(size = 10)) &
-    plot_annotation(tag_levels = 'A')
-```
+Figure 2A shows a moderate growth in the proportion of open access
+articles in hybrid journals, comparing the overall open access uptake
+and the impact of transformative agreements on this trend. Over the
+five-years period from 2018 to 2022, open access increased from 4.3\% (n
+= 65,486) to 15\% (n = 249,511). At the same time, the total article
+volume of the investigated journals grew from 1,528,051 in 2018 to
+1,676,928 in 2022.
 
-```{r}
-# in text data
-oa_by_year <- oa_ta_plot_df |>
-  group_by(cr_year) |>
-  summarise(oa = sum(value),
-            articles = n) |>
-  mutate(prop = oa / articles * 100) |>
-  distinct()
-```
+Figure 2B highlights that the majority of open access articles in hybrid
+journals were made available through transformative agreements in 2021
+and 2022, contributing 58\% of the total open access article volume in
+2022. However, there was also a notable growth in open access provision
+through individual publication fees, which increased from 4.1\% (n =
+62,625) in 2018 to 6.3\% (n = 105,896). This suggests that publishers
+were able to gain equally from individual and institutional open access
+publishing options.
 
-Figure 2A shows a moderate growth in the proportion of open access articles in hybrid journals, comparing the overall open access uptake and the impact of transformative agreements on this trend.
-Over the five-years period from 2018 to 2022, open access increased from `r oa_by_year |> filter(cr_year == "2018") |> pull(prop)`% (n = `r oa_by_year |> filter(cr_year == "2018") |> pull(oa)`) to `r oa_by_year |> filter(cr_year == "2022") |> pull(prop)`% (n = `r oa_by_year |> filter(cr_year == "2022") |> pull(oa)`).
-At the same time, the total article volume of the investigated journals grew from `r oa_by_year |> filter(cr_year == "2018") |> pull(articles)` in 2018 to `r oa_by_year |> filter(cr_year == "2022") |> pull(articles)` in 2022.
+Figure 2C depicts the substantial variations among the hybrid journals
+included in transformative agreements in terms of open access uptake.
+Although the median generally follows the trend shown in Figure 2A, the
+farther stretch of upper quartiles and whiskers over the years
+illustrates that an increasing number of journals published an
+above-average proportion of open access articles. In 2022, 25\% of
+hybrid journals (n = 2,576) had an open access uptake of 29\%, and 6.6\%
+of journals (n = 744) provided the majority of their articles under a
+Creative Commons license in the same year. These journals were, on
+average, smaller (M = 75, SD = 186) than those with an open access share
+below 50\% (M = 164, SD = 347). Notable exception of large journals with
+an above-average open access proportion were Physical Review D, a
+high-energy physics journal covered by the SCOAP3 consortium that
+provided open access to 2,341 out of 4,074 articles in 2022, Astronomy
+and Physics (1,396 out of 2,230 articles in 2022 were open access),
+which shifted to a subscribe to open business model for all accepted
+articles as of April 2022, the Journal of Fluid Mechanics (577 of 1,077
+articles in 2022 were open access) and Bioinformatics ( out of articles
+in 2022 were open access), which flipped to full open access as of
+January 2023.
 
-Figure 2B highlights that the majority of open access articles in hybrid journals were made available through transformative agreements in 2021 and 2022, contributing `r oa_ta_plot_df |> filter(cr_year == "2022", type == "ta_oa") |> pull(value) / oa_by_year |> filter(cr_year == "2022") |> pull(oa) * 100`% of the total open access article volume in 2022.
-However, there was also a notable growth in open access provision through individual publication fees, which increased from `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> pull(value)`) in 2018 to  `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> pull(value)`).
-This suggests that publishers were able to gain equally from individual and institutional open access publishing options.
+When comparing the impact of open access trough transformative
+agreements across journals, it shows that for many journals these
+agreements substantially contributed to the growth of open access over
+the years (Figure 2D). Examples of such journals include those with a
+scope on specific countries or regions, where also transformative
+agreements were implemented. For instance, in 2022, the Germany-based
+journals Zeitschrift für Erziehungswissenschaft and Zeitschrift für
+Politikwissenschaft, as well as the Scandinavian Political Studies
+adressing the Nordic countries, achieved an overall open access uptake
+of more than 90\% just through transformative agreements. Despite the
+rise in transformative agreements, it is worth noting that other means
+of publishing open access in hybrid journals remained common. In total,
+9,153 journals published open access articles from authors affiliated
+with institutions without transformative agreements in place, while
+8,780 journals published at least one open access article through a
+transformative agreement in the same year.
 
-Figure 2C depicts the substantial variations among the hybrid journals included in transformative agreements in terms of open access uptake.
-Although the median generally follows the trend shown in Figure 2A, the farther stretch of upper quartiles and whiskers over the years illustrates that an increasing number of journals published an above-average proportion of open access articles.
-In 2022, 25% of hybrid journals (n = `r oa_jn_box_df |> filter(cr_year == "2022") |> mutate(top25 = quantile(prop, .75)) |> filter(prop >= top25) |> nrow()`) had an open access uptake of `r oa_jn_box_df |> filter(cr_year == "2022") |> pull(prop) |> quantile(0.75)*100`%, and `r (oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> nrow()) / length(unique(oa_jn_box_df$issn_l)) *100`% of journals (n = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> nrow()`) provided the majority of their articles under a Creative Commons license in the same year.
-These journals were, on average, smaller (M = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> pull(n) |> mean()`, SD = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> pull(n) |> sd()`) than those with an open access share below 50% (M = `r oa_jn_box_df |> filter(cr_year == "2022", prop <= 0.5) |> pull(n) |> mean()`, SD = `r oa_jn_box_df |> filter(cr_year == "2022", prop <= 0.5) |> pull(n) |> sd()`).
-Notable exception of large journals with an above-average open access proportion were Physical Review D, a high-energy physics journal covered by the SCOAP3 consortium that provided open access to `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "2470-0010") |> pull(oa_all)` out of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "2470-0010") |> pull(n)` articles in 2022, Astronomy and Physics (`r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0004-6361") |> pull(oa_all)` out of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0004-6361") |> pull(n)` articles in 2022 were open access), which shifted to a subscribe to open business model for all accepted articles as of April 2022, the Journal of Fluid Mechanics (`r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0022-1120") |> pull(oa_all)` of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0022-1120") |> pull(n)` articles in 2022 were open access) and Bioinformatics (`r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "1367-4803") |> pull(oa_all)` out of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "1367-4803") |> pull(n)` articles in 2022 were open access), which flipped to full open access as of January 2023.
+\hypertarget{publishing-market}{%
+\subsection{Publishing market}\label{publishing-market}}
 
-When comparing the impact of open access trough transformative agreements across journals, it shows that for many journals these agreements substantially contributed to the growth of open access over the years (Figure 2D).
-Examples of such journals include those with a scope on specific countries or regions, where also transformative agreements were implemented. 
-For instance, in 2022, the Germany-based journals Zeitschrift für Erziehungswissenschaft and Zeitschrift für Politikwissenschaft, as well as the Scandinavian Political Studies adressing the Nordic countries, achieved an overall open access uptake of more than 90% just through transformative agreements.
-Despite the rise in transformative agreements, it is worth noting that other means of publishing open access in hybrid journals remained common. 
-In total, `r summary_all |> filter(cr_year == "2022", value > 0, name == "prop_no_ta") |> nrow()` journals published open access articles from authors affiliated with institutions without transformative agreements in place, while `r summary_all |> filter(cr_year == "2022", value > 0, name == "prop_ta_oa") |> nrow()` journals published at least one open access article through a transformative agreement in the same year.
+\begin{table}[H]
 
-## Publishing market
+\caption{\label{tab:unnamed-chunk-5}Hybrid open access through transformative agreements market shares 2018-2022}
+\centering
+\fontsize{10}{12}\selectfont
+\begin{tabular}[t]{lrllrllrllrll}
+\toprule
+\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Hybrid journals} & \multicolumn{3}{c}{Articles} & \multicolumn{3}{c}{OA articles} & \multicolumn{3}{c}{TA OA articles} \\
+\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7} \cmidrule(l{3pt}r{3pt}){8-10} \cmidrule(l{3pt}r{3pt}){11-13}
+Publisher & Total & \% & C\% & Total & \% & C\% & Total & \% & C\% & Total & \% & C\%\\
+\midrule
+Elsevier & 1,936 & 17 & 17 & 2,770,826 & 33.8 & 33.8 & 172,723 & 22.9 & 22.9 & 60,440 & 18.3 & 18.3\\
+Springer Nature & 2,274 & 20 & 37 & 1,330,430 & 16.2 & 50 & 175,432 & 23.3 & 46.2 & 100,008 & 30.3 & 48.6\\
+Wiley & 1,410 & 12.4 & 49.4 & 1,043,052 & 12.7 & 62.7 & 152,723 & 20.3 & 66.5 & 83,443 & 25.3 & 73.9\\
+Other & 5,767 & 50.6 & 100 & 3,061,337 & 37.3 & 100 & 252,523 & 33.5 & 100 & 86,294 & 26.1 & 100\\
+\bottomrule
+\end{tabular}
+\end{table}
 
-```{r}
-jn_ind_jct <- hoaddata::jct_hybrid_jns |>
-  inner_join(hoaddata::jn_ind, by = "issn_l") |>
-  filter(issn_l != "0027-8424")
+Analysing hybrid open access across publishers between 2018 and 2022
+reveals a large market concentration. Although 48 publishers offered
+transformative agreements, the big three commercial publishers Elsevier,
+Springer Nature, and Wiley, accounted for 49\% of total article volume
+published (see Table ). Together, they published 500,878 open access
+articles, representing 66\% of the open access articles in hybrid
+journals. Elsevier, Springer Nature, and Wiley made 243,891 articles
+open access in hybrid journals through transformative agreements,
+resulting in an even larger market share of 74\%. However, there are
+differences among the three large publishers. Although Elsevier
+published the largest volume of articles (n = 2,770,826, 34\%), it
+published a relatively low number of open access articles, including
+those that can be associated with transformative agreements. In
+contrast, Springer Nature and Wiley provided open access to a larger
+proportion of their articles (13\% of Springer Nature articles and 15\%
+of Wiley articles were open acccess), leading to higher open access
+market shares (23\% Springer Nature resp. 23\% Wiley). This difference
+between Elsevier on the one hand and Springer Nature and Wiley on the
+other can be attributed to transformative agreements, as the latter made
+the majority of their open access articles available through such deals
+(Springer Nature 57\% resp. Wiley 55\%).
 
-ta_active <- ta_oa_inst |>
-  filter(issn_l != "0027-8424") |>
-  filter(ta_active == TRUE, !is.na(cc)) |>
-  group_by(cr_year, issn_l, esac_publisher) |>
-  summarise(ta_oa = n_distinct(doi)) |>
-  mutate(cr_year = as.character(cr_year))
+\begin{center}\includegraphics[width=0.99\linewidth]{fig/unnamed-chunk-6-1} \end{center}
 
-jn_ind_jct_oa <- jn_ind_jct |>
-  filter(!is.na(cc)) |>
-  distinct(cr_year, issn_l, esac_publisher, cc, cc_total, jn_all) |>
-  group_by(cr_year, issn_l, esac_publisher, jn_all) |>
-  summarise(oa = sum(cc_total, na.rm = TRUE))
+Figure 2 takes a closer look into the growth of hybrid open access
+across publishers by year with a focus on open articles enabled by
+transformative agreements. Although all publishers show a general
+long-term trend towards transformative agreements, Figure 2A and 2B
+indicate that, in particular, Wiley's has experienced a substantial
+increase in its open access share from 5.9\% (n = 11,628) in 2018 to
+26\% (n = 53,503) in 2022, rperesenting an 4.5-fold increase. In
+contrast, Elsevier's hybrid journals deonstrated a more modest increase,
+from 3.3\% (n = 16,872) in 2018 to 10\% (n = 60,821) in 2022, which is a
+relatively low open access share compared to the general trend. In 2018,
+Springer Nature had the largest open access proportion among the three
+publishers of 8.4\% (n = 19,701), but experienced a relatively slower
+growth, resulting in 18\% (n = 52,616) of articles being open access in
+Springer Nature hybrid journals in 2022.
 
-jn_ind_jct_all <- jn_ind_jct |>
-  filter(issn_l %in% hoaddata::cc_articles$issn_l) |>
-  distinct(cr_year, issn_l, esac_publisher, jn_all)
+The varying degrees of uptake of open access across the three major
+publishers can be attributed to distinct approaches to transformative
+agreements. Springer Nature, for example, began in 2015 offering
+selected consortia, such as the Max Planck Society, the Swedish Bibsam
+consortium, and the Finnish FinELib consortium, open access agreements
+for its hybrid journal portfolio under the name Springer
+Compact\footnote{\url{https://web.archive.org/web/20180414062853id_/http://www.liber2015.org.uk/wp-content/uploads/2015/03/Springer-Compact.pdf}}.
+However, these agreements were not included in the data as they
+concluded prior to the start of the transformative agreement data
+collection in June 2021. Nonetheless, the results suggest the importance
+of central agreements for Springer Nature's hybrid open access business
+over the past five years (Figure 2B). In 2022, 66\% (n = 34,725) of open
+access articles in f Springer Nature hybrid journals were enabled
+through transformative agreements. In the same year, 70\% (n = 37,316)
+of Wiley's open access articles could be linked to transformative
+agreements in 2022. In contrast, Elsevier published fewer than half of
+its open access articles through transformative agreements (n = 32,627;
+54\%).
 
-publisher_df <- left_join(jn_ind_jct_all, jn_ind_jct_oa) |>
-  mutate(cr_year = as.character(cr_year)) |>
-  left_join(ta_active) |>
-  mutate(across(everything(), .fns = ~replace_na(.,0))) |>
-   mutate(no_ta = oa - ta_oa) |>
-    mutate(
-    publisher = case_when(
-      esac_publisher == "Elsevier" ~ "Elsevier",
-      esac_publisher == "Springer Nature" ~ "Springer Nature",
-      esac_publisher == "Wiley" ~ "Wiley",
-      is.character(esac_publisher) ~ "Other"
-    )
-  ) |>
-  mutate(publisher =
-           forcats::fct_relevel(
-             publisher,
-             c("Elsevier", "Springer Nature", "Wiley", "Other") 
-           )) |>
-  filter(!cr_year %in%  c("2017", "2023")) |>
-  # Journal transfers across publishers grouped in publisher
-  select(-esac_publisher) |>
-  distinct()
+The increasing trend towards transformative agreements can be also
+observed at the journal-level (Figure 2). While no substantial
+differences between open access enabled through transformative
+agreements and other revenue source could observed across Elsevier
+journals, the distribution of open access across Springer Nature and
+Wiley hybrid journals indicates that the growth is not limited to a few
+journals, but extends across the portfolio. In particular, Wiley's upper
+quantile, which represents the top 25\% of journals in terms of the
+proportion of open access articles from transformative agreements,
+increased markedly from 13\% in 2020 to 31\% in 2022. At the same time,
+the median proportion grew from 7.5\% to 19\%. It is interesting to note
+that a small but increasing number of journals from these two publishers
+are providing open access to the majority of articles through
+transformative agreements. Wiley recorded 68 and Springer Nature 102
+hybrid journals with an open access share above 50\% that could be
+solely attributed to transformative agreements. Upon inspection, these
+journals were mainly society or local language journals with a small
+yearly article volume.
 
-## plot
-publisher_plot_df <- publisher_df |> 
-  pivot_longer(cols = c(ta_oa, no_ta), names_to = "type", values_to = "value")
+\hypertarget{journal-subjects}{%
+\subsection{Journal subjects}\label{journal-subjects}}
 
-#+ publisher ta shares
-publisher_ta_prop <- publisher_plot_df |>
-  filter(type == "ta_oa") |>
-  group_by(cr_year, publisher) |>
-  summarize(ta_oa = sum(value), oa = sum(oa), jn_all = sum(jn_all)) |>
-  mutate(prop = ta_oa / oa)
+Table presents a high-level overview of hybrid open access by AJCS
+subject area using fractionalised counting to account for journals
+belonging to more than one category. Between 2018 and 2022, most hybrid
+journals with at least one open articles could be attributed to the
+social sciences including the humanities. However, these journals
+published the fewest number of articles, whereas physical sciences
+journals recorded most articles, followed by the health sciences and the
+life sciences. In terms of open access, physical sciences journals
+accounted for more than one third of articles published in the
+five-years period, followed by the health science, the social sciences
+and the life sciences.
 
-oa_publisher <- publisher_plot_df |>
-  group_by(cr_year, publisher, type) |>
-  summarise(value = sum(value))
-all_publisher <- publisher_plot_df |>
-  distinct(cr_year, issn_l, publisher, jn_all) |>
-  group_by(cr_year, publisher) |>
-  summarise(jn_all = sum(jn_all))
+\begin{table}[H]
 
-publisher_summary_df <- inner_join(oa_publisher, all_publisher) |>
-  mutate(prop = value / jn_all)
+\caption{\label{tab:subject_summary_table}Hybrid open access through transformative agreements by journal subject 2018-2022}
+\centering
+\begin{tabular}[t]{lrllrllrllrll}
+\toprule
+\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Hybrid journals} & \multicolumn{3}{c}{Articles} & \multicolumn{3}{c}{OA articles} & \multicolumn{3}{c}{TA OA articles} \\
+\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7} \cmidrule(l{3pt}r{3pt}){8-10} \cmidrule(l{3pt}r{3pt}){11-13}
+Journal subject & Total & \% & C\% & Total & \% & C\% & Total & \% & C\% & Total & \% & C\%\\
+\midrule
+Health Sciences & 2,376 & 22.5 & 22.5 & 2,709,906 & 27.8 & 27.8 & 286,592 & 27.3 & 27.3 & 117,746 & 25 & 25\\
+Life Sciences & 1,403 & 13.3 & 35.8 & 1,477,808 & 15.1 & 42.9 & 191,880 & 18.3 & 45.6 & 71,593 & 15.2 & 40.2\\
+Physical Sciences & 2,732 & 25.9 & 61.7 & 4,291,833 & 44 & 86.9 & 366,794 & 35 & 80.6 & 167,686 & 35.6 & 75.8\\
+Social Sciences & 4,050 & 38.3 & 100 & 1,280,460 & 13.1 & 100 & 203,461 & 19.4 & 100 & 114,190 & 24.2 & 100\\
+\bottomrule
+\end{tabular}
+\end{table}
 
-publisher_prop <-
-  ggplot(publisher_summary_df, aes(
-    gsub("^20", "'", as.character(cr_year)),
-    prop,
-    fill = type,
-    group = type
-  )) +
-  geom_area(color = "#000000",
-            linewidth = 0.15) +
-  facet_wrap( ~ publisher, nrow = 1) +
-  scale_fill_manual(
-    "Hybrid OA",
-    values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"),
-    labels = c(no_ta = "Other", ta_oa = "Agreement"),
-    guide = guide_legend(reverse = TRUE)
-  ) +
-  scale_y_continuous(
-    expand = expansion(mult = c(0, 0.05)),
-    labels = scales::percent_format(),
-    breaks = c(0.05, 0.1, 0.15, 0.2, 0.25)
-  ) +
-  theme_minimal() +
-  labs(y = "Hybrid OA Uptake", x = NULL) +
-  theme(
-    legend.position = "top",
-    legend.justification = "left",
-    axis.text.x = element_blank()
-  )
+Figure presents the relative growth of hybrid open access by subject
+area between 2018-2022. In particular, Social Sciences and Humanties
+journals accounted for the strongest growth in the five-years period
+from 6.4\% (n = 8,361) to 23\% (n = 51,938), followed by the Life
+Science from 7.6\% (n = 15,003) to 18\% (n = 39,494) , Health Science
+from 5.3\% (n = 18,279) to 16\% (n = 63,089) and Physical Sciences from
+4.5\% (n = 22,364) to 12\% (n = 85,428). This growth in the social
+sciences can be largely attributed to transformative agreements. In
+2022, two-third of open access articles (67\%, n = 34,759) were
+published by first authors affiliated with participating institutions.
+Figure C shows that this trend is consistent across Social Sciences
+journals. In 2022, 25\% of Social Science journals provided open access
+to at least every fourth article exclusively through transformative
+agreements. However, hybrid open access through transformative
+agreements played a comparable lesser role in the Life Sciences and
+Health Sciences. In these two subject areas, only about half of the open
+access articles can be linked to these agreements, both overall and on
+median average across journals. In contrast, the majority of Physical
+Science Journals, shows an increase of open access through
+transformative agreements compared to other options to publish open
+access in hybrid journals.
 
-publisher_boxplot <- publisher_plot_df |>
-  filter(value > 0) |>
-  ggplot(aes(gsub("^20", "'", as.character(cr_year)), value / jn_all, fill = type)) +
-  geom_boxplot(outlier.shape = NA,
-               show.legend = FALSE,
-               varwidth = TRUE) +
-  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
-                     labels = scales::percent) +
-  scale_fill_manual(
-    NULL,
-    values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"),
-    labels = c(no_ta = "Other", ta_oa = "Agreement")
-  ) +
-  coord_cartesian(ylim = c(0, 0.65)) +
-  facet_wrap( ~ publisher, nrow = 1) +
-  labs(y = "OA Uptake\nby Hybrid Journal", x = "Publication Year") +
-  theme_minimal() +
-  theme(strip.background = element_blank(),
-        strip.text.x = element_blank())
+\begin{center}\includegraphics[width=0.99\linewidth]{fig/subject_panel-1} \end{center}
 
-publisher_ta_prop_plot <- publisher_ta_prop |>
-  ggplot(aes(gsub("^20", "'", as.character(cr_year)), prop)) +
-  geom_bar(stat = "identity",
-           fill = "#56B4E9",
-            color = "#000000",
-           linewidth = 0.15) +
-  scale_y_continuous(
-    expand = expansion(mult = c(0, 0.05)),
-    breaks = c(0, 0.25, 0.5, 0.75),
-    limits = c(0, .75),
-    labels = scales::percent
-  ) +
-  facet_wrap( ~ publisher, nrow = 1) +
-  theme_minimal() +
-  labs(y = "Percentage\nof Hybrid OA\nvia Agreement", x = NULL) +
-  theme(
-    axis.text.x = element_blank(),
-    strip.background = element_blank(),
-    strip.text.x = element_blank()
-  )
-```
+\hypertarget{comparing-countries}{%
+\subsection{Comparing countries}\label{comparing-countries}}
 
-```{r publisher_summary_table}
-publisher_article_total_table <- publisher_df |>
-  group_by(publisher) |>
-  summarise(oa_journals = n_distinct(issn_l),
-            publisher_total = sum(jn_all)) |>
-  mutate(
-    jn_perc = oa_journals / sum(oa_journals),
-    jn_cum = cumsum(oa_journals) / sum(oa_journals),
-    perc = publisher_total / sum(publisher_total),
-    cum = cumsum(publisher_total) / sum(publisher_total)
-  ) |>
-  relocate(publisher_total, .after = jn_cum)
+Between 2018 and 2022, Western economies almost exclusively dominated
+hybrid open access publishing through transformative agreements. During
+this period, first-authors affiliated with institutions from
+Organisation for Economic Co-operation and Development (OECD) member
+countries published 602,050 open access articles in hybrid journals,
+representing 81\% of the investigated open access articles. This
+disparity between OECD nations and other countries becomes even more
+evident when considering open access through transformative agreements,
+as 310,712 of 328,957, or 94\% of open access articles were associated
+with such agreements.
 
-publisher_article_oa_table <- publisher_summary_df |>
-  group_by(publisher) |>
-  summarise(oa_total = sum(value)) |>
-  mutate(oa_perc = oa_total / sum(oa_total),
-         oa_cum = cumsum(oa_total) / sum(oa_total))
+\begin{center}\includegraphics[width=0.99\linewidth]{fig/country_patch-1} \end{center}
 
-publisher_article_ta_oa_table <- publisher_summary_df |>
-  filter(type == "ta_oa") |>
-  group_by(publisher) |>
-  summarise(ta_oa_total = sum(value)) |>
-  mutate(
-    ta_oa_perc = ta_oa_total / sum(ta_oa_total),
-    ta_oa_cum = cumsum(ta_oa_total) / sum(ta_oa_total)
-  )
+Figure 2A shows the development of hybrid open access publishing by
+countries, comparing the OECD area with the BRICS, an intergovernmental
+organisation, which comprised the countries Brazil, Russia, India, China
+and South Africa in 2022. The residual category ``Other'' includes the
+remaining countries. From 2018 to 2022, the proportion of open access in
+hybrid journals increased from 6.1\% in 2018 to 26\% in 2022. On the
+other hand, BRICS recorded an low uptake, from 1.6\% in 2018 to 3.7\% in
+2022.\\
+Despite rise of open access across OECD countries, the overall
+publication output decreased sharply, dropping to 786,903 in 2022 after
+peaking 892,197 articles in 2020. In stark contrast, the number of
+articles published in hybrid journals by first authors affiliated with
+institutions from BRICS countries increased steadily over the years,
+more than doubling from 356,632 in 2018 to 786,903 in 2022. Upon closer
+examination, this trend can be observed across all big three publishers,
+although the shift towards BRICS is particularly evident in Elsevier's
+hybrid journal portfolio, in particular with regard to articles
+published in Physical Sciences journals (see Supplement). While OECD
+publication output in Elsevier's Physical Sciences journals declined
+from 112,822 articles in 2018 to 103,766 in 2022, BRICS output increased
+from 104,654 to 171,713 in the same five-year period. Furthermore, OECD
+publication output in Health Science Journals and Life Science journals
+stagnated across the investigated hybrid journal portfolios after a peak
+in 2020.
 
-publisher_league_tbl <-
-  inner_join(publisher_article_total_table, publisher_article_oa_table) |>
-  inner_join(publisher_article_ta_oa_table)
-```
+To illustrate the situation in 2022, Figure B compares total publication
+output with the number of open access articles. With 391,530 articles,
+China was the most productive country, followed by the United States
+(268,965 articles) and India (87,428 articles). In contrast, West and
+Nord European countries published a considerable high number of open
+access articles, mainly due to transformative agreements. Particularly,
+Germany, Great Britain, the Netherlands, Sweden, Switzerland and Spain
+recorded an above-average open access share as indicated by the linear
+trend line. As represented by the point size, as well as it can been
+seen in Figure C, transformative agreements contributed to this market
+position of these countries. Interestingly, the United States had a
+notable open access market share of 15\%, although transformative
+agreements contributed to a lesser extent. Similarly, China's open
+access market share of 7.2 in 2022 was comparable to that of the
+Netherlands, which was (7.1\%,).
 
-```{r}
-publisher_league_tbl |>
-  mutate(across(contains(c("perc", "cum")), ~ paste(round(., 3) * 100))) |>
-  kbl(
-    caption ="Hybrid open access through transformative agreements market shares 2018-2022",
-    col.names = c(
-      "Publisher",
-      "Total",
-      "%",
-      "C%",
-      "Total",
-      "%",
-      "C%",
-      "Total",
-      "%",
-      "C%",
-      "Total",
-      "%",
-      "C%"
-    ),
-    format.args = list(big.mark = ',')
-  ) %>%
-  kable_classic() %>%
-  add_header_above(c(
-    " " = 1,
-    "Hybrid journals" = 3,
-    "Articles" = 3,
-    "OA articles" = 3,
-    "TA OA articles" = 3
-  ))
-```
+\begin{center}\includegraphics[width=0.99\linewidth]{fig/country_top_20_plot-1} \end{center}
 
-Analysing hybrid open access across publishers between 2018 and 2022 reveals a large market concentration. 
-Although `r length(unique(jn_ind_jct_all$esac_publisher))` publishers offered transformative agreements, the big three commercial publishers Elsevier, Springer Nature, and Wiley, accounted for `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(jn_cum) * 100`% of total article volume published (see Table ).
-Together, they published `r publisher_league_tbl  |> filter(publisher != "Other") |> pull(oa_total) |> sum()` open access articles, representing `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(oa_cum) * 100`% of the open access articles in hybrid journals.
-Elsevier, Springer Nature, and Wiley made `r publisher_league_tbl  |> filter(publisher != "Other") |> pull(ta_oa_total) |> sum()` articles open access in hybrid journals through transformative agreements, resulting in an even larger market share of `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(ta_oa_cum)*100`%.
-However, there are differences among the three large publishers.
-Although Elsevier published the largest volume of articles (n = `r publisher_league_tbl  |> filter(publisher == "Elsevier") |> pull(publisher_total)`, `r publisher_league_tbl  |> filter(publisher == "Elsevier") |> pull(cum)*100`%), it published a relatively low number of open access articles, including those that can be associated with transformative agreements.
-In contrast, Springer Nature and Wiley provided open access to a larger proportion of their articles (`r (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_total)) / (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(publisher_total)) * 100`% of Springer Nature articles and `r (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(oa_total)) / (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(publisher_total)) * 100`% of Wiley articles were open acccess), leading to higher open access market shares (`r publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_perc)*100`% Springer Nature resp.
-`r publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_perc)*100`% Wiley).
-This  difference between Elsevier on the one hand and Springer Nature and Wiley on the other can  be attributed to transformative agreements, as the latter made the majority of their open access articles available through such deals (Springer Nature `r ((publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(ta_oa_total)) / (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_total)) * 100)`% resp.
-Wiley `r (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(ta_oa_total)) / (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(oa_total)) * 100`%).
+Figure illustrates the development of hybrid open access from 2018 to
+2022, highlighting the top 20 most productive countries in terms of
+articles published in hybrid journals that were included in
+transformative agreements over the five-year period. Notably, The
+Netherlands (27\%), Sweden (24\%), Poland (17\%) and Great Britain
+(17\%)) exhibited a relatively high level of uptake in 2018 which
+continued to increase in the following years. In 2022, Sweden had the
+highest proportion of open-access articles relative to its publication
+output (78\%), followed by the Netherlands (67\%) and Switzerland
+(57\%), with these countries benefiting from transformative agreements.
+In Germany, however, hybrid open access only began to increase from 2019
+onwards after the successful negotiation of nationwide agreements with
+Wiley (July 2019) and Springer Nature (January 2020). Prior to this,
+only a few organisations had agreements in place, for the example the
+Max Planck Society with Springer Compact.
 
-```{r, fig.asp=0.8}
-library(patchwork)
-design = "E
-A
-B
-C"
-wrap_plots(
-  E = guide_area(),
-  A = publisher_prop,
-  B = publisher_ta_prop_plot,
-  C = publisher_boxplot,
-  design = design,
-  guides = "collect",
-  heights = c(0, 0.4, 0.3, 0.7)
-) &
-  theme(
-    legend.position = "top",
-    legend.justification = "right",
-    legend.direction = "horizontal",
-    panel.grid.minor = element_blank(),
-                panel.border = element_rect(color = "grey50", fill = NA),
-    axis.title=element_text(size = 10)) &
-    plot_annotation(tag_levels = 'A') 
-```
+Since 2021, there has been a general trend towards hybrid open access
+among many western countries, primarily driven by transformative
+agreements. However, proliferation of transformative agreements differe
+across these countries. For instance, Germany successfully negotiated an
+agreement with Elsevier not until 2023. Additionally, publication limits
+or eligibility criteria for institutions and article types may explain
+why even countries with widespread agreement implementation do not
+achieve 100\% hybrid open access. Interestingly, in Japan and the US
+other options than transformative agreements were the main driver for
+the increase in hybrid open access. Once again, the graph highlights
+countries with low hybrid open access, particularly non-OECD countries,
+where only a few or no agreements were in place.
 
-Figure 2 takes a closer look into the growth of hybrid open access across publishers by year with a focus on open articles enabled by transformative agreements. 
-Although all publishers show a general long-term trend towards transformative agreements, Figure 2A and 2B indicate that, in particular, Wiley's has experienced a substantial increase in its open access share from `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(value) |> sum()`) in 2018 to `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(prop) |> sum()*100`%  (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()`) in 2022, rperesenting an 4.5-fold increase.
-In contrast, Elsevier's hybrid journals deonstrated a more modest increase, from `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Elsevier") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Elsevier") |> pull(value) |> sum()`) in 2018 to `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(value) |> sum()`) in 2022, which is a relatively low open access share compared to the general trend.
-In 2018, Springer Nature had the largest open access proportion among the three publishers of `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Springer Nature") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Springer Nature") |> pull(value) |> sum()`), but experienced a relatively slower growth, resulting in `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature") |> pull(value) |> sum()`) of articles being open access in Springer Nature hybrid journals in 2022.
+\hypertarget{supplements}{%
+\subsubsection{Supplements}\label{supplements}}
 
-The varying degrees of uptake of open access across the three major publishers can be attributed to distinct approaches to transformative agreements. 
-Springer Nature, for example, began in 2015 offering selected consortia, such as the Max Planck Society, the Swedish Bibsam consortium, and the Finnish FinELib consortium, open access agreements for its hybrid journal portfolio under the name Springer Compact[^springercompact].
-However, these agreements were not included in the data as they concluded prior to the start of the transformative agreement data collection in June 2021. 
-Nonetheless, the results suggest the importance of central agreements for Springer Nature's hybrid open access business over the past five years (Figure 2B).
-In 2022, `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature") |> pull(value) |> sum()*100`% (n = `r  publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature", type == "ta_oa") |> pull(value)`) of open access articles in f Springer Nature  hybrid journals were enabled through transformative agreements.
-In the same year, `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley", type == "ta_oa") |> pull(value)`) of Wiley's open access articles could be linked to transformative agreements in 2022.
-In contrast, Elsevier published fewer than half of its open access articles through transformative agreements (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier", type == "ta_oa") |> pull(value) `; `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(value) |> sum()*100`%).
+\begin{center}\includegraphics[width=0.99\linewidth]{fig/unnamed-chunk-10-1} \end{center}
 
-```{r}
-pub_75 <- publisher_plot_df |> filter(value > 0) |> mutate(prop = value /jn_all) |> group_by(publisher, type, cr_year) |> summarise(
-  top25 = quantile(prop, 0.75),
-  median = median(prop),
-  top10 = quantile(prop, 0.9))
-```
-
-The increasing trend towards transformative agreements can be also observed at the journal-level (Figure 2).
-While no substantial differences between open access enabled through transformative agreements and other revenue source could observed across Elsevier journals, the distribution of open access across Springer Nature and Wiley hybrid journals indicates that the growth is not limited to a few journals, but extends across the portfolio.
-In particular, Wiley's upper quantile, which represents the top 25% of journals in terms of the proportion of open access articles from transformative agreements, increased markedly from `r pub_75 |> filter(publisher == "Wiley", type == "ta_oa", cr_year == "2020") |> pull(top25) * 100`% in 2020 to `r pub_75 |> filter(publisher == "Wiley", type == "ta_oa", cr_year == "2022") |> pull(top25) * 100`% in 2022.
-At the same time, the median proportion grew from `r pub_75 |> filter(publisher == "Wiley", type == "ta_oa", cr_year == "2020") |> pull(median) * 100`% to  `r pub_75 |> filter(publisher == "Wiley", type == "ta_oa", cr_year == "2022") |> pull(median) * 100`%.
-It is interesting to note that a small but increasing number of journals from these two publishers are providing open access to the majority of articles through transformative agreements.
-Wiley recorded `r publisher_plot_df |>  mutate(prop = value / jn_all)  |> filter(cr_year == "2022", prop > 0.5, type == "ta_oa", publisher == "Wiley") |> nrow()` and Springer Nature `r publisher_plot_df |>  mutate(prop = value / jn_all)  |> filter(cr_year == "2022", prop > 0.5, type == "ta_oa", publisher == "Springer Nature") |> nrow()` hybrid journals with an open access share above 50% that could be solely attributed to transformative agreements.
-Upon inspection, these journals were mainly society or local language journals with a small yearly article volume.
-
-```{r subject_data_prep}
-jn_issnl_ajsc <-
-  readr::read_csv(here::here("data", "jn_issnl_ajsc.csv")) |>
-  select(issn_l, top_level, subject_area, ajcs)
-subject_df <- left_join(jn_ind_jct_all, jn_ind_jct_oa) |>
-  mutate(cr_year = as.character(cr_year)) |>
-  left_join(ta_active) |>
-  mutate(across(everything(), .fns = ~ replace_na(., 0))) |>
-  mutate(no_ta = oa - ta_oa) |>
-  left_join(jn_issnl_ajsc, by = "issn_l")
-
-top_level_by_year_jn <- subject_df |>
-  # with at least one
-  filter(oa > 0) |>
-  distinct(cr_year, issn_l, jn_all, top_level) |>
-  group_by(cr_year, issn_l, top_level) |>
-  summarise(n = sum(jn_all)) |>
-  mutate(fractions_factor = n / sum(n)) |>
-  mutate(fractions = n * fractions_factor) |>
-  ungroup() |> 
-  group_by(cr_year, top_level) |>
-  summarise(articles_n = sum(n),
-            articles_frac = sum(fractions)) |>
-  arrange(top_level, cr_year)
-
-top_level_by_year_oa_frac <- subject_df |>
-   # with at least one
-  filter(oa > 0) |>
-  select(-ajcs, -subject_area, -esac_publisher) |>
-  distinct(cr_year, issn_l, top_level, .keep_all = TRUE) |>
-  group_by(cr_year, issn_l, top_level) |>
-  summarise(n = sum(jn_all)) |>
-  mutate(fractions_factor = n / sum(n)) |>
-  select(-n)
-
-top_level_by_year_oa <- top_level_by_year_oa_frac |>
-  inner_join(subject_df) |>
-     # with at least one
-  filter(oa > 0) |>
-  select(-ajcs, -subject_area, -esac_publisher) |>
-  distinct(cr_year, issn_l, top_level, .keep_all = TRUE) |>
-  mutate(across(c(oa, ta_oa, no_ta),  function(x) x * fractions_factor, .names = "frac_{.col}")) |>
-  ungroup() |> 
-  group_by(cr_year, top_level) |>
-  summarise(oa_n = sum(oa),
-            oa_frac = sum(frac_oa),
-            ta_oa = sum(ta_oa),
-            ta_oa_frac= sum(frac_ta_oa),
-            no_ta = sum(no_ta),
-            no_ta_fractions = sum(frac_no_ta)) |>
-  arrange(top_level, cr_year) |>
-  inner_join(top_level_by_year_jn)
-```
-
-## Journal subjects
-
-Table presents a high-level overview of hybrid open access by AJCS subject area using fractionalised counting to account for journals belonging to more than one category. 
-Between 2018 and 2022, most hybrid journals with at least one open articles could be attributed to the social sciences including the humanities.
-However, these journals published the fewest number of articles, whereas physical sciences journals recorded most articles, followed by the health sciences and the life sciences. 
-In terms of open access,  physical sciences journals accounted for more than one third of articles published in the five-years period, followed by the health science, the social sciences and the life sciences.
-
-```{r subject_summary_table}
-jn_fract <- subject_df |>
-  # with at least one
-  filter(oa > 0) |>
-  distinct(issn_l, top_level) |>
-  group_by(issn_l, top_level) |>
-  summarise(n = n_distinct(issn_l)) |>
-  mutate(fractions_factor = n / sum(n)) |>
-  mutate(fractions = n * fractions_factor) |>
-  ungroup() |> 
-  group_by(top_level) |>
-  summarise(jn_frac = sum(fractions)) |>
-  arrange(top_level)
-
-subject_tbl <- top_level_by_year_oa |>
-  inner_join(jn_fract, by = "top_level") |>
-  filter(!top_level %in% c("Multidisciplinary", NA)) |>
-  group_by(top_level, jn_frac) |>
-  summarise(articles_frac = sum(articles_frac),
-            oa_frac = sum(oa_frac),
-            ta_oa_frac = sum(ta_oa_frac)) |>
-  ungroup() |>
-  mutate(
-    jn_perc = jn_frac / sum(jn_frac),
-    jn_cum = cumsum(jn_frac) / sum(jn_frac),
-    perc = articles_frac / sum(articles_frac),
-    cum = cumsum(articles_frac) / sum(articles_frac),
-    oa_perc = oa_frac / sum(oa_frac),
-    oa_cum = cumsum(oa_frac) / sum(oa_frac),
-    ta_oa_perc = ta_oa_frac / sum(ta_oa_frac),
-    ta_oa_cum = cumsum(ta_oa_frac) / sum(ta_oa_frac)
-  ) |>
-  select(top_level, jn_frac, jn_perc, jn_cum, articles_frac, perc, cum, 
-         oa_frac, oa_perc, oa_cum, ta_oa_frac,ta_oa_perc,  ta_oa_cum)
-
-subject_tbl |>
-  mutate(across(contains(c("perc", "cum")), ~ paste(round(., 3) * 100))) |>
-  kbl(
-    caption ="Hybrid open access through transformative agreements by journal subject 2018-2022",
-    col.names = c(
-      "Journal subject",
-      "Total",
-      "%",
-      "C%",
-      "Total",
-      "%",
-      "C%",
-      "Total",
-      "%",
-      "C%",
-      "Total",
-      "%",
-      "C%"
-    ),
-    format.args = list(big.mark = ',')
-  ) %>%
-  kable_classic() %>%
-  add_header_above(c(
-    " " = 1,
-    "Hybrid journals" = 3,
-    "Articles" = 3,
-    "OA articles" = 3,
-    "TA OA articles" = 3
-  ))
-```
-
-```{r subject_plots}
-subject_trend_plot <- top_level_by_year_oa |>
-  filter(top_level != "Multidisciplinary") |>
-  filter(!cr_year %in%  c("2017", "2023")) |>
-  select(cr_year, top_level, contains("frac"),-oa_frac) |>
-  pivot_longer(
-    cols = c(ta_oa_frac, no_ta_fractions),
-    names_to = "type",
-    values_to = "value"
-  ) |>
-  mutate(prop = value / articles_frac) |>
-  ggplot(aes(
-    gsub("^20", "'", as.character(cr_year)),
-    prop,
-    fill = type,
-    group = type
-  )) +
-    geom_area(color = "#000000",
-            linewidth = 0.15) +
-  facet_wrap( ~ top_level, nrow = 1) +
-  scale_fill_manual(
-    NULL,
-    values = c(
-      no_ta_fractions = "#b3b3b3a0",
-      ta_oa_frac = "#56B4E9"
-    ),
-    labels = c(no_ta_fractions = "Other", ta_oa_frac = "Agreement"),
-    guide = guide_legend(reverse = TRUE)
-  ) +
-  scale_y_continuous(
-    expand = expansion(mult = c(0, 0.05)),
-    labels = scales::percent_format(),
-    breaks = c(0.05, 0.1, 0.15, 0.2, 0.25)
-  ) +
-  theme_minimal() +
-  labs(y = "Hybrid OA Uptake", x = NULL) +
-  theme(
-    legend.position = "top",
-    legend.justification = "left",
-    axis.text.x = element_blank()
-  )
-
-subject_ta_trend_plot <- top_level_by_year_oa |>
-  filter(top_level != "Multidisciplinary") |>
-  filter(!cr_year %in%  c("2017", "2023")) |>
-  mutate(prop = ta_oa_frac / oa_frac) |>
-  ggplot(aes(gsub("^20", "'", as.character(cr_year)), prop)) +
-  geom_bar(stat = "identity", fill = "#56B4E9", color = "#000000",
-            linewidth = 0.15) +
-  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), breaks = c(0, 0.25,0.5,0.75), limits = c(0, .75),
-                     labels = scales::percent) +
-  facet_wrap(~top_level, nrow = 1) +
-  theme_minimal() +
-  labs(y = "Percentage\nof Hybrid OA\nvia Agreement", x = NULL) +
-  theme(
-    axis.text.x = element_blank(),
-    strip.background = element_blank(),
-    strip.text.x = element_blank())
-subject_box_trend <- subject_df |>
-   # with at least one
-  select(-ajcs, -subject_area, -esac_publisher) |>
-  distinct(cr_year, issn_l, top_level, .keep_all = TRUE) |>
-  pivot_longer(cols = c(ta_oa, no_ta), names_to = "type", values_to = "value") |>
-  filter(value > 0) |>
-  filter(top_level != "Multidisciplinary", !cr_year %in% c("2017", "2023")) |>
-  ggplot(aes(gsub("^20", "'", as.character(cr_year)), value / jn_all, fill = type)) +
-  geom_boxplot(outlier.shape = NA, show.legend = FALSE, varwidth = TRUE) +
-  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
-                     labels = scales::percent) +
-  scale_fill_manual("Hybrid OA", values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"), labels = c(no_ta = "Other", ta_oa = "Agreement")) +
-  coord_cartesian(ylim = c(0,0.55)) +
-  facet_wrap(~top_level, nrow = 1) +
-  labs(y = "OA Uptake\nby Hybrid Journal", x = "Publication Year") +
-  theme_minimal() +
-    theme(strip.background = element_blank(),
-    strip.text.x = element_blank()
-    )
-```
-
-Figure presents the relative growth of hybrid open access by subject area between 2018-2022.
-In particular, Social Sciences and Humanties journals accounted for the strongest growth in the five-years period from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Social Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Social Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(value) |> sum()`), followed by the Life Science from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Life Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Life Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Life Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Life Sciences") |> pull(value) |> sum()`) , Health Science from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Health Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Health Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Health Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Health Sciences") |> pull(value) |> sum()`) and Physical Sciences from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Physical Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Physical Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Physical Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Physical Sciences") |> pull(value) |> sum()`).
-<!-- In particular, journals in the subject areas recorded large high rate (see supplement)-->
-This growth in the social sciences can be largely attributed to transformative agreements.
-In 2022, two-third of open access articles (`r subject_ta_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(prop) * 100`%, n = `r subject_ta_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(ta_oa_frac)`) were published by first authors affiliated with participating institutions.
-Figure C shows that this trend is consistent across Social Sciences journals.
-In 2022, 25% of Social Science journals provided open access to at least every fourth article exclusively through transformative agreements.
-However, hybrid open access through transformative agreements played a comparable lesser role in the Life Sciences and Health Sciences. 
-In these two subject areas, only about half of the open access articles can be linked to these agreements, both overall and on median average across journals.
-In contrast, the majority of Physical Science Journals, shows an increase of open access through transformative agreements compared to other options to publish open access in hybrid journals.
-
-```{r subject_panel, fig.asp=0.8}
-library(patchwork)
-design = "E
-A
-B
-C"
-wrap_plots(
-  E = guide_area(),
-  A = subject_trend_plot,
-  B = subject_ta_trend_plot,
-  C = subject_box_trend,
-  design = design,
-  guides = "collect",
-  heights = c(0, 0.4, 0.3, 0.7)
-) &
-  theme(
-    legend.position = "top",
-    legend.justification = "right",
-    legend.direction = "horizontal",
-    panel.grid.minor = element_blank(),
-                panel.border = element_rect(color = "grey50", fill = NA),
-    axis.title=element_text(size = 10)) &
-    plot_annotation(tag_levels = 'A') 
-```
-
-```{r coutnry_data_prep}
-country_all <- hoaddata::jn_aff |>
-  filter(!cr_year %in% c(2017, 2023), issn_l != "0027-8424") |>
-  # only with cc
-  filter(issn_l %in% hoaddata::cc_articles$issn_l) |>
-  distinct(issn_l, cr_year, country_code, articles_total) |>
-  group_by(cr_year, country_code) |>
-  summarise(all = sum(articles_total))
-
-country_oa <- hoaddata::cc_articles |>
-  filter(!cr_year %in% c(2017, 2023), issn_l != "0027-8424") |>
-  group_by(cr_year, country_code) |>
-  summarise(oa = n_distinct(doi))
-
-country_codes <- readr::read_csv(here::here("data", "ror_country_codes.csv"))
-country_ta <-  ta_oa_inst |>
-  filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
-  left_join(country_codes) |>
-  group_by(cr_year, country_code) |>
-  summarise(ta_oa = n_distinct(doi))
-
-country_df <- country_all |>
-  left_join(country_oa, by = c("cr_year", "country_code")) |>
-  left_join(country_ta, by = c("cr_year", "country_code")) |>
-  distinct() |>
-  ungroup() |>
-  mutate(cr_year = as.character(cr_year)) |>
-  mutate(across(c(oa, ta_oa), .fns = ~replace_na(.,0))) |>
-  filter(!is.na(country_code)) |>
-  mutate(
-    country_name = countrycode::countrycode(country_code, origin = "iso2c", destination = "country.name")
-  )
-
-# Top 20 in terms of articles published
-top_20 <- country_df |>
-  filter(!is.na(country_code)) |>
-  group_by(country_code, country_name) |>
-  summarise(n = sum(all)) |>
-  arrange(desc(n)) |>
-  head(20)
-
-## Order of multiples, sort by the most productive (Top 20)
-
-country_multiple_df <- country_df |>
-  mutate(country_name_fct = forcats::fct_other(country_code, keep = top_20$country_code)) |>
-  mutate(country_name_fct = forcats::fct_relevel(country_name_fct, top_20$country_code, "Other")) |>
-  group_by(country_name_fct, cr_year) |>
-  summarise(
-    all_total = sum(all),
-    oa_total = sum(oa),
-    ta_oa_total = sum(ta_oa)
-  ) |>
-  mutate(prop = oa_total / all_total) |>
-  # remove cases where we could not retrieve a country name |>
-  filter(!is.na(country_name_fct))
-
-# Prepare facet heading
-facet_labels <- country_multiple_df |>
-  group_by(country_name_fct) |>
-  summarise(total = sum(all_total)) |>
-  mutate(prop = total / sum(total)) |>
-  ungroup() |>
-  mutate(total_string = case_when(
-    total < 1000 ~ as.character(total),
-    total < 1000000 ~ paste0(format(round(total / 1e3, 1), trim = TRUE), "K"),
-    total >= 1000000 ~ paste0(format(round(total / 1e6, 2), trim = TRUE), "M")
-  )) |>
-  mutate(
-    facet_labels = glue::glue(
-      '**{country_name_fct}**<br>
-      {total_string} | {paste(round(prop * 100, 1), "%")}'
-    )
-  ) |>
-  distinct(country_name_fct, facet_labels)
-
-country_facet_labels <- as.character(facet_labels$facet_labels)
-names(country_facet_labels) <- facet_labels$country_name_fct
-
-top_20_plot_df <- 
-  country_multiple_df |> 
-  mutate(no_ta = oa_total - ta_oa_total) |>
-  rename(ta_oa = ta_oa_total) |>
-  pivot_longer(
-  cols = c(no_ta, ta_oa),
-  values_to = "value",
-  names_to = "type") 
-```
-
-```{r, country_plot, fig.asp=0.8}
-country_plot <- ggplot(top_20_plot_df, aes(
-    gsub("^20", "'", as.character(cr_year)),
-    value / all_total,
-    group = type,
-    fill = type
-  )) +
-  geom_area(color = "#000000",
-            linewidth = 0.15,
-            stat = "identity") +
-scale_fill_manual(
-    NULL,
-    values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"),
-    labels = c(no_ta = "Other", ta_oa = "Agreement"),
-    guide = guide_legend(reverse = TRUE)
-  ) +
-  facet_wrap( ~ country_name_fct,
-              nrow = 3,
-              labeller = as_labeller(country_facet_labels)) +
-  scale_y_continuous(
-    labels = scales::percent_format(accuracy = 1L),
-    expand = expansion(mult = c(0, 0.05)),
-    breaks = scales::breaks_extended(4)
-  ) +
-  theme_minimal() +
-  labs(y = "Hybrid OA Uptake", x = "2018 - 2022") +
-  theme(
-    legend.position = "top",
-    legend.justification = "right",
-    legend.direction = "horizontal",
-    axis.text.x = element_blank(),
-    panel.grid.minor = element_blank(),
-                panel.border = element_rect(color = "grey50", fill = NA),
-    axis.title=element_text(size = 10),
-    strip.text = ggtext::element_markdown(size = 7),
-    plot.subtitle = ggtext::element_markdown(hjust = 0.5),
-    panel.spacing = unit(1, "lines")
-  )
-```
-
-## Comparing countries
-
-```{r}
-oecd <- readr::read_csv(here::here("data", "oecd.csv")) |>
-  mutate(country_code = countrycode::countrycode(iso3, "iso3c", "iso2c"))
-
-country_scatter_df <- country_df |>
-#  filter(cr_year == "2022", !is.na(country_code)) |>
-  mutate(cat = case_when(
-    country_code %in% oecd$country_code ~ "OECD",
-    country_code %in% c("BR", "CN", "IN", "RU", "ZA") ~ "BRICS",
-    .default = "Other"
-  ))
-
-oecd_stats <- readr::read_csv(here::here("data", "oecd_stats.csv"))
-oecd_stats_global <- oecd_stats |> 
-  filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
-  filter(issn_l %in% hoaddata::cc_articles$issn_l) |>
-  group_by(cat) |> summarise(articles = sum(articles), oa = sum(oa), ta_oa = sum(ta_oa))
-```
-
-```{r country_bar_chart_top_20}
-top_20_df <- country_scatter_df |>
-  filter(cr_year == "2022") |>
-  mutate(country_name_fct = forcats::fct_other(country_name, keep = top_20$country_name)) |>
-  mutate(country_name_fct = forcats::fct_relevel(country_name_fct, top_20$country_name, "Other")) |>
-  group_by(country_name_fct, country_code, cat) |>
-  summarise(all = sum(all),
-            oa = sum(oa),
-            ta_oa = sum(ta_oa), .groups = "drop") |>
-  head(20) |>
-  mutate(cat_labels = case_when(
-   cat == "OECD" ~ glue::glue("<span style = 'color: #3fc1c9'>{country_code}</span>"),
-   cat == "BRICS" ~ glue::glue("<span style = 'color: #fc5185'>{country_code}</span>"),
-   cat == "Other" ~ glue::glue("<span style = 'color: #e1e1e1'>{country_code}</span>")
-  )) |>
-  mutate(cat_labels = fct_inorder(cat_labels))
-
-top_20_plot <- top_20_df |>
-  mutate(no_ta = oa - ta_oa) |>
-  pivot_longer(
-  cols = c(no_ta, ta_oa),
-  values_to = "value",
-  names_to = "type") |>
-  ggplot(aes(reorder(as.character(country_code),  value), value / sum(value), fill = type)) +
-    geom_bar(stat = "identity", color = "#000000", linewidth = 0.15) +
-   scale_fill_manual(
-    NULL,
-    values = c(no_ta = "#b3b3b3a0", ta_oa = "#56B4E9"),
-    labels = c(no_ta = "Other", ta_oa = "Agreement"),
-    guide = FALSE) +
-  coord_flip() +
-  scale_y_continuous(
-    labels = scales::percent_format(accuracy = 1L),
-    expand = expansion(mult = c(0, 0.05)),
-    breaks = scales::breaks_extended(4)
-  ) +
-  theme_minimal() +
-#  theme(axis.text.y = ggtext::element_markdown()) +
-  labs(x = NULL, y = "% of Hybrid OA published\nin 2022")
-
-```
-
-```{r}
-country_scatter_22 <- country_scatter_df |>
-  filter(cr_year == "2022", !is.na(country_code), all > 10000) |>
-  ggplot(aes(all, oa)) +
-  geom_point(aes(fill = cat, size = ta_oa), pch = 21, alpha = 0.7) +
-  scale_fill_manual(NULL, values = c("OECD" = "#d01c8b", "BRICS" = "#4dac26", "Other" = "#f7f7f7")) +
-  geom_smooth(aes(all, oa), method=lm) +
-  scale_size("Agreement", guide = FALSE) +
-  ggrepel::geom_text_repel(
-    data = filter(country_scatter_df, cr_year == 2022, ta_oa > 5000 | country_code %in% c("CN", "IN", "BR", "RU", "IR")),
-    aes(x = all, y = oa, label = country_code),
-            min.segment.length = .15,
-    force = 5,
-    segment.color = "grey10",
-    segment.size = 0.5,
-    segment.alpha = 0.8,
-    color = "grey10", 
-    size = 3,
-    box.padding = 0.3) +
-  scale_y_log10(labels =  scales::number_format(big.mark = ",")) +
-  scale_x_log10(labels =  scales::number_format(big.mark = ",")) +
-  theme_minimal() +
-  labs(x = "Total articles in hybrid journals in 2022", y = "OA articles")
-
-oecd_stats <- readr::read_csv(here::here("data", "oecd_stats.csv"))
-oecd_stats_df <- oecd_stats |>
-  filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
-  filter(issn_l %in% hoaddata::cc_articles$issn_l) |>
-  group_by(cr_year, cat) |>
-  summarise(all = sum(articles, na.rm = TRUE), oa = sum(oa,na.rm = TRUE), ta_oa = sum(ta_oa, na.rm = TRUE)) |>
-  mutate(closed = all - oa,
-         oa_other = oa - ta_oa) |>
-  pivot_longer(cols = c(ta_oa, oa_other, closed), names_to = "oa_type") |>
-  mutate(cat = ifelse(is.na(cat), "Other", cat))
-
-country_bar <- ggplot(oecd_stats_df, aes(gsub("^20", "'", as.character(cr_year)), value, group = oa_type, fill = oa_type)) +
-  geom_area(color = "#000000",
-            linewidth = 0.15) +
-  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels =  scales::number_format(big.mark = ",")) +
-  facet_wrap(~ cat) +
-  scale_fill_manual(NULL, values = c(oa_other = "#b3b3b3a0", ta_oa = "#56B4E9", closed = "#e1e1e1a0"),
-                    labels = c(oa_other = "Other", ta_oa = "Agreement", closed = "NA"),  guide = guide_legend(reverse = TRUE)) +
-  theme_minimal() +
-  labs(x = "Publication year", y = "Total articles")
-```
-
-```{r country_by_publisher}
-publisher_mapping <-  hoaddata::jct_hybrid_jns |>
-  distinct(issn_l, esac_publisher) 
-subject_mapping <- subject_df |>
-  distinct(issn_l, top_level)
-country_jn_all <- oecd_stats |>
-  filter(!cr_year %in% c(2017, 2023), issn_l != "0027-8424") |>
-  # only with cc
-  filter(issn_l %in% hoaddata::cc_articles$issn_l) |>
-  inner_join(publisher_mapping, by = "issn_l") |>
-  inner_join(subject_mapping, by = "issn_l") |>
-  # Publisher encoding
-   mutate(
-    publisher = case_when(
-      esac_publisher == "Elsevier" ~ "Elsevier",
-      esac_publisher == "Springer Nature" ~ "Springer Nature",
-      esac_publisher == "Wiley" ~ "Wiley",
-      is.character(esac_publisher) ~ "Other"
-    )
-  ) |>
-  mutate(publisher =
-           forcats::fct_relevel(
-             publisher,
-             c("Elsevier", "Springer Nature", "Wiley", "Other") 
-           ))  |> 
-  group_by(cr_year, publisher, cat, top_level) |>
-  summarise(all = sum(articles),
-            oa = sum(oa),
-            ta_oa = sum(ta_oa), .groups = "drop") |>
-  mutate(closed = all - oa,
-         oa_other = oa - ta_oa)
-
-publ_subject_output_df <- country_jn_all |>
-  distinct(cr_year, publisher, top_level, cat, all) |>
-  filter(!top_level %in% c("Multidisciplinary", NA))
-
-pub_subject_output_plot <-  ggplot(publ_subject_output_df, aes(gsub("^20", "'", as.character(cr_year)), all, fill = cat)) +
-  geom_bar(stat = "identity", position = position_dodge2(preserve = "single")) +
-  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels =  scales::number_format(big.mark = ",")) +
-  scale_fill_manual(NULL, values = c("OECD" = "#d01c8b", "BRICS" = "#4dac26", "Other" = "#f7f7f7")) +
-  facet_grid(publisher ~ top_level) +
-  theme_minimal() +
-  labs(x = "Publication year", y = "Total articles")
-
-subject_output_plot <- publ_subject_output_df |>
-  group_by(cr_year, top_level, publisher) |>
-  summarise(all = sum(all)) |>
-  ggplot(aes(cr_year, all)) +
-  geom_bar(stat = "identity") +
-  facet_grid(top_level ~ publisher) +
-  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels =  scales::number_format(big.mark = ",")) +
-   theme_minimal() +
-  labs(x = "Publication year", y = "Total articles")
-
-publisher_output_plot <-
-  publ_subject_output_df |>
-  group_by(cr_year, cat, publisher) |>
-  summarise(all = sum(all)) |>
-  ggplot(aes(cr_year, all)) +
-  geom_bar(stat = "identity") +
-  facet_grid(cat ~ publisher) +
-  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels =  scales::number_format(big.mark = ",")) +
-   theme_minimal() +
-  labs(x = "Publication year", y = "Total articles")
-```
-
-Between 2018 and 2022, Western economies almost exclusively dominated hybrid open access publishing through transformative agreements. 
-During this period, first-authors affiliated with institutions from Organisation for Economic Co-operation and Development (OECD) member countries published `r oecd_stats_global |> filter(cat == "OECD") |> pull(oa)` open access articles in hybrid journals, representing `r oecd_stats_global |> filter(cat == "OECD") |> pull(oa) / n_oa * 100`% of the investigated open access articles.
-This disparity between OECD nations and other countries becomes even more evident when considering open access through transformative agreements, as  `r oecd_stats_global |> filter(cat == "OECD") |> pull(ta_oa)` of `r n_ta_oa`, or `r oecd_stats_global |> filter(cat == "OECD") |> pull(ta_oa) / n_ta_oa*100`% of open access articles were associated with such agreements.
-
-```{r country_patch, fig.asp=0.9}
-library(patchwork)
-wrap_plots(
-  A = country_bar,
-  B = country_scatter_22,
-  C = top_20_plot,
-  design = "
-  AC
-  BC",
- # guides = "collect",
-  widths = c(0.7, 0.3),
-  heights = c(0.5,0.5)
-) &
-  theme(
-    panel.grid.minor = element_blank(),
-                panel.border = element_rect(color = "grey50", fill = NA),
-    axis.title=element_text(size = 10),
-    legend.position = "top", 
-    legend.justification = "right") &
-    plot_annotation(tag_levels = 'A') 
-```
-
-Figure 2A shows the development of hybrid open access publishing by countries, comparing the OECD area with the BRICS, an intergovernmental organisation,  which comprised the countries Brazil, Russia, India, China and South Africa in 2022.
-The residual category "Other" includes the remaining countries. 
-From 2018 to 2022, the proportion of open access in hybrid journals increased from `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2018) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2018 to `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2022) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2022.
-On the other hand, BRICS recorded an low uptake, from `r country_bar$data |> filter(cat == "BRICS", oa_type != "closed", cr_year == 2018) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2018 to `r country_bar$data |> filter(cat == "BRICS", oa_type != "closed", cr_year == 2022) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2022.
-Despite rise of open access across OECD countries, the overall publication output decreased sharply, dropping to `r country_bar$data |> filter(cat == "OECD", cr_year == "2022") |> distinct(all) |> pull(all)` in 2022 after peaking `r country_bar$data |> filter(cat == "OECD", cr_year == "2020") |> distinct(all) |> pull(all)` articles in 2020.
-In stark contrast, the number of articles published in hybrid journals by first authors affiliated with institutions from BRICS countries increased steadily over the years, more than doubling from `r country_bar$data |> filter(cat == "BRICS", cr_year == "2018") |> distinct(all) |> pull(all)` in 2018 to `r country_bar$data |> filter(cat == "OECD", cr_year == "2022") |> distinct(all) |> pull(all)` in 2022.
-Upon closer examination, this trend can be observed across all big three publishers, although the shift towards BRICS is particularly evident in Elsevier's hybrid journal portfolio, in particular with regard to articles published in Physical Sciences journals (see Supplement).
-While OECD publication output in Elsevier's Physical Sciences journals declined from `r publ_subject_output_df |> filter(cr_year == 2018, top_level == "Physical Sciences", publisher == "Elsevier", cat == "OECD") |> pull(all)` articles in 2018 to `r publ_subject_output_df |> filter(cr_year == 2022, top_level == "Physical Sciences", publisher == "Elsevier", cat == "OECD") |> pull(all)` in 2022, BRICS output increased from `r publ_subject_output_df |> filter(cr_year == 2018, top_level == "Physical Sciences", publisher == "Elsevier", cat == "BRICS") |> pull(all)` to `r publ_subject_output_df |> filter(cr_year == 2022, top_level == "Physical Sciences", publisher == "Elsevier", cat == "BRICS") |> pull(all)` in the same five-year period.
-Furthermore, OECD publication output in Health Science Journals and Life Science journals stagnated across the investigated hybrid journal portfolios after a peak in 2020.
-
-To illustrate the situation in 2022, Figure B compares total publication output with the number of open access articles. 
-With `r country_scatter_df |> filter(cr_year == "2022", country_code == "CN") |> pull(all)` articles, China was the most productive country, followed by the United States (`r country_scatter_df |> filter(cr_year == "2022", country_code == "US") |> pull(all)` articles) and India (`r country_scatter_df |> filter(cr_year == "2022", country_code == "IN") |> pull(all)` articles).
-In contrast, West and Nord European countries published a considerable high number of open access articles, mainly due to transformative agreements. 
-Particularly, Germany, Great Britain, the Netherlands, Sweden, Switzerland and Spain recorded an above-average open access share as indicated by the linear trend line.
-As represented by the point size, as well as it can been seen in Figure C, transformative agreements contributed to this market position of these countries. 
-Interestingly, the United States had a notable open access market share of `r top_20_plot$data |> distinct(country_code, oa, all) |> mutate(prop = oa / sum(oa)) |> filter(country_code == "US") |> pull(prop) * 100`%, although transformative agreements contributed to a lesser extent.
-Similarly, China's open access market share of `r top_20_plot$data |> distinct(country_code, oa, all) |> mutate(prop = oa / sum(oa)) |> filter(country_code == "CN") |> pull(prop) * 100` in 2022 was comparable to that of the Netherlands, which was (`r  top_20_plot$data |> distinct(country_code, oa, all) |> mutate(prop = oa / sum(oa)) |> filter(country_code == "NL") |> pull(prop) * 100`%,).
-
-```{r country_top_20_plot, fig.asp=0.8}
-country_plot
-```
-
-Figure illustrates the development of hybrid open access from 2018 to 2022, highlighting the top 20 most productive countries in terms of articles published in hybrid journals that were included in transformative agreements over the five-year period.
-Notably, The Netherlands (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "NL") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), Sweden (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "SE") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), Poland (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "PL") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%) and Great Britain (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "GB") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%)) exhibited a relatively high level of uptake in 2018 which continued to increase in the following years.
-In 2022, Sweden had the highest proportion of open-access articles relative to its publication output (`r country_plot$data |> filter(cr_year == "2022", country_name_fct == "SE") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), followed by the Netherlands (`r country_plot$data |> filter(cr_year == "2022", country_name_fct == "NL") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%) and Switzerland (`r country_plot$data |> filter(cr_year == "2022", country_name_fct == "CH") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), with these countries benefiting from transformative agreements.
-In Germany, however, hybrid open access only began to increase from 2019 onwards after the successful negotiation of nationwide agreements with Wiley (July 2019) and Springer Nature (January 2020).
-Prior to this, only a few organisations had agreements in place, for the example the Max Planck Society with Springer Compact.
-
-Since 2021, there has been a general trend towards hybrid open access among many western countries, primarily driven by transformative agreements.
-However, proliferation of transformative agreements differe across these countries.
-For instance, Germany successfully negotiated an agreement with Elsevier not until 2023. 
-Additionally, publication limits or eligibility criteria for institutions and article types may explain why even countries with widespread agreement implementation do not achieve 100% hybrid open access.
-Interestingly, in Japan and the US other options than transformative agreements were the main driver for the increase in hybrid open access.
-Once again, the graph highlights countries with low hybrid open access, particularly non-OECD countries, where only a few or no agreements were in place.
-
-### Supplements
-
-```{r}
-pub_subject_output_plot
-```
-
-[^springercompact]: <https://web.archive.org/web/20180414062853id_/http://www.liber2015.org.uk/wp-content/uploads/2015/03/Springer-Compact.pdf>
+\end{document}

### 2024-01-28 data/003_oa_journal_status.R

@@ -1,0 +1,38 @@
+#' Exclusion of full OA journals
+library(readr)
+library(here)
+library(bigrquery)
+#' Full OA data sources
+#' 
+#' - `is_doaj` DOAJ (13 December 2023)
+#' - `openalex` OpenAlex (Source)
+#' - `gold oa` Bielefeld Gold OA (version 5) (https://pub.uni-bielefeld.de/download/2961544/2961545/issn_gold_oa_version_5.csv)
+#' 
+#' Download matched journal data from Google BigQuery
+oa_status_sql <- "SELECT DISTINCT
+  issn_l,
+  CASE
+    WHEN is_oa = TRUE OR is_in_doaj = TRUE THEN TRUE
+  ELSE
+  FALSE
+END
+  AS openalex,
+  is_doaj AS doaj,
+  gold_oa AS bielefeld
+FROM (
+  SELECT
+    jct.issn_l,
+    is_oa,
+    is_in_doaj,
+    is_doaj,
+    gold_oa
+  FROM
+    `subugoe-collaborative.openalex.sources` AS openalex
+  INNER JOIN
+    `hoa-article.jct.jct_hybrid_jns_new` AS jct
+  ON
+    openalex.issn_l = jct.issn_l )"
+
+tb <- bq_project_query("subugoe-collaborative", oa_status_sql)
+oa_journal_status <- bq_table_download(tb)
+readr::write_csv(oa_journal_status, here::here("data", "oa_journal_status.csv"))

### 2024-01-29 analysis/jn_classification.Rmd
Use data in methods

@@
   unnest(ajcs) |>
-  mutate(top_level_code = str_extract(ajcs, "\\d{2}")) |>
+  mutate(top_level_code = str_extract(ajcs, "\\d{2}"),
+         code =  str_extract(ajcs, "\\d{4}")) |>
   inner_join(ajsc_mapped, by = "top_level_code") |>

### 2024-01-29 analysis/manuscript.Rmd
Use data in methods

@@
 library(kableExtra)
+library(patchwork)
 ```
@@
 This study combines data from multiple publicly available data sources as visualised in Figure 1. 
-Data gathering involved transformative agreement data retrieved from the cOAlition S Journal Checker Tool, which provided information about journal portfolios and participating institutions. 
+Transformative agreement data retrieved from the cOAlition S Journal Checker Tool provided information about journal portfolios and participating institutions. 
 After identification of hybrid journals, Crossref served as the primary data source for article-level metadata including Creative Commons (CC) license information to determine open access articles on publisher websites.
@@
 ```{r journal_oa_exclusion}
+# ESAC 13 December snapshot
+esac_df <- readxl::read_xlsx(here::here("data", "esac.xlsx"), skip = 2) 
+#' OA journal check
 oa_journal_status <- readr::read_csv(here::here("data", "oa_journal_status.csv")) 
+oa_prop <- readr::read_csv(here::here("data", "oa_prop.csv"))
 
@@
     my_venn,
-    fill_color = c("grey55",
-          "steelblue2",
-          "lightsalmon"),
+    fill_color = c("#8dd3c7",
+          "#ffffb3",
+          "#bebada"),
     digits = 0,
+    stroke_color = "white",
     text_color = "black",
-    text_size = 3,
-    set_name_size = 4
-  )
-```
-
-Data gathering started with obtaining journals included in transformative agreements using the cOAlition S’ Journal Checker Tool publicly available Transformative Agreement Data dump. 
-The dump consisted of multiple online Google spreadsheets where each data file represented one agreement listed in the ESAC Transformative Agreement Registry. 
+    text_size = 2.5,
+    set_name_size = 2.5
+  ) +
+  labs(xlab = NULL, xlab = NULL)
+```
+
+Data gathering started with obtaining journals included in transformative agreements using the cOAlition S’ Journal Checker Tool publicly available Transformative Agreement Data dump.[^tadump] 
+The dump consisted of multiple online Google spreadsheets where each data file represented one agreement listed in the ESAC Transformative Agreement Registry.[^esacregistry] 
 Each file comprised information about journals and institutions involved per agreement. 
-Because the Journal Checker Tool only collected data about active transformative agreements, four different snapshots were used for this analysis:the archived version from July 2021, July 2022, May 2023 and the most current dump downloaded on 11 December 2022.
-This ensured that transformative agreements, which ended in 2021, were also included in the analysis. 
+Because the Journal Checker Tool only only provides data about active transformative agreements, removing data representing expired transformative agreements, four different snapshots were used for this analysis: self-archived version from July 2021, July 2022, and May 2023, as well as the most current dump downloaded on 11 December 2023.
+This ensured that transformative agreements, which ended in 2021, were included in the analysis.
 Agreements ending before 2021 could not be obtained.
-Overall, the combined Transformative Agreement Data dumps contained `r  length(unique(hoaddata::jct_hybrid_jns$esac_id))` agreements.
-
-The Transformative Agreement Data dumps linked agreements to journals represented by journal names and ISSNs. 
-After mapping ISSN variants to the corresponding linking ISSN (ISSN-L) as provided by The ISSN International Centre.
-Journals were associated to publishers using the ESAC ID, a unique identifier for transformative agreements in the ESAC Transformative Agreement Registry.
-
-Because transformative agreements can include both full open access and hybrid journals, journal data were complemented with information about a journal's open access status using multiple sources: the Directory of Open Access Journals (DOAJ) downloaded on 4 October 2022, OpenAlex (September 2022 dump) and the the Bielefeld list of GOLD OA journals.
-In total, `r n_oa_journals` full open access journals were excluded.
-Figure shows a substantial overlap between the three data sources of
+Overall, the combined Transformative Agreement Data dumps contained `r  length(unique(hoaddata::jct_hybrid_jns$esac_id))` out of `r length(unique(esac_df$ID))` agreements listed in the ESAC registry by December 2023.
+
+The Transformative Agreement Data dumps linked agreements to journals represented by journal names and ISSN. 
+After mapping ISSN variants to the corresponding linking ISSN (ISSN-L) as provided by The ISSN International Centre, journals were associated to publishers using the ESAC ID, a unique identifier for transformative agreements in the ESAC Transformative Agreement Registry.
+Because transformative agreements can include both full open access and hybrid journals, journal data were complemented with information about a journal's open access status using multiple sources: the Directory of Open Access Journals (DOAJ) downloaded on 12 December 2023[^doaj], OpenAlex (November 2023) and the the Bielefeld list of GOLD OA journals [@goldoa].
+In total, `r n_oa_journals` full open access journals were excluded based on journal title matching.
+Figure shows an overlap between the three data sources of
 `r round(length(unique(Reduce(intersect, my_venn))) / length(unique(Reduce(union, my_venn)))* 100,0)`%.
@@
 ## Article metadata
+
+After identifying hybrid journals included in transformative agreements, article metadata was retrieved from Crossref and mapped to OpenAlex to obtain affiliation metadata. 
+The Crossref November 2023 database snapshot was used to determine the journals’ yearly publication volume for the 5-year period 2018 to 2022. 
+Because Crossref metadata did not distinguish between research articles and other types of journal content, only included articles published in regular issues indicated by non-numeric pagination were included.
+Furthermore, an expanded version of Unpaywall’s paratext recognition approach was applied to exclude non-scholarly journal content such as table of contents.
+
+Open access articles in hybrid journals were identified through Creative Commons (CC) license information in Crossref metadata and mapped them to the different CC variants. 
+License information relative to the "accepted manuscript (AM)" version were not considered. 
+Crossref was used for open access identification because transformative agreements generally require publishers to deposit CC license information.
+Comparing Crossref license coverage with OpenAlex, which re-uses open access evidence from Unpaywall, a widely used open access discovery service, revealed a slight discrepancy between these two data sources.
+Overall, `r hoaddata::cr_upw |> filter(cr_year %in% 2018:2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, cr_hybrid_total) |> pull(cr_hybrid_total) |> sum()` articles with CC license were retrieved using Crossref, while `r hoaddata::cr_upw |> filter(cr_year %in% 2018:2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, upw_hybrid_total) |> pull(upw_hybrid_total) |> sum()` articles were tagged as "hybrid" in OpenAlex.
+The biggest differences concerned articles published between 2018 and 2020, and among some publishers that presumably did not provide CC license information to Crossref including AIP Publishing, American Physiological Society and Emerald.
+Crossref license metadata was more complete with regard to articles from the publisher Wiley.
+During the data collection, inconsistent open access status information in previous OpenAlex version was observed [@jahn2023], which might explain this discrepancy.
+
+Next, the number and proportion of open access articles per journal was calculated. 
+To improve the identification of hybrid journals, journals with an open access proportion above 95% were excluded.
+This further step allowed to remove additional `r length(unique(oa_prop$issn_l))` full open access journals.
+
+Affiliation data is a crucial data point for the planning and evaluation of transformative agreements [@Schimmer_2015; @Geschuhn_2017]. 
+Author affiliations were retrieved from OpenAlex.
+Generally, corresponding authors are responsible to arrange open access publication [@Borrego_2020].
+However, because of low coverage, this study focuses on first author and their affiliations.
+First authors typically contribute most to a paper and are often considered lead author research papers [@Larivi_re_2016]. 
+Overall, around 90% of studied articles had first author affiliation metadata in OpenAlex, whereas the coverage of articles with corresponding author information was around 54%.
+
+```{r cr_stats_overview}
+cr_stats <- readr::read_csv(here::here("data", "cr_stats.csv"))
+cr_stats_plot <- cr_stats |>
+  mutate(cat = case_when(
+    is.na(src) & oa_status == "gold" ~ "gold",
+    src == "hoad" ~ "hoad",
+    .default = "other"
+  )) |>
+  group_by(cr_year, cat) |>
+  summarise(n = sum(n_articles)) |>
+  mutate(cat = factor(cat, levels = c("hoad", "other", "gold"))) |>
+  mutate(prop = n / sum(n)) |>
+  ggplot(aes(gsub("^20", "'", as.character(cr_year)), n, fill = cat)) +
+  geom_bar(
+    position = position_stack(reverse = TRUE),
+    color = "black",
+    stat = "identity"
+  ) +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.05)),
+                     labels =  scales::number_format(big.mark = ",")) +
+  scale_fill_manual(
+    "Journal",
+    values = c(
+      "hoad" = "#5789B6",
+      "other" = "#D8D4C9",
+      "gold" = "#DA6524"
+    ),
+    labels = c(
+      "hoad" = "Hybrid (TA)",
+      "other" = "Other",
+      "gold" = "Full OA"
+    )
+  ) +
+  theme_minimal() +
+  labs(x = "Publication year", y = "Total articles") +
+  guides(fill = guide_legend(reverse = TRUE)) +
+  theme(
+    panel.grid.minor = element_blank(),
+    panel.border = element_rect(color = "grey50", fill = NA),
+    axis.title = element_text(size = 10)#,
+  #  legend.position = "right",
+  #  legend.justification = "right"
+  ) 
+```
+
+In total, the compiled data set consists of `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l, cr_year, jn_all) |> pull(jn_all) |> sum()` articles published in `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l) |> nrow()` hybrid journals between 2018 and 2022.
+They represented around
+
+```{r method_fig, fig.asp=0.55}
+(p | cr_stats_plot) + plot_annotation(tag_levels = 'A') +  plot_layout(widths = c(0.6, 0.4))
+```
+<!-- To reflect the increasing trend towards multiple and cross-country affiliations (Hottenrott et al., 2021; Larivière et al., 2015), articles were counted once per country. -->
 
@@
 
-library(patchwork)
 layout = "
@@
 [^springercompact]: <https://web.archive.org/web/20180414062853id_/http://www.liber2015.org.uk/wp-content/uploads/2015/03/Springer-Compact.pdf>
+
+[^tadump]: <https://journalcheckertool.org/transformative-agreements/>
+
+[^esacregistry]: <https://esac-initiative.org/about/transformative-agreements/agreement-registry/>
+
+[^doaj]: <https://doaj.org/csv>

### 2024-01-29 analysis/manuscript.tex
Use data in methods

@@
 
+% number lines
+% \usepackage{lineno}
+% \linenumbers
+
 % bibliography
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newenvironment{CSLReferences}%
+  {}%
+  {\par}
 
@@
 
+% QSS wants numbered sections
+%\setcounter{secnumdepth}{2}
+
 \hypertarget{introduction}{%
@@
 This study combines data from multiple publicly available data sources
-as visualised in Figure 1. Data gathering involved transformative
-agreement data retrieved from the cOAlition S Journal Checker Tool,
-which provided information about journal portfolios and participating
-institutions. After identification of hybrid journals, Crossref served
-as the primary data source for article-level metadata including Creative
-Commons (CC) license information to determine open access articles on
-publisher websites. OpenAlex provided first author affiliation metadata
-that were subsequently linked to the transformative agreement data to
-identify open access articles published through these agreements. In
-total, the compiled data set consists of 8,942,669 articles published in
-12,858 hybrid journals between 2018 and 2022.
+as visualised in Figure 1. Transformative agreement data retrieved from
+the cOAlition S Journal Checker Tool provided information about journal
+portfolios and participating institutions. After identification of
+hybrid journals, Crossref served as the primary data source for
+article-level metadata including Creative Commons (CC) license
+information to determine open access articles on publisher websites.
+OpenAlex provided first author affiliation metadata that were
+subsequently linked to the transformative agreement data to identify
+open access articles published through these agreements. In total, the
+compiled data set consists of 8,942,669 articles published in 12,858
+hybrid journals between 2018 and 2022.
 
@@
 transformative agreements using the cOAlition S' Journal Checker Tool
-publicly available Transformative Agreement Data dump. The dump
-consisted of multiple online Google spreadsheets where each data file
-represented one agreement listed in the ESAC Transformative Agreement
-Registry. Each file comprised information about journals and
-institutions involved per agreement. Because the Journal Checker Tool
-only collected data about active transformative agreements, four
-different snapshots were used for this analysis:the archived version
-from July 2021, July 2022, May 2023 and the most current dump downloaded
-on 11 December 2022. This ensured that transformative agreements, which
-ended in 2021, were also included in the analysis. Agreements ending
-before 2021 could not be obtained. Overall, the combined Transformative
-Agreement Data dumps contained 729 agreements.
+publicly available Transformative Agreement Data dump.\footnote{\url{https://journalcheckertool.org/transformative-agreements/}}
+The dump consisted of multiple online Google spreadsheets where each
+data file represented one agreement listed in the ESAC Transformative
+Agreement Registry.\footnote{\url{https://esac-initiative.org/about/transformative-agreements/agreement-registry/}}
+Each file comprised information about journals and institutions involved
+per agreement. Because the Journal Checker Tool only only provides data
+about active transformative agreements, removing data representing
+expired transformative agreements, four different snapshots were used
+for this analysis: self-archived version from July 2021, July 2022, and
+May 2023, as well as the most current dump downloaded on 11 December
+2023. This ensured that transformative agreements, which ended in 2021,
+were included in the analysis. Agreements ending before 2021 could not
+be obtained. Overall, the combined Transformative Agreement Data dumps
+contained 729 out of 869 agreements listed in the ESAC registry by
+December 2023.
 
 The Transformative Agreement Data dumps linked agreements to journals
-represented by journal names and ISSNs. After mapping ISSN variants to
+represented by journal names and ISSN. After mapping ISSN variants to
 the corresponding linking ISSN (ISSN-L) as provided by The ISSN
-International Centre. Journals were associated to publishers using the
+International Centre, journals were associated to publishers using the
 ESAC ID, a unique identifier for transformative agreements in the ESAC
-Transformative Agreement Registry.
-
-Because transformative agreements can include both full open access and
-hybrid journals, journal data were complemented with information about a
-journal's open access status using multiple sources: the Directory of
-Open Access Journals (DOAJ) downloaded on 4 October 2022, OpenAlex
-(September 2022 dump) and the the Bielefeld list of GOLD OA journals. In
-total, 3,439 full open access journals were excluded. Figure shows a
-substantial overlap between the three data sources of 72\%. However,
-combining different data sources considerably extended the journal
-matching to exclude full open access journals from transformative
-agreements. The Gold OA journals dataset alone adds 176 journals, while
-the DOAJ comprised 10 full open access journals not listed in either of
-the other two sources. These full open access journals were mostly
-launched in 2022.
+Transformative Agreement Registry. Because transformative agreements can
+include both full open access and hybrid journals, journal data were
+complemented with information about a journal's open access status using
+multiple sources: the Directory of Open Access Journals (DOAJ)
+downloaded on 12 December 2023\footnote{\url{https://doaj.org/csv}},
+OpenAlex (November 2023) and the the Bielefeld list of GOLD OA journals
+(Bruns et al., 2022). In total, 3,439 full open access journals were
+excluded based on journal title matching. Figure shows an overlap
+between the three data sources of 72\%. However, combining different
+data sources considerably extended the journal matching to exclude full
+open access journals from transformative agreements. The Gold OA
+journals dataset alone adds 176 journals, while the DOAJ comprised 10
+full open access journals not listed in either of the other two sources.
+These full open access journals were mostly launched in 2022.
 
@@
 \subsection{Article metadata}\label{article-metadata}}
+
+After identifying hybrid journals included in transformative agreements,
+article metadata was retrieved from Crossref and mapped to OpenAlex to
+obtain affiliation metadata. The Crossref November 2023 database
+snapshot was used to determine the journals' yearly publication volume
+for the 5-year period 2018 to 2022. Because Crossref metadata did not
+distinguish between research articles and other types of journal
+content, only included articles published in regular issues indicated by
+non-numeric pagination were included. Furthermore, an expanded version
+of Unpaywall's paratext recognition approach was applied to exclude
+non-scholarly journal content such as table of contents.
+
+Open access articles in hybrid journals were identified through Creative
+Commons (CC) license information in Crossref metadata and mapped them to
+the different CC variants. License information relative to the
+``accepted manuscript (AM)'' version were not considered. Crossref was
+used for open access identification because transformative agreement
+standards require publishers to deposit CC license information.
+Comparing Crossref license coverage with OpenAlex, which re-uses open
+access evidence from Unpaywall, a widely used open access discovery
+service, revealed a slight discrepancy between these two data sources.
+Overall, 847,107 articles with CC license were retrieved using Crossref,
+while 1,055,765 articles were tagged as ``hybrid'' in OpenAlex. The
+biggest differences concerned articles published between 2018 and 2020,
+and among some publishers that presumably did not provide CC license
+information to Crossref including AIP Publishing, American Physiological
+Society and Emerald (\textbf{ludo?}). Crossref license metadata was more
+complete with regard to articles from the publisher Wiley. During the
+data collection, inconsistent open access status information in previous
+OpenAlex version was observed (Jahn et al., 2023), which might explain
+this discrepancy.
+
+Next, the number and proportion of open access articles per journal was
+calculated. To improve the identification of hybrid journals, journals
+with an open access proportion above 95\% were excluded. This further
+step allowed to remove additional 241 full open access journals.
+
+Affiliation data is a crucial data point for the planning and evaluation
+of transformative agreements (Geschuhn \& Stone, 2017; Schimmer et al.,
+2015). Author affiliations were retrieved from OpenAlex. Generally,
+corresponding authors are responsible to arrange open access publication
+(Borrego et al., 2020). However, because of low coverage, this study
+focuses on first author and their affiliations. First authors typically
+contribute most to a paper and are often considered lead author research
+papers (Larivière et al., 2016). Overall, around 90\% of studied
+articles had first author affiliation metadata in OpenAlex, whereas the
+coverage of articles with corresponding author information was around
+54\%.
+
+In total, the compiled data set consists of 8,922,146 articles published
+in 12,857 hybrid journals between 2018 and 2022. They represented around
+
+\begin{center}\includegraphics[width=0.99\linewidth]{fig/method_fig-1} \end{center}
 
@@
 
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-Borrego_2020}{}}%
+Borrego, Á., Anglada, L., \& Abadal, E. (2020). Transformative
+agreements: Do they pave the way to open access? \emph{Learned
+Publishing}. \url{https://doi.org/10.1002/leap.1347}
+
+\leavevmode\vadjust pre{\hypertarget{ref-goldoa}{}}%
+Bruns, A., Cakir, Y., Kaya, S., \& Beidaghi, S. (2022).
+\emph{{ISSN-Matching of Gold OA Journals (ISSN-GOLD-OA) 5.0}}. Bielefeld
+University. \url{https://doi.org/10.4119/unibi/2961544}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Geschuhn_2017}{}}%
+Geschuhn, K., \& Stone, G. (2017). It's the workflows, stupid! What is
+required to make {``offsetting''} work for the open access transition.
+\emph{Insights the {UKSG} Journal}, \emph{30}(3), 103--114.
+\url{https://doi.org/10.1629/uksg.391}
+
+\leavevmode\vadjust pre{\hypertarget{ref-jahn2023}{}}%
+Jahn, N., Haupka, N., \& Hobert, A. (2023). \emph{Analysing and
+reclassifying open access information in OpenAlex}. Blog post.
+\url{https://subugoe.github.io/scholcomm_analytics/posts/oalex_oa_status/}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Larivi_re_2016}{}}%
+Larivière, V., Desrochers, N., Macaluso, B., Mongeon, P., Paul-Hus, A.,
+\& Sugimoto, C. R. (2016). Contributorship and division of labor in
+knowledge production. \emph{Social Studies of Science}, \emph{46}(3),
+417--435. \url{https://doi.org/10.1177/0306312716650046}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Schimmer_2015}{}}%
+Schimmer, R., Geschuhn, K., \& Vogler, A. (2015). \emph{{Disrupting the
+subscription journals'business model for the necessary large-scale
+transformation to open access}}. Max Planck Digital Library.
+\url{https://doi.org/10.17617/1.3}
+
+\end{CSLReferences}
+
 \end{document}

### 2024-01-29 data/002_total_article_volume.R

@@ -7,6 +7,7 @@ my_sql <- "WITH
   hoad AS (
   SELECT
     doi,
+    issn_l,
     'hoad' AS src
   FROM
     `hoa-article.hoaddata_nov23.cc_md` AS hoad ),
@@ -35,7 +36,7 @@ SELECT
   COUNT(DISTINCT crossref.doi) AS n_articles,
   cr_year,
   src,
-  open_access.oa_status
+  open_access.oa_status 
 FROM
   crossref
 LEFT JOIN
@@ -43,8 +44,8 @@ LEFT JOIN
 ON
   crossref.doi = openalex.doi
 WHERE
-  cr_year BETWEEN 2018
-  AND 2022
+  (cr_year BETWEEN 2018
+  AND 2022) AND NOT primary_location.source.issn_l = '0027-8424'
 GROUP BY
   cr_year,
   src,
@@ -54,31 +55,5 @@ ORDER BY
 
 tb <- bq_project_query("subugoe-collaborative", my_sql)
 cr_stats <- bq_table_download(tb)
-
-my_df <- cr_stats |>
-  mutate(cat = case_when(
-    is.na(src) & oa_status == "gold" ~ "gold",
-    src == "hoad" ~ "hoad",
-    .default = "other"
-  )) |> 
-  group_by(cr_year, cat) |>
-  summarise(n = sum(n_articles)) |>
-  mutate(cat = factor(cat, levels = c("hoad", "other", "gold"))) |>
-  mutate(prop = n / sum(n) )
-
-ggplot(my_df, aes(cr_year, n, fill = cat)) +
-  geom_bar(position = position_stack(reverse = TRUE), color = "black", stat = "identity") +
-  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels =  scales::number_format(big.mark = ",")) +
-  scale_fill_manual("Journal type", values = c("hoad" = "#5789B6", "other" = "#D8D4C9", "gold" = "#DA6524"),
-                    labels = c("hoad" = "Hybrid in TA", "other" = "Other", "gold" = "Full OA")) +
-  theme_minimal() +
-  labs(x = "Publication year", y = "Total articles") +
-  guides(fill = guide_legend(reverse = TRUE)) +
-  theme(
-    panel.grid.minor = element_blank(),
-    panel.border = element_rect(color = "grey50", fill = NA),
-    axis.title=element_text(size = 10),
-    legend.position = "top", 
[…]

### 2024-01-29 data/003_oa_journal_status.R

@@ -36,3 +36,10 @@ FROM (
 tb <- bq_project_query("subugoe-collaborative", oa_status_sql)
 oa_journal_status <- bq_table_download(tb)
 readr::write_csv(oa_journal_status, here::here("data", "oa_journal_status.csv"))
+
+#' Obtain journals with an proportion of above 0.95
+tb <- bq_project_query("subugoe-collaborative", 
+                       query = "SELECT * FROM `hoa-article.hoaddata_nov23.cc_oa_prop`")
+oa_prop <- bq_table_download(tb)
+readr::write_csv(oa_prop, here::here("data", "oa_prop.csv"))
+

### 2024-02-06 analysis/manuscript.Rmd
re compile

@@
 OpenAlex provided first author affiliation metadata that were subsequently linked to the transformative agreement data to identify open access articles published through these agreements.
-In total, the compiled data set consists of `r hoaddata::jn_ind |> filter(!cr_year %in% c("2017", "2023")) |> distinct(issn_l, cr_year, jn_all) |> pull(jn_all) |> sum()` articles published in `r hoaddata::jn_ind |> filter(!cr_year %in% c("2017", "2023")) |> distinct(issn_l) |> nrow()` hybrid journals between 2018 and 2022.
 
@@
 
-Data gathering started with obtaining journals included in transformative agreements using the cOAlition S’ Journal Checker Tool publicly available Transformative Agreement Data dump.[^tadump] 
+Data gathering started with obtaining journals included in transformative agreements using the publicly available Transformative Agreement Data dump[^tadump] used by the cOAlition S Journal Checker Tool.[^checkertool]
 The dump consisted of multiple online Google spreadsheets where each data file represented one agreement listed in the ESAC Transformative Agreement Registry.[^esacregistry] 
 Each file comprised information about journals and institutions involved per agreement. 
-Because the Journal Checker Tool only only provides data about active transformative agreements, removing data representing expired transformative agreements, four different snapshots were used for this analysis: self-archived version from July 2021, July 2022, and May 2023, as well as the most current dump downloaded on 11 December 2023.
+Because the Journal Checker Tool only provided data about active transformative agreements, data representing expired transformative agreements were constantly removed.
+Therefore, four different snapshots were used for this analysis: self-archived version from July 2021, July 2022, and May 2023, as well as the most current dump downloaded on 11 December 2023.
 This ensured that transformative agreements, which ended in 2021, were included in the analysis.
@@
 
-The Transformative Agreement Data dumps linked agreements to journals represented by journal names and ISSN. 
+The Transformative Agreement Data dumps link agreements to journals represented by journal names and ISSN. 
 After mapping ISSN variants to the corresponding linking ISSN (ISSN-L) as provided by The ISSN International Centre, journals were associated to publishers using the ESAC ID, a unique identifier for transformative agreements in the ESAC Transformative Agreement Registry.
-Because transformative agreements can include both full open access and hybrid journals, journal data were complemented with information about a journal's open access status using multiple sources: the Directory of Open Access Journals (DOAJ) downloaded on 12 December 2023[^doaj], OpenAlex (November 2023) and the the Bielefeld list of GOLD OA journals [@goldoa].
-In total, `r n_oa_journals` full open access journals were excluded based on journal title matching.
+Because transformative agreements can include both full open access and hybrid journals, the data were complemented with information about a journal's open access status using multiple sources: the Directory of Open Access Journals (DOAJ) downloaded on 12 December 2023[^doaj], OpenAlex (November 2023) and the the Bielefeld list of GOLD OA journals [@goldoa].
+In total, `r n_oa_journals` full open access journals were excluded based on ISSN matching.
 Figure shows an overlap between the three data sources of
@@
 
-After identifying hybrid journals included in transformative agreements, article metadata was retrieved from Crossref and mapped to OpenAlex to obtain affiliation metadata. 
-The Crossref November 2023 database snapshot was used to determine the journals’ yearly publication volume for the 5-year period 2018 to 2022. 
-Because Crossref metadata did not distinguish between research articles and other types of journal content, only included articles published in regular issues indicated by non-numeric pagination were included.
+After identifying hybrid journals included in transformative agreements, article metadata was retrieved from Crossref and mapped to OpenAlex for obtaining author metadata including affiliations. 
+The Crossref November 2023 database snapshot was used to determine the journals’ yearly publication volume for the five-year period 2018 to 2022. 
+Because Crossref metadata lacked information to distinguish between research articles and other types of journal content, only articles published in regular issues indicated by non-numeric pagination were included.
 Furthermore, an expanded version of Unpaywall’s paratext recognition approach was applied to exclude non-scholarly journal content such as table of contents.
 
-Open access articles in hybrid journals were identified through Creative Commons (CC) license information in Crossref metadata and mapped them to the different CC variants. 
+Open access articles in hybrid journals were identified through Creative Commons (CC) license information in Crossref metadata. 
 License information relative to the "accepted manuscript (AM)" version were not considered. 
@@
 Overall, `r hoaddata::cr_upw |> filter(cr_year %in% 2018:2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, cr_hybrid_total) |> pull(cr_hybrid_total) |> sum()` articles with CC license were retrieved using Crossref, while `r hoaddata::cr_upw |> filter(cr_year %in% 2018:2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, upw_hybrid_total) |> pull(upw_hybrid_total) |> sum()` articles were tagged as "hybrid" in OpenAlex.
-The biggest differences concerned articles published between 2018 and 2020, and among some publishers that presumably did not provide CC license information to Crossref including AIP Publishing, American Physiological Society and Emerald.
-Crossref license metadata was more complete with regard to articles from the publisher Wiley.
-During the data collection, inconsistent open access status information in previous OpenAlex version was observed [@jahn2023], which might explain this discrepancy.
-
-Next, the number and proportion of open access articles per journal was calculated. 
+The biggest differences concerned articles published between 2018 and 2020.
+In 2022, however, Crossref and OpenAlex open access numbers only differ slightly (`r hoaddata::cr_upw |> filter(cr_year %in% 2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, cr_hybrid_total) |> pull(cr_hybrid_total) |> sum()` records using Crossref vs.
+`r hoaddata::cr_upw |> filter(cr_year %in% 2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, upw_hybrid_total) |> pull(upw_hybrid_total) |> sum()` in OpenAlex).
+Upon inspection, notable difference could be furthermore observed among some publishers that presumably did not provide CC license information to Crossref including AIP Publishing, American Physiological Society and Emerald.
+Crossref license metadata was more complete with regard to articles from the publisher Wiley and American Chemical Society.
+<!-- During the data collection, inconsistent open access status information in previous OpenAlex versions was observed [@jahn2023], which might explain this discrepancy.-->
+
+After retrieving article metadata, the publication volume including open access was calculated per journal. 
 To improve the identification of hybrid journals, journals with an open access proportion above 95% were excluded.
@@
 
-Affiliation data is a crucial data point for the planning and evaluation of transformative agreements [@Schimmer_2015; @Geschuhn_2017]. 
+Author metadata are crucial for the planning and evaluation of transformative agreements [@Schimmer_2015; @Geschuhn_2017], in particular information about corresponding authors who are generally responsible to arrange open access publication [@Borrego_2020].
 Author affiliations were retrieved from OpenAlex.
-Generally, corresponding authors are responsible to arrange open access publication [@Borrego_2020].
-However, because of low coverage, this study focuses on first author and their affiliations.
+However, because of low coverage, this study focused on first authors and their affiliations.
 First authors typically contribute most to a paper and are often considered lead author research papers [@Larivi_re_2016]. 
@@
 
+To assess the impact of transformative agreements to hybrid open access, eligible institutions per agreement as listed in the Transformative Agreement Data dump. 
+According to the documentation, data about participating institutions were crowd-sourced from the agreements and consortia that successfully negotiated an agreement.
+Eligible were automatically matched to first author affiliations recorded by OpenAlex using the ROR-ID.
+The matching also took into account the duration of an agreement according to the ESAC registry.
+Upon inspection, Transformative Agreement Data did not cover  associated institutions comprehensively like university hospitals or institutes of large research organisations like the Max Planck Society.
+To improve the matching, Transformative Agreement Data was automatically enriched with associated institutions.
+
 In total, the compiled data set consists of `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l, cr_year, jn_all) |> pull(jn_all) |> sum()` articles published in `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l) |> nrow()` hybrid journals between 2018 and 2022.
-They represented around
+Hybrid journals included in transformative agreements represented `r cr_stats |> filter(src == "hoad") |> pull(n_articles) |> sum() / cr_stats |> pull(n_articles) |> sum() * 100`%  of total global output over the same time period according to Crossref, while full open access journals
 
@@
 [^doaj]: <https://doaj.org/csv>
+
+[^checkertool]: <https://www.coalition-s.org/blog/enabling-accurate-results-within-the-journal-checker-tool/>

### 2024-02-06 analysis/manuscript.tex
re compile

@@
 subsequently linked to the transformative agreement data to identify
-open access articles published through these agreements. In total, the
-compiled data set consists of 8,942,669 articles published in 12,858
-hybrid journals between 2018 and 2022.
+open access articles published through these agreements.
 
@@
 Data gathering started with obtaining journals included in
-transformative agreements using the cOAlition S' Journal Checker Tool
-publicly available Transformative Agreement Data dump.\footnote{\url{https://journalcheckertool.org/transformative-agreements/}}
+transformative agreements using the publicly available Transformative
+Agreement Data dump\footnote{\url{https://journalcheckertool.org/transformative-agreements/}}
+used by the cOAlition S Journal Checker Tool.\footnote{\url{https://www.coalition-s.org/blog/enabling-accurate-results-within-the-journal-checker-tool/}}
 The dump consisted of multiple online Google spreadsheets where each
@@
 Each file comprised information about journals and institutions involved
-per agreement. Because the Journal Checker Tool only only provides data
-about active transformative agreements, removing data representing
-expired transformative agreements, four different snapshots were used
-for this analysis: self-archived version from July 2021, July 2022, and
-May 2023, as well as the most current dump downloaded on 11 December
-2023. This ensured that transformative agreements, which ended in 2021,
-were included in the analysis. Agreements ending before 2021 could not
-be obtained. Overall, the combined Transformative Agreement Data dumps
-contained 729 out of 869 agreements listed in the ESAC registry by
-December 2023.
+per agreement. Because the Journal Checker Tool only provided data about
+active transformative agreements, data representing expired
+transformative agreements were constantly removed. Therefore, four
+different snapshots were used for this analysis: self-archived version
+from July 2021, July 2022, and May 2023, as well as the most current
+dump downloaded on 11 December 2023. This ensured that transformative
+agreements, which ended in 2021, were included in the analysis.
+Agreements ending before 2021 could not be obtained. Overall, the
+combined Transformative Agreement Data dumps contained 729 out of 869
+agreements listed in the ESAC registry by December 2023.
 
-The Transformative Agreement Data dumps linked agreements to journals
+The Transformative Agreement Data dumps link agreements to journals
 represented by journal names and ISSN. After mapping ISSN variants to
@@
 Transformative Agreement Registry. Because transformative agreements can
-include both full open access and hybrid journals, journal data were
+include both full open access and hybrid journals, the data were
 complemented with information about a journal's open access status using
@@
 (Bruns et al., 2022). In total, 3,439 full open access journals were
-excluded based on journal title matching. Figure shows an overlap
-between the three data sources of 72\%. However, combining different
-data sources considerably extended the journal matching to exclude full
-open access journals from transformative agreements. The Gold OA
-journals dataset alone adds 176 journals, while the DOAJ comprised 10
-full open access journals not listed in either of the other two sources.
-These full open access journals were mostly launched in 2022.
+excluded based on ISSN matching. Figure shows an overlap between the
+three data sources of 72\%. However, combining different data sources
+considerably extended the journal matching to exclude full open access
+journals from transformative agreements. The Gold OA journals dataset
+alone adds 176 journals, while the DOAJ comprised 10 full open access
+journals not listed in either of the other two sources. These full open
+access journals were mostly launched in 2022.
 
@@
 After identifying hybrid journals included in transformative agreements,
-article metadata was retrieved from Crossref and mapped to OpenAlex to
-obtain affiliation metadata. The Crossref November 2023 database
-snapshot was used to determine the journals' yearly publication volume
-for the 5-year period 2018 to 2022. Because Crossref metadata did not
-distinguish between research articles and other types of journal
-content, only included articles published in regular issues indicated by
-non-numeric pagination were included. Furthermore, an expanded version
-of Unpaywall's paratext recognition approach was applied to exclude
-non-scholarly journal content such as table of contents.
+article metadata was retrieved from Crossref and mapped to OpenAlex for
+obtaining author metadata including affiliations. The Crossref November
+2023 database snapshot was used to determine the journals' yearly
+publication volume for the five-year period 2018 to 2022. Because
+Crossref metadata lacked information to distinguish between research
+articles and other types of journal content, only articles published in
+regular issues indicated by non-numeric pagination were included.
+Furthermore, an expanded version of Unpaywall's paratext recognition
+approach was applied to exclude non-scholarly journal content such as
+table of contents.
 
 Open access articles in hybrid journals were identified through Creative
-Commons (CC) license information in Crossref metadata and mapped them to
-the different CC variants. License information relative to the
-``accepted manuscript (AM)'' version were not considered. Crossref was
-used for open access identification because transformative agreement
-standards require publishers to deposit CC license information.
-Comparing Crossref license coverage with OpenAlex, which re-uses open
-access evidence from Unpaywall, a widely used open access discovery
-service, revealed a slight discrepancy between these two data sources.
-Overall, 847,107 articles with CC license were retrieved using Crossref,
-while 1,055,765 articles were tagged as ``hybrid'' in OpenAlex. The
-biggest differences concerned articles published between 2018 and 2020,
-and among some publishers that presumably did not provide CC license
-information to Crossref including AIP Publishing, American Physiological
-Society and Emerald (\textbf{ludo?}). Crossref license metadata was more
-complete with regard to articles from the publisher Wiley. During the
-data collection, inconsistent open access status information in previous
-OpenAlex version was observed (Jahn et al., 2023), which might explain
-this discrepancy.
+Commons (CC) license information in Crossref metadata. License
+information relative to the ``accepted manuscript (AM)'' version were
+not considered. Crossref was used for open access identification because
+transformative agreements generally require publishers to deposit CC
+license information. Comparing Crossref license coverage with OpenAlex,
+which re-uses open access evidence from Unpaywall, a widely used open
+access discovery service, revealed a slight discrepancy between these
+two data sources. Overall, 742,369 articles with CC license were
+retrieved using Crossref, while 950,260 articles were tagged as
+``hybrid'' in OpenAlex. The biggest differences concerned articles
+published between 2018 and 2020. In 2022, however, Crossref and OpenAlex
+open access numbers only differ slightly (249,511 records using Crossref
+vs.~255,344 in OpenAlex). Upon inspection, notable difference could be
+furthermore observed among some publishers that presumably did not
+provide CC license information to Crossref including AIP Publishing,
+American Physiological Society and Emerald. Crossref license metadata
+was more complete with regard to articles from the publisher Wiley and
+American Chemical Society.
 
-Next, the number and proportion of open access articles per journal was
-calculated. To improve the identification of hybrid journals, journals
-with an open access proportion above 95\% were excluded. This further
-step allowed to remove additional 241 full open access journals.
+After retrieving article metadata, the publication volume including open
+access was calculated per journal. To improve the identification of
+hybrid journals, journals with an open access proportion above 95\% were
+excluded. This further step allowed to remove additional 241 full open
+access journals.
 
-Affiliation data is a crucial data point for the planning and evaluation
-of transformative agreements (Geschuhn \& Stone, 2017; Schimmer et al.,
-2015). Author affiliations were retrieved from OpenAlex. Generally,
-corresponding authors are responsible to arrange open access publication
-(Borrego et al., 2020). However, because of low coverage, this study
-focuses on first author and their affiliations. First authors typically
-contribute most to a paper and are often considered lead author research
-papers (Larivière et al., 2016). Overall, around 90\% of studied
-articles had first author affiliation metadata in OpenAlex, whereas the
-coverage of articles with corresponding author information was around
-54\%.
+Author metadata are crucial for the planning and evaluation of
+transformative agreements (Geschuhn \& Stone, 2017; Schimmer et al.,
+2015), in particular information about corresponding authors who are
+generally responsible to arrange open access publication (Borrego et
+al., 2020). Author affiliations were retrieved from OpenAlex. However,
+because of low coverage, this study focused on first authors and their
+affiliations. First authors typically contribute most to a paper and are
+often considered lead author research papers (Larivière et al., 2016).
+Overall, around 90\% of studied articles had first author affiliation
+metadata in OpenAlex, whereas the coverage of articles with
+corresponding author information was around 54\%.
+
+To assess the impact of transformative agreements to hybrid open access,
+eligible institutions per agreement as listed in the Transformative
+Agreement Data dump. According to the documentation, data about
+participating institutions were crowd-sourced from the agreements and
+consortia that successfully negotiated an agreement. Eligible were
+automatically matched to first author affiliations recorded by OpenAlex
+using the ROR-ID. The matching also took into account the duration of an
+agreement according to the ESAC registry. Upon inspection,
+Transformative Agreement Data did not cover associated institutions
+comprehensively like university hospitals or institutes of large
+research organisations like the Max Planck Society. To improve the
+matching, Transformative Agreement Data was automatically enriched with
+associated institutions.
 
 In total, the compiled data set consists of 8,922,146 articles published
-in 12,857 hybrid journals between 2018 and 2022. They represented around
+in 12,857 hybrid journals between 2018 and 2022. Hybrid journals
+included in transformative agreements represented 40\% of total global
+output over the same time period according to Crossref, while full open
+access journals
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-jahn2023}{}}%
-Jahn, N., Haupka, N., \& Hobert, A. (2023). \emph{Analysing and
-reclassifying open access information in OpenAlex}. Blog post.
-\url{https://subugoe.github.io/scholcomm_analytics/posts/oalex_oa_status/}
-
 \leavevmode\vadjust pre{\hypertarget{ref-Larivi_re_2016}{}}%

### 2024-02-11 analysis/manuscript.Rmd
Recompile

@@
 
-This study combines data from multiple publicly available data sources as visualised in Figure 1. 
-Transformative agreement data retrieved from the cOAlition S Journal Checker Tool provided information about journal portfolios and participating institutions. 
-After identification of hybrid journals, Crossref served as the primary data source for article-level metadata including Creative Commons (CC) license information to determine open access articles on publisher websites.
-OpenAlex provided first author affiliation metadata that were subsequently linked to the transformative agreement data to identify open access articles published through these agreements.
+This study combines data from multiple publicly available data sources as diagrammatically shown in Figure \ref{fig:data_workflow}. 
+Initially, transformative agreement data retrieved from the cOAlition S Journal Checker Tool provided information about journal portfolios and participating institutions. 
+After identification of hybrid journals, Crossref served as the primary data source for article-level metadata including Creative Commons (CC) license information to determine open access availability on publisher websites.
+To identify open access articles published through transformative agreements, first author affiliations from OpenAlex were subsequently linked to eligible institutions according to the transformative agreement data.
 
@@
 
-Data gathering started with obtaining journals included in transformative agreements using the publicly available Transformative Agreement Data dump[^tadump] used by the cOAlition S Journal Checker Tool.[^checkertool]
+Data gathering started with obtaining journals included in transformative agreements from the publicly available Transformative Agreement Data dump[^tadump] used by the cOAlition S Journal Checker Tool.[^checkertool]
 The dump consisted of multiple online Google spreadsheets where each data file represented one agreement listed in the ESAC Transformative Agreement Registry.[^esacregistry]
-Each file comprised information about journals and institutions involved per agreement. 
-Because the Journal Checker Tool only provided data about active transformative agreements, data representing expired transformative agreements were constantly removed.
-Therefore, four different snapshots were used for this analysis: self-archived version from July 2021, July 2022, and May 2023, as well as the most current dump downloaded on 11 December 2023.
-This ensured that transformative agreements, which ended in 2021, were included in the analysis.
-Agreements ending before 2021 could not be obtained.
-Overall, the combined Transformative Agreement Data dumps contained `r  length(unique(hoaddata::jct_hybrid_jns$esac_id))` out of `r length(unique(esac_df$ID))` agreements listed in the ESAC registry by December 2023.
+From the retrieved spreadsheet files, journals and institutions involved per agreement were obtained.
+
+A limitation of using the Journal Checker Tool to study the development of transformative agreements over time is that it only provided data about active transformative agreements.
+As a result, expired transformative agreements were constantly removed from the publicly available data dump. 
+To address this, four different snapshots were safeguarded and combined for this study: self-archived versions from July 2021, July 2022, and May 2023, as well as the most current dump downloaded on 11 December 2023.
+This ensured that transformative agreements, which ended from 2021 onwards, were included.
+<!-- Agreements ending before 2021 could not be obtained.-->
+Overall, the four combined Transformative Agreement Data dumps used in this study contained `r  length(unique(hoaddata::jct_hybrid_jns$esac_id))` out of `r length(unique(esac_df$ID))` agreements listed in the ESAC registry by December 2023.
 
 The Transformative Agreement Data dumps link agreements to journals represented by journal names and ISSN. 
-After mapping ISSN variants to the corresponding linking ISSN (ISSN-L) as provided by The ISSN International Centre, journals were associated to publishers using the ESAC ID, a unique identifier for transformative agreements in the ESAC Transformative Agreement Registry.
+After mapping ISSN variants to the corresponding linking ISSN (ISSN-L) as provided by the ISSN International Centre, journals were associated to publishers using the ESAC ID, a unique identifier for transformative agreements in the ESAC Transformative Agreement Registry.
 Because transformative agreements can include both full open access and hybrid journals, the data were complemented with information about a journal's open access status using multiple sources: the Directory of Open Access Journals (DOAJ) downloaded on 12 December 2023[^doaj], OpenAlex (November 2023) and the the Bielefeld list of GOLD OA journals [@goldoa].
+As shown in Figure \ref{fig:method_fig}A, combining different data sources considerably extended the journal matching to exclude full open access journals from transformative agreements.
 In total, `r n_oa_journals` full open access journals were excluded based on ISSN matching.
-Figure shows an overlap between the three data sources of
-`r round(length(unique(Reduce(intersect, my_venn))) / length(unique(Reduce(union, my_venn)))* 100,0)`%.
-However, combining different data sources considerably extended the journal matching to exclude full open access journals from transformative agreements.
-The Gold OA journals dataset alone adds `r length(unique(Reduce(setdiff, my_venn)))` journals, while the DOAJ comprised `r length(setdiff(my_venn$DOAJ, union(my_venn[["GOLD OA"]], my_venn$OpenAlex)))` full open access journals not listed in either of the other two sources.
+The overlap between the three data sources was `r round(length(unique(Reduce(intersect, my_venn))) / length(unique(Reduce(union, my_venn)))* 100,0)`%.
+The Gold OA journals dataset alone added `r length(unique(Reduce(setdiff, my_venn)))` journals, while the DOAJ comprised `r length(setdiff(my_venn$DOAJ, union(my_venn[["GOLD OA"]], my_venn$OpenAlex)))` full open access journals not listed in either of the other two sources.
 These full open access journals were mostly launched in 2022.
 
-## Article metadata
-
-After identifying hybrid journals included in transformative agreements, article metadata was retrieved from Crossref and mapped to OpenAlex for obtaining author metadata including affiliations. 
-The Crossref November 2023 database snapshot was used to determine the journals’ yearly publication volume for the five-year period 2018 to 2022. 
-Because Crossref metadata lacked information to distinguish between research articles and other types of journal content, only articles published in regular issues indicated by non-numeric pagination were included.
+## Article and author metadata
+
+After identifying hybrid journals included in transformative agreements, article metadata was retrieved from Crossref<!.
+<!-- and mapped to OpenAlex for obtaining author metadata including affiliations.
+--> November 2023 database snapshot for the five-year period 2018 to 2022.
+Because Crossref metadata lacked information to distinguish between original research articles including review and other types of journal content, which are often not covered by transformative agreements [@Borrego_2020], only articles published in regular issues indicated by non-numeric pagination were included.
 Furthermore, an expanded version of Unpaywall’s paratext recognition approach was applied to exclude non-scholarly journal content such as table of contents.
@@
 License information relative to the "accepted manuscript (AM)" version were not considered. 
-Crossref was used for open access identification because transformative agreements generally require publishers to deposit CC license information.
-Comparing Crossref license coverage with OpenAlex, which re-uses open access evidence from Unpaywall, a widely used open access discovery service, revealed a slight discrepancy between these two data sources.
+Crossref was used for open access identification because transformative agreements workflows generally require publishers to deliver CC license information to this DOI registration agency [@Geschuhn_2017].
+Comparing Crossref license coverage with OpenAlex, which re-uses open access evidence from Unpaywall, not all publisher provide such license information a widely used open access discovery service that also parses journal websites for open content licenses [@Piwowar_2018], demonstrates the ongoing challenges to identify open access articles in hybrid journals [@Mart_n_Mart_n_2018; @Jahn_2021].
 Overall, `r hoaddata::cr_upw |> filter(cr_year %in% 2018:2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, cr_hybrid_total) |> pull(cr_hybrid_total) |> sum()` articles with CC license were retrieved using Crossref, while `r hoaddata::cr_upw |> filter(cr_year %in% 2018:2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, upw_hybrid_total) |> pull(upw_hybrid_total) |> sum()` articles were tagged as "hybrid" in OpenAlex.
@@
 Crossref license metadata was more complete with regard to articles from the publisher Wiley and American Chemical Society.
-<!-- During the data collection, inconsistent open access status information in previous OpenAlex versions was observed [@jahn2023], which might explain this discrepancy.-->
+Finally, inconsistent open access status information in previous OpenAlex versions was observed [@jahn2023].
+According to the OpenAlex release notes, fixing this issue was still ongoing, which might also explain this discrepancy.
 
@@
 Author metadata are crucial for the planning and evaluation of transformative agreements [@Schimmer_2015; @Geschuhn_2017], in particular information about corresponding authors who are generally responsible to arrange open access publication [@Borrego_2020].
-Author affiliations were retrieved from OpenAlex.
+Here, affiliations were retrieved from OpenAlex.
 However, because of low coverage, this study focused on first authors and their affiliations.
@@
 
-To assess the impact of transformative agreements to hybrid open access, eligible institutions per agreement as listed in the Transformative Agreement Data dump. 
-According to the documentation, data about participating institutions were crowd-sourced from the agreements and consortia that successfully negotiated an agreement.
-Eligible were automatically matched to first author affiliations recorded by OpenAlex using the ROR-ID.
+To assess the impact of transformative agreements to hybrid open access, participating institutions from the Transformative Agreement Data dump, which were crowd-sourced from the agreements and consortia that successfully negotiated an agreement, were matched with first author affiliations recorded by OpenAlex using the ROR-ID.
 The matching also took into account the duration of an agreement according to the ESAC registry.
 Upon inspection, Transformative Agreement Data did not cover associated institutions comprehensively like university hospitals or institutes of large research organisations like the Max Planck Society.
-To improve the matching, Transformative Agreement Data was automatically enriched with associated institutions.
-
-In total, the compiled data set consists of `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l, cr_year, jn_all) |> pull(jn_all) |> sum()` articles published in `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l) |> nrow()` hybrid journals between 2018 and 2022.
-Hybrid journals included in transformative agreements represented `r cr_stats |> filter(src == "hoad") |> pull(n_articles) |> sum() / cr_stats |> pull(n_articles) |> sum() * 100`%  of total global output over the same time period according to Crossref, while full open access journals
-
-```{r method_fig, fig.asp=0.55}
+To improve the matching, Transformative Agreement Data was automatically enriched with associated institutions using OpenAlex's insitution data.
+
+In total, the compiled data set consists of `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l, cr_year, jn_all) |> pull(jn_all) |> sum()` articles published in `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l) |> nrow()` hybrid journals between 2018 and 2022 (see Figure \ref{fig:method_fig}.
+Hybrid journals included in transformative agreements represented `r cr_stats |> filter(src == "hoad") |> pull(n_articles) |> sum() / cr_stats |> pull(n_articles) |> sum() * 100`%  of total global output over the same time period according to Crossref, while full open access journals accounted for
+`r cr_stats |> filter(src == "gold") |> pull(n_articles) |> sum() / cr_stats |> pull(n_articles) |> sum() * 100`% of total article volume.
+
+```{r method_fig, fig.asp=0.55, fig.cap="Initial data characteristics.
+(A) Full open access journals included in transformative agreements by evidence source Directory of Open Access Journals (DOAJ), OpenAlex and the Bielefeld GOLD OA list.
+(B) Number of articles in Crossref by journal types.
+The blue bars show the overall article volume of hybrid journals in transformative agreements, which were initially included in the study, in comaprision with full open access journals according to OpenAlex.
+The remainder represents closed access journals not covered by transformative agreements."}
 (p | cr_stats_plot) + plot_annotation(tag_levels = 'A') +  plot_layout(widths = c(0.6, 0.4))
@@
 ## Data Analysis
+
+Throughout this mostly automated data gathering and analysis process, we used tools from the Tidyverse (Wickham et al., 2019) for the R programming language (R Core Team, 2020). 
+The resulting data is openly available through an R package, hoaddata. 
+Following @Marwick_2018, hoaddata contains not only the datasets used in the data analysis.
+It also includes codeused to compile the data by connecting it the cloud-based Google Big Query data warehouse, where the big scholarly data from Crossref, OpenAlex and Unpaywall were imported.
+To increase the computational reproducibility, the R package was automatically built using GitHub Actions, a continuous integration service.
 

### 2024-02-11 analysis/manuscript.tex
Recompile

@@
 This study combines data from multiple publicly available data sources
-as visualised in Figure 1. Transformative agreement data retrieved from
-the cOAlition S Journal Checker Tool provided information about journal
-portfolios and participating institutions. After identification of
-hybrid journals, Crossref served as the primary data source for
-article-level metadata including Creative Commons (CC) license
-information to determine open access articles on publisher websites.
-OpenAlex provided first author affiliation metadata that were
-subsequently linked to the transformative agreement data to identify
-open access articles published through these agreements.
+as diagrammatically shown in Figure \ref{fig:data_workflow}. Initially,
+transformative agreement data retrieved from the cOAlition S Journal
+Checker Tool provided information about journal portfolios and
+participating institutions. After identification of hybrid journals,
+Crossref served as the primary data source for article-level metadata
+including Creative Commons (CC) license information to determine open
+access availability on publisher websites. To identify open access
+articles published through transformative agreements, first author
+affiliations from OpenAlex were subsequently linked to eligible
+institutions according to the transformative agreement data.
 
@@
 Data gathering started with obtaining journals included in
-transformative agreements using the publicly available Transformative
+transformative agreements from the publicly available Transformative
 Agreement Data dump\footnote{\url{https://journalcheckertool.org/transformative-agreements/}}
@@
 Agreement Registry.\footnote{\url{https://esac-initiative.org/about/transformative-agreements/agreement-registry/}}
-Each file comprised information about journals and institutions involved
-per agreement. Because the Journal Checker Tool only provided data about
-active transformative agreements, data representing expired
-transformative agreements were constantly removed. Therefore, four
-different snapshots were used for this analysis: self-archived version
-from July 2021, July 2022, and May 2023, as well as the most current
-dump downloaded on 11 December 2023. This ensured that transformative
-agreements, which ended in 2021, were included in the analysis.
-Agreements ending before 2021 could not be obtained. Overall, the
-combined Transformative Agreement Data dumps contained 729 out of 869
-agreements listed in the ESAC registry by December 2023.
+From the retrieved spreadsheet files, journals and institutions involved
+per agreement were obtained.
+
+A limitation of using the Journal Checker Tool to study the development
+of transformative agreements over time is that it only provided data
+about active transformative agreements. As a result, expired
+transformative agreements were constantly removed from the publicly
+available data dump. To address this, four different snapshots were
+safeguarded and combined for this study: self-archived versions from
+July 2021, July 2022, and May 2023, as well as the most current dump
+downloaded on 11 December 2023. This ensured that transformative
+agreements, which ended from 2021 onwards, were included. Overall, the
+four combined Transformative Agreement Data dumps used in this study
+contained 729 out of 869 agreements listed in the ESAC registry by
+December 2023.
 
@@
 represented by journal names and ISSN. After mapping ISSN variants to
-the corresponding linking ISSN (ISSN-L) as provided by The ISSN
+the corresponding linking ISSN (ISSN-L) as provided by the ISSN
 International Centre, journals were associated to publishers using the
@@
 OpenAlex (November 2023) and the the Bielefeld list of GOLD OA journals
-(Bruns et al., 2022). In total, 3,439 full open access journals were
-excluded based on ISSN matching. Figure shows an overlap between the
-three data sources of 72\%. However, combining different data sources
-considerably extended the journal matching to exclude full open access
-journals from transformative agreements. The Gold OA journals dataset
-alone adds 176 journals, while the DOAJ comprised 10 full open access
-journals not listed in either of the other two sources. These full open
-access journals were mostly launched in 2022.
+(Bruns et al., 2022). As shown in Figure \ref{fig:method_fig}A,
+combining different data sources considerably extended the journal
+matching to exclude full open access journals from transformative
+agreements. In total, 3,439 full open access journals were excluded
+based on ISSN matching. The overlap between the three data sources was
+72\%. The Gold OA journals dataset alone added 176 journals, while the
+DOAJ comprised 10 full open access journals not listed in either of the
+other two sources. These full open access journals were mostly launched
+in 2022.
 
-\hypertarget{article-metadata}{%
-\subsection{Article metadata}\label{article-metadata}}
+\hypertarget{article-and-author-metadata}{%
+\subsection{Article and author
+metadata}\label{article-and-author-metadata}}
 
 After identifying hybrid journals included in transformative agreements,
-article metadata was retrieved from Crossref and mapped to OpenAlex for
-obtaining author metadata including affiliations. The Crossref November
-2023 database snapshot was used to determine the journals' yearly
-publication volume for the five-year period 2018 to 2022. Because
-Crossref metadata lacked information to distinguish between research
-articles and other types of journal content, only articles published in
-regular issues indicated by non-numeric pagination were included.
-Furthermore, an expanded version of Unpaywall's paratext recognition
-approach was applied to exclude non-scholarly journal content such as
-table of contents.
+article metadata was retrieved from Crossref\textless!. November 2023
+database snapshot for the five-year period 2018 to 2022. Because
+Crossref metadata lacked information to distinguish between original
+research articles including review and other types of journal content,
+which are often not covered by transformative agreements (Borrego et
+al., 2020), only articles published in regular issues indicated by
+non-numeric pagination were included. Furthermore, an expanded version
+of Unpaywall's paratext recognition approach was applied to exclude
+non-scholarly journal content such as table of contents.
 
@@
 not considered. Crossref was used for open access identification because
-transformative agreements generally require publishers to deposit CC
-license information. Comparing Crossref license coverage with OpenAlex,
-which re-uses open access evidence from Unpaywall, a widely used open
-access discovery service, revealed a slight discrepancy between these
-two data sources. Overall, 742,369 articles with CC license were
-retrieved using Crossref, while 950,260 articles were tagged as
+transformative agreements workflows generally require publishers to
+deliver CC license information to this DOI registration agency (Geschuhn
+\& Stone, 2017). Comparing Crossref license coverage with OpenAlex,
+which re-uses open access evidence from Unpaywall, not all publisher
+provide such license information a widely used open access discovery
+service that also parses journal websites for open content licenses
+(Piwowar et al., 2018), demonstrates the ongoing challenges to identify
+open access articles in hybrid journals (Jahn et al., 2021;
+Martín-Martín et al., 2018). Overall, 742,369 articles with CC license
+were retrieved using Crossref, while 950,260 articles were tagged as
 ``hybrid'' in OpenAlex. The biggest differences concerned articles
@@
 was more complete with regard to articles from the publisher Wiley and
-American Chemical Society.
+American Chemical Society. Finally, inconsistent open access status
+information in previous OpenAlex versions was observed (Jahn et al.,
+2023). According to the OpenAlex release notes, fixing this issue was
+still ongoing, which might also explain this discrepancy.
 
@@
 generally responsible to arrange open access publication (Borrego et
-al., 2020). Author affiliations were retrieved from OpenAlex. However,
+al., 2020). Here, affiliations were retrieved from OpenAlex. However,
 because of low coverage, this study focused on first authors and their
@@
 To assess the impact of transformative agreements to hybrid open access,
-eligible institutions per agreement as listed in the Transformative
-Agreement Data dump. According to the documentation, data about
-participating institutions were crowd-sourced from the agreements and
-consortia that successfully negotiated an agreement. Eligible were
-automatically matched to first author affiliations recorded by OpenAlex
-using the ROR-ID. The matching also took into account the duration of an
-agreement according to the ESAC registry. Upon inspection,
-Transformative Agreement Data did not cover associated institutions
-comprehensively like university hospitals or institutes of large
-research organisations like the Max Planck Society. To improve the
-matching, Transformative Agreement Data was automatically enriched with
-associated institutions.
+participating institutions from the Transformative Agreement Data dump,
+which were crowd-sourced from the agreements and consortia that
+successfully negotiated an agreement, were matched with first author
+affiliations recorded by OpenAlex using the ROR-ID. The matching also
+took into account the duration of an agreement according to the ESAC
+registry. Upon inspection, Transformative Agreement Data did not cover
+associated institutions comprehensively like university hospitals or
+institutes of large research organisations like the Max Planck Society.
+To improve the matching, Transformative Agreement Data was automatically
+enriched with associated institutions using OpenAlex's insitution data.
 
 In total, the compiled data set consists of 8,922,146 articles published
-in 12,857 hybrid journals between 2018 and 2022. Hybrid journals
-included in transformative agreements represented 40\% of total global
-output over the same time period according to Crossref, while full open
-access journals
+in 12,857 hybrid journals between 2018 and 2022 (see Figure
+\ref{fig:method_fig}. Hybrid journals included in transformative
+agreements represented 40\% of total global output over the same time
+period according to Crossref, while full open access journals accounted
+for 0\% of total article volume.
 
-\begin{center}\includegraphics[width=0.99\linewidth]{fig/method_fig-1} \end{center}
+\begin{figure}
+
+{\centering \includegraphics[width=0.99\linewidth]{fig/method_fig-1}
+
+}
+
+\caption{Initial data characteristics.
+(A) Full open access journals included in transformative agreements by evidence source Directory of Open Access Journals (DOAJ), OpenAlex and the Bielefeld GOLD OA list.
+(B) Number of articles in Crossref by journal types.
+The blue bars show the overall article volume of hybrid journals in transformative agreements, which were initially included in the study, in comaprision with full open access journals according to OpenAlex.
+The remainder represents closed access journals not covered by transformative agreements.}\label{fig:method_fig}
+\end{figure}
 
@@
 \subsection{Data Analysis}\label{data-analysis}}
+
+Throughout this mostly automated data gathering and analysis process, we
+used tools from the Tidyverse (Wickham et al., 2019) for the R
+programming language (R Core Team, 2020). The resulting data is openly
+available through an R package, hoaddata. Following Marwick et al.
+(2018), hoaddata contains not only the datasets used in the data
+analysis. It also includes codeused to compile the data by connecting it
+the cloud-based Google Big Query data warehouse, where the big scholarly
+data from Crossref, OpenAlex and Unpaywall were imported. To increase
+the computational reproducibility, the R package was automatically built
+using GitHub Actions, a continuous integration service.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-jahn2023}{}}%
+Jahn, N., Haupka, N., \& Hobert, A. (2023). \emph{Analysing and
+reclassifying open access information in OpenAlex}. Blog post.
+\url{https://subugoe.github.io/scholcomm_analytics/posts/oalex_oa_status/}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Jahn_2021}{}}%
+Jahn, N., Matthias, L., \& Laakso, M. (2021). Toward transparency of
+hybrid open access through publisher-provided metadata: An article-level
+study of elsevier. \emph{Journal of the Association for Information
+Science and Technology}, \emph{73}(1), 104--118.
+\url{https://doi.org/10.1002/asi.24549}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Larivi_re_2016}{}}%
@@
 417--435. \url{https://doi.org/10.1177/0306312716650046}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Mart_n_Mart_n_2018}{}}%
+Martín-Martín, A., Costas, R., Leeuwen, T. van, \& López-Cózar, E. D.
+(2018). Evidence of open access of scientific publications in google
+scholar: A large-scale analysis. \emph{Journal of Informetrics},
+\emph{12}(3), 819--841. \url{https://doi.org/10.1016/j.joi.2018.06.012}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Marwick_2018}{}}%
+Marwick, B., Boettiger, C., \& Mullen, L. (2018). Packaging data
+analytical work reproducibly using r (and friends). \emph{The American
+Statistician}, \emph{72}(1), 80--88.
+\url{https://doi.org/10.1080/00031305.2017.1375986}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Piwowar_2018}{}}%
+Piwowar, H., Priem, J., Larivière, V., Alperin, J. P., Matthias, L.,
+Norlander, B., Farley, A., West, J., \& Haustein, S. (2018). The state
+of {OA}: A large-scale analysis of the prevalence and impact of open
+access articles. \emph{{PeerJ}}, \emph{6}, e4375.
+\url{https://doi.org/10.7717/peerj.4375}
 

### 2024-02-12 analysis/manuscript.Rmd
refs

@@
     template: apa6.sty
+    extra_dependencies: ["flafter"]
 csl: apa.csl
@@
   dev = c("png", "pdf"),
-  fig.path = "fig/"
+  fig.path = "fig/",
+  fig.pos = "ht!", out.extra = ""
 )
@@
 
+For over two decades, hybrid open access journal publishing, which makes some articles freely available under an open license while others remain behind a paywall, has been discussed as a  model for transitioning the subscription system to full open access [@Prosser_2003].
+The idea was that when journals publish more and more open access articles, they could reduce revenues from subscriptions, while libraries and funders could change their funding models and shift expenditures from subscription to open access.
+However, initial approaches, mainly based on publication fees, also called article processing charges (APCs), did not contribute substantially to an large open access uptake. 
+In 2012 ddd - at the same time
+
+Tthe introduction of transformative agreements, a collective term for agreements starting from mid-2015, has resulted in a new push for hybrid open access.
+In particular, the Max Planck Society's Berlin 2020 initiative and plan has advocated these types of agreements as a transitional means to progress with open access.
+As a result, more than 800 transformative agreements have been tracked so far, which have been instrumental in funding activities and controversies.
+today, all publsiehrs offer such agreements
+
+Advocates of these agreements argue that they not only overcome the issue of double-dipping but also help to shift library spending from subscriptions to publications, streamline expenditures, and offer more value than big deals, which focus on large publishers.
+A few agreements can impact an institution's open access share, resulting in a significant proportion of open access content.
+increase transparency
+
+However, opponents of transformative agreements perceive continued reliance on big deals as problematic.
+They argue that the focus on large publishers perpetuates the business relationships that led to the serial crisis and lacks evidence that this leads to a reduction of In addition, concerns over inequality abound.
+could not benn mass removal
+
+Effects can be seen, mixed consequences.
+some countries who have introduced this fundign effects netherlands, sweden.
+no in germany, elsevier deal to price reduction and first steps towards mroe privacy.
+however, main goals associated woth trnasofrtiomative, the transition of the journal landscape not achivned, publsiehrs were not able to meet growtht raget, sweden did not see the flipping and has
+
+This study contributes
+
 # Methods and data
@@
 Initially, transformative agreement data retrieved from the cOAlition S Journal Checker Tool provided information about journal portfolios and participating institutions. 
-After identification of hybrid journals, Crossref served as the primary data source for article-level metadata including Creative Commons (CC) license information to determine open access availability on publisher websites.
-To identify open access articles published through transformative agreements, first author affiliations from OpenAlex were subsequently linked to eligible institutions according to the transformative agreement data.
-
-```{r data_workflow, echo=FALSE, fig.cap="Data collection workflow to obtain article-level transformative agreement data"}
+After identification of hybrid journals by excluding full open access journals, Crossref served as the primary data source for article-level metadata including Creative Commons (CC) license information to indicate open access availability on publisher websites.
+To determine open access articles published through transformative agreements, first author affiliations from OpenAlex [@priem2022openalex] were subsequently linked to eligible institutions according to the transformative agreement data.
+
+```{r data_workflow, echo=FALSE, fig.cap="Data collection workflow"}
 knitr::include_graphics(here::here("analysis", "data_collection_workflow.pdf"))
@@
 Data gathering started with obtaining journals included in transformative agreements from the publicly available Transformative Agreement Data dump[^tadump] used by the cOAlition S Journal Checker Tool.[^checkertool]
-The dump consisted of multiple online Google spreadsheets where each data file represented one agreement listed in the ESAC Transformative Agreement Registry.[^esacregistry]
+The dump consists of multiple online Google spreadsheets where each data file represents one agreement listed in the ESAC Transformative Agreement Registry.[^esacregistry]
 From the retrieved spreadsheet files, journals and institutions involved per agreement were obtained.
 
-A limitation of using the Journal Checker Tool to study the development of transformative agreements over time is that it only provided data about active transformative agreements.
-As a result, expired transformative agreements were constantly removed from the publicly available data dump. 
+A limitation of using the Journal Checker Tool and its underlying publicly available data dump to study the development of transformative agreements over time is that expired transformative agreements are constantly removed.
 To address this, four different snapshots were safeguarded and combined for this study: self-archived versions from July 2021, July 2022, and May 2023, as well as the most current dump downloaded on 11 December 2023.
-This ensured that transformative agreements, which ended from 2021 onwards, were included.
+This ensured that transformative agreements, which ended from 2021 onwards, were included, representing the majority of transformative agreements.
 <!-- Agreements ending before 2021 could not be obtained.-->
@@
 After mapping ISSN variants to the corresponding linking ISSN (ISSN-L) as provided by the ISSN International Centre, journals were associated to publishers using the ESAC ID, a unique identifier for transformative agreements in the ESAC Transformative Agreement Registry.
+Furthermore, journal subjects according to the All Science Journal Classification code (ASJC) were added from the Scopus journal source list as of August 2023.
+
 Because transformative agreements can include both full open access and hybrid journals, the data were complemented with information about a journal's open access status using multiple sources: the Directory of Open Access Journals (DOAJ) downloaded on 12 December 2023[^doaj], OpenAlex (November 2023) and the the Bielefeld list of GOLD OA journals [@goldoa].
@@
 
-After identifying hybrid journals included in transformative agreements, article metadata was retrieved from Crossref<!.
-<!-- and mapped to OpenAlex for obtaining author metadata including affiliations.
---> November 2023 database snapshot for the five-year period 2018 to 2022.
+After identifying hybrid journals included in transformative agreements, article metadata was retrieved from the Crossref <!-- and mapped to OpenAlex for obtaining author metadata including affiliations.
+--> November 2023 database snapshot for the five-year period 2018 to 2022 according to the issued date, representing the earliest known publication date.
 Because Crossref metadata lacked information to distinguish between original research articles including review and other types of journal content, which are often not covered by transformative agreements [@Borrego_2020], only articles published in regular issues indicated by non-numeric pagination were included.
@@
 Crossref was used for open access identification because transformative agreements workflows generally require publishers to deliver CC license information to this DOI registration agency [@Geschuhn_2017].
-Comparing Crossref license coverage with OpenAlex, which re-uses open access evidence from Unpaywall, not all publisher provide such license information a widely used open access discovery service that also parses journal websites for open content licenses [@Piwowar_2018], demonstrates the ongoing challenges to identify open access articles in hybrid journals [@Mart_n_Mart_n_2018; @Jahn_2021].
-Overall, `r hoaddata::cr_upw |> filter(cr_year %in% 2018:2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, cr_hybrid_total) |> pull(cr_hybrid_total) |> sum()` articles with CC license were retrieved using Crossref, while `r hoaddata::cr_upw |> filter(cr_year %in% 2018:2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, upw_hybrid_total) |> pull(upw_hybrid_total) |> sum()` articles were tagged as "hybrid" in OpenAlex.
+Comparing Crossref license coverage with OpenAlex, which re-uses open access evidence from Unpaywall, a widely used open access discovery service that also parses journal websites for open content licenses [@Piwowar_2018], highlighted ongoing challenges to identify hybrid open access [@Mart_n_Mart_n_2018; @Jahn_2021; @Zhang_2022; @Butler_2023].
+Here, `r hoaddata::cr_upw |> filter(cr_year %in% 2018:2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, cr_hybrid_total) |> pull(cr_hybrid_total) |> sum()` articles with CC license were retrieved using Crossref, while `r hoaddata::cr_upw |> filter(cr_year %in% 2018:2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, upw_hybrid_total) |> pull(upw_hybrid_total) |> sum()` articles were tagged as "hybrid" according to the OpenAlex November 2023 release, which was used throughout this study.
 The biggest differences concerned articles published between 2018 and 2020.
@@
 `r hoaddata::cr_upw |> filter(cr_year %in% 2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, upw_hybrid_total) |> pull(upw_hybrid_total) |> sum()` in OpenAlex).
-Upon inspection, notable difference could be furthermore observed among some publishers that presumably did not provide CC license information to Crossref including AIP Publishing, American Physiological Society and Emerald.
+Notable difference could be furthermore observed among some publishers that presumably did not provide CC license information to Crossref including AIP Publishing, American Physiological Society and Emerald.
 Crossref license metadata was more complete with regard to articles from the publisher Wiley and American Chemical Society.
@@
 
-Author metadata are crucial for the planning and evaluation of transformative agreements [@Schimmer_2015; @Geschuhn_2017], in particular information about corresponding authors who are generally responsible to arrange open access publication [@Borrego_2020].
-Here, affiliations were retrieved from OpenAlex.
-However, because of low coverage, this study focused on first authors and their affiliations.
-First authors typically contribute most to a paper and are often considered lead author research papers [@Larivi_re_2016]. 
+Affiliation metadata about corresponding authors are crucial for the planning and evaluation of transformative agreements, because they are considered to be responsible to arrange open access publication [@Schimmer_2015; @Geschuhn_2017; @Borrego_2020].
+Here, country and institutional affiliations were retrieved from OpenAlex.
+However, because of low coverage in OpenAlex, this study focused on first authors and their affiliations instead.
+First authors typically contribute most to a paper and are often considered lead author research papers [@Larivi_re_2016], and can be therefore assumed as a proxy to measure to open access payments [@Zhang_2022].
 Overall, around 90% of studied articles had first author affiliation metadata in OpenAlex, whereas the coverage of articles with corresponding author information was around 54%.
@@
 
-To assess the impact of transformative agreements to hybrid open access, participating institutions from the Transformative Agreement Data dump, which were crowd-sourced from the agreements and consortia that successfully negotiated an agreement, were matched with first author affiliations recorded by OpenAlex using the ROR-ID.
+To assess the impact of transformative agreements to hybrid open access,
+participating institutions from the Transformative Agreement Data dump,
+which were crowd-sourced from the agreements and consortia that
+successfully negotiated an agreement, were matched with first author
+affiliations recorded by OpenAlex using the ROR-ID.
 The matching also took into account the duration of an agreement according to the ESAC registry.
 Upon inspection, Transformative Agreement Data did not cover associated institutions comprehensively like university hospitals or institutes of large research organisations like the Max Planck Society.
-To improve the matching, Transformative Agreement Data was automatically enriched with associated institutions using OpenAlex's insitution data.
-
-In total, the compiled data set consists of `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l, cr_year, jn_all) |> pull(jn_all) |> sum()` articles published in `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l) |> nrow()` hybrid journals between 2018 and 2022 (see Figure \ref{fig:method_fig}.
+To improve the matching, Transformative Agreement Data was automatically enriched with associated organisations using OpenAlex's institution entity.
+
+In total, the compiled data set consists of `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l, cr_year, jn_all) |> pull(jn_all) |> sum()` articles published in `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l) |> nrow()` hybrid journals between 2018 and 2022 (see Figure \ref{fig:method_fig}B).
 Hybrid journals included in transformative agreements represented `r cr_stats |> filter(src == "hoad") |> pull(n_articles) |> sum() / cr_stats |> pull(n_articles) |> sum() * 100`%  of total global output over the same time period according to Crossref, while full open access journals accounted for
-`r cr_stats |> filter(src == "gold") |> pull(n_articles) |> sum() / cr_stats |> pull(n_articles) |> sum() * 100`% of total article volume.
+`r cr_stats |> filter(is.na(src) & oa_status == "gold") |> pull(n_articles) |> sum() / cr_stats |> pull(n_articles) |> sum() * 100`% of total article volume.
 
@@
 Following @Marwick_2018, hoaddata contains not only the datasets used in the data analysis.
-It also includes codeused to compile the data by connecting it the cloud-based Google Big Query data warehouse, where the big scholarly data from Crossref, OpenAlex and Unpaywall were imported.
+It also includes code used to compile the data by connecting it the cloud-based Google Big Query data warehouse, where the big scholarly data from Crossref, OpenAlex and Unpaywall were imported.
 To increase the computational reproducibility, the R package was automatically built using GitHub Actions, a continuous integration service.
@@
 
-```{r}
-#| fig-cap: "Relative growth of open access in hybrid journals through transformative agreements between 2018 and 2022 per publication year  according to the total number of articles indexed in Crossref.
-The blue areas represent open access through transformative agreements, as indicated by the affiliation with institutions that had a corresponding transformative agreement in place at the time of publication according to the first author affiliation data in OpenAlex matched with journal and institutional data from cOAlition S Transformative Agreements Public Data.
-Grey areas depict open access articles where no link to an agreement could be established.
+```{r results_overview}
+#| fig-cap: "Relative growth of open access in hybrid journals in transformative agreements between 2018 and 2022 per publication year.
+The blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established (according to matching OpenAlex first author affiliations matched with cOAlition S transformative agreement data).
 (A) Proportion of open access articles in hybrid journals per year.
@@
 Boxplots show the proportion of open access articles by individual hybrid journals (C) and individual open access uptake rates by individual hybrid journals and open access funding (D) per publication year.
-Horizontal lines correspond to the lower quartile, median, and upper quartile, with whiskers extending to ±1.5 of interquartile range.
 The individual outliers are not shown.
-Note that data on transformative agreements ending before June 2021 were not available at the time of this study."
+Note that data on transformative agreements ending before June 2021 were not available for this study."
 
@@
 
-Figure 2A shows a moderate growth in the proportion of open access articles in hybrid journals, comparing the overall open access uptake and the impact of transformative agreements on this trend.
+Figure \ref{fig:results_overview}A shows a moderate growth in the proportion of open access articles in hybrid journals, comparing the overall open access uptake and the impact of transformative agreements on this trend.
 Over the five-years period from 2018 to 2022, open access increased from `r oa_by_year |> filter(cr_year == "2018") |> pull(prop)`% (n = `r oa_by_year |> filter(cr_year == "2018") |> pull(oa)`) to `r oa_by_year |> filter(cr_year == "2022") |> pull(prop)`% (n = `r oa_by_year |> filter(cr_year == "2022") |> pull(oa)`).
@@
 
-Figure 2B highlights that the majority of open access articles in hybrid journals were made available through transformative agreements in 2021 and 2022, contributing `r oa_ta_plot_df |> filter(cr_year == "2022", type == "ta_oa") |> pull(value) / oa_by_year |> filter(cr_year == "2022") |> pull(oa) * 100`% of the total open access article volume in 2022.
+Figure \ref{fig:results_overview}B highlights that the majority of open access articles in hybrid journals were made available through transformative agreements in 2021 and 2022, contributing `r oa_ta_plot_df |> filter(cr_year == "2022", type == "ta_oa") |> pull(value) / oa_by_year |> filter(cr_year == "2022") |> pull(oa) * 100`% of the total open access article volume in 2022.
 However, there was also a notable growth in open access provision through individual publication fees, which increased from `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> pull(value)`) in 2018 to  `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> pull(value)`).
@@
 
-Figure 2C depicts the substantial variations among the hybrid journals included in transformative agreements in terms of open access uptake.
-Although the median generally follows the trend shown in Figure 2A, the farther stretch of upper quartiles and whiskers over the years illustrates that an increasing number of journals published an above-average proportion of open access articles.
+Figure \ref{fig:results_overview}C depicts the substantial variations among the hybrid journals included in transformative agreements in terms of open access uptake.
+Although the median generally follows the trend shown in Figure \ref{fig:results_overview}A, the farther stretch of upper quartiles and whiskers over the years illustrates that an increasing number of journals published an above-average proportion of open access articles.
 In 2022, 25% of hybrid journals (n = `r oa_jn_box_df |> filter(cr_year == "2022") |> mutate(top25 = quantile(prop, .75)) |> filter(prop >= top25) |> nrow()`) had an open access uptake of `r oa_jn_box_df |> filter(cr_year == "2022") |> pull(prop) |> quantile(0.75)*100`%, and `r (oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> nrow()) / length(unique(oa_jn_box_df$issn_l)) *100`% of journals (n = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> nrow()`) provided the majority of their articles under a Creative Commons license in the same year.
 These journals were, on average, smaller (M = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> pull(n) |> mean()`, SD = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> pull(n) |> sd()`) than those with an open access share below 50% (M = `r oa_jn_box_df |> filter(cr_year == "2022", prop <= 0.5) |> pull(n) |> mean()`, SD = `r oa_jn_box_df |> filter(cr_year == "2022", prop <= 0.5) |> pull(n) |> sd()`).
-Notable exception of large journals with an above-average open access proportion were Physical Review D, a high-energy physics journal covered by the SCOAP3 consortium that provided open access to `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "2470-0010") |> pull(oa_all)` out of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "2470-0010") |> pull(n)` articles in 2022, Astronomy and Physics (`r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0004-6361") |> pull(oa_all)` out of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0004-6361") |> pull(n)` articles in 2022 were open access), which shifted to a subscribe to open business model for all accepted articles as of April 2022, the Journal of Fluid Mechanics (`r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0022-1120") |> pull(oa_all)` of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0022-1120") |> pull(n)` articles in 2022 were open access) and Bioinformatics (`r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "1367-4803") |> pull(oa_all)` out of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "1367-4803") |> pull(n)` articles in 2022 were open access), which flipped to full open access as of January 2023.
-
-When comparing the impact of open access trough transformative agreements across journals, it shows that for many journals these agreements substantially contributed to the growth of open access over the years (Figure 2D).
+Notable exception of large journals with an above-average open access proportion were Physical Review D, a high-energy physics journal covered by the SCOAP3 consortium that provided open access to `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "2470-0010") |> pull(oa_all)` out of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "2470-0010") |> pull(n)` articles in 2022, Astronomy and Physics (`r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0004-6361") |> pull(oa_all)` out of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0004-6361") |> pull(n)` articles in 2022 were open access), which shifted to a subscribe to open business model for all accepted articles as of April 2022, and the Journal of Fluid Mechanics (`r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0022-1120") |> pull(oa_all)` of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0022-1120") |> pull(n)` articles in 2022 were open access).
+
+When comparing the impact of open access trough transformative agreements across journals, it shows that for many journals these agreements substantially contributed to the growth of open access over the years (Figure \ref{fig:results_overview}D).
 Examples of such journals include those with a scope on specific countries or regions, where also transformative agreements were implemented. 
@@
 
-```{r}
+```{r publisher_league_table}
 publisher_league_tbl |>
-  mutate(across(contains(c("perc", "cum")), ~ paste(round(., 3) * 100))) |>
+  mutate(across(contains(c("perc")), ~ paste(round(., 3) * 100))) |>
+  select(-contains("cum")) |>
   kbl(
@@
       "%",
-      "C%",
       "Total",
       "%",
-      "C%",
       "Total",
       "%",
-      "C%",
       "Total",
-      "%",
-      "C%"
+      "%"
     )
   ) |>
-  kable_styling(font_size = 10, position = "center") |>
+  kable_styling(position = "center") |>
   add_header_above(c(
     " " = 1,
-    "Hybrid journals" = 3,
-    "Articles" = 3,
-    "OA articles" = 3,
-    "TA OA articles" = 3
+    "Hybrid journals" = 2,
+    "Articles" = 2,
+    "OA articles" = 2,
+    "TA OA articles" = 2
   ))
@@
 Analysing hybrid open access across publishers between 2018 and 2022 reveals a large market concentration. 
-Although `r length(unique(jn_ind_jct_all$esac_publisher))` publishers offered transformative agreements, the big three commercial publishers Elsevier, Springer Nature, and Wiley, accounted for `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(jn_cum) * 100`% of total article volume published (see Table ).
+Although `r length(unique(jn_ind_jct_all$esac_publisher))` publishers offered transformative agreements, the big three commercial publishers Elsevier, Springer Nature, and Wiley, accounted for `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(jn_cum) * 100`% of total article volume published (see Table \ref{tab:publisher_league_table}).
 Together, they published `r publisher_league_tbl  |> filter(publisher != "Other") |> pull(oa_total) |> sum()` open access articles, representing `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(oa_cum) * 100`% of the open access articles in hybrid journals.
@@
 
-```{r, fig.asp=0.8}
+```{r publisher_figure, fig.asp=0.8, fig.cap="Developement of open access in hybrid journals included in transformative agreements between 2018 and 2022 by publishers.
+The blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established (according to matching OpenAlex first author affiliations matched with cOAlition S transformative agreement data).
+(A) Proportion of open access articles in hybrid journals per year and publisher.
+(B) Percentage of hybrid open access via agreements per year and publisher.
+Boxplots (C) show individual open access uptake rates by individual hybrid journals and open access funding per publication year and publisher.
+The individual outliers are not shown.
+Note that data on transformative agreements ending before June 2021 were not available for this study."}
 library(patchwork)
@@
 
-Figure 2 takes a closer look into the growth of hybrid open access across publishers by year with a focus on open articles enabled by transformative agreements. 
-Although all publishers show a general long-term trend towards transformative agreements, Figure 2A and 2B indicate that, in particular, Wiley's has experienced a substantial increase in its open access share from `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(value) |> sum()`) in 2018 to `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(prop) |> sum()*100`%  (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()`) in 2022, rperesenting an 4.5-fold increase.
-In contrast, Elsevier's hybrid journals deonstrated a more modest increase, from `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Elsevier") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Elsevier") |> pull(value) |> sum()`) in 2018 to `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(value) |> sum()`) in 2022, which is a relatively low open access share compared to the general trend.
+Figure \ref{fig:publisher_figure} takes a closer look into the growth of hybrid open access across publishers by year with a focus on open articles enabled by transformative agreements. 
+Although all publishers show a general long-term trend towards transformative agreements, Figure \ref{fig:publisher_figure}A and B indicate that, in particular, Wiley's has experienced a substantial increase in its open access share from `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(value) |> sum()`) in 2018 to `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(prop) |> sum()*100`%  (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()`) in 2022, representing an 4.5-fold increase.
+In contrast, Elsevier's hybrid journals demonstrated a more modest increase, from `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Elsevier") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Elsevier") |> pull(value) |> sum()`) in 2018 to `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(value) |> sum()`) in 2022, which is a relatively low open access share compared to the general trend.
 In 2018, Springer Nature had the largest open access proportion among the three publishers of `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Springer Nature") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Springer Nature") |> pull(value) |> sum()`), but experienced a relatively slower growth, resulting in `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature") |> pull(value) |> sum()`) of articles being open access in Springer Nature hybrid journals in 2022.
@@
 
-The increasing trend towards transformative agreements can be also observed at the journal-level (Figure 2).
+The increasing trend towards transformative agreements can be also observed at the journal-level (Figure \ref{fig:publisher_figure}C).
 While no substantial differences between open access enabled through transformative agreements and other revenue source could observed across Elsevier journals, the distribution of open access across Springer Nature and Wiley hybrid journals indicates that the growth is not limited to a few journals, but extends across the portfolio.
@@
 
-Table presents a high-level overview of hybrid open access by AJCS subject area using fractionalised counting to account for journals belonging to more than one category. 
+Table \ref{tab:subject_summary_table} presents a high-level overview of hybrid open access by AJCS subject area using fractionalised counting to account for journals belonging to more than one category.
 Between 2018 and 2022, most hybrid journals with at least one open articles could be attributed to the social sciences including the humanities.
@@
   mutate(across(contains(c("perc", "cum")), ~ paste(round(., 3) * 100))) |>
+  select(-contains("cum")) |>
   kbl(
@@
       "%",
-      "C%",
       "Total",
       "%",
-      "C%",
       "Total",
       "%",
-      "C%",
       "Total",
-      "%",
-      "C%"
+      "%"
     )
@@
     " " = 1,
-    "Hybrid journals" = 3,
-    "Articles" = 3,
-    "OA articles" = 3,
-    "TA OA articles" = 3
+    "Hybrid journals" = 2,
+    "Articles" = 2,
+    "OA articles" = 2,
+    "TA OA articles" = 2
   ))
@@
 
-Figure presents the relative growth of hybrid open access by subject area between 2018-2022.
+Figure \ref{fig:subject_panel} presents the relative growth of hybrid open access by subject area between 2018-2022.
 In particular, Social Sciences and Humanties journals accounted for the strongest growth in the five-years period from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Social Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Social Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(value) |> sum()`), followed by the Life Science from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Life Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Life Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Life Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Life Sciences") |> pull(value) |> sum()`) , Health Science from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Health Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Health Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Health Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Health Sciences") |> pull(value) |> sum()`) and Physical Sciences from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Physical Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Physical Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Physical Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Physical Sciences") |> pull(value) |> sum()`).
@@
 This growth in the social sciences can be largely attributed to transformative agreements.
-In 2022, two-third of open access articles (`r subject_ta_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(prop) * 100`%, n = `r subject_ta_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(ta_oa_frac)`) were published by first authors affiliated with participating institutions.
-Figure C shows that this trend is consistent across Social Sciences journals.
+In 2022, two-third of open access articles (`r subject_ta_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(prop) * 100`%, n = `r subject_ta_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(ta_oa_frac)`) were published by first authors affiliated with participating institutions (see \ref{fig:subject_panel}B).
+Figure \ref{fig:subject_panel}C shows that this trend is consistent across Social Sciences journals.
 In 2022, 25% of Social Science journals provided open access to at least every fourth article exclusively through transformative agreements.
@@
 
-```{r country_patch, fig.asp=0.9}
+```{r countrypatch, fig.asp=0.9}
 library(patchwork)
@@
   BC",
- # guides = "collect",
+ # guides = "c  ollect",
   widths = c(0.7, 0.3),
@@
 
-Figure 2A shows the development of hybrid open access publishing by countries, comparing the OECD area with the BRICS, an intergovernmental organisation,  which comprised the countries Brazil, Russia, India, China and South Africa in 2022.
+Figure \ref{fig:countrypatch}A shows the development of hybrid open access publishing by countries, comparing the OECD area with the BRICS, an intergovernmental organisation, which comprised the countries Brazil, Russia, India, China and South Africa in 2022.
 The residual category "Other" includes the remaining countries. 
@@
 
-To illustrate the situation in 2022, Figure B compares total publication output with the number of open access articles. 
+To illustrate the situation in 2022, \ref{fig:countrypatch}B compares total publication output with the number of open access articles. 
 With `r country_scatter_df |> filter(cr_year == "2022", country_code == "CN") |> pull(all)` articles, China was the most productive country, followed by the United States (`r country_scatter_df |> filter(cr_year == "2022", country_code == "US") |> pull(all)` articles) and India (`r country_scatter_df |> filter(cr_year == "2022", country_code == "IN") |> pull(all)` articles).
@@
 Particularly, Germany, Great Britain, the Netherlands, Sweden, Switzerland and Spain recorded an above-average open access share as indicated by the linear trend line.
-As represented by the point size, as well as it can been seen in Figure C, transformative agreements contributed to this market position of these countries. 
+As represented by the point size, as well as it can been seen in Figure \ref{fig:country_patch}C, transformative agreements contributed to this market position of these countries. 
 Interestingly, the United States had a notable open access market share of `r top_20_plot$data |> distinct(country_code, oa, all) |> mutate(prop = oa / sum(oa)) |> filter(country_code == "US") |> pull(prop) * 100`%, although transformative agreements contributed to a lesser extent.
@@
 
-```{r country_top_20_plot, fig.asp=0.8}
+```{r countrytop20plot, fig.asp=0.8}
 country_plot
@@
 
-Figure illustrates the development of hybrid open access from 2018 to 2022, highlighting the top 20 most productive countries in terms of articles published in hybrid journals that were included in transformative agreements over the five-year period.
+Figure \ref{fig:countrytop20plot} illustrates the development of hybrid open access from 2018 to 2022, highlighting the top 20 most productive countries in terms of articles published in hybrid journals that were included in transformative agreements over the five-year period.
 Notably, The Netherlands (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "NL") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), Sweden (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "SE") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), Poland (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "PL") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%) and Great Britain (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "GB") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%)) exhibited a relatively high level of uptake in 2018 which continued to increase in the following years.
@@
 Since 2021, there has been a general trend towards hybrid open access among many western countries, primarily driven by transformative agreements.
-However, proliferation of transformative agreements differe across these countries.
+However, proliferation of transformative agreements differed across these countries.
 For instance, Germany successfully negotiated an agreement with Elsevier not until 2023. 
@@
 
-### Supplements
+# Supplements
 

### 2024-02-12 analysis/manuscript.tex
refs

@@
 
+For over two decades, hybrid open access journal publishing, which makes
+some articles freely available under an open license while others remain
+behind a paywall, has been discussed as a model for transitioning the
+subscription system to full open access (Prosser, 2003). The idea was
+that when journals publish more and more open access articles, they
+could reduce revenues from subscriptions, while libraries and funders
+could change their funding models and shift expenditures from
+subscription to open access. However, initial approaches, mainly based
+on publication fees, also called article processing charges (APCs), did
+not contribute substantially to an large open access uptake. In 2012 ddd
+- at the same time
+
+Tthe introduction of transformative agreements, a collective term for
+agreements starting from mid-2015, has resulted in a new push for hybrid
+open access. In particular, the Max Planck Society's Berlin 2020
+initiative and plan has advocated these types of agreements as a
+transitional means to progress with open access. As a result, more than
+800 transformative agreements have been tracked so far, which have been
+instrumental in funding activities and controversies. today, all
+publsiehrs offer such agreements
+
+Advocates of these agreements argue that they not only overcome the
+issue of double-dipping but also help to shift library spending from
+subscriptions to publications, streamline expenditures, and offer more
+value than big deals, which focus on large publishers. A few agreements
+can impact an institution's open access share, resulting in a
+significant proportion of open access content. increase transparency
+
+However, opponents of transformative agreements perceive continued
+reliance on big deals as problematic. They argue that the focus on large
+publishers perpetuates the business relationships that led to the serial
+crisis and lacks evidence that this leads to a reduction of In addition,
+concerns over inequality abound. could not benn mass removal
+
+Effects can be seen, mixed consequences. some countries who have
+introduced this fundign effects netherlands, sweden. no in germany,
+elsevier deal to price reduction and first steps towards mroe privacy.
+however, main goals associated woth trnasofrtiomative, the transition of
+the journal landscape not achivned, publsiehrs were not able to meet
+growtht raget, sweden did not see the flipping and has
+
+This study contributes
+
 \hypertarget{methods-and-data}{%
@@
 Checker Tool provided information about journal portfolios and
-participating institutions. After identification of hybrid journals,
-Crossref served as the primary data source for article-level metadata
-including Creative Commons (CC) license information to determine open
-access availability on publisher websites. To identify open access
-articles published through transformative agreements, first author
-affiliations from OpenAlex were subsequently linked to eligible
-institutions according to the transformative agreement data.
+participating institutions. After identification of hybrid journals by
+excluding full open access journals, Crossref served as the primary data
+source for article-level metadata including Creative Commons (CC)
+license information to indicate open access availability on publisher
+websites. To determine open access articles published through
+transformative agreements, first author affiliations from OpenAlex
+(Priem et al., 2022) were subsequently linked to eligible institutions
+according to the transformative agreement data.
 
-\begin{figure}
+\begin{figure}[ht!]
 
-{\centering \includegraphics[width=0.99\linewidth]{data_collection_workflow}
+{\centering \includegraphics[width=0.99\linewidth,]{data_collection_workflow}
 
@@
 
-\caption{Data collection workflow to obtain article-level transformative agreement data}\label{fig:data_workflow}
+\caption{Data collection workflow}\label{fig:data_workflow}
 \end{figure}
@@
 used by the cOAlition S Journal Checker Tool.\footnote{\url{https://www.coalition-s.org/blog/enabling-accurate-results-within-the-journal-checker-tool/}}
-The dump consisted of multiple online Google spreadsheets where each
-data file represented one agreement listed in the ESAC Transformative
+The dump consists of multiple online Google spreadsheets where each data
+file represents one agreement listed in the ESAC Transformative
 Agreement Registry.\footnote{\url{https://esac-initiative.org/about/transformative-agreements/agreement-registry/}}
@@
 
-A limitation of using the Journal Checker Tool to study the development
-of transformative agreements over time is that it only provided data
-about active transformative agreements. As a result, expired
-transformative agreements were constantly removed from the publicly
-available data dump. To address this, four different snapshots were
+A limitation of using the Journal Checker Tool and its underlying
+publicly available data dump to study the development of transformative
+agreements over time is that expired transformative agreements are
+constantly removed. To address this, four different snapshots were
 safeguarded and combined for this study: self-archived versions from
@@
 downloaded on 11 December 2023. This ensured that transformative
-agreements, which ended from 2021 onwards, were included. Overall, the
-four combined Transformative Agreement Data dumps used in this study
-contained 729 out of 869 agreements listed in the ESAC registry by
-December 2023.
+agreements, which ended from 2021 onwards, were included, representing
+the majority of transformative agreements. Overall, the four combined
+Transformative Agreement Data dumps used in this study contained 729 out
+of 869 agreements listed in the ESAC registry by December 2023.
 
@@
 ESAC ID, a unique identifier for transformative agreements in the ESAC
-Transformative Agreement Registry. Because transformative agreements can
-include both full open access and hybrid journals, the data were
-complemented with information about a journal's open access status using
-multiple sources: the Directory of Open Access Journals (DOAJ)
-downloaded on 12 December 2023\footnote{\url{https://doaj.org/csv}},
+Transformative Agreement Registry. Furthermore, journal subjects
+according to the All Science Journal Classification code (ASJC) were
+added from the Scopus journal source list as of August 2023.
+
+Because transformative agreements can include both full open access and
+hybrid journals, the data were complemented with information about a
+journal's open access status using multiple sources: the Directory of
+Open Access Journals (DOAJ) downloaded on 12 December 2023\footnote{\url{https://doaj.org/csv}},
 OpenAlex (November 2023) and the the Bielefeld list of GOLD OA journals
@@
 After identifying hybrid journals included in transformative agreements,
-article metadata was retrieved from Crossref\textless!. November 2023
-database snapshot for the five-year period 2018 to 2022. Because
-Crossref metadata lacked information to distinguish between original
-research articles including review and other types of journal content,
-which are often not covered by transformative agreements (Borrego et
-al., 2020), only articles published in regular issues indicated by
-non-numeric pagination were included. Furthermore, an expanded version
-of Unpaywall's paratext recognition approach was applied to exclude
+article metadata was retrieved from the Crossref November 2023 database
+snapshot for the five-year period 2018 to 2022 according to the issued
+date, representing the earliest known publication date. Because Crossref
+metadata lacked information to distinguish between original research
+articles including review and other types of journal content, which are
+often not covered by transformative agreements (Borrego et al., 2020),
+only articles published in regular issues indicated by non-numeric
+pagination were included. Furthermore, an expanded version of
+Unpaywall's paratext recognition approach was applied to exclude
 non-scholarly journal content such as table of contents.
@@
 \& Stone, 2017). Comparing Crossref license coverage with OpenAlex,
-which re-uses open access evidence from Unpaywall, not all publisher
-provide such license information a widely used open access discovery
-service that also parses journal websites for open content licenses
-(Piwowar et al., 2018), demonstrates the ongoing challenges to identify
-open access articles in hybrid journals (Jahn et al., 2021;
-Martín-Martín et al., 2018). Overall, 742,369 articles with CC license
-were retrieved using Crossref, while 950,260 articles were tagged as
-``hybrid'' in OpenAlex. The biggest differences concerned articles
-published between 2018 and 2020. In 2022, however, Crossref and OpenAlex
-open access numbers only differ slightly (249,511 records using Crossref
-vs.~255,344 in OpenAlex). Upon inspection, notable difference could be
-furthermore observed among some publishers that presumably did not
-provide CC license information to Crossref including AIP Publishing,
+which re-uses open access evidence from Unpaywall, a widely used open
+access discovery service that also parses journal websites for open
+content licenses (Piwowar et al., 2018), highlighted ongoing challenges
+to identify hybrid open access (Butler et al., 2023; Jahn et al., 2021;
+Martín-Martín et al., 2018; Zhang et al., 2022). Here, 742,369 articles
+with CC license were retrieved using Crossref, while 950,260 articles
+were tagged as ``hybrid'' according to the OpenAlex November 2023
+release, which was used throughout this study. The biggest differences
+concerned articles published between 2018 and 2020. In 2022, however,
+Crossref and OpenAlex open access numbers only differ slightly (249,511
+records using Crossref vs.~255,344 in OpenAlex). Notable difference
+could be furthermore observed among some publishers that presumably did
+not provide CC license information to Crossref including AIP Publishing,
 American Physiological Society and Emerald. Crossref license metadata
@@
 
-Author metadata are crucial for the planning and evaluation of
-transformative agreements (Geschuhn \& Stone, 2017; Schimmer et al.,
-2015), in particular information about corresponding authors who are
-generally responsible to arrange open access publication (Borrego et
-al., 2020). Here, affiliations were retrieved from OpenAlex. However,
-because of low coverage, this study focused on first authors and their
-affiliations. First authors typically contribute most to a paper and are
-often considered lead author research papers (Larivière et al., 2016).
-Overall, around 90\% of studied articles had first author affiliation
-metadata in OpenAlex, whereas the coverage of articles with
-corresponding author information was around 54\%.
+Affiliation metadata about corresponding authors are crucial for the
+planning and evaluation of transformative agreements, because they are
+considered to be responsible to arrange open access publication (Borrego
+et al., 2020; Geschuhn \& Stone, 2017; Schimmer et al., 2015). Here,
+country and institutional affiliations were retrieved from OpenAlex.
+However, because of low coverage in OpenAlex, this study focused on
+first authors and their affiliations instead. First authors typically
+contribute most to a paper and are often considered lead author research
+papers (Larivière et al., 2016), and can be therefore assumed as a proxy
+to measure to open access payments (Zhang et al., 2022). Overall, around
+90\% of studied articles had first author affiliation metadata in
+OpenAlex, whereas the coverage of articles with corresponding author
+information was around 54\%.
 
@@
 To improve the matching, Transformative Agreement Data was automatically
-enriched with associated institutions using OpenAlex's insitution data.
+enriched with associated organisations using OpenAlex's institution
+entity.
 
@@
 in 12,857 hybrid journals between 2018 and 2022 (see Figure
-\ref{fig:method_fig}. Hybrid journals included in transformative
+\ref{fig:method_fig}B). Hybrid journals included in transformative
 agreements represented 40\% of total global output over the same time
 period according to Crossref, while full open access journals accounted
-for 0\% of total article volume.
+for 35\% of total article volume.
 
-\begin{figure}
+\begin{figure}[ht!]
 
-{\centering \includegraphics[width=0.99\linewidth]{fig/method_fig-1}
+{\centering \includegraphics[width=0.99\linewidth,]{fig/method_fig-1}
 
@@
 (2018), hoaddata contains not only the datasets used in the data
-analysis. It also includes codeused to compile the data by connecting it
-the cloud-based Google Big Query data warehouse, where the big scholarly
-data from Crossref, OpenAlex and Unpaywall were imported. To increase
-the computational reproducibility, the R package was automatically built
-using GitHub Actions, a continuous integration service.
+analysis. It also includes code used to compile the data by connecting
+it the cloud-based Google Big Query data warehouse, where the big
+scholarly data from Crossref, OpenAlex and Unpaywall were imported. To
+increase the computational reproducibility, the R package was
+automatically built using GitHub Actions, a continuous integration
+service.
 
@@
 
-\begin{figure}
+\begin{figure}[ht!]
 
-{\centering \includegraphics[width=0.99\linewidth]{fig/unnamed-chunk-2-1}
+{\centering \includegraphics[width=0.99\linewidth,]{fig/results_overview-1}
 
@@
 
-\caption{Relative growth of open access in hybrid journals through transformative agreements between 2018 and 2022 per publication year  according to the total number of articles indexed in Crossref.
-The blue areas represent open access through transformative agreements, as indicated by the affiliation with institutions that had a corresponding transformative agreement in place at the time of publication according to the first author affiliation data in OpenAlex matched with journal and institutional data from cOAlition S Transformative Agreements Public Data.
-Grey areas depict open access articles where no link to an agreement could be established.
+\caption{Relative growth of open access in hybrid journals in transformative agreements between 2018 and 2022 per publication year.
+The blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established (according to matching OpenAlex first author affiliations matched with cOAlition S transformative agreement data).
 (A) Proportion of open access articles in hybrid journals per year.
@@
 Boxplots show the proportion of open access articles by individual hybrid journals (C) and individual open access uptake rates by individual hybrid journals and open access funding (D) per publication year.
-Horizontal lines correspond to the lower quartile, median, and upper quartile, with whiskers extending to ±1.5 of interquartile range.
 The individual outliers are not shown.
-Note that data on transformative agreements ending before June 2021 were not available at the time of this study.}\label{fig:unnamed-chunk-2}
+Note that data on transformative agreements ending before June 2021 were not available for this study.}\label{fig:results_overview}
 \end{figure}
 
-Figure 2A shows a moderate growth in the proportion of open access
-articles in hybrid journals, comparing the overall open access uptake
-and the impact of transformative agreements on this trend. Over the
-five-years period from 2018 to 2022, open access increased from 4.3\% (n
-= 65,486) to 15\% (n = 249,511). At the same time, the total article
-volume of the investigated journals grew from 1,528,051 in 2018 to
-1,676,928 in 2022.
+Figure \ref{fig:results_overview}A shows a moderate growth in the
+proportion of open access articles in hybrid journals, comparing the
+overall open access uptake and the impact of transformative agreements
+on this trend. Over the five-years period from 2018 to 2022, open access
+increased from 4.3\% (n = 65,486) to 15\% (n = 249,511). At the same
+time, the total article volume of the investigated journals grew from
+1,528,051 in 2018 to 1,676,928 in 2022.
 
-Figure 2B highlights that the majority of open access articles in hybrid
-journals were made available through transformative agreements in 2021
-and 2022, contributing 58\% of the total open access article volume in
-2022. However, there was also a notable growth in open access provision
-through individual publication fees, which increased from 4.1\% (n =
-62,625) in 2018 to 6.3\% (n = 105,896). This suggests that publishers
-were able to gain equally from individual and institutional open access
-publishing options.
+Figure \ref{fig:results_overview}B highlights that the majority of open
+access articles in hybrid journals were made available through
+transformative agreements in 2021 and 2022, contributing 58\% of the
+total open access article volume in 2022. However, there was also a
+notable growth in open access provision through individual publication
+fees, which increased from 4.1\% (n = 62,625) in 2018 to 6.3\% (n =
+105,896). This suggests that publishers were able to gain equally from
+individual and institutional open access publishing options.
 
-Figure 2C depicts the substantial variations among the hybrid journals
-included in transformative agreements in terms of open access uptake.
-Although the median generally follows the trend shown in Figure 2A, the
-farther stretch of upper quartiles and whiskers over the years
-illustrates that an increasing number of journals published an
-above-average proportion of open access articles. In 2022, 25\% of
-hybrid journals (n = 2,576) had an open access uptake of 29\%, and 6.6\%
-of journals (n = 744) provided the majority of their articles under a
-Creative Commons license in the same year. These journals were, on
-average, smaller (M = 75, SD = 186) than those with an open access share
-below 50\% (M = 164, SD = 347). Notable exception of large journals with
-an above-average open access proportion were Physical Review D, a
-high-energy physics journal covered by the SCOAP3 consortium that
-provided open access to 2,341 out of 4,074 articles in 2022, Astronomy
-and Physics (1,396 out of 2,230 articles in 2022 were open access),
-which shifted to a subscribe to open business model for all accepted
-articles as of April 2022, the Journal of Fluid Mechanics (577 of 1,077
-articles in 2022 were open access) and Bioinformatics ( out of articles
-in 2022 were open access), which flipped to full open access as of
-January 2023.
+Figure \ref{fig:results_overview}C depicts the substantial variations
+among the hybrid journals included in transformative agreements in terms
+of open access uptake. Although the median generally follows the trend
+shown in Figure \ref{fig:results_overview}A, the farther stretch of
+upper quartiles and whiskers over the years illustrates that an
+increasing number of journals published an above-average proportion of
+open access articles. In 2022, 25\% of hybrid journals (n = 2,576) had
+an open access uptake of 29\%, and 6.6\% of journals (n = 744) provided
+the majority of their articles under a Creative Commons license in the
+same year. These journals were, on average, smaller (M = 75, SD = 186)
+than those with an open access share below 50\% (M = 164, SD = 347).
+Notable exception of large journals with an above-average open access
+proportion were Physical Review D, a high-energy physics journal covered
+by the SCOAP3 consortium that provided open access to 2,341 out of 4,074
+articles in 2022, Astronomy and Physics (1,396 out of 2,230 articles in
+2022 were open access), which shifted to a subscribe to open business
+model for all accepted articles as of April 2022, and the Journal of
+Fluid Mechanics (577 of 1,077 articles in 2022 were open access).
 
@@
 agreements substantially contributed to the growth of open access over
-the years (Figure 2D). Examples of such journals include those with a
-scope on specific countries or regions, where also transformative
-agreements were implemented. For instance, in 2022, the Germany-based
-journals Zeitschrift für Erziehungswissenschaft and Zeitschrift für
-Politikwissenschaft, as well as the Scandinavian Political Studies
-adressing the Nordic countries, achieved an overall open access uptake
-of more than 90\% just through transformative agreements. Despite the
-rise in transformative agreements, it is worth noting that other means
-of publishing open access in hybrid journals remained common. In total,
-9,153 journals published open access articles from authors affiliated
-with institutions without transformative agreements in place, while
-8,780 journals published at least one open access article through a
-transformative agreement in the same year.
+the years (Figure \ref{fig:results_overview}D). Examples of such
+journals include those with a scope on specific countries or regions,
+where also transformative agreements were implemented. For instance, in
+2022, the Germany-based journals Zeitschrift für Erziehungswissenschaft
+and Zeitschrift für Politikwissenschaft, as well as the Scandinavian
+Political Studies adressing the Nordic countries, achieved an overall
+open access uptake of more than 90\% just through transformative
+agreements. Despite the rise in transformative agreements, it is worth
+noting that other means of publishing open access in hybrid journals
+remained common. In total, 9,153 journals published open access articles
+from authors affiliated with institutions without transformative
+agreements in place, while 8,780 journals published at least one open
+access article through a transformative agreement in the same year.
 
@@
 
-\caption{\label{tab:unnamed-chunk-5}Hybrid open access through transformative agreements market shares 2018-2022}
+\caption{\label{tab:publisher_league_table}Hybrid open access through transformative agreements market shares 2018-2022}
 \centering
-\fontsize{10}{12}\selectfont
-\begin{tabular}[t]{lrllrllrllrll}
+\begin{tabular}[t]{lrlrlrlrl}
 \toprule
-\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Hybrid journals} & \multicolumn{3}{c}{Articles} & \multicolumn{3}{c}{OA articles} & \multicolumn{3}{c}{TA OA articles} \\
-\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7} \cmidrule(l{3pt}r{3pt}){8-10} \cmidrule(l{3pt}r{3pt}){11-13}
-Publisher & Total & \% & C\% & Total & \% & C\% & Total & \% & C\% & Total & \% & C\%\\
+\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Hybrid journals} & \multicolumn{2}{c}{Articles} & \multicolumn{2}{c}{OA articles} & \multicolumn{2}{c}{TA OA articles} \\
+\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9}
+Publisher & Total & \% & Total & \% & Total & \% & Total & \%\\
 \midrule
-Elsevier & 1,936 & 17 & 17 & 2,770,826 & 33.8 & 33.8 & 172,723 & 22.9 & 22.9 & 60,440 & 18.3 & 18.3\\
-Springer Nature & 2,274 & 20 & 37 & 1,330,430 & 16.2 & 50 & 175,432 & 23.3 & 46.2 & 100,008 & 30.3 & 48.6\\
-Wiley & 1,410 & 12.4 & 49.4 & 1,043,052 & 12.7 & 62.7 & 152,723 & 20.3 & 66.5 & 83,443 & 25.3 & 73.9\\
-Other & 5,767 & 50.6 & 100 & 3,061,337 & 37.3 & 100 & 252,523 & 33.5 & 100 & 86,294 & 26.1 & 100\\
+Elsevier & 1,936 & 17 & 2,770,826 & 33.8 & 172,723 & 22.9 & 60,440 & 18.3\\
+Springer Nature & 2,274 & 20 & 1,330,430 & 16.2 & 175,432 & 23.3 & 100,008 & 30.3\\
+Wiley & 1,410 & 12.4 & 1,043,052 & 12.7 & 152,723 & 20.3 & 83,443 & 25.3\\
+Other & 5,767 & 50.6 & 3,061,337 & 37.3 & 252,523 & 33.5 & 86,294 & 26.1\\
 \bottomrule
@@
 Springer Nature, and Wiley, accounted for 49\% of total article volume
-published (see Table ). Together, they published 500,878 open access
-articles, representing 66\% of the open access articles in hybrid
-journals. Elsevier, Springer Nature, and Wiley made 243,891 articles
-open access in hybrid journals through transformative agreements,
-resulting in an even larger market share of 74\%. However, there are
-differences among the three large publishers. Although Elsevier
-published the largest volume of articles (n = 2,770,826, 34\%), it
-published a relatively low number of open access articles, including
-those that can be associated with transformative agreements. In
-contrast, Springer Nature and Wiley provided open access to a larger
-proportion of their articles (13\% of Springer Nature articles and 15\%
-of Wiley articles were open acccess), leading to higher open access
-market shares (23\% Springer Nature resp. 23\% Wiley). This difference
-between Elsevier on the one hand and Springer Nature and Wiley on the
-other can be attributed to transformative agreements, as the latter made
-the majority of their open access articles available through such deals
-(Springer Nature 57\% resp. Wiley 55\%).
+published (see Table \ref{tab:publisher_league_table}). Together, they
+published 500,878 open access articles, representing 66\% of the open
+access articles in hybrid journals. Elsevier, Springer Nature, and Wiley
+made 243,891 articles open access in hybrid journals through
+transformative agreements, resulting in an even larger market share of
+74\%. However, there are differences among the three large publishers.
+Although Elsevier published the largest volume of articles (n =
+2,770,826, 34\%), it published a relatively low number of open access
+articles, including those that can be associated with transformative
+agreements. In contrast, Springer Nature and Wiley provided open access
+to a larger proportion of their articles (13\% of Springer Nature
+articles and 15\% of Wiley articles were open acccess), leading to
+higher open access market shares (23\% Springer Nature resp. 23\%
+Wiley). This difference between Elsevier on the one hand and Springer
+Nature and Wiley on the other can be attributed to transformative
+agreements, as the latter made the majority of their open access
+articles available through such deals (Springer Nature 57\% resp. Wiley
+55\%).
 
-\begin{center}\includegraphics[width=0.99\linewidth]{fig/unnamed-chunk-6-1} \end{center}
+\begin{figure}[ht!]
 
-Figure 2 takes a closer look into the growth of hybrid open access
-across publishers by year with a focus on open articles enabled by
-transformative agreements. Although all publishers show a general
-long-term trend towards transformative agreements, Figure 2A and 2B
-indicate that, in particular, Wiley's has experienced a substantial
-increase in its open access share from 5.9\% (n = 11,628) in 2018 to
-26\% (n = 53,503) in 2022, rperesenting an 4.5-fold increase. In
-contrast, Elsevier's hybrid journals deonstrated a more modest increase,
-from 3.3\% (n = 16,872) in 2018 to 10\% (n = 60,821) in 2022, which is a
-relatively low open access share compared to the general trend. In 2018,
-Springer Nature had the largest open access proportion among the three
-publishers of 8.4\% (n = 19,701), but experienced a relatively slower
-growth, resulting in 18\% (n = 52,616) of articles being open access in
-Springer Nature hybrid journals in 2022.
+{\centering \includegraphics[width=0.99\linewidth,]{fig/publisher_figure-1}
+
+}
+
+\caption{Developement of open access in hybrid journals included in transformative agreements between 2018 and 2022 by publishers.
+The blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established (according to matching OpenAlex first author affiliations matched with cOAlition S transformative agreement data).
+(A) Proportion of open access articles in hybrid journals per year and publisher.
+(B) Percentage of hybrid open access via agreements per year and publisher.
+Boxplots (C) show individual open access uptake rates by individual hybrid journals and open access funding per publication year and publisher.
+The individual outliers are not shown.
+Note that data on transformative agreements ending before June 2021 were not available for this study.}\label{fig:publisher_figure}
+\end{figure}
+
+Figure \ref{fig:publisher_figure} takes a closer look into the growth of
+hybrid open access across publishers by year with a focus on open
+articles enabled by transformative agreements. Although all publishers
+show a general long-term trend towards transformative agreements, Figure
+\ref{fig:publisher_figure}A and B indicate that, in particular, Wiley's
+has experienced a substantial increase in its open access share from
+5.9\% (n = 11,628) in 2018 to 26\% (n = 53,503) in 2022, representing an
+4.5-fold increase. In contrast, Elsevier's hybrid journals demonstrated
+a more modest increase, from 3.3\% (n = 16,872) in 2018 to 10\% (n =
+60,821) in 2022, which is a relatively low open access share compared to
+the general trend. In 2018, Springer Nature had the largest open access
+proportion among the three publishers of 8.4\% (n = 19,701), but
+experienced a relatively slower growth, resulting in 18\% (n = 52,616)
+of articles being open access in Springer Nature hybrid journals in
+2022.
 
@@
 The increasing trend towards transformative agreements can be also
-observed at the journal-level (Figure 2). While no substantial
-differences between open access enabled through transformative
-agreements and other revenue source could observed across Elsevier
-journals, the distribution of open access across Springer Nature and
-Wiley hybrid journals indicates that the growth is not limited to a few
-journals, but extends across the portfolio. In particular, Wiley's upper
-quantile, which represents the top 25\% of journals in terms of the
-proportion of open access articles from transformative agreements,
-increased markedly from 13\% in 2020 to 31\% in 2022. At the same time,
-the median proportion grew from 7.5\% to 19\%. It is interesting to note
-that a small but increasing number of journals from these two publishers
-are providing open access to the majority of articles through
-transformative agreements. Wiley recorded 68 and Springer Nature 102
-hybrid journals with an open access share above 50\% that could be
-solely attributed to transformative agreements. Upon inspection, these
-journals were mainly society or local language journals with a small
-yearly article volume.
+observed at the journal-level (Figure \ref{fig:publisher_figure}C).
+While no substantial differences between open access enabled through
+transformative agreements and other revenue source could observed across
+Elsevier journals, the distribution of open access across Springer
+Nature and Wiley hybrid journals indicates that the growth is not
+limited to a few journals, but extends across the portfolio. In
+particular, Wiley's upper quantile, which represents the top 25\% of
+journals in terms of the proportion of open access articles from
+transformative agreements, increased markedly from 13\% in 2020 to 31\%
+in 2022. At the same time, the median proportion grew from 7.5\% to
+19\%. It is interesting to note that a small but increasing number of
+journals from these two publishers are providing open access to the
+majority of articles through transformative agreements. Wiley recorded
+68 and Springer Nature 102 hybrid journals with an open access share
+above 50\% that could be solely attributed to transformative agreements.
+Upon inspection, these journals were mainly society or local language
+journals with a small yearly article volume.
 
@@
 
-Table presents a high-level overview of hybrid open access by AJCS
-subject area using fractionalised counting to account for journals
-belonging to more than one category. Between 2018 and 2022, most hybrid
-journals with at least one open articles could be attributed to the
-social sciences including the humanities. However, these journals
-published the fewest number of articles, whereas physical sciences
-journals recorded most articles, followed by the health sciences and the
-life sciences. In terms of open access, physical sciences journals
-accounted for more than one third of articles published in the
-five-years period, followed by the health science, the social sciences
-and the life sciences.
+Table \ref{tab:subject_summary_table} presents a high-level overview of
+hybrid open access by AJCS subject area using fractionalised counting to
+account for journals belonging to more than one category. Between 2018
+and 2022, most hybrid journals with at least one open articles could be
+attributed to the social sciences including the humanities. However,
+these journals published the fewest number of articles, whereas physical
+sciences journals recorded most articles, followed by the health
+sciences and the life sciences. In terms of open access, physical
+sciences journals accounted for more than one third of articles
+published in the five-years period, followed by the health science, the
+social sciences and the life sciences.
 
@@
 \centering
-\begin{tabular}[t]{lrllrllrllrll}
+\begin{tabular}[t]{lrlrlrlrl}
 \toprule
-\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Hybrid journals} & \multicolumn{3}{c}{Articles} & \multicolumn{3}{c}{OA articles} & \multicolumn{3}{c}{TA OA articles} \\
-\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7} \cmidrule(l{3pt}r{3pt}){8-10} \cmidrule(l{3pt}r{3pt}){11-13}
-Journal subject & Total & \% & C\% & Total & \% & C\% & Total & \% & C\% & Total & \% & C\%\\
+\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Hybrid journals} & \multicolumn{2}{c}{Articles} & \multicolumn{2}{c}{OA articles} & \multicolumn{2}{c}{TA OA articles} \\
+\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9}
+Journal subject & Total & \% & Total & \% & Total & \% & Total & \%\\
 \midrule
-Health Sciences & 2,376 & 22.5 & 22.5 & 2,709,906 & 27.8 & 27.8 & 286,592 & 27.3 & 27.3 & 117,746 & 25 & 25\\
-Life Sciences & 1,403 & 13.3 & 35.8 & 1,477,808 & 15.1 & 42.9 & 191,880 & 18.3 & 45.6 & 71,593 & 15.2 & 40.2\\
-Physical Sciences & 2,732 & 25.9 & 61.7 & 4,291,833 & 44 & 86.9 & 366,794 & 35 & 80.6 & 167,686 & 35.6 & 75.8\\
-Social Sciences & 4,050 & 38.3 & 100 & 1,280,460 & 13.1 & 100 & 203,461 & 19.4 & 100 & 114,190 & 24.2 & 100\\
+Health Sciences & 2,376 & 22.5 & 2,709,906 & 27.8 & 286,592 & 27.3 & 117,746 & 25\\
+Life Sciences & 1,403 & 13.3 & 1,477,808 & 15.1 & 191,880 & 18.3 & 71,593 & 15.2\\
+Physical Sciences & 2,732 & 25.9 & 4,291,833 & 44 & 366,794 & 35 & 167,686 & 35.6\\
+Social Sciences & 4,050 & 38.3 & 1,280,460 & 13.1 & 203,461 & 19.4 & 114,190 & 24.2\\
 \bottomrule
@@
 
-Figure presents the relative growth of hybrid open access by subject
-area between 2018-2022. In particular, Social Sciences and Humanties
-journals accounted for the strongest growth in the five-years period
-from 6.4\% (n = 8,361) to 23\% (n = 51,938), followed by the Life
-Science from 7.6\% (n = 15,003) to 18\% (n = 39,494) , Health Science
-from 5.3\% (n = 18,279) to 16\% (n = 63,089) and Physical Sciences from
-4.5\% (n = 22,364) to 12\% (n = 85,428). This growth in the social
-sciences can be largely attributed to transformative agreements. In
-2022, two-third of open access articles (67\%, n = 34,759) were
-published by first authors affiliated with participating institutions.
-Figure C shows that this trend is consistent across Social Sciences
-journals. In 2022, 25\% of Social Science journals provided open access
-to at least every fourth article exclusively through transformative
-agreements. However, hybrid open access through transformative
-agreements played a comparable lesser role in the Life Sciences and
-Health Sciences. In these two subject areas, only about half of the open
-access articles can be linked to these agreements, both overall and on
-median average across journals. In contrast, the majority of Physical
-Science Journals, shows an increase of open access through
+Figure \ref{fig:subject_panel} presents the relative growth of hybrid
+open access by subject area between 2018-2022. In particular, Social
+Sciences and Humanties journals accounted for the strongest growth in
+the five-years period from 6.4\% (n = 8,361) to 23\% (n = 51,938),
+followed by the Life Science from 7.6\% (n = 15,003) to 18\% (n =
+39,494) , Health Science from 5.3\% (n = 18,279) to 16\% (n = 63,089)
+and Physical Sciences from 4.5\% (n = 22,364) to 12\% (n = 85,428). This
+growth in the social sciences can be largely attributed to
+transformative agreements. In 2022, two-third of open access articles
+(67\%, n = 34,759) were published by first authors affiliated with
+participating institutions (see \ref{fig:subject_panel}B). Figure
+\ref{fig:subject_panel}C shows that this trend is consistent across
+Social Sciences journals. In 2022, 25\% of Social Science journals
+provided open access to at least every fourth article exclusively
+through transformative agreements. However, hybrid open access through
+transformative agreements played a comparable lesser role in the Life
+Sciences and Health Sciences. In these two subject areas, only about
+half of the open access articles can be linked to these agreements, both
+overall and on median average across journals. In contrast, the majority
+of Physical Science Journals, shows an increase of open access through
 transformative agreements compared to other options to publish open
@@
 
-\begin{center}\includegraphics[width=0.99\linewidth]{fig/subject_panel-1} \end{center}
+\begin{center}\includegraphics[width=0.99\linewidth,]{fig/subject_panel-1} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=0.99\linewidth]{fig/country_patch-1} \end{center}
+\begin{center}\includegraphics[width=0.99\linewidth,]{fig/countrypatch-1} \end{center}
 
-Figure 2A shows the development of hybrid open access publishing by
-countries, comparing the OECD area with the BRICS, an intergovernmental
-organisation, which comprised the countries Brazil, Russia, India, China
-and South Africa in 2022. The residual category ``Other'' includes the
-remaining countries. From 2018 to 2022, the proportion of open access in
-hybrid journals increased from 6.1\% in 2018 to 26\% in 2022. On the
-other hand, BRICS recorded an low uptake, from 1.6\% in 2018 to 3.7\% in
-2022.\\
+Figure \ref{fig:countrypatch}A shows the development of hybrid open
+access publishing by countries, comparing the OECD area with the BRICS,
+an intergovernmental organisation, which comprised the countries Brazil,
+Russia, India, China and South Africa in 2022. The residual category
+``Other'' includes the remaining countries. From 2018 to 2022, the
+proportion of open access in hybrid journals increased from 6.1\% in
+2018 to 26\% in 2022. On the other hand, BRICS recorded an low uptake,
+from 1.6\% in 2018 to 3.7\% in 2022.\\
 Despite rise of open access across OECD countries, the overall
@@
 
-To illustrate the situation in 2022, Figure B compares total publication
-output with the number of open access articles. With 391,530 articles,
-China was the most productive country, followed by the United States
-(268,965 articles) and India (87,428 articles). In contrast, West and
-Nord European countries published a considerable high number of open
-access articles, mainly due to transformative agreements. Particularly,
-Germany, Great Britain, the Netherlands, Sweden, Switzerland and Spain
-recorded an above-average open access share as indicated by the linear
-trend line. As represented by the point size, as well as it can been
-seen in Figure C, transformative agreements contributed to this market
-position of these countries. Interestingly, the United States had a
-notable open access market share of 15\%, although transformative
-agreements contributed to a lesser extent. Similarly, China's open
-access market share of 7.2 in 2022 was comparable to that of the
-Netherlands, which was (7.1\%,).
+To illustrate the situation in 2022, \ref{fig:countrypatch}B compares
+total publication output with the number of open access articles. With
+391,530 articles, China was the most productive country, followed by the
+United States (268,965 articles) and India (87,428 articles). In
+contrast, West and Nord European countries published a considerable high
+number of open access articles, mainly due to transformative agreements.
+Particularly, Germany, Great Britain, the Netherlands, Sweden,
+Switzerland and Spain recorded an above-average open access share as
+indicated by the linear trend line. As represented by the point size, as
+well as it can been seen in Figure \ref{fig:country_patch}C,
+transformative agreements contributed to this market position of these
+countries. Interestingly, the United States had a notable open access
+market share of 15\%, although transformative agreements contributed to
+a lesser extent. Similarly, China's open access market share of 7.2 in
+2022 was comparable to that of the Netherlands, which was (7.1\%,).
 
-\begin{center}\includegraphics[width=0.99\linewidth]{fig/country_top_20_plot-1} \end{center}
+\begin{center}\includegraphics[width=0.99\linewidth,]{fig/countrytop20plot-1} \end{center}
 
-Figure illustrates the development of hybrid open access from 2018 to
-2022, highlighting the top 20 most productive countries in terms of
-articles published in hybrid journals that were included in
-transformative agreements over the five-year period. Notably, The
-Netherlands (27\%), Sweden (24\%), Poland (17\%) and Great Britain
-(17\%)) exhibited a relatively high level of uptake in 2018 which
-continued to increase in the following years. In 2022, Sweden had the
-highest proportion of open-access articles relative to its publication
-output (78\%), followed by the Netherlands (67\%) and Switzerland
-(57\%), with these countries benefiting from transformative agreements.
-In Germany, however, hybrid open access only began to increase from 2019
-onwards after the successful negotiation of nationwide agreements with
-Wiley (July 2019) and Springer Nature (January 2020). Prior to this,
-only a few organisations had agreements in place, for the example the
-Max Planck Society with Springer Compact.
+Figure \ref{fig:countrytop20plot} illustrates the development of hybrid
+open access from 2018 to 2022, highlighting the top 20 most productive
+countries in terms of articles published in hybrid journals that were
+included in transformative agreements over the five-year period.
+Notably, The Netherlands (27\%), Sweden (24\%), Poland (17\%) and Great
+Britain (17\%)) exhibited a relatively high level of uptake in 2018
+which continued to increase in the following years. In 2022, Sweden had
+the highest proportion of open-access articles relative to its
+publication output (78\%), followed by the Netherlands (67\%) and
+Switzerland (57\%), with these countries benefiting from transformative
+agreements. In Germany, however, hybrid open access only began to
+increase from 2019 onwards after the successful negotiation of
+nationwide agreements with Wiley (July 2019) and Springer Nature
+(January 2020). Prior to this, only a few organisations had agreements
+in place, for the example the Max Planck Society with Springer Compact.
 
@@
 among many western countries, primarily driven by transformative
-agreements. However, proliferation of transformative agreements differe
+agreements. However, proliferation of transformative agreements differed
 across these countries. For instance, Germany successfully negotiated an
@@
 \hypertarget{supplements}{%
-\subsubsection{Supplements}\label{supplements}}
+\section{Supplements}\label{supplements}}
 
-\begin{center}\includegraphics[width=0.99\linewidth]{fig/unnamed-chunk-10-1} \end{center}
+\begin{center}\includegraphics[width=0.99\linewidth,]{fig/unnamed-chunk-7-1} \end{center}
 
@@
 University. \url{https://doi.org/10.4119/unibi/2961544}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Butler_2023}{}}%
+Butler, L.-A., Matthias, L., Simard, M.-A., Mongeon, P., \& Haustein, S.
+(2023). The oligopoly's shift to open access: How the big five academic
+publishers profit from article processing charges. \emph{Quantitative
+Science Studies}, 1--22. \url{https://doi.org/10.1162/qss_a_00272}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-priem2022openalex}{}}%
+Priem, J., Piwowar, H., \& Orr, R. (2022). \emph{OpenAlex: A fully-open
+index of scholarly works, authors, venues, institutions, and concepts}.
+\url{https://arxiv.org/abs/2205.01833}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Prosser_2003}{}}%
+Prosser, D. C. (2003). From here to there: A proposed mechanism for
+transforming journals from closed to open access. \emph{Learned
+Publishing}, \emph{16}(3), 163--166.
+\url{https://doi.org/10.1087/095315103322110923}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Schimmer_2015}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Zhang_2022}{}}%
+Zhang, L., Wei, Y., Huang, Y., \& Sivertsen, G. (2022). Should open
+access lead to closed research? The trends towards paying to perform
+research. \emph{Scientometrics}, \emph{127}(12), 7653--7679.
+\url{https://doi.org/10.1007/s11192-022-04407-5}
+
 \end{CSLReferences}

### 2024-02-13 analysis/manuscript.Rmd
continue introduction

@@
 ```
+
 # Introduction
 
-For over two decades, hybrid open access journal publishing, which makes some articles freely available under an open license while others remain behind a paywall, has been discussed as a  model for transitioning the subscription system to full open access [@Prosser_2003].
+For over two decades, hybrid open access journal publishing, which makes some articles openly available while others remain behind a paywall, has been discussed as a  model for transitioning the subscription system to full open access [@Prosser_2003].
 The idea was that when journals publish more and more open access articles, they could reduce revenues from subscriptions, while libraries and funders could change their funding models and shift expenditures from subscription to open access.
 However, initial approaches, mainly based on publication fees, also called article processing charges (APCs), did not contribute substantially to an large open access uptake. 
-In 2012 ddd - at the same time
-
-Tthe introduction of transformative agreements, a collective term for agreements starting from mid-2015, has resulted in a new push for hybrid open access.
-In particular, the Max Planck Society's Berlin 2020 initiative and plan has advocated these types of agreements as a transitional means to progress with open access.
-As a result, more than 800 transformative agreements have been tracked so far, which have been instrumental in funding activities and controversies.
-today, all publsiehrs offer such agreements
-
-Advocates of these agreements argue that they not only overcome the issue of double-dipping but also help to shift library spending from subscriptions to publications, streamline expenditures, and offer more value than big deals, which focus on large publishers.
+In 2009, Springer reported open access to 1% of articles in hybrid journals [@Dallmeier_2010].
+Other publishers also recorded a low uptake.
+In 2011, only 1-2% of articles in over 4,300 hybrid journals from large publishers were open access [@Bj_rk_2012], growing to around 4% between 2011 and 2013 [@Laakso_2016].
+
+With the introduction of central funding mechanisms for hybrid open access publication fees in a few countries since 2012, notably in the UK, an increase in open access could be observed [@Bj_rk_2017; @Jubb_2017; @Robinson_Garcia_2020].
+However, this shift in funding policy towards hybrid open access also added to the overall cost of publishing, which includes subscriptions spending and the administrative efforts required to handle payments [@Pinfield_2016].
+Moreover, established large commercial publishers, which already dominated the publishing market [@Larivi_re_2015], disproportionately benefited from hybrid open access funding in comparison to full open access publishers [@Jahn_2016; @Butler_2023; @Shu_2023].
+
+As a consequence, libraries and their consortia began to develop licensing strategies aimed at avoiding such ‘double dipping’ scenarios, in which well-established commercial publishers gain twice from reading and publication fees, as well as to increase publisher-provided immediate open access [@BJ_RK_2014; @Schimmer_2015].
+These considerations resulted into transformative agreements, which cover a broad range of successfully negotiated contracts between library consortia and publishers from mid-2010s onwards where institutional spending for subscriptions are re-purposed for open access publishing [@Borrego_2020].
+These transformative agreements seek to control costs while allowing a transitional phase for publishing more open access articles [@Hinchliffe_2019]. 
+At the same time, they aim at a higher degree of transparency than previous subscription-only Big Deals, which bundled access to large journal portfolios of commercial publishers [@Bergstrom_2014].
+However, the focus of transformative agreements remains primarily on these publishers with large hybrid journals and subscription-only portfolios [@Borrego_2020]. 
+<!-- national negotiations, oa2020 -->
+For example, the nationwide DEAL consortium in Germany, which includes universities and academic institutions, collectively cancelled their subscriptions with Elsevier from 2017 onwards to negotiate a transformative agreement [@Fraser_2023].
+
+The introduction of transformative agreements aligns with funding policy changes, such as  the decision made by the  cOAlition S, a consortium of national funders including the European Commission, to no longer provide financial support for individual publication fees when publishing in hybrid journals.
+ According ot its Plan S, they only accept hybrid open access through transformative agreements "during a transition period that should be as short as possible" [@Schiltz_2018]. 
+Specifically, cOAlition S funders agreed to support hybrid open access solely through transformative agreements from 1 January 2021, until the end of 2024.
+The German Research Foundation (DFG) has also extended its financial support for hybrid open access through transformative agreements, despite not being part of cOAlition S.
+Previously, the DFG only provided funding for full open access journals [@Jahn_2016].
+
+Advocates of these agreements argue that they not only overcome the issue standardiazation of double-dipping but also help to shift library spending from subscriptions to publications, streamline expenditures, and offer more value than big deals, which focus on large publishers.
 A few agreements can impact an institution's open access share, resulting in a significant proportion of open access content.
@@
 They argue that the focus on large publishers perpetuates the business relationships that led to the serial crisis and lacks evidence that this leads to a reduction of In addition, concerns over inequality abound.
+trnasparency differ and now nsns
+
 could not benn mass removal
@@
 
-```{r countrypatch, fig.asp=0.9}
+```{r country_patch, fig.asp=0.9, fig.cap="Development of hybrid open access publishing by countries"}
 library(patchwork)
@@
   BC",
- # guides = "c  ollect",
   widths = c(0.7, 0.3),
@@
 
-Figure \ref{fig:countrypatch}A shows the development of hybrid open access publishing by countries, comparing the OECD area with the BRICS, an intergovernmental organisation, which comprised the countries Brazil, Russia, India, China and South Africa in 2022.
+Figure \ref{fig:country_patch}A shows the development of hybrid open access publishing by countries, comparing the OECD area with the BRICS, an intergovernmental organisation, which comprised the countries Brazil, Russia, India, China and South Africa in 2022.
 The residual category "Other" includes the remaining countries. 
@@
 
-To illustrate the situation in 2022, \ref{fig:countrypatch}B compares total publication output with the number of open access articles. 
+To illustrate the situation in 2022, \ref{fig:country_patch}B compares total publication output with the number of open access articles. 
 With `r country_scatter_df |> filter(cr_year == "2022", country_code == "CN") |> pull(all)` articles, China was the most productive country, followed by the United States (`r country_scatter_df |> filter(cr_year == "2022", country_code == "US") |> pull(all)` articles) and India (`r country_scatter_df |> filter(cr_year == "2022", country_code == "IN") |> pull(all)` articles).
@@
 
-```{r countrytop20plot, fig.asp=0.8}
+```{r country_top_20_plot, fig.asp=0.8, fig.cap="Top 20 countries"}
 country_plot
@@
 
-Figure \ref{fig:countrytop20plot} illustrates the development of hybrid open access from 2018 to 2022, highlighting the top 20 most productive countries in terms of articles published in hybrid journals that were included in transformative agreements over the five-year period.
+Figure \ref{fig:country_top_20_plot} illustrates the development of hybrid open access from 2018 to 2022, highlighting the top 20 most productive countries in terms of articles published in hybrid journals that were included in transformative agreements over the five-year period.
 Notably, The Netherlands (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "NL") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), Sweden (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "SE") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), Poland (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "PL") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%) and Great Britain (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "GB") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%)) exhibited a relatively high level of uptake in 2018 which continued to increase in the following years.

### 2024-02-13 analysis/manuscript.tex
continue introduction

@@
 For over two decades, hybrid open access journal publishing, which makes
-some articles freely available under an open license while others remain
-behind a paywall, has been discussed as a model for transitioning the
-subscription system to full open access (Prosser, 2003). The idea was
-that when journals publish more and more open access articles, they
-could reduce revenues from subscriptions, while libraries and funders
-could change their funding models and shift expenditures from
-subscription to open access. However, initial approaches, mainly based
-on publication fees, also called article processing charges (APCs), did
-not contribute substantially to an large open access uptake. In 2012 ddd
-- at the same time
+some articles openly available while others remain behind a paywall, has
+been discussed as a model for transitioning the subscription system to
+full open access (Prosser, 2003). The idea was that when journals
+publish more and more open access articles, they could reduce revenues
+from subscriptions, while libraries and funders could change their
+funding models and shift expenditures from subscription to open access.
+However, initial approaches, mainly based on publication fees, also
+called article processing charges (APCs), did not contribute
+substantially to an large open access uptake. In 2009, Springer reported
+open access to 1\% of articles in hybrid journals (Dallmeier-Tiessen et
+al., 2010). Other publishers also recorded a low uptake. In 2011, only
+1-2\% of articles in over 4,300 hybrid journals from large publishers
+were open access (Björk, 2012), growing to around 4\% between 2011 and
+2013 (Laakso \& Björk, 2016).
 
-Tthe introduction of transformative agreements, a collective term for
-agreements starting from mid-2015, has resulted in a new push for hybrid
-open access. In particular, the Max Planck Society's Berlin 2020
-initiative and plan has advocated these types of agreements as a
-transitional means to progress with open access. As a result, more than
-800 transformative agreements have been tracked so far, which have been
-instrumental in funding activities and controversies. today, all
-publsiehrs offer such agreements
+With the introduction of central funding mechanisms for hybrid open
+access publication fees in a few countries since 2012, notably in the
+UK, an increase in open access could be observed (Björk, 2017; Jubb et
+al., 2017; Robinson-Garcia et al., 2020). However, this shift in funding
+policy towards hybrid open access also added to the overall cost of
+publishing, which includes subscriptions spending and the administrative
+efforts required to handle payments (Pinfield et al., 2016). Moreover,
+established large commercial publishers, which already dominated the
+publishing market (Larivière et al., 2015), disproportionately benefited
+from hybrid open access funding in comparison to full open access
+publishers (Butler et al., 2023; Jahn \& Tullney, 2016; Shu \&
+Larivière, 2023).
+
+As a consequence, libraries and their consortia began to develop
+licensing strategies aimed at avoiding such `double dipping' scenarios,
+in which well-established commercial publishers gain twice from reading
+and publication fees, as well as to increase publisher-provided
+immediate open access (BJÖRK \& SOLOMON, 2014; Schimmer et al., 2015).
+These considerations resulted into transformative agreements, which
+cover a broad range of successfully negotiated contracts between library
+consortia and publishers from mid-2010s onwards where institutional
+spending for subscriptions are re-purposed for open access publishing
+(Borrego et al., 2020). These transformative agreements seek to control
+costs while allowing a transitional phase for publishing more open
+access articles (Hinchliffe, 2019). At the same time, they aim at a
+higher degree of transparency than previous subscription-only Big Deals,
+which bundled access to large journal portfolios of commercial
+publishers (Bergstrom et al., 2014). However, the focus of
+transformative agreements remains primarily on these publishers with
+large hybrid journals and subscription-only portfolios (Borrego et al.,
+2020). For example, the nationwide DEAL consortium in Germany, which
+includes universities and academic institutions, collectively cancelled
+their subscriptions with Elsevier from 2017 onwards to negotiate a
+transformative agreement (Fraser et al., 2023).
+
+The introduction of transformative agreements aligns with funding policy
+changes, such as the decision made by the cOAlition S, a consortium of
+national funders including the European Commission, to no longer provide
+financial support for individual publication fees when publishing in
+hybrid journals. According ot its Plan S, they only accept hybrid open
+access through transformative agreements ``during a transition period
+that should be as short as possible'' (Schiltz, 2018). Specifically,
+cOAlition S funders agreed to support hybrid open access solely through
+transformative agreements from 1 January 2021, until the end of 2024.
+The German Research Foundation (DFG) has also extended its financial
+support for hybrid open access through transformative agreements,
+despite not being part of cOAlition S. Previously, the DFG only provided
+funding for full open access journals (Jahn \& Tullney, 2016).
 
 Advocates of these agreements argue that they not only overcome the
-issue of double-dipping but also help to shift library spending from
-subscriptions to publications, streamline expenditures, and offer more
-value than big deals, which focus on large publishers. A few agreements
-can impact an institution's open access share, resulting in a
-significant proportion of open access content. increase transparency
+issue standardiazation of double-dipping but also help to shift library
+spending from subscriptions to publications, streamline expenditures,
+and offer more value than big deals, which focus on large publishers. A
+few agreements can impact an institution's open access share, resulting
+in a significant proportion of open access content. increase
+transparency
 
@@
 crisis and lacks evidence that this leads to a reduction of In addition,
-concerns over inequality abound. could not benn mass removal
+concerns over inequality abound. trnasparency differ and now nsns
+
+could not benn mass removal
 
@@
 
-\begin{center}\includegraphics[width=0.99\linewidth,]{fig/countrypatch-1} \end{center}
+\begin{figure}[ht!]
 
-Figure \ref{fig:countrypatch}A shows the development of hybrid open
+{\centering \includegraphics[width=0.99\linewidth,]{fig/country_patch-1}
+
+}
+
+\caption{Development of hybrid open access publishing by countries}\label{fig:country_patch}
+\end{figure}
+
+Figure \ref{fig:country_patch}A shows the development of hybrid open
 access publishing by countries, comparing the OECD area with the BRICS,
@@
 
-To illustrate the situation in 2022, \ref{fig:countrypatch}B compares
+To illustrate the situation in 2022, \ref{fig:country_patch}B compares
 total publication output with the number of open access articles. With
@@
 
-\begin{center}\includegraphics[width=0.99\linewidth,]{fig/countrytop20plot-1} \end{center}
+\begin{figure}[ht!]
 
-Figure \ref{fig:countrytop20plot} illustrates the development of hybrid
-open access from 2018 to 2022, highlighting the top 20 most productive
-countries in terms of articles published in hybrid journals that were
-included in transformative agreements over the five-year period.
-Notably, The Netherlands (27\%), Sweden (24\%), Poland (17\%) and Great
-Britain (17\%)) exhibited a relatively high level of uptake in 2018
-which continued to increase in the following years. In 2022, Sweden had
-the highest proportion of open-access articles relative to its
+{\centering \includegraphics[width=0.99\linewidth,]{fig/country_top_20_plot-1}
+
+}
+
+\caption{Top 20 countries}\label{fig:country_top_20_plot}
+\end{figure}
+
+Figure \ref{fig:country_top_20_plot} illustrates the development of
+hybrid open access from 2018 to 2022, highlighting the top 20 most
+productive countries in terms of articles published in hybrid journals
+that were included in transformative agreements over the five-year
+period. Notably, The Netherlands (27\%), Sweden (24\%), Poland (17\%)
+and Great Britain (17\%)) exhibited a relatively high level of uptake in
+2018 which continued to increase in the following years. In 2022, Sweden
+had the highest proportion of open-access articles relative to its
 publication output (78\%), followed by the Netherlands (67\%) and
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-Bergstrom_2014}{}}%
+Bergstrom, T. C., Courant, P. N., McAfee, R. P., \& Williams, M. A.
+(2014). Evaluating big deal journal bundles. \emph{Proceedings of the
+National Academy of Sciences}, \emph{111}(26), 9425--9430.
+\url{https://doi.org/10.1073/pnas.1403006111}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Bj_rk_2012}{}}%
+Björk, B.-C. (2012). The hybrid model for open access publication of
+scholarly articles: A failed experiment? \emph{Journal of the American
+Society for Information Science and Technology}, \emph{63}(8),
+1496--1504. \url{https://doi.org/10.1002/asi.22709}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Bj_rk_2017}{}}%
+Björk, B.-C. (2017). Growth of hybrid open access, 2009--2016.
+\emph{PeerJ}, \emph{5}, e3878. \url{https://doi.org/10.7717/peerj.3878}
+
+\leavevmode\vadjust pre{\hypertarget{ref-BJ_RK_2014}{}}%
+BJÖRK, B., \& SOLOMON, D. (2014). How research funders can finance APCs
+in full OA and hybrid journals. \emph{Learned Publishing}, \emph{27}(2),
+93--103. \url{https://doi.org/10.1087/20140203}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Borrego_2020}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Dallmeier_2010}{}}%
+Dallmeier-Tiessen, S., Goerner, B., Darby, R., Hyppoelae, J.,
+Igo-Kemenes, P., Kahn, D., Lambert, S., Lengenfelder, A., Leonard, C.,
+Mele, S., Polydoratou, P., Ross, D., Ruiz-Perez, S., Schimmer, R.,
+Swaisland, M., \& Stelt, W. van der. (2010). \emph{{Open Access
+Publishing - Models and Attributes}} (T. S. consortium, Ed.).
+
+\leavevmode\vadjust pre{\hypertarget{ref-Fraser_2023}{}}%
+Fraser, N., Hobert, A., Jahn, N., Mayr, P., \& Peters, I. (2023). No
+deal: German researchers' publishing and citing behaviors after big deal
+negotiations with elsevier. \emph{Quantitative Science Studies},
+\emph{4}(2), 325--352. \url{https://doi.org/10.1162/qss_a_00255}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Geschuhn_2017}{}}%
@@
 \url{https://doi.org/10.1629/uksg.391}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Hinchliffe_2019}{}}%
+Hinchliffe, L. J. (2019). \emph{Transformative agreements: A primer}.
+\url{https://web.archive.org/web/20210128170342/https://scholarlykitchen.sspnet.org/2019/04/23/transformative-agreements/};
+The Scholarly Kitchen.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Jahn_2016}{}}%
+Jahn, N., \& Tullney, M. (2016). A study of institutional spending on
+open access publication fees in germany. \emph{{PeerJ}}, \emph{4},
+e2323. \url{https://doi.org/10.7717/peerj.2323}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Jubb_2017}{}}%
+Jubb, M., Plume, A., Oeben, S., Brammer, L., Johnson, R., Bütün, C., \&
+Pinfield, S. (2017). \emph{Monitoring the transition to open access:
+December 2017}.
+\url{https://web.archive.org/web/20200212015524/https://www.universitiesuk.ac.uk/policy-and-analysis/reports/Documents/2017/monitoring-transition-open-access-2017.pdf}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Laakso_2016}{}}%
+Laakso, M., \& Björk, B.-C. (2016). Hybrid open access--a longitudinal
+study. \emph{Journal of Informetrics}, \emph{10}(4), 919--932.
+\url{https://doi.org/10.1016/j.joi.2016.08.002}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Larivi_re_2016}{}}%
@@
 417--435. \url{https://doi.org/10.1177/0306312716650046}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Larivi_re_2015}{}}%
+Larivière, V., Haustein, S., \& Mongeon, P. (2015). The oligopoly of
+academic publishers in the digital era. \emph{{PLOS} {ONE}},
+\emph{10}(6), e0127502.
+\url{https://doi.org/10.1371/journal.pone.0127502}
 
@@
 \url{https://doi.org/10.1080/00031305.2017.1375986}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Pinfield_2016}{}}%
+Pinfield, S., Salter, J., \& Bath, P. A. (2016). The "total cost of
+publication" in a hybrid open-access environment: Institutional
+approaches to funding journal article-processing charges in combination
+with subscriptions. \emph{Journal of the Association for Information
+Science and Technology}, \emph{67}(7), 1751--1766.
+\url{https://doi.org/10.1002/asi.23446}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Robinson_Garcia_2020}{}}%
+Robinson-Garcia, N., Costas, R., \& Leeuwen, T. N. van. (2020). Open
+access uptake by universities worldwide. \emph{{PeerJ}}, \emph{8},
+e9410. \url{https://doi.org/10.7717/peerj.9410}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Schiltz_2018}{}}%
+Schiltz, M. (2018). Science without publication paywalls: cOAlition s
+for the realisation of full and immediate open access. \emph{PLOS
+Biology}, \emph{16}(9), e3000031.
+\url{https://doi.org/10.1371/journal.pbio.3000031}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Schimmer_2015}{}}%
@@
 \url{https://doi.org/10.17617/1.3}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Shu_2023}{}}%
+Shu, F., \& Larivière, V. (2023). The oligopoly of open access
+publishing. \emph{Scientometrics}, \emph{129}(1), 519--536.
+\url{https://doi.org/10.1007/s11192-023-04876-2}
 

### 2024-02-14 analysis/manuscript.Rmd
continue introduction ii

@@
 For over two decades, hybrid open access journal publishing, which makes some articles openly available while others remain behind a paywall, has been discussed as a  model for transitioning the subscription system to full open access [@Prosser_2003].
-The idea was that when journals publish more and more open access articles, they could reduce revenues from subscriptions, while libraries and funders could change their funding models and shift expenditures from subscription to open access.
+The idea was that when journals increasingly publish open access articles, they could reduce revenues from subscriptions, while libraries and funders could change their funding models and shift expenditures from subscription to open access.
 However, initial approaches, mainly based on publication fees, also called article processing charges (APCs), did not contribute substantially to an large open access uptake. 
@@
 
-With the introduction of central funding mechanisms for hybrid open access publication fees in a few countries since 2012, notably in the UK, an increase in open access could be observed [@Bj_rk_2017; @Jubb_2017; @Robinson_Garcia_2020].
-However, this shift in funding policy towards hybrid open access also added to the overall cost of publishing, which includes subscriptions spending and the administrative efforts required to handle payments [@Pinfield_2016].
+With the introduction of central funding mechanisms for hybrid open access publication fees in some European countries since 2012 an increase in open access could be observed [@Bj_rk_2017; @Piwowar_2018; @Jubb_2017; @Huang_2020].
+For example, studying university output, @Robinson_Garcia_2020 estimated a median uptake of 7.1% in the period 2014-2017. 
+In particular, British (17%), Austrian (15%) and Dutch (13%) universities contributed to this trend. 
+However, this shift in funding policy towards hybrid open access also added to the overall cost of publishing, which includes subscription spending and the administrative efforts required to handle payments [@Pinfield_2016].
 Moreover, established large commercial publishers, which already dominated the publishing market [@Larivi_re_2015], disproportionately benefited from hybrid open access funding in comparison to full open access publishers [@Jahn_2016; @Butler_2023; @Shu_2023].
@@
 As a consequence, libraries and their consortia began to develop licensing strategies aimed at avoiding such ‘double dipping’ scenarios, in which well-established commercial publishers gain twice from reading and publication fees, as well as to increase publisher-provided immediate open access [@BJ_RK_2014; @Schimmer_2015].
-These considerations resulted into transformative agreements, which cover a broad range of successfully negotiated contracts between library consortia and publishers from mid-2010s onwards where institutional spending for subscriptions are re-purposed for open access publishing [@Borrego_2020].
-These transformative agreements seek to control costs while allowing a transitional phase for publishing more open access articles [@Hinchliffe_2019]. 
-At the same time, they aim at a higher degree of transparency than previous subscription-only Big Deals, which bundled access to large journal portfolios of commercial publishers [@Bergstrom_2014].
-However, the focus of transformative agreements remains primarily on these publishers with large hybrid journals and subscription-only portfolios [@Borrego_2020]. 
-<!-- national negotiations, oa2020 -->
-For example, the nationwide DEAL consortium in Germany, which includes universities and academic institutions, collectively cancelled their subscriptions with Elsevier from 2017 onwards to negotiate a transformative agreement [@Fraser_2023].
+These considerations resulted into transformative agreements, which cover a broad range of successfully negotiated contracts between library consortia and publishers from mid-2010s onwards where institutional spending for subscriptions are re-purposed for open access publishing [@Borrego_2020; @Hinchliffe_2019].
+Transformative agreements seek to control costs while allowing a transitional phase for publishing more open access articles. 
+Although transformative agreements mainly focus on large hybrid journals and subscription-only portfolios of commercial publishers,  they aim at a higher degree of transparency than previous big deals, where contracts including payments were confidential [@Bergstrom_2014].
+<!-- For example, the nationwide DEAL consortium in Germany, which includes universities and academic institutions, collectively cancelled their subscriptions with Elsevier from 2017 onwards to negotiate a transformative agreement [@Fraser_2023].
+-->
 
@@
 Specifically, cOAlition S funders agreed to support hybrid open access solely through transformative agreements from 1 January 2021, until the end of 2024.
-The German Research Foundation (DFG) has also extended its financial support for hybrid open access through transformative agreements, despite not being part of cOAlition S.
+Notably, the German Research Foundation (DFG), despite not being part of cOAlition S, has also extended its financial support for hybrid open access through transformative agreements [@Mittermaier_2021].
 Previously, the DFG only provided funding for full open access journals [@Jahn_2016].
 
-Advocates of these agreements argue that they not only overcome the issue standardiazation of double-dipping but also help to shift library spending from subscriptions to publications, streamline expenditures, and offer more value than big deals, which focus on large publishers.
-A few agreements can impact an institution's open access share, resulting in a significant proportion of open access content.
-increase transparency
-
-However, opponents of transformative agreements perceive continued reliance on big deals as problematic.
-They argue that the focus on large publishers perpetuates the business relationships that led to the serial crisis and lacks evidence that this leads to a reduction of In addition, concerns over inequality abound.
-trnasparency differ and now nsns
-
-could not benn mass removal
-
-Effects can be seen, mixed consequences.
-some countries who have introduced this fundign effects netherlands, sweden.
-no in germany, elsevier deal to price reduction and first steps towards mroe privacy.
-however, main goals associated woth trnasofrtiomative, the transition of the journal landscape not achivned, publsiehrs were not able to meet growtht raget, sweden did not see the flipping and has
-
-This study contributes
+By the end of 2023, many transformative agreements were implemented, but the outcomes were mixed.
+The ESAC Transformative Agreement Registry[^esacregistry], the largest source of disclosure, recorded more than 800 transformative agreements, resulting in up to 900.000 open access articles published in both full open access and hybrid journals according to the accompanying ESAC Market Watch[^esacmarketwatch].
+National consortia reported increased open access volume, streamlined payment and monitoring procedures, as well as extensive utilization of open access options by the researchers they serve [@Marques_2020; @Hinchliffe_2019; @Parmhed_2023].
+The ongoing standardisation of transformative agreements contributed to improved transparency in terms of contracts and publisher-provided article metadata [@Marques_2019].
+However, with the growing trend toward transformative agreements, continued reliance on big deals is perceived as problematic, perpetuating the business relationships that led to the serial crisis. 
+Whether transformative agreements lead to reduced pricing remains uncertain [@Borrego_2023].
+A substantial transition of hybrid journals towards full OA could not be observed [@Momeni_2021; @Matthias_2019], and the focus on large commercial publishers has led to concerns regarding inequality, because transformative agreements continue to target institutions from high-income countries [@Klebel_2023; @Butler_2023].
+Besides, an editorial-board resignation raised concerns that publishers' desire to maximize journal publication volume "without regard to quality" is a consequence of transformative agreements [@Goodin_2023].
+
+The controversies surrounding hybrid open access and transformative agreements have led to varying practical conclusions. 
+For instance, the Swedish BIBSAM consortium recommended to only support agreements for publishing in full open access journals.
+Likewise, most cOAlition S funders will discontinue financial support for transformative agreements from 2024 onwards [@Liverpool_2023].
+The consortium also removed most hybrid journals from its Transformative Journal program in 2023 due to their failure to meet self-defined open access growth targets. 
+In contrast, the German DEAL consortium signed a cost-effective five-year transformative agreement with Elsevier starting in 2024, while also renewing its contracts with Springer Nature and Wiley until the end of 2028.
+
+Against this background, the research objective of this study is to investigate the open access uptake of hybrid journals between 2018 and 2022, and the extent to which open access in hybrid journals can be attributed to transformative agreements.
+Previous quantitative science studies were able to investigate transformative agreements relative to specific countries [@P_l_nen_2020; @Huang_2020; @Haucap_2021; @Wenaas_2022; @Taubert_2023] and journal portfolios [@Pieper_2018; @Bakker_2024; @Fraser_2023; @Jahn_2022; @Momeni_2023].
+Others have focused soley on APC list prices [@Zhang_2022; @Shu_2023], not accounting for the peculiarities of transformative agreements. 
+Similarly, bibliometric databases, even though many allow the retrieval of open access articles in hybrid journals, do not directly attribute them to specific transformative agreements.
+
+The present study aims at a large-scale investigation using openly available data sources to answer the following research questions:
+
+- What are the number and proportion of open access articles in hybrid journals published under transformative agreements between 2018-2022?
+- Who is publishing open access in hybrid journals enabled by transformative agreements, and who is not?
+
+The results of the study will be group by publishers, subjects and countries to account for the controversies.
 
@@
 
+[^esacmarketwatch]: <https://esac-initiative.org/market-watch/>
+
 [^doaj]: <https://doaj.org/csv>

### 2024-02-14 analysis/manuscript.tex
continue introduction ii

@@
 full open access (Prosser, 2003). The idea was that when journals
-publish more and more open access articles, they could reduce revenues
+increasingly publish open access articles, they could reduce revenues
 from subscriptions, while libraries and funders could change their
@@
 1-2\% of articles in over 4,300 hybrid journals from large publishers
-were open access (Björk, 2012), growing to around 4\% between 2011 and
-2013 (Laakso \& Björk, 2016).
+were open access (B.-C. Björk, 2012), growing to around 4\% between 2011
+and 2013 (Laakso \& Björk, 2016).
 
 With the introduction of central funding mechanisms for hybrid open
-access publication fees in a few countries since 2012, notably in the
-UK, an increase in open access could be observed (Björk, 2017; Jubb et
-al., 2017; Robinson-Garcia et al., 2020). However, this shift in funding
-policy towards hybrid open access also added to the overall cost of
-publishing, which includes subscriptions spending and the administrative
-efforts required to handle payments (Pinfield et al., 2016). Moreover,
-established large commercial publishers, which already dominated the
-publishing market (Larivière et al., 2015), disproportionately benefited
-from hybrid open access funding in comparison to full open access
-publishers (Butler et al., 2023; Jahn \& Tullney, 2016; Shu \&
-Larivière, 2023).
+access publication fees in some European countries since 2012 an
+increase in open access could be observed (B.-C. Björk, 2017; Huang et
+al., 2020; Jubb et al., 2017; Piwowar et al., 2018). For example,
+studying university output, Robinson-Garcia et al. (2020) estimated a
+median uptake of 7.1\% in the period 2014-2017. In particular, British
+(17\%), Austrian (15\%) and Dutch (13\%) universities contributed to
+this trend. However, this shift in funding policy towards hybrid open
+access also added to the overall cost of publishing, which includes
+subscription spending and the administrative efforts required to handle
+payments (Pinfield et al., 2016). Moreover, established large commercial
+publishers, which already dominated the publishing market (Larivière et
+al., 2015), disproportionately benefited from hybrid open access funding
+in comparison to full open access publishers (Butler et al., 2023; Jahn
+\& Tullney, 2016; Shu \& Larivière, 2023).
 
@@
 and publication fees, as well as to increase publisher-provided
-immediate open access (BJÖRK \& SOLOMON, 2014; Schimmer et al., 2015).
-These considerations resulted into transformative agreements, which
-cover a broad range of successfully negotiated contracts between library
-consortia and publishers from mid-2010s onwards where institutional
-spending for subscriptions are re-purposed for open access publishing
-(Borrego et al., 2020). These transformative agreements seek to control
-costs while allowing a transitional phase for publishing more open
-access articles (Hinchliffe, 2019). At the same time, they aim at a
-higher degree of transparency than previous subscription-only Big Deals,
-which bundled access to large journal portfolios of commercial
-publishers (Bergstrom et al., 2014). However, the focus of
-transformative agreements remains primarily on these publishers with
-large hybrid journals and subscription-only portfolios (Borrego et al.,
-2020). For example, the nationwide DEAL consortium in Germany, which
-includes universities and academic institutions, collectively cancelled
-their subscriptions with Elsevier from 2017 onwards to negotiate a
-transformative agreement (Fraser et al., 2023).
+immediate open access (B. Björk \& Solomon, 2014; Schimmer et al.,
+2015). These considerations resulted into transformative agreements,
+which cover a broad range of successfully negotiated contracts between
+library consortia and publishers from mid-2010s onwards where
+institutional spending for subscriptions are re-purposed for open access
+publishing (Borrego et al., 2020; Hinchliffe, 2019). Transformative
+agreements seek to control costs while allowing a transitional phase for
+publishing more open access articles. Although transformative agreements
+mainly focus on large hybrid journals and subscription-only portfolios
+of commercial publishers, they aim at a higher degree of transparency
+than previous big deals, where contracts including payments were
+confidential (Bergstrom et al., 2014).
 
@@
 transformative agreements from 1 January 2021, until the end of 2024.
-The German Research Foundation (DFG) has also extended its financial
-support for hybrid open access through transformative agreements,
-despite not being part of cOAlition S. Previously, the DFG only provided
-funding for full open access journals (Jahn \& Tullney, 2016).
+Notably, the German Research Foundation (DFG), despite not being part of
+cOAlition S, has also extended its financial support for hybrid open
+access through transformative agreements (Mittermaier, 2021).
+Previously, the DFG only provided funding for full open access journals
+(Jahn \& Tullney, 2016).
 
-Advocates of these agreements argue that they not only overcome the
-issue standardiazation of double-dipping but also help to shift library
-spending from subscriptions to publications, streamline expenditures,
-and offer more value than big deals, which focus on large publishers. A
-few agreements can impact an institution's open access share, resulting
-in a significant proportion of open access content. increase
-transparency
+By the end of 2023, many transformative agreements were implemented, but
+the outcomes are mixed. The ESAC Transformative Agreement
+Registry\footnote{\url{https://esac-initiative.org/about/transformative-agreements/agreement-registry/}},
+the largest source of disclosure, recorded more than 800 transformative
+agreements, resulting in up to 900.000 open access articles published in
+both full open access and hybrid journals according to the accompanying
+ESAC Market Watch\footnote{\url{https://esac-initiative.org/market-watch/}}.
+National consortia reported increased open access volume, streamlined
+payment and monitoring procedures, as well as extensive utilization of
+open access options by the researchers they serve (Hinchliffe, 2019;
+Marques \& Stone, 2020; Parmhed \& Säll, 2023). The ongoing
+standardisation of transformative agreements contributed to improved
+transparency in terms of contracts and publisher-provided article
+metadata (Marques et al., 2019). However, with the growing trend toward
+transformative agreements, continued reliance on big deals is perceived
+as problematic, perpetuating the business relationships that led to the
+serial crisis. Whether transformative agreements lead to reduced pricing
+remains uncertain (\textbf{Borrego\_2023?}). A substantial transition of
+hybrid journals towards full OA could not be observed (Matthias et al.,
+2019; Momeni et al., 2021), and the focus on large commercial publishers
+has led to concerns regarding inequality, because transformative
+agreements continue to target institutions from high-income countries
+(Butler et al., 2023; Klebel \& Ross-Hellauer, 2023). Besides, an
+editorial-board resignation raised concerns that publishers' desire to
+maximize journal publication volume ``without regard to quality'' is a
+consequence of transformative agreements (Rasmussen, 2023).
 
-However, opponents of transformative agreements perceive continued
-reliance on big deals as problematic. They argue that the focus on large
-publishers perpetuates the business relationships that led to the serial
-crisis and lacks evidence that this leads to a reduction of In addition,
-concerns over inequality abound. trnasparency differ and now nsns
+The controversies surrounding hybrid open access and transformative
+agreements have led to varying practical conclusions. For instance, the
+Swedish BIBSAM consortium recommended to only support agreements for
+publishing in full open access journals. Likewise, most cOAlition S
+funders will discontinue financial support for transformative agreements
+from 2024 onwards (Liverpool, 2023). The consortium also removed most
+hybrid journals from its Transformative Journal program in 2023 due to
+their failure to meet self-defined open access growth targets. In
+contrast, the German DEAL consortium signed a cost-effective five-year
+transformative agreement with Elsevier starting in 2024, while also
+renewing its contracts with Springer Nature and Wiley until the end of
+2028.
 
-could not benn mass removal
+Against this background, the research objective of this study is to
+investigate the open access uptake of hybrid journals between 2018 and
+2022, and the extent to which open access in hybrid journals can be
+attributed to transformative agreements. Previous quantitative science
+studies were able to investigate transformative agreements relative to
+specific countries (Haucap et al., 2021; Huang et al., 2020; Pölönen et
+al., 2020; Taubert et al., 2023; Wenaas, 2022) and journal portfolios
+(Bakker et al., 2024; Fraser et al., 2023; Momeni et al., 2023; Pieper
+\& Broschinski, 2018; \textbf{Jahn\_2022?}). Others have focused soley
+on APC list prices (Shu \& Larivière, 2023; Zhang et al., 2022), not
+accounting for the peculiarities of transformative agreements.
+Similarly, bibliometric databases, even though many allow the retrieval
+of open access articles in hybrid journals, do not directly attribute
+them to specific transformative agreements.
 
-Effects can be seen, mixed consequences. some countries who have
-introduced this fundign effects netherlands, sweden. no in germany,
-elsevier deal to price reduction and first steps towards mroe privacy.
-however, main goals associated woth trnasofrtiomative, the transition of
-the journal landscape not achivned, publsiehrs were not able to meet
-growtht raget, sweden did not see the flipping and has
+The present study aims at a large-scale investigation using openly
+available data sources to answer the following research questions:
 
-This study contributes
+\begin{itemize}
+\tightlist
+\item
+  What are the number and proportion of open access articles in hybrid
+  journals published under transformative agreements between 2018-2022?
+\item
+  Who is publishing open access in hybrid journals enabled by
+  transformative agreements, and who is not?
+\end{itemize}
+
+The results of the study will be group by publishers, subjects and
+countries to account for the controversies.
 
@@
 content licenses (Piwowar et al., 2018), highlighted ongoing challenges
-to identify hybrid open access (Butler et al., 2023; Jahn et al., 2021;
+to identify hybrid open access (Butler et al., 2023; Jahn et al., 2022;
 Martín-Martín et al., 2018; Zhang et al., 2022). Here, 742,369 articles
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-Bakker_2024}{}}%
+Bakker, C., Langham-Putrow, A., \& Riegelman, A. (2024). Impact of
+transformative agreements on publication patterns: An analysis based on
+agreements from the ESAC registry. \emph{International Journal of
+Librarianship}, \emph{8}(4), 67--96.
+\url{https://doi.org/10.23974/ijol.2024.vol8.4.341}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Bergstrom_2014}{}}%
@@
 \leavevmode\vadjust pre{\hypertarget{ref-BJ_RK_2014}{}}%
-BJÖRK, B., \& SOLOMON, D. (2014). How research funders can finance APCs
+Björk, B., \& Solomon, D. (2014). How research funders can finance APCs
 in full OA and hybrid journals. \emph{Learned Publishing}, \emph{27}(2),
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Haucap_2021}{}}%
+Haucap, J., Moshgbar, N., \& Schmal, W. B. (2021). The impact of the
+german {``DEAL''} on competition in the academic publishing market.
+\emph{Managerial and Decision Economics}, \emph{42}(8), 2027--2049.
+\url{https://doi.org/10.1002/mde.3493}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Hinchliffe_2019}{}}%
@@
 The Scholarly Kitchen.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Huang_2020}{}}%
+Huang, C.-K. (Karl), Neylon, C., Hosking, R., Montgomery, L., Wilson, K.
+S., Ozaygen, A., \& Brookes-Kenworthy, C. (2020). Evaluating the impact
+of open access policies on research institutions. \emph{{eLife}},
+\emph{9}. \url{https://doi.org/10.7554/elife.57067}
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Jahn_2021}{}}%
-Jahn, N., Matthias, L., \& Laakso, M. (2021). Toward transparency of
+Jahn, N., Matthias, L., \& Laakso, M. (2022). Toward transparency of
 hybrid open access through publisher-provided metadata: An article-level
@@
 \url{https://web.archive.org/web/20200212015524/https://www.universitiesuk.ac.uk/policy-and-analysis/reports/Documents/2017/monitoring-transition-open-access-2017.pdf}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Klebel_2023}{}}%
+Klebel, T., \& Ross-Hellauer, T. (2023). The APC-barrier and its effect
+on stratification in open access publishing. \emph{Quantitative Science
+Studies}, \emph{4}(1), 22--43. \url{https://doi.org/10.1162/qss_a_00245}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Liverpool_2023}{}}%
+Liverpool, L. (2023). Open-access reformers launch next bold publishing
+plan. \emph{Nature}, \emph{623}(7986), 238--240.
+\url{https://doi.org/10.1038/d41586-023-03342-6}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Marques_2020}{}}%
+Marques, M., \& Stone, G. (2020). Transitioning to open access: An
+evaluation of the {UK} springer compact agreement pilot 2016--2018.
+\emph{College {\&} Research Libraries}, \emph{81}(6), 913--927.
+\url{https://doi.org/10.5860/crl.81.6.913}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Marques_2019}{}}%
+Marques, M., Woutersen-Windhouwer, S., \& Tuuliniemi, A. (2019).
+Monitoring agreements with open access elements: Why article-level
+metadata are important. \emph{Insights the {UKSG} Journal}, \emph{32}.
+\url{https://doi.org/10.1629/uksg.489}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Mart_n_Mart_n_2018}{}}%
@@
 \url{https://doi.org/10.1080/00031305.2017.1375986}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Matthias_2019}{}}%
+Matthias, L., Jahn, N., \& Laakso, M. (2019). The two-way street of open
+access journal publishing: Flip it and reverse it. \emph{Publications},
+\emph{7}(2), 23. \url{https://doi.org/10.3390/publications7020023}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Mittermaier_2021}{}}%
+Mittermaier, B. (2021). {Rolle des Open Access Monitor Deutschland bei
+der Antragstellung im DFG-Förderprogramm
+Open-Access-Publikationskosten}. \emph{{O-Bib. Das Offene
+Bibliotheksjournal / Herausgeber VDB}}, \emph{8}.
+\url{https://doi.org/10.5282/O-BIB/5731}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Momeni_2023}{}}%
+Momeni, F., Dietze, S., Mayr, P., Biesenbender, K., \& Peters, I.
+(2023). Which factors are associated with open access publishing? A
+springer nature case study. \emph{Quantitative Science Studies},
+\emph{4}(2), 353--371. \url{https://doi.org/10.1162/qss_a_00253}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Momeni_2021}{}}%
+Momeni, F., Mayr, P., Fraser, N., \& Peters, I. (2021). What happens
+when a journal converts to open access? A bibliometric analysis.
+\emph{Scientometrics}, \emph{126}(12), 9811--9827.
+\url{https://doi.org/10.1007/s11192-021-03972-5}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Parmhed_2023}{}}%
+Parmhed, S., \& Säll, J. (2023). Transformative agreements and their
+practical impact: A librarian perspective. \emph{Insights the UKSG
+Journal}, \emph{36}. \url{https://doi.org/10.1629/uksg.612}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Pieper_2018}{}}%
+Pieper, D., \& Broschinski, C. (2018). {OpenAPC}: A contribution to a
+transparent and reproducible monitoring of fee-based open access
+publishing across institutions and nations. \emph{Insights the {UKSG}
+Journal}, \emph{31}. \url{https://doi.org/10.1629/uksg.439}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-P_l_nen_2020}{}}%
+Pölönen, J., Laakso, M., Guns, R., Kulczycki, E., \& Sivertsen, G.
+(2020). Open access at the national level: A comprehensive analysis of
+publications by finnish researchers. \emph{Quantitative Science
+Studies}, \emph{1}(4), 1396--1428.
+\url{https://doi.org/10.1162/qss_a_00084}
+
 \leavevmode\vadjust pre{\hypertarget{ref-priem2022openalex}{}}%
@@
 \url{https://doi.org/10.1087/095315103322110923}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Goodin_2023}{}}%
+Rasmussen, K. B. (2023). Interview with robert {``bob''} e. goodin.
+\emph{Tidskrift För Politisk Filosofi}.
+\url{https://www.politiskfilosofi.se/fulltext/2023-2/pdf/TPF_2023-2_interview_with_robert_bob_e_goodin.pdf}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Taubert_2023}{}}%
+Taubert, N., Hobert, A., Jahn, N., Bruns, A., \& Iravani, E. (2023).
+Understanding differences of the OA uptake within the german university
+landscape (2010--2020): Part 1---journal-based OA.
+\emph{Scientometrics}, \emph{128}(6), 3601--3625.
+\url{https://doi.org/10.1007/s11192-023-04716-3}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Wenaas_2022}{}}%
+Wenaas, L. (2022). Choices of immediate open access and the relationship
+to journal ranking and publish-and-read deals. \emph{Frontiers in
+Research Metrics and Analytics}, \emph{7}.
+\url{https://doi.org/10.3389/frma.2022.943932}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Zhang_2022}{}}%

### 2024-02-15 analysis/manuscript.Rmd
Finish introduction

@@
 The idea was that when journals increasingly publish open access articles, they could reduce revenues from subscriptions, while libraries and funders could change their funding models and shift expenditures from subscription to open access.
-However, initial approaches, mainly based on publication fees, also called article processing charges (APCs), did not contribute substantially to an large open access uptake. 
+However, initial approaches, mainly based on publication fees, also called article processing charges (APCs), did not contribute substantially to a large open access uptake. 
 In 2009, Springer reported open access to 1% of articles in hybrid journals [@Dallmeier_2010].
 Other publishers also recorded a low uptake.
-In 2011, only 1-2% of articles in over 4,300 hybrid journals from large publishers were open access [@Bj_rk_2012], growing to around 4% between 2011 and 2013 [@Laakso_2016].
-
-With the introduction of central funding mechanisms for hybrid open access publication fees in some European countries since 2012 an increase in open access could be observed [@Bj_rk_2017; @Piwowar_2018; @Jubb_2017; @Huang_2020].
+In 2011, only 1-2% of articles were open access [@Bj_rk_2012], growing to around 4% between 2011 and 2013 [@Laakso_2016].
+
+With the introduction of central funding mechanisms for hybrid open access publication fees in some European countries since 2012, an increase in open access could be observed [@Bj_rk_2017; @Piwowar_2018; @Jubb_2017; @Huang_2020].
 For example, studying university output, @Robinson_Garcia_2020 estimated a median uptake of 7.1% in the period 2014-2017. 
@@
 Transformative agreements seek to control costs while allowing a transitional phase for publishing more open access articles. 
-Although transformative agreements mainly focus on large hybrid journals and subscription-only portfolios of commercial publishers,  they aim at a higher degree of transparency than previous big deals, where contracts including payments were confidential [@Bergstrom_2014].
+Although transformative agreements mainly bundle hybrid and subscription-only journals from commercial publishers, they aim at a higher degree of transparency than previous big deals, where contracts including payments were confidential [@Bergstrom_2014].
 <!-- For example, the nationwide DEAL consortium in Germany, which includes universities and academic institutions, collectively cancelled their subscriptions with Elsevier from 2017 onwards to negotiate a transformative agreement [@Fraser_2023].
@@
 The introduction of transformative agreements aligns with funding policy changes, such as the decision made by the cOAlition S, a consortium of national funders including the European Commission, to no longer provide financial support for individual publication fees when publishing in hybrid journals.
-According ot its Plan S, they only accept hybrid open access through transformative agreements "during a transition period that should be as short as possible" [@Schiltz_2018]. 
-Specifically, cOAlition S funders agreed to support hybrid open access solely through transformative agreements from 1 January 2021, until the end of 2024.
+According ot its Plan S launched in 2018, the cOAlition S funders only accept hybrid open access through transformative agreements "during a transition period that should be as short as possible" [@Schiltz_2018].
+Specifically, they agreed to support hybrid open access solely through transformative agreements from 1 January 2021, until the end of 2024.
 Notably, the German Research Foundation (DFG), despite not being part of cOAlition S, has also extended its financial support for hybrid open access through transformative agreements [@Mittermaier_2021].
@@
 
-By the end of 2023, many transformative agreements were implemented, but the outcomes were mixed.
+By the end of 2023, many transformative agreements were implemented, but outcomes were mixed.
 The ESAC Transformative Agreement Registry[^esacregistry], the largest source of disclosure, recorded more than 800 transformative agreements, resulting in up to 900.000 open access articles published in both full open access and hybrid journals according to the accompanying ESAC Market Watch[^esacmarketwatch].
-National consortia reported increased open access volume, streamlined payment and monitoring procedures, as well as extensive utilization of open access options by the researchers they serve [@Marques_2020; @Hinchliffe_2019; @Parmhed_2023].
+Library consortia reported increased open access volume, streamlined payment and monitoring procedures, as well as extensive utilization of open access options by the researchers they serve [@Marques_2020; @Pinhasi_2020; @Parmhed_2023].
 The ongoing standardisation of transformative agreements contributed to improved transparency in terms of contracts and publisher-provided article metadata [@Marques_2019].
-However, with the growing trend toward transformative agreements, continued reliance on big deals is perceived as problematic, perpetuating the business relationships that led to the serial crisis. 
-Whether transformative agreements lead to reduced pricing remains uncertain [@Borrego_2023].
-A substantial transition of hybrid journals towards full OA could not be observed [@Momeni_2021; @Matthias_2019], and the focus on large commercial publishers has led to concerns regarding inequality, because transformative agreements continue to target institutions from high-income countries [@Klebel_2023; @Butler_2023].
+However, with the growing trend toward transformative agreements, continued reliance on big deals is perceived as problematic, which might perpetuate market concentration [@Butler_2023; @Shu_2023]. 
+Whether transformative agreements lead to reduced pricing remains uncertain [@Borrego_2023] and
+a substantial transition of hybrid journals towards full open access could not be observed [@Momeni_2021; @Matthias_2019].
+Furthermore, the focus on large commercial publishers might increase inequality, because transformative agreements continue to target institutions from high-income countries [@Klebel_2023]. 
 Besides, an editorial-board resignation raised concerns that publishers' desire to maximize journal publication volume "without regard to quality" is a consequence of transformative agreements [@Goodin_2023].
 
-The controversies surrounding hybrid open access and transformative agreements have led to varying practical conclusions. 
-For instance, the Swedish BIBSAM consortium recommended to only support agreements for publishing in full open access journals.
+The controversies surrounding hybrid open access and transformative agreements have led to varying policy conclusions. 
+For instance, the  Association of Swedish Higher Education Institutions (Sveriges universitets- och
+högskoleförbund, SUHF) recommended to only support agreements for publishing in full open access journals [^sweden].
 Likewise, most cOAlition S funders will discontinue financial support for transformative agreements from 2024 onwards [@Liverpool_2023].
-The consortium also removed most hybrid journals from its Transformative Journal program in 2023 due to their failure to meet self-defined open access growth targets. 
-In contrast, the German DEAL consortium signed a cost-effective five-year transformative agreement with Elsevier starting in 2024, while also renewing its contracts with Springer Nature and Wiley until the end of 2028.
-
-Against this background, the research objective of this study is to investigate the open access uptake of hybrid journals between 2018 and 2022, and the extent to which open access in hybrid journals can be attributed to transformative agreements.
-Previous quantitative science studies were able to investigate transformative agreements relative to specific countries [@P_l_nen_2020; @Huang_2020; @Haucap_2021; @Wenaas_2022; @Taubert_2023] and journal portfolios [@Pieper_2018; @Bakker_2024; @Fraser_2023; @Jahn_2022; @Momeni_2023].
-Others have focused soley on APC list prices [@Zhang_2022; @Shu_2023], not accounting for the peculiarities of transformative agreements. 
-Similarly, bibliometric databases, even though many allow the retrieval of open access articles in hybrid journals, do not directly attribute them to specific transformative agreements.
-
-The present study aims at a large-scale investigation using openly available data sources to answer the following research questions:
-
-- What are the number and proportion of open access articles in hybrid journals published under transformative agreements between 2018-2022?
-- Who is publishing open access in hybrid journals enabled by transformative agreements, and who is not?
-
-The results of the study will be group by publishers, subjects and countries to account for the controversies.
+The consortium also removed most hybrid journals from its Transformative Journal program in 2023 due to their failure to meet self-defined open access growth targets [@Brainard_2023]. 
+In contrast, the German DEAL consortium signed a five-year transformative agreement with Elsevier starting in 2024, while also renewing its contracts with Springer Nature and Wiley until the end of 2028.
+
+Despite these controversies around transformative agreements as a means of transitioning journal publishing to full open access, there is limited evidence available on the adoption of open access in hybrid journals, and the extent to which this uptake can be attributed to transformative agreements.
+Previous studies have focused on specific countries [@P_l_nen_2020; @Huang_2020; @Haucap_2021; @Wenaas_2022; @Taubert_2023] or publisher portfolios [@Pieper_2018; @Bakker_2024; @Fraser_2023; @Jahn_2021; @Momeni_2023], while related large-scale studies relied on self-reported agreement data [@Moskovkin_2022], or used APC pricing lists [@Shu_2023].
+Particularly, data availability is a limiting factor when studying the impact of transformative agreements [@Bakker_2024], because bibliometric databases, even though many allow the retrieval of open access articles in hybrid journals, do not directly attribute them to specific transformative agreements.
+
+The present study aims to address these limitations by combining multiple openly available data sources to determine the open access uptake in hybrid journals, while distinguishing between open access through transformative agreements and other means.
+With this novel and open approach, this study aims to answer the following questions:
+
+- What was the number and proportion of open access articles in hybrid journals included in transformative agreements between 2018 and 2022?
+- How did transformative agreements impact open access uptake in hybrid journals?
+
+For both of these research questions, this study will analyse the variability by publisher, journal subject, and country.
 
@@
 However, because of low coverage in OpenAlex, this study focused on first authors and their affiliations instead.
-First authors typically contribute most to a paper and are often considered lead author research papers [@Larivi_re_2016], and can be therefore assumed as a proxy to measure to open access payments [@Zhang_2022].
+First authors typically contribute most to a paper and are often considered lead author research papers [@Larivi_re_2016], and can be therefore assumed as a proxy to measure to open access payments [@Zhang_2022; @Shu_2023].
 Overall, around 90% of studied articles had first author affiliation metadata in OpenAlex, whereas the coverage of articles with corresponding author information was around 54%.
@@
 [^checkertool]: <https://www.coalition-s.org/blog/enabling-accurate-results-within-the-journal-checker-tool/>
+
+[^sweden]: <https://www.su.se/english/news/open-access-need-to-move-away-from-transformative-agreements-1.683787>

### 2024-02-15 analysis/manuscript.tex
Finish introduction

@@
 called article processing charges (APCs), did not contribute
-substantially to an large open access uptake. In 2009, Springer reported
+substantially to a large open access uptake. In 2009, Springer reported
 open access to 1\% of articles in hybrid journals (Dallmeier-Tiessen et
 al., 2010). Other publishers also recorded a low uptake. In 2011, only
-1-2\% of articles in over 4,300 hybrid journals from large publishers
-were open access (B.-C. Björk, 2012), growing to around 4\% between 2011
-and 2013 (Laakso \& Björk, 2016).
+1-2\% of articles were open access (Björk, 2012), growing to around 4\%
+between 2011 and 2013 (Laakso \& Björk, 2016).
 
 With the introduction of central funding mechanisms for hybrid open
-access publication fees in some European countries since 2012 an
-increase in open access could be observed (B.-C. Björk, 2017; Huang et
-al., 2020; Jubb et al., 2017; Piwowar et al., 2018). For example,
-studying university output, Robinson-Garcia et al. (2020) estimated a
-median uptake of 7.1\% in the period 2014-2017. In particular, British
-(17\%), Austrian (15\%) and Dutch (13\%) universities contributed to
-this trend. However, this shift in funding policy towards hybrid open
-access also added to the overall cost of publishing, which includes
-subscription spending and the administrative efforts required to handle
-payments (Pinfield et al., 2016). Moreover, established large commercial
+access publication fees in some European countries since 2012, an
+increase in open access could be observed (Björk, 2017; Huang et al.,
+2020; Jubb et al., 2017; Piwowar et al., 2018). For example, studying
+university output, Robinson-Garcia et al. (2020) estimated a median
+uptake of 7.1\% in the period 2014-2017. In particular, British (17\%),
+Austrian (15\%) and Dutch (13\%) universities contributed to this trend.
+However, this shift in funding policy towards hybrid open access also
+added to the overall cost of publishing, which includes subscription
+spending and the administrative efforts required to handle payments
+(Pinfield et al., 2016). Moreover, established large commercial
 publishers, which already dominated the publishing market (Larivière et
@@
 and publication fees, as well as to increase publisher-provided
-immediate open access (B. Björk \& Solomon, 2014; Schimmer et al.,
-2015). These considerations resulted into transformative agreements,
-which cover a broad range of successfully negotiated contracts between
-library consortia and publishers from mid-2010s onwards where
-institutional spending for subscriptions are re-purposed for open access
-publishing (Borrego et al., 2020; Hinchliffe, 2019). Transformative
-agreements seek to control costs while allowing a transitional phase for
-publishing more open access articles. Although transformative agreements
-mainly focus on large hybrid journals and subscription-only portfolios
-of commercial publishers, they aim at a higher degree of transparency
-than previous big deals, where contracts including payments were
-confidential (Bergstrom et al., 2014).
+immediate open access (Björk \& Solomon, 2014; Schimmer et al., 2015).
+These considerations resulted into transformative agreements, which
+cover a broad range of successfully negotiated contracts between library
+consortia and publishers from mid-2010s onwards where institutional
+spending for subscriptions are re-purposed for open access publishing
+(Borrego et al., 2020; Hinchliffe, 2019). Transformative agreements seek
+to control costs while allowing a transitional phase for publishing more
+open access articles. Although transformative agreements mainly bundle
+hybrid and subscription-only journals from commercial publishers, they
+aim at a higher degree of transparency than previous big deals, where
+contracts including payments were confidential (Bergstrom et al., 2014).
 
@@
 financial support for individual publication fees when publishing in
-hybrid journals. According ot its Plan S, they only accept hybrid open
-access through transformative agreements ``during a transition period
-that should be as short as possible'' (Schiltz, 2018). Specifically,
-cOAlition S funders agreed to support hybrid open access solely through
-transformative agreements from 1 January 2021, until the end of 2024.
-Notably, the German Research Foundation (DFG), despite not being part of
-cOAlition S, has also extended its financial support for hybrid open
-access through transformative agreements (Mittermaier, 2021).
-Previously, the DFG only provided funding for full open access journals
-(Jahn \& Tullney, 2016).
+hybrid journals. According ot its Plan S launched in 2018, the cOAlition
+S funders only accept hybrid open access through transformative
+agreements ``during a transition period that should be as short as
+possible'' (Schiltz, 2018). Specifically, they agreed to support hybrid
+open access solely through transformative agreements from 1 January
+2021, until the end of 2024. Notably, the German Research Foundation
+(DFG), despite not being part of cOAlition S, has also extended its
+financial support for hybrid open access through transformative
+agreements (Mittermaier, 2021). Previously, the DFG only provided
+funding for full open access journals (Jahn \& Tullney, 2016).
 
 By the end of 2023, many transformative agreements were implemented, but
-the outcomes are mixed. The ESAC Transformative Agreement
+outcomes were mixed. The ESAC Transformative Agreement
 Registry\footnote{\url{https://esac-initiative.org/about/transformative-agreements/agreement-registry/}},
@@
 ESAC Market Watch\footnote{\url{https://esac-initiative.org/market-watch/}}.
-National consortia reported increased open access volume, streamlined
+Library consortia reported increased open access volume, streamlined
 payment and monitoring procedures, as well as extensive utilization of
-open access options by the researchers they serve (Hinchliffe, 2019;
-Marques \& Stone, 2020; Parmhed \& Säll, 2023). The ongoing
+open access options by the researchers they serve (Marques \& Stone,
+2020; Parmhed \& Säll, 2023; Pinhasi et al., 2020). The ongoing
 standardisation of transformative agreements contributed to improved
@@
 transformative agreements, continued reliance on big deals is perceived
-as problematic, perpetuating the business relationships that led to the
-serial crisis. Whether transformative agreements lead to reduced pricing
-remains uncertain (\textbf{Borrego\_2023?}). A substantial transition of
-hybrid journals towards full OA could not be observed (Matthias et al.,
-2019; Momeni et al., 2021), and the focus on large commercial publishers
-has led to concerns regarding inequality, because transformative
-agreements continue to target institutions from high-income countries
-(Butler et al., 2023; Klebel \& Ross-Hellauer, 2023). Besides, an
-editorial-board resignation raised concerns that publishers' desire to
-maximize journal publication volume ``without regard to quality'' is a
-consequence of transformative agreements (Rasmussen, 2023).
+as problematic, which might perpetuate market concentration (Butler et
+al., 2023; Shu \& Larivière, 2023). Whether transformative agreements
+lead to reduced pricing remains uncertain (Borrego, 2023) and a
+substantial transition of hybrid journals towards full open access could
+not be observed (Matthias et al., 2019; Momeni et al., 2021).
+Furthermore, the focus on large commercial publishers might increase
+inequality, because transformative agreements continue to target
+institutions from high-income countries (Klebel \& Ross-Hellauer, 2023).
+Besides, an editorial-board resignation raised concerns that publishers'
+desire to maximize journal publication volume ``without regard to
+quality'' is a consequence of transformative agreements (Rasmussen,
+2023).
 
 The controversies surrounding hybrid open access and transformative
-agreements have led to varying practical conclusions. For instance, the
-Swedish BIBSAM consortium recommended to only support agreements for
-publishing in full open access journals. Likewise, most cOAlition S
-funders will discontinue financial support for transformative agreements
-from 2024 onwards (Liverpool, 2023). The consortium also removed most
-hybrid journals from its Transformative Journal program in 2023 due to
-their failure to meet self-defined open access growth targets. In
-contrast, the German DEAL consortium signed a cost-effective five-year
-transformative agreement with Elsevier starting in 2024, while also
-renewing its contracts with Springer Nature and Wiley until the end of
-2028.
-
-Against this background, the research objective of this study is to
-investigate the open access uptake of hybrid journals between 2018 and
-2022, and the extent to which open access in hybrid journals can be
-attributed to transformative agreements. Previous quantitative science
-studies were able to investigate transformative agreements relative to
-specific countries (Haucap et al., 2021; Huang et al., 2020; Pölönen et
-al., 2020; Taubert et al., 2023; Wenaas, 2022) and journal portfolios
-(Bakker et al., 2024; Fraser et al., 2023; Momeni et al., 2023; Pieper
-\& Broschinski, 2018; \textbf{Jahn\_2022?}). Others have focused soley
-on APC list prices (Shu \& Larivière, 2023; Zhang et al., 2022), not
-accounting for the peculiarities of transformative agreements.
-Similarly, bibliometric databases, even though many allow the retrieval
-of open access articles in hybrid journals, do not directly attribute
-them to specific transformative agreements.
-
-The present study aims at a large-scale investigation using openly
-available data sources to answer the following research questions:
+agreements have led to varying policy conclusions. For instance, the
+Association of Swedish Higher Education Institutions (Sveriges
+universitets- och högskoleförbund, SUHF) recommended to only support
+agreements for publishing in full open access journals \footnote{\url{https://www.su.se/english/news/open-access-need-to-move-away-from-transformative-agreements-1.683787}}.
+Likewise, most cOAlition S funders will discontinue financial support
+for transformative agreements from 2024 onwards (Liverpool, 2023). The
+consortium also removed most hybrid journals from its Transformative
+Journal program in 2023 due to their failure to meet self-defined open
+access growth targets (Brainard, 2023). In contrast, the German DEAL
+consortium signed a five-year transformative agreement with Elsevier
+starting in 2024, while also renewing its contracts with Springer Nature
+and Wiley until the end of 2028.
+
+Despite these controversies around transformative agreements as a means
+of transitioning journal publishing to full open access, there is
+limited evidence available on the adoption of open access in hybrid
+journals, and the extent to which this uptake can be attributed to
+transformative agreements. Previous studies have focused on specific
+countries (Haucap et al., 2021; Huang et al., 2020; Pölönen et al.,
+2020; Taubert et al., 2023; Wenaas, 2022) or publisher portfolios
+(Bakker et al., 2024; Fraser et al., 2023; Jahn et al., 2022; Momeni et
+al., 2023; Pieper \& Broschinski, 2018), while related large-scale
+studies relied on self-reported agreement data (Moskovkin et al., 2022),
+or used APC pricing lists (Shu \& Larivière, 2023). Particularly, data
+availability is a limiting factor when studying the impact of
+transformative agreements (Bakker et al., 2024), because bibliometric
+databases, even though many allow the retrieval of open access articles
+in hybrid journals, do not directly attribute them to specific
+transformative agreements.
+
+The present study aims to address these limitations by combining
+multiple openly available data sources to determine the open access
+uptake in hybrid journals, while distinguishing between open access
+through transformative agreements and other means. With this novel and
+open approach, this study aims to answer the following questions:
 
@@
 \item
-  What are the number and proportion of open access articles in hybrid
-  journals published under transformative agreements between 2018-2022?
+  What was the number and proportion of open access articles in hybrid
+  journals included in transformative agreements between 2018 and 2022?
 \item
-  Who is publishing open access in hybrid journals enabled by
-  transformative agreements, and who is not?
+  How did transformative agreements impact open access uptake in hybrid
+  journals?
 \end{itemize}
 
-The results of the study will be group by publishers, subjects and
-countries to account for the controversies.
+For both of these research questions, this study will analyse the
+variability by publisher, journal subject, and country.
 
@@
 papers (Larivière et al., 2016), and can be therefore assumed as a proxy
-to measure to open access payments (Zhang et al., 2022). Overall, around
-90\% of studied articles had first author affiliation metadata in
-OpenAlex, whereas the coverage of articles with corresponding author
-information was around 54\%.
+to measure to open access payments (Shu \& Larivière, 2023; Zhang et
+al., 2022). Overall, around 90\% of studied articles had first author
+affiliation metadata in OpenAlex, whereas the coverage of articles with
+corresponding author information was around 54\%.
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-BJ_RK_2014}{}}%
-Björk, B., \& Solomon, D. (2014). How research funders can finance APCs
-in full OA and hybrid journals. \emph{Learned Publishing}, \emph{27}(2),
-93--103. \url{https://doi.org/10.1087/20140203}
+Björk, B.-C., \& Solomon, D. (2014). How research funders can finance
+APCs in full OA and hybrid journals. \emph{Learned Publishing},
+\emph{27}(2), 93--103. \url{https://doi.org/10.1087/20140203}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Borrego_2023}{}}%
+Borrego, Á. (2023). Article processing charges for open access journal
+publishing: A review. \emph{Learned Publishing}, \emph{36}(3), 359--378.
+\url{https://doi.org/10.1002/leap.1558}
 
@@
 Publishing}. \url{https://doi.org/10.1002/leap.1347}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Brainard_2023}{}}%
+Brainard, J. (2023). {``Transformative''} journals get booted for
+switching to open access too slowly. In \emph{Science: SCIENCEINSIDER}.
+\url{https://doi.org/10.1126/science.adj3282}
 
@@
 \url{https://doi.org/10.1007/s11192-021-03972-5}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Moskovkin_2022}{}}%
+Moskovkin, V. M., Saprykina, T. V., \& Boichuk, I. V. (2022).
+Transformative agreements in the development of open access.
+\emph{Journal of Electronic Resources Librarianship}, \emph{34}(3),
+165--207. \url{https://doi.org/10.1080/1941126x.2022.2099000}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Pinhasi_2020}{}}%
+Pinhasi, R., Kromp, B., Blechl, G., \& Hölbling, L. (2020). The impact
+of open access publishing agreements at the {University of Vienna} in
+light of the plan s requirements: A review of current status, challenges
+and perspectives. \emph{Insights the UKSG Journal}, \emph{33}.
+\url{https://doi.org/10.1629/uksg.523}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Piwowar_2018}{}}%
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Goodin_2023}{}}%
-Rasmussen, K. B. (2023). Interview with robert {``bob''} e. goodin.
+Rasmussen, K. B. (2023). Interview with {Robert {``Bob''} E. Goodin}.
 \emph{Tidskrift För Politisk Filosofi}.

### 2024-02-15 analysis/manuscript.Rmd
minor

@@
 
-For over two decades, hybrid open access journal publishing, which makes some articles openly available while others remain behind a paywall, has been discussed as a  model for transitioning the subscription system to full open access [@Prosser_2003].
+For over two decades, hybrid open access journal publishing, which makes some articles openly available while others remain behind a paywall, has been discussed as a means for transitioning the subscription system to full open access [@Prosser_2003].
 The idea was that when journals increasingly publish open access articles, they could reduce revenues from subscriptions, while libraries and funders could change their funding models and shift expenditures from subscription to open access.
@@
 
-Throughout this mostly automated data gathering and analysis process, we used tools from the Tidyverse (Wickham et al., 2019) for the R programming language (R Core Team, 2020). 
-The resulting data is openly available through an R package, hoaddata. 
+Throughout this mostly automated data gathering and analysis process, tools from the Tidyverse (Wickham et al., 2019) for the R programming language (R Core Team, 2020) were used. 
+The resulting data is openly available through an R package, hoaddata, version 0.2.91 
 Following @Marwick_2018, hoaddata contains not only the datasets used in the data analysis. 
@@
 #| fig-cap: "Relative growth of open access in hybrid journals in transformative agreements between 2018 and 2022 per publication year.
-The blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established (according to matching OpenAlex first author affiliations matched with cOAlition S transformative agreement data).
+The blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established.
 (A) Proportion of open access articles in hybrid journals per year.
@@
 ```
+
+Analysing hybrid open access across publishers between 2018 and 2022 reveals a large market concentration. 
+Although `r length(unique(jn_ind_jct_all$esac_publisher))` publishers offered transformative agreements, the big three commercial publishers Elsevier, Springer Nature, and Wiley, accounted for `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(jn_cum) * 100`% of total article volume published (see Table \ref{tab:publisher_league_table}).
+Together, they published `r publisher_league_tbl  |> filter(publisher != "Other") |> pull(oa_total) |> sum()` open access articles, representing `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(oa_cum) * 100`% of the open access articles in hybrid journals.
+Elsevier, Springer Nature, and Wiley made `r publisher_league_tbl  |> filter(publisher != "Other") |> pull(ta_oa_total) |> sum()` articles open access in hybrid journals through transformative agreements, resulting in an even larger market share of `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(ta_oa_cum)*100`%.
 
@@
 
-Analysing hybrid open access across publishers between 2018 and 2022 reveals a large market concentration. 
-Although `r length(unique(jn_ind_jct_all$esac_publisher))` publishers offered transformative agreements, the big three commercial publishers Elsevier, Springer Nature, and Wiley, accounted for `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(jn_cum) * 100`% of total article volume published (see Table \ref{tab:publisher_league_table}).
-Together, they published `r publisher_league_tbl  |> filter(publisher != "Other") |> pull(oa_total) |> sum()` open access articles, representing `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(oa_cum) * 100`% of the open access articles in hybrid journals.
-Elsevier, Springer Nature, and Wiley made `r publisher_league_tbl  |> filter(publisher != "Other") |> pull(ta_oa_total) |> sum()` articles open access in hybrid journals through transformative agreements, resulting in an even larger market share of `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(ta_oa_cum)*100`%.
-However, there are differences among the three large publishers.
+However, there are notable differences among the three large publishers.
 Although Elsevier published the largest volume of articles (n = `r publisher_league_tbl  |> filter(publisher == "Elsevier") |> pull(publisher_total)`, `r publisher_league_tbl  |> filter(publisher == "Elsevier") |> pull(cum)*100`%), it published a relatively low number of open access articles, including those that can be associated with transformative agreements.
@@
 
+Figure \ref{fig:publisher_figure} takes a closer look into the growth of hybrid open access across publishers by year with a focus on open articles enabled by transformative agreements. 
+Although all publishers show a general long-term trend towards transformative agreements, Figure \ref{fig:publisher_figure}A and B indicate that, in particular, Wiley's has experienced a substantial increase in its open access share from `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(value) |> sum()`) in 2018 to `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(prop) |> sum()*100`%  (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()`) in 2022, representing an 4.5-fold increase.
+In contrast, Elsevier's hybrid journals demonstrated a more modest increase, from `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Elsevier") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Elsevier") |> pull(value) |> sum()`) in 2018 to `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(value) |> sum()`) in 2022, which is a relatively low open access share compared to the general trend.
+In 2018, Springer Nature had the largest open access proportion among the three publishers of `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Springer Nature") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Springer Nature") |> pull(value) |> sum()`), but experienced a relatively slower growth, resulting in `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature") |> pull(value) |> sum()`) of articles being open access in Springer Nature hybrid journals in 2022.
+
 ```{r publisher_figure, fig.asp=0.8, fig.cap="Developement of open access in hybrid journals included in transformative agreements between 2018 and 2022 by publishers.
-The blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established (according to matching OpenAlex first author affiliations matched with cOAlition S transformative agreement data).
+The blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established.
 (A) Proportion of open access articles in hybrid journals per year and publisher.
@@
 
-Figure \ref{fig:publisher_figure} takes a closer look into the growth of hybrid open access across publishers by year with a focus on open articles enabled by transformative agreements. 
-Although all publishers show a general long-term trend towards transformative agreements, Figure \ref{fig:publisher_figure}A and B indicate that, in particular, Wiley's has experienced a substantial increase in its open access share from `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(value) |> sum()`) in 2018 to `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(prop) |> sum()*100`%  (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()`) in 2022, representing an 4.5-fold increase.
-In contrast, Elsevier's hybrid journals demonstrated a more modest increase, from `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Elsevier") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Elsevier") |> pull(value) |> sum()`) in 2018 to `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(value) |> sum()`) in 2022, which is a relatively low open access share compared to the general trend.
-In 2018, Springer Nature had the largest open access proportion among the three publishers of `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Springer Nature") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Springer Nature") |> pull(value) |> sum()`), but experienced a relatively slower growth, resulting in `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature") |> pull(value) |> sum()`) of articles being open access in Springer Nature hybrid journals in 2022.
-
 The varying degrees of uptake of open access across the three major publishers can be attributed to distinct approaches to transformative agreements. 
@@
 Nonetheless, the results suggest the importance of central agreements for Springer Nature's hybrid open access business over the past five years (Figure 2B).
-In 2022, `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature") |> pull(value) |> sum()*100`% (n = `r  publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature", type == "ta_oa") |> pull(value)`) of open access articles in f Springer Nature  hybrid journals were enabled through transformative agreements.
+In 2022, `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature") |> pull(value) |> sum()*100`% (n = `r  publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature", type == "ta_oa") |> pull(value)`) of open access articles in Springer Nature hybrid journals were enabled through transformative agreements.
 In the same year, `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley", type == "ta_oa") |> pull(value)`) of Wiley's open access articles could be linked to transformative agreements in 2022.
@@
 The increasing trend towards transformative agreements can be also observed at the journal-level (Figure \ref{fig:publisher_figure}C).
-While no substantial differences between open access enabled through transformative agreements and other revenue source could observed across Elsevier journals, the distribution of open access across Springer Nature and Wiley hybrid journals indicates that the growth is not limited to a few journals, but extends across the portfolio.
+While no substantial differences between open access enabled through transformative agreements and other revenue sources could observed across Elsevier journals, the distribution of open access across Springer Nature and Wiley hybrid journals indicates that the growth is not limited to a few journals, but extends across the portfolio.
 In particular, Wiley's upper quantile, which represents the top 25% of journals in terms of the proportion of open access articles from transformative agreements, increased markedly from `r pub_75 |> filter(publisher == "Wiley", type == "ta_oa", cr_year == "2020") |> pull(top25) * 100`% in 2020 to `r pub_75 |> filter(publisher == "Wiley", type == "ta_oa", cr_year == "2022") |> pull(top25) * 100`% in 2022.
@@
 Table \ref{tab:subject_summary_table} presents a high-level overview of hybrid open access by AJCS subject area using fractionalised counting to account for journals belonging to more than one category.
-Between 2018 and 2022, most hybrid journals with at least one open articles could be attributed to the social sciences including the humanities.
-However, these journals published the fewest number of articles, whereas physical sciences journals recorded most articles, followed by the health sciences and the life sciences. 
-In terms of open access,  physical sciences journals accounted for more than one third of articles published in the five-years period, followed by the health science, the social sciences and the life sciences.
+Between 2018 and 2022, most hybrid journals with at least one open articles could be attributed to the category Social Sciences, which also includes the humanities.
+However, these journals published the fewest number of articles, whereas Physical Sciences journals recorded most articles, followed by the Health Sciences and the Life Sciences. 
+In terms of open access,  Physical Sciences journals accounted for more than one third of articles published in the five-years period, followed by the Health Science, the Social Sciences and the Life Sciences.
 
@@
 In 2022, two-third of open access articles (`r subject_ta_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(prop) * 100`%, n = `r subject_ta_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(ta_oa_frac)`) were published by first authors affiliated with participating institutions (see \ref{fig:subject_panel}B).
-Figure \ref{fig:subject_panel}C shows that this trend is consistent across Social Sciences journals.
-In 2022, 25% of Social Science journals provided open access to at least every fourth article exclusively through transformative agreements.
-However, hybrid open access through transformative agreements played a comparable lesser role in the Life Sciences and Health Sciences. 
-In these two subject areas, only about half of the open access articles can be linked to these agreements, both overall and on median average across journals.
-In contrast, the majority of Physical Science Journals, shows an increase of open access through transformative agreements compared to other options to publish open access in hybrid journals.
-
-```{r subject_panel, fig.asp=0.8}
+
+```{r subject_panel, fig.asp=0.8, fig.cap="Developement of open access in hybrid journals in transformative agreements between 2018 and 2022 by AJCS subject area.
+The blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established.
+(A) Proportion of open access articles in hybrid journals per year and publisher.
+(B) Percentage of hybrid open access via agreements per year and publisher.
+Boxplots (C) show individual open access uptake rates by individual hybrid journals and open access funding per publication year and publisher.
+The individual outliers are not shown.
+Note that data on transformative agreements ending before June 2021 were not available for this study."}
 library(patchwork)
@@
 ```
+
+Figure \ref{fig:subject_panel}C shows that this trend is consistent across Social Sciences journals.
+In 2022, 25% of Social Science journals provided open access to at least every fourth article exclusively through transformative agreements.
+However, hybrid open access through transformative agreements played a comparable lesser role in the Life Sciences and Health Sciences. 
+In these two subject areas, only about half of the open access articles can be linked to these agreements, both overall and on median average across journals.
+In contrast, the majority of Physical Science Journals, shows an increase of open access through transformative agreements compared to other options to publish open access in hybrid journals.
 
@@
 
-Between 2018 and 2022, Western economies almost exclusively dominated hybrid open access publishing through transformative agreements. 
+Between 2018 and 2022, high-income economies almost exclusively dominated hybrid open access publishing through transformative agreements. 
 During this period, first-authors affiliated with institutions from Organisation for Economic Co-operation and Development (OECD) member countries published `r oecd_stats_global |> filter(cat == "OECD") |> pull(oa)` open access articles in hybrid journals, representing `r oecd_stats_global |> filter(cat == "OECD") |> pull(oa) / n_oa * 100`% of the investigated open access articles.
@@
 
-```{r country_patch, fig.asp=0.9, fig.cap="Development of hybrid open access publishing by countries"}
+Figure \ref{fig:country_patch}A shows the development of hybrid open access publishing by countries, comparing the OECD area with the BRICS, an intergovernmental organisation, which comprised the countries Brazil, Russia, India, China and South Africa in 2022.
+The residual category "Other" includes the remaining countries. 
+From 2018 to 2022, the proportion of open access in hybrid journals increased from `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2018) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2018 to `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2022) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2022.
+On the other hand, BRICS recorded an low uptake, from `r country_bar$data |> filter(cat == "BRICS", oa_type != "closed", cr_year == 2018) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2018 to `r country_bar$data |> filter(cat == "BRICS", oa_type != "closed", cr_year == 2022) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2022.
+Despite rise of open access across OECD countries, the overall publication output decreased sharply, dropping to `r country_bar$data |> filter(cat == "OECD", cr_year == "2022") |> distinct(all) |> pull(all)` in 2022 after peaking `r country_bar$data |> filter(cat == "OECD", cr_year == "2020") |> distinct(all) |> pull(all)` articles in 2020.
+In stark contrast, the number of articles published in hybrid journals by first authors affiliated with institutions from BRICS countries increased steadily over the years, more than doubling from `r country_bar$data |> filter(cat == "BRICS", cr_year == "2018") |> distinct(all) |> pull(all)` in 2018 to `r country_bar$data |> filter(cat == "OECD", cr_year == "2022") |> distinct(all) |> pull(all)` in 2022.
+Upon closer examination, this trend can be observed across all big three publishers, although the shift towards BRICS is particularly evident in Elsevier's hybrid journal portfolio, in particular with regard to articles published in Physical Sciences journals.
+While OECD publication output in Elsevier's Physical Sciences journals declined from `r publ_subject_output_df |> filter(cr_year == 2018, top_level == "Physical Sciences", publisher == "Elsevier", cat == "OECD") |> pull(all)` articles in 2018 to `r publ_subject_output_df |> filter(cr_year == 2022, top_level == "Physical Sciences", publisher == "Elsevier", cat == "OECD") |> pull(all)` in 2022, BRICS output increased from `r publ_subject_output_df |> filter(cr_year == 2018, top_level == "Physical Sciences", publisher == "Elsevier", cat == "BRICS") |> pull(all)` to `r publ_subject_output_df |> filter(cr_year == 2022, top_level == "Physical Sciences", publisher == "Elsevier", cat == "BRICS") |> pull(all)` in the same five-year period.
+Furthermore, OECD publication output in Health Science Journals and Life Science journals stagnated across the investigated hybrid journal portfolios after a peak in 2020.
+
+```{r country_patch, fig.asp=0.9, fig.cap="Development of hybrid open access publishing by country.
+(A) presents the number of articles published in hybrid journals included in transformative agreement, distinguishing between BRICS as of 20222, OECD and other countries.
+(B) Scatterplot contrasting total articles with open access article volume in 2022, by country.
+Point size represents the number of articles enabled through transformative agreements.
+(C) Hybrid open access market share in 2022 by country.
+In (A) and (C), the blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established.
+Country names are represented as ISO two-letter country codes."}
 library(patchwork)
@@
 
-Figure \ref{fig:country_patch}A shows the development of hybrid open access publishing by countries, comparing the OECD area with the BRICS, an intergovernmental organisation, which comprised the countries Brazil, Russia, India, China and South Africa in 2022.
-The residual category "Other" includes the remaining countries. 
-From 2018 to 2022, the proportion of open access in hybrid journals increased from `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2018) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2018 to `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2022) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2022.
-On the other hand, BRICS recorded an low uptake, from `r country_bar$data |> filter(cat == "BRICS", oa_type != "closed", cr_year == 2018) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2018 to `r country_bar$data |> filter(cat == "BRICS", oa_type != "closed", cr_year == 2022) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2022.
-Despite rise of open access across OECD countries, the overall publication output decreased sharply, dropping to `r country_bar$data |> filter(cat == "OECD", cr_year == "2022") |> distinct(all) |> pull(all)` in 2022 after peaking `r country_bar$data |> filter(cat == "OECD", cr_year == "2020") |> distinct(all) |> pull(all)` articles in 2020.
-In stark contrast, the number of articles published in hybrid journals by first authors affiliated with institutions from BRICS countries increased steadily over the years, more than doubling from `r country_bar$data |> filter(cat == "BRICS", cr_year == "2018") |> distinct(all) |> pull(all)` in 2018 to `r country_bar$data |> filter(cat == "OECD", cr_year == "2022") |> distinct(all) |> pull(all)` in 2022.
-Upon closer examination, this trend can be observed across all big three publishers, although the shift towards BRICS is particularly evident in Elsevier's hybrid journal portfolio, in particular with regard to articles published in Physical Sciences journals (see Supplement).
-While OECD publication output in Elsevier's Physical Sciences journals declined from `r publ_subject_output_df |> filter(cr_year == 2018, top_level == "Physical Sciences", publisher == "Elsevier", cat == "OECD") |> pull(all)` articles in 2018 to `r publ_subject_output_df |> filter(cr_year == 2022, top_level == "Physical Sciences", publisher == "Elsevier", cat == "OECD") |> pull(all)` in 2022, BRICS output increased from `r publ_subject_output_df |> filter(cr_year == 2018, top_level == "Physical Sciences", publisher == "Elsevier", cat == "BRICS") |> pull(all)` to `r publ_subject_output_df |> filter(cr_year == 2022, top_level == "Physical Sciences", publisher == "Elsevier", cat == "BRICS") |> pull(all)` in the same five-year period.
-Furthermore, OECD publication output in Health Science Journals and Life Science journals stagnated across the investigated hybrid journal portfolios after a peak in 2020.
-
 To illustrate the situation in 2022, \ref{fig:country_patch}B compares total publication output with the number of open access articles. 
@@
 
-```{r country_top_20_plot, fig.asp=0.8, fig.cap="Top 20 countries"}
-country_plot
-```
-
 Figure \ref{fig:country_top_20_plot} illustrates the development of hybrid open access from 2018 to 2022, highlighting the top 20 most productive countries in terms of articles published in hybrid journals that were included in transformative agreements over the five-year period.
@@
 Prior to this, only a few organisations had agreements in place, for the example the Max Planck Society with Springer Compact.
+
+```{r country_top_20_plot, fig.asp=0.8, fig.cap="Development of open access in hybrid journals in transformative agreements between 2018 and 2022, by the Top 20 most productive countries in terms of total articles published in the five-years period.
+Blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established.
+Country names are represented as ISO two-letter country codes"}
+country_plot
+```
 
@@
 
+<!--
 # Supplements
@@
 ```{r}
-pub_subject_output_plot
-```
+#pub_subject_output_plot
+```
+
+-->
+# Discussion
+
+# Conclusions
+
+# References
 

### 2024-02-15 analysis/manuscript.tex
minor

@@
 
-Throughout this mostly automated data gathering and analysis process, we
-used tools from the Tidyverse (Wickham et al., 2019) for the R
-programming language (R Core Team, 2020). The resulting data is openly
-available through an R package, hoaddata. Following Marwick et al.
-(2018), hoaddata contains not only the datasets used in the data
-analysis. It also includes code used to compile the data by connecting
-it the cloud-based Google Big Query data warehouse, where the big
-scholarly data from Crossref, OpenAlex and Unpaywall were imported. To
-increase the computational reproducibility, the R package was
+Throughout this mostly automated data gathering and analysis process,
+tools from the Tidyverse (Wickham et al., 2019) for the R programming
+language (R Core Team, 2020) were used. The resulting data is openly
+available through an R package, hoaddata, version 0.2.91 Following
+Marwick et al. (2018), hoaddata contains not only the datasets used in
+the data analysis. It also includes code used to compile the data by
+connecting it the cloud-based Google Big Query data warehouse, where the
+big scholarly data from Crossref, OpenAlex and Unpaywall were imported.
+To increase the computational reproducibility, the R package was
 automatically built using GitHub Actions, a continuous integration
@@
 \caption{Relative growth of open access in hybrid journals in transformative agreements between 2018 and 2022 per publication year.
-The blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established (according to matching OpenAlex first author affiliations matched with cOAlition S transformative agreement data).
+The blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established.
 (A) Proportion of open access articles in hybrid journals per year.
@@
 
+Analysing hybrid open access across publishers between 2018 and 2022
+reveals a large market concentration. Although 48 publishers offered
+transformative agreements, the big three commercial publishers Elsevier,
+Springer Nature, and Wiley, accounted for 49\% of total article volume
+published (see Table \ref{tab:publisher_league_table}). Together, they
+published 500,878 open access articles, representing 66\% of the open
+access articles in hybrid journals. Elsevier, Springer Nature, and Wiley
+made 243,891 articles open access in hybrid journals through
+transformative agreements, resulting in an even larger market share of
+74\%.
+
 \begin{table}[H]
@@
 
-Analysing hybrid open access across publishers between 2018 and 2022
-reveals a large market concentration. Although 48 publishers offered
-transformative agreements, the big three commercial publishers Elsevier,
-Springer Nature, and Wiley, accounted for 49\% of total article volume
-published (see Table \ref{tab:publisher_league_table}). Together, they
-published 500,878 open access articles, representing 66\% of the open
-access articles in hybrid journals. Elsevier, Springer Nature, and Wiley
-made 243,891 articles open access in hybrid journals through
-transformative agreements, resulting in an even larger market share of
-74\%. However, there are differences among the three large publishers.
+However, there are notable differences among the three large publishers.
 Although Elsevier published the largest volume of articles (n =
@@
 55\%).
-
-\begin{figure}[ht!]
-
-{\centering \includegraphics[width=0.99\linewidth,]{fig/publisher_figure-1}
-
-}
-
-\caption{Developement of open access in hybrid journals included in transformative agreements between 2018 and 2022 by publishers.
-The blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established (according to matching OpenAlex first author affiliations matched with cOAlition S transformative agreement data).
-(A) Proportion of open access articles in hybrid journals per year and publisher.
-(B) Percentage of hybrid open access via agreements per year and publisher.
-Boxplots (C) show individual open access uptake rates by individual hybrid journals and open access funding per publication year and publisher.
-The individual outliers are not shown.
-Note that data on transformative agreements ending before June 2021 were not available for this study.}\label{fig:publisher_figure}
-\end{figure}
 
@@
 
+\begin{figure}[ht!]
+
+{\centering \includegraphics[width=0.99\linewidth,]{fig/publisher_figure-1}
+
+}
+
+\caption{Developement of open access in hybrid journals included in transformative agreements between 2018 and 2022 by publishers.
+The blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established.
+(A) Proportion of open access articles in hybrid journals per year and publisher.
+(B) Percentage of hybrid open access via agreements per year and publisher.
+Boxplots (C) show individual open access uptake rates by individual hybrid journals and open access funding per publication year and publisher.
+The individual outliers are not shown.
+Note that data on transformative agreements ending before June 2021 were not available for this study.}\label{fig:publisher_figure}
+\end{figure}
+
 The varying degrees of uptake of open access across the three major
@@
 over the past five years (Figure 2B). In 2022, 66\% (n = 34,725) of open
-access articles in f Springer Nature hybrid journals were enabled
-through transformative agreements. In the same year, 70\% (n = 37,316)
-of Wiley's open access articles could be linked to transformative
+access articles in Springer Nature hybrid journals were enabled through
+transformative agreements. In the same year, 70\% (n = 37,316) of
+Wiley's open access articles could be linked to transformative
 agreements in 2022. In contrast, Elsevier published fewer than half of
@@
 While no substantial differences between open access enabled through
-transformative agreements and other revenue source could observed across
-Elsevier journals, the distribution of open access across Springer
-Nature and Wiley hybrid journals indicates that the growth is not
-limited to a few journals, but extends across the portfolio. In
+transformative agreements and other revenue sources could observed
+across Elsevier journals, the distribution of open access across
+Springer Nature and Wiley hybrid journals indicates that the growth is
+not limited to a few journals, but extends across the portfolio. In
 particular, Wiley's upper quantile, which represents the top 25\% of
@@
 and 2022, most hybrid journals with at least one open articles could be
-attributed to the social sciences including the humanities. However,
-these journals published the fewest number of articles, whereas physical
-sciences journals recorded most articles, followed by the health
-sciences and the life sciences. In terms of open access, physical
-sciences journals accounted for more than one third of articles
-published in the five-years period, followed by the health science, the
-social sciences and the life sciences.
+attributed to the category Social Sciences, which also includes the
+humanities. However, these journals published the fewest number of
+articles, whereas Physical Sciences journals recorded most articles,
+followed by the Health Sciences and the Life Sciences. In terms of open
+access, Physical Sciences journals accounted for more than one third of
+articles published in the five-years period, followed by the Health
+Science, the Social Sciences and the Life Sciences.
 
@@
 (67\%, n = 34,759) were published by first authors affiliated with
-participating institutions (see \ref{fig:subject_panel}B). Figure
-\ref{fig:subject_panel}C shows that this trend is consistent across
-Social Sciences journals. In 2022, 25\% of Social Science journals
-provided open access to at least every fourth article exclusively
-through transformative agreements. However, hybrid open access through
-transformative agreements played a comparable lesser role in the Life
-Sciences and Health Sciences. In these two subject areas, only about
-half of the open access articles can be linked to these agreements, both
-overall and on median average across journals. In contrast, the majority
-of Physical Science Journals, shows an increase of open access through
-transformative agreements compared to other options to publish open
-access in hybrid journals.
-
-\begin{center}\includegraphics[width=0.99\linewidth,]{fig/subject_panel-1} \end{center}
+participating institutions (see \ref{fig:subject_panel}B).
+
+\begin{figure}[ht!]
+
+{\centering \includegraphics[width=0.99\linewidth,]{fig/subject_panel-1}
+
+}
+
+\caption{Developement of open access in hybrid journals in transformative agreements between 2018 and 2022 by AJCS subject area.
+The blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established.
+(A) Proportion of open access articles in hybrid journals per year and publisher.
+(B) Percentage of hybrid open access via agreements per year and publisher.
+Boxplots (C) show individual open access uptake rates by individual hybrid journals and open access funding per publication year and publisher.
+The individual outliers are not shown.
+Note that data on transformative agreements ending before June 2021 were not available for this study.}\label{fig:subject_panel}
+\end{figure}
+
+Figure \ref{fig:subject_panel}C shows that this trend is consistent
+across Social Sciences journals. In 2022, 25\% of Social Science
+journals provided open access to at least every fourth article
+exclusively through transformative agreements. However, hybrid open
+access through transformative agreements played a comparable lesser role
+in the Life Sciences and Health Sciences. In these two subject areas,
+only about half of the open access articles can be linked to these
+agreements, both overall and on median average across journals. In
+contrast, the majority of Physical Science Journals, shows an increase
+of open access through transformative agreements compared to other
+options to publish open access in hybrid journals.
 
@@
 
-Between 2018 and 2022, Western economies almost exclusively dominated
-hybrid open access publishing through transformative agreements. During
-this period, first-authors affiliated with institutions from
-Organisation for Economic Co-operation and Development (OECD) member
-countries published 602,050 open access articles in hybrid journals,
-representing 81\% of the investigated open access articles. This
-disparity between OECD nations and other countries becomes even more
-evident when considering open access through transformative agreements,
-as 310,712 of 328,957, or 94\% of open access articles were associated
-with such agreements.
-
-\begin{figure}[ht!]
-
-{\centering \includegraphics[width=0.99\linewidth,]{fig/country_patch-1}
-
-}
-
-\caption{Development of hybrid open access publishing by countries}\label{fig:country_patch}
-\end{figure}
+Between 2018 and 2022, high-income economies almost exclusively
+dominated hybrid open access publishing through transformative
+agreements. During this period, first-authors affiliated with
+institutions from Organisation for Economic Co-operation and Development
+(OECD) member countries published 602,050 open access articles in hybrid
+journals, representing 81\% of the investigated open access articles.
+This disparity between OECD nations and other countries becomes even
+more evident when considering open access through transformative
+agreements, as 310,712 of 328,957, or 94\% of open access articles were
+associated with such agreements.
 
@@
 hybrid journal portfolio, in particular with regard to articles
-published in Physical Sciences journals (see Supplement). While OECD
-publication output in Elsevier's Physical Sciences journals declined
-from 112,822 articles in 2018 to 103,766 in 2022, BRICS output increased
-from 104,654 to 171,713 in the same five-year period. Furthermore, OECD
-publication output in Health Science Journals and Life Science journals
-stagnated across the investigated hybrid journal portfolios after a peak
-in 2020.
+published in Physical Sciences journals. While OECD publication output
+in Elsevier's Physical Sciences journals declined from 112,822 articles
+in 2018 to 103,766 in 2022, BRICS output increased from 104,654 to
+171,713 in the same five-year period. Furthermore, OECD publication
+output in Health Science Journals and Life Science journals stagnated
+across the investigated hybrid journal portfolios after a peak in 2020.
+
+\begin{figure}[ht!]
+
+{\centering \includegraphics[width=0.99\linewidth,]{fig/country_patch-1}
+
+}
+
+\caption{Development of hybrid open access publishing by country.
+(A) presents the number of articles published in hybrid journals included in transformative agreement, distinguishing between BRICS as of 20222, OECD and other countries.
+(B) Scatterplot contrasting total articles with open access article volume in 2022, by country.
+Point size represents the number of articles enabled through transformative agreements.
+(C) Hybrid open access market share in 2022 by country.
+In (A) and (C), the blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established.
+Country names are represented as ISO two-letter country codes.}\label{fig:country_patch}
+\end{figure}
 
@@
 
-\begin{figure}[ht!]
-
-{\centering \includegraphics[width=0.99\linewidth,]{fig/country_top_20_plot-1}
-
-}
-
-\caption{Top 20 countries}\label{fig:country_top_20_plot}
-\end{figure}
-
 Figure \ref{fig:country_top_20_plot} illustrates the development of
@@
 
+\begin{figure}[ht!]
+
+{\centering \includegraphics[width=0.99\linewidth,]{fig/country_top_20_plot-1}
+
+}
+
+\caption{Development of open access in hybrid journals in transformative agreements between 2018 and 2022, by the Top 20 most productive countries in terms of total articles published in the five-years period.
+Blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established.
+Country names are represented as ISO two-letter country codes}\label{fig:country_top_20_plot}
+\end{figure}
+
 Since 2021, there has been a general trend towards hybrid open access
@@
 
-\hypertarget{supplements}{%
-\section{Supplements}\label{supplements}}
-
-\begin{center}\includegraphics[width=0.99\linewidth,]{fig/unnamed-chunk-7-1} \end{center}
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+\hypertarget{conclusions}{%
+\section{Conclusions}\label{conclusions}}
+
+\hypertarget{references}{%
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
 

### 2024-02-15 analysis/manuscript.tex
re-render

@@
 some articles openly available while others remain behind a paywall, has
-been discussed as a model for transitioning the subscription system to
+been discussed as a means for transitioning the subscription system to
 full open access (Prosser, 2003). The idea was that when journals

### 2024-02-19 analysis/manuscript.Rmd
starting discussions

@@
 In 2009, Springer reported open access to 1% of articles in hybrid journals [@Dallmeier_2010].
-Other publishers also recorded a low uptake.
+Other studies also recorded a low uptake.
 In 2011, only 1-2% of articles were open access [@Bj_rk_2012], growing to around 4% between 2011 and 2013 [@Laakso_2016].
 
-With the introduction of central funding mechanisms for hybrid open access publication fees in some European countries since 2012, an increase in open access could be observed [@Bj_rk_2017; @Piwowar_2018; @Jubb_2017; @Huang_2020].
+With the introduction of central funding mechanisms for publication fees in some European countries since 2012, an increase in hybrid open access could be observed [@Bj_rk_2017; @Piwowar_2018; @Jubb_2017; @Huang_2020].
 For example, studying university output, @Robinson_Garcia_2020 estimated a median uptake of 7.1% in the period 2014-2017. 
@@
 
-As a consequence, libraries and their consortia began to develop licensing strategies aimed at avoiding such ‘double dipping’ scenarios, in which well-established commercial publishers gain twice from reading and publication fees, as well as to increase publisher-provided immediate open access [@BJ_RK_2014; @Schimmer_2015].
-These considerations resulted into transformative agreements, which cover a broad range of successfully negotiated contracts between library consortia and publishers from mid-2010s onwards where institutional spending for subscriptions are re-purposed for open access publishing [@Borrego_2020; @Hinchliffe_2019].
+As a consequence, libraries and their consortia began to develop licensing strategies aimed at avoiding such ‘double dipping’ scenarios, in which well-established commercial publishers gain twice from reading and open access publishing, as well as to increase publisher-provided immediate open access [@BJ_RK_2014; @Schimmer_2015].
+These considerations resulted into transformative agreements, which cover a broad range of contracts between library consortia and publishers from mid-2010s onwards where institutional spending for subscriptions and open access publishing are considered together [@Borrego_2020; @Hinchliffe_2019].
 Transformative agreements seek to control costs while allowing a transitional phase for publishing more open access articles. 
-Although transformative agreements mainly bundle hybrid and subscription-only journals from commercial publishers, they aim at a higher degree of transparency than previous big deals, where contracts including payments were confidential [@Bergstrom_2014].
+Similar to big deals, transformative agreements mainly bundle hybrid and subscription-only journals from commercial publishers, but aim at a higher degree of transparency than previous big deals, where contracts including payments were confidential [@Bergstrom_2014].
 <!-- For example, the nationwide DEAL consortium in Germany, which includes universities and academic institutions, collectively cancelled their subscriptions with Elsevier from 2017 onwards to negotiate a transformative agreement [@Fraser_2023].
@@
 According ot its Plan S launched in 2018, the cOAlition S funders only accept hybrid open access through transformative agreements "during a transition period that should be as short as possible" [@Schiltz_2018].
-Specifically, they agreed to support hybrid open access solely through transformative agreements from 1 January 2021, until the end of 2024.
+Specifically, they agreed to support hybrid open access only through transformative agreements from 2021, until the end of 2024.
 Notably, the German Research Foundation (DFG), despite not being part of cOAlition S, has also extended its financial support for hybrid open access through transformative agreements [@Mittermaier_2021].
@@
 
-By the end of 2023, many transformative agreements were implemented, but outcomes were mixed.
+By the end of 2023, many transformative agreements were implemented, but interim outcomes are mixed.
 The ESAC Transformative Agreement Registry[^esacregistry], the largest source of disclosure, recorded more than 800 transformative agreements, resulting in up to 900.000 open access articles published in both full open access and hybrid journals according to the accompanying ESAC Market Watch[^esacmarketwatch].
 Library consortia reported increased open access volume, streamlined payment and monitoring procedures, as well as extensive utilization of open access options by the researchers they serve [@Marques_2020; @Pinhasi_2020; @Parmhed_2023].
-The ongoing standardisation of transformative agreements contributed to improved transparency in terms of contracts and publisher-provided article metadata [@Marques_2019].
-However, with the growing trend toward transformative agreements, continued reliance on big deals is perceived as problematic, which might perpetuate market concentration [@Butler_2023; @Shu_2023]. 
+The ongoing standardisation of transformative agreements contributed to improved transparency in terms of contracts and publisher-provided article metadata [@Marques_2019; @Pinhasi_2021].
+However, with the growing trend toward transformative agreements, continued reliance on big deals is perceived as problematic, because it perpetuate market concentration [@Butler_2023; @Shu_2023]. 
 Whether transformative agreements lead to reduced pricing remains uncertain [@Borrego_2023] and
 a substantial transition of hybrid journals towards full open access could not be observed [@Momeni_2021; @Matthias_2019].
-Furthermore, the focus on large commercial publishers might increase inequality, because transformative agreements continue to target institutions from high-income countries [@Klebel_2023]. 
+The focus on large commercial publishers might also increase inequality, because transformative agreements focus on pay to publish open access mainly targets institutions from high-income countries, furthering a questionable journal prestige culture [@budapest].
 Besides, an editorial-board resignation raised concerns that publishers' desire to maximize journal publication volume "without regard to quality" is a consequence of transformative agreements [@Goodin_2023].
@@
 högskoleförbund, SUHF) recommended to only support agreements for publishing in full open access journals [^sweden].
-Likewise, most cOAlition S funders will discontinue financial support for transformative agreements from 2024 onwards [@Liverpool_2023].
-The consortium also removed most hybrid journals from its Transformative Journal program in 2023 due to their failure to meet self-defined open access growth targets [@Brainard_2023]. 
+Likewise, most cOAlition S funders will discontinue financial support for transformative agreements by the end of 2024 [@Liverpool_2023].
+The consortium also removed most hybrid journals from its Transformative Journal program in 2023 due to publishers' failure to meet self-defined open access growth targets [@Brainard_2023]. 
 In contrast, the German DEAL consortium signed a five-year transformative agreement with Elsevier starting in 2024, while also renewing its contracts with Springer Nature and Wiley until the end of 2028.
@@
 Despite these controversies around transformative agreements as a means of transitioning journal publishing to full open access, there is limited evidence available on the adoption of open access in hybrid journals, and the extent to which this uptake can be attributed to transformative agreements.
-Previous studies have focused on specific countries [@P_l_nen_2020; @Huang_2020; @Haucap_2021; @Wenaas_2022; @Taubert_2023] or publisher portfolios [@Pieper_2018; @Bakker_2024; @Fraser_2023; @Jahn_2021; @Momeni_2023], while related large-scale studies relied on self-reported agreement data [@Moskovkin_2022], or used APC pricing lists [@Shu_2023].
+Previous studies have focused on specific countries [@P_l_nen_2020; @Huang_2020; @Haucap_2021; @Wenaas_2022; @Taubert_2023] or publisher portfolios [@Pieper_2018; @Bakker_2024; @Fraser_2023; @Jahn_2021; @Momeni_2023], while large-scale studies relied on self-reported agreement data [@Moskovkin_2022], or used APC pricing lists [@Shu_2023].
 Particularly, data availability is a limiting factor when studying the impact of transformative agreements [@Bakker_2024], because bibliometric databases, even though many allow the retrieval of open access articles in hybrid journals, do not directly attribute them to specific transformative agreements.
@@
 The present study aims to address these limitations by combining multiple openly available data sources to determine the open access uptake in hybrid journals, while distinguishing between open access through transformative agreements and other means.
-With this novel and open approach, this study aims to answer the following questions:
-
-- What was the number and proportion of open access articles in hybrid journals included in transformative agreements between 2018 and 2022?
-- How did transformative agreements impact open access uptake in hybrid journals?
+With this novel and open approach, this the first large-scale analysis aims to answer the following questions:
+
+- What was the number and proportion of open access articles in hybrid journals in transformative agreements between 2018 and 2022?
+- To what extent did institutions with a transformation agreement contribute to the adoption of open access in hybrid journals?
 
@@
 
-# Methods and data
+# Methods
 
 This study combines data from multiple publicly available data sources as diagrammatically shown in Figure \ref{fig:data_workflow}. 
-Initially, transformative agreement data retrieved from the cOAlition S Journal Checker Tool provided information about journal portfolios and participating institutions. 
+Initially, transformative agreement data retrieved from the cOAlition S Journal Checker Tool[^checkertool]  provided information about journal portfolios and participating institutions. 
 After identification of hybrid journals by excluding full open access journals, Crossref served as the primary data source for article-level metadata including Creative Commons (CC) license information to indicate open access availability on publisher websites.
 To determine open access articles published through transformative agreements, first author affiliations from OpenAlex [@priem2022openalex] were subsequently linked to eligible institutions according to the transformative agreement data.
+In the following, these steps are described in more detail.
 
@@
 Because transformative agreements can include both full open access and hybrid journals, the data were complemented with information about a journal's open access status using multiple sources: the Directory of Open Access Journals (DOAJ) downloaded on 12 December 2023[^doaj], OpenAlex (November 2023) and the the Bielefeld list of GOLD OA journals [@goldoa].
-As shown in Figure \ref{fig:method_fig}A, combining different data sources considerably extended the journal matching to exclude full open access journals from transformative agreements.
+As shown in Figure \ref{fig:method_fig}A, combining different data sources considerably extended the journal matching<!-- to exclude full open access journals from transformative agreements-->.
 In total, `r n_oa_journals` full open access journals were excluded based on ISSN matching.
@@
 `r hoaddata::cr_upw |> filter(cr_year %in% 2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, upw_hybrid_total) |> pull(upw_hybrid_total) |> sum()` in OpenAlex).
-Notable difference could be furthermore observed among some publishers that presumably did not provide CC license information to Crossref including AIP Publishing, American Physiological Society and Emerald.
+Notable difference could be furthermore observed among some publishers that presumably did not provide CC license information to Crossref including AIP Publishing, American Physiological Society, Emerald and the Royal Society.
 Crossref license metadata was more complete with regard to articles from the publisher Wiley and American Chemical Society.
 Finally, inconsistent open access status information in previous OpenAlex versions was observed [@jahn2023].
-According to the OpenAlex release notes, fixing this issue was still ongoing, which might also explain this discrepancy.
+After reporting it to OpenAlex, according to the release notes, fixing this issue was still ongoing, which might also explain part of the discrepancy.
 
@@
 However, because of low coverage in OpenAlex, this study focused on first authors and their affiliations instead.
-First authors typically contribute most to a paper and are often considered lead author research papers [@Larivi_re_2016], and can be therefore assumed as a proxy to measure to open access payments [@Zhang_2022; @Shu_2023].
+First authors typically contribute most to a paper and are often considered lead author research papers [@Larivi_re_2016]. 
+Related studies therefore assumed first authors as a proxy to measure to open access payments and the impact of transformative agreement [@Zhang_2022; @Shu_2023; @Haucap_2021].
 Overall, around 90% of studied articles had first author affiliation metadata in OpenAlex, whereas the coverage of articles with corresponding author information was around 54%.
@@
 Upon inspection, Transformative Agreement Data did not cover associated institutions comprehensively like university hospitals or institutes of large research organisations like the Max Planck Society.
-To improve the matching, Transformative Agreement Data was automatically enriched with associated organisations using OpenAlex's institution entity.
+To improve the matching, Transformative Agreement Data was automatically enriched with associated organisations using OpenAlex's institution entity data.
 
 In total, the compiled data set consists of `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l, cr_year, jn_all) |> pull(jn_all) |> sum()` articles published in `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l) |> nrow()` hybrid journals between 2018 and 2022 (see Figure \ref{fig:method_fig}B).
-Hybrid journals included in transformative agreements represented `r cr_stats |> filter(src == "hoad") |> pull(n_articles) |> sum() / cr_stats |> pull(n_articles) |> sum() * 100`%  of total global output over the same time period according to Crossref, while full open access journals accounted for
-`r cr_stats |> filter(is.na(src) & oa_status == "gold") |> pull(n_articles) |> sum() / cr_stats |> pull(n_articles) |> sum() * 100`% of total article volume.
+Hybrid journals in transformative agreements represented `r cr_stats |> filter(src == "hoad") |> pull(n_articles) |> sum() / cr_stats |> pull(n_articles) |> sum() * 100`%  of total global output over the same time period according to Crossref, while full open access journals recorded
+`r cr_stats |> filter(is.na(src) & oa_status == "gold") |> pull(n_articles) |> sum() / cr_stats |> pull(n_articles) |> sum() * 100`%.
 
@@
 Throughout this mostly automated data gathering and analysis process, tools from the Tidyverse (Wickham et al., 2019) for the R programming language (R Core Team, 2020) were used. 
-The resulting data is openly available through an R package, hoaddata, version 0.2.91 
+The resulting data is openly available through an R data package, hoaddata, version 0.2.91 
 Following @Marwick_2018, hoaddata contains not only the datasets used in the data analysis. 
 It also includes code used to compile the data by connecting it the cloud-based Google Big Query data warehouse, where the big scholarly data from Crossref, OpenAlex and Unpaywall were imported.
-To increase the computational reproducibility, the R package was automatically built using GitHub Actions, a continuous integration service.
+To increase the computational reproducibility, the data aggregation through hoaddata was automatically carried out using GitHub Actions, a continuous integration service.
 
@@
 
-Figure \ref{fig:results_overview}B highlights that the majority of open access articles in hybrid journals were made available through transformative agreements in 2021 and 2022, contributing `r oa_ta_plot_df |> filter(cr_year == "2022", type == "ta_oa") |> pull(value) / oa_by_year |> filter(cr_year == "2022") |> pull(oa) * 100`% of the total open access article volume in 2022.
-However, there was also a notable growth in open access provision through individual publication fees, which increased from `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> pull(value)`) in 2018 to  `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> pull(value)`).
-This suggests that publishers were able to gain equally from individual and institutional open access publishing options.
+Figure \ref{fig:results_overview}B highlights that the majority of hybrid open access was made available through transformative agreements in 2021 and 2022, contributing `r oa_ta_plot_df |> filter(cr_year == "2022", type == "ta_oa") |> pull(value) / oa_by_year |> filter(cr_year == "2022") |> pull(oa) * 100`% of the total open access article volume in 2022.
+However, there was also a notable growth in open access provision through other means, most likely publication fees being not invoiced through transformative agreements, which increased from `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> pull(value)`) in 2018 to  `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> pull(value)`).
+<!-- This suggests that publishers were able to gain equally from individual and institutional open access publishing options.-->
 
@@
 These journals were, on average, smaller (M = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> pull(n) |> mean()`, SD = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> pull(n) |> sd()`) than those with an open access share below 50% (M = `r oa_jn_box_df |> filter(cr_year == "2022", prop <= 0.5) |> pull(n) |> mean()`, SD = `r oa_jn_box_df |> filter(cr_year == "2022", prop <= 0.5) |> pull(n) |> sd()`).
-Notable exception of large journals with an above-average open access proportion were Physical Review D, a high-energy physics journal covered by the SCOAP3 consortium that provided open access to `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "2470-0010") |> pull(oa_all)` out of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "2470-0010") |> pull(n)` articles in 2022, Astronomy and Physics (`r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0004-6361") |> pull(oa_all)` out of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0004-6361") |> pull(n)` articles in 2022 were open access), which shifted to a subscribe to open business model for all accepted articles as of April 2022, and the Journal of Fluid Mechanics (`r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0022-1120") |> pull(oa_all)` of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0022-1120") |> pull(n)` articles in 2022 were open access).
+<!-- Notable exception of large journals with an above-average open access proportion were Physical Review D, a high-energy physics journal covered by the SCOAP3 consortium that provided open access to `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "2470-0010") |> pull(oa_all)` out of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "2470-0010") |> pull(n)` articles in 2022, Astronomy and Physics (`r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0004-6361") |> pull(oa_all)` out of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0004-6361") |> pull(n)` articles in 2022 were open access), which shifted to a subscribe to open business model for all accepted articles as of April 2022, and the Journal of Fluid Mechanics (`r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0022-1120") |> pull(oa_all)` of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0022-1120") |> pull(n)` articles in 2022 were open access).
+-->
 
 When comparing the impact of open access trough transformative agreements across journals, it shows that for many journals these agreements substantially contributed to the growth of open access over the years (Figure \ref{fig:results_overview}D).
-Examples of such journals include those with a scope on specific countries or regions, where also transformative agreements were implemented. 
+<!-- Examples of such journals include those with a scope on specific countries or regions, where also transformative agreements were implemented. 
 For instance, in 2022, the Germany-based journals Zeitschrift für Erziehungswissenschaft and Zeitschrift für Politikwissenschaft, as well as the Scandinavian Political Studies adressing the Nordic countries, achieved an overall open access uptake of more than 90% just through transformative agreements.
-Despite the rise in transformative agreements, it is worth noting that other means of publishing open access in hybrid journals remained common. 
+-->
+Despite the rise in transformative agreements, it is worth noting that other means of publishing open access remained common across the investiagted hybrid journals. 
 In total, `r summary_all |> filter(cr_year == "2022", value > 0, name == "prop_no_ta") |> nrow()` journals published open access articles from authors affiliated with institutions without transformative agreements in place, while `r summary_all |> filter(cr_year == "2022", value > 0, name == "prop_ta_oa") |> nrow()` journals published at least one open access article through a transformative agreement in the same year.
@@
 Analysing hybrid open access across publishers between 2018 and 2022 reveals a large market concentration. 
-Although `r length(unique(jn_ind_jct_all$esac_publisher))` publishers offered transformative agreements, the big three commercial publishers Elsevier, Springer Nature, and Wiley, accounted for `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(jn_cum) * 100`% of total article volume published (see Table \ref{tab:publisher_league_table}).
-Together, they published `r publisher_league_tbl  |> filter(publisher != "Other") |> pull(oa_total) |> sum()` open access articles, representing `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(oa_cum) * 100`% of the open access articles in hybrid journals.
+Although `r length(unique(jn_ind_jct_all$esac_publisher))` publishers offered transformative agreements, the big three commercial publishers Elsevier, Springer Nature, and Wiley accounted for `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(jn_cum) * 100`% of total article volume published (see Table \ref{tab:publisher_league_table}).
+Together, they published `r publisher_league_tbl  |> filter(publisher != "Other") |> pull(oa_total) |> sum()` or `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(oa_cum) * 100`% of open access articles in hybrid journals.
 Elsevier, Springer Nature, and Wiley made `r publisher_league_tbl  |> filter(publisher != "Other") |> pull(ta_oa_total) |> sum()` articles open access in hybrid journals through transformative agreements, resulting in an even larger market share of `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(ta_oa_cum)*100`%.
@@
 
-However, there are notable differences among the three large publishers.
-Although Elsevier published the largest volume of articles (n = `r publisher_league_tbl  |> filter(publisher == "Elsevier") |> pull(publisher_total)`, `r publisher_league_tbl  |> filter(publisher == "Elsevier") |> pull(cum)*100`%), it published a relatively low number of open access articles, including those that can be associated with transformative agreements.
+However, there are notable differences among the three big publishers.
+Although Elsevier published the largest volume of articles (n = `r publisher_league_tbl  |> filter(publisher == "Elsevier") |> pull(publisher_total)`, `r publisher_league_tbl  |> filter(publisher == "Elsevier") |> pull(cum)*100`%), it recorded a comparable low number of open access articles, including those that can be associated with transformative agreements.
 In contrast, Springer Nature and Wiley provided open access to a larger proportion of their articles (`r (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_total)) / (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(publisher_total)) * 100`% of Springer Nature articles and `r (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(oa_total)) / (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(publisher_total)) * 100`% of Wiley articles were open acccess), leading to higher open access market shares (`r publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_perc)*100`% Springer Nature resp.
@@
 Figure \ref{fig:publisher_figure} takes a closer look into the growth of hybrid open access across publishers by year with a focus on open articles enabled by transformative agreements. 
-Although all publishers show a general long-term trend towards transformative agreements, Figure \ref{fig:publisher_figure}A and B indicate that, in particular, Wiley's has experienced a substantial increase in its open access share from `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(value) |> sum()`) in 2018 to `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(prop) |> sum()*100`%  (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()`) in 2022, representing an 4.5-fold increase.
+Although all publishers show a general long-term trend towards transformative agreements, Figure \ref{fig:publisher_figure}A and B indicate that, in particular, Wiley experienced a substantial increase in its open access share from `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(value) |> sum()`) in 2018 to `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(prop) |> sum()*100`%  (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()`) in 2022, representing an 4.5-fold increase.
 In contrast, Elsevier's hybrid journals demonstrated a more modest increase, from `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Elsevier") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Elsevier") |> pull(value) |> sum()`) in 2018 to `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(value) |> sum()`) in 2022, which is a relatively low open access share compared to the general trend.
@@
 
-The varying degrees of uptake of open access across the three major publishers can be attributed to distinct approaches to transformative agreements. 
+The varying degrees of adoption of open access across the three major publishers can be attributed to distinct approaches to transformative agreements. 
 Springer Nature, for example, began in 2015 offering selected consortia, such as the Max Planck Society, the Swedish Bibsam consortium, and the Finnish FinELib consortium, open access agreements for its hybrid journal portfolio under the name Springer Compact[^springercompact].
 However, these agreements were not included in the data as they concluded prior to the start of the transformative agreement data collection in June 2021. 
-Nonetheless, the results suggest the importance of central agreements for Springer Nature's hybrid open access business over the past five years (Figure 2B).
-In 2022, `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature") |> pull(value) |> sum()*100`% (n = `r  publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature", type == "ta_oa") |> pull(value)`) of open access articles in Springer Nature hybrid journals were enabled through transformative agreements.
+Nonetheless, the results demonstrate the importance of agreements for Springer Nature's hybrid open access business over the past five years (Figure 2B).
+In 2022, `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature") |> pull(value) |> sum()*100`% (n = `r  publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature", type == "ta_oa") |> pull(value)`) of open access in Springer Nature hybrid journals were enabled through transformative agreements.
 In the same year, `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley", type == "ta_oa") |> pull(value)`) of Wiley's open access articles could be linked to transformative agreements in 2022.
@@
 The increasing trend towards transformative agreements can be also observed at the journal-level (Figure \ref{fig:publisher_figure}C).
-While no substantial differences between open access enabled through transformative agreements and other revenue sources could observed across Elsevier journals, the distribution of open access across Springer Nature and Wiley hybrid journals indicates that the growth is not limited to a few journals, but extends across the portfolio.
+While no substantial differences between open access enabled through transformative agreements and other revenue sources could observed across Elsevier's portfolio, the distribution of open access across Springer Nature and Wiley hybrid journals indicates that the growth is not limited to a few journals, but extends across the portfolio.
 In particular, Wiley's upper quantile, which represents the top 25% of journals in terms of the proportion of open access articles from transformative agreements, increased markedly from `r pub_75 |> filter(publisher == "Wiley", type == "ta_oa", cr_year == "2020") |> pull(top25) * 100`% in 2020 to `r pub_75 |> filter(publisher == "Wiley", type == "ta_oa", cr_year == "2022") |> pull(top25) * 100`% in 2022.
@@
 Table \ref{tab:subject_summary_table} presents a high-level overview of hybrid open access by AJCS subject area using fractionalised counting to account for journals belonging to more than one category.
-Between 2018 and 2022, most hybrid journals with at least one open articles could be attributed to the category Social Sciences, which also includes the humanities.
+Between 2018 and 2022, most hybrid journals with at least one open articles could be attributed to the category Social Sciences, which also includes the Arts and Humanities.
 However, these journals published the fewest number of articles, whereas Physical Sciences journals recorded most articles, followed by the Health Sciences and the Life Sciences. 
@@
 <!-- In particular, journals in the subject areas recorded large high rate (see supplement)-->
-This growth in the social sciences can be largely attributed to transformative agreements.
+Growth in the Social Sciences can be largely attributed to transformative agreements. 
 In 2022, two-third of open access articles (`r subject_ta_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(prop) * 100`%, n = `r subject_ta_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(ta_oa_frac)`) were published by first authors affiliated with participating institutions (see \ref{fig:subject_panel}B).
@@
 
-Figure \ref{fig:subject_panel}C shows that this trend is consistent across Social Sciences journals.
-In 2022, 25% of Social Science journals provided open access to at least every fourth article exclusively through transformative agreements.
+Figure \ref{fig:subject_panel}C shows that this trend is consistent across hybrid journals belonging to the ASJC subject area Social Sciences.
+In 2022, 25% of Social Sciences journals provided open access to at least every fourth article exclusively through transformative agreements.
 However, hybrid open access through transformative agreements played a comparable lesser role in the Life Sciences and Health Sciences. 
@@
     facet_labels = glue::glue(
-      '**{country_name_fct}**<br>
-      {total_string} | {paste(round(prop * 100, 1), "%")}'
+      '{country_name_fct}<br>
+      {total_string} | {paste0(round(prop * 100, 1), "%")}'
     )
@@
 
-Between 2018 and 2022, high-income economies almost exclusively dominated hybrid open access publishing through transformative agreements. 
+Between 2018 and 2022, high-income countries almost exclusively dominated hybrid open access publishing through transformative agreements. 
 During this period, first-authors affiliated with institutions from Organisation for Economic Co-operation and Development (OECD) member countries published `r oecd_stats_global |> filter(cat == "OECD") |> pull(oa)` open access articles in hybrid journals, representing `r oecd_stats_global |> filter(cat == "OECD") |> pull(oa) / n_oa * 100`% of the investigated open access articles.
@@
 
-Figure \ref{fig:country_patch}A shows the development of hybrid open access publishing by countries, comparing the OECD area with the BRICS, an intergovernmental organisation, which comprised the countries Brazil, Russia, India, China and South Africa in 2022.
+Figure \ref{fig:country_patch}A shows the development of hybrid open access publishing by countries, comparing the OECD area with the BRICS, an intergovernmental organisation, which comprised the countries Brazil, Russia, India, China and South Africa as of 2022.
 The residual category "Other" includes the remaining countries. 
@@
 
-To illustrate the situation in 2022, \ref{fig:country_patch}B compares total publication output with the number of open access articles. 
+To illustrate the situation in 2022, Figure \ref{fig:country_patch}B compares total publication output with the number of open access articles. 
 With `r country_scatter_df |> filter(cr_year == "2022", country_code == "CN") |> pull(all)` articles, China was the most productive country, followed by the United States (`r country_scatter_df |> filter(cr_year == "2022", country_code == "US") |> pull(all)` articles) and India (`r country_scatter_df |> filter(cr_year == "2022", country_code == "IN") |> pull(all)` articles).
@@
 Particularly, Germany, Great Britain, the Netherlands, Sweden, Switzerland and Spain recorded an above-average open access share as indicated by the linear trend line.
-As represented by the point size, as well as it can been seen in Figure \ref{fig:country_patch}C, transformative agreements contributed to this market position of these countries. 
+As represented by the point size, as well as it can been seen in Figure \ref{fig:country_patch}C, transformative agreements contributed to these market positions. 
 Interestingly, the United States had a notable open access market share of `r top_20_plot$data |> distinct(country_code, oa, all) |> mutate(prop = oa / sum(oa)) |> filter(country_code == "US") |> pull(prop) * 100`%, although transformative agreements contributed to a lesser extent.
-Similarly, China's open access market share of `r top_20_plot$data |> distinct(country_code, oa, all) |> mutate(prop = oa / sum(oa)) |> filter(country_code == "CN") |> pull(prop) * 100` in 2022 was comparable to that of the Netherlands, which was (`r  top_20_plot$data |> distinct(country_code, oa, all) |> mutate(prop = oa / sum(oa)) |> filter(country_code == "NL") |> pull(prop) * 100`%,).
+Similarly, China's open access market share of `r top_20_plot$data |> distinct(country_code, oa, all) |> mutate(prop = oa / sum(oa)) |> filter(country_code == "CN") |> pull(prop) * 100`% in 2022 was comparable to that of the Netherlands, which was `r  top_20_plot$data |> distinct(country_code, oa, all) |> mutate(prop = oa / sum(oa)) |> filter(country_code == "NL") |> pull(prop) * 100`%.
 
@@
 Notably, The Netherlands (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "NL") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), Sweden (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "SE") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), Poland (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "PL") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%) and Great Britain (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "GB") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%)) exhibited a relatively high level of uptake in 2018 which continued to increase in the following years.
-In 2022, Sweden had the highest proportion of open-access articles relative to its publication output (`r country_plot$data |> filter(cr_year == "2022", country_name_fct == "SE") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), followed by the Netherlands (`r country_plot$data |> filter(cr_year == "2022", country_name_fct == "NL") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%) and Switzerland (`r country_plot$data |> filter(cr_year == "2022", country_name_fct == "CH") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), with these countries benefiting from transformative agreements.
+In 2022, Sweden had the highest proportion of open access relative to its publication output (`r country_plot$data |> filter(cr_year == "2022", country_name_fct == "SE") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), followed by the Netherlands (`r country_plot$data |> filter(cr_year == "2022", country_name_fct == "NL") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%) and Switzerland (`r country_plot$data |> filter(cr_year == "2022", country_name_fct == "CH") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), with these countries benefiting from transformative agreements.
 In Germany, however, hybrid open access only began to increase from 2019 onwards after the successful negotiation of nationwide agreements with Wiley (July 2019) and Springer Nature (January 2020).
@@
 
-Since 2021, there has been a general trend towards hybrid open access among many western countries, primarily driven by transformative agreements.
+Since 2021, there has been a general trend towards hybrid open access among many high-income countries, primarily driven by transformative agreements.
 However, proliferation of transformative agreements differed across these countries.
@@
 
+The primary aim of this study was to investigate the adoption of open access in hybrid journals through transformative agreements, presenting a novel approach based on open data. 
+The results indicate a steady increase, estimating that the majority of open access in hybrid journals in 2022 originate from institutions that had an transformative agreement in place. 
+Yet, the majority of research literature published in hybrid journals between 2018 and 2022 remained behind a publisher paywall. 
+Growth in the adoption of open access in hybrid journals through transformative agreements can be largely attributed to big three commercial publishers -- Elsevier, Springer Nature, and Wiley -- but varies across journals, publishers, subjects, and country affiliations.
+Despite limitations regarding the data used, the findings indicate that the current level of implementation of transformative agreements is insufficient to bring about a large-scale transformation to full open access.
+
+Presenting an initial overview, some journals recorded substantial open access uptake rates through transformative agreements.
+Previous research indicate that these journals highlight specific areas where publisher portfolios already had strong market positions [@].
+
 # Conclusions

### 2024-02-19 analysis/manuscript.tex
starting discussions

@@
 open access to 1\% of articles in hybrid journals (Dallmeier-Tiessen et
-al., 2010). Other publishers also recorded a low uptake. In 2011, only
+al., 2010). Other studies also recorded a low uptake. In 2011, only
 1-2\% of articles were open access (Björk, 2012), growing to around 4\%
@@
 
-With the introduction of central funding mechanisms for hybrid open
-access publication fees in some European countries since 2012, an
-increase in open access could be observed (Björk, 2017; Huang et al.,
-2020; Jubb et al., 2017; Piwowar et al., 2018). For example, studying
-university output, Robinson-Garcia et al. (2020) estimated a median
-uptake of 7.1\% in the period 2014-2017. In particular, British (17\%),
-Austrian (15\%) and Dutch (13\%) universities contributed to this trend.
-However, this shift in funding policy towards hybrid open access also
-added to the overall cost of publishing, which includes subscription
-spending and the administrative efforts required to handle payments
-(Pinfield et al., 2016). Moreover, established large commercial
-publishers, which already dominated the publishing market (Larivière et
-al., 2015), disproportionately benefited from hybrid open access funding
-in comparison to full open access publishers (Butler et al., 2023; Jahn
-\& Tullney, 2016; Shu \& Larivière, 2023).
+With the introduction of central funding mechanisms for publication fees
+in some European countries since 2012, an increase in hybrid open access
+could be observed (Björk, 2017; Huang et al., 2020; Jubb et al., 2017;
+Piwowar et al., 2018). For example, studying university output,
+Robinson-Garcia et al. (2020) estimated a median uptake of 7.1\% in the
+period 2014-2017. In particular, British (17\%), Austrian (15\%) and
+Dutch (13\%) universities contributed to this trend. However, this shift
+in funding policy towards hybrid open access also added to the overall
+cost of publishing, which includes subscription spending and the
+administrative efforts required to handle payments (Pinfield et al.,
+2016). Moreover, established large commercial publishers, which already
+dominated the publishing market (Larivière et al., 2015),
+disproportionately benefited from hybrid open access funding in
+comparison to full open access publishers (Butler et al., 2023; Jahn \&
+Tullney, 2016; Shu \& Larivière, 2023).
 
@@
 in which well-established commercial publishers gain twice from reading
-and publication fees, as well as to increase publisher-provided
+and open access publishing, as well as to increase publisher-provided
 immediate open access (Björk \& Solomon, 2014; Schimmer et al., 2015).
 These considerations resulted into transformative agreements, which
-cover a broad range of successfully negotiated contracts between library
-consortia and publishers from mid-2010s onwards where institutional
-spending for subscriptions are re-purposed for open access publishing
+cover a broad range of contracts between library consortia and
+publishers from mid-2010s onwards where institutional spending for
+subscriptions and open access publishing are considered together
 (Borrego et al., 2020; Hinchliffe, 2019). Transformative agreements seek
 to control costs while allowing a transitional phase for publishing more
-open access articles. Although transformative agreements mainly bundle
-hybrid and subscription-only journals from commercial publishers, they
-aim at a higher degree of transparency than previous big deals, where
-contracts including payments were confidential (Bergstrom et al., 2014).
+open access articles. Similar to big deals, transformative agreements
+mainly bundle hybrid and subscription-only journals from commercial
+publishers, but aim at a higher degree of transparency than previous big
+deals, where contracts including payments were confidential (Bergstrom
+et al., 2014).
 
@@
 possible'' (Schiltz, 2018). Specifically, they agreed to support hybrid
-open access solely through transformative agreements from 1 January
-2021, until the end of 2024. Notably, the German Research Foundation
-(DFG), despite not being part of cOAlition S, has also extended its
-financial support for hybrid open access through transformative
-agreements (Mittermaier, 2021). Previously, the DFG only provided
-funding for full open access journals (Jahn \& Tullney, 2016).
+open access only through transformative agreements from 2021, until the
+end of 2024. Notably, the German Research Foundation (DFG), despite not
+being part of cOAlition S, has also extended its financial support for
+hybrid open access through transformative agreements (Mittermaier,
+2021). Previously, the DFG only provided funding for full open access
+journals (Jahn \& Tullney, 2016).
 
 By the end of 2023, many transformative agreements were implemented, but
-outcomes were mixed. The ESAC Transformative Agreement
+interim outcomes are mixed. The ESAC Transformative Agreement
 Registry\footnote{\url{https://esac-initiative.org/about/transformative-agreements/agreement-registry/}},
@@
 transparency in terms of contracts and publisher-provided article
-metadata (Marques et al., 2019). However, with the growing trend toward
-transformative agreements, continued reliance on big deals is perceived
-as problematic, which might perpetuate market concentration (Butler et
-al., 2023; Shu \& Larivière, 2023). Whether transformative agreements
-lead to reduced pricing remains uncertain (Borrego, 2023) and a
-substantial transition of hybrid journals towards full open access could
-not be observed (Matthias et al., 2019; Momeni et al., 2021).
-Furthermore, the focus on large commercial publishers might increase
-inequality, because transformative agreements continue to target
-institutions from high-income countries (Klebel \& Ross-Hellauer, 2023).
+metadata (Marques et al., 2019; Pinhasi et al., 2021). However, with the
+growing trend toward transformative agreements, continued reliance on
+big deals is perceived as problematic, because it perpetuate market
+concentration (Butler et al., 2023; Shu \& Larivière, 2023). Whether
+transformative agreements lead to reduced pricing remains uncertain
+(Borrego, 2023) and a substantial transition of hybrid journals towards
+full open access could not be observed (Matthias et al., 2019; Momeni et
+al., 2021). The focus on large commercial publishers might also increase
+inequality, because transformative agreements focus on pay to publish
+open access mainly targets institutions from high-income countries,
+furthering a questionable journal prestige culture (\textbf{budapest?}).
 Besides, an editorial-board resignation raised concerns that publishers'
@@
 Likewise, most cOAlition S funders will discontinue financial support
-for transformative agreements from 2024 onwards (Liverpool, 2023). The
+for transformative agreements by the end of 2024 (Liverpool, 2023). The
 consortium also removed most hybrid journals from its Transformative
-Journal program in 2023 due to their failure to meet self-defined open
-access growth targets (Brainard, 2023). In contrast, the German DEAL
-consortium signed a five-year transformative agreement with Elsevier
-starting in 2024, while also renewing its contracts with Springer Nature
-and Wiley until the end of 2028.
+Journal program in 2023 due to publishers' failure to meet self-defined
+open access growth targets (Brainard, 2023). In contrast, the German
+DEAL consortium signed a five-year transformative agreement with
+Elsevier starting in 2024, while also renewing its contracts with
+Springer Nature and Wiley until the end of 2028.
 
@@
 (Bakker et al., 2024; Fraser et al., 2023; Jahn et al., 2022; Momeni et
-al., 2023; Pieper \& Broschinski, 2018), while related large-scale
-studies relied on self-reported agreement data (Moskovkin et al., 2022),
-or used APC pricing lists (Shu \& Larivière, 2023). Particularly, data
+al., 2023; Pieper \& Broschinski, 2018), while large-scale studies
+relied on self-reported agreement data (Moskovkin et al., 2022), or used
+APC pricing lists (Shu \& Larivière, 2023). Particularly, data
 availability is a limiting factor when studying the impact of
@@
 through transformative agreements and other means. With this novel and
-open approach, this study aims to answer the following questions:
+open approach, this the first large-scale analysis aims to answer the
+following questions:
 
@@
   What was the number and proportion of open access articles in hybrid
-  journals included in transformative agreements between 2018 and 2022?
+  journals in transformative agreements between 2018 and 2022?
 \item
-  How did transformative agreements impact open access uptake in hybrid
-  journals?
+  To what extent did institutions with a transformation agreement
+  contribute to the adoption of open access in hybrid journals?
 \end{itemize}
@@
 
-\hypertarget{methods-and-data}{%
-\section{Methods and data}\label{methods-and-data}}
+\hypertarget{methods}{%
+\section{Methods}\label{methods}}
 
@@
 transformative agreement data retrieved from the cOAlition S Journal
-Checker Tool provided information about journal portfolios and
-participating institutions. After identification of hybrid journals by
-excluding full open access journals, Crossref served as the primary data
-source for article-level metadata including Creative Commons (CC)
-license information to indicate open access availability on publisher
-websites. To determine open access articles published through
-transformative agreements, first author affiliations from OpenAlex
-(Priem et al., 2022) were subsequently linked to eligible institutions
-according to the transformative agreement data.
+Checker Tool\footnote{\url{https://www.coalition-s.org/blog/enabling-accurate-results-within-the-journal-checker-tool/}}
+provided information about journal portfolios and participating
+institutions. After identification of hybrid journals by excluding full
+open access journals, Crossref served as the primary data source for
+article-level metadata including Creative Commons (CC) license
+information to indicate open access availability on publisher websites.
+To determine open access articles published through transformative
+agreements, first author affiliations from OpenAlex (Priem et al., 2022)
+were subsequently linked to eligible institutions according to the
+transformative agreement data. In the following, these steps are
+described in more detail.
 
@@
 combining different data sources considerably extended the journal
-matching to exclude full open access journals from transformative
-agreements. In total, 3,439 full open access journals were excluded
-based on ISSN matching. The overlap between the three data sources was
-72\%. The Gold OA journals dataset alone added 176 journals, while the
-DOAJ comprised 10 full open access journals not listed in either of the
-other two sources. These full open access journals were mostly launched
-in 2022.
+matching. In total, 3,439 full open access journals were excluded based
+on ISSN matching. The overlap between the three data sources was 72\%.
+The Gold OA journals dataset alone added 176 journals, while the DOAJ
+comprised 10 full open access journals not listed in either of the other
+two sources. These full open access journals were mostly launched in
+2022.
 
@@
 not provide CC license information to Crossref including AIP Publishing,
-American Physiological Society and Emerald. Crossref license metadata
-was more complete with regard to articles from the publisher Wiley and
-American Chemical Society. Finally, inconsistent open access status
-information in previous OpenAlex versions was observed (Jahn et al.,
-2023). According to the OpenAlex release notes, fixing this issue was
-still ongoing, which might also explain this discrepancy.
+American Physiological Society, Emerald and the Royal Society. Crossref
+license metadata was more complete with regard to articles from the
+publisher Wiley and American Chemical Society. Finally, inconsistent
+open access status information in previous OpenAlex versions was
+observed (Jahn et al., 2023). After reporting it to OpenAlex, according
+to the release notes, fixing this issue was still ongoing, which might
+also explain part of the discrepancy.
 
@@
 contribute most to a paper and are often considered lead author research
-papers (Larivière et al., 2016), and can be therefore assumed as a proxy
-to measure to open access payments (Shu \& Larivière, 2023; Zhang et
-al., 2022). Overall, around 90\% of studied articles had first author
-affiliation metadata in OpenAlex, whereas the coverage of articles with
-corresponding author information was around 54\%.
+papers (Larivière et al., 2016). Related studies therefore assumed first
+authors as a proxy to measure to open access payments and the impact of
+transformative agreement (Haucap et al., 2021; Shu \& Larivière, 2023;
+Zhang et al., 2022). Overall, around 90\% of studied articles had first
+author affiliation metadata in OpenAlex, whereas the coverage of
+articles with corresponding author information was around 54\%.
 
@@
 enriched with associated organisations using OpenAlex's institution
-entity.
+entity data.
 
@@
 in 12,857 hybrid journals between 2018 and 2022 (see Figure
-\ref{fig:method_fig}B). Hybrid journals included in transformative
-agreements represented 40\% of total global output over the same time
-period according to Crossref, while full open access journals accounted
-for 35\% of total article volume.
+\ref{fig:method_fig}B). Hybrid journals in transformative agreements
+represented 40\% of total global output over the same time period
+according to Crossref, while full open access journals recorded 35\%.
 
@@
 language (R Core Team, 2020) were used. The resulting data is openly
-available through an R package, hoaddata, version 0.2.91 Following
+available through an R data package, hoaddata, version 0.2.91 Following
 Marwick et al. (2018), hoaddata contains not only the datasets used in
@@
 big scholarly data from Crossref, OpenAlex and Unpaywall were imported.
-To increase the computational reproducibility, the R package was
-automatically built using GitHub Actions, a continuous integration
-service.
+To increase the computational reproducibility, the data aggregation
+through hoaddata was automatically carried out using GitHub Actions, a
+continuous integration service.
 
@@
 
-Figure \ref{fig:results_overview}B highlights that the majority of open
-access articles in hybrid journals were made available through
-transformative agreements in 2021 and 2022, contributing 58\% of the
-total open access article volume in 2022. However, there was also a
-notable growth in open access provision through individual publication
-fees, which increased from 4.1\% (n = 62,625) in 2018 to 6.3\% (n =
-105,896). This suggests that publishers were able to gain equally from
-individual and institutional open access publishing options.
+Figure \ref{fig:results_overview}B highlights that the majority of
+hybrid open access was made available through transformative agreements
+in 2021 and 2022, contributing 58\% of the total open access article
+volume in 2022. However, there was also a notable growth in open access
+provision through other means, most likely publication fees being not
+invoiced through transformative agreements, which increased from 4.1\%
+(n = 62,625) in 2018 to 6.3\% (n = 105,896).
 
@@
 than those with an open access share below 50\% (M = 164, SD = 347).
-Notable exception of large journals with an above-average open access
-proportion were Physical Review D, a high-energy physics journal covered
-by the SCOAP3 consortium that provided open access to 2,341 out of 4,074
-articles in 2022, Astronomy and Physics (1,396 out of 2,230 articles in
-2022 were open access), which shifted to a subscribe to open business
-model for all accepted articles as of April 2022, and the Journal of
-Fluid Mechanics (577 of 1,077 articles in 2022 were open access).
 
@@
 agreements substantially contributed to the growth of open access over
-the years (Figure \ref{fig:results_overview}D). Examples of such
-journals include those with a scope on specific countries or regions,
-where also transformative agreements were implemented. For instance, in
-2022, the Germany-based journals Zeitschrift für Erziehungswissenschaft
-and Zeitschrift für Politikwissenschaft, as well as the Scandinavian
-Political Studies adressing the Nordic countries, achieved an overall
-open access uptake of more than 90\% just through transformative
-agreements. Despite the rise in transformative agreements, it is worth
-noting that other means of publishing open access in hybrid journals
-remained common. In total, 9,153 journals published open access articles
-from authors affiliated with institutions without transformative
-agreements in place, while 8,780 journals published at least one open
-access article through a transformative agreement in the same year.
+the years (Figure \ref{fig:results_overview}D). Despite the rise in
+transformative agreements, it is worth noting that other means of
+publishing open access remained common across the investiagted hybrid
+journals. In total, 9,153 journals published open access articles from
+authors affiliated with institutions without transformative agreements
+in place, while 8,780 journals published at least one open access
+article through a transformative agreement in the same year.
 
@@
 transformative agreements, the big three commercial publishers Elsevier,
-Springer Nature, and Wiley, accounted for 49\% of total article volume
+Springer Nature, and Wiley accounted for 49\% of total article volume
 published (see Table \ref{tab:publisher_league_table}). Together, they
-published 500,878 open access articles, representing 66\% of the open
-access articles in hybrid journals. Elsevier, Springer Nature, and Wiley
-made 243,891 articles open access in hybrid journals through
-transformative agreements, resulting in an even larger market share of
-74\%.
+published 500,878 or 66\% of open access articles in hybrid journals.
+Elsevier, Springer Nature, and Wiley made 243,891 articles open access
+in hybrid journals through transformative agreements, resulting in an
+even larger market share of 74\%.
 
@@
 
-However, there are notable differences among the three large publishers.
+However, there are notable differences among the three big publishers.
 Although Elsevier published the largest volume of articles (n =
-2,770,826, 34\%), it published a relatively low number of open access
+2,770,826, 34\%), it recorded a comparable low number of open access
 articles, including those that can be associated with transformative
@@
 show a general long-term trend towards transformative agreements, Figure
-\ref{fig:publisher_figure}A and B indicate that, in particular, Wiley's
-has experienced a substantial increase in its open access share from
-5.9\% (n = 11,628) in 2018 to 26\% (n = 53,503) in 2022, representing an
+\ref{fig:publisher_figure}A and B indicate that, in particular, Wiley
+experienced a substantial increase in its open access share from 5.9\%
+(n = 11,628) in 2018 to 26\% (n = 53,503) in 2022, representing an
 4.5-fold increase. In contrast, Elsevier's hybrid journals demonstrated
@@
 
-The varying degrees of uptake of open access across the three major
+The varying degrees of adoption of open access across the three major
 publishers can be attributed to distinct approaches to transformative
@@
 concluded prior to the start of the transformative agreement data
-collection in June 2021. Nonetheless, the results suggest the importance
-of central agreements for Springer Nature's hybrid open access business
-over the past five years (Figure 2B). In 2022, 66\% (n = 34,725) of open
-access articles in Springer Nature hybrid journals were enabled through
-transformative agreements. In the same year, 70\% (n = 37,316) of
-Wiley's open access articles could be linked to transformative
+collection in June 2021. Nonetheless, the results demonstrate the
+importance of agreements for Springer Nature's hybrid open access
+business over the past five years (Figure 2B). In 2022, 66\% (n =
+34,725) of open access in Springer Nature hybrid journals were enabled
+through transformative agreements. In the same year, 70\% (n = 37,316)
+of Wiley's open access articles could be linked to transformative
 agreements in 2022. In contrast, Elsevier published fewer than half of
@@
 transformative agreements and other revenue sources could observed
-across Elsevier journals, the distribution of open access across
+across Elsevier's portfolio, the distribution of open access across
 Springer Nature and Wiley hybrid journals indicates that the growth is
@@
 and 2022, most hybrid journals with at least one open articles could be
-attributed to the category Social Sciences, which also includes the
-humanities. However, these journals published the fewest number of
+attributed to the category Social Sciences, which also includes the Arts
+and Humanities. However, these journals published the fewest number of
 articles, whereas Physical Sciences journals recorded most articles,
@@
 39,494) , Health Science from 5.3\% (n = 18,279) to 16\% (n = 63,089)
-and Physical Sciences from 4.5\% (n = 22,364) to 12\% (n = 85,428). This
-growth in the social sciences can be largely attributed to
+and Physical Sciences from 4.5\% (n = 22,364) to 12\% (n = 85,428).
+Growth in the Social Sciences can be largely attributed to
 transformative agreements. In 2022, two-third of open access articles
@@
 Figure \ref{fig:subject_panel}C shows that this trend is consistent
-across Social Sciences journals. In 2022, 25\% of Social Science
-journals provided open access to at least every fourth article
-exclusively through transformative agreements. However, hybrid open
-access through transformative agreements played a comparable lesser role
-in the Life Sciences and Health Sciences. In these two subject areas,
-only about half of the open access articles can be linked to these
-agreements, both overall and on median average across journals. In
-contrast, the majority of Physical Science Journals, shows an increase
-of open access through transformative agreements compared to other
-options to publish open access in hybrid journals.
+across hybrid journals belonging to the ASJC subject area Social
+Sciences. In 2022, 25\% of Social Sciences journals provided open access
+to at least every fourth article exclusively through transformative
+agreements. However, hybrid open access through transformative
+agreements played a comparable lesser role in the Life Sciences and
+Health Sciences. In these two subject areas, only about half of the open
+access articles can be linked to these agreements, both overall and on
+median average across journals. In contrast, the majority of Physical
+Science Journals, shows an increase of open access through
+transformative agreements compared to other options to publish open
+access in hybrid journals.
 
@@
 
-Between 2018 and 2022, high-income economies almost exclusively
+Between 2018 and 2022, high-income countries almost exclusively
 dominated hybrid open access publishing through transformative
@@
 an intergovernmental organisation, which comprised the countries Brazil,
-Russia, India, China and South Africa in 2022. The residual category
+Russia, India, China and South Africa as of 2022. The residual category
 ``Other'' includes the remaining countries. From 2018 to 2022, the
@@
 
-To illustrate the situation in 2022, \ref{fig:country_patch}B compares
-total publication output with the number of open access articles. With
-391,530 articles, China was the most productive country, followed by the
-United States (268,965 articles) and India (87,428 articles). In
-contrast, West and Nord European countries published a considerable high
-number of open access articles, mainly due to transformative agreements.
-Particularly, Germany, Great Britain, the Netherlands, Sweden,
-Switzerland and Spain recorded an above-average open access share as
-indicated by the linear trend line. As represented by the point size, as
-well as it can been seen in Figure \ref{fig:country_patch}C,
-transformative agreements contributed to this market position of these
-countries. Interestingly, the United States had a notable open access
-market share of 15\%, although transformative agreements contributed to
-a lesser extent. Similarly, China's open access market share of 7.2 in
-2022 was comparable to that of the Netherlands, which was (7.1\%,).
+To illustrate the situation in 2022, Figure \ref{fig:country_patch}B
+compares total publication output with the number of open access
+articles. With 391,530 articles, China was the most productive country,
+followed by the United States (268,965 articles) and India (87,428
+articles). In contrast, West and Nord European countries published a
+considerable high number of open access articles, mainly due to
+transformative agreements. Particularly, Germany, Great Britain, the
+Netherlands, Sweden, Switzerland and Spain recorded an above-average
+open access share as indicated by the linear trend line. As represented
+by the point size, as well as it can been seen in Figure
+\ref{fig:country_patch}C, transformative agreements contributed to these
+market positions. Interestingly, the United States had a notable open
+access market share of 15\%, although transformative agreements
+contributed to a lesser extent. Similarly, China's open access market
+share of 7.2\% in 2022 was comparable to that of the Netherlands, which
+was 7.1\%.
 
@@
 2018 which continued to increase in the following years. In 2022, Sweden
-had the highest proportion of open-access articles relative to its
-publication output (78\%), followed by the Netherlands (67\%) and
-Switzerland (57\%), with these countries benefiting from transformative
-agreements. In Germany, however, hybrid open access only began to
-increase from 2019 onwards after the successful negotiation of
-nationwide agreements with Wiley (July 2019) and Springer Nature
-(January 2020). Prior to this, only a few organisations had agreements
-in place, for the example the Max Planck Society with Springer Compact.
+had the highest proportion of open access relative to its publication
+output (78\%), followed by the Netherlands (67\%) and Switzerland
+(57\%), with these countries benefiting from transformative agreements.
+In Germany, however, hybrid open access only began to increase from 2019
+onwards after the successful negotiation of nationwide agreements with
+Wiley (July 2019) and Springer Nature (January 2020). Prior to this,
+only a few organisations had agreements in place, for the example the
+Max Planck Society with Springer Compact.
 
@@
 Since 2021, there has been a general trend towards hybrid open access
-among many western countries, primarily driven by transformative
+among many high-income countries, primarily driven by transformative
 agreements. However, proliferation of transformative agreements differed
@@
 
+The primary aim of this study was to investigate the adoption of open
+access in hybrid journals through transformative agreements, presenting
+a novel approach based on open data. The results indicate a steady
+increase, estimating that the majority of open access in hybrid journals
+in 2022 originate from institutions that had an transformative agreement
+in place. Yet, the majority of research literature published in hybrid
+journals between 2018 and 2022 remained behind a publisher paywall.
+Growth in the adoption of open access in hybrid journals through
+transformative agreements can be largely attributed to big three
+commercial publishers -- Elsevier, Springer Nature, and Wiley -- but
+varies across journals, publishers, subjects, and country affiliations.
+Despite limitations regarding the data used, the findings indicate that
+the current level of implementation of transformative agreements is
+insufficient to bring about a large-scale transformation to full open
+access.
+
+Presenting an initial overview, some journals recorded substantial open
+access uptake rates through transformative agreements. Previous research
+indicate that these journals highlight specific areas where publisher
+portfolios already had strong market positions {[}@{]}.
+
 \hypertarget{conclusions}{%
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Brainard_2023}{}}%
-Brainard, J. (2023). {``Transformative''} journals get booted for
+Brainard, J. (2023). {`Transformative'} journals get booted for
 switching to open access too slowly. In \emph{Science: SCIENCEINSIDER}.
@@
 Geschuhn, K., \& Stone, G. (2017). It's the workflows, stupid! What is
-required to make {``offsetting''} work for the open access transition.
+required to make {`offsetting'} work for the open access transition.
 \emph{Insights the {UKSG} Journal}, \emph{30}(3), 103--114.
@@
 \url{https://web.archive.org/web/20200212015524/https://www.universitiesuk.ac.uk/policy-and-analysis/reports/Documents/2017/monitoring-transition-open-access-2017.pdf}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Klebel_2023}{}}%
-Klebel, T., \& Ross-Hellauer, T. (2023). The APC-barrier and its effect
-on stratification in open access publishing. \emph{Quantitative Science
-Studies}, \emph{4}(1), 22--43. \url{https://doi.org/10.1162/qss_a_00245}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Pinhasi_2021}{}}%
+Pinhasi, R., Hölbling, L., \& Kromp, B. (2021). Austrian transition to
+open access: A collaborative approach. \emph{Insights the UKSG Journal},
+\emph{34}. \url{https://doi.org/10.1629/uksg.561}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Pinhasi_2020}{}}%
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Goodin_2023}{}}%
-Rasmussen, K. B. (2023). Interview with {Robert {``Bob''} E. Goodin}.
+Rasmussen, K. B. (2023). Interview with {Robert {`Bob'} E. Goodin}.
 \emph{Tidskrift För Politisk Filosofi}.

### 2024-02-21 analysis/manuscript.Rmd
continuing discussions

@@
 As a consequence, libraries and their consortia began to develop licensing strategies aimed at avoiding such ‘double dipping’ scenarios, in which well-established commercial publishers gain twice from reading and open access publishing, as well as to increase publisher-provided immediate open access [@BJ_RK_2014; @Schimmer_2015].
-These considerations resulted into transformative agreements, which cover a broad range of contracts between library consortia and publishers from mid-2010s onwards where institutional spending for subscriptions and open access publishing are considered together [@Borrego_2020; @Hinchliffe_2019].
+These considerations resulted into transformative agreements[^agreements], which cover a broad range of contracts between library consortia and publishers from mid-2010s onwards where institutional spending for subscriptions and open access publishing are considered together [@Borrego_2020; @Hinchliffe_2019].
 Transformative agreements seek to control costs while allowing a transitional phase for publishing more open access articles. 
@@
 In contrast, the German DEAL consortium signed a five-year transformative agreement with Elsevier starting in 2024, while also renewing its contracts with Springer Nature and Wiley until the end of 2028.
+At the same time, the Colombia Consortium signed the first transformative agreements in Latin America [@Mu_oz_V_lez_2024].
 
@@
 
-The primary aim of this study was to investigate the adoption of open access in hybrid journals through transformative agreements, presenting a novel approach based on open data. 
-The results indicate a steady increase, estimating that the majority of open access in hybrid journals in 2022 originate from institutions that had an transformative agreement in place. 
-Yet, the majority of research literature published in hybrid journals between 2018 and 2022 remained behind a publisher paywall. 
-Growth in the adoption of open access in hybrid journals through transformative agreements can be largely attributed to big three commercial publishers -- Elsevier, Springer Nature, and Wiley -- but varies across journals, publishers, subjects, and country affiliations.
-Despite limitations regarding the data used, the findings indicate that the current level of implementation of transformative agreements is insufficient to bring about a large-scale transformation to full open access.
-
-Presenting an initial overview, some journals recorded substantial open access uptake rates through transformative agreements.
-Previous research indicate that these journals highlight specific areas where publisher portfolios already had strong market positions [@].
-
-# Conclusions
+The primary aim of this study was to investigate the adoption of open access in hybrid journals in transformative agreements, which were introduced as a temporal means to support the transition of subscription-based academic publishing to full open access.
+This study presents a novel approach based on open data, which leverages metadata on over 700 agreements and nine million journal articles to estimate the extent to which transformation agreement contribute to the transition of this journal business model based on institutional affiliation.
+The results highlight a strong growth in open access between 2018 and 2022, particularly driven by an increasing number of institutions that had transformative agreements in place. 
+Yet, the majority of research literature published in hybrid journals in this five-year period remained behind a publisher's paywall. 
+Growth in the adoption of open access in hybrid journals, in particular through transformative agreements, can be largely attributed to three large commercial publishers -- Elsevier, Springer Nature, and Wiley -- but varies substantially across journals, publishers, subjects, and country affiliations.
+Despite the limitations of the data used, the findings indicate that the current level of implementation of transformative agreements is insufficient to bring about a large-scale transition to fully open access.
+
+A key finding of the analysis is that transformative agreements increase market concentration. 
+Specifically, the three largest commercial publishers—Elsevier, Springer Nature, and Wiley—have increased their market share in the context of open access, particularly in regard to open access provided through transformative agreements.
+Together, the three publishers accounted for three-fourths of open access articles through transformative agreements, despite publishing less than half of the total publication volume published between 2018 and 2022.
+This observation aligns with previous research on shifts in the publishing market following the introduction of funding opportunities for hybrid open access [@Jahn_2016; @Butler_2023; @Shu_2023]. 
+Additionally, the results confirm previously observed variations by publisher, with Elsevier exhibiting a different development than Springer Nature and Wiley  [@Butler_2023; @Jahn_2021]. 
+However, it must be noted that the focus of transformative agreements on publishers with large hybrid journal portfolios is intentional [@ifla].
+Because of transformative agreements with a few large publishers, national consortia were able to substantially increased their country’s annual open access article volume [@Pinhasi_2021; @Huang_2020; @Taubert_2023; @Wenaas_2022].
+
+Moreover, this study presents varying levels of open access uptake through transformative agreements across journals, which can be attributed to the alignment of authors' affiliations and the availability of such agreements at their institutions.
+In line with previous research findings [@Butler_2023; @Jahn_2021], high uptake rates were observed across hybrid journals in the Social Sciences and Humanities.
+However, it is important to emphasise that these hybrid journals do not encompass the entire field.
+For example, a study by @Khanna_2022 found that approximately 60% of journals utilizing the open-source publishing platform Open Journal Systems (OJS) fall within the Social Sciences and Humanities. 
+In these fields, the majority of full open access journals are so-called "Diamond OA journals" that do not charge publication fees [@bosman_2021_4558704].
+Rather, this result can be more accurately attributed to the substantial proportion of authors from high-income countries who publish in these hybrid journals included in transformative agreements, particularly when journals are of local relevance, such as through the belonging to national societies or regional research scope.
+
+ <!--   [funding options in the humanities and social sicences] -->
+Surprisingly, the total publication output of high-income countries belonging to the OECD declined susbtantially between 2018 and 2022, while the BRICS countries recorded an steady increase. 
+Because the BRICS expansion can be exclusivley attributed to closed access articles, this developement has the potential to hinder the transition of academic journal publishing to full open access through transformative agreements.
+In China, the most productive country according to the total article volume in hybrid journals, 
+(shu und zhang und das indische paper)
+
+other countries less represented and brazil diamond russia The Eastern Europe Expanded Collection
+
+but not coherent acorss all countries, in particular in the us. neverthelles china and us have susbtantial market shares when it comes to hybrid oa.
+
+There are at least three potential limitations
+concerning the results of this study. A first limitation
+concerns … A second potential limitation is that …
+
+These data have some potential intervention implications.
+forst and formesot, the supprot recent pocliy decision highlighting a grew in oa at country-level, but questioning the.
+if thouse contracts, a transiliteration is ttoo short and might require a longer transiltierional period.
+the growtrh in brics however would require too impleis unboundling, other payment means and focus on alternative, in particular innovations like preprinting, diamond oa
+
+In terms of future research, it would be useful to extend the current findings by examining other oa types, combine oa with mobility.
+fianncials tudies are also required to assess claims of cost-sensitive, here, publisher and publsiher are required to increase transparency.
+at the same time, studies needed to assess these claim economically.
+data can be improved to entail more, in apritcular consortia could start to disclose funded articles as , with openapc already a cwors-orucing platform available.
+also libraries can contrbute to an improved catalogiuing of transformative agreements including participating insitutions, which is aligned with standards used in schoalrly communications like ror id.
+
+In summary, after more than tow decades, hybrid open access has not demonstrated a succesful menas.
+stduying transofrmative agreemnts, effects can be found at the -level of participating countries, but overlla uptake remains low, in particular with regard to publications from brocs countries, which country share rise, exclusively on the expense of closed access.
+means
 
@@
 [^sweden]: <https://www.su.se/english/news/open-access-need-to-move-away-from-transformative-agreements-1.683787>
+
+[^agreements]: In this paper I use the term "transformative agreement", addressing also offsetting, read-and-publish or publish-and-read deals, and other variants [@Borrego_2020; @Hinchliffe_2019].
+Although the term is critised as misleading and not useful to describe these agreements between library consortia and publishers [@budapest], it is widely used in policy discussions and in the research literature.

### 2024-02-21 analysis/manuscript.tex
continuing discussions

@@
 immediate open access (Björk \& Solomon, 2014; Schimmer et al., 2015).
-These considerations resulted into transformative agreements, which
-cover a broad range of contracts between library consortia and
+These considerations resulted into transformative agreements\footnote{In
+  this paper I use the term ``transformative agreement'', addressing
+  also offsetting, read-and-publish or publish-and-read deals, and other
+  variants (Borrego et al., 2020; Hinchliffe, 2019). Although the term
+  is critised as misleading and not useful to describe these agreements
+  between library consortia and publishers (Babini et al., 2022), it is
+  widely used in policy discussions and in the research literature.},
+which cover a broad range of contracts between library consortia and
 publishers from mid-2010s onwards where institutional spending for
@@
 open access mainly targets institutions from high-income countries,
-furthering a questionable journal prestige culture (\textbf{budapest?}).
-Besides, an editorial-board resignation raised concerns that publishers'
-desire to maximize journal publication volume ``without regard to
-quality'' is a consequence of transformative agreements (Rasmussen,
-2023).
+furthering a questionable journal prestige culture (Babini et al.,
+2022). Besides, an editorial-board resignation raised concerns that
+publishers' desire to maximize journal publication volume ``without
+regard to quality'' is a consequence of transformative agreements
+(Rasmussen, 2023).
 
@@
 Elsevier starting in 2024, while also renewing its contracts with
-Springer Nature and Wiley until the end of 2028.
+Springer Nature and Wiley until the end of 2028. At the same time, the
+Colombia Consortium signed the first transformative agreements in Latin
+America (Muñoz-Vélez et al., 2024).
 
@@
 The primary aim of this study was to investigate the adoption of open
-access in hybrid journals through transformative agreements, presenting
-a novel approach based on open data. The results indicate a steady
-increase, estimating that the majority of open access in hybrid journals
-in 2022 originate from institutions that had an transformative agreement
-in place. Yet, the majority of research literature published in hybrid
-journals between 2018 and 2022 remained behind a publisher paywall.
-Growth in the adoption of open access in hybrid journals through
-transformative agreements can be largely attributed to big three
-commercial publishers -- Elsevier, Springer Nature, and Wiley -- but
-varies across journals, publishers, subjects, and country affiliations.
-Despite limitations regarding the data used, the findings indicate that
-the current level of implementation of transformative agreements is
-insufficient to bring about a large-scale transformation to full open
-access.
-
-Presenting an initial overview, some journals recorded substantial open
-access uptake rates through transformative agreements. Previous research
-indicate that these journals highlight specific areas where publisher
-portfolios already had strong market positions {[}@{]}.
-
-\hypertarget{conclusions}{%
-\section{Conclusions}\label{conclusions}}
+access in hybrid journals in transformative agreements, which were
+introduced as a temporal means to support the transition of
+subscription-based academic publishing to full open access. This study
+presents a novel approach based on open data, which leverages metadata
+on over 700 agreements and nine million journal articles to estimate the
+extent to which transformation agreement contribute to the transition of
+this journal business model based on institutional affiliation. The
+results highlight a strong growth in open access between 2018 and 2022,
+particularly driven by an increasing number of institutions that had
+transformative agreements in place. Yet, the majority of research
+literature published in hybrid journals in this five-year period
+remained behind a publisher's paywall. Growth in the adoption of open
+access in hybrid journals, in particular through transformative
+agreements, can be largely attributed to three large commercial
+publishers -- Elsevier, Springer Nature, and Wiley -- but varies
+substantially across journals, publishers, subjects, and country
+affiliations. Despite the limitations of the data used, the findings
+indicate that the current level of implementation of transformative
+agreements is insufficient to bring about a large-scale transition to
+fully open access.
+
+A key finding of the analysis is that transformative agreements increase
+market concentration. Specifically, the three largest commercial
+publishers---Elsevier, Springer Nature, and Wiley---have increased their
+market share in the context of open access, particularly in regard to
+open access provided through transformative agreements. Together, the
+three publishers accounted for three-fourths of open access articles
+through transformative agreements, despite publishing less than half of
+the total publication volume published between 2018 and 2022. This
+observation aligns with previous research on shifts in the publishing
+market following the introduction of funding opportunities for hybrid
+open access (Butler et al., 2023; Jahn \& Tullney, 2016; Shu \&
+Larivière, 2023). Additionally, the results confirm previously observed
+variations by publisher, with Elsevier exhibiting a different
+development than Springer Nature and Wiley (Butler et al., 2023; Jahn et
+al., 2022). However, it must be noted that the focus of transformative
+agreements on publishers with large hybrid journal portfolios is
+intentional (Campbell et al., 2022). Because of transformative
+agreements with a few large publishers, national consortia were able to
+substantially increased their country's annual open access article
+volume (Huang et al., 2020; Pinhasi et al., 2021; Taubert et al., 2023;
+Wenaas, 2022).
+
+Moreover, this study presents varying levels of open access uptake
+through transformative agreements across journals, which can be
+attributed to the alignment of authors' affiliations and the
+availability of such agreements at their institutions. In line with
+previous research findings (Butler et al., 2023; Jahn et al., 2022),
+high uptake rates were observed across hybrid journals in the Social
+Sciences and Humanities. However, it is important to emphasise that
+these hybrid journals do not encompass the entire field. For example, a
+study by Khanna et al. (2022) found that approximately 60\% of journals
+utilizing the open-source publishing platform Open Journal Systems (OJS)
+fall within the Social Sciences and Humanities. In these fields, the
+majority of full open access journals are so-called ``Diamond OA
+journals'' that do not charge publication fees (Bosman et al., 2021).
+Rather, this result can be more accurately attributed to the substantial
+proportion of authors from high-income countries who publish in these
+hybrid journals included in transformative agreements, particularly when
+journals are of local relevance, such as through the belonging to
+national societies or regional research scope.
+
+Surprisingly, the total publication output of high-income countries
+belonging to the OECD declined susbtantially between 2018 and 2022,
+while the BRICS countries recorded an steady increase. Because the BRICS
+expansion can be exclusivley attributed to closed access articles, this
+developement has the potential to hinder the transition of academic
+journal publishing to full open access through transformative
+agreements. In China, the most productive country according to the total
+article volume in hybrid journals, (shu und zhang und das indische
+paper)
+
+other countries less represented and brazil diamond russia The Eastern
+Europe Expanded Collection
+
+but not coherent acorss all countries, in particular in the us.
+neverthelles china and us have susbtantial market shares when it comes
+to hybrid oa.
+
+There are at least three potential limitations concerning the results of
+this study. A first limitation concerns \ldots{} A second potential
+limitation is that \ldots{}
+
+These data have some potential intervention implications. forst and
+formesot, the supprot recent pocliy decision highlighting a grew in oa
+at country-level, but questioning the. if thouse contracts, a
+transiliteration is ttoo short and might require a longer
+transiltierional period. the growtrh in brics however would require too
+impleis unboundling, other payment means and focus on alternative, in
+particular innovations like preprinting, diamond oa
+
+In terms of future research, it would be useful to extend the current
+findings by examining other oa types, combine oa with mobility.
+fianncials tudies are also required to assess claims of cost-sensitive,
+here, publisher and publsiher are required to increase transparency. at
+the same time, studies needed to assess these claim economically. data
+can be improved to entail more, in apritcular consortia could start to
+disclose funded articles as , with openapc already a cwors-orucing
+platform available. also libraries can contrbute to an improved
+catalogiuing of transformative agreements including participating
+insitutions, which is aligned with standards used in schoalrly
+communications like ror id.
+
+In summary, after more than tow decades, hybrid open access has not
+demonstrated a succesful menas. stduying transofrmative agreemnts,
+effects can be found at the -level of participating countries, but
+overlla uptake remains low, in particular with regard to publications
+from brocs countries, which country share rise, exclusively on the
+expense of closed access. means
 
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-budapest}{}}%
+Babini, D., Chan, L., Hagemann, M., Joseph, H., Kuchma, I., \& Suber, P.
+(2022). \emph{{The Budapest Open Access Initiative-20th. Anniversary
+recommendations (BOAI20)}}.
+\url{https://www.budapestopenaccessinitiative.org/boai20/}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Bakker_2024}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-bosman_2021_4558704}{}}%
+Bosman, J., Frantsvåg, J. E., Kramer, B., Langlais, P.-C., \& Proudman,
+V. (2021). \emph{OA diamond journals study. Part 1: findings}. Zenodo.
+\url{https://doi.org/10.5281/zenodo.4558704}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Brainard_2023}{}}%
@@
 Science Studies}, 1--22. \url{https://doi.org/10.1162/qss_a_00272}
+
+\leavevmode\vadjust pre{\hypertarget{ref-ifla}{}}%
+Campbell, C., Dér, Á., Geschuhn, K., \& Valente, A. (2022). How are
+transformative agreements transforming libraries? In I. F. of Library
+Associations \& I. (IFLA) (Eds.), \emph{87th IFLA world library and
+information congress (WLIC) / 2022 in dublin, ireland}. IFLA.
+\url{https://repository.ifla.org/handle/123456789/1973}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Khanna_2022}{}}%
+Khanna, S., Ball, J., Alperin, J. P., \& Willinsky, J. (2022).
+Recalibrating the scope of scholarly publishing: A modest step in a vast
+decolonization process. \emph{Quantitative Science Studies},
+\emph{3}(4), 912--930. \url{https://doi.org/10.1162/qss_a_00228}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Laakso_2016}{}}%
@@
 165--207. \url{https://doi.org/10.1080/1941126x.2022.2099000}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Mu_oz_V_lez_2024}{}}%
+Muñoz-Vélez, H., Pallares, C., Echavarría, A. F., Contreras, J., Pavas,
+A., Bello, D., Rendón, C., Calderón-Rojas, J., \& Garzón, F. (2024).
+Strategies for negotiating and signing transformative agreements in the
+global south: The colombia consortium experience. \emph{Journal of
+Library Administration}, \emph{64}(1), 80--98.
+\url{https://doi.org/10.1080/01930826.2023.2287945}
 

### 2024-02-22 analysis/manuscript.Rmd
first draft discussion

@@
 For example, studying university output, @Robinson_Garcia_2020 estimated a median uptake of 7.1% in the period 2014-2017. 
-In particular, British (17%), Austrian (15%) and Dutch (13%) universities contributed to this trend. 
+In particular, British (17%), Austrian (15%) and Dutch (13%) universities stand out. 
 However, this shift in funding policy towards hybrid open access also added to the overall cost of publishing, which includes subscription spending and the administrative efforts required to handle payments [@Pinfield_2016].
@@
 
-As a consequence, libraries and their consortia began to develop licensing strategies aimed at avoiding such ‘double dipping’ scenarios, in which well-established commercial publishers gain twice from reading and open access publishing, as well as to increase publisher-provided immediate open access [@BJ_RK_2014; @Schimmer_2015].
+As a consequence, libraries and their consortia began to develop licensing strategies aimed at avoiding such ‘double dipping’ scenarios, in which well-established commercial publishers gain twice from reading and open access publishing [@Mittermaier_2015], as well as to increase publisher-provided immediate open access [@BJ_RK_2014; @Schimmer_2015].
 These considerations resulted into transformative agreements[^agreements], which cover a broad range of contracts between library consortia and publishers from mid-2010s onwards where institutional spending for subscriptions and open access publishing are considered together [@Borrego_2020; @Hinchliffe_2019].
@@
 a substantial transition of hybrid journals towards full open access could not be observed [@Momeni_2021; @Matthias_2019].
-The focus on large commercial publishers might also increase inequality, because transformative agreements focus on pay to publish open access mainly targets institutions from high-income countries, furthering a questionable journal prestige culture [@budapest].
+The focus on large commercial publishers might also increase inequality [@Klebel_2023], because transformative agreements focus on pay to publish open access mainly targets institutions from high-income countries, furthering a questionable journal prestige culture [@budapest].
 Besides, an editorial-board resignation raised concerns that publishers' desire to maximize journal publication volume "without regard to quality" is a consequence of transformative agreements [@Goodin_2023].
@@
 The consortium also removed most hybrid journals from its Transformative Journal program in 2023 due to publishers' failure to meet self-defined open access growth targets [@Brainard_2023]. 
-In contrast, the German DEAL consortium signed a five-year transformative agreement with Elsevier starting in 2024, while also renewing its contracts with Springer Nature and Wiley until the end of 2028.
+In contrast, the German DEAL consortium announced a five-year transformative agreement with Elsevier starting in 2024, while also renewing its contracts with Springer Nature and Wiley until the end of 2028.
 At the same time, the Colombia Consortium signed the first transformative agreements in Latin America [@Mu_oz_V_lez_2024].
@@
 Particularly, data availability is a limiting factor when studying the impact of transformative agreements [@Bakker_2024], because bibliometric databases, even though many allow the retrieval of open access articles in hybrid journals, do not directly attribute them to specific transformative agreements.
+Likewise, article-level open access invoicing data is only partly available [@Jahn_2021].
 
 The present study aims to address these limitations by combining multiple openly available data sources to determine the open access uptake in hybrid journals, while distinguishing between open access through transformative agreements and other means.
-With this novel and open approach, this the first large-scale analysis aims to answer the following questions:
+With this novel and open approach, this first large-scale analysis aims to answer the following questions:
 
@@
 
-This study combines data from multiple publicly available data sources as diagrammatically shown in Figure \ref{fig:data_workflow}. 
+This study combines data from multiple publicly available data sources as shown in Figure \ref{fig:data_workflow}. 
 Initially, transformative agreement data retrieved from the cOAlition S Journal Checker Tool[^checkertool]  provided information about journal portfolios and participating institutions. 
@@
 Crossref was used for open access identification because transformative agreements workflows generally require publishers to deliver CC license information to this DOI registration agency [@Geschuhn_2017].
+
 Comparing Crossref license coverage with OpenAlex, which re-uses open access evidence from Unpaywall, a widely used open access discovery service that also parses journal websites for open content licenses [@Piwowar_2018], highlighted ongoing challenges to identify hybrid open access [@Mart_n_Mart_n_2018; @Jahn_2021; @Zhang_2022; @Butler_2023].
@@
 Finally, inconsistent open access status information in previous OpenAlex versions was observed [@jahn2023].
-After reporting it to OpenAlex, according to the release notes, fixing this issue was still ongoing, which might also explain part of the discrepancy.
+After reporting it to OpenAlex, fixing this issue was still ongoing according to the release notes, which might also explain part of the discrepancy.
 
@@
 
-## Data Analysis
+## Data analysis
 
@@
 Figure \ref{fig:results_overview}B highlights that the majority of hybrid open access was made available through transformative agreements in 2021 and 2022, contributing `r oa_ta_plot_df |> filter(cr_year == "2022", type == "ta_oa") |> pull(value) / oa_by_year |> filter(cr_year == "2022") |> pull(oa) * 100`% of the total open access article volume in 2022.
-However, there was also a notable growth in open access provision through other means, most likely publication fees being not invoiced through transformative agreements, which increased from `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> pull(value)`) in 2018 to  `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> pull(value)`).
+However, there was also a notable growth in open access provision through other means, presumably publication fees being not invoiced through transformative agreements, which increased from `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> pull(value)`) in 2018 to  `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> pull(value)`).
 <!-- This suggests that publishers were able to gain equally from individual and institutional open access publishing options.-->
@@
 
-Figure \ref{fig:publisher_figure} takes a closer look into the growth of hybrid open access across publishers by year with a focus on open articles enabled by transformative agreements. 
+Figure \ref{fig:publisher_figure} takes a closer look into the growth of hybrid open access across publishers by year with a focus on open access enabled by transformative agreements. 
 Although all publishers show a general long-term trend towards transformative agreements, Figure \ref{fig:publisher_figure}A and B indicate that, in particular, Wiley experienced a substantial increase in its open access share from `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(value) |> sum()`) in 2018 to `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(prop) |> sum()*100`%  (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()`) in 2022, representing an 4.5-fold increase.
@@
 From 2018 to 2022, the proportion of open access in hybrid journals increased from `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2018) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2018 to `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2022) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2022.
-On the other hand, BRICS recorded an low uptake, from `r country_bar$data |> filter(cat == "BRICS", oa_type != "closed", cr_year == 2018) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2018 to `r country_bar$data |> filter(cat == "BRICS", oa_type != "closed", cr_year == 2022) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2022.
-Despite rise of open access across OECD countries, the overall publication output decreased sharply, dropping to `r country_bar$data |> filter(cat == "OECD", cr_year == "2022") |> distinct(all) |> pull(all)` in 2022 after peaking `r country_bar$data |> filter(cat == "OECD", cr_year == "2020") |> distinct(all) |> pull(all)` articles in 2020.
+On the other hand, BRICS recorded a low uptake, from `r country_bar$data |> filter(cat == "BRICS", oa_type != "closed", cr_year == 2018) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2018 to `r country_bar$data |> filter(cat == "BRICS", oa_type != "closed", cr_year == 2022) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2022.
+Despite the rise of open access across OECD countries, the overall publication output decreased sharply, dropping to `r country_bar$data |> filter(cat == "OECD", cr_year == "2022") |> distinct(all) |> pull(all)` in 2022 after peaking `r country_bar$data |> filter(cat == "OECD", cr_year == "2020") |> distinct(all) |> pull(all)` articles in 2020.
 In stark contrast, the number of articles published in hybrid journals by first authors affiliated with institutions from BRICS countries increased steadily over the years, more than doubling from `r country_bar$data |> filter(cat == "BRICS", cr_year == "2018") |> distinct(all) |> pull(all)` in 2018 to `r country_bar$data |> filter(cat == "OECD", cr_year == "2022") |> distinct(all) |> pull(all)` in 2022.
@@
 With `r country_scatter_df |> filter(cr_year == "2022", country_code == "CN") |> pull(all)` articles, China was the most productive country, followed by the United States (`r country_scatter_df |> filter(cr_year == "2022", country_code == "US") |> pull(all)` articles) and India (`r country_scatter_df |> filter(cr_year == "2022", country_code == "IN") |> pull(all)` articles).
-In contrast, West and Nord European countries published a considerable high number of open access articles, mainly due to transformative agreements. 
+In contrast, West and Nord European countries published a considerable high number of open access articles. 
 Particularly, Germany, Great Britain, the Netherlands, Sweden, Switzerland and Spain recorded an above-average open access share as indicated by the linear trend line.
@@
 Figure \ref{fig:country_top_20_plot} illustrates the development of hybrid open access from 2018 to 2022, highlighting the top 20 most productive countries in terms of articles published in hybrid journals that were included in transformative agreements over the five-year period.
-Notably, The Netherlands (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "NL") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), Sweden (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "SE") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), Poland (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "PL") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%) and Great Britain (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "GB") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%)) exhibited a relatively high level of uptake in 2018 which continued to increase in the following years.
+Notably, The Netherlands (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "NL") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), Sweden (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "SE") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), Poland (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "PL") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%) and Great Britain (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "GB") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%) exhibited a relatively high level of uptake in 2018 which continued to increase in the following years.
 In 2022, Sweden had the highest proportion of open access relative to its publication output (`r country_plot$data |> filter(cr_year == "2022", country_name_fct == "SE") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), followed by the Netherlands (`r country_plot$data |> filter(cr_year == "2022", country_name_fct == "NL") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%) and Switzerland (`r country_plot$data |> filter(cr_year == "2022", country_name_fct == "CH") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), with these countries benefiting from transformative agreements.
@@
 For instance, Germany successfully negotiated an agreement with Elsevier not until 2023. 
-Additionally, publication limits or eligibility criteria for institutions and article types may explain why even countries with widespread agreement implementation do not achieve 100% hybrid open access.
+Additionally, publication limits or eligibility criteria for institutions and article types may explain why even countries with widespread agreement implementation like Sweden or the Nethterlands did not achieve 100% hybrid open access.
 Interestingly, in Japan and the US other options than transformative agreements were the main driver for the increase in hybrid open access.
@@
 The primary aim of this study was to investigate the adoption of open access in hybrid journals in transformative agreements, which were introduced as a temporal means to support the transition of subscription-based academic publishing to full open access.
-This study presents a novel approach based on open data, which leverages metadata on over 700 agreements and nine million journal articles to estimate the extent to which transformation agreement contribute to the transition of this journal business model based on institutional affiliation.
+This study presents a novel approach based on open data, which leverages metadata on over 700 agreements and nine million journal articles to estimate the extent to which transformation agreement contribute to the transition of this journal business model.
 The results highlight a strong growth in open access between 2018 and 2022, particularly driven by an increasing number of institutions that had transformative agreements in place. 
@@
 A key finding of the analysis is that transformative agreements increase market concentration. 
-Specifically, the three largest commercial publishers—Elsevier, Springer Nature, and Wiley—have increased their market share in the context of open access, particularly in regard to open access provided through transformative agreements.
+Specifically, the three largest commercial publishers Elsevier, Springer Nature, and Wiley dominate, particularly in regard to open access provided through transformative agreements.
 Together, the three publishers accounted for three-fourths of open access articles through transformative agreements, despite publishing less than half of the total publication volume published between 2018 and 2022.
@@
 However, it must be noted that the focus of transformative agreements on publishers with large hybrid journal portfolios is intentional [@ifla].
-Because of transformative agreements with a few large publishers, national consortia were able to substantially increased their country’s annual open access article volume [@Pinhasi_2021; @Huang_2020; @Taubert_2023; @Wenaas_2022].
+Because of transformative agreements with a few large publishers, national consortia were able to substantially increase their country’s annual open access article volume [@Bosman_2021; @Pinhasi_2021; @Huang_2020; @Taubert_2023; @Wenaas_2022].
 
@@
 Rather, this result can be more accurately attributed to the substantial proportion of authors from high-income countries who publish in these hybrid journals included in transformative agreements, particularly when journals are of local relevance, such as through the belonging to national societies or regional research scope.
-
  <!--   [funding options in the humanities and social sicences] -->
-Surprisingly, the total publication output of high-income countries belonging to the OECD declined susbtantially between 2018 and 2022, while the BRICS countries recorded an steady increase. 
-Because the BRICS expansion can be exclusivley attributed to closed access articles, this developement has the potential to hinder the transition of academic journal publishing to full open access through transformative agreements.
-In China, the most productive country according to the total article volume in hybrid journals, 
-(shu und zhang und das indische paper)
-
-other countries less represented and brazil diamond russia The Eastern Europe Expanded Collection
-
-but not coherent acorss all countries, in particular in the us. neverthelles china and us have susbtantial market shares when it comes to hybrid oa.
-
-There are at least three potential limitations
-concerning the results of this study. A first limitation
-concerns … A second potential limitation is that …
-
-These data have some potential intervention implications.
-forst and formesot, the supprot recent pocliy decision highlighting a grew in oa at country-level, but questioning the.
-if thouse contracts, a transiliteration is ttoo short and might require a longer transiltierional period.
-the growtrh in brics however would require too impleis unboundling, other payment means and focus on alternative, in particular innovations like preprinting, diamond oa
-
-In terms of future research, it would be useful to extend the current findings by examining other oa types, combine oa with mobility.
-fianncials tudies are also required to assess claims of cost-sensitive, here, publisher and publsiher are required to increase transparency.
-at the same time, studies needed to assess these claim economically.
-data can be improved to entail more, in apritcular consortia could start to disclose funded articles as , with openapc already a cwors-orucing platform available.
-also libraries can contrbute to an improved catalogiuing of transformative agreements including participating insitutions, which is aligned with standards used in schoalrly communications like ror id.
-
-In summary, after more than tow decades, hybrid open access has not demonstrated a succesful menas.
-stduying transofrmative agreemnts, effects can be found at the -level of participating countries, but overlla uptake remains low, in particular with regard to publications from brocs countries, which country share rise, exclusively on the expense of closed access.
-means
+
+Surprisingly, the total publication output of high-income countries belonging to the OECD in hybrid journals declined substantially after peaking in 2020, whereas that of BRICS countries doubled from 2018 to 2022.
+Because the BRICS expansion can be exclusively attributed to closed access articles, this development has the potential to hinder the transition of academic journal publishing to full open access through transformative agreements and demands discussion.
+In China, the country with the highest volume of articles in hybrid journals, limited research funding to pay for open access is available, with expenditures for open access publishing already surpassing subscription costs [@Shu_2023].
+Furthermore, the focus of Chinese authors on established journals may contribute to this trend [@Zhang_2022]. 
+Although China supports Plan S [@Schiermeier_2018], this is not reflected in the data in terms of open access uptake in hybrid journals.
+As highlighted by @Koley_2023, India faces similar challenges in terms of resources to pay for open access.
+At the same time, access to research literature is a pressing, which is adressed by the "Indian one nation, one subscription" policy proposal. 
+However, this policy focuses on centrally negotiated subscriptions and does not entail open access.
+
+But open access uptake also differs among OECD countries. 
+In the United States, for example, hybrid open access including transformative agreements plays a lesser role compared to some European countries. 
+The smallest share of federally funded research was hybrid open access between 2017 and 2021 [@Schares_2023]. 
+While some university consortia like the California Digital Library signed transformative agreements, others attempt to depart from big deals and unbundle large journal portfolios to address cost increases [@Brainard_2021; @Schares_2022].
+Despite the relatively low penetration of transformative agreements in China and the United States compared to Europe, 22% of open access in hybrid journals in 2022 originates from these two countries, indicating the availability of funding sources for publication fees.
+
+This large-scale study provides first empricial evidence about the influence of transformative agreements on the transition of hybrid journals to full open access. 
+However, several limitations needs to be acknowledged.
+From a data perspective, estimations of open access through transformative agreements were established by linking first author affiliations with publicly available agreement data from cOAlition S, and not through invoicing data because these data a usually not shared by publishers.
+Moreover, this study is unable to account for the various types of transformative agreements, particularly with regards to article types and  capping.
+Similarly, assessing the quality of the OpenAlex and Crossref data used, particularly in terms of affiliations and article types, in comparison with established bibliometric databases was beyond the scope of this analysis.
+Overall, the methodology is designed in such a way that the results underestimate rather than overestimate the adoption of open access through transformative agreements. 
+In addition, the data used in this study are openly available, along with the code used for this analysis.
+
+Additionally, it must be noted that the study period was significantly impacted by the COVID-19 pandemic, which led to an unprecedented number of publications as well as restrictions on migration flows, which could have influenced the country-specific results [@Aviv_Reuven_2021; @Zhao_2023].
+Furthermore, the study design did not consider emerging publication practices such as preprints [@Fraser_2021] and questionable special issues [@hanson2023strain], which have grown particularly rapidly since 2020.
+Lastly, it should be emphasized that the study aims to estimate articles enabled through transformative agreements but does not address financial shifts due to a lack of data on expenditures.
+
+This present study allows multiple strands of further research. 
+One is to complement this large-scale study with more specific evidence from individual countries or subjects, particularly those with low hybrid open access  rates.
+Incorporating full open access and subscription-based journals, as well as considering global trends in scholarly migration and collaboration could also be promising.
+Financial studies could build upon the presented study design and incorporate subscription and open access expenditures to evaluate the cost-effectiveness of transformative agreements.
+
+But the study also has practival implications for research funders and libraries.
+Of primary concern should be observed differences across countries, in particular the relationship between economic development and open access adoption.
+From a data perspective, the reporting of open access funding is not harmonised, but often crowd-sourced from various sources.
+To improve the assessment of transformative agreements, libraries and publishers could contribute by providing information about respective journal portfolios, participating institutions and open access invoicing.
+
+In summary, the aim of this study was to provide empirical insights into the developement of hybrid open access following the introduction of transformative agreements. 
+The results are important for both researchers and stakeholders engaged in negotiating and evaluating these agreements. 
+The presented approach relies on open data, which enables follow-up studies and open access monitoring acitvties to further explore the role of transformative agreements in facilitating the shift towards full open access in academic publishing.
+
+-->
 
@@
 [^agreements]: In this paper I use the term "transformative agreement", addressing also offsetting, read-and-publish or publish-and-read deals, and other variants [@Borrego_2020; @Hinchliffe_2019].
-Although the term is critised as misleading and not useful to describe these agreements between library consortia and publishers [@budapest], it is widely used in policy discussions and in the research literature.
+Although the term is critised as misleading and not useful to describe the different types of open access agreements between library consortia and commercial publishers [@budapest], it is widely used in policy discussions and in the research literature.

### 2024-02-22 analysis/manuscript.tex
first draft discussion

@@
 period 2014-2017. In particular, British (17\%), Austrian (15\%) and
-Dutch (13\%) universities contributed to this trend. However, this shift
-in funding policy towards hybrid open access also added to the overall
-cost of publishing, which includes subscription spending and the
-administrative efforts required to handle payments (Pinfield et al.,
-2016). Moreover, established large commercial publishers, which already
-dominated the publishing market (Larivière et al., 2015),
-disproportionately benefited from hybrid open access funding in
-comparison to full open access publishers (Butler et al., 2023; Jahn \&
-Tullney, 2016; Shu \& Larivière, 2023).
+Dutch (13\%) universities stand out. However, this shift in funding
+policy towards hybrid open access also added to the overall cost of
+publishing, which includes subscription spending and the administrative
+efforts required to handle payments (Pinfield et al., 2016). Moreover,
+established large commercial publishers, which already dominated the
+publishing market (Larivière et al., 2015), disproportionately benefited
+from hybrid open access funding in comparison to full open access
+publishers (Butler et al., 2023; Jahn \& Tullney, 2016; Shu \&
+Larivière, 2023).
 
@@
 in which well-established commercial publishers gain twice from reading
-and open access publishing, as well as to increase publisher-provided
-immediate open access (Björk \& Solomon, 2014; Schimmer et al., 2015).
-These considerations resulted into transformative agreements\footnote{In
-  this paper I use the term ``transformative agreement'', addressing
-  also offsetting, read-and-publish or publish-and-read deals, and other
-  variants (Borrego et al., 2020; Hinchliffe, 2019). Although the term
-  is critised as misleading and not useful to describe these agreements
-  between library consortia and publishers (Babini et al., 2022), it is
-  widely used in policy discussions and in the research literature.},
-which cover a broad range of contracts between library consortia and
-publishers from mid-2010s onwards where institutional spending for
-subscriptions and open access publishing are considered together
-(Borrego et al., 2020; Hinchliffe, 2019). Transformative agreements seek
-to control costs while allowing a transitional phase for publishing more
-open access articles. Similar to big deals, transformative agreements
-mainly bundle hybrid and subscription-only journals from commercial
-publishers, but aim at a higher degree of transparency than previous big
-deals, where contracts including payments were confidential (Bergstrom
-et al., 2014).
+and open access publishing (Mittermaier, 2015), as well as to increase
+publisher-provided immediate open access (Björk \& Solomon, 2014;
+Schimmer et al., 2015). These considerations resulted into
+transformative agreements\footnote{In this paper I use the term
+  ``transformative agreement'', addressing also offsetting,
+  read-and-publish or publish-and-read deals, and other variants
+  (Borrego et al., 2020; Hinchliffe, 2019). Although the term is
+  critised as misleading and not useful to describe the different types
+  of open access agreements between library consortia and commercial
+  publishers (Babini et al., 2022), it is widely used in policy
+  discussions and in the research literature.}, which cover a broad
+range of contracts between library consortia and publishers from
+mid-2010s onwards where institutional spending for subscriptions and
+open access publishing are considered together (Borrego et al., 2020;
+Hinchliffe, 2019). Transformative agreements seek to control costs while
+allowing a transitional phase for publishing more open access articles.
+Similar to big deals, transformative agreements mainly bundle hybrid and
+subscription-only journals from commercial publishers, but aim at a
+higher degree of transparency than previous big deals, where contracts
+including payments were confidential (Bergstrom et al., 2014).
 
@@
 al., 2021). The focus on large commercial publishers might also increase
-inequality, because transformative agreements focus on pay to publish
-open access mainly targets institutions from high-income countries,
-furthering a questionable journal prestige culture (Babini et al.,
-2022). Besides, an editorial-board resignation raised concerns that
-publishers' desire to maximize journal publication volume ``without
-regard to quality'' is a consequence of transformative agreements
-(Rasmussen, 2023).
+inequality (Klebel \& Ross-Hellauer, 2023), because transformative
+agreements focus on pay to publish open access mainly targets
+institutions from high-income countries, furthering a questionable
+journal prestige culture (Babini et al., 2022). Besides, an
+editorial-board resignation raised concerns that publishers' desire to
+maximize journal publication volume ``without regard to quality'' is a
+consequence of transformative agreements (Rasmussen, 2023).
 
@@
 open access growth targets (Brainard, 2023). In contrast, the German
-DEAL consortium signed a five-year transformative agreement with
+DEAL consortium announced a five-year transformative agreement with
 Elsevier starting in 2024, while also renewing its contracts with
@@
 in hybrid journals, do not directly attribute them to specific
-transformative agreements.
+transformative agreements. Likewise, article-level open access invoicing
+data is only partly available (Jahn et al., 2022).
 
@@
 through transformative agreements and other means. With this novel and
-open approach, this the first large-scale analysis aims to answer the
+open approach, this first large-scale analysis aims to answer the
 following questions:
@@
 This study combines data from multiple publicly available data sources
-as diagrammatically shown in Figure \ref{fig:data_workflow}. Initially,
-transformative agreement data retrieved from the cOAlition S Journal
-Checker Tool\footnote{\url{https://www.coalition-s.org/blog/enabling-accurate-results-within-the-journal-checker-tool/}}
+as shown in Figure \ref{fig:data_workflow}. Initially, transformative
+agreement data retrieved from the cOAlition S Journal Checker
+Tool\footnote{\url{https://www.coalition-s.org/blog/enabling-accurate-results-within-the-journal-checker-tool/}}
 provided information about journal portfolios and participating
@@
 deliver CC license information to this DOI registration agency (Geschuhn
-\& Stone, 2017). Comparing Crossref license coverage with OpenAlex,
-which re-uses open access evidence from Unpaywall, a widely used open
-access discovery service that also parses journal websites for open
-content licenses (Piwowar et al., 2018), highlighted ongoing challenges
-to identify hybrid open access (Butler et al., 2023; Jahn et al., 2022;
+\& Stone, 2017).
+
+Comparing Crossref license coverage with OpenAlex, which re-uses open
+access evidence from Unpaywall, a widely used open access discovery
+service that also parses journal websites for open content licenses
+(Piwowar et al., 2018), highlighted ongoing challenges to identify
+hybrid open access (Butler et al., 2023; Jahn et al., 2022;
 Martín-Martín et al., 2018; Zhang et al., 2022). Here, 742,369 articles
@@
 open access status information in previous OpenAlex versions was
-observed (Jahn et al., 2023). After reporting it to OpenAlex, according
-to the release notes, fixing this issue was still ongoing, which might
+observed (Jahn et al., 2023). After reporting it to OpenAlex, fixing
+this issue was still ongoing according to the release notes, which might
 also explain part of the discrepancy.
@@
 \hypertarget{data-analysis}{%
-\subsection{Data Analysis}\label{data-analysis}}
+\subsection{Data analysis}\label{data-analysis}}
 
@@
 volume in 2022. However, there was also a notable growth in open access
-provision through other means, most likely publication fees being not
+provision through other means, presumably publication fees being not
 invoiced through transformative agreements, which increased from 4.1\%
@@
 Figure \ref{fig:publisher_figure} takes a closer look into the growth of
-hybrid open access across publishers by year with a focus on open
-articles enabled by transformative agreements. Although all publishers
-show a general long-term trend towards transformative agreements, Figure
+hybrid open access across publishers by year with a focus on open access
+enabled by transformative agreements. Although all publishers show a
+general long-term trend towards transformative agreements, Figure
 \ref{fig:publisher_figure}A and B indicate that, in particular, Wiley
@@
 proportion of open access in hybrid journals increased from 6.1\% in
-2018 to 26\% in 2022. On the other hand, BRICS recorded an low uptake,
+2018 to 26\% in 2022. On the other hand, BRICS recorded a low uptake,
 from 1.6\% in 2018 to 3.7\% in 2022.\\
-Despite rise of open access across OECD countries, the overall
+Despite the rise of open access across OECD countries, the overall
 publication output decreased sharply, dropping to 786,903 in 2022 after
@@
 articles). In contrast, West and Nord European countries published a
-considerable high number of open access articles, mainly due to
-transformative agreements. Particularly, Germany, Great Britain, the
-Netherlands, Sweden, Switzerland and Spain recorded an above-average
-open access share as indicated by the linear trend line. As represented
-by the point size, as well as it can been seen in Figure
-\ref{fig:country_patch}C, transformative agreements contributed to these
-market positions. Interestingly, the United States had a notable open
-access market share of 15\%, although transformative agreements
-contributed to a lesser extent. Similarly, China's open access market
-share of 7.2\% in 2022 was comparable to that of the Netherlands, which
-was 7.1\%.
+considerable high number of open access articles. Particularly, Germany,
+Great Britain, the Netherlands, Sweden, Switzerland and Spain recorded
+an above-average open access share as indicated by the linear trend
+line. As represented by the point size, as well as it can been seen in
+Figure \ref{fig:country_patch}C, transformative agreements contributed
+to these market positions. Interestingly, the United States had a
+notable open access market share of 15\%, although transformative
+agreements contributed to a lesser extent. Similarly, China's open
+access market share of 7.2\% in 2022 was comparable to that of the
+Netherlands, which was 7.1\%.
 
@@
 period. Notably, The Netherlands (27\%), Sweden (24\%), Poland (17\%)
-and Great Britain (17\%)) exhibited a relatively high level of uptake in
+and Great Britain (17\%) exhibited a relatively high level of uptake in
 2018 which continued to increase in the following years. In 2022, Sweden
@@
 or eligibility criteria for institutions and article types may explain
-why even countries with widespread agreement implementation do not
-achieve 100\% hybrid open access. Interestingly, in Japan and the US
-other options than transformative agreements were the main driver for
-the increase in hybrid open access. Once again, the graph highlights
-countries with low hybrid open access, particularly non-OECD countries,
-where only a few or no agreements were in place.
+why even countries with widespread agreement implementation like Sweden
+or the Nethterlands did not achieve 100\% hybrid open access.
+Interestingly, in Japan and the US other options than transformative
+agreements were the main driver for the increase in hybrid open access.
+Once again, the graph highlights countries with low hybrid open access,
+particularly non-OECD countries, where only a few or no agreements were
+in place.
 
@@
 extent to which transformation agreement contribute to the transition of
-this journal business model based on institutional affiliation. The
-results highlight a strong growth in open access between 2018 and 2022,
-particularly driven by an increasing number of institutions that had
-transformative agreements in place. Yet, the majority of research
-literature published in hybrid journals in this five-year period
-remained behind a publisher's paywall. Growth in the adoption of open
-access in hybrid journals, in particular through transformative
-agreements, can be largely attributed to three large commercial
-publishers -- Elsevier, Springer Nature, and Wiley -- but varies
-substantially across journals, publishers, subjects, and country
+this journal business model. The results highlight a strong growth in
+open access between 2018 and 2022, particularly driven by an increasing
+number of institutions that had transformative agreements in place. Yet,
+the majority of research literature published in hybrid journals in this
+five-year period remained behind a publisher's paywall. Growth in the
+adoption of open access in hybrid journals, in particular through
+transformative agreements, can be largely attributed to three large
+commercial publishers -- Elsevier, Springer Nature, and Wiley -- but
+varies substantially across journals, publishers, subjects, and country
 affiliations. Despite the limitations of the data used, the findings
@@
 market concentration. Specifically, the three largest commercial
-publishers---Elsevier, Springer Nature, and Wiley---have increased their
-market share in the context of open access, particularly in regard to
-open access provided through transformative agreements. Together, the
-three publishers accounted for three-fourths of open access articles
-through transformative agreements, despite publishing less than half of
-the total publication volume published between 2018 and 2022. This
-observation aligns with previous research on shifts in the publishing
-market following the introduction of funding opportunities for hybrid
-open access (Butler et al., 2023; Jahn \& Tullney, 2016; Shu \&
-Larivière, 2023). Additionally, the results confirm previously observed
-variations by publisher, with Elsevier exhibiting a different
-development than Springer Nature and Wiley (Butler et al., 2023; Jahn et
-al., 2022). However, it must be noted that the focus of transformative
-agreements on publishers with large hybrid journal portfolios is
-intentional (Campbell et al., 2022). Because of transformative
-agreements with a few large publishers, national consortia were able to
-substantially increased their country's annual open access article
-volume (Huang et al., 2020; Pinhasi et al., 2021; Taubert et al., 2023;
-Wenaas, 2022).
+publishers Elsevier, Springer Nature, and Wiley dominate, particularly
+in regard to open access provided through transformative agreements.
+Together, the three publishers accounted for three-fourths of open
+access articles through transformative agreements, despite publishing
+less than half of the total publication volume published between 2018
+and 2022. This observation aligns with previous research on shifts in
+the publishing market following the introduction of funding
+opportunities for hybrid open access (Butler et al., 2023; Jahn \&
+Tullney, 2016; Shu \& Larivière, 2023). Additionally, the results
+confirm previously observed variations by publisher, with Elsevier
+exhibiting a different development than Springer Nature and Wiley
+(Butler et al., 2023; Jahn et al., 2022). However, it must be noted that
+the focus of transformative agreements on publishers with large hybrid
+journal portfolios is intentional (Campbell et al., 2022). Because of
+transformative agreements with a few large publishers, national
+consortia were able to substantially increase their country's annual
+open access article volume (Bosman, Jonge, et al., 2021; Huang et al.,
+2020; Pinhasi et al., 2021; Taubert et al., 2023; Wenaas, 2022).
 
@@
 majority of full open access journals are so-called ``Diamond OA
-journals'' that do not charge publication fees (Bosman et al., 2021).
-Rather, this result can be more accurately attributed to the substantial
-proportion of authors from high-income countries who publish in these
-hybrid journals included in transformative agreements, particularly when
-journals are of local relevance, such as through the belonging to
-national societies or regional research scope.
+journals'' that do not charge publication fees (Bosman, Frantsvåg, et
+al., 2021). Rather, this result can be more accurately attributed to the
+substantial proportion of authors from high-income countries who publish
+in these hybrid journals included in transformative agreements,
+particularly when journals are of local relevance, such as through the
+belonging to national societies or regional research scope.
 
 Surprisingly, the total publication output of high-income countries
-belonging to the OECD declined susbtantially between 2018 and 2022,
-while the BRICS countries recorded an steady increase. Because the BRICS
-expansion can be exclusivley attributed to closed access articles, this
-developement has the potential to hinder the transition of academic
-journal publishing to full open access through transformative
-agreements. In China, the most productive country according to the total
-article volume in hybrid journals, (shu und zhang und das indische
-paper)
-
-other countries less represented and brazil diamond russia The Eastern
-Europe Expanded Collection
-
-but not coherent acorss all countries, in particular in the us.
-neverthelles china and us have susbtantial market shares when it comes
-to hybrid oa.
-
-There are at least three potential limitations concerning the results of
-this study. A first limitation concerns \ldots{} A second potential
-limitation is that \ldots{}
-
-These data have some potential intervention implications. forst and
-formesot, the supprot recent pocliy decision highlighting a grew in oa
-at country-level, but questioning the. if thouse contracts, a
-transiliteration is ttoo short and might require a longer
-transiltierional period. the growtrh in brics however would require too
-impleis unboundling, other payment means and focus on alternative, in
-particular innovations like preprinting, diamond oa
-
-In terms of future research, it would be useful to extend the current
-findings by examining other oa types, combine oa with mobility.
-fianncials tudies are also required to assess claims of cost-sensitive,
-here, publisher and publsiher are required to increase transparency. at
-the same time, studies needed to assess these claim economically. data
-can be improved to entail more, in apritcular consortia could start to
-disclose funded articles as , with openapc already a cwors-orucing
-platform available. also libraries can contrbute to an improved
-catalogiuing of transformative agreements including participating
-insitutions, which is aligned with standards used in schoalrly
-communications like ror id.
-
-In summary, after more than tow decades, hybrid open access has not
-demonstrated a succesful menas. stduying transofrmative agreemnts,
-effects can be found at the -level of participating countries, but
-overlla uptake remains low, in particular with regard to publications
-from brocs countries, which country share rise, exclusively on the
-expense of closed access. means
+belonging to the OECD in hybrid journals declined substantially after
+peaking in 2020, whereas that of BRICS countries doubled from 2018 to
+2022. Because the BRICS expansion can be exclusively attributed to
+closed access articles, this development has the potential to hinder the
+transition of academic journal publishing to full open access through
+transformative agreements and demands discussion. In China, the country
+with the highest volume of articles in hybrid journals, limited research
+funding to pay for open access is available, with expenditures for open
+access publishing already surpassing subscription costs (Shu \&
+Larivière, 2023). Furthermore, the focus of Chinese authors on
+established journals may contribute to this trend (Zhang et al., 2022).
+Although China supports Plan S (Schiermeier, 2018), this is not
+reflected in the data in terms of open access uptake in hybrid journals.
+As highlighted by Koley \& Lala (2023), India faces similar challenges
+in terms of resources to pay for open access. At the same time, access
+to research literature is a pressing, which is adressed by the ``Indian
+one nation, one subscription'' policy proposal. However, this policy
+focuses on centrally negotiated subscriptions and does not entail open
+access.
+
+But open access uptake also differs among OECD countries. In the United
+States, for example, hybrid open access including transformative
+agreements plays a lesser role compared to some European countries. The
+smallest share of federally funded research was hybrid open access
+between 2017 and 2021 (Schares, 2023). While some university consortia
+like the California Digital Library signed transformative agreements,
+others attempt to depart from big deals and unbundle large journal
+portfolios to address cost increases (Brainard, 2021; Schares, 2022).
+Despite the relatively low penetration of transformative agreements in
+China and the United States compared to Europe, 22\% of open access in
+hybrid journals in 2022 originates from these two countries, indicating
+the availability of funding sources for publication fees.
+
+This large-scale study provides first empricial evidence about the
+influence of transformative agreements on the transition of hybrid
+journals to full open access. However, several limitations needs to be
+acknowledged. From a data perspective, estimations of open access
+through transformative agreements were established by linking first
+author affiliations with publicly available agreement data from
+cOAlition S, and not through invoicing data because these data a usually
+not shared by publishers. Moreover, this study is unable to account for
+the various types of transformative agreements, particularly with
+regards to article types and capping. Similarly, assessing the quality
+of the OpenAlex and Crossref data used, particularly in terms of
+affiliations and article types, in comparison with established
+bibliometric databases was beyond the scope of this analysis. Overall,
+the methodology is designed in such a way that the results underestimate
+rather than overestimate the adoption of open access through
+transformative agreements. In addition, the data used in this study are
+openly available, along with the code used for this analysis.
+
+Additionally, it must be noted that the study period was significantly
+impacted by the COVID-19 pandemic, which led to an unprecedented number
+of publications as well as restrictions on migration flows, which could
+have influenced the country-specific results (Aviv-Reuven \& Rosenfeld,
+2021; Zhao et al., 2023). Furthermore, the study design did not consider
+emerging publication practices such as preprints (Fraser et al., 2021)
+and questionable special issues (Hanson et al., 2023), which have grown
+particularly rapidly since 2020. Lastly, it should be emphasized that
+the study aims to estimate articles enabled through transformative
+agreements but does not address financial shifts due to a lack of data
+on expenditures.
+
+This present study allows multiple strands of further research. One is
+to complement this large-scale study with more specific evidence from
+individual countries or subjects, particularly those with low hybrid
+open access rates. Incorporating full open access and subscription-based
+journals, as well as considering global trends in scholarly migration
+and collaboration could also be promising. Financial studies could build
+upon the presented study design and incorporate subscription and open
+access expenditures to evaluate the cost-effectiveness of transformative
+agreements.
+
+But the study also has practival implications for research funders and
+libraries. Of primary concern should be observed differences across
+countries, in particular the relationship between economic development
+and open access adoption. From a data perspective, the reporting of open
+access funding is not harmonised, but often crowd-sourced from various
+sources. To improve the assessment of transformative agreements,
+libraries and publishers could contribute by providing information about
+respective journal portfolios, participating institutions and open
+access invoicing.
+
+In summary, the aim of this study was to provide empirical insights into
+the developement of hybrid open access following the introduction of
+transformative agreements. The results are important for both
+researchers and stakeholders engaged in negotiating and evaluating these
+agreements. The presented approach relies on open data, which enables
+follow-up studies and open access monitoring acitvties to further
+explore the role of transformative agreements in facilitating the shift
+towards full open access in academic publishing.
+
+--\textgreater{}
 
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-Aviv_Reuven_2021}{}}%
+Aviv-Reuven, S., \& Rosenfeld, A. (2021). Publication patterns' changes
+due to the COVID-19 pandemic: A longitudinal and short-term
+scientometric analysis. \emph{Scientometrics}, \emph{126}(8),
+6761--6784. \url{https://doi.org/10.1007/s11192-021-04059-x}
+
 \leavevmode\vadjust pre{\hypertarget{ref-budapest}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Bosman_2021}{}}%
+Bosman, J., Jonge, H. de, Kramer, B., \& Sondervan, J. (2021). Advancing
+open access in the netherlands after 2020: From quantity to quality.
+\emph{Insights the UKSG Journal}, \emph{34}.
+\url{https://doi.org/10.1629/uksg.545}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Brainard_2021}{}}%
+Brainard, J. (2021). California universities and elsevier make up, ink
+big open-access deal. \emph{Science}.
+\url{https://doi.org/10.1126/science.abi5505}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Brainard_2023}{}}%
 Brainard, J. (2023). {`Transformative'} journals get booted for
-switching to open access too slowly. In \emph{Science: SCIENCEINSIDER}.
+switching to open access too slowly. In \emph{Science}.
 \url{https://doi.org/10.1126/science.adj3282}
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Fraser_2021}{}}%
+Fraser, N., Brierley, L., Dey, G., Polka, J. K., Pálfy, M., Nanni, F.,
+\& Coates, J. A. (2021). The evolving role of preprints in the
+dissemination of COVID-19 research and their impact on the science
+communication landscape. \emph{PLOS Biology}, \emph{19}(4), e3000959.
+\url{https://doi.org/10.1371/journal.pbio.3000959}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Fraser_2023}{}}%
@@
 \url{https://doi.org/10.1629/uksg.391}
+
+\leavevmode\vadjust pre{\hypertarget{ref-hanson2023strain}{}}%
+Hanson, M. A., Barreiro, P. G., Crosetto, P., \& Brockington, D. (2023).
+\emph{The strain on scientific publishing}.
+\url{https://arxiv.org/abs/2309.15884}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Klebel_2023}{}}%
+Klebel, T., \& Ross-Hellauer, T. (2023). The APC-barrier and its effect
+on stratification in open access publishing. \emph{Quantitative Science
+Studies}, \emph{4}(1), 22--43. \url{https://doi.org/10.1162/qss_a_00245}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Koley_2023}{}}%
+Koley, M., \& Lala, K. (2023). Limitations of the {``indian one nation,
+one subscription''} policy proposal and a way forward. \emph{Journal of
+Librarianship and Information Science}, 096100062211467.
+\url{https://doi.org/10.1177/09610006221146771}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Laakso_2016}{}}%
@@
 \emph{7}(2), 23. \url{https://doi.org/10.3390/publications7020023}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Mittermaier_2015}{}}%
+Mittermaier, B. (2015). Double dipping in hybrid open access -- chimera
+or reality? \emph{{ScienceOpen} Research}.
+\url{https://doi.org/10.14293/s2199-1006.1.sor-socsci.aowntu.v1}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Schares_2022}{}}%
+Schares, E. (2022). Unsub extender: A python-based web application for
+visualizing unsub data. \emph{Quantitative Science Studies},
+\emph{3}(3), 600--623. \url{https://doi.org/10.1162/qss_a_00200}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Schares_2023}{}}%
+Schares, E. (2023). Impact of the 2022 OSTP memo: A bibliometric
+analysis of US federally funded publications, 2017--2021.
+\emph{Quantitative Science Studies}, \emph{4}(1), 1--21.
+\url{https://doi.org/10.1162/qss_a_00237}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Schiermeier_2018}{}}%
+Schiermeier, Q. (2018). China backs bold plan to tear down journal
+paywalls. \emph{Nature}, \emph{564}(7735), 171--172.
+\url{https://doi.org/10.1038/d41586-018-07659-5}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Schiltz_2018}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Zhao_2023}{}}%
+Zhao, X., Akbaritabar, A., Kashyap, R., \& Zagheni, E. (2023). A gender
+perspective on the global migration of scholars. \emph{Proceedings of
+the National Academy of Sciences}, \emph{120}(10).
+\url{https://doi.org/10.1073/pnas.2214664120}
+
 \end{CSLReferences}

### 2024-02-22 analysis/manuscript.Rmd
Shared version for internal review

@@
 Specifically, the three largest commercial publishers Elsevier, Springer Nature, and Wiley dominate, particularly in regard to open access provided through transformative agreements.
-Together, the three publishers accounted for three-fourths of open access articles through transformative agreements, despite publishing less than half of the total publication volume published between 2018 and 2022.
+Together, the three publishers accounted for three-fourths of open access articles through transformative agreements, while recording less than half of total publication volume published between 2018 and 2022.
 This observation aligns with previous research on shifts in the publishing market following the introduction of funding opportunities for hybrid open access [@Jahn_2016; @Butler_2023; @Shu_2023]. 
@@
 In the United States, for example, hybrid open access including transformative agreements plays a lesser role compared to some European countries. 
-The smallest share of federally funded research was hybrid open access between 2017 and 2021 [@Schares_2023]. 
-While some university consortia like the California Digital Library signed transformative agreements, others attempt to depart from big deals and unbundle large journal portfolios to address cost increases [@Brainard_2021; @Schares_2022].
+Between 2017 and 2021, hybrid open access contributed the least to make federally funded research articles openly available [@Schares_2023]. 
+While some university consortia like the California Digital Library have signed transformative agreements, others attempt to depart from big deals and unbundle large journal portfolios to address cost increases [@Brainard_2021; @Schares_2022].
 Despite the relatively low penetration of transformative agreements in China and the United States compared to Europe, 22% of open access in hybrid journals in 2022 originates from these two countries, indicating the availability of funding sources for publication fees.
@@
 However, several limitations needs to be acknowledged.
-From a data perspective, estimations of open access through transformative agreements were established by linking first author affiliations with publicly available agreement data from cOAlition S, and not through invoicing data because these data a usually not shared by publishers.
+From a data perspective, estimations of open access through transformative agreements were established by linking first author affiliations with publicly available agreement data from cOAlition S, and not through invoicing data, which is usually not shared.
 Moreover, this study is unable to account for the various types of transformative agreements, particularly with regards to article types and capping.
-Similarly, assessing the quality of the OpenAlex and Crossref data used, particularly in terms of affiliations and article types, in comparison with established bibliometric databases was beyond the scope of this analysis.
+Furthermore, assessing the quality of the OpenAlex and Crossref data used, particularly in terms of affiliations and article types, by comparing it with established bibliometric databases such as Scopus or Web of Science was beyond the scope of this analysis.
 Overall, the methodology is designed in such a way that the results underestimate rather than overestimate the adoption of open access through transformative agreements. 
@@
 
-Additionally, it must be noted that the study period was significantly impacted by the COVID-19 pandemic, which led to an unprecedented number of publications as well as restrictions on migration flows, which could have influenced the country-specific results [@Aviv_Reuven_2021; @Zhao_2023].
-Furthermore, the study design did not consider emerging publication practices such as preprints [@Fraser_2021] and questionable special issues [@hanson2023strain], which have grown particularly rapidly since 2020.
-Lastly, it should be emphasized that the study aims to estimate articles enabled through transformative agreements but does not address financial shifts due to a lack of data on expenditures.
+Additionally, it must be noted that the study period was significantly impacted by the COVID-19 pandemic, which led to an unprecedented number of publications as well as an reduction of international collaboration [@Aviv_Reuven_2021], which could explain the observed contrasting developements in OECD and BRICS countries.
+Even before the pandemic, growth in publications in Europe were only due to internationally co-authored journal articles [@Kwiek_2020]. 
+Likewise, inflows from China to the US and European countries already declined before 2020 [@Zhao_2023].
+Furthermore, the study design did not consider emerging publication practices such as preprints [@Fraser_2021] and special issues [@hanson2023strain], which have grown rapidly since 2020. 
+Lastly, it should be emphasized that the study did not address financial shifts between suscriptiuons spending and open acces spayemnts while analysing hybrid open access through transformative agreements due to a lack of data on expenditures.
 
@@
 
-But the study also has practival implications for research funders and libraries.
-Of primary concern should be observed differences across countries, in particular the relationship between economic development and open access adoption.
-From a data perspective, the reporting of open access funding is not harmonised, but often crowd-sourced from various sources.
-To improve the assessment of transformative agreements, libraries and publishers could contribute by providing information about respective journal portfolios, participating institutions and open access invoicing.
-
-In summary, the aim of this study was to provide empirical insights into the developement of hybrid open access following the introduction of transformative agreements. 
+This study also has practival implications for research funders and libraries.
+Of primary concern should be the observed differences across countries, in particular the relationship between economic development and open access adoption.
+From a data perspective, the reporting of open access funding including transformative agreements is not harmonised, but often crowd-sourced from various sources.
+To improve the assessment of transformative agreements, libraries and publishers could agree on standards and services to publicly share information about respective journal portfolios, participating institutions and open access invoicing.
+
+In summary, this study provides empirical insights into the developement of hybrid open access following the introduction of transformative agreements. 
 The results are important for both researchers and stakeholders engaged in negotiating and evaluating these agreements. 
-The presented approach relies on open data, which enables follow-up studies and open access monitoring acitvties to further explore the role of transformative agreements in facilitating the shift towards full open access in academic publishing.
-
--->
+The presented approach relies on open data, which enables follow-up studies and open access monitoring acitvties to further explore the role of transformative agreements in transitioning academic publishing to full and equitable open access.
 

### 2024-02-22 analysis/manuscript.tex
Shared version for internal review

@@
   read-and-publish or publish-and-read deals, and other variants
-  (Borrego et al., 2020; Hinchliffe, 2019). Although the term is
+  (Borrego et al., 2021; Hinchliffe, 2019). Although the term is
   critised as misleading and not useful to describe the different types
@@
 mid-2010s onwards where institutional spending for subscriptions and
-open access publishing are considered together (Borrego et al., 2020;
+open access publishing are considered together (Borrego et al., 2021;
 Hinchliffe, 2019). Transformative agreements seek to control costs while
@@
 articles including review and other types of journal content, which are
-often not covered by transformative agreements (Borrego et al., 2020),
+often not covered by transformative agreements (Borrego et al., 2021),
 only articles published in regular issues indicated by non-numeric
@@
 considered to be responsible to arrange open access publication (Borrego
-et al., 2020; Geschuhn \& Stone, 2017; Schimmer et al., 2015). Here,
+et al., 2021; Geschuhn \& Stone, 2017; Schimmer et al., 2015). Here,
 country and institutional affiliations were retrieved from OpenAlex.
@@
 Together, the three publishers accounted for three-fourths of open
-access articles through transformative agreements, despite publishing
-less than half of the total publication volume published between 2018
-and 2022. This observation aligns with previous research on shifts in
-the publishing market following the introduction of funding
-opportunities for hybrid open access (Butler et al., 2023; Jahn \&
-Tullney, 2016; Shu \& Larivière, 2023). Additionally, the results
-confirm previously observed variations by publisher, with Elsevier
-exhibiting a different development than Springer Nature and Wiley
-(Butler et al., 2023; Jahn et al., 2022). However, it must be noted that
-the focus of transformative agreements on publishers with large hybrid
-journal portfolios is intentional (Campbell et al., 2022). Because of
-transformative agreements with a few large publishers, national
-consortia were able to substantially increase their country's annual
-open access article volume (Bosman, Jonge, et al., 2021; Huang et al.,
-2020; Pinhasi et al., 2021; Taubert et al., 2023; Wenaas, 2022).
+access articles through transformative agreements, while recording less
+than half of total publication volume published between 2018 and 2022.
+This observation aligns with previous research on shifts in the
+publishing market following the introduction of funding opportunities
+for hybrid open access (Butler et al., 2023; Jahn \& Tullney, 2016; Shu
+\& Larivière, 2023). Additionally, the results confirm previously
+observed variations by publisher, with Elsevier exhibiting a different
+development than Springer Nature and Wiley (Butler et al., 2023; Jahn et
+al., 2022). However, it must be noted that the focus of transformative
+agreements on publishers with large hybrid journal portfolios is
+intentional (Campbell et al., 2022). Because of transformative
+agreements with a few large publishers, national consortia were able to
+substantially increase their country's annual open access article volume
+(Bosman, Jonge, et al., 2021; Huang et al., 2020; Pinhasi et al., 2021;
+Taubert et al., 2023; Wenaas, 2022).
 
@@
 States, for example, hybrid open access including transformative
-agreements plays a lesser role compared to some European countries. The
-smallest share of federally funded research was hybrid open access
-between 2017 and 2021 (Schares, 2023). While some university consortia
-like the California Digital Library signed transformative agreements,
-others attempt to depart from big deals and unbundle large journal
-portfolios to address cost increases (Brainard, 2021; Schares, 2022).
-Despite the relatively low penetration of transformative agreements in
-China and the United States compared to Europe, 22\% of open access in
-hybrid journals in 2022 originates from these two countries, indicating
-the availability of funding sources for publication fees.
+agreements plays a lesser role compared to some European countries.
+Between 2017 and 2021, hybrid open access contributed the least to make
+federally funded research articles openly available (Schares, 2023).
+While some university consortia like the California Digital Library have
+signed transformative agreements, others attempt to depart from big
+deals and unbundle large journal portfolios to address cost increases
+(Brainard, 2021; Schares, 2022). Despite the relatively low penetration
+of transformative agreements in China and the United States compared to
+Europe, 22\% of open access in hybrid journals in 2022 originates from
+these two countries, indicating the availability of funding sources for
+publication fees.
 
@@
 author affiliations with publicly available agreement data from
-cOAlition S, and not through invoicing data because these data a usually
-not shared by publishers. Moreover, this study is unable to account for
-the various types of transformative agreements, particularly with
-regards to article types and capping. Similarly, assessing the quality
-of the OpenAlex and Crossref data used, particularly in terms of
-affiliations and article types, in comparison with established
-bibliometric databases was beyond the scope of this analysis. Overall,
+cOAlition S, and not through invoicing data, which is usually not
+shared. Moreover, this study is unable to account for the various types
+of transformative agreements, particularly with regards to article types
+and capping. Furthermore, assessing the quality of the OpenAlex and
+Crossref data used, particularly in terms of affiliations and article
+types, by comparing it with established bibliometric databases such as
+Scopus or Web of Science was beyond the scope of this analysis. Overall,
 the methodology is designed in such a way that the results underestimate
@@
 impacted by the COVID-19 pandemic, which led to an unprecedented number
-of publications as well as restrictions on migration flows, which could
-have influenced the country-specific results (Aviv-Reuven \& Rosenfeld,
-2021; Zhao et al., 2023). Furthermore, the study design did not consider
-emerging publication practices such as preprints (Fraser et al., 2021)
-and questionable special issues (Hanson et al., 2023), which have grown
-particularly rapidly since 2020. Lastly, it should be emphasized that
-the study aims to estimate articles enabled through transformative
-agreements but does not address financial shifts due to a lack of data
-on expenditures.
+of publications as well as an reduction of international collaboration
+(Aviv-Reuven \& Rosenfeld, 2021), which could explain the observed
+contrasting developements in OECD and BRICS countries. Even before the
+pandemic, growth in publications in Europe were only due to
+internationally co-authored journal articles (Kwiek, 2021). Likewise,
+inflows from China to the US and European countries already declined
+before 2020 (Zhao et al., 2023). Furthermore, the study design did not
+consider emerging publication practices such as preprints (Fraser et
+al., 2021) and special issues (Hanson et al., 2023), which have grown
+rapidly since 2020. Lastly, it should be emphasized that the study did
+not address financial shifts between suscriptiuons spending and open
+acces spayemnts while analysing hybrid open access through
+transformative agreements due to a lack of data on expenditures.
 
@@
 
-But the study also has practival implications for research funders and
-libraries. Of primary concern should be observed differences across
+This study also has practival implications for research funders and
+libraries. Of primary concern should be the observed differences across
 countries, in particular the relationship between economic development
 and open access adoption. From a data perspective, the reporting of open
-access funding is not harmonised, but often crowd-sourced from various
-sources. To improve the assessment of transformative agreements,
-libraries and publishers could contribute by providing information about
-respective journal portfolios, participating institutions and open
-access invoicing.
-
-In summary, the aim of this study was to provide empirical insights into
-the developement of hybrid open access following the introduction of
-transformative agreements. The results are important for both
-researchers and stakeholders engaged in negotiating and evaluating these
-agreements. The presented approach relies on open data, which enables
-follow-up studies and open access monitoring acitvties to further
-explore the role of transformative agreements in facilitating the shift
-towards full open access in academic publishing.
-
---\textgreater{}
+access funding including transformative agreements is not harmonised,
+but often crowd-sourced from various sources. To improve the assessment
+of transformative agreements, libraries and publishers could agree on
+standards and services to publicly share information about respective
+journal portfolios, participating institutions and open access
+invoicing.
+
+In summary, this study provides empirical insights into the developement
+of hybrid open access following the introduction of transformative
+agreements. The results are important for both researchers and
+stakeholders engaged in negotiating and evaluating these agreements. The
+presented approach relies on open data, which enables follow-up studies
+and open access monitoring acitvties to further explore the role of
+transformative agreements in transitioning academic publishing to full
+and equitable open access.
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Borrego_2020}{}}%
-Borrego, Á., Anglada, L., \& Abadal, E. (2020). Transformative
+Borrego, Á., Anglada, L., \& Abadal, E. (2021). Transformative
 agreements: Do they pave the way to open access? \emph{Learned
-Publishing}. \url{https://doi.org/10.1002/leap.1347}
+Publishing}, \emph{34}(2), 216--232.
+\url{https://doi.org/10.1002/leap.1347}
 
@@
 Bosman, J., Frantsvåg, J. E., Kramer, B., Langlais, P.-C., \& Proudman,
-V. (2021). \emph{OA diamond journals study. Part 1: findings}. Zenodo.
+V. (2021). \emph{{OA Diamond Journals Study}. Part 1: findings}. Zenodo.
 \url{https://doi.org/10.5281/zenodo.4558704}
@@
 Bosman, J., Jonge, H. de, Kramer, B., \& Sondervan, J. (2021). Advancing
-open access in the netherlands after 2020: From quantity to quality.
+open access in the {Netherlands} after 2020: From quantity to quality.
 \emph{Insights the UKSG Journal}, \emph{34}.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Brainard_2021}{}}%
-Brainard, J. (2021). California universities and elsevier make up, ink
+Brainard, J. (2021). California universities and {Elsevier} make up, ink
 big open-access deal. \emph{Science}.
@@
 Swaisland, M., \& Stelt, W. van der. (2010). \emph{{Open Access
-Publishing - Models and Attributes}} (T. S. consortium, Ed.).
+Publishing - Models and Attributes}}. The SOAP consortium.
+\url{https://hdl.handle.net/11858/00-001M-0000-0013-838A-6}
 
@@
 deal: German researchers' publishing and citing behaviors after big deal
-negotiations with elsevier. \emph{Quantitative Science Studies},
+negotiations with {Elsevier}. \emph{Quantitative Science Studies},
 \emph{4}(2), 325--352. \url{https://doi.org/10.1162/qss_a_00255}
@@
 Haucap, J., Moshgbar, N., \& Schmal, W. B. (2021). The impact of the
-german {``DEAL''} on competition in the academic publishing market.
+{German {``DEAL''}} on competition in the academic publishing market.
 \emph{Managerial and Decision Economics}, \emph{42}(8), 2027--2049.
@@
 hybrid open access through publisher-provided metadata: An article-level
-study of elsevier. \emph{Journal of the Association for Information
+study of {Elsevier}. \emph{Journal of the Association for Information
 Science and Technology}, \emph{73}(1), 104--118.
@@
 Jahn, N., \& Tullney, M. (2016). A study of institutional spending on
-open access publication fees in germany. \emph{{PeerJ}}, \emph{4},
+open access publication fees in {Germany}. \emph{{PeerJ}}, \emph{4},
 e2323. \url{https://doi.org/10.7717/peerj.2323}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Koley_2023}{}}%
-Koley, M., \& Lala, K. (2023). Limitations of the {``indian one nation,
-one subscription''} policy proposal and a way forward. \emph{Journal of
+Koley, M., \& Lala, K. (2023). Limitations of the {{``Indian one nation,
+one subscription''}} policy proposal and a way forward. \emph{Journal of
 Librarianship and Information Science}, 096100062211467.
 \url{https://doi.org/10.1177/09610006221146771}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Kwiek_2020}{}}%
+Kwiek, M. (2021). What large-scale publication and citation data tell us
+about international research collaboration in europe: Changing national
+patterns in global contexts. \emph{Studies in Higher Education},
+\emph{46}(12), 2629--2649.
+\url{https://doi.org/10.1080/03075079.2020.1749254}
 
@@
 Marques, M., \& Stone, G. (2020). Transitioning to open access: An
-evaluation of the {UK} springer compact agreement pilot 2016--2018.
+evaluation of the {UK} {Springer Compact} agreement pilot 2016--2018.
 \emph{College {\&} Research Libraries}, \emph{81}(6), 913--927.
@@
 Marwick, B., Boettiger, C., \& Mullen, L. (2018). Packaging data
-analytical work reproducibly using r (and friends). \emph{The American
+analytical work reproducibly using {R} (and friends). \emph{The American
 Statistician}, \emph{72}(1), 80--88.
@@
 Open-Access-Publikationskosten}. \emph{{O-Bib. Das Offene
-Bibliotheksjournal / Herausgeber VDB}}, \emph{8}.
-\url{https://doi.org/10.5282/O-BIB/5731}
+Bibliotheksjournal}}, \emph{8}. \url{https://doi.org/10.5282/O-BIB/5731}
 
@@
 (2023). Which factors are associated with open access publishing? A
-springer nature case study. \emph{Quantitative Science Studies},
+{Springer Nature} case study. \emph{Quantitative Science Studies},
 \emph{4}(2), 353--371. \url{https://doi.org/10.1162/qss_a_00253}
@@
 Strategies for negotiating and signing transformative agreements in the
-global south: The colombia consortium experience. \emph{Journal of
+{Global South}: The {Colombia Consortium} experience. \emph{Journal of
 Library Administration}, \emph{64}(1), 80--98.
@@
 Taubert, N., Hobert, A., Jahn, N., Bruns, A., \& Iravani, E. (2023).
-Understanding differences of the OA uptake within the german university
-landscape (2010--2020): Part 1---journal-based OA.
+Understanding differences of the OA uptake within the {German}
+university landscape (2010--2020): Part 1---journal-based OA.
 \emph{Scientometrics}, \emph{128}(6), 3601--3625.

### 2024-02-26 analysis/manuscript.Rmd
Birgits suggestions

@@
 In 2009, Springer reported open access to 1% of articles in hybrid journals [@Dallmeier_2010].
-Other studies also recorded a low uptake.
-In 2011, only 1-2% of articles were open access [@Bj_rk_2012], growing to around 4% between 2011 and 2013 [@Laakso_2016].
-
-With the introduction of central funding mechanisms for publication fees in some European countries since 2012, an increase in hybrid open access could be observed [@Bj_rk_2017; @Piwowar_2018; @Jubb_2017; @Huang_2020].
+Other studies have also recorded low uptake.
+In 2011, only 1-2% of articles were open access [@Bj_rk_2012], increasing to around 4% between 2011 and 2013 [@Laakso_2016].
+
+With the introduction of central funding mechanisms for publication fees in some European countries since 2012, an substantial increase in hybrid open access has been observed [@Bj_rk_2017; @Piwowar_2018; @Jubb_2017; @Huang_2020].
 For example, studying university output, @Robinson_Garcia_2020 estimated a median uptake of 7.1% in the period 2014-2017. 
-In particular, British (17%), Austrian (15%) and Dutch (13%) universities stand out. 
+In particular, British (17%), Austrian (15%) and Dutch (13%) universities stood out. 
 However, this shift in funding policy towards hybrid open access also added to the overall cost of publishing, which includes subscription spending and the administrative efforts required to handle payments [@Pinfield_2016].
-Moreover, established large commercial publishers, which already dominated the publishing market [@Larivi_re_2015], disproportionately benefited from hybrid open access funding in comparison to full open access publishers [@Jahn_2016; @Butler_2023; @Shu_2023].
+Moreover, large commercial publishers, which already dominated the publishing market [@Larivi_re_2015], disproportionately benefited from hybrid open access funding in comparison to full open access publishers [@Jahn_2016; @Butler_2023; @Shu_2023].
 
 As a consequence, libraries and their consortia began to develop licensing strategies aimed at avoiding such ‘double dipping’ scenarios, in which well-established commercial publishers gain twice from reading and open access publishing [@Mittermaier_2015], as well as to increase publisher-provided immediate open access [@BJ_RK_2014; @Schimmer_2015].
-These considerations resulted into transformative agreements[^agreements], which cover a broad range of contracts between library consortia and publishers from mid-2010s onwards where institutional spending for subscriptions and open access publishing are considered together [@Borrego_2020; @Hinchliffe_2019].
+These considerations resulted in transformative agreements[^agreements], which cover a broad range of contracts between library consortia and publishers from the mid-2010s onwards, where institutional spending for subscriptions and open access publishing are considered together [@Borrego_2020; @Hinchliffe_2019].
 Transformative agreements seek to control costs while allowing a transitional phase for publishing more open access articles. 
@@
 
-The introduction of transformative agreements aligns with funding policy changes, such as the decision made by the cOAlition S, a consortium of national funders including the European Commission, to no longer provide financial support for individual publication fees when publishing in hybrid journals.
-According ot its Plan S launched in 2018, the cOAlition S funders only accept hybrid open access through transformative agreements "during a transition period that should be as short as possible" [@Schiltz_2018].
+The introduction of transformative agreements aligns with funding policy changes, such as the decision made by cOAlition S, a consortium of funders and research organisations including the European Commission, to no longer provide financial support for individual publication fees when publishing in hybrid journals.
+According to its Plan S launched in 2018, cOAlition S members only accept hybrid open access through transformative agreements "during a transition period that should be as short as possible" [@Schiltz_2018].
 Specifically, they agreed to support hybrid open access only through transformative agreements from 2021, until the end of 2024.
-Notably, the German Research Foundation (DFG), despite not being part of cOAlition S, has also extended its financial support for hybrid open access through transformative agreements [@Mittermaier_2021].
-Previously, the DFG only provided funding for full open access journals [@Jahn_2016].
-
-By the end of 2023, many transformative agreements were implemented, but interim outcomes are mixed.
-The ESAC Transformative Agreement Registry[^esacregistry], the largest source of disclosure, recorded more than 800 transformative agreements, resulting in up to 900.000 open access articles published in both full open access and hybrid journals according to the accompanying ESAC Market Watch[^esacmarketwatch].
-Library consortia reported increased open access volume, streamlined payment and monitoring procedures, as well as extensive utilization of open access options by the researchers they serve [@Marques_2020; @Pinhasi_2020; @Parmhed_2023].
-The ongoing standardisation of transformative agreements contributed to improved transparency in terms of contracts and publisher-provided article metadata [@Marques_2019; @Pinhasi_2021].
-However, with the growing trend toward transformative agreements, continued reliance on big deals is perceived as problematic, because it perpetuate market concentration [@Butler_2023; @Shu_2023]. 
+Notably, despite not being part of cOAlition S, the German Research Foundation (DFG), has also extended its financial support for hybrid open access through transformative agreements [@Mittermaier_2021].
+Previously, the DFG only provided funding for fully open access journals [@Jahn_2016].
+
+By the end of 2023, many transformative agreements had been implemented, but the interim outcomes were mixed.
+The ESAC Transformative Agreement Registry[^esacregistry], the largest source of disclosure, recorded more than 800 transformative agreements, resulting in up to 900.000 open access articles published in both fully open access and hybrid journals according to the accompanying ESAC Market Watch[^esacmarketwatch].
+Library consortia reported increased open access volume, streamlined payment and monitoring procedures, as well as extensive utilisation of open access options by the researchers they serve [@Marques_2020; @Pinhasi_2020; @Parmhed_2023].
+The ongoing standardisation of transformative agreements has contributed to improved transparency in terms of contracts and publisher-provided article metadata [@Marques_2019; @Pinhasi_2021].
+However, with the growing trend toward transformative agreements, continued reliance on big deals is perceived as problematic, because it perpetuates market concentration [@Butler_2023; @Shu_2023]. 
 Whether transformative agreements lead to reduced pricing remains uncertain [@Borrego_2023] and
 a substantial transition of hybrid journals towards full open access could not be observed [@Momeni_2021; @Matthias_2019].
-The focus on large commercial publishers might also increase inequality [@Klebel_2023], because transformative agreements focus on pay to publish open access mainly targets institutions from high-income countries, furthering a questionable journal prestige culture [@budapest].
-Besides, an editorial-board resignation raised concerns that publishers' desire to maximize journal publication volume "without regard to quality" is a consequence of transformative agreements [@Goodin_2023].
+The focus on large commercial publishers might also increase inequality [@Klebel_2023], because transformative agreements' focus on pay to publish mainly targets institutions from high-income countries, furthering a questionable journal prestige culture [@budapest].
+Besides, an editorial-board resignation raised concerns that transformative agreements might encourage publishers to maximize journal publication volume "without regard to quality" [@Goodin_2023].
 
@@
 For instance, the Association of Swedish Higher Education Institutions (Sveriges universitets- och
-högskoleförbund, SUHF) recommended to only support agreements for publishing in full open access journals [^sweden].
+högskoleförbund, SUHF) recommended only supporting agreements for publishing in fully open access journals [^sweden].
 Likewise, most cOAlition S funders will discontinue financial support for transformative agreements by the end of 2024 [@Liverpool_2023].
-The consortium also removed most hybrid journals from its Transformative Journal program in 2023 due to publishers' failure to meet self-defined open access growth targets [@Brainard_2023]. 
-In contrast, the German DEAL consortium announced a five-year transformative agreement with Elsevier starting in 2024, while also renewing its contracts with Springer Nature and Wiley until the end of 2028.
-At the same time, the Colombia Consortium signed the first transformative agreements in Latin America [@Mu_oz_V_lez_2024].
-
-Despite these controversies around transformative agreements as a means of transitioning journal publishing to full open access, there is limited evidence available on the adoption of open access in hybrid journals, and the extent to which this uptake can be attributed to transformative agreements.
+The consortium also removed the majority of hybrid journals from its Transformative Journal program in 2023 due to publishers' failure to meet self-defined open access growth targets [@Brainard_2023].
+By contrast, the German DEAL consortium announced a five-year transformative agreement with Elsevier starting in 2024 and also renewed its contracts with Springer Nature and Wiley until the end of 2028.
+Similarly, the Colombia Consortium signed the first transformative agreements in Latin America [@Mu_oz_V_lez_2024].
+
+Despite these controversies around transformative agreements as a means of transitioning journal publishing to full open access, there is limited evidence available on the adoption of open access in hybrid journals, and the extent to which this can be attributed to transformative agreements.
 Previous studies have focused on specific countries [@P_l_nen_2020; @Huang_2020; @Haucap_2021; @Wenaas_2022; @Taubert_2023] or publisher portfolios [@Pieper_2018; @Bakker_2024; @Fraser_2023; @Jahn_2021; @Momeni_2023], while large-scale studies relied on self-reported agreement data [@Moskovkin_2022], or used APC pricing lists [@Shu_2023].
-Particularly, data availability is a limiting factor when studying the impact of transformative agreements [@Bakker_2024], because bibliometric databases, even though many allow the retrieval of open access articles in hybrid journals, do not directly attribute them to specific transformative agreements.
-Likewise, article-level open access invoicing data is only partly available [@Jahn_2021].
-
-The present study aims to address these limitations by combining multiple openly available data sources to determine the open access uptake in hybrid journals, while distinguishing between open access through transformative agreements and other means.
-With this novel and open approach, this first large-scale analysis aims to answer the following questions:
+In particular, data availability is a limiting factor when studying the impact of transformative agreements [@Bakker_2024], because bibliometric databases, even though many allow the retrieval of open access articles in hybrid journals, do not directly attribute them to specific transformative agreements.
+Likewise, article-level open access invoicing data are only partly available [@Jahn_2021].
+
+The present study aims to address these limitations by combining multiple openly available data sources to determine open access uptake in hybrid journals, while distinguishing between open access through transformative agreements and other means.
+With this novel and open approach, this first large-scale analysis will answer the following questions:
 
@@
 
-For both of these research questions, this study will analyse the variability by publisher, journal subject, and country.
+For both research questions, this study will analyse the variability by publisher, journal subject, and country.
 
@@
 
-This study combines data from multiple publicly available data sources as shown in Figure \ref{fig:data_workflow}. 
-Initially, transformative agreement data retrieved from the cOAlition S Journal Checker Tool[^checkertool]  provided information about journal portfolios and participating institutions. 
-After identification of hybrid journals by excluding full open access journals, Crossref served as the primary data source for article-level metadata including Creative Commons (CC) license information to indicate open access availability on publisher websites.
+This study combines data from multiple publicly available sources, as shown in Figure \ref{fig:data_workflow}. 
+Initially, transformative agreement data retrieved from cOAlition S Journal Checker Tool[^checkertool]  provided information about journal portfolios and participating institutions. 
+After identifying hybrid journals by excluding fully open access journals, Crossref served as the primary data source for article-level metadata including Creative Commons (CC) license information to indicate open access availability on publisher websites.
 To determine open access articles published through transformative agreements, first author affiliations from OpenAlex [@priem2022openalex] were subsequently linked to eligible institutions according to the transformative agreement data.
-In the following, these steps are described in more detail.
+In the following, the steps are described in detail.
 
@@
 <!-- Agreements ending before 2021 could not be obtained.-->
-Overall, the four combined Transformative Agreement Data dumps used in this study contained `r  length(unique(hoaddata::jct_hybrid_jns$esac_id))` out of `r length(unique(esac_df$ID))` agreements listed in the ESAC registry by December 2023.
-
-The Transformative Agreement Data dumps link agreements to journals represented by journal names and ISSN. 
-After mapping ISSN variants to the corresponding linking ISSN (ISSN-L) as provided by the ISSN International Centre, journals were associated to publishers using the ESAC ID, a unique identifier for transformative agreements in the ESAC Transformative Agreement Registry.
+Overall, the four combined Transformative Agreement Data dumps used in this study contained `r  length(unique(hoaddata::jct_hybrid_jns$esac_id))` of `r length(unique(esac_df$ID))` agreements listed in the ESAC registry by December 2023.
+
+The Transformative Agreement Data dumps link agreements to journals represented by journal names and the ISSN. 
+After mapping ISSN variants to the corresponding linking ISSN (ISSN-L), as provided by the ISSN International Centre, journals were associated with publishers using the ESAC ID, a unique identifier for transformative agreements in the ESAC Transformative Agreement Registry.
 Furthermore, journal subjects according to the All Science Journal Classification code (ASJC) were added from the Scopus journal source list as of August 2023.
 
-Because transformative agreements can include both full open access and hybrid journals, the data were complemented with information about a journal's open access status using multiple sources: the Directory of Open Access Journals (DOAJ) downloaded on 12 December 2023[^doaj], OpenAlex (November 2023) and the the Bielefeld list of GOLD OA journals [@goldoa].
-As shown in Figure \ref{fig:method_fig}A, combining different data sources considerably extended the journal matching<!-- to exclude full open access journals from transformative agreements-->.
+Because transformative agreements can include both fully open access and hybrid journals, the data were complemented with information about a journal's open access status using multiple sources: the Directory of Open Access Journals (DOAJ) downloaded on 12 December 2023[^doaj], OpenAlex (November 2023) and the Bielefeld list of GOLD OA journals [@goldoa].
+As shown in Figure \ref{fig:method_fig}A, combining different data sources considerably extended journal matching<!-- to exclude full open access journals from transformative agreements-->.
 In total, `r n_oa_journals` full open access journals were excluded based on ISSN matching.
 The overlap between the three data sources was `r round(length(unique(Reduce(intersect, my_venn))) / length(unique(Reduce(union, my_venn)))* 100,0)`%.
-The Gold OA journals dataset alone added `r length(unique(Reduce(setdiff, my_venn)))` journals, while the DOAJ comprised `r length(setdiff(my_venn$DOAJ, union(my_venn[["GOLD OA"]], my_venn$OpenAlex)))` full open access journals not listed in either of the other two sources.
-These full open access journals were mostly launched in 2022.
+The Gold OA journals dataset alone added `r length(unique(Reduce(setdiff, my_venn)))` journals, while the DOAJ comprised `r length(setdiff(my_venn$DOAJ, union(my_venn[["GOLD OA"]], my_venn$OpenAlex)))` fully open access journals not listed in either of the other two sources.
+These fully open access journals were launched in 2022.
 
@@
 
-After identifying hybrid journals included in transformative agreements, article metadata was retrieved from the Crossref <!-- and mapped to OpenAlex for obtaining author metadata including affiliations.
---> November 2023 database snapshot for the five-year period 2018 to 2022 according to the issued date, representing the earliest known publication date.
-Because Crossref metadata lacked information to distinguish between original research articles including review and other types of journal content, which are often not covered by transformative agreements [@Borrego_2020], only articles published in regular issues indicated by non-numeric pagination were included.
+After identifying hybrid journals included in transformative agreements, article metadata were retrieved from the Crossref <!-- and mapped to OpenAlex for obtaining author metadata including affiliations.
+--> November 2023 database snapshot for the five-year period from 2018 to 2022, according to the issued date, representing the earliest known publication date.
+Because Crossref metadata lacked information to distinguish between original research articles, including reviews, and other types of journal content, which are often not covered by transformative agreements [@Borrego_2020], only articles published in regular issues indicated by non-numeric pagination were included.
 Furthermore, an expanded version of Unpaywall’s paratext recognition approach was applied to exclude non-scholarly journal content such as table of contents.
 
-Open access articles in hybrid journals were identified through Creative Commons (CC) license information in Crossref metadata. 
-License information relative to the "accepted manuscript (AM)" version were not considered. 
-Crossref was used for open access identification because transformative agreements workflows generally require publishers to deliver CC license information to this DOI registration agency [@Geschuhn_2017].
-
-Comparing Crossref license coverage with OpenAlex, which re-uses open access evidence from Unpaywall, a widely used open access discovery service that also parses journal websites for open content licenses [@Piwowar_2018], highlighted ongoing challenges to identify hybrid open access [@Mart_n_Mart_n_2018; @Jahn_2021; @Zhang_2022; @Butler_2023].
-Here, `r hoaddata::cr_upw |> filter(cr_year %in% 2018:2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, cr_hybrid_total) |> pull(cr_hybrid_total) |> sum()` articles with CC license were retrieved using Crossref, while `r hoaddata::cr_upw |> filter(cr_year %in% 2018:2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, upw_hybrid_total) |> pull(upw_hybrid_total) |> sum()` articles were tagged as "hybrid" according to the OpenAlex November 2023 release, which was used throughout this study.
-The biggest differences concerned articles published between 2018 and 2020.
-In 2022, however, Crossref and OpenAlex open access numbers only differ slightly (`r hoaddata::cr_upw |> filter(cr_year %in% 2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, cr_hybrid_total) |> pull(cr_hybrid_total) |> sum()` records using Crossref vs.
+Open access articles in hybrid journals were identified using the Creative Commons (CC) license information in Crossref metadata. 
+License information relative to the "accepted manuscript (AM)" version was not considered. 
+Crossref was used for open access identification because transformative agreement workflows generally require publishers to deliver CC license information to this DOI registration agency [@Geschuhn_2017].
+
+Comparing Crossref license coverage with OpenAlex, which re-uses open access evidence from Unpaywall, a widely used open access discovery service that also parses journal websites for open content licenses [@Piwowar_2018], highlighted ongoing challenges in identifying hybrid open access [@Mart_n_Mart_n_2018; @Jahn_2021; @Zhang_2022; @Butler_2023].
+Here, `r hoaddata::cr_upw |> filter(cr_year %in% 2018:2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, cr_hybrid_total) |> pull(cr_hybrid_total) |> sum()` articles under CC license were retrieved using Crossref, while `r hoaddata::cr_upw |> filter(cr_year %in% 2018:2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, upw_hybrid_total) |> pull(upw_hybrid_total) |> sum()` articles were tagged as "hybrid" according to the OpenAlex November 2023 release, which was used throughout this study.
+The largest differences concerned articles published between 2018 and 2020.
+In 2022, however, Crossref and OpenAlex open access numbers differ only slightly (`r hoaddata::cr_upw |> filter(cr_year %in% 2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, cr_hybrid_total) |> pull(cr_hybrid_total) |> sum()` records using Crossref vs.
 `r hoaddata::cr_upw |> filter(cr_year %in% 2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, upw_hybrid_total) |> pull(upw_hybrid_total) |> sum()` in OpenAlex).
-Notable difference could be furthermore observed among some publishers that presumably did not provide CC license information to Crossref including AIP Publishing, American Physiological Society, Emerald and the Royal Society.
-Crossref license metadata was more complete with regard to articles from the publisher Wiley and American Chemical Society.
+Notable differences could be observed among some publishers that presumably did not provide CC license information to Crossref, including AIP Publishing, the American Physiological Society, Emerald, and the Royal Society.
+Crossref license metadata was more complete with regard to the articles published by Wiley and the American Chemical Society.
 Finally, inconsistent open access status information in previous OpenAlex versions was observed [@jahn2023].
-After reporting it to OpenAlex, fixing this issue was still ongoing according to the release notes, which might also explain part of the discrepancy.
-
-After retrieving article metadata, the publication volume including open access was calculated per journal. 
+After reporting this to OpenAlex, fixing this issue was still ongoing according to the release notes, which might also explain part of the discrepancy.
+
+After retrieving the article metadata, the publication volume, including open access, was calculated for each journal. 
 To improve the identification of hybrid journals, journals with an open access proportion above 95% were excluded.
-This further step allowed to remove additional `r length(unique(oa_prop$issn_l))` full open access journals.
-
-Affiliation metadata about corresponding authors are crucial for the planning and evaluation of transformative agreements, because they are considered to be responsible to arrange open access publication [@Schimmer_2015; @Geschuhn_2017; @Borrego_2020].
-Here, country and institutional affiliations were retrieved from OpenAlex.
-However, because of low coverage in OpenAlex, this study focused on first authors and their affiliations instead.
-First authors typically contribute most to a paper and are often considered lead author research papers [@Larivi_re_2016]. 
-Related studies therefore assumed first authors as a proxy to measure to open access payments and the impact of transformative agreement [@Zhang_2022; @Shu_2023; @Haucap_2021].
-Overall, around 90% of studied articles had first author affiliation metadata in OpenAlex, whereas the coverage of articles with corresponding author information was around 54%.
+This step allowed removing additional `r length(unique(oa_prop$issn_l))` fully open access journals.
+
+Affiliation metadata about corresponding authors are crucial for the planning and evaluation of transformative agreements because they are considered responsible for arranging open access publication [@Schimmer_2015; @Geschuhn_2017; @Borrego_2020].
+For this study, country and institutional affiliations were retrieved from OpenAlex.
+However, because corresponding author affiliation was only available for 54% of the investigated articles in OpenAlex, this study focused on first authors and their affiliations instead; approximately 90% of the investigated articles had first author affiliation metadata in OpenAlex.
+First authors typically contribute the most to a paper and are often considered being lead authors [@Larivi_re_2016]. 
+Related studies also assumed first authors as a proxy to measure open access payments and the impact of transformative agreements [@Zhang_2022; @Shu_2023; @Haucap_2021].
 
@@
 
-To assess the impact of transformative agreements to hybrid open access,
-participating institutions from the Transformative Agreement Data dump,
-which were crowd-sourced from the agreements and consortia that
-successfully negotiated an agreement, were matched with first author
-affiliations recorded by OpenAlex using the ROR-ID.
-The matching also took into account the duration of an agreement according to the ESAC registry.
-Upon inspection, Transformative Agreement Data did not cover associated institutions comprehensively like university hospitals or institutes of large research organisations like the Max Planck Society.
-To improve the matching, Transformative Agreement Data was automatically enriched with associated organisations using OpenAlex's institution entity data.
-
-In total, the compiled data set consists of `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l, cr_year, jn_all) |> pull(jn_all) |> sum()` articles published in `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l) |> nrow()` hybrid journals between 2018 and 2022 (see Figure \ref{fig:method_fig}B).
-Hybrid journals in transformative agreements represented `r cr_stats |> filter(src == "hoad") |> pull(n_articles) |> sum() / cr_stats |> pull(n_articles) |> sum() * 100`%  of total global output over the same time period according to Crossref, while full open access journals recorded
+To assess the impact of transformative agreements on hybrid open access, participating institutions from the Transformative Agreement Data dump, which were crowd-sourced from the agreements and consortia that successfully negotiated an agreement, were matched with the first author affiliations recorded by OpenAlex using the ROR-ID.
+Matching also considered the duration of agreements according to the ESAC registry.
+Upon inspection, Transformative Agreement Data did not comprehensively cover associated institutions, such as university hospitals or institutes of large research organisations such as the Max Planck Society.
+To improve the matching, Transformative Agreement Data were automatically enriched with associated organisations using OpenAlex's institution entity data.
+
+The so compiled data set consists of `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l, cr_year, jn_all) |> pull(jn_all) |> sum()` articles published in `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l) |> nrow()` hybrid journals between 2018 and 2022 (see Figure \ref{fig:method_fig}B).
+Hybrid journals in transformative agreements represented `r cr_stats |> filter(src == "hoad") |> pull(n_articles) |> sum() / cr_stats |> pull(n_articles) |> sum() * 100`%  of the total global output over the same time period, according to Crossref, while fully open access journals recorded
 `r cr_stats |> filter(is.na(src) & oa_status == "gold") |> pull(n_articles) |> sum() / cr_stats |> pull(n_articles) |> sum() * 100`%.
@@
 (B) Number of articles in Crossref by journal types.
-The blue bars show the overall article volume of hybrid journals in transformative agreements, which were initially included in the study, in comaprision with full open access journals according to OpenAlex.
-The remainder represents closed access journals not covered by transformative agreements."}
+The blue bars show the overall article volume of hybrid journals in transformative agreements, which were initially included in the study, in comparision with full open access journals according to OpenAlex.
+The remainder represents subscription-based journals not covered by transformative agreements."}
 (p | cr_stats_plot) + plot_annotation(tag_levels = 'A') +  plot_layout(widths = c(0.6, 0.4))
@@
 
-Throughout this mostly automated data gathering and analysis process, tools from the Tidyverse (Wickham et al., 2019) for the R programming language (R Core Team, 2020) were used. 
-The resulting data is openly available through an R data package, hoaddata, version 0.2.91 
+Throughout this mostly automated data gathering and analysis process, Tidyverse tools (Wickham et al., 2019) for the R programming language (R Core Team, 2020) were used. 
+The resulting data are openly available through an R data package, hoaddata, version 0.2.91 
 Following @Marwick_2018, hoaddata contains not only the datasets used in the data analysis. 
-It also includes code used to compile the data by connecting it the cloud-based Google Big Query data warehouse, where the big scholarly data from Crossref, OpenAlex and Unpaywall were imported.
-To increase the computational reproducibility, the data aggregation through hoaddata was automatically carried out using GitHub Actions, a continuous integration service.
+It also includes code used to compile the data by connecting it to the cloud-based Google Big Query data warehouse, where scholarly big data from Crossref, OpenAlex and Unpaywall were imported.
+To increase computational reproducibility, data aggregation through hoaddata was automatically carried out using GitHub Actions, a continuous integration service.
 
@@
 
-Between 2018 and 2022, a total of `r n_hybrid_jns` out of `r n_hybrid_jns_all` hybrid journals in transformative agreements published at least one open access article under a Creative Commons license.
-During this period, these hybrid journals provided open access to `r n_oa` out of `r n_total` articles, representing a five-year open access proportion of `r n_oa / n_total * 100`%.
-Authors who could make use of transformative agreements at the time of publication contributed `r n_ta_oa` open access articles to the total.
+Between 2018 and 2022, `r n_hybrid_jns` out of `r n_hybrid_jns_all` hybrid journals in transformative agreements published at least one open access article under a Creative Commons license.
+During this period, these hybrid journals provided open access to `r n_oa` of the `r n_total` articles, representing a five-year open access proportion of `r n_oa / n_total * 100`%.
+Authors who could make use of transformative agreements at the time of publication contributed `r n_ta_oa` open access articles.
 
@@
 Over the five-years period from 2018 to 2022, open access increased from `r oa_by_year |> filter(cr_year == "2018") |> pull(prop)`% (n = `r oa_by_year |> filter(cr_year == "2018") |> pull(oa)`) to `r oa_by_year |> filter(cr_year == "2022") |> pull(prop)`% (n = `r oa_by_year |> filter(cr_year == "2022") |> pull(oa)`).
-At the same time, the total article volume of the investigated journals grew from `r oa_by_year |> filter(cr_year == "2018") |> pull(articles)` in 2018 to `r oa_by_year |> filter(cr_year == "2022") |> pull(articles)` in 2022.
+Simultaneously, the total article volume of the investigated journals increased from `r oa_by_year |> filter(cr_year == "2018") |> pull(articles)` in 2018 to `r oa_by_year |> filter(cr_year == "2022") |> pull(articles)` in 2022.
 
 Figure \ref{fig:results_overview}B highlights that the majority of hybrid open access was made available through transformative agreements in 2021 and 2022, contributing `r oa_ta_plot_df |> filter(cr_year == "2022", type == "ta_oa") |> pull(value) / oa_by_year |> filter(cr_year == "2022") |> pull(oa) * 100`% of the total open access article volume in 2022.
-However, there was also a notable growth in open access provision through other means, presumably publication fees being not invoiced through transformative agreements, which increased from `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> pull(value)`) in 2018 to  `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> pull(value)`).
+However, there was also a notable increase in open access provision through other means, presumably publication fees being not invoiced through transformative agreements, which increased from `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> pull(value)`) in 2018 to  `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> pull(value)`).
 <!-- This suggests that publishers were able to gain equally from individual and institutional open access publishing options.-->
@@
 In 2022, 25% of hybrid journals (n = `r oa_jn_box_df |> filter(cr_year == "2022") |> mutate(top25 = quantile(prop, .75)) |> filter(prop >= top25) |> nrow()`) had an open access uptake of `r oa_jn_box_df |> filter(cr_year == "2022") |> pull(prop) |> quantile(0.75)*100`%, and `r (oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> nrow()) / length(unique(oa_jn_box_df$issn_l)) *100`% of journals (n = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> nrow()`) provided the majority of their articles under a Creative Commons license in the same year.
-These journals were, on average, smaller (M = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> pull(n) |> mean()`, SD = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> pull(n) |> sd()`) than those with an open access share below 50% (M = `r oa_jn_box_df |> filter(cr_year == "2022", prop <= 0.5) |> pull(n) |> mean()`, SD = `r oa_jn_box_df |> filter(cr_year == "2022", prop <= 0.5) |> pull(n) |> sd()`).
+On average, these journals were smaller (M = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> pull(n) |> mean()`, SD = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> pull(n) |> sd()`) than those with an open access share below 50% (M = `r oa_jn_box_df |> filter(cr_year == "2022", prop <= 0.5) |> pull(n) |> mean()`, SD = `r oa_jn_box_df |> filter(cr_year == "2022", prop <= 0.5) |> pull(n) |> sd()`).
 <!-- Notable exception of large journals with an above-average open access proportion were Physical Review D, a high-energy physics journal covered by the SCOAP3 consortium that provided open access to `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "2470-0010") |> pull(oa_all)` out of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "2470-0010") |> pull(n)` articles in 2022, Astronomy and Physics (`r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0004-6361") |> pull(oa_all)` out of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0004-6361") |> pull(n)` articles in 2022 were open access), which shifted to a subscribe to open business model for all accepted articles as of April 2022, and the Journal of Fluid Mechanics (`r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0022-1120") |> pull(oa_all)` of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0022-1120") |> pull(n)` articles in 2022 were open access).
@@
 
-When comparing the impact of open access trough transformative agreements across journals, it shows that for many journals these agreements substantially contributed to the growth of open access over the years (Figure \ref{fig:results_overview}D).
+When comparing the impact of open access through transformative agreements across journals, it shows that for many journals, these agreements substantially contributed to the growth of open access over the years (Figure \ref{fig:results_overview}D).
 <!-- Examples of such journals include those with a scope on specific countries or regions, where also transformative agreements were implemented. 
@@
 -->
-Despite the rise in transformative agreements, it is worth noting that other means of publishing open access remained common across the investiagted hybrid journals. 
+Despite the rise in transformative agreements, it is worth noting that other means of publishing open access remained common across the investigated hybrid journals. 
 In total, `r summary_all |> filter(cr_year == "2022", value > 0, name == "prop_no_ta") |> nrow()` journals published open access articles from authors affiliated with institutions without transformative agreements in place, while `r summary_all |> filter(cr_year == "2022", value > 0, name == "prop_ta_oa") |> nrow()` journals published at least one open access article through a transformative agreement in the same year.
@@
 Analysing hybrid open access across publishers between 2018 and 2022 reveals a large market concentration. 
-Although `r length(unique(jn_ind_jct_all$esac_publisher))` publishers offered transformative agreements, the big three commercial publishers Elsevier, Springer Nature, and Wiley accounted for `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(jn_cum) * 100`% of total article volume published (see Table \ref{tab:publisher_league_table}).
-Together, they published `r publisher_league_tbl  |> filter(publisher != "Other") |> pull(oa_total) |> sum()` or `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(oa_cum) * 100`% of open access articles in hybrid journals.
+Although `r length(unique(jn_ind_jct_all$esac_publisher))` publishers offered transformative agreements, the three commercial publishers Elsevier, Springer Nature, and Wiley, accounted for `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(jn_cum) * 100`% of the total article volume published (see Table \ref{tab:publisher_league_table}).
+Together, they published `r publisher_league_tbl  |> filter(publisher != "Other") |> pull(oa_total) |> sum()` or `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(oa_cum) * 100`% of the open access articles in hybrid journals.
 Elsevier, Springer Nature, and Wiley made `r publisher_league_tbl  |> filter(publisher != "Other") |> pull(ta_oa_total) |> sum()` articles open access in hybrid journals through transformative agreements, resulting in an even larger market share of `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(ta_oa_cum)*100`%.
@@
 
-However, there are notable differences among the three big publishers.
-Although Elsevier published the largest volume of articles (n = `r publisher_league_tbl  |> filter(publisher == "Elsevier") |> pull(publisher_total)`, `r publisher_league_tbl  |> filter(publisher == "Elsevier") |> pull(cum)*100`%), it recorded a comparable low number of open access articles, including those that can be associated with transformative agreements.
-In contrast, Springer Nature and Wiley provided open access to a larger proportion of their articles (`r (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_total)) / (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(publisher_total)) * 100`% of Springer Nature articles and `r (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(oa_total)) / (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(publisher_total)) * 100`% of Wiley articles were open acccess), leading to higher open access market shares (`r publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_perc)*100`% Springer Nature resp.
+However, there are notable differences between the three large publishers.
+Although Elsevier published the largest volume of articles (n = `r publisher_league_tbl  |> filter(publisher == "Elsevier") |> pull(publisher_total)`, `r publisher_league_tbl  |> filter(publisher == "Elsevier") |> pull(cum)*100`%), it recorded a comparably low number of open access articles, including those that are associated with transformative agreements.
+In contrast, Springer Nature and Wiley provided open access to a larger proportion of their articles (`r (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_total)) / (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(publisher_total)) * 100`% of Springer Nature articles and `r (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(oa_total)) / (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(publisher_total)) * 100`% of Wiley articles were open access), leading to higher open access market share (`r publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_perc)*100`% Springer Nature resp.
 `r publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_perc)*100`% Wiley).
@@
 
-Figure \ref{fig:publisher_figure} takes a closer look into the growth of hybrid open access across publishers by year with a focus on open access enabled by transformative agreements. 
-Although all publishers show a general long-term trend towards transformative agreements, Figure \ref{fig:publisher_figure}A and B indicate that, in particular, Wiley experienced a substantial increase in its open access share from `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(value) |> sum()`) in 2018 to `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(prop) |> sum()*100`%  (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()`) in 2022, representing an 4.5-fold increase.
+Figure \ref{fig:publisher_figure} takes a closer look into the growth of hybrid open access across publishers by year, with a focus on open access enabled by transformative agreements. 
+Although all publishers show a general long-term trend towards transformative agreements, Figure \ref{fig:publisher_figure}A and B indicate that Wiley experienced a substantial increase in its open access share from `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(value) |> sum()`) in 2018 to `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(prop) |> sum()*100`%  (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()`) in 2022, representing a 4.5-fold increase.
 In contrast, Elsevier's hybrid journals demonstrated a more modest increase, from `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Elsevier") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Elsevier") |> pull(value) |> sum()`) in 2018 to `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(value) |> sum()`) in 2022, which is a relatively low open access share compared to the general trend.
@@
 The varying degrees of adoption of open access across the three major publishers can be attributed to distinct approaches to transformative agreements. 
-Springer Nature, for example, began in 2015 offering selected consortia, such as the Max Planck Society, the Swedish Bibsam consortium, and the Finnish FinELib consortium, open access agreements for its hybrid journal portfolio under the name Springer Compact[^springercompact].
-However, these agreements were not included in the data as they concluded prior to the start of the transformative agreement data collection in June 2021. 
+Springer Nature, for example, started offering open access agreements for its hybrid journal portfolio to selected consortia such as the Max Planck Society, the Swedish Bibsam Consortium and the Finnish FinELib consortium in 2015 under the name Springer Compact[^springercompact].
+Prior to this, Springer discontinued its hybrid OA pilot agreements with a small number of institutions, including the University of Göttingen [@Schmidt_2012].
+However, the Springer Compact agreements were not included in the data as they concluded prior to the start of the transformative agreement data collection in June 2021. 
 Nonetheless, the results demonstrate the importance of agreements for Springer Nature's hybrid open access business over the past five years (Figure 2B).
-In 2022, `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature") |> pull(value) |> sum()*100`% (n = `r  publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature", type == "ta_oa") |> pull(value)`) of open access in Springer Nature hybrid journals were enabled through transformative agreements.
+In 2022, `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature") |> pull(value) |> sum()*100`% (n = `r  publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature", type == "ta_oa") |> pull(value)`) of open access in Springer Nature hybrid journals was enabled through transformative agreements.
 In the same year, `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley", type == "ta_oa") |> pull(value)`) of Wiley's open access articles could be linked to transformative agreements in 2022.
-In contrast, Elsevier published fewer than half of its open access articles through transformative agreements (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier", type == "ta_oa") |> pull(value) `; `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(value) |> sum()*100`%).
+By contrast, Elsevier published fewer than half of its open access articles through transformative agreements (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier", type == "ta_oa") |> pull(value) `; `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(value) |> sum()*100`%).
 
@@
 
-The increasing trend towards transformative agreements can be also observed at the journal-level (Figure \ref{fig:publisher_figure}C).
-While no substantial differences between open access enabled through transformative agreements and other revenue sources could observed across Elsevier's portfolio, the distribution of open access across Springer Nature and Wiley hybrid journals indicates that the growth is not limited to a few journals, but extends across the portfolio.
+The increasing trend towards transformative agreements can also be observed at the journal-level (Figure \ref{fig:publisher_figure}C).
+While no substantial differences between open access enabled through transformative agreements and other revenue sources could be seen across Elsevier's portfolio, the distribution of open access across Springer Nature and Wiley hybrid journals indicates that the growth is not limited to a few journals but extends across the portfolio.
 In particular, Wiley's upper quantile, which represents the top 25% of journals in terms of the proportion of open access articles from transformative agreements, increased markedly from `r pub_75 |> filter(publisher == "Wiley", type == "ta_oa", cr_year == "2020") |> pull(top25) * 100`% in 2020 to `r pub_75 |> filter(publisher == "Wiley", type == "ta_oa", cr_year == "2022") |> pull(top25) * 100`% in 2022.
-At the same time, the median proportion grew from `r pub_75 |> filter(publisher == "Wiley", type == "ta_oa", cr_year == "2020") |> pull(median) * 100`% to  `r pub_75 |> filter(publisher == "Wiley", type == "ta_oa", cr_year == "2022") |> pull(median) * 100`%.
-It is interesting to note that a small but increasing number of journals from these two publishers are providing open access to the majority of articles through transformative agreements.
-Wiley recorded `r publisher_plot_df |>  mutate(prop = value / jn_all)  |> filter(cr_year == "2022", prop > 0.5, type == "ta_oa", publisher == "Wiley") |> nrow()` and Springer Nature `r publisher_plot_df |>  mutate(prop = value / jn_all)  |> filter(cr_year == "2022", prop > 0.5, type == "ta_oa", publisher == "Springer Nature") |> nrow()` hybrid journals with an open access share above 50% that could be solely attributed to transformative agreements.
-Upon inspection, these journals were mainly society or local language journals with a small yearly article volume.
+Simultaneously, the median proportion increased from `r pub_75 |> filter(publisher == "Wiley", type == "ta_oa", cr_year == "2020") |> pull(median) * 100`% to  `r pub_75 |> filter(publisher == "Wiley", type == "ta_oa", cr_year == "2022") |> pull(median) * 100`%.
+It is interesting to note that a small but increasing number of journals from these two publishers provide open access to the majority of articles through transformative agreements.
+Wiley recorded `r publisher_plot_df |>  mutate(prop = value / jn_all)  |> filter(cr_year == "2022", prop > 0.5, type == "ta_oa", publisher == "Wiley") |> nrow()` and Springer Nature `r publisher_plot_df |>  mutate(prop = value / jn_all)  |> filter(cr_year == "2022", prop > 0.5, type == "ta_oa", publisher == "Springer Nature") |> nrow()` hybrid journals with an open access share above 50%, which could be attributed solely to transformative agreements.
+Upon inspection, these journals were mainly society or local language journals with small annual article volumes.
 
@@
 Table \ref{tab:subject_summary_table} presents a high-level overview of hybrid open access by AJCS subject area using fractionalised counting to account for journals belonging to more than one category.
-Between 2018 and 2022, most hybrid journals with at least one open articles could be attributed to the category Social Sciences, which also includes the Arts and Humanities.
-However, these journals published the fewest number of articles, whereas Physical Sciences journals recorded most articles, followed by the Health Sciences and the Life Sciences. 
-In terms of open access,  Physical Sciences journals accounted for more than one third of articles published in the five-years period, followed by the Health Science, the Social Sciences and the Life Sciences.
+Between 2018 and 2022, most hybrid journals with at least one open article could be attributed to the Social Sciences category, which also includes Arts and Humanities.
+However, these journals published the fewest number of articles, whereas Physical Sciences journals recorded the most articles, followed by Health Sciences and Life Sciences. 
+In terms of open access,  Physical Sciences journals accounted for more than one-third of the articles published in the five-year period, followed by Health Sciences, Social Sciences and Life Sciences.
 
@@
 Figure \ref{fig:subject_panel} presents the relative growth of hybrid open access by subject area between 2018-2022.
-In particular, Social Sciences and Humanties journals accounted for the strongest growth in the five-years period from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Social Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Social Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(value) |> sum()`), followed by the Life Science from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Life Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Life Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Life Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Life Sciences") |> pull(value) |> sum()`) , Health Science from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Health Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Health Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Health Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Health Sciences") |> pull(value) |> sum()`) and Physical Sciences from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Physical Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Physical Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Physical Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Physical Sciences") |> pull(value) |> sum()`).
+In particular, Social Sciences inbcluding Arts and Humanties journals accounted for the strongest growth in the five-years period from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Social Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Social Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(value) |> sum()`), followed by the Life Sciences from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Life Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Life Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Life Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Life Sciences") |> pull(value) |> sum()`) , Health Sciences from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Health Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Health Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Health Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Health Sciences") |> pull(value) |> sum()`) and Physical Sciences from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Physical Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Physical Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Physical Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Physical Sciences") |> pull(value) |> sum()`).
 <!-- In particular, journals in the subject areas recorded large high rate (see supplement)-->
-Growth in the Social Sciences can be largely attributed to transformative agreements. 
-In 2022, two-third of open access articles (`r subject_ta_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(prop) * 100`%, n = `r subject_ta_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(ta_oa_frac)`) were published by first authors affiliated with participating institutions (see \ref{fig:subject_panel}B).
+Growth in Social Sciences category can be largely attributed to transformative agreements. 
+In 2022, two-thirds of open access articles (`r subject_ta_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(prop) * 100`%, n = `r subject_ta_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(ta_oa_frac)`) were published by the first authors affiliated with participating institutions (see \ref{fig:subject_panel}B).
 
@@
 
-Figure \ref{fig:subject_panel}C shows that this trend is consistent across hybrid journals belonging to the ASJC subject area Social Sciences.
+Figure \ref{fig:subject_panel}C shows that this trend was consistent across hybrid journals belonging to the ASJC Social Sciences category.
 In 2022, 25% of Social Sciences journals provided open access to at least every fourth article exclusively through transformative agreements.
-However, hybrid open access through transformative agreements played a comparable lesser role in the Life Sciences and Health Sciences. 
-In these two subject areas, only about half of the open access articles can be linked to these agreements, both overall and on median average across journals.
-In contrast, the majority of Physical Science Journals, shows an increase of open access through transformative agreements compared to other options to publish open access in hybrid journals.
+However, hybrid open access through transformative agreements played a comparably lesser role in Life Sciences and Health Sciences. 
+In these two subject areas, only about half of the open access articles can be linked to these agreements, both overall and on a median average across journals.
+In contrast, the majority of Physical Science Journals show an increase of open access through transformative agreements compared to other options to publish open access in hybrid journals.
 
@@
 
-Figure \ref{fig:country_patch}A shows the development of hybrid open access publishing by countries, comparing the OECD area with the BRICS, an intergovernmental organisation, which comprised the countries Brazil, Russia, India, China and South Africa as of 2022.
+Figure \ref{fig:country_patch}A shows the development of hybrid open access publishing by country, comparing the OECD area with the BRICS, an intergovernmental organisation, which comprised  Brazil, Russia, India, China, and South Africa as of 2022.
 The residual category "Other" includes the remaining countries. 
+Full counting was applied to account for multiple institutional affiliations [@Hottenrott_2021].
 From 2018 to 2022, the proportion of open access in hybrid journals increased from `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2018) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2018 to `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2022) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2022.
-On the other hand, BRICS recorded a low uptake, from `r country_bar$data |> filter(cat == "BRICS", oa_type != "closed", cr_year == 2018) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2018 to `r country_bar$data |> filter(cat == "BRICS", oa_type != "closed", cr_year == 2022) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2022.
+In contrast, BRICS recorded a low uptake, moderately growing from `r country_bar$data |> filter(cat == "BRICS", oa_type != "closed", cr_year == 2018) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2018 to `r country_bar$data |> filter(cat == "BRICS", oa_type != "closed", cr_year == 2022) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2022.
+
 Despite the rise of open access across OECD countries, the overall publication output decreased sharply, dropping to `r country_bar$data |> filter(cat == "OECD", cr_year == "2022") |> distinct(all) |> pull(all)` in 2022 after peaking `r country_bar$data |> filter(cat == "OECD", cr_year == "2020") |> distinct(all) |> pull(all)` articles in 2020.
 In stark contrast, the number of articles published in hybrid journals by first authors affiliated with institutions from BRICS countries increased steadily over the years, more than doubling from `r country_bar$data |> filter(cat == "BRICS", cr_year == "2018") |> distinct(all) |> pull(all)` in 2018 to `r country_bar$data |> filter(cat == "OECD", cr_year == "2022") |> distinct(all) |> pull(all)` in 2022.
-Upon closer examination, this trend can be observed across all big three publishers, although the shift towards BRICS is particularly evident in Elsevier's hybrid journal portfolio, in particular with regard to articles published in Physical Sciences journals.
+Upon closer examination, this trend can be observed across all of the three largest publishers, although the shift towards BRICS is particularly evident in Elsevier's hybrid journal portfolio, particularly with regard to articles published in Physical Sciences journals.
 While OECD publication output in Elsevier's Physical Sciences journals declined from `r publ_subject_output_df |> filter(cr_year == 2018, top_level == "Physical Sciences", publisher == "Elsevier", cat == "OECD") |> pull(all)` articles in 2018 to `r publ_subject_output_df |> filter(cr_year == 2022, top_level == "Physical Sciences", publisher == "Elsevier", cat == "OECD") |> pull(all)` in 2022, BRICS output increased from `r publ_subject_output_df |> filter(cr_year == 2018, top_level == "Physical Sciences", publisher == "Elsevier", cat == "BRICS") |> pull(all)` to `r publ_subject_output_df |> filter(cr_year == 2022, top_level == "Physical Sciences", publisher == "Elsevier", cat == "BRICS") |> pull(all)` in the same five-year period.
-Furthermore, OECD publication output in Health Science Journals and Life Science journals stagnated across the investigated hybrid journal portfolios after a peak in 2020.
+Furthermore, OECD publication output in Health Science Journals and Life Science journals stagnated across the investigated hybrid journal portfolios after peaking in 2020.
 
@@
 
-To illustrate the situation in 2022, Figure \ref{fig:country_patch}B compares total publication output with the number of open access articles. 
+To illustrate the situation in 2022, Figure \ref{fig:country_patch}B compares the total publication output with the number of open access articles. 
 With `r country_scatter_df |> filter(cr_year == "2022", country_code == "CN") |> pull(all)` articles, China was the most productive country, followed by the United States (`r country_scatter_df |> filter(cr_year == "2022", country_code == "US") |> pull(all)` articles) and India (`r country_scatter_df |> filter(cr_year == "2022", country_code == "IN") |> pull(all)` articles).
-In contrast, West and Nord European countries published a considerable high number of open access articles. 
-Particularly, Germany, Great Britain, the Netherlands, Sweden, Switzerland and Spain recorded an above-average open access share as indicated by the linear trend line.
-As represented by the point size, as well as it can been seen in Figure \ref{fig:country_patch}C, transformative agreements contributed to these market positions. 
+In contrast, Western and Northern European countries published a considerably high number of open access articles. 
+Particularly, Germany, Great Britain, the Netherlands, Sweden, Switzerland, and Spain recorded an above-average open access share, as indicated by the linear trend line.
+As represented by the point size, as well as that shown in Figure \ref{fig:country_patch}C, transformative agreements contributed to these market positions. 
 Interestingly, the United States had a notable open access market share of `r top_20_plot$data |> distinct(country_code, oa, all) |> mutate(prop = oa / sum(oa)) |> filter(country_code == "US") |> pull(prop) * 100`%, although transformative agreements contributed to a lesser extent.
@@
 Figure \ref{fig:country_top_20_plot} illustrates the development of hybrid open access from 2018 to 2022, highlighting the top 20 most productive countries in terms of articles published in hybrid journals that were included in transformative agreements over the five-year period.
-Notably, The Netherlands (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "NL") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), Sweden (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "SE") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), Poland (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "PL") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%) and Great Britain (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "GB") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%) exhibited a relatively high level of uptake in 2018 which continued to increase in the following years.
+Notably, the Netherlands (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "NL") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), Sweden (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "SE") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), Poland (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "PL") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%) and Great Britain (`r country_plot$data |> filter(cr_year == "2018", country_name_fct == "GB") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%) exhibited a relatively high level of uptake in 2018 which continued to increase in the following years.
 In 2022, Sweden had the highest proportion of open access relative to its publication output (`r country_plot$data |> filter(cr_year == "2022", country_name_fct == "SE") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), followed by the Netherlands (`r country_plot$data |> filter(cr_year == "2022", country_name_fct == "NL") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%) and Switzerland (`r country_plot$data |> filter(cr_year == "2022", country_name_fct == "CH") |> distinct(all_total, oa_total) |> mutate(prop = oa_total / all_total) |> pull(prop)*100`%), with these countries benefiting from transformative agreements.
 In Germany, however, hybrid open access only began to increase from 2019 onwards after the successful negotiation of nationwide agreements with Wiley (July 2019) and Springer Nature (January 2020).
-Prior to this, only a few organisations had agreements in place, for the example the Max Planck Society with Springer Compact.
+Prior to this, only a few organisations had agreements in place, for example, the Max Planck Society with Springer Compact.
 
@@
 
-Since 2021, there has been a general trend towards hybrid open access among many high-income countries, primarily driven by transformative agreements.
-However, proliferation of transformative agreements differed across these countries.
+Since 2021, there has been a general trend towards hybrid open access among many high-income countries, driven primarily by transformative agreements.
+However, the proliferation of transformative agreements differed across these countries.
 For instance, Germany successfully negotiated an agreement with Elsevier not until 2023. 
-Additionally, publication limits or eligibility criteria for institutions and article types may explain why even countries with widespread agreement implementation like Sweden or the Nethterlands did not achieve 100% hybrid open access.
-Interestingly, in Japan and the US other options than transformative agreements were the main driver for the increase in hybrid open access.
+Additionally, publication limits or eligibility criteria for institutions and article types may explain why even countries with widespread agreement implementation such as Sweden or the Netherlands did not achieve 100% hybrid open access.
+Interestingly, in Japan and the US, options other than transformative agreements were the main drivers for the increase in hybrid open access.
 Once again, the graph highlights countries with low hybrid open access, particularly non-OECD countries, where only a few or no agreements were in place.
@@
 
-The primary aim of this study was to investigate the adoption of open access in hybrid journals in transformative agreements, which were introduced as a temporal means to support the transition of subscription-based academic publishing to full open access.
-This study presents a novel approach based on open data, which leverages metadata on over 700 agreements and nine million journal articles to estimate the extent to which transformation agreement contribute to the transition of this journal business model.
-The results highlight a strong growth in open access between 2018 and 2022, particularly driven by an increasing number of institutions that had transformative agreements in place. 
-Yet, the majority of research literature published in hybrid journals in this five-year period remained behind a publisher's paywall.
-Growth in the adoption of open access in hybrid journals, in particular through transformative agreements, can be largely attributed to three large commercial publishers -- Elsevier, Springer Nature, and Wiley -- but varies substantially across journals, publishers, subjects, and country affiliations.
-Despite the limitations of the data used, the findings indicate that the current level of implementation of transformative agreements is insufficient to bring about a large-scale transition to fully open access.
-
-A key finding of the analysis is that transformative agreements increase market concentration. 
-Specifically, the three largest commercial publishers Elsevier, Springer Nature, and Wiley dominate, particularly in regard to open access provided through transformative agreements.
-Together, the three publishers accounted for three-fourths of open access articles through transformative agreements, while recording less than half of total publication volume published between 2018 and 2022.
+The primary aim of this study was to assess the uptake of open access in hybrid journals included in transformative agreements, which were introduced as a temporal means to support the transition of subscription-based academic publishing to full open access.
+This study presents a novel approach based on open data, which leverages metadata on over 700 agreements and nine million journal articles to estimate the extent to which transformation agreements contribute to the transition of this journal business model.
+The results highlight a strong growth in open access between 2018 and 2022, driven by an increasing number of institutions that had transformative agreements in place. 
+However, the majority of research literature published in hybrid journals in this five-year period remained behind the publisher's paywall.
+Growth in the adoption of open access in hybrid journals, in particular through transformative agreements, can be largely attributed to three large commercial publishers -- Elsevier, Springer Nature, and Wiley -- but varies substantially across journals, publishers, disciplines, and country affiliations.
+Despite the limitations of the data, the findings indicate that the current level of implementation of transformative agreements is insufficient to bring about a large-scale transition to fully open access.
+
+A key finding of this analysis is that transformative agreements maintain market concentration. 
+Specifically, the three largest commercial publishers Elsevier, Springer Nature, and Wiley dominate, particularly with regard to open access provided through transformative agreements.
+Together, the three publishers accounted for three-fourths of open access articles through transformative agreements, while recording less than half of the total publication volume published between 2018 and 2022.
 This observation aligns with previous research on shifts in the publishing market following the introduction of funding opportunities for hybrid open access [@Jahn_2016; @Butler_2023; @Shu_2023]. 
@@
 Moreover, this study presents varying levels of open access uptake through transformative agreements across journals, which can be attributed to the alignment of authors' affiliations and the availability of such agreements at their institutions.
-In line with previous research findings [@Butler_2023; @Jahn_2021], high uptake rates were observed across hybrid journals in the Social Sciences and Humanities.
+In line with previous research findings [@Butler_2023; @Jahn_2021], high uptake rates were observed across hybrid journals in  Social Sciences and Humanities.
 However, it is important to emphasise that these hybrid journals do not encompass the entire field.
-For example, a study by @Khanna_2022 found that approximately 60% of journals utilizing the open-source publishing platform Open Journal Systems (OJS) fall within the Social Sciences and Humanities. 
-In these fields, the majority of full open access journals are so-called "Diamond OA journals" that do not charge publication fees [@bosman_2021_4558704].
-Rather, this result can be more accurately attributed to the substantial proportion of authors from high-income countries who publish in these hybrid journals included in transformative agreements, particularly when journals are of local relevance, such as through the belonging to national societies or regional research scope.
+For example, @Khanna_2022 found that approximately 60% of journals utilizing the open-source publishing platform Open Journal Systems (OJS) fall within the Social Sciences and Humanities. 
+In these fields, the majority of full open access journals are so-called "Diamond OA journals", which do not charge publication fees [@bosman_2021_4558704].
+Rather, this result can be more accurately attributed to the substantial proportion of authors from high-income countries who publish in these hybrid journals included in transformative agreements, particularly when journals are of local relevance, such as by belonging to national societies or regional research scope.
  <!--   [funding options in the humanities and social sicences] -->
@@
 Because the BRICS expansion can be exclusively attributed to closed access articles, this development has the potential to hinder the transition of academic journal publishing to full open access through transformative agreements and demands discussion.
-In China, the country with the highest volume of articles in hybrid journals, limited research funding to pay for open access is available, with expenditures for open access publishing already surpassing subscription costs [@Shu_2023].
+In China, the country with the highest volume of articles in hybrid journals, only limited research funding to pay for open access is available, with expenditures for open access publishing  surpassing subscription costs [@Shu_2023].
 Furthermore, the focus of Chinese authors on established journals may contribute to this trend [@Zhang_2022]. 
 Although China supports Plan S [@Schiermeier_2018], this is not reflected in the data in terms of open access uptake in hybrid journals.
-As highlighted by @Koley_2023, India faces similar challenges in terms of resources to pay for open access.
-At the same time, access to research literature is a pressing, which is adressed by the "Indian one nation, one subscription" policy proposal. 
+As highlighted by @Koley_2023, India faces similar challenges in terms of the availablility of resources to pay for open access.
+At the same time, access to research literature is a pressing issue, which is addressed by the "Indian one nation, one subscription" policy proposal. 
 However, this policy focuses on centrally negotiated subscriptions and does not entail open access.
@@
 But open access uptake also differs among OECD countries. 
-In the United States, for example, hybrid open access including transformative agreements plays a lesser role compared to some European countries. 
-Between 2017 and 2021, hybrid open access contributed the least to make federally funded research articles openly available [@Schares_2023]. 
-While some university consortia like the California Digital Library have signed transformative agreements, others attempt to depart from big deals and unbundle large journal portfolios to address cost increases [@Brainard_2021; @Schares_2022].
+In the United States, for example, hybrid open access, including transformative agreements, plays a lesser role than in some European countries. 
+Between 2017 and 2021, hybrid open access contributed the least to openly available federally funded research articles [@Schares_2023]. 
+While some university consortia, such as the California Digital Library, have signed transformative agreements, others have attempted to depart from big deals and unbundle large journal portfolios to address cost increases [@Brainard_2021; @Schares_2022].
 Despite the relatively low penetration of transformative agreements in China and the United States compared to Europe, 22% of open access in hybrid journals in 2022 originates from these two countries, indicating the availability of funding sources for publication fees.
 
-This large-scale study provides first empricial evidence about the influence of transformative agreements on the transition of hybrid journals to full open access. 
-However, several limitations needs to be acknowledged.
+This large-scale study provides the first empiricial evidence of the influence of transformative agreements on the transition of hybrid journals to full open access. 
+However, that study has several limitations that need to be acknowledged.
 From a data perspective, estimations of open access through transformative agreements were established by linking first author affiliations with publicly available agreement data from cOAlition S, and not through invoicing data, which is usually not shared.
-Moreover, this study is unable to account for the various types of transformative agreements, particularly with regards to article types and capping.
-Furthermore, assessing the quality of the OpenAlex and Crossref data used, particularly in terms of affiliations and article types, by comparing it with established bibliometric databases such as Scopus or Web of Science was beyond the scope of this analysis.
+Moreover, this study is unable to account for the various types of transformative agreements, particularly with regard to article types and caps that limit the number of open access articles covered.
+Furthermore, assessing the quality of the OpenAlex and Crossref data used, particularly in terms of affiliations and article types, by comparing them with established bibliometric databases such as Scopus and Web of Science was beyond the scope of this analysis.
 Overall, the methodology is designed in such a way that the results underestimate rather than overestimate the adoption of open access through transformative agreements. 
@@
 
-Additionally, it must be noted that the study period was significantly impacted by the COVID-19 pandemic, which led to an unprecedented number of publications as well as an reduction of international collaboration [@Aviv_Reuven_2021], which could explain the observed contrasting developements in OECD and BRICS countries.
-Even before the pandemic, growth in publications in Europe were only due to internationally co-authored journal articles [@Kwiek_2020]. 
-Likewise, inflows from China to the US and European countries already declined before 2020 [@Zhao_2023].
+Additionally, it must be noted that the study period was significantly impacted by the COVID-19 pandemic, which led to an unprecedented number of publications and a reduction in international collaboration [@Aviv_Reuven_2021], which could explain the observed contrasting developments in OECD and BRICS countries.
+Even before the pandemic, growth in publications in Europe was only due to internationally co-authored journal articles [@Kwiek_2020]. 
+Likewise, inflows from China to the US and European countries already declined by 2020 [@Zhao_2023].
 Furthermore, the study design did not consider emerging publication practices such as preprints [@Fraser_2021] and special issues [@hanson2023strain], which have grown rapidly since 2020. 
@@
 
-This present study allows multiple strands of further research. 
+This study allows for multiple strands of further research. 
 One is to complement this large-scale study with more specific evidence from individual countries or subjects, particularly those with low hybrid open access  rates.
 Incorporating full open access and subscription-based journals, as well as considering global trends in scholarly migration and collaboration could also be promising.
-Financial studies could build upon the presented study design and incorporate subscription and open access expenditures to evaluate the cost-effectiveness of transformative agreements.
-
-This study also has practival implications for research funders and libraries.
-Of primary concern should be the observed differences across countries, in particular the relationship between economic development and open access adoption.
-From a data perspective, the reporting of open access funding including transformative agreements is not harmonised, but often crowd-sourced from various sources.
+Financial studies could build upon the study design and incorporate subscription and open access expenditures to evaluate the cost-effectiveness of transformative agreements.
+
+This study has practical implications for research funding and libraries.
+Of primary concern should be the observed differences across countries, particularly the relationship between economic development and open access adoption.
+From a data perspective, the reporting of open access funding, including transformative agreements, is not harmonised, but often crowd-sourced from various sources.
 To improve the assessment of transformative agreements, libraries and publishers could agree on standards and services to publicly share information about respective journal portfolios, participating institutions and open access invoicing.
 
-In summary, this study provides empirical insights into the developement of hybrid open access following the introduction of transformative agreements. 
-The results are important for both researchers and stakeholders engaged in negotiating and evaluating these agreements. 
-The presented approach relies on open data, which enables follow-up studies and open access monitoring acitvties to further explore the role of transformative agreements in transitioning academic publishing to full and equitable open access.
+In summary, this study provides empirical insights into the development of hybrid open access following the introduction of transformative agreements. 
+These results are important for both researchers and stakeholders engaged in negotiating and evaluating these agreements. 
+The presented approach relies on open data, which enables follow-up studies and open access monitoring activties to further explore the role of transformative agreements in transitioning academic publishing to full and equitable open access.
 

### 2024-02-26 analysis/manuscript.tex
Birgits suggestions

@@
 open access to 1\% of articles in hybrid journals (Dallmeier-Tiessen et
-al., 2010). Other studies also recorded a low uptake. In 2011, only
-1-2\% of articles were open access (Björk, 2012), growing to around 4\%
-between 2011 and 2013 (Laakso \& Björk, 2016).
+al., 2010). Other studies have also recorded low uptake. In 2011, only
+1-2\% of articles were open access (Björk, 2012), increasing to around
+4\% between 2011 and 2013 (Laakso \& Björk, 2016).
 
 With the introduction of central funding mechanisms for publication fees
-in some European countries since 2012, an increase in hybrid open access
-could be observed (Björk, 2017; Huang et al., 2020; Jubb et al., 2017;
-Piwowar et al., 2018). For example, studying university output,
-Robinson-Garcia et al. (2020) estimated a median uptake of 7.1\% in the
-period 2014-2017. In particular, British (17\%), Austrian (15\%) and
-Dutch (13\%) universities stand out. However, this shift in funding
+in some European countries since 2012, an substantial increase in hybrid
+open access has been observed (Björk, 2017; Huang et al., 2020; Jubb et
+al., 2017; Piwowar et al., 2018). For example, studying university
+output, Robinson-Garcia et al. (2020) estimated a median uptake of 7.1\%
+in the period 2014-2017. In particular, British (17\%), Austrian (15\%)
+and Dutch (13\%) universities stood out. However, this shift in funding
 policy towards hybrid open access also added to the overall cost of
@@
 efforts required to handle payments (Pinfield et al., 2016). Moreover,
-established large commercial publishers, which already dominated the
-publishing market (Larivière et al., 2015), disproportionately benefited
-from hybrid open access funding in comparison to full open access
-publishers (Butler et al., 2023; Jahn \& Tullney, 2016; Shu \&
-Larivière, 2023).
+large commercial publishers, which already dominated the publishing
+market (Larivière et al., 2015), disproportionately benefited from
+hybrid open access funding in comparison to full open access publishers
+(Butler et al., 2023; Jahn \& Tullney, 2016; Shu \& Larivière, 2023).
 
@@
 publisher-provided immediate open access (Björk \& Solomon, 2014;
-Schimmer et al., 2015). These considerations resulted into
-transformative agreements\footnote{In this paper I use the term
-  ``transformative agreement'', addressing also offsetting,
-  read-and-publish or publish-and-read deals, and other variants
-  (Borrego et al., 2021; Hinchliffe, 2019). Although the term is
-  critised as misleading and not useful to describe the different types
-  of open access agreements between library consortia and commercial
-  publishers (Babini et al., 2022), it is widely used in policy
-  discussions and in the research literature.}, which cover a broad
-range of contracts between library consortia and publishers from
-mid-2010s onwards where institutional spending for subscriptions and
-open access publishing are considered together (Borrego et al., 2021;
-Hinchliffe, 2019). Transformative agreements seek to control costs while
-allowing a transitional phase for publishing more open access articles.
-Similar to big deals, transformative agreements mainly bundle hybrid and
-subscription-only journals from commercial publishers, but aim at a
-higher degree of transparency than previous big deals, where contracts
-including payments were confidential (Bergstrom et al., 2014).
+Schimmer et al., 2015). These considerations resulted in transformative
+agreements\footnote{In this paper I use the term ``transformative
+  agreement'', addressing also offsetting, read-and-publish or
+  publish-and-read deals, and other variants (Borrego et al., 2021;
+  Hinchliffe, 2019). Although the term is critised as misleading and not
+  useful to describe the different types of open access agreements
+  between library consortia and commercial publishers (Babini et al.,
+  2022), it is widely used in policy discussions and in the research
+  literature.}, which cover a broad range of contracts between library
+consortia and publishers from the mid-2010s onwards, where institutional
+spending for subscriptions and open access publishing are considered
+together (Borrego et al., 2021; Hinchliffe, 2019). Transformative
+agreements seek to control costs while allowing a transitional phase for
+publishing more open access articles. Similar to big deals,
+transformative agreements mainly bundle hybrid and subscription-only
+journals from commercial publishers, but aim at a higher degree of
+transparency than previous big deals, where contracts including payments
+were confidential (Bergstrom et al., 2014).
 
 The introduction of transformative agreements aligns with funding policy
-changes, such as the decision made by the cOAlition S, a consortium of
-national funders including the European Commission, to no longer provide
-financial support for individual publication fees when publishing in
-hybrid journals. According ot its Plan S launched in 2018, the cOAlition
-S funders only accept hybrid open access through transformative
-agreements ``during a transition period that should be as short as
-possible'' (Schiltz, 2018). Specifically, they agreed to support hybrid
-open access only through transformative agreements from 2021, until the
-end of 2024. Notably, the German Research Foundation (DFG), despite not
-being part of cOAlition S, has also extended its financial support for
-hybrid open access through transformative agreements (Mittermaier,
-2021). Previously, the DFG only provided funding for full open access
-journals (Jahn \& Tullney, 2016).
-
-By the end of 2023, many transformative agreements were implemented, but
-interim outcomes are mixed. The ESAC Transformative Agreement
+changes, such as the decision made by cOAlition S, a consortium of
+funders and research organisations including the European Commission, to
+no longer provide financial support for individual publication fees when
+publishing in hybrid journals. According to its Plan S launched in 2018,
+cOAlition S members only accept hybrid open access through
+transformative agreements ``during a transition period that should be as
+short as possible'' (Schiltz, 2018). Specifically, they agreed to
+support hybrid open access only through transformative agreements from
+2021, until the end of 2024. Notably, despite not being part of
+cOAlition S, the German Research Foundation (DFG), has also extended its
+financial support for hybrid open access through transformative
+agreements (Mittermaier, 2021). Previously, the DFG only provided
+funding for fully open access journals (Jahn \& Tullney, 2016).
+
+By the end of 2023, many transformative agreements had been implemented,
+but the interim outcomes were mixed. The ESAC Transformative Agreement
 Registry\footnote{\url{https://esac-initiative.org/about/transformative-agreements/agreement-registry/}},
@@
 agreements, resulting in up to 900.000 open access articles published in
-both full open access and hybrid journals according to the accompanying
+both fully open access and hybrid journals according to the accompanying
 ESAC Market Watch\footnote{\url{https://esac-initiative.org/market-watch/}}.
 Library consortia reported increased open access volume, streamlined
-payment and monitoring procedures, as well as extensive utilization of
+payment and monitoring procedures, as well as extensive utilisation of
 open access options by the researchers they serve (Marques \& Stone,
 2020; Parmhed \& Säll, 2023; Pinhasi et al., 2020). The ongoing
-standardisation of transformative agreements contributed to improved
+standardisation of transformative agreements has contributed to improved
 transparency in terms of contracts and publisher-provided article
@@
 growing trend toward transformative agreements, continued reliance on
-big deals is perceived as problematic, because it perpetuate market
+big deals is perceived as problematic, because it perpetuates market
 concentration (Butler et al., 2023; Shu \& Larivière, 2023). Whether
@@
 inequality (Klebel \& Ross-Hellauer, 2023), because transformative
-agreements focus on pay to publish open access mainly targets
-institutions from high-income countries, furthering a questionable
-journal prestige culture (Babini et al., 2022). Besides, an
-editorial-board resignation raised concerns that publishers' desire to
-maximize journal publication volume ``without regard to quality'' is a
-consequence of transformative agreements (Rasmussen, 2023).
+agreements' focus on pay to publish mainly targets institutions from
+high-income countries, furthering a questionable journal prestige
+culture (Babini et al., 2022). Besides, an editorial-board resignation
+raised concerns that transformative agreements might encourage
+publishers to maximize journal publication volume ``without regard to
+quality'' (Rasmussen, 2023).
 
@@
 Association of Swedish Higher Education Institutions (Sveriges
-universitets- och högskoleförbund, SUHF) recommended to only support
-agreements for publishing in full open access journals \footnote{\url{https://www.su.se/english/news/open-access-need-to-move-away-from-transformative-agreements-1.683787}}.
+universitets- och högskoleförbund, SUHF) recommended only supporting
+agreements for publishing in fully open access journals \footnote{\url{https://www.su.se/english/news/open-access-need-to-move-away-from-transformative-agreements-1.683787}}.
 Likewise, most cOAlition S funders will discontinue financial support
 for transformative agreements by the end of 2024 (Liverpool, 2023). The
-consortium also removed most hybrid journals from its Transformative
-Journal program in 2023 due to publishers' failure to meet self-defined
-open access growth targets (Brainard, 2023). In contrast, the German
-DEAL consortium announced a five-year transformative agreement with
-Elsevier starting in 2024, while also renewing its contracts with
-Springer Nature and Wiley until the end of 2028. At the same time, the
-Colombia Consortium signed the first transformative agreements in Latin
-America (Muñoz-Vélez et al., 2024).
+consortium also removed the majority of hybrid journals from its
+Transformative Journal program in 2023 due to publishers' failure to
+meet self-defined open access growth targets (Brainard, 2023). By
+contrast, the German DEAL consortium announced a five-year
+transformative agreement with Elsevier starting in 2024 and also renewed
+its contracts with Springer Nature and Wiley until the end of 2028.
+Similarly, the Colombia Consortium signed the first transformative
+agreements in Latin America (Muñoz-Vélez et al., 2024).
 
@@
 limited evidence available on the adoption of open access in hybrid
-journals, and the extent to which this uptake can be attributed to
+journals, and the extent to which this can be attributed to
 transformative agreements. Previous studies have focused on specific
@@
 relied on self-reported agreement data (Moskovkin et al., 2022), or used
-APC pricing lists (Shu \& Larivière, 2023). Particularly, data
+APC pricing lists (Shu \& Larivière, 2023). In particular, data
 availability is a limiting factor when studying the impact of
@@
 transformative agreements. Likewise, article-level open access invoicing
-data is only partly available (Jahn et al., 2022).
+data are only partly available (Jahn et al., 2022).
 
 The present study aims to address these limitations by combining
-multiple openly available data sources to determine the open access
-uptake in hybrid journals, while distinguishing between open access
-through transformative agreements and other means. With this novel and
-open approach, this first large-scale analysis aims to answer the
-following questions:
+multiple openly available data sources to determine open access uptake
+in hybrid journals, while distinguishing between open access through
+transformative agreements and other means. With this novel and open
+approach, this first large-scale analysis will answer the following
+questions:
 
@@
 
-For both of these research questions, this study will analyse the
-variability by publisher, journal subject, and country.
+For both research questions, this study will analyse the variability by
+publisher, journal subject, and country.
 
@@
 
-This study combines data from multiple publicly available data sources
-as shown in Figure \ref{fig:data_workflow}. Initially, transformative
-agreement data retrieved from the cOAlition S Journal Checker
+This study combines data from multiple publicly available sources, as
+shown in Figure \ref{fig:data_workflow}. Initially, transformative
+agreement data retrieved from cOAlition S Journal Checker
 Tool\footnote{\url{https://www.coalition-s.org/blog/enabling-accurate-results-within-the-journal-checker-tool/}}
 provided information about journal portfolios and participating
-institutions. After identification of hybrid journals by excluding full
-open access journals, Crossref served as the primary data source for
+institutions. After identifying hybrid journals by excluding fully open
+access journals, Crossref served as the primary data source for
 article-level metadata including Creative Commons (CC) license
@@
 were subsequently linked to eligible institutions according to the
-transformative agreement data. In the following, these steps are
-described in more detail.
+transformative agreement data. In the following, the steps are described
+in detail.
 
@@
 the majority of transformative agreements. Overall, the four combined
-Transformative Agreement Data dumps used in this study contained 729 out
-of 869 agreements listed in the ESAC registry by December 2023.
+Transformative Agreement Data dumps used in this study contained 729 of
+869 agreements listed in the ESAC registry by December 2023.
 
 The Transformative Agreement Data dumps link agreements to journals
-represented by journal names and ISSN. After mapping ISSN variants to
-the corresponding linking ISSN (ISSN-L) as provided by the ISSN
-International Centre, journals were associated to publishers using the
+represented by journal names and the ISSN. After mapping ISSN variants
+to the corresponding linking ISSN (ISSN-L), as provided by the ISSN
+International Centre, journals were associated with publishers using the
 ESAC ID, a unique identifier for transformative agreements in the ESAC
@@
 
-Because transformative agreements can include both full open access and
+Because transformative agreements can include both fully open access and
 hybrid journals, the data were complemented with information about a
@@
 Open Access Journals (DOAJ) downloaded on 12 December 2023\footnote{\url{https://doaj.org/csv}},
-OpenAlex (November 2023) and the the Bielefeld list of GOLD OA journals
+OpenAlex (November 2023) and the Bielefeld list of GOLD OA journals
 (Bruns et al., 2022). As shown in Figure \ref{fig:method_fig}A,
-combining different data sources considerably extended the journal
-matching. In total, 3,439 full open access journals were excluded based
-on ISSN matching. The overlap between the three data sources was 72\%.
-The Gold OA journals dataset alone added 176 journals, while the DOAJ
-comprised 10 full open access journals not listed in either of the other
-two sources. These full open access journals were mostly launched in
-2022.
+combining different data sources considerably extended journal matching.
+In total, 3,439 full open access journals were excluded based on ISSN
+matching. The overlap between the three data sources was 72\%. The Gold
+OA journals dataset alone added 176 journals, while the DOAJ comprised
+10 fully open access journals not listed in either of the other two
+sources. These fully open access journals were launched in 2022.
 
@@
 After identifying hybrid journals included in transformative agreements,
-article metadata was retrieved from the Crossref November 2023 database
-snapshot for the five-year period 2018 to 2022 according to the issued
-date, representing the earliest known publication date. Because Crossref
-metadata lacked information to distinguish between original research
-articles including review and other types of journal content, which are
-often not covered by transformative agreements (Borrego et al., 2021),
-only articles published in regular issues indicated by non-numeric
-pagination were included. Furthermore, an expanded version of
-Unpaywall's paratext recognition approach was applied to exclude
-non-scholarly journal content such as table of contents.
-
-Open access articles in hybrid journals were identified through Creative
-Commons (CC) license information in Crossref metadata. License
-information relative to the ``accepted manuscript (AM)'' version were
-not considered. Crossref was used for open access identification because
-transformative agreements workflows generally require publishers to
+article metadata were retrieved from the Crossref November 2023 database
+snapshot for the five-year period from 2018 to 2022, according to the
+issued date, representing the earliest known publication date. Because
+Crossref metadata lacked information to distinguish between original
+research articles, including reviews, and other types of journal
+content, which are often not covered by transformative agreements
+(Borrego et al., 2021), only articles published in regular issues
+indicated by non-numeric pagination were included. Furthermore, an
+expanded version of Unpaywall's paratext recognition approach was
+applied to exclude non-scholarly journal content such as table of
+contents.
+
+Open access articles in hybrid journals were identified using the
+Creative Commons (CC) license information in Crossref metadata. License
+information relative to the ``accepted manuscript (AM)'' version was not
+considered. Crossref was used for open access identification because
+transformative agreement workflows generally require publishers to
 deliver CC license information to this DOI registration agency (Geschuhn
@@
 service that also parses journal websites for open content licenses
-(Piwowar et al., 2018), highlighted ongoing challenges to identify
+(Piwowar et al., 2018), highlighted ongoing challenges in identifying
 hybrid open access (Butler et al., 2023; Jahn et al., 2022;
 Martín-Martín et al., 2018; Zhang et al., 2022). Here, 742,369 articles
-with CC license were retrieved using Crossref, while 950,260 articles
+under CC license were retrieved using Crossref, while 950,260 articles
 were tagged as ``hybrid'' according to the OpenAlex November 2023
-release, which was used throughout this study. The biggest differences
+release, which was used throughout this study. The largest differences
 concerned articles published between 2018 and 2020. In 2022, however,
-Crossref and OpenAlex open access numbers only differ slightly (249,511
-records using Crossref vs.~255,344 in OpenAlex). Notable difference
-could be furthermore observed among some publishers that presumably did
-not provide CC license information to Crossref including AIP Publishing,
-American Physiological Society, Emerald and the Royal Society. Crossref
-license metadata was more complete with regard to articles from the
-publisher Wiley and American Chemical Society. Finally, inconsistent
-open access status information in previous OpenAlex versions was
-observed (Jahn et al., 2023). After reporting it to OpenAlex, fixing
-this issue was still ongoing according to the release notes, which might
-also explain part of the discrepancy.
-
-After retrieving article metadata, the publication volume including open
-access was calculated per journal. To improve the identification of
-hybrid journals, journals with an open access proportion above 95\% were
-excluded. This further step allowed to remove additional 241 full open
-access journals.
+Crossref and OpenAlex open access numbers differ only slightly (249,511
+records using Crossref vs.~255,344 in OpenAlex). Notable differences
+could be observed among some publishers that presumably did not provide
+CC license information to Crossref, including AIP Publishing, the
+American Physiological Society, Emerald, and the Royal Society. Crossref
+license metadata was more complete with regard to the articles published
+by Wiley and the American Chemical Society. Finally, inconsistent open
+access status information in previous OpenAlex versions was observed
+(Jahn et al., 2023). After reporting this to OpenAlex, fixing this issue
+was still ongoing according to the release notes, which might also
+explain part of the discrepancy.
+
+After retrieving the article metadata, the publication volume, including
+open access, was calculated for each journal. To improve the
+identification of hybrid journals, journals with an open access
+proportion above 95\% were excluded. This step allowed removing
+additional 241 fully open access journals.
 
 Affiliation metadata about corresponding authors are crucial for the
-planning and evaluation of transformative agreements, because they are
-considered to be responsible to arrange open access publication (Borrego
-et al., 2021; Geschuhn \& Stone, 2017; Schimmer et al., 2015). Here,
-country and institutional affiliations were retrieved from OpenAlex.
-However, because of low coverage in OpenAlex, this study focused on
-first authors and their affiliations instead. First authors typically
-contribute most to a paper and are often considered lead author research
-papers (Larivière et al., 2016). Related studies therefore assumed first
-authors as a proxy to measure to open access payments and the impact of
-transformative agreement (Haucap et al., 2021; Shu \& Larivière, 2023;
-Zhang et al., 2022). Overall, around 90\% of studied articles had first
-author affiliation metadata in OpenAlex, whereas the coverage of
-articles with corresponding author information was around 54\%.
-
-To assess the impact of transformative agreements to hybrid open access,
+planning and evaluation of transformative agreements because they are
+considered responsible for arranging open access publication (Borrego et
+al., 2021; Geschuhn \& Stone, 2017; Schimmer et al., 2015). For this
+study, country and institutional affiliations were retrieved from
+OpenAlex. However, because corresponding author affiliation was only
+available for 54\% of the investigated articles in OpenAlex, this study
+focused on first authors and their affiliations instead; approximately
+90\% of the investigated articles had first author affiliation metadata
+in OpenAlex. First authors typically contribute the most to a paper and
+are often considered being lead authors (Larivière et al., 2016).
+Related studies also assumed first authors as a proxy to measure open
+access payments and the impact of transformative agreements (Haucap et
+al., 2021; Shu \& Larivière, 2023; Zhang et al., 2022).
+
+To assess the impact of transformative agreements on hybrid open access,
 participating institutions from the Transformative Agreement Data dump,
 which were crowd-sourced from the agreements and consortia that
-successfully negotiated an agreement, were matched with first author
-affiliations recorded by OpenAlex using the ROR-ID. The matching also
-took into account the duration of an agreement according to the ESAC
-registry. Upon inspection, Transformative Agreement Data did not cover
-associated institutions comprehensively like university hospitals or
-institutes of large research organisations like the Max Planck Society.
-To improve the matching, Transformative Agreement Data was automatically
-enriched with associated organisations using OpenAlex's institution
-entity data.
-
-In total, the compiled data set consists of 8,922,146 articles published
-in 12,857 hybrid journals between 2018 and 2022 (see Figure
+successfully negotiated an agreement, were matched with the first author
+affiliations recorded by OpenAlex using the ROR-ID. Matching also
+considered the duration of agreements according to the ESAC registry.
+Upon inspection, Transformative Agreement Data did not comprehensively
+cover associated institutions, such as university hospitals or
+institutes of large research organisations such as the Max Planck
+Society. To improve the matching, Transformative Agreement Data were
+automatically enriched with associated organisations using OpenAlex's
+institution entity data.
+
+The so compiled data set consists of 8,922,146 articles published in
+12,857 hybrid journals between 2018 and 2022 (see Figure
 \ref{fig:method_fig}B). Hybrid journals in transformative agreements
-represented 40\% of total global output over the same time period
-according to Crossref, while full open access journals recorded 35\%.
+represented 40\% of the total global output over the same time period,
+according to Crossref, while fully open access journals recorded 35\%.
 
@@
 (B) Number of articles in Crossref by journal types.
-The blue bars show the overall article volume of hybrid journals in transformative agreements, which were initially included in the study, in comaprision with full open access journals according to OpenAlex.
-The remainder represents closed access journals not covered by transformative agreements.}\label{fig:method_fig}
+The blue bars show the overall article volume of hybrid journals in transformative agreements, which were initially included in the study, in comparision with full open access journals according to OpenAlex.
+The remainder represents subscription-based journals not covered by transformative agreements.}\label{fig:method_fig}
 \end{figure}
@@
 Throughout this mostly automated data gathering and analysis process,
-tools from the Tidyverse (Wickham et al., 2019) for the R programming
-language (R Core Team, 2020) were used. The resulting data is openly
-available through an R data package, hoaddata, version 0.2.91 Following
-Marwick et al. (2018), hoaddata contains not only the datasets used in
-the data analysis. It also includes code used to compile the data by
-connecting it the cloud-based Google Big Query data warehouse, where the
-big scholarly data from Crossref, OpenAlex and Unpaywall were imported.
-To increase the computational reproducibility, the data aggregation
-through hoaddata was automatically carried out using GitHub Actions, a
+Tidyverse tools (Wickham et al., 2019) for the R programming language (R
+Core Team, 2020) were used. The resulting data are openly available
+through an R data package, hoaddata, version 0.2.91 Following Marwick et
+al. (2018), hoaddata contains not only the datasets used in the data
+analysis. It also includes code used to compile the data by connecting
+it to the cloud-based Google Big Query data warehouse, where scholarly
+big data from Crossref, OpenAlex and Unpaywall were imported. To
+increase computational reproducibility, data aggregation through
+hoaddata was automatically carried out using GitHub Actions, a
 continuous integration service.
@@
 
-Between 2018 and 2022, a total of 11,189 out of 12,857 hybrid journals
-in transformative agreements published at least one open access article
+Between 2018 and 2022, 11,189 out of 12,857 hybrid journals in
+transformative agreements published at least one open access article
 under a Creative Commons license. During this period, these hybrid
-journals provided open access to 742,369 out of 8,146,958 articles,
+journals provided open access to 742,369 of the 8,146,958 articles,
 representing a five-year open access proportion of 9.1\%. Authors who
 could make use of transformative agreements at the time of publication
-contributed 328,957 open access articles to the total.
+contributed 328,957 open access articles.
 
@@
 on this trend. Over the five-years period from 2018 to 2022, open access
-increased from 4.3\% (n = 65,486) to 15\% (n = 249,511). At the same
-time, the total article volume of the investigated journals grew from
+increased from 4.3\% (n = 65,486) to 15\% (n = 249,511). Simultaneously,
+the total article volume of the investigated journals increased from
 1,528,051 in 2018 to 1,676,928 in 2022.
@@
 in 2021 and 2022, contributing 58\% of the total open access article
-volume in 2022. However, there was also a notable growth in open access
-provision through other means, presumably publication fees being not
-invoiced through transformative agreements, which increased from 4.1\%
-(n = 62,625) in 2018 to 6.3\% (n = 105,896).
+volume in 2022. However, there was also a notable increase in open
+access provision through other means, presumably publication fees being
+not invoiced through transformative agreements, which increased from
+4.1\% (n = 62,625) in 2018 to 6.3\% (n = 105,896).
 
@@
 the majority of their articles under a Creative Commons license in the
-same year. These journals were, on average, smaller (M = 75, SD = 186)
+same year. On average, these journals were smaller (M = 75, SD = 186)
 than those with an open access share below 50\% (M = 164, SD = 347).
 
-When comparing the impact of open access trough transformative
-agreements across journals, it shows that for many journals these
+When comparing the impact of open access through transformative
+agreements across journals, it shows that for many journals, these
 agreements substantially contributed to the growth of open access over
@@
 transformative agreements, it is worth noting that other means of
-publishing open access remained common across the investiagted hybrid
+publishing open access remained common across the investigated hybrid
 journals. In total, 9,153 journals published open access articles from
@@
 reveals a large market concentration. Although 48 publishers offered
-transformative agreements, the big three commercial publishers Elsevier,
-Springer Nature, and Wiley accounted for 49\% of total article volume
-published (see Table \ref{tab:publisher_league_table}). Together, they
-published 500,878 or 66\% of open access articles in hybrid journals.
-Elsevier, Springer Nature, and Wiley made 243,891 articles open access
-in hybrid journals through transformative agreements, resulting in an
-even larger market share of 74\%.
+transformative agreements, the three commercial publishers Elsevier,
+Springer Nature, and Wiley, accounted for 49\% of the total article
+volume published (see Table \ref{tab:publisher_league_table}). Together,
+they published 500,878 or 66\% of the open access articles in hybrid
+journals. Elsevier, Springer Nature, and Wiley made 243,891 articles
+open access in hybrid journals through transformative agreements,
+resulting in an even larger market share of 74\%.
 
@@
 
-However, there are notable differences among the three big publishers.
-Although Elsevier published the largest volume of articles (n =
-2,770,826, 34\%), it recorded a comparable low number of open access
-articles, including those that can be associated with transformative
+However, there are notable differences between the three large
+publishers. Although Elsevier published the largest volume of articles
+(n = 2,770,826, 34\%), it recorded a comparably low number of open
+access articles, including those that are associated with transformative
 agreements. In contrast, Springer Nature and Wiley provided open access
 to a larger proportion of their articles (13\% of Springer Nature
-articles and 15\% of Wiley articles were open acccess), leading to
-higher open access market shares (23\% Springer Nature resp. 23\%
-Wiley). This difference between Elsevier on the one hand and Springer
-Nature and Wiley on the other can be attributed to transformative
-agreements, as the latter made the majority of their open access
-articles available through such deals (Springer Nature 57\% resp. Wiley
-55\%).
+articles and 15\% of Wiley articles were open access), leading to higher
+open access market share (23\% Springer Nature resp. 23\% Wiley). This
+difference between Elsevier on the one hand and Springer Nature and
+Wiley on the other can be attributed to transformative agreements, as
+the latter made the majority of their open access articles available
+through such deals (Springer Nature 57\% resp. Wiley 55\%).
 
 Figure \ref{fig:publisher_figure} takes a closer look into the growth of
-hybrid open access across publishers by year with a focus on open access
-enabled by transformative agreements. Although all publishers show a
-general long-term trend towards transformative agreements, Figure
-\ref{fig:publisher_figure}A and B indicate that, in particular, Wiley
-experienced a substantial increase in its open access share from 5.9\%
-(n = 11,628) in 2018 to 26\% (n = 53,503) in 2022, representing an
-4.5-fold increase. In contrast, Elsevier's hybrid journals demonstrated
-a more modest increase, from 3.3\% (n = 16,872) in 2018 to 10\% (n =
-60,821) in 2022, which is a relatively low open access share compared to
-the general trend. In 2018, Springer Nature had the largest open access
-proportion among the three publishers of 8.4\% (n = 19,701), but
-experienced a relatively slower growth, resulting in 18\% (n = 52,616)
-of articles being open access in Springer Nature hybrid journals in
-2022.
+hybrid open access across publishers by year, with a focus on open
+access enabled by transformative agreements. Although all publishers
+show a general long-term trend towards transformative agreements, Figure
+\ref{fig:publisher_figure}A and B indicate that Wiley experienced a
+substantial increase in its open access share from 5.9\% (n = 11,628) in
+2018 to 26\% (n = 53,503) in 2022, representing a 4.5-fold increase. In
+contrast, Elsevier's hybrid journals demonstrated a more modest
+increase, from 3.3\% (n = 16,872) in 2018 to 10\% (n = 60,821) in 2022,
+which is a relatively low open access share compared to the general
+trend. In 2018, Springer Nature had the largest open access proportion
+among the three publishers of 8.4\% (n = 19,701), but experienced a
+relatively slower growth, resulting in 18\% (n = 52,616) of articles
+being open access in Springer Nature hybrid journals in 2022.
 
@@
 publishers can be attributed to distinct approaches to transformative
-agreements. Springer Nature, for example, began in 2015 offering
-selected consortia, such as the Max Planck Society, the Swedish Bibsam
-consortium, and the Finnish FinELib consortium, open access agreements
-for its hybrid journal portfolio under the name Springer
-Compact\footnote{\url{https://web.archive.org/web/20180414062853id_/http://www.liber2015.org.uk/wp-content/uploads/2015/03/Springer-Compact.pdf}}.
-However, these agreements were not included in the data as they
-concluded prior to the start of the transformative agreement data
-collection in June 2021. Nonetheless, the results demonstrate the
-importance of agreements for Springer Nature's hybrid open access
-business over the past five years (Figure 2B). In 2022, 66\% (n =
-34,725) of open access in Springer Nature hybrid journals were enabled
-through transformative agreements. In the same year, 70\% (n = 37,316)
-of Wiley's open access articles could be linked to transformative
-agreements in 2022. In contrast, Elsevier published fewer than half of
-its open access articles through transformative agreements (n = 32,627;
-54\%).
-
-The increasing trend towards transformative agreements can be also
+agreements. Springer Nature, for example, started offering open access
+agreements for its hybrid journal portfolio to selected consortia such
+as the Max Planck Society, the Swedish Bibsam Consortium and the Finnish
+FinELib consortium in 2015 under the name Springer Compact\footnote{\url{https://web.archive.org/web/20180414062853id_/http://www.liber2015.org.uk/wp-content/uploads/2015/03/Springer-Compact.pdf}}.
+Prior to this, Springer discontinued its hybrid OA pilot agreements with
+a small number of institutions, including the University of Göttingen
+(Schmidt \& Shearer, 2012). However, the Springer Compact agreements
+were not included in the data as they concluded prior to the start of
+the transformative agreement data collection in June 2021. Nonetheless,
+the results demonstrate the importance of agreements for Springer
+Nature's hybrid open access business over the past five years (Figure
+2B). In 2022, 66\% (n = 34,725) of open access in Springer Nature hybrid
+journals was enabled through transformative agreements. In the same
+year, 70\% (n = 37,316) of Wiley's open access articles could be linked
+to transformative agreements in 2022. By contrast, Elsevier published
+fewer than half of its open access articles through transformative
+agreements (n = 32,627; 54\%).
+
+The increasing trend towards transformative agreements can also be
 observed at the journal-level (Figure \ref{fig:publisher_figure}C).
 While no substantial differences between open access enabled through
-transformative agreements and other revenue sources could observed
-across Elsevier's portfolio, the distribution of open access across
-Springer Nature and Wiley hybrid journals indicates that the growth is
-not limited to a few journals, but extends across the portfolio. In
+transformative agreements and other revenue sources could be seen across
+Elsevier's portfolio, the distribution of open access across Springer
+Nature and Wiley hybrid journals indicates that the growth is not
+limited to a few journals but extends across the portfolio. In
 particular, Wiley's upper quantile, which represents the top 25\% of
@@
 transformative agreements, increased markedly from 13\% in 2020 to 31\%
-in 2022. At the same time, the median proportion grew from 7.5\% to
+in 2022. Simultaneously, the median proportion increased from 7.5\% to
 19\%. It is interesting to note that a small but increasing number of
-journals from these two publishers are providing open access to the
-majority of articles through transformative agreements. Wiley recorded
-68 and Springer Nature 102 hybrid journals with an open access share
-above 50\% that could be solely attributed to transformative agreements.
+journals from these two publishers provide open access to the majority
+of articles through transformative agreements. Wiley recorded 68 and
+Springer Nature 102 hybrid journals with an open access share above
+50\%, which could be attributed solely to transformative agreements.
 Upon inspection, these journals were mainly society or local language
-journals with a small yearly article volume.
+journals with small annual article volumes.
 
@@
 account for journals belonging to more than one category. Between 2018
-and 2022, most hybrid journals with at least one open articles could be
-attributed to the category Social Sciences, which also includes the Arts
-and Humanities. However, these journals published the fewest number of
-articles, whereas Physical Sciences journals recorded most articles,
-followed by the Health Sciences and the Life Sciences. In terms of open
-access, Physical Sciences journals accounted for more than one third of
-articles published in the five-years period, followed by the Health
-Science, the Social Sciences and the Life Sciences.
+and 2022, most hybrid journals with at least one open article could be
+attributed to the Social Sciences category, which also includes Arts and
+Humanities. However, these journals published the fewest number of
+articles, whereas Physical Sciences journals recorded the most articles,
+followed by Health Sciences and Life Sciences. In terms of open access,
+Physical Sciences journals accounted for more than one-third of the
+articles published in the five-year period, followed by Health Sciences,
+Social Sciences and Life Sciences.
 
@@
 open access by subject area between 2018-2022. In particular, Social
-Sciences and Humanties journals accounted for the strongest growth in
-the five-years period from 6.4\% (n = 8,361) to 23\% (n = 51,938),
-followed by the Life Science from 7.6\% (n = 15,003) to 18\% (n =
-39,494) , Health Science from 5.3\% (n = 18,279) to 16\% (n = 63,089)
-and Physical Sciences from 4.5\% (n = 22,364) to 12\% (n = 85,428).
-Growth in the Social Sciences can be largely attributed to
-transformative agreements. In 2022, two-third of open access articles
-(67\%, n = 34,759) were published by first authors affiliated with
+Sciences inbcluding Arts and Humanties journals accounted for the
+strongest growth in the five-years period from 6.4\% (n = 8,361) to 23\%
+(n = 51,938), followed by the Life Sciences from 7.6\% (n = 15,003) to
+18\% (n = 39,494) , Health Sciences from 5.3\% (n = 18,279) to 16\% (n =
+63,089) and Physical Sciences from 4.5\% (n = 22,364) to 12\% (n =
+85,428). Growth in Social Sciences category can be largely attributed to
+transformative agreements. In 2022, two-thirds of open access articles
+(67\%, n = 34,759) were published by the first authors affiliated with
 participating institutions (see \ref{fig:subject_panel}B).
@@
 
-Figure \ref{fig:subject_panel}C shows that this trend is consistent
-across hybrid journals belonging to the ASJC subject area Social
-Sciences. In 2022, 25\% of Social Sciences journals provided open access
-to at least every fourth article exclusively through transformative
+Figure \ref{fig:subject_panel}C shows that this trend was consistent
+across hybrid journals belonging to the ASJC Social Sciences category.
+In 2022, 25\% of Social Sciences journals provided open access to at
+least every fourth article exclusively through transformative
 agreements. However, hybrid open access through transformative
-agreements played a comparable lesser role in the Life Sciences and
-Health Sciences. In these two subject areas, only about half of the open
-access articles can be linked to these agreements, both overall and on
-median average across journals. In contrast, the majority of Physical
-Science Journals, shows an increase of open access through
-transformative agreements compared to other options to publish open
-access in hybrid journals.
+agreements played a comparably lesser role in Life Sciences and Health
+Sciences. In these two subject areas, only about half of the open access
+articles can be linked to these agreements, both overall and on a median
+average across journals. In contrast, the majority of Physical Science
+Journals show an increase of open access through transformative
+agreements compared to other options to publish open access in hybrid
+journals.
 
@@
 Figure \ref{fig:country_patch}A shows the development of hybrid open
-access publishing by countries, comparing the OECD area with the BRICS,
-an intergovernmental organisation, which comprised the countries Brazil,
-Russia, India, China and South Africa as of 2022. The residual category
-``Other'' includes the remaining countries. From 2018 to 2022, the
-proportion of open access in hybrid journals increased from 6.1\% in
-2018 to 26\% in 2022. On the other hand, BRICS recorded a low uptake,
-from 1.6\% in 2018 to 3.7\% in 2022.\\
+access publishing by country, comparing the OECD area with the BRICS, an
+intergovernmental organisation, which comprised Brazil, Russia, India,
+China, and South Africa as of 2022. The residual category ``Other''
+includes the remaining countries. Full counting was applied to account
+for multiple institutional affiliations (Hottenrott et al., 2021). From
+2018 to 2022, the proportion of open access in hybrid journals increased
+from 6.1\% in 2018 to 26\% in 2022. In contrast, BRICS recorded a low
+uptake, moderately growing from 1.6\% in 2018 to 3.7\% in 2022.
+
 Despite the rise of open access across OECD countries, the overall
@@
 more than doubling from 356,632 in 2018 to 786,903 in 2022. Upon closer
-examination, this trend can be observed across all big three publishers,
-although the shift towards BRICS is particularly evident in Elsevier's
-hybrid journal portfolio, in particular with regard to articles
-published in Physical Sciences journals. While OECD publication output
-in Elsevier's Physical Sciences journals declined from 112,822 articles
-in 2018 to 103,766 in 2022, BRICS output increased from 104,654 to
-171,713 in the same five-year period. Furthermore, OECD publication
+examination, this trend can be observed across all of the three largest
+publishers, although the shift towards BRICS is particularly evident in
+Elsevier's hybrid journal portfolio, particularly with regard to
+articles published in Physical Sciences journals. While OECD publication
+output in Elsevier's Physical Sciences journals declined from 112,822
+articles in 2018 to 103,766 in 2022, BRICS output increased from 104,654
+to 171,713 in the same five-year period. Furthermore, OECD publication
 output in Health Science Journals and Life Science journals stagnated
-across the investigated hybrid journal portfolios after a peak in 2020.
+across the investigated hybrid journal portfolios after peaking in 2020.
 
@@
 To illustrate the situation in 2022, Figure \ref{fig:country_patch}B
-compares total publication output with the number of open access
+compares the total publication output with the number of open access
 articles. With 391,530 articles, China was the most productive country,
 followed by the United States (268,965 articles) and India (87,428
-articles). In contrast, West and Nord European countries published a
-considerable high number of open access articles. Particularly, Germany,
-Great Britain, the Netherlands, Sweden, Switzerland and Spain recorded
-an above-average open access share as indicated by the linear trend
-line. As represented by the point size, as well as it can been seen in
-Figure \ref{fig:country_patch}C, transformative agreements contributed
-to these market positions. Interestingly, the United States had a
-notable open access market share of 15\%, although transformative
-agreements contributed to a lesser extent. Similarly, China's open
-access market share of 7.2\% in 2022 was comparable to that of the
-Netherlands, which was 7.1\%.
+articles). In contrast, Western and Northern European countries
+published a considerably high number of open access articles.
+Particularly, Germany, Great Britain, the Netherlands, Sweden,
+Switzerland, and Spain recorded an above-average open access share, as
+indicated by the linear trend line. As represented by the point size, as
+well as that shown in Figure \ref{fig:country_patch}C, transformative
+agreements contributed to these market positions. Interestingly, the
+United States had a notable open access market share of 15\%, although
+transformative agreements contributed to a lesser extent. Similarly,
+China's open access market share of 7.2\% in 2022 was comparable to that
+of the Netherlands, which was 7.1\%.
 
@@
 that were included in transformative agreements over the five-year
-period. Notably, The Netherlands (27\%), Sweden (24\%), Poland (17\%)
+period. Notably, the Netherlands (27\%), Sweden (24\%), Poland (17\%)
 and Great Britain (17\%) exhibited a relatively high level of uptake in
@@
 Wiley (July 2019) and Springer Nature (January 2020). Prior to this,
-only a few organisations had agreements in place, for the example the
-Max Planck Society with Springer Compact.
+only a few organisations had agreements in place, for example, the Max
+Planck Society with Springer Compact.
 
@@
 Since 2021, there has been a general trend towards hybrid open access
-among many high-income countries, primarily driven by transformative
-agreements. However, proliferation of transformative agreements differed
-across these countries. For instance, Germany successfully negotiated an
-agreement with Elsevier not until 2023. Additionally, publication limits
-or eligibility criteria for institutions and article types may explain
-why even countries with widespread agreement implementation like Sweden
-or the Nethterlands did not achieve 100\% hybrid open access.
-Interestingly, in Japan and the US other options than transformative
-agreements were the main driver for the increase in hybrid open access.
-Once again, the graph highlights countries with low hybrid open access,
-particularly non-OECD countries, where only a few or no agreements were
-in place.
+among many high-income countries, driven primarily by transformative
+agreements. However, the proliferation of transformative agreements
+differed across these countries. For instance, Germany successfully
+negotiated an agreement with Elsevier not until 2023. Additionally,
+publication limits or eligibility criteria for institutions and article
+types may explain why even countries with widespread agreement
+implementation such as Sweden or the Netherlands did not achieve 100\%
+hybrid open access. Interestingly, in Japan and the US, options other
+than transformative agreements were the main drivers for the increase in
+hybrid open access. Once again, the graph highlights countries with low
+hybrid open access, particularly non-OECD countries, where only a few or
+no agreements were in place.
 
@@
 
-The primary aim of this study was to investigate the adoption of open
-access in hybrid journals in transformative agreements, which were
+The primary aim of this study was to assess the uptake of open access in
+hybrid journals included in transformative agreements, which were
 introduced as a temporal means to support the transition of
@@
 on over 700 agreements and nine million journal articles to estimate the
-extent to which transformation agreement contribute to the transition of
-this journal business model. The results highlight a strong growth in
-open access between 2018 and 2022, particularly driven by an increasing
-number of institutions that had transformative agreements in place. Yet,
-the majority of research literature published in hybrid journals in this
-five-year period remained behind a publisher's paywall. Growth in the
+extent to which transformation agreements contribute to the transition
+of this journal business model. The results highlight a strong growth in
+open access between 2018 and 2022, driven by an increasing number of
+institutions that had transformative agreements in place. However, the
+majority of research literature published in hybrid journals in this
+five-year period remained behind the publisher's paywall. Growth in the
 adoption of open access in hybrid journals, in particular through
@@
 commercial publishers -- Elsevier, Springer Nature, and Wiley -- but
-varies substantially across journals, publishers, subjects, and country
-affiliations. Despite the limitations of the data used, the findings
+varies substantially across journals, publishers, disciplines, and
+country affiliations. Despite the limitations of the data, the findings
 indicate that the current level of implementation of transformative
@@
 
-A key finding of the analysis is that transformative agreements increase
-market concentration. Specifically, the three largest commercial
-publishers Elsevier, Springer Nature, and Wiley dominate, particularly
-in regard to open access provided through transformative agreements.
-Together, the three publishers accounted for three-fourths of open
-access articles through transformative agreements, while recording less
-than half of total publication volume published between 2018 and 2022.
-This observation aligns with previous research on shifts in the
-publishing market following the introduction of funding opportunities
-for hybrid open access (Butler et al., 2023; Jahn \& Tullney, 2016; Shu
-\& Larivière, 2023). Additionally, the results confirm previously
-observed variations by publisher, with Elsevier exhibiting a different
-development than Springer Nature and Wiley (Butler et al., 2023; Jahn et
-al., 2022). However, it must be noted that the focus of transformative
-agreements on publishers with large hybrid journal portfolios is
-intentional (Campbell et al., 2022). Because of transformative
-agreements with a few large publishers, national consortia were able to
-substantially increase their country's annual open access article volume
-(Bosman, Jonge, et al., 2021; Huang et al., 2020; Pinhasi et al., 2021;
-Taubert et al., 2023; Wenaas, 2022).
+A key finding of this analysis is that transformative agreements
+maintain market concentration. Specifically, the three largest
+commercial publishers Elsevier, Springer Nature, and Wiley dominate,
+particularly with regard to open access provided through transformative
+agreements. Together, the three publishers accounted for three-fourths
+of open access articles through transformative agreements, while
+recording less than half of the total publication volume published
+between 2018 and 2022. This observation aligns with previous research on
+shifts in the publishing market following the introduction of funding
+opportunities for hybrid open access (Butler et al., 2023; Jahn \&
+Tullney, 2016; Shu \& Larivière, 2023). Additionally, the results
+confirm previously observed variations by publisher, with Elsevier
+exhibiting a different development than Springer Nature and Wiley
+(Butler et al., 2023; Jahn et al., 2022). However, it must be noted that
+the focus of transformative agreements on publishers with large hybrid
+journal portfolios is intentional (Campbell et al., 2022). Because of
+transformative agreements with a few large publishers, national
+consortia were able to substantially increase their country's annual
+open access article volume (Bosman, Jonge, et al., 2021; Huang et al.,
+2020; Pinhasi et al., 2021; Taubert et al., 2023; Wenaas, 2022).
 
@@
 previous research findings (Butler et al., 2023; Jahn et al., 2022),
-high uptake rates were observed across hybrid journals in the Social
+high uptake rates were observed across hybrid journals in Social
 Sciences and Humanities. However, it is important to emphasise that
-these hybrid journals do not encompass the entire field. For example, a
-study by Khanna et al. (2022) found that approximately 60\% of journals
-utilizing the open-source publishing platform Open Journal Systems (OJS)
-fall within the Social Sciences and Humanities. In these fields, the
-majority of full open access journals are so-called ``Diamond OA
-journals'' that do not charge publication fees (Bosman, Frantsvåg, et
-al., 2021). Rather, this result can be more accurately attributed to the
-substantial proportion of authors from high-income countries who publish
-in these hybrid journals included in transformative agreements,
-particularly when journals are of local relevance, such as through the
-belonging to national societies or regional research scope.
+these hybrid journals do not encompass the entire field. For example,
+Khanna et al. (2022) found that approximately 60\% of journals utilizing
+the open-source publishing platform Open Journal Systems (OJS) fall
+within the Social Sciences and Humanities. In these fields, the majority
+of full open access journals are so-called ``Diamond OA journals'',
+which do not charge publication fees (Bosman, Frantsvåg, et al., 2021).
+Rather, this result can be more accurately attributed to the substantial
+proportion of authors from high-income countries who publish in these
+hybrid journals included in transformative agreements, particularly when
+journals are of local relevance, such as by belonging to national
+societies or regional research scope.
 
@@
 transformative agreements and demands discussion. In China, the country
-with the highest volume of articles in hybrid journals, limited research
-funding to pay for open access is available, with expenditures for open
-access publishing already surpassing subscription costs (Shu \&
+with the highest volume of articles in hybrid journals, only limited
+research funding to pay for open access is available, with expenditures
+for open access publishing surpassing subscription costs (Shu \&
 Larivière, 2023). Furthermore, the focus of Chinese authors on
@@
 As highlighted by Koley \& Lala (2023), India faces similar challenges
-in terms of resources to pay for open access. At the same time, access
-to research literature is a pressing, which is adressed by the ``Indian
-one nation, one subscription'' policy proposal. However, this policy
-focuses on centrally negotiated subscriptions and does not entail open
-access.
+in terms of the availablility of resources to pay for open access. At
+the same time, access to research literature is a pressing issue, which
+is addressed by the ``Indian one nation, one subscription'' policy
+proposal. However, this policy focuses on centrally negotiated
+subscriptions and does not entail open access.
 
 But open access uptake also differs among OECD countries. In the United
-States, for example, hybrid open access including transformative
-agreements plays a lesser role compared to some European countries.
-Between 2017 and 2021, hybrid open access contributed the least to make
-federally funded research articles openly available (Schares, 2023).
-While some university consortia like the California Digital Library have
-signed transformative agreements, others attempt to depart from big
-deals and unbundle large journal portfolios to address cost increases
-(Brainard, 2021; Schares, 2022). Despite the relatively low penetration
-of transformative agreements in China and the United States compared to
-Europe, 22\% of open access in hybrid journals in 2022 originates from
-these two countries, indicating the availability of funding sources for
-publication fees.
-
-This large-scale study provides first empricial evidence about the
+States, for example, hybrid open access, including transformative
+agreements, plays a lesser role than in some European countries. Between
+2017 and 2021, hybrid open access contributed the least to openly
+available federally funded research articles (Schares, 2023). While some
+university consortia, such as the California Digital Library, have
+signed transformative agreements, others have attempted to depart from
+big deals and unbundle large journal portfolios to address cost
+increases (Brainard, 2021; Schares, 2022). Despite the relatively low
+penetration of transformative agreements in China and the United States
+compared to Europe, 22\% of open access in hybrid journals in 2022
+originates from these two countries, indicating the availability of
+funding sources for publication fees.
+
+This large-scale study provides the first empiricial evidence of the
 influence of transformative agreements on the transition of hybrid
-journals to full open access. However, several limitations needs to be
-acknowledged. From a data perspective, estimations of open access
-through transformative agreements were established by linking first
-author affiliations with publicly available agreement data from
-cOAlition S, and not through invoicing data, which is usually not
-shared. Moreover, this study is unable to account for the various types
-of transformative agreements, particularly with regards to article types
-and capping. Furthermore, assessing the quality of the OpenAlex and
-Crossref data used, particularly in terms of affiliations and article
-types, by comparing it with established bibliometric databases such as
-Scopus or Web of Science was beyond the scope of this analysis. Overall,
-the methodology is designed in such a way that the results underestimate
+journals to full open access. However, that study has several
+limitations that need to be acknowledged. From a data perspective,
+estimations of open access through transformative agreements were
+established by linking first author affiliations with publicly available
+agreement data from cOAlition S, and not through invoicing data, which
+is usually not shared. Moreover, this study is unable to account for the
+various types of transformative agreements, particularly with regard to
+article types and caps that limit the number of open access articles
+covered. Furthermore, assessing the quality of the OpenAlex and Crossref
+data used, particularly in terms of affiliations and article types, by
+comparing them with established bibliometric databases such as Scopus
+and Web of Science was beyond the scope of this analysis. Overall, the
+methodology is designed in such a way that the results underestimate
 rather than overestimate the adoption of open access through
@@
 impacted by the COVID-19 pandemic, which led to an unprecedented number
-of publications as well as an reduction of international collaboration
+of publications and a reduction in international collaboration
 (Aviv-Reuven \& Rosenfeld, 2021), which could explain the observed
-contrasting developements in OECD and BRICS countries. Even before the
-pandemic, growth in publications in Europe were only due to
+contrasting developments in OECD and BRICS countries. Even before the
+pandemic, growth in publications in Europe was only due to
 internationally co-authored journal articles (Kwiek, 2021). Likewise,
-inflows from China to the US and European countries already declined
-before 2020 (Zhao et al., 2023). Furthermore, the study design did not
-consider emerging publication practices such as preprints (Fraser et
-al., 2021) and special issues (Hanson et al., 2023), which have grown
-rapidly since 2020. Lastly, it should be emphasized that the study did
-not address financial shifts between suscriptiuons spending and open
-acces spayemnts while analysing hybrid open access through
-transformative agreements due to a lack of data on expenditures.
-
-This present study allows multiple strands of further research. One is
-to complement this large-scale study with more specific evidence from
+inflows from China to the US and European countries already declined by
+2020 (Zhao et al., 2023). Furthermore, the study design did not consider
+emerging publication practices such as preprints (Fraser et al., 2021)
+and special issues (Hanson et al., 2023), which have grown rapidly since
+2020. Lastly, it should be emphasized that the study did not address
+financial shifts between suscriptiuons spending and open acces spayemnts
+while analysing hybrid open access through transformative agreements due
+to a lack of data on expenditures.
+
+This study allows for multiple strands of further research. One is to
+complement this large-scale study with more specific evidence from
 individual countries or subjects, particularly those with low hybrid
@@
 and collaboration could also be promising. Financial studies could build
-upon the presented study design and incorporate subscription and open
-access expenditures to evaluate the cost-effectiveness of transformative
+upon the study design and incorporate subscription and open access
+expenditures to evaluate the cost-effectiveness of transformative
 agreements.
 
-This study also has practival implications for research funders and
+This study has practical implications for research funding and
 libraries. Of primary concern should be the observed differences across
-countries, in particular the relationship between economic development
+countries, particularly the relationship between economic development
 and open access adoption. From a data perspective, the reporting of open
-access funding including transformative agreements is not harmonised,
+access funding, including transformative agreements, is not harmonised,
 but often crowd-sourced from various sources. To improve the assessment
@@
 
-In summary, this study provides empirical insights into the developement
+In summary, this study provides empirical insights into the development
 of hybrid open access following the introduction of transformative
-agreements. The results are important for both researchers and
+agreements. These results are important for both researchers and
 stakeholders engaged in negotiating and evaluating these agreements. The
 presented approach relies on open data, which enables follow-up studies
-and open access monitoring acitvties to further explore the role of
+and open access monitoring activties to further explore the role of
 transformative agreements in transitioning academic publishing to full
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Hottenrott_2021}{}}%
+Hottenrott, H., Rose, M. E., \& Lawson, C. (2021). The rise of multiple
+institutional affiliations in academia. \emph{Journal of the Association
+for Information Science and Technology}, \emph{72}(8), 1039--1058.
+\url{https://doi.org/10.1002/asi.24472}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Huang_2020}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Schmidt_2012}{}}%
+Schmidt, B., \& Shearer, K. (2012). Licensing revisited: Open access
+clauses in practice. \emph{LIBER Quarterly: The Journal of the
+Association of European Research Libraries}, \emph{22}(3), 176--189.
+\url{https://doi.org/10.18352/lq.8055}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Shu_2023}{}}%

### 2024-02-27 analysis/manuscript.Rmd
final glitches

@@
 a substantial transition of hybrid journals towards full open access could not be observed [@Momeni_2021; @Matthias_2019].
-The focus on large commercial publishers might also increase inequality [@Klebel_2023], because transformative agreements' focus on pay to publish mainly targets institutions from high-income countries, furthering a questionable journal prestige culture [@budapest].
+The focus on large commercial publishers might also increase inequality [@Ross_Hellauer_2022], because transformative agreements' focus on pay to publish mainly targets institutions from high-income countries, furthering a questionable journal prestige culture [@budapest].
 Besides, an editorial-board resignation raised concerns that transformative agreements might encourage publishers to maximize journal publication volume "without regard to quality" [@Goodin_2023].
@@
 Despite these controversies around transformative agreements as a means of transitioning journal publishing to full open access, there is limited evidence available on the adoption of open access in hybrid journals, and the extent to which this can be attributed to transformative agreements.
-Previous studies have focused on specific countries [@P_l_nen_2020; @Huang_2020; @Haucap_2021; @Wenaas_2022; @Taubert_2023] or publisher portfolios [@Pieper_2018; @Bakker_2024; @Fraser_2023; @Jahn_2021; @Momeni_2023], while large-scale studies relied on self-reported agreement data [@Moskovkin_2022], or used APC pricing lists [@Shu_2023].
+Previous studies have focused on specific countries [@P_l_nen_2020; @Huang_2020; @Haucap_2021; @Wenaas_2022; @Taubert_2023] or publisher portfolios [@Pieper_2018; @Bakker_2024; @Fraser_2023; @Jahn_2021; @Momeni_2023; @Schmal_2024], while large-scale studies relied on self-reported agreement data [@Moskovkin_2022], or used APC pricing lists [@Shu_2023].
 In particular, data availability is a limiting factor when studying the impact of transformative agreements [@Bakker_2024], because bibliometric databases, even though many allow the retrieval of open access articles in hybrid journals, do not directly attribute them to specific transformative agreements.
@@
 Initially, transformative agreement data retrieved from cOAlition S Journal Checker Tool[^checkertool]  provided information about journal portfolios and participating institutions. 
-After identifying hybrid journals by excluding fully open access journals, Crossref served as the primary data source for article-level metadata including Creative Commons (CC) license information to indicate open access availability on publisher websites.
+After identifying hybrid journals by excluding fully open access journals, Crossref [@Hendricks_2020] served as the primary data source for article-level metadata including Creative Commons (CC) license information to indicate open access availability on publisher websites.
 To determine open access articles published through transformative agreements, first author affiliations from OpenAlex [@priem2022openalex] were subsequently linked to eligible institutions according to the transformative agreement data.
@@
 Comparing Crossref license coverage with OpenAlex, which re-uses open access evidence from Unpaywall, a widely used open access discovery service that also parses journal websites for open content licenses [@Piwowar_2018], highlighted ongoing challenges in identifying hybrid open access [@Mart_n_Mart_n_2018; @Jahn_2021; @Zhang_2022; @Butler_2023].
-Here, `r hoaddata::cr_upw |> filter(cr_year %in% 2018:2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, cr_hybrid_total) |> pull(cr_hybrid_total) |> sum()` articles under CC license were retrieved using Crossref, while `r hoaddata::cr_upw |> filter(cr_year %in% 2018:2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, upw_hybrid_total) |> pull(upw_hybrid_total) |> sum()` articles were tagged as "hybrid" according to the OpenAlex November 2023 release, which was used throughout this study.
+For the purpose of this study, `r hoaddata::cr_upw |> filter(cr_year %in% 2018:2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, cr_hybrid_total) |> pull(cr_hybrid_total) |> sum()` articles under CC license were retrieved using Crossref, while `r hoaddata::cr_upw |> filter(cr_year %in% 2018:2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, upw_hybrid_total) |> pull(upw_hybrid_total) |> sum()` articles were tagged as "hybrid" according to the OpenAlex November 2023 release, which was used throughout this study.
 The largest differences concerned articles published between 2018 and 2020.
@@
 Throughout this mostly automated data gathering and analysis process, Tidyverse tools (Wickham et al., 2019) for the R programming language (R Core Team, 2020) were used. 
-The resulting data are openly available through an R data package, hoaddata, version 0.2.91 
+The resulting data are openly available through an R data package, hoaddata [@hoaddata]. 
 Following @Marwick_2018, hoaddata contains not only the datasets used in the data analysis. 
-It also includes code used to compile the data by connecting it to the cloud-based Google Big Query data warehouse, where scholarly big data from Crossref, OpenAlex and Unpaywall were imported.
+It also includes code used to compile the data by connecting it to the cloud-based Google Big Query data warehouse, where scholarly big data from Crossref, OpenAlex and Unpaywall were imported, using bigrquery [@bigrquery].
 To increase computational reproducibility, data aggregation through hoaddata was automatically carried out using GitHub Actions, a continuous integration service.
@@
 Authors who could make use of transformative agreements at the time of publication contributed `r n_ta_oa` open access articles.
+Overall, `r n_tas_with_oa` transformative agreements recorded at least one open access article from an eligible institution.
 
@@
 Over the five-years period from 2018 to 2022, open access increased from `r oa_by_year |> filter(cr_year == "2018") |> pull(prop)`% (n = `r oa_by_year |> filter(cr_year == "2018") |> pull(oa)`) to `r oa_by_year |> filter(cr_year == "2022") |> pull(prop)`% (n = `r oa_by_year |> filter(cr_year == "2022") |> pull(oa)`).
-Simultaneously, the total article volume of the investigated journals increased from `r oa_by_year |> filter(cr_year == "2018") |> pull(articles)` in 2018 to `r oa_by_year |> filter(cr_year == "2022") |> pull(articles)` in 2022.
+Simultaneously, the total article volume of the investigated journals rose from `r oa_by_year |> filter(cr_year == "2018") |> pull(articles)` in 2018 to `r oa_by_year |> filter(cr_year == "2022") |> pull(articles)` in 2022.
 
 Figure \ref{fig:results_overview}B highlights that the majority of hybrid open access was made available through transformative agreements in 2021 and 2022, contributing `r oa_ta_plot_df |> filter(cr_year == "2022", type == "ta_oa") |> pull(value) / oa_by_year |> filter(cr_year == "2022") |> pull(oa) * 100`% of the total open access article volume in 2022.
-However, there was also a notable increase in open access provision through other means, presumably publication fees being not invoiced through transformative agreements, which increased from `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> pull(value)`) in 2018 to  `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> pull(value)`).
+However, there was also a notable increase in open access provision through other means, presumably publication fees being not invoiced through transformative agreements, from `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> pull(value)`) in 2018 to  `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> pull(value)`).
 <!-- This suggests that publishers were able to gain equally from individual and institutional open access publishing options.-->
@@
   inner_join(hoaddata::jn_ind, by = "issn_l") |>
-  filter(issn_l != "0027-8424")
+  filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
+  filter(issn_l %in% hoaddata::cc_articles$issn_l)
 
 ta_active <- ta_oa_inst |>
-  filter(issn_l != "0027-8424") |>
+  filter(issn_l %in% jn_ind_jct$issn_l) |>
   filter(ta_active == TRUE, !is.na(cc)) |>
@@
 Springer Nature, for example, started offering open access agreements for its hybrid journal portfolio to selected consortia such as the Max Planck Society, the Swedish Bibsam Consortium and the Finnish FinELib consortium in 2015 under the name Springer Compact[^springercompact].
-Prior to this, Springer discontinued its hybrid OA pilot agreements with a small number of institutions, including the University of Göttingen [@Schmidt_2012].
-However, the Springer Compact agreements were not included in the data as they concluded prior to the start of the transformative agreement data collection in June 2021. 
+Prior to this, Springer had some pilot agreements with a small number of institutions, including the University of Göttingen [@Schmidt_2012].
+However, the Springer Compact agreements were not included in the data as they concluded before the start of the transformative agreement data collection in June 2021. 
 Nonetheless, the results demonstrate the importance of agreements for Springer Nature's hybrid open access business over the past five years (Figure 2B).
@@
 Upon inspection, these journals were mainly society or local language journals with small annual article volumes.
+
+## Journal subjects
 
@@
 
-## Journal subjects
-
 Table \ref{tab:subject_summary_table} presents a high-level overview of hybrid open access by AJCS subject area using fractionalised counting to account for journals belonging to more than one category.
-Between 2018 and 2022, most hybrid journals with at least one open article could be attributed to the Social Sciences category, which also includes Arts and Humanities.
+Between 2018 and 2022, most hybrid journals with at least one open access article could be attributed to the Social Sciences category, which also includes Arts and Humanities.
 However, these journals published the fewest number of articles, whereas Physical Sciences journals recorded the most articles, followed by Health Sciences and Life Sciences. 
@@
 
-Figure \ref{fig:subject_panel} presents the relative growth of hybrid open access by subject area between 2018-2022.
-In particular, Social Sciences inbcluding Arts and Humanties journals accounted for the strongest growth in the five-years period from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Social Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Social Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(value) |> sum()`), followed by the Life Sciences from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Life Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Life Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Life Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Life Sciences") |> pull(value) |> sum()`) , Health Sciences from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Health Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Health Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Health Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Health Sciences") |> pull(value) |> sum()`) and Physical Sciences from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Physical Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Physical Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Physical Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Physical Sciences") |> pull(value) |> sum()`).
+Figure \ref{fig:subject_panel} presents the relative growth of hybrid open access by subject area between 2018 and 2022.
+In particular, Social Sciences including Arts and Humanties journals accounted for the strongest growth in the five-years period from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Social Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Social Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(value) |> sum()`), followed by the Life Sciences from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Life Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Life Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Life Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Life Sciences") |> pull(value) |> sum()`) , Health Sciences from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Health Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Health Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Health Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Health Sciences") |> pull(value) |> sum()`) and Physical Sciences from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Physical Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Physical Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Physical Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Physical Sciences") |> pull(value) |> sum()`).
 <!-- In particular, journals in the subject areas recorded large high rate (see supplement)-->
@@
 In these two subject areas, only about half of the open access articles can be linked to these agreements, both overall and on a median average across journals.
-In contrast, the majority of Physical Science Journals show an increase of open access through transformative agreements compared to other options to publish open access in hybrid journals.
+In contrast, the majority of Physical Science journals show an increase of open access through transformative agreements compared to other options to publish open access in hybrid journals.
+
+## Comparing countries
 
@@
 country_all <- hoaddata::jn_aff |>
+  filter(issn_l %in% hoaddata::jn_ind$issn_l) |>
   filter(!cr_year %in% c(2017, 2023), issn_l != "0027-8424") |>
@@
 country_oa <- hoaddata::cc_articles |>
+  filter(issn_l %in% hoaddata::jn_ind$issn_l) |>
   filter(!cr_year %in% c(2017, 2023), issn_l != "0027-8424") |>
@@
 country_ta <-  ta_oa_inst |>
+  filter(issn_l %in% hoaddata::jn_ind$issn_l) |>
   filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
@@
     axis.title=element_text(size = 10),
-    strip.text = ggtext::element_markdown(size = 7),
+    strip.text = ggtext::element_markdown(size = 6),
     plot.subtitle = ggtext::element_markdown(hjust = 0.5),
@@
 ```
-
-## Comparing countries
 
@@
 country_scatter_22 <- country_scatter_df |>
-  filter(cr_year == "2022", !is.na(country_code), all > 10000) |>
-  ggplot(aes(all, oa)) +
-  geom_point(aes(fill = cat, size = ta_oa), pch = 21, alpha = 0.7) +
+  filter(cr_year == "2022", !is.na(country_code), all > 5000) |>
+  ggplot(aes(all, oa / all)) +
+  geom_point(aes(fill = cat), pch = 21, alpha = 0.7) +
   scale_fill_manual(NULL, values = c("OECD" = "#d01c8b", "BRICS" = "#4dac26", "Other" = "#f7f7f7")) +
-  geom_smooth(aes(all, oa), method=lm) +
+  geom_smooth(aes(all, oa / all), method = lm) +
   scale_size("Agreement", guide = FALSE) +
   ggrepel::geom_text_repel(
-    data = filter(country_scatter_df, cr_year == 2022, ta_oa > 5000 | country_code %in% c("CN", "IN", "BR", "RU", "IR")),
-    aes(x = all, y = oa, label = country_code),
+    data = filter(country_scatter_df, cr_year == 2022, ta_oa > 2000 | country_code %in% c("CN", "IN", "BR", "RU", "IR")),
+    aes(x = all, y = oa / all, label = country_code),
             min.segment.length = .15,
@@
     box.padding = 0.3) +
-  scale_y_log10(labels =  scales::number_format(big.mark = ",")) +
+  scale_y_continuous(labels =  scales::percent) +
   scale_x_log10(labels =  scales::number_format(big.mark = ",")) +
   theme_minimal() +
-  labs(x = "Total articles in hybrid journals in 2022", y = "OA articles")
+  labs(x = "Total articles in hybrid journals in 2022\n(log-scale)", y = "OA Uptake")
 
@@
 Between 2018 and 2022, high-income countries almost exclusively dominated hybrid open access publishing through transformative agreements. 
-During this period, first-authors affiliated with institutions from Organisation for Economic Co-operation and Development (OECD) member countries published `r oecd_stats_global |> filter(cat == "OECD") |> pull(oa)` open access articles in hybrid journals, representing `r oecd_stats_global |> filter(cat == "OECD") |> pull(oa) / n_oa * 100`% of the investigated open access articles.
+During this period, first authors affiliated with institutions from Organisation for Economic Co-operation and Development (OECD) member countries published `r oecd_stats_global |> filter(cat == "OECD") |> pull(oa)` open access articles in hybrid journals, representing `r oecd_stats_global |> filter(cat == "OECD") |> pull(oa) / n_oa * 100`% of the investigated open access articles.
 This disparity between OECD nations and other countries becomes even more evident when considering open access through transformative agreements, as  `r oecd_stats_global |> filter(cat == "OECD") |> pull(ta_oa)` of `r n_ta_oa`, or `r oecd_stats_global |> filter(cat == "OECD") |> pull(ta_oa) / n_ta_oa*100`% of open access articles were associated with such agreements.
@@
 The residual category "Other" includes the remaining countries. 
-Full counting was applied to account for multiple institutional affiliations [@Hottenrott_2021].
+Full counting was applied to account for multiple country affiliations [@Hottenrott_2021].
 From 2018 to 2022, the proportion of open access in hybrid journals increased from `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2018) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2018 to `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2022) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2022.
@@
 In contrast, Western and Northern European countries published a considerably high number of open access articles. 
-Particularly, Germany, Great Britain, the Netherlands, Sweden, Switzerland, and Spain recorded an above-average open access share, as indicated by the linear trend line.
-As represented by the point size, as well as that shown in Figure \ref{fig:country_patch}C, transformative agreements contributed to these market positions. 
+Particularly, Scandinavian countries, the Netherlands and Austria recorded above-average open access shares, as indicated by the linear trend line.
+As shown in Figure \ref{fig:country_patch}C, transformative agreements contributed to these market positions. 
 Interestingly, the United States had a notable open access market share of `r top_20_plot$data |> distinct(country_code, oa, all) |> mutate(prop = oa / sum(oa)) |> filter(country_code == "US") |> pull(prop) * 100`%, although transformative agreements contributed to a lesser extent.
@@
 However, the proliferation of transformative agreements differed across these countries.
-For instance, Germany successfully negotiated an agreement with Elsevier not until 2023. 
+# For instance, Germany successfully negotiated an agreement with Elsevier not until 2023. 
 Additionally, publication limits or eligibility criteria for institutions and article types may explain why even countries with widespread agreement implementation such as Sweden or the Netherlands did not achieve 100% hybrid open access.
-Interestingly, in Japan and the US, options other than transformative agreements were the main drivers for the increase in hybrid open access.
+Interestingly, in Japan and the United States, options other than transformative agreements were the main drivers for the increase in hybrid open access.
 Once again, the graph highlights countries with low hybrid open access, particularly non-OECD countries, where only a few or no agreements were in place.
@@
 The results highlight a strong growth in open access between 2018 and 2022, driven by an increasing number of institutions that had transformative agreements in place. 
-However, the majority of research literature published in hybrid journals in this five-year period remained behind the publisher's paywall.
+However, the majority of research literature published in hybrid journals in this five-year period remained behind publisher paywalls.
 Growth in the adoption of open access in hybrid journals, in particular through transformative agreements, can be largely attributed to three large commercial publishers -- Elsevier, Springer Nature, and Wiley -- but varies substantially across journals, publishers, disciplines, and country affiliations.
@@
 Moreover, this study presents varying levels of open access uptake through transformative agreements across journals, which can be attributed to the alignment of authors' affiliations and the availability of such agreements at their institutions.
-In line with previous research findings [@Butler_2023; @Jahn_2021], high uptake rates were observed across hybrid journals in  Social Sciences and Humanities.
+In line with previous research findings [@Butler_2023; @Jahn_2021; @Wenaas_2022], high uptake rates were observed across hybrid journals in  Social Sciences and Humanities.
 However, it is important to emphasise that these hybrid journals do not encompass the entire field.
@@
 
-This large-scale study provides the first empiricial evidence of the influence of transformative agreements on the transition of hybrid journals to full open access. 
-However, that study has several limitations that need to be acknowledged.
+This large-scale study provides the first empirical evidence of the influence of transformative agreements on the transition of hybrid journals to full open access. 
+However, several limitations needs to be acknowledged.
 From a data perspective, estimations of open access through transformative agreements were established by linking first author affiliations with publicly available agreement data from cOAlition S, and not through invoicing data, which is usually not shared.
@@
 Furthermore, the study design did not consider emerging publication practices such as preprints [@Fraser_2021] and special issues [@hanson2023strain], which have grown rapidly since 2020. 
-Lastly, it should be emphasized that the study did not address financial shifts between suscriptiuons spending and open acces spayemnts while analysing hybrid open access through transformative agreements due to a lack of data on expenditures.
+Lastly, it should be emphasised that the study did not address financial shifts between subscriptions spending and open access payments while analysing hybrid open access through transformative agreements due to a lack of data on expenditures.
 
@@
 Incorporating full open access and subscription-based journals, as well as considering global trends in scholarly migration and collaboration could also be promising.
-Financial studies could build upon the study design and incorporate subscription and open access expenditures to evaluate the cost-effectiveness of transformative agreements.
+Financial studies could build on the study design and include subscription and open access expenditure to assess the cost-effectiveness of transformative agreements, in particular whether transformative agreements can integrate the substantial amounts of individual payments for publication fees [@Wenaas_2022; @Butler_2023], as well as potential changes in authors' behaviour following the introduction of these agreements [@Schmal_2024].
 
@@
 These results are important for both researchers and stakeholders engaged in negotiating and evaluating these agreements. 
-The presented approach relies on open data, which enables follow-up studies and open access monitoring activties to further explore the role of transformative agreements in transitioning academic publishing to full and equitable open access.
+The presented approach relies on open data, which enables follow-up studies and open access monitoring activities to further explore the role of transformative agreements in transitioning academic publishing to full open access.
 

### 2024-02-27 analysis/manuscript.tex
final glitches

@@
 al., 2021). The focus on large commercial publishers might also increase
-inequality (Klebel \& Ross-Hellauer, 2023), because transformative
+inequality (Ross-Hellauer et al., 2022), because transformative
 agreements' focus on pay to publish mainly targets institutions from
@@
 (Bakker et al., 2024; Fraser et al., 2023; Jahn et al., 2022; Momeni et
-al., 2023; Pieper \& Broschinski, 2018), while large-scale studies
-relied on self-reported agreement data (Moskovkin et al., 2022), or used
-APC pricing lists (Shu \& Larivière, 2023). In particular, data
+al., 2023; Pieper \& Broschinski, 2018; Schmal, 2024), while large-scale
+studies relied on self-reported agreement data (Moskovkin et al., 2022),
+or used APC pricing lists (Shu \& Larivière, 2023). In particular, data
 availability is a limiting factor when studying the impact of
@@
 institutions. After identifying hybrid journals by excluding fully open
-access journals, Crossref served as the primary data source for
-article-level metadata including Creative Commons (CC) license
-information to indicate open access availability on publisher websites.
-To determine open access articles published through transformative
-agreements, first author affiliations from OpenAlex (Priem et al., 2022)
-were subsequently linked to eligible institutions according to the
-transformative agreement data. In the following, the steps are described
-in detail.
+access journals, Crossref (Hendricks et al., 2020) served as the primary
+data source for article-level metadata including Creative Commons (CC)
+license information to indicate open access availability on publisher
+websites. To determine open access articles published through
+transformative agreements, first author affiliations from OpenAlex
+(Priem et al., 2022) were subsequently linked to eligible institutions
+according to the transformative agreement data. In the following, the
+steps are described in detail.
 
@@
 hybrid open access (Butler et al., 2023; Jahn et al., 2022;
-Martín-Martín et al., 2018; Zhang et al., 2022). Here, 742,369 articles
-under CC license were retrieved using Crossref, while 950,260 articles
-were tagged as ``hybrid'' according to the OpenAlex November 2023
-release, which was used throughout this study. The largest differences
-concerned articles published between 2018 and 2020. In 2022, however,
-Crossref and OpenAlex open access numbers differ only slightly (249,511
-records using Crossref vs.~255,344 in OpenAlex). Notable differences
-could be observed among some publishers that presumably did not provide
-CC license information to Crossref, including AIP Publishing, the
-American Physiological Society, Emerald, and the Royal Society. Crossref
-license metadata was more complete with regard to the articles published
-by Wiley and the American Chemical Society. Finally, inconsistent open
-access status information in previous OpenAlex versions was observed
-(Jahn et al., 2023). After reporting this to OpenAlex, fixing this issue
-was still ongoing according to the release notes, which might also
-explain part of the discrepancy.
+Martín-Martín et al., 2018; Zhang et al., 2022). For the purpose of this
+study, 742,369 articles under CC license were retrieved using Crossref,
+while 950,260 articles were tagged as ``hybrid'' according to the
+OpenAlex November 2023 release, which was used throughout this study.
+The largest differences concerned articles published between 2018 and
+2020. In 2022, however, Crossref and OpenAlex open access numbers differ
+only slightly (249,511 records using Crossref vs.~255,344 in OpenAlex).
+Notable differences could be observed among some publishers that
+presumably did not provide CC license information to Crossref, including
+AIP Publishing, the American Physiological Society, Emerald, and the
+Royal Society. Crossref license metadata was more complete with regard
+to the articles published by Wiley and the American Chemical Society.
+Finally, inconsistent open access status information in previous
+OpenAlex versions was observed (Jahn et al., 2023). After reporting this
+to OpenAlex, fixing this issue was still ongoing according to the
+release notes, which might also explain part of the discrepancy.
 
@@
 Core Team, 2020) were used. The resulting data are openly available
-through an R data package, hoaddata, version 0.2.91 Following Marwick et
+through an R data package, hoaddata (Jahn, 2023). Following Marwick et
 al. (2018), hoaddata contains not only the datasets used in the data
@@
 it to the cloud-based Google Big Query data warehouse, where scholarly
-big data from Crossref, OpenAlex and Unpaywall were imported. To
-increase computational reproducibility, data aggregation through
-hoaddata was automatically carried out using GitHub Actions, a
-continuous integration service.
+big data from Crossref, OpenAlex and Unpaywall were imported, using
+bigrquery (Wickham \& Bryan, 2020). To increase computational
+reproducibility, data aggregation through hoaddata was automatically
+carried out using GitHub Actions, a continuous integration service.
 
@@
 could make use of transformative agreements at the time of publication
-contributed 328,957 open access articles.
+contributed 328,957 open access articles. Overall, 394 transformative
+agreements recorded at least one open access article from an eligible
+institution.
 
@@
 increased from 4.3\% (n = 65,486) to 15\% (n = 249,511). Simultaneously,
-the total article volume of the investigated journals increased from
+the total article volume of the investigated journals rose from
 1,528,051 in 2018 to 1,676,928 in 2022.
@@
 access provision through other means, presumably publication fees being
-not invoiced through transformative agreements, which increased from
-4.1\% (n = 62,625) in 2018 to 6.3\% (n = 105,896).
+not invoiced through transformative agreements, from 4.1\% (n = 62,625)
+in 2018 to 6.3\% (n = 105,896).
 
@@
 FinELib consortium in 2015 under the name Springer Compact\footnote{\url{https://web.archive.org/web/20180414062853id_/http://www.liber2015.org.uk/wp-content/uploads/2015/03/Springer-Compact.pdf}}.
-Prior to this, Springer discontinued its hybrid OA pilot agreements with
-a small number of institutions, including the University of Göttingen
-(Schmidt \& Shearer, 2012). However, the Springer Compact agreements
-were not included in the data as they concluded prior to the start of
-the transformative agreement data collection in June 2021. Nonetheless,
-the results demonstrate the importance of agreements for Springer
-Nature's hybrid open access business over the past five years (Figure
-2B). In 2022, 66\% (n = 34,725) of open access in Springer Nature hybrid
-journals was enabled through transformative agreements. In the same
-year, 70\% (n = 37,316) of Wiley's open access articles could be linked
-to transformative agreements in 2022. By contrast, Elsevier published
-fewer than half of its open access articles through transformative
-agreements (n = 32,627; 54\%).
+Prior to this, Springer had some pilot agreements with a small number of
+institutions, including the University of Göttingen (Schmidt \& Shearer,
+2012). However, the Springer Compact agreements were not included in the
+data as they concluded before the start of the transformative agreement
+data collection in June 2021. Nonetheless, the results demonstrate the
+importance of agreements for Springer Nature's hybrid open access
+business over the past five years (Figure 2B). In 2022, 66\% (n =
+34,725) of open access in Springer Nature hybrid journals was enabled
+through transformative agreements. In the same year, 70\% (n = 37,316)
+of Wiley's open access articles could be linked to transformative
+agreements in 2022. By contrast, Elsevier published fewer than half of
+its open access articles through transformative agreements (n = 32,627;
+54\%).
 
@@
 account for journals belonging to more than one category. Between 2018
-and 2022, most hybrid journals with at least one open article could be
-attributed to the Social Sciences category, which also includes Arts and
-Humanities. However, these journals published the fewest number of
-articles, whereas Physical Sciences journals recorded the most articles,
-followed by Health Sciences and Life Sciences. In terms of open access,
-Physical Sciences journals accounted for more than one-third of the
-articles published in the five-year period, followed by Health Sciences,
-Social Sciences and Life Sciences.
+and 2022, most hybrid journals with at least one open access article
+could be attributed to the Social Sciences category, which also includes
+Arts and Humanities. However, these journals published the fewest number
+of articles, whereas Physical Sciences journals recorded the most
+articles, followed by Health Sciences and Life Sciences. In terms of
+open access, Physical Sciences journals accounted for more than
+one-third of the articles published in the five-year period, followed by
+Health Sciences, Social Sciences and Life Sciences.
 
@@
 \midrule
-Health Sciences & 2,376 & 22.5 & 2,709,906 & 27.8 & 286,592 & 27.3 & 117,746 & 25\\
-Life Sciences & 1,403 & 13.3 & 1,477,808 & 15.1 & 191,880 & 18.3 & 71,593 & 15.2\\
-Physical Sciences & 2,732 & 25.9 & 4,291,833 & 44 & 366,794 & 35 & 167,686 & 35.6\\
-Social Sciences & 4,050 & 38.3 & 1,280,460 & 13.1 & 203,461 & 19.4 & 114,190 & 24.2\\
+Health Sciences & 2,342 & 22.5 & 1,998,045 & 28 & 199,265 & 27.7 & 81,913 & 25.7\\
+Life Sciences & 1,399 & 13.4 & 1,080,346 & 15.1 & 133,526 & 18.6 & 48,570 & 15.2\\
+Physical Sciences & 2,693 & 25.9 & 3,111,711 & 43.6 & 247,515 & 34.4 & 110,933 & 34.8\\
+Social Sciences & 3,967 & 38.1 & 953,084 & 13.3 & 138,388 & 19.3 & 77,496 & 24.3\\
 \bottomrule
@@
 Figure \ref{fig:subject_panel} presents the relative growth of hybrid
-open access by subject area between 2018-2022. In particular, Social
-Sciences inbcluding Arts and Humanties journals accounted for the
+open access by subject area between 2018 and 2022. In particular, Social
+Sciences including Arts and Humanties journals accounted for the
 strongest growth in the five-years period from 6.4\% (n = 8,361) to 23\%
@@
 average across journals. In contrast, the majority of Physical Science
-Journals show an increase of open access through transformative
+journals show an increase of open access through transformative
 agreements compared to other options to publish open access in hybrid
@@
 dominated hybrid open access publishing through transformative
-agreements. During this period, first-authors affiliated with
+agreements. During this period, first authors affiliated with
 institutions from Organisation for Economic Co-operation and Development
@@
 includes the remaining countries. Full counting was applied to account
-for multiple institutional affiliations (Hottenrott et al., 2021). From
-2018 to 2022, the proportion of open access in hybrid journals increased
-from 6.1\% in 2018 to 26\% in 2022. In contrast, BRICS recorded a low
-uptake, moderately growing from 1.6\% in 2018 to 3.7\% in 2022.
+for multiple country affiliations (Hottenrott et al., 2021). From 2018
+to 2022, the proportion of open access in hybrid journals increased from
+6.1\% in 2018 to 26\% in 2022. In contrast, BRICS recorded a low uptake,
+moderately growing from 1.6\% in 2018 to 3.7\% in 2022.
 
@@
 published a considerably high number of open access articles.
-Particularly, Germany, Great Britain, the Netherlands, Sweden,
-Switzerland, and Spain recorded an above-average open access share, as
-indicated by the linear trend line. As represented by the point size, as
-well as that shown in Figure \ref{fig:country_patch}C, transformative
+Particularly, Scandinavian countries, the Netherlands and Austria
+recorded above-average open access shares, as indicated by the linear
+trend line. As shown in Figure \ref{fig:country_patch}C, transformative
 agreements contributed to these market positions. Interestingly, the
@@
 agreements. However, the proliferation of transformative agreements
-differed across these countries. For instance, Germany successfully
+differed across these countries. \# For instance, Germany successfully
 negotiated an agreement with Elsevier not until 2023. Additionally,
@@
 implementation such as Sweden or the Netherlands did not achieve 100\%
-hybrid open access. Interestingly, in Japan and the US, options other
-than transformative agreements were the main drivers for the increase in
-hybrid open access. Once again, the graph highlights countries with low
-hybrid open access, particularly non-OECD countries, where only a few or
-no agreements were in place.
+hybrid open access. Interestingly, in Japan and the United States,
+options other than transformative agreements were the main drivers for
+the increase in hybrid open access. Once again, the graph highlights
+countries with low hybrid open access, particularly non-OECD countries,
+where only a few or no agreements were in place.
 
@@
 majority of research literature published in hybrid journals in this
-five-year period remained behind the publisher's paywall. Growth in the
+five-year period remained behind publisher paywalls. Growth in the
 adoption of open access in hybrid journals, in particular through
@@
 availability of such agreements at their institutions. In line with
-previous research findings (Butler et al., 2023; Jahn et al., 2022),
-high uptake rates were observed across hybrid journals in Social
-Sciences and Humanities. However, it is important to emphasise that
-these hybrid journals do not encompass the entire field. For example,
-Khanna et al. (2022) found that approximately 60\% of journals utilizing
-the open-source publishing platform Open Journal Systems (OJS) fall
-within the Social Sciences and Humanities. In these fields, the majority
-of full open access journals are so-called ``Diamond OA journals'',
-which do not charge publication fees (Bosman, Frantsvåg, et al., 2021).
-Rather, this result can be more accurately attributed to the substantial
-proportion of authors from high-income countries who publish in these
-hybrid journals included in transformative agreements, particularly when
-journals are of local relevance, such as by belonging to national
-societies or regional research scope.
+previous research findings (Butler et al., 2023; Jahn et al., 2022;
+Wenaas, 2022), high uptake rates were observed across hybrid journals in
+Social Sciences and Humanities. However, it is important to emphasise
+that these hybrid journals do not encompass the entire field. For
+example, Khanna et al. (2022) found that approximately 60\% of journals
+utilizing the open-source publishing platform Open Journal Systems (OJS)
+fall within the Social Sciences and Humanities. In these fields, the
+majority of full open access journals are so-called ``Diamond OA
+journals'', which do not charge publication fees (Bosman, Frantsvåg, et
+al., 2021). Rather, this result can be more accurately attributed to the
+substantial proportion of authors from high-income countries who publish
+in these hybrid journals included in transformative agreements,
+particularly when journals are of local relevance, such as by belonging
+to national societies or regional research scope.
 
@@
 
-This large-scale study provides the first empiricial evidence of the
+This large-scale study provides the first empirical evidence of the
 influence of transformative agreements on the transition of hybrid
-journals to full open access. However, that study has several
-limitations that need to be acknowledged. From a data perspective,
-estimations of open access through transformative agreements were
-established by linking first author affiliations with publicly available
-agreement data from cOAlition S, and not through invoicing data, which
-is usually not shared. Moreover, this study is unable to account for the
-various types of transformative agreements, particularly with regard to
-article types and caps that limit the number of open access articles
-covered. Furthermore, assessing the quality of the OpenAlex and Crossref
-data used, particularly in terms of affiliations and article types, by
+journals to full open access. However, several limitations needs to be
+acknowledged. From a data perspective, estimations of open access
+through transformative agreements were established by linking first
+author affiliations with publicly available agreement data from
+cOAlition S, and not through invoicing data, which is usually not
+shared. Moreover, this study is unable to account for the various types
+of transformative agreements, particularly with regard to article types
+and caps that limit the number of open access articles covered.
+Furthermore, assessing the quality of the OpenAlex and Crossref data
+used, particularly in terms of affiliations and article types, by
 comparing them with established bibliometric databases such as Scopus
@@
 and special issues (Hanson et al., 2023), which have grown rapidly since
-2020. Lastly, it should be emphasized that the study did not address
-financial shifts between suscriptiuons spending and open acces spayemnts
+2020. Lastly, it should be emphasised that the study did not address
+financial shifts between subscriptions spending and open access payments
 while analysing hybrid open access through transformative agreements due
@@
 and collaboration could also be promising. Financial studies could build
-upon the study design and incorporate subscription and open access
-expenditures to evaluate the cost-effectiveness of transformative
-agreements.
+on the study design and include subscription and open access expenditure
+to assess the cost-effectiveness of transformative agreements, in
+particular whether transformative agreements can integrate the
+substantial amounts of individual payments for publication fees (Butler
+et al., 2023; Wenaas, 2022), as well as potential changes in authors'
+behaviour following the introduction of these agreements (Schmal, 2024).
 
@@
 presented approach relies on open data, which enables follow-up studies
-and open access monitoring activties to further explore the role of
+and open access monitoring activities to further explore the role of
 transformative agreements in transitioning academic publishing to full
-and equitable open access.
+open access.
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Brainard_2023}{}}%
-Brainard, J. (2023). {`Transformative'} journals get booted for
-switching to open access too slowly. In \emph{Science}.
+Brainard, J. (2023). {``Transformative''} journals get booted for
+switching to open access too slowly. \emph{Science}.
 \url{https://doi.org/10.1126/science.adj3282}
@@
 Campbell, C., Dér, Á., Geschuhn, K., \& Valente, A. (2022). How are
-transformative agreements transforming libraries? In I. F. of Library
-Associations \& I. (IFLA) (Eds.), \emph{87th IFLA world library and
-information congress (WLIC) / 2022 in dublin, ireland}. IFLA.
+transformative agreements transforming libraries? \emph{87th IFLA World
+Library and Information Congress (WLIC) / 2022 in Dublin, Ireland}.
 \url{https://repository.ifla.org/handle/123456789/1973}
@@
 Geschuhn, K., \& Stone, G. (2017). It's the workflows, stupid! What is
-required to make {`offsetting'} work for the open access transition.
+required to make {``offsetting''} work for the open access transition.
 \emph{Insights the {UKSG} Journal}, \emph{30}(3), 103--114.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Hendricks_2020}{}}%
+Hendricks, G., Tkaczyk, D., Lin, J., \& Feeney, P. (2020). Crossref: The
+sustainable source of community-owned scholarly metadata.
+\emph{Quantitative Science Studies}, \emph{1}(1), 414--427.
+\url{https://doi.org/10.1162/qss_a_00022}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Hinchliffe_2019}{}}%
 Hinchliffe, L. J. (2019). \emph{Transformative agreements: A primer}.
-\url{https://web.archive.org/web/20210128170342/https://scholarlykitchen.sspnet.org/2019/04/23/transformative-agreements/};
 The Scholarly Kitchen.
+\url{https://web.archive.org/web/20210128170342/https://scholarlykitchen.sspnet.org/2019/04/23/transformative-agreements/}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-hoaddata}{}}%
+Jahn, N. (2023). \emph{{hoaddata}: Data about hybrid open access journal
+publishing}.
+\url{https://github.com/subugoe/hoaddata/releases/tag/v0.2.91}
+
 \leavevmode\vadjust pre{\hypertarget{ref-jahn2023}{}}%
 Jahn, N., Haupka, N., \& Hobert, A. (2023). \emph{Analysing and
-reclassifying open access information in OpenAlex}. Blog post.
+reclassifying open access information in OpenAlex}.
 \url{https://subugoe.github.io/scholcomm_analytics/posts/oalex_oa_status/}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Klebel_2023}{}}%
-Klebel, T., \& Ross-Hellauer, T. (2023). The APC-barrier and its effect
-on stratification in open access publishing. \emph{Quantitative Science
-Studies}, \emph{4}(1), 22--43. \url{https://doi.org/10.1162/qss_a_00245}
-
 \leavevmode\vadjust pre{\hypertarget{ref-Koley_2023}{}}%
@@
 Kwiek, M. (2021). What large-scale publication and citation data tell us
-about international research collaboration in europe: Changing national
-patterns in global contexts. \emph{Studies in Higher Education},
-\emph{46}(12), 2629--2649.
+about international research collaboration in {Europe}: Changing
+national patterns in global contexts. \emph{Studies in Higher
+Education}, \emph{46}(12), 2629--2649.
 \url{https://doi.org/10.1080/03075079.2020.1749254}
@@
 of open access publishing agreements at the {University of Vienna} in
-light of the plan s requirements: A review of current status, challenges
-and perspectives. \emph{Insights the UKSG Journal}, \emph{33}.
-\url{https://doi.org/10.1629/uksg.523}
+light of the {Plan S} requirements: A review of current status,
+challenges and perspectives. \emph{Insights the UKSG Journal},
+\emph{33}. \url{https://doi.org/10.1629/uksg.523}
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Goodin_2023}{}}%
-Rasmussen, K. B. (2023). Interview with {Robert {`Bob'} E. Goodin}.
+Rasmussen, K. B. (2023). Interview with {Robert {``Bob''} E. Goodin}.
 \emph{Tidskrift För Politisk Filosofi}.
@@
 e9410. \url{https://doi.org/10.7717/peerj.9410}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Ross_Hellauer_2022}{}}%
+Ross-Hellauer, T., Reichmann, S., Cole, N. L., Fessl, A., Klebel, T., \&
+Pontika, N. (2022). Dynamics of cumulative advantage and threats to
+equity in open science: A scoping review. \emph{Royal Society Open
+Science}, \emph{9}(1). \url{https://doi.org/10.1098/rsos.211032}
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Schiltz_2018}{}}%
-Schiltz, M. (2018). Science without publication paywalls: cOAlition s
+Schiltz, M. (2018). Science without publication paywalls: {cOAlition S}
 for the realisation of full and immediate open access. \emph{PLOS
@@
 \url{https://doi.org/10.17617/1.3}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Schmal_2024}{}}%
+Schmal, W. B. (2024). How transformative are transformative agreements?
+Evidence from germany across disciplines. \emph{Scientometrics}.
+\url{https://doi.org/10.1007/s11192-024-04955-y}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-bigrquery}{}}%
+Wickham, H., \& Bryan, J. (2020). \emph{{bigrquery}: An interface to
+{Google's BigQuery API}}.
+\url{https://CRAN.R-project.org/package=bigrquery}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Zhang_2022}{}}%

### 2024-02-27 analysis/manuscript.Rmd
one more glitch
Drafting abstract and declarations

@@
 Blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established.
-Country names are represented as ISO two-letter country codes"}
+Country names are represented as ISO two-letter country codes.
+Facet subheadings show the total number of articles and corresponding market share."}
 country_plot
@@
 
+# Competing interests
+
+The author declares no competing interests.
+
+# Funding information
+
+This work was supported by the Deutsche Forschungsgemeinschaft (Grant number 416115939).
+
+# Data and code availability
+
+Source code analysis including data used is available on GitHub: <https://github.com/njahn82/hoa_ta_effects>.
+
 # References

### 2024-02-27 analysis/manuscript.tex
one more glitch
Drafting abstract and declarations

@@
 
-%\authornote{Correspondence concerning this article should be addressed to Najko Jahn}
-
-\abstract{}
+\authornote{ORCID: https://orcid.org/0000-0001-5105-1463}
+
+\abstract{Ongoing controversies surrounding transformative agreements as a means of transitioning journal publishing to full open access highlight the need for large-scale studies assessing the adoption of open access in hybrid journals, including the extent to which transformative agreements contributed to it.
+Combining publicly available data from various sources, including cOAlition S Journal Checker, Crossref, and OpenAlex, this study presents a novel approach that leverages metadata from over 700 agreements and nine million journal articles.
+Estimates suggest a strong growth in open access between 2018 and 2022 from 4.3\% to 15\%.
+In 2022, 58\% of open access articles could be attributed to transformative agreements.
+Growth can be largely attributed to three large commercial publishers: Elsevier, Springer Nature, and Wiley, but it varies substantially across journals, publishers, disciplines, and country affiliations.
+These estimates indicate that the current level of implementation of transformative agreements is insufficient to bring about a large-scale transition to full open access.
+In particular, different developments in the OECD and BRICS areas have the potential to hinder the transition of academic journal publishing.
+These findings are important for researchers and stakeholders engaged in negotiating and evaluating these agreements.}
+
+\keywords{hybrid open access, transformative agreements, scholarly publishing, big deals, bibliometrics}
 
@@
 Blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established.
-Country names are represented as ISO two-letter country codes}\label{fig:country_top_20_plot}
+Country names are represented as ISO two-letter country codes.
+Facet subheadings show the total number of articles and corresponding market share.}\label{fig:country_top_20_plot}
 \end{figure}
@@
 
+\hypertarget{competing-interests}{%
+\section{Competing interests}\label{competing-interests}}
+
+The author declares no competing interests.
+
+\hypertarget{funding-information}{%
+\section{Funding information}\label{funding-information}}
+
+This work was supported by the Deutsche Forschungsgemeinschaft (Grant
+number 416115939).
+
+\hypertarget{data-and-code-availability}{%
+\section{Data and code availability}\label{data-and-code-availability}}
+
+Source code analysis including data used is available on GitHub:
+\url{https://github.com/njahn82/hoa_ta_effects}.
+
 \hypertarget{references}{%

### 2024-02-27 data/000_data_gathering.R
update data metadata

@@ -2,23 +2,23 @@
 #' 
 #' ## Data
 #' 
-#' *{hoaddata} version v0.2.82* provides journal-level and article-level data used.
+#' *{hoaddata} version v0.2.91* provides journal-level and article-level data used.
 #' 
-#' <https://github.com/subugoe/hoaddata/releases/tag/v0.2.82>
+#' <https://github.com/subugoe/hoaddata/releases/tag/v0.2.91>
 #' 
 #' This R package contains: 
 #' 
 #' *Journal-level data*, retrieved from the cOAlition S public transformative
-#' agreement data on 8 May 2023
+#' agreement data on 11 December 2023
 #' 
-#' <https://github.com/njahn82/jct_data/tree/4a017bba5ea72e86840d701420afc9d922e00266/data>
+#' <https://github.com/njahn82/jct_data/tree/a00138fa78bca769cc103caed715b2fab1375b6e/data>
 #' 
 #' The code repository also contains code used to obtain the data.
 #' 
 #' {hoaddata} also represents *article-level data* retrieved from
 #' 
-#'  - Crossref release 2023/10
-#'  - OpenAlex release 2023-10-18
+#'  - Crossref release 2023/11
+#'  - OpenAlex release 2023-11-21
 #'  
 #'  as well as the code used to compile the data from SUB Göttingen Scholarly
 #'  data warehouse.
@@ -26,7 +26,7 @@
 #'  The main dataset, providing article-level data about publications linked to 
 #'  transformative agreements and institutions, is available as {hoaddata} 
 #'  release asset: 
-#'  <https://github.com/subugoe/hoaddata/releases/download/v0.2.82/ta_oa_inst.csv.gz>
+#'  <https://github.com/subugoe/hoaddata/releases/download/v0.2.91/ta_oa_inst.csv.gz>
 #'  
 #'  A version is provided in `data/` folder
 download.file("https://github.com/subugoe/hoaddata/releases/download/v0.2.91/ta_oa_inst.csv.gz",

### 2024-02-27 install.R
Try binder config with renv

@@ -1,0 +1,1 @@
+install.packages("remotes")

### 2024-02-27 install.R
add latex dep

@@ -1,1 +1,2 @@
-install.packages("remotes")
+install.packages(c("remotes", "tinytex"), dependencies=TRUE)
+tinytex::install_tinytex()

### 2024-02-27 install.R
Remove binder related stuff

@@ -1,2 +1,0 @@
-install.packages(c("remotes", "tinytex"), dependencies=TRUE)
-tinytex::install_tinytex()

### 2024-02-28 analysis/manuscript.Rmd
update data metadata
minor

@@
 
-Despite these controversies around transformative agreements as a means of transitioning journal publishing to full open access, there is limited evidence available on the adoption of open access in hybrid journals, and the extent to which this can be attributed to transformative agreements.
+Despite these controversies around transformative agreements as a means of transitioning journal publishing to full open access, there is limited evidence available on the uptake of open access in hybrid journals, and the extent to which this can be attributed to transformative agreements.
 Previous studies have focused on specific countries [@P_l_nen_2020; @Huang_2020; @Haucap_2021; @Wenaas_2022; @Taubert_2023] or publisher portfolios [@Pieper_2018; @Bakker_2024; @Fraser_2023; @Jahn_2021; @Momeni_2023; @Schmal_2024], while large-scale studies relied on self-reported agreement data [@Moskovkin_2022], or used APC pricing lists [@Shu_2023].
@@
 - What was the number and proportion of open access articles in hybrid journals in transformative agreements between 2018 and 2022?
-- To what extent did institutions with a transformation agreement contribute to the adoption of open access in hybrid journals?
+- To what extent did institutions with a transformation agreement contribute to open access in hybrid journals?
 
@@
 After mapping ISSN variants to the corresponding linking ISSN (ISSN-L), as provided by the ISSN International Centre, journals were associated with publishers using the ESAC ID, a unique identifier for transformative agreements in the ESAC Transformative Agreement Registry.
-Furthermore, journal subjects according to the All Science Journal Classification code (ASJC) were added from the Scopus journal source list as of August 2023.
+Furthermore, journal subjects according to the All Science Journal Classification (ASJC) were added from the Scopus journal source list as of August 2023.
 
@@
 Because Crossref metadata lacked information to distinguish between original research articles, including reviews, and other types of journal content, which are often not covered by transformative agreements [@Borrego_2020], only articles published in regular issues indicated by non-numeric pagination were included.
-Furthermore, an expanded version of Unpaywall’s paratext recognition approach was applied to exclude non-scholarly journal content such as table of contents.
+Furthermore, paratext recognition was applied to exclude non-scholarly journal content such as table of contents.
 
@@
 The largest differences concerned articles published between 2018 and 2020.
-In 2022, however, Crossref and OpenAlex open access numbers differ only slightly (`r hoaddata::cr_upw |> filter(cr_year %in% 2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, cr_hybrid_total) |> pull(cr_hybrid_total) |> sum()` records using Crossref vs.
+With regard to the publication year 2022, however, Crossref and OpenAlex open access numbers differ only slightly (`r hoaddata::cr_upw |> filter(cr_year %in% 2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, cr_hybrid_total) |> pull(cr_hybrid_total) |> sum()` records using Crossref vs.
 `r hoaddata::cr_upw |> filter(cr_year %in% 2022, cat == "Global", issn_l != "0027-8424") |> distinct(issn_l, cr_year, upw_hybrid_total) |> pull(upw_hybrid_total) |> sum()` in OpenAlex).
@@
 For this study, country and institutional affiliations were retrieved from OpenAlex.
-However, because corresponding author affiliation was only available for 54% of the investigated articles in OpenAlex, this study focused on first authors and their affiliations instead; approximately 90% of the investigated articles had first author affiliation metadata in OpenAlex.
+However, because corresponding author affiliation was only available for 54% of the investigated articles in OpenAlex, this study focused on first authors and their affiliations instead; approximately 90% of the investigated articles had first author affiliation metadata in OpenAlex, which is a much larger proportion than previoulsy reported for the October 2022 snapshot [@Zhang_2024].
 First authors typically contribute the most to a paper and are often considered being lead authors [@Larivi_re_2016]. 
-Related studies also assumed first authors as a proxy to measure open access payments and the impact of transformative agreements [@Zhang_2022; @Shu_2023; @Haucap_2021].
+Related studies also assumed first authors as a proxy to examine open access payments and transformative agreements [@Zhang_2022; @Shu_2023; @Haucap_2021].
 
@@
 
-To assess the impact of transformative agreements on hybrid open access, participating institutions from the Transformative Agreement Data dump, which were crowd-sourced from the agreements and consortia that successfully negotiated an agreement, were matched with the first author affiliations recorded by OpenAlex using the ROR-ID.
+To estimate the impact of transformative agreements on hybrid open access, participating institutions from the Transformative Agreement Data dump, which the cOAlition S crowd-sourced from  agreement documents and consortia, were matched with the first author affiliations recorded by OpenAlex using the ROR-ID.
 Matching also considered the duration of agreements according to the ESAC registry.
@@
 Between 2018 and 2022, `r n_hybrid_jns` out of `r n_hybrid_jns_all` hybrid journals in transformative agreements published at least one open access article under a Creative Commons license.
-During this period, these hybrid journals provided open access to `r n_oa` of the `r n_total` articles, representing a five-year open access proportion of `r n_oa / n_total * 100`%.
+During this period, these hybrid journals provided open access to `r n_oa` out of `r n_total` articles, representing a five-year open access proportion of `r n_oa / n_total * 100`%.
 Authors who could make use of transformative agreements at the time of publication contributed `r n_ta_oa` open access articles.
@@
 Figure \ref{fig:results_overview}B highlights that the majority of hybrid open access was made available through transformative agreements in 2021 and 2022, contributing `r oa_ta_plot_df |> filter(cr_year == "2022", type == "ta_oa") |> pull(value) / oa_by_year |> filter(cr_year == "2022") |> pull(oa) * 100`% of the total open access article volume in 2022.
-However, there was also a notable increase in open access provision through other means, presumably publication fees being not invoiced through transformative agreements, from `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> pull(value)`) in 2018 to  `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> pull(value)`).
+However, there was also a notable increase in open access provision through other means, presumably publication fees being not invoiced through transformative agreements, from `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> pull(value)`) in 2018 to  `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> pull(value)`) in 2022.
 <!-- This suggests that publishers were able to gain equally from individual and institutional open access publishing options.-->
@@
 While OECD publication output in Elsevier's Physical Sciences journals declined from `r publ_subject_output_df |> filter(cr_year == 2018, top_level == "Physical Sciences", publisher == "Elsevier", cat == "OECD") |> pull(all)` articles in 2018 to `r publ_subject_output_df |> filter(cr_year == 2022, top_level == "Physical Sciences", publisher == "Elsevier", cat == "OECD") |> pull(all)` in 2022, BRICS output increased from `r publ_subject_output_df |> filter(cr_year == 2018, top_level == "Physical Sciences", publisher == "Elsevier", cat == "BRICS") |> pull(all)` to `r publ_subject_output_df |> filter(cr_year == 2022, top_level == "Physical Sciences", publisher == "Elsevier", cat == "BRICS") |> pull(all)` in the same five-year period.
-Furthermore, OECD publication output in Health Science Journals and Life Science journals stagnated across the investigated hybrid journal portfolios after peaking in 2020.
+Furthermore, OECD publication output in Health Sciences and Life Sciences journals stagnated after peaking in 2020.
 
@@
 To illustrate the situation in 2022, Figure \ref{fig:country_patch}B compares the total publication output with the number of open access articles. 
-With `r country_scatter_df |> filter(cr_year == "2022", country_code == "CN") |> pull(all)` articles, China was the most productive country, followed by the United States (`r country_scatter_df |> filter(cr_year == "2022", country_code == "US") |> pull(all)` articles) and India (`r country_scatter_df |> filter(cr_year == "2022", country_code == "IN") |> pull(all)` articles).
+With `r country_scatter_df |> filter(cr_year == "2022", country_code == "CN") |> pull(all)` articles, China was the most productive country, followed by the United States (n = `r country_scatter_df |> filter(cr_year == "2022", country_code == "US") |> pull(all)`) and India (n = `r country_scatter_df |> filter(cr_year == "2022", country_code == "IN") |> pull(all)`).
 In contrast, Western and Northern European countries published a considerably high number of open access articles. 
@@
 However, the proliferation of transformative agreements differed across these countries.
-# For instance, Germany successfully negotiated an agreement with Elsevier not until 2023. 
+<!-- For instance, Germany successfully negotiated an agreement with Elsevier not until 2023. -->
 Additionally, publication limits or eligibility criteria for institutions and article types may explain why even countries with widespread agreement implementation such as Sweden or the Netherlands did not achieve 100% hybrid open access.
@@
 Moreover, this study is unable to account for the various types of transformative agreements, particularly with regard to article types and caps that limit the number of open access articles covered.
-Furthermore, assessing the quality of the OpenAlex and Crossref data used, particularly in terms of affiliations and article types, by comparing them with established bibliometric databases such as Scopus and Web of Science was beyond the scope of this analysis.
+Furthermore, assessing the quality of the OpenAlex and Crossref data used, particularly in terms of affiliations and article types, by combining them with established bibliometric databases such as Scopus and Web of Science was beyond the scope of this analysis [@Visser_2021].
 Overall, the methodology is designed in such a way that the results underestimate rather than overestimate the adoption of open access through transformative agreements. 
@@
 Additionally, it must be noted that the study period was significantly impacted by the COVID-19 pandemic, which led to an unprecedented number of publications and a reduction in international collaboration [@Aviv_Reuven_2021], which could explain the observed contrasting developments in OECD and BRICS countries.
-Even before the pandemic, growth in publications in Europe was only due to internationally co-authored journal articles [@Kwiek_2020]. 
-Likewise, inflows from China to the US and European countries already declined by 2020 [@Zhao_2023].
+However, even before the pandemic, growth in publications in Europe was only due to internationally co-authored journal articles [@Kwiek_2020]. 
+Likewise, inflows from China to the United States and European countries already declined by 2020 [@Zhao_2023].
 Furthermore, the study design did not consider emerging publication practices such as preprints [@Fraser_2021] and special issues [@hanson2023strain], which have grown rapidly since 2020. 
@@
 From a data perspective, the reporting of open access funding, including transformative agreements, is not harmonised, but often crowd-sourced from various sources.
-To improve the assessment of transformative agreements, libraries and publishers could agree on standards and services to publicly share information about respective journal portfolios, participating institutions and open access invoicing.
+To improve the assessment of transformative agreements, libraries and publishers need to collaborate on standards and services to publicly share information about respective journal portfolios, participating institutions and open access invoicing.
 

### 2024-02-28 analysis/manuscript.tex
minor

--- 2024-02-27 analysis/manuscript.tex
@@
 
-%\authornote{Correspondence concerning this article should be addressed to Najko Jahn}
-
-\abstract{}
+\authornote{ORCID: https://orcid.org/0000-0001-5105-1463}
+
+\abstract{The ongoing controversy surrounding transformative agreements, which aim to transition journal publishing to full open access, highlight the need for large-scale studies assessing the uptake of open access in hybrid journals.
+This includes evaluating the extent to which transformative agreements enabled open access.
+By combining publicly available data from various sources, including cOAlition S Journal Checker, Crossref, and OpenAlex, this study presents a novel approach that analyses over 700 agreements and nine million journal articles published in more than 11.000 hybrid journals.
+Estimates suggest a strong growth in open access between 2018 and 2022 from 4.3\% to 15\%.
+In 2022, 58\% of hybrid open access was enabled by transformative agreements.
+This trend was largely driven by the three commercial publishers Elsevier, Springer Nature, and Wiley, but the open access uptake varied substantially across journals, publishers, disciplines, and country affiliations.
+In particular, comparing the developments in the OECD and BRICS areas revealed different publication trends relative to hybrid open access.
+In conclusion, estimates suggest that current levels of implementation of transformative agreements is insufficient to bring about a large-scale transition to full open access.}
+
+\keywords{hybrid open access, transformative agreements, scholarly publishing, big deals, bibliometrics}
 
@@
 of transitioning journal publishing to full open access, there is
-limited evidence available on the adoption of open access in hybrid
+limited evidence available on the uptake of open access in hybrid
 journals, and the extent to which this can be attributed to
@@
   To what extent did institutions with a transformation agreement
-  contribute to the adoption of open access in hybrid journals?
+  contribute to open access in hybrid journals?
 \end{itemize}
@@
 Transformative Agreement Registry. Furthermore, journal subjects
-according to the All Science Journal Classification code (ASJC) were
-added from the Scopus journal source list as of August 2023.
+according to the All Science Journal Classification (ASJC) were added
+from the Scopus journal source list as of August 2023.
 
@@
 (Borrego et al., 2021), only articles published in regular issues
-indicated by non-numeric pagination were included. Furthermore, an
-expanded version of Unpaywall's paratext recognition approach was
-applied to exclude non-scholarly journal content such as table of
-contents.
+indicated by non-numeric pagination were included. Furthermore, paratext
+recognition was applied to exclude non-scholarly journal content such as
+table of contents.
 
@@
 The largest differences concerned articles published between 2018 and
-2020. In 2022, however, Crossref and OpenAlex open access numbers differ
-only slightly (249,511 records using Crossref vs.~255,344 in OpenAlex).
-Notable differences could be observed among some publishers that
-presumably did not provide CC license information to Crossref, including
-AIP Publishing, the American Physiological Society, Emerald, and the
-Royal Society. Crossref license metadata was more complete with regard
-to the articles published by Wiley and the American Chemical Society.
-Finally, inconsistent open access status information in previous
-OpenAlex versions was observed (Jahn et al., 2023). After reporting this
-to OpenAlex, fixing this issue was still ongoing according to the
-release notes, which might also explain part of the discrepancy.
+2020. With regard to the publication year 2022, however, Crossref and
+OpenAlex open access numbers differ only slightly (249,511 records using
+Crossref vs.~255,344 in OpenAlex). Notable differences could be observed
+among some publishers that presumably did not provide CC license
+information to Crossref, including AIP Publishing, the American
+Physiological Society, Emerald, and the Royal Society. Crossref license
+metadata was more complete with regard to the articles published by
+Wiley and the American Chemical Society. Finally, inconsistent open
+access status information in previous OpenAlex versions was observed
+(Jahn et al., 2023). After reporting this to OpenAlex, fixing this issue
+was still ongoing according to the release notes, which might also
+explain part of the discrepancy.
 
@@
 90\% of the investigated articles had first author affiliation metadata
-in OpenAlex. First authors typically contribute the most to a paper and
-are often considered being lead authors (Larivière et al., 2016).
-Related studies also assumed first authors as a proxy to measure open
-access payments and the impact of transformative agreements (Haucap et
-al., 2021; Shu \& Larivière, 2023; Zhang et al., 2022).
-
-To assess the impact of transformative agreements on hybrid open access,
-participating institutions from the Transformative Agreement Data dump,
-which were crowd-sourced from the agreements and consortia that
-successfully negotiated an agreement, were matched with the first author
-affiliations recorded by OpenAlex using the ROR-ID. Matching also
-considered the duration of agreements according to the ESAC registry.
-Upon inspection, Transformative Agreement Data did not comprehensively
-cover associated institutions, such as university hospitals or
-institutes of large research organisations such as the Max Planck
-Society. To improve the matching, Transformative Agreement Data were
-automatically enriched with associated organisations using OpenAlex's
-institution entity data.
+in OpenAlex, which is a much larger proportion than previoulsy reported
+for the October 2022 snapshot (Zhang et al., 2024). First authors
+typically contribute the most to a paper and are often considered being
+lead authors (Larivière et al., 2016). Related studies also assumed
+first authors as a proxy to examine open access payments and
+transformative agreements (Haucap et al., 2021; Shu \& Larivière, 2023;
+Zhang et al., 2022).
+
+To estimate the impact of transformative agreements on hybrid open
+access, participating institutions from the Transformative Agreement
+Data dump, which the cOAlition S crowd-sourced from agreement documents
+and consortia, were matched with the first author affiliations recorded
+by OpenAlex using the ROR-ID. Matching also considered the duration of
+agreements according to the ESAC registry. Upon inspection,
+Transformative Agreement Data did not comprehensively cover associated
+institutions, such as university hospitals or institutes of large
+research organisations such as the Max Planck Society. To improve the
+matching, Transformative Agreement Data were automatically enriched with
+associated organisations using OpenAlex's institution entity data.
 
@@
 big data from Crossref, OpenAlex and Unpaywall were imported, using
-bigrquery (Wickham \& Bryan, 2020). To increase computational
+bigrquery (Wickham \& Bryan, 2023). To increase computational
 reproducibility, data aggregation through hoaddata was automatically
@@
 under a Creative Commons license. During this period, these hybrid
-journals provided open access to 742,369 of the 8,146,958 articles,
+journals provided open access to 742,369 out of 8,146,958 articles,
 representing a five-year open access proportion of 9.1\%. Authors who
@@
 not invoiced through transformative agreements, from 4.1\% (n = 62,625)
-in 2018 to 6.3\% (n = 105,896).
+in 2018 to 6.3\% (n = 105,896) in 2022.
 
@@
 to 171,713 in the same five-year period. Furthermore, OECD publication
-output in Health Science Journals and Life Science journals stagnated
-across the investigated hybrid journal portfolios after peaking in 2020.
+output in Health Sciences and Life Sciences journals stagnated after
+peaking in 2020.
 
@@
 articles. With 391,530 articles, China was the most productive country,
-followed by the United States (268,965 articles) and India (87,428
-articles). In contrast, Western and Northern European countries
-published a considerably high number of open access articles.
-Particularly, Scandinavian countries, the Netherlands and Austria
-recorded above-average open access shares, as indicated by the linear
-trend line. As shown in Figure \ref{fig:country_patch}C, transformative
-agreements contributed to these market positions. Interestingly, the
-United States had a notable open access market share of 15\%, although
-transformative agreements contributed to a lesser extent. Similarly,
-China's open access market share of 7.2\% in 2022 was comparable to that
-of the Netherlands, which was 7.1\%.
+followed by the United States (n = 268,965) and India (n = 87,428). In
+contrast, Western and Northern European countries published a
+considerably high number of open access articles. Particularly,
+Scandinavian countries, the Netherlands and Austria recorded
+above-average open access shares, as indicated by the linear trend line.
+As shown in Figure \ref{fig:country_patch}C, transformative agreements
+contributed to these market positions. Interestingly, the United States
+had a notable open access market share of 15\%, although transformative
+agreements contributed to a lesser extent. Similarly, China's open
+access market share of 7.2\% in 2022 was comparable to that of the
+Netherlands, which was 7.1\%.
 
@@
 agreements. However, the proliferation of transformative agreements
-differed across these countries. \# For instance, Germany successfully
-negotiated an agreement with Elsevier not until 2023. Additionally,
-publication limits or eligibility criteria for institutions and article
-types may explain why even countries with widespread agreement
-implementation such as Sweden or the Netherlands did not achieve 100\%
-hybrid open access. Interestingly, in Japan and the United States,
-options other than transformative agreements were the main drivers for
-the increase in hybrid open access. Once again, the graph highlights
-countries with low hybrid open access, particularly non-OECD countries,
-where only a few or no agreements were in place.
+differed across these countries. Additionally, publication limits or
+eligibility criteria for institutions and article types may explain why
+even countries with widespread agreement implementation such as Sweden
+or the Netherlands did not achieve 100\% hybrid open access.
+Interestingly, in Japan and the United States, options other than
+transformative agreements were the main drivers for the increase in
+hybrid open access. Once again, the graph highlights countries with low
+hybrid open access, particularly non-OECD countries, where only a few or
+no agreements were in place.
 
@@
 used, particularly in terms of affiliations and article types, by
-comparing them with established bibliometric databases such as Scopus
-and Web of Science was beyond the scope of this analysis. Overall, the
-methodology is designed in such a way that the results underestimate
-rather than overestimate the adoption of open access through
-transformative agreements. In addition, the data used in this study are
-openly available, along with the code used for this analysis.
+combining them with established bibliometric databases such as Scopus
+and Web of Science was beyond the scope of this analysis (Visser et al.,
+2021). Overall, the methodology is designed in such a way that the
+results underestimate rather than overestimate the adoption of open
+access through transformative agreements. In addition, the data used in
+this study are openly available, along with the code used for this
+analysis.
 
@@
 (Aviv-Reuven \& Rosenfeld, 2021), which could explain the observed
-contrasting developments in OECD and BRICS countries. Even before the
-pandemic, growth in publications in Europe was only due to
+contrasting developments in OECD and BRICS countries. However, even
+before the pandemic, growth in publications in Europe was only due to
 internationally co-authored journal articles (Kwiek, 2021). Likewise,
-inflows from China to the US and European countries already declined by
-2020 (Zhao et al., 2023). Furthermore, the study design did not consider
-emerging publication practices such as preprints (Fraser et al., 2021)
-and special issues (Hanson et al., 2023), which have grown rapidly since
-2020. Lastly, it should be emphasised that the study did not address
-financial shifts between subscriptions spending and open access payments
-while analysing hybrid open access through transformative agreements due
-to a lack of data on expenditures.
+inflows from China to the United States and European countries already
+declined by 2020 (Zhao et al., 2023). Furthermore, the study design did
+not consider emerging publication practices such as preprints (Fraser et
+al., 2021) and special issues (Hanson et al., 2023), which have grown
+rapidly since 2020. Lastly, it should be emphasised that the study did
+not address financial shifts between subscriptions spending and open
+access payments while analysing hybrid open access through
+transformative agreements due to a lack of data on expenditures.
 
@@
 but often crowd-sourced from various sources. To improve the assessment
-of transformative agreements, libraries and publishers could agree on
-standards and services to publicly share information about respective
-journal portfolios, participating institutions and open access
-invoicing.
+of transformative agreements, libraries and publishers need to
+collaborate on standards and services to publicly share information
+about respective journal portfolios, participating institutions and open
+access invoicing.
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Brainard_2023}{}}%
-Brainard, J. (2023). {``Transformative''} journals get booted for
+Brainard, J. (2023). {`Transformative'} journals get booted for
 switching to open access too slowly. \emph{Science}.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Visser_2021}{}}%
+Visser, M., Eck, N. J. van, \& Waltman, L. (2021). Large-scale
+comparison of bibliographic data sources: {Scopus, Web of Science,
+Dimensions, Crossref, and Microsoft Academic}. \emph{Quantitative
+Science Studies}, \emph{2}(1), 20--41.
+\url{https://doi.org/10.1162/qss_a_00112}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Wenaas_2022}{}}%
@@
 \leavevmode\vadjust pre{\hypertarget{ref-bigrquery}{}}%
-Wickham, H., \& Bryan, J. (2020). \emph{{bigrquery}: An interface to
-{Google's BigQuery API}}.
+Wickham, H., \& Bryan, J. (2023). \emph{Bigrquery: An interface to
+google's 'BigQuery' 'API'}.
 \url{https://CRAN.R-project.org/package=bigrquery}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Zhang_2024}{}}%
+Zhang, L., Cao, Z., Shang, Y., Sivertsen, G., \& Huang, Y. (2024).
+Missing institutions in OpenAlex: Possible reasons, implications, and
+solutions. \emph{Scientometrics}.
+\url{https://doi.org/10.1007/s11192-023-04923-y}
 

### 2024-03-18 analysis/manuscript.Rmd

@@
 However, initial approaches, mainly based on publication fees, also called article processing charges (APCs), did not contribute substantially to a large open access uptake. 
-In 2009, Springer reported open access to 1% of articles in hybrid journals [@Dallmeier_2010].
+In 2009, the publisher Springer reported that 1% of its articles in hybrid journals were open access [@Dallmeier_2010].
 Other studies have also recorded low uptake.
@@
 By the end of 2023, many transformative agreements had been implemented, but the interim outcomes were mixed.
-The ESAC Transformative Agreement Registry[^esacregistry], the largest source of disclosure, recorded more than 800 transformative agreements, resulting in up to 900.000 open access articles published in both fully open access and hybrid journals according to the accompanying ESAC Market Watch[^esacmarketwatch].
+The ESAC Transformative Agreement Registry[^esacregistry], the largest resource for library consortia to publicly share their agreements, recorded more than 800 transformative agreements.
+These agreements resulted in the publication of up to 900.000 open access articles published in both fully open access and hybrid journals, according to the accompanying ESAC Market Watch[^esacmarketwatch].
 Library consortia reported increased open access volume, streamlined payment and monitoring procedures, as well as extensive utilisation of open access options by the researchers they serve [@Marques_2020; @Pinhasi_2020; @Parmhed_2023].
@@
 For instance, the Association of Swedish Higher Education Institutions (Sveriges universitets- och
-högskoleförbund, SUHF) recommended only supporting agreements for publishing in fully open access journals [^sweden].
+högskoleförbund, SUHF) recommended only supporting agreements for publishing in fully open access journals by 2026 [^sweden].
 Likewise, most cOAlition S funders will discontinue financial support for transformative agreements by the end of 2024 [@Liverpool_2023].
-The consortium also removed the majority of hybrid journals from its Transformative Journal program in 2023 due to publishers' failure to meet self-defined open access growth targets [@Brainard_2023].
+cOAlition S also removed the majority of hybrid journals from its Transformative Journal program in 2023 due to publishers' failure to meet self-defined open access growth targets [@Brainard_2023]. 
 By contrast, the German DEAL consortium announced a five-year transformative agreement with Elsevier starting in 2024 and also renewed its contracts with Springer Nature and Wiley until the end of 2028.
@@
 Despite these controversies around transformative agreements as a means of transitioning journal publishing to full open access, there is limited evidence available on the uptake of open access in hybrid journals, and the extent to which this can be attributed to transformative agreements.
-Previous studies have focused on specific countries [@P_l_nen_2020; @Huang_2020; @Haucap_2021; @Wenaas_2022; @Taubert_2023] or publisher portfolios [@Pieper_2018; @Bakker_2024; @Fraser_2023; @Jahn_2021; @Momeni_2023; @Schmal_2024], while large-scale studies relied on self-reported agreement data [@Moskovkin_2022], or used APC pricing lists [@Shu_2023].
+Previous studies have focused on specific countries [@P_l_nen_2020; @Huang_2020; @Haucap_2021; @Wenaas_2022; @Taubert_2023; @Jisc_2024] or publisher portfolios [@Pieper_2018; @Bakker_2024; @Fraser_2023; @Jahn_2021; @Momeni_2023; @Schmal_2024], while large-scale studies relied on self-reported agreement data [@Moskovkin_2022], or used APC pricing lists [@Shu_2023].
 In particular, data availability is a limiting factor when studying the impact of transformative agreements [@Bakker_2024], because bibliometric databases, even though many allow the retrieval of open access articles in hybrid journals, do not directly attribute them to specific transformative agreements.
-Likewise, article-level open access invoicing data are only partly available [@Jahn_2021].
+Likewise, article-level open access invoicing and cost data are only partly available [@Jahn_2021; @Kramer_2024].
 
@@
 oa_prop <- readr::read_csv(here::here("data", "oa_prop.csv"))
+oa_prop_yearly <- readr::read_csv(here::here("data", "oa_prop_yearly.csv"))
 
@@
 <!-- Agreements ending before 2021 could not be obtained.-->
-Overall, the four combined Transformative Agreement Data dumps used in this study contained `r  length(unique(hoaddata::jct_hybrid_jns$esac_id))` of `r length(unique(esac_df$ID))` agreements listed in the ESAC registry by December 2023.
+Overall, the four combined Transformative Agreement Data dumps used in this study contained `r  length(unique(hoaddata::jct_hybrid_jns$esac_id))` of `r length(unique(esac_df$ID))` agreements listed in the ESAC Transformative Agreement Registry by December 2023.
 
 The Transformative Agreement Data dumps link agreements to journals represented by journal names and the ISSN. 
-After mapping ISSN variants to the corresponding linking ISSN (ISSN-L), as provided by the ISSN International Centre, journals were associated with publishers using the ESAC ID, a unique identifier for transformative agreements in the ESAC Transformative Agreement Registry.
+After mapping ISSN variants to the corresponding linking ISSN (ISSN-L), as provided by the ISSN International Centre, journals were associated with publishers according to the ESAC Transformative Agreement Registry.
+This reflects that some portfolios may include imprints.
 Furthermore, journal subjects according to the All Science Journal Classification (ASJC) were added from the Scopus journal source list as of August 2023.
@@
 --> November 2023 database snapshot for the five-year period from 2018 to 2022, according to the issued date, representing the earliest known publication date.
-Because Crossref metadata lacked information to distinguish between original research articles, including reviews, and other types of journal content, which are often not covered by transformative agreements [@Borrego_2020], only articles published in regular issues indicated by non-numeric pagination were included.
+Because Crossref metadata lacked information to distinguish between original research articles, including reviews, and other types of journal content, which are often not covered by transformative agreements [@Borrego_2020], only articles published in regular issues indicated by numeric pagination were included.
 Furthermore, paratext recognition was applied to exclude non-scholarly journal content such as table of contents.
@@
 To improve the identification of hybrid journals, journals with an open access proportion above 95% were excluded.
-This step allowed removing additional `r length(unique(oa_prop$issn_l))` fully open access journals.
+This step allowed removing additional `r length(unique(oa_prop_yearly$issn_l))` fully open access journals.
+Together, these journals published `r sum(oa_prop_yearly$articles)` articles between 2018 and 2022.
 
@@
 For this study, country and institutional affiliations were retrieved from OpenAlex.
-However, because corresponding author affiliation was only available for 54% of the investigated articles in OpenAlex, this study focused on first authors and their affiliations instead; approximately 90% of the investigated articles had first author affiliation metadata in OpenAlex, which is a much larger proportion than previoulsy reported for the October 2022 snapshot [@Zhang_2024].
+However, because corresponding author affiliation was only available for 54% of the investigated articles in OpenAlex, this study focused on first authors and their affiliations instead; approximately 90% of the investigated articles had first author affiliation metadata in OpenAlex, which is a much larger proportion than previously reported for the October 2022 snapshot [@Zhang_2024].
 First authors typically contribute the most to a paper and are often considered being lead authors [@Larivi_re_2016]. 
@@
 Following @Marwick_2018, hoaddata contains not only the datasets used in the data analysis. 
-It also includes code used to compile the data by connecting it to the cloud-based Google Big Query data warehouse, where scholarly big data from Crossref, OpenAlex and Unpaywall were imported, using bigrquery [@bigrquery].
+It also includes code used to compile the data by connecting it to a cloud-based Google BigQuery data warehouse, where scholarly big data from Crossref, OpenAlex and Unpaywall were made available, using bigrquery [@bigrquery].
 To increase computational reproducibility, data aggregation through hoaddata was automatically carried out using GitHub Actions, a continuous integration service.
@@
 
-Figure \ref{fig:results_overview}B highlights that the majority of hybrid open access was made available through transformative agreements in 2021 and 2022, contributing `r oa_ta_plot_df |> filter(cr_year == "2022", type == "ta_oa") |> pull(value) / oa_by_year |> filter(cr_year == "2022") |> pull(oa) * 100`% of the total open access article volume in 2022.
+Figure \ref{fig:results_overview}B highlights that the majority of hybrid open access was made available through transformative agreements in 2021 and 2022.
+In 2022, `r  oa_ta_plot_df |> filter(cr_year == "2022", type == "ta_oa") |> pull(value)` out of `r oa_by_year |> filter(cr_year == "2022") |> pull(oa)` open access articles were from eligible authors, representing `r oa_ta_plot_df |> filter(cr_year == "2022", type == "ta_oa") |> pull(value) / oa_by_year |> filter(cr_year == "2022") |> pull(oa) * 100`%.
 However, there was also a notable increase in open access provision through other means, presumably publication fees being not invoiced through transformative agreements, from `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> pull(value)`) in 2018 to  `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> pull(value)`) in 2022.
@@
 `r publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_perc)*100`% Wiley).
-This difference between Elsevier on the one hand and Springer Nature and Wiley on the other can be attributed to transformative agreements, as the latter made the majority of their open access articles available through such deals (Springer Nature `r ((publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(ta_oa_total)) / (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_total)) * 100)`% resp.
-Wiley `r (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(ta_oa_total)) / (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(oa_total)) * 100`%).
+This difference between Elsevier on the one hand and Springer Nature and Wiley on the other can be attributed to transformative agreements, as the latter made the majority of their open access articles available through such deals (`r ((publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(ta_oa_total)) / (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_total)) * 100)`% Springer Nature  resp.
+`r (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(ta_oa_total)) / (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(oa_total)) * 100`% Wiley).
 
@@
 Between 2018 and 2022, high-income countries almost exclusively dominated hybrid open access publishing through transformative agreements. 
-During this period, first authors affiliated with institutions from Organisation for Economic Co-operation and Development (OECD) member countries published `r oecd_stats_global |> filter(cat == "OECD") |> pull(oa)` open access articles in hybrid journals, representing `r oecd_stats_global |> filter(cat == "OECD") |> pull(oa) / n_oa * 100`% of the investigated open access articles.
+To discern socio-economic differences, these countries were grouped according to their membership in the Organisation for Economic Co-operation and Development (OECD).
+During 2018 and 2022, first authors affiliated with institutions from OECD member countries published `r oecd_stats_global |> filter(cat == "OECD") |> pull(oa)` open access articles in hybrid journals, representing `r oecd_stats_global |> filter(cat == "OECD") |> pull(oa) / n_oa * 100`% of the investigated open access articles.
 This disparity between OECD nations and other countries becomes even more evident when considering open access through transformative agreements, as  `r oecd_stats_global |> filter(cat == "OECD") |> pull(ta_oa)` of `r n_ta_oa`, or `r oecd_stats_global |> filter(cat == "OECD") |> pull(ta_oa) / n_ta_oa*100`% of open access articles were associated with such agreements.
@@
 This study presents a novel approach based on open data, which leverages metadata on over 700 agreements and nine million journal articles to estimate the extent to which transformation agreements contribute to the transition of this journal business model.
-The results highlight a strong growth in open access between 2018 and 2022, driven by an increasing number of institutions that had transformative agreements in place. 
+The results highlight a strong growth in open access between 2018 and 2022, driven by an increasing number transformative agreements. 
 However, the majority of research literature published in hybrid journals in this five-year period remained behind publisher paywalls.
 Growth in the adoption of open access in hybrid journals, in particular through transformative agreements, can be largely attributed to three large commercial publishers -- Elsevier, Springer Nature, and Wiley -- but varies substantially across journals, publishers, disciplines, and country affiliations.
-Despite the limitations of the data, the findings indicate that the current level of implementation of transformative agreements is insufficient to bring about a large-scale transition to fully open access.
+Despite the limitations of the data, the findings indicate that the current level of implementation of transformative agreements is insufficient to bring about a large-scale transition to full open access.
 
@@
 Moreover, this study presents varying levels of open access uptake through transformative agreements across journals, which can be attributed to the alignment of authors' affiliations and the availability of such agreements at their institutions.
-In line with previous research findings [@Butler_2023; @Jahn_2021; @Wenaas_2022], high uptake rates were observed across hybrid journals in  Social Sciences and Humanities.
+In line with previous research findings [@Butler_2023; @Jahn_2021; @Wenaas_2022], high uptake rates were observed across Social Sciences and Humanities hybrid journals.
 However, it is important to emphasise that these hybrid journals do not encompass the entire field.
-For example, @Khanna_2022 found that approximately 60% of journals utilizing the open-source publishing platform Open Journal Systems (OJS) fall within the Social Sciences and Humanities. 
+For example, @Khanna_2022 found that approximately 60% of journals utilising the open-source publishing platform Open Journal Systems (OJS) fall within the Social Sciences and Humanities. 
 In these fields, the majority of full open access journals are so-called "Diamond OA journals", which do not charge publication fees [@bosman_2021_4558704].
@@
 This large-scale study provides the first empirical evidence of the influence of transformative agreements on the transition of hybrid journals to full open access. 
-However, several limitations needs to be acknowledged.
+However, several limitations need to be acknowledged.
 From a data perspective, estimations of open access through transformative agreements were established by linking first author affiliations with publicly available agreement data from cOAlition S, and not through invoicing data, which is usually not shared.
@@
 Furthermore, assessing the quality of the OpenAlex and Crossref data used, particularly in terms of affiliations and article types, by combining them with established bibliometric databases such as Scopus and Web of Science was beyond the scope of this analysis [@Visser_2021].
-Overall, the methodology is designed in such a way that the results underestimate rather than overestimate the adoption of open access through transformative agreements. 
+Overall, the methodology is designed to underestimate, rather than overestimate, the adoption of open access through transformative agreements. 
 In addition, the data used in this study are openly available, along with the code used for this analysis.
@@
 This study has practical implications for research funding and libraries.
-Of primary concern should be the observed differences across countries, particularly the relationship between economic development and open access adoption.
+Of primary concern should be the observed differences across countries, particularly the relationship between socio-economic development and open access adoption.
 From a data perspective, the reporting of open access funding, including transformative agreements, is not harmonised, but often crowd-sourced from various sources.

### 2024-03-18 analysis/manuscript.tex
arxiv version

@@
 % number lines
-% \usepackage{lineno}
-% \linenumbers
+ \usepackage{lineno}
+ \linenumbers
 
@@
 
-\abstract{The ongoing controversy surrounding transformative agreements, which aim to transition journal publishing to full open access, highlight the need for large-scale studies assessing the uptake of open access in hybrid journals.
-This includes evaluating the extent to which transformative agreements enabled open access.
-By combining publicly available data from various sources, including cOAlition S Journal Checker, Crossref, and OpenAlex, this study presents a novel approach that analyses over 700 agreements and nine million journal articles published in more than 11.000 hybrid journals.
-Estimates suggest a strong growth in open access between 2018 and 2022 from 4.3\% to 15\%.
-In 2022, 58\% of hybrid open access was enabled by transformative agreements.
+\abstract{The ongoing controversy surrounding transformative agreements, which aim to transition journal publishing to full open access, highlight the need for large-scale studies assessing the impact of these agreements on hybrid open access.
+By combining publicly available data from various sources, including cOAlition S Journal Checker, Crossref, and OpenAlex, this study presents a novel approach that analyses over 700 agreements.
+Results suggest a strong growth in open access between 2018 and 2022 from 4.3\% to 15\%.
+During this period, 11,189 hybrid journals provided open access to 742,369 out of 8,146,958 articles, representing a five-year open access proportion of 9.1\%.
+Authors who could make use of transformative agreements at the time of publication contributed 328,957 open access articles.
+In 2022, 143,615 out of 249,511 open access articles in hybrid journals (58\%) were enabled by transformative agreements.
 This trend was largely driven by the three commercial publishers Elsevier, Springer Nature, and Wiley, but the open access uptake varied substantially across journals, publishers, disciplines, and country affiliations.
-In particular, comparing the developments in the OECD and BRICS areas revealed different publication trends relative to hybrid open access.
-In conclusion, estimates suggest that current levels of implementation of transformative agreements is insufficient to bring about a large-scale transition to full open access.}
+In particular, the OECD and BRICS areas revealed different publication trends.
+In conclusion, this study suggests that current levels of implementation of transformative agreements is insufficient to bring about a large-scale transition to full open access.}
 
@@
 % QSS wants numbered sections
-%\setcounter{secnumdepth}{2}
+\setcounter{secnumdepth}{2}
 
@@
 called article processing charges (APCs), did not contribute
-substantially to a large open access uptake. In 2009, Springer reported
-open access to 1\% of articles in hybrid journals (Dallmeier-Tiessen et
-al., 2010). Other studies have also recorded low uptake. In 2011, only
-1-2\% of articles were open access (Björk, 2012), increasing to around
-4\% between 2011 and 2013 (Laakso \& Björk, 2016).
+substantially to a large open access uptake. In 2009, the publisher
+Springer reported that 1\% of its articles in hybrid journals were open
+access (Dallmeier-Tiessen et al., 2010). Other studies have also
+recorded low uptake. In 2011, only 1-2\% of articles were open access
+(Björk, 2012), increasing to around 4\% between 2011 and 2013 (Laakso \&
+Björk, 2016).
 
@@
 Registry\footnote{\url{https://esac-initiative.org/about/transformative-agreements/agreement-registry/}},
-the largest source of disclosure, recorded more than 800 transformative
-agreements, resulting in up to 900.000 open access articles published in
-both fully open access and hybrid journals according to the accompanying
-ESAC Market Watch\footnote{\url{https://esac-initiative.org/market-watch/}}.
+the largest resource for library consortia to publicly share their
+agreements, recorded more than 800 transformative agreements. These
+agreements resulted in the publication of up to 900.000 open access
+articles published in both fully open access and hybrid journals,
+according to the accompanying ESAC Market Watch\footnote{\url{https://esac-initiative.org/market-watch/}}.
 Library consortia reported increased open access volume, streamlined
@@
 universitets- och högskoleförbund, SUHF) recommended only supporting
-agreements for publishing in fully open access journals \footnote{\url{https://www.su.se/english/news/open-access-need-to-move-away-from-transformative-agreements-1.683787}}.
+agreements for publishing in fully open access journals by 2026
+\footnote{\url{https://www.su.se/english/news/open-access-need-to-move-away-from-transformative-agreements-1.683787}}.
 Likewise, most cOAlition S funders will discontinue financial support
-for transformative agreements by the end of 2024 (Liverpool, 2023). The
-consortium also removed the majority of hybrid journals from its
+for transformative agreements by the end of 2024 (Liverpool, 2023).
+cOAlition S also removed the majority of hybrid journals from its
 Transformative Journal program in 2023 due to publishers' failure to
@@
 transformative agreements. Previous studies have focused on specific
-countries (Haucap et al., 2021; Huang et al., 2020; Pölönen et al.,
-2020; Taubert et al., 2023; Wenaas, 2022) or publisher portfolios
-(Bakker et al., 2024; Fraser et al., 2023; Jahn et al., 2022; Momeni et
-al., 2023; Pieper \& Broschinski, 2018; Schmal, 2024), while large-scale
-studies relied on self-reported agreement data (Moskovkin et al., 2022),
-or used APC pricing lists (Shu \& Larivière, 2023). In particular, data
-availability is a limiting factor when studying the impact of
-transformative agreements (Bakker et al., 2024), because bibliometric
-databases, even though many allow the retrieval of open access articles
-in hybrid journals, do not directly attribute them to specific
-transformative agreements. Likewise, article-level open access invoicing
-data are only partly available (Jahn et al., 2022).
+countries (Brayman et al., 2024; Haucap et al., 2021; Huang et al.,
+2020; Pölönen et al., 2020; Taubert et al., 2023; Wenaas, 2022) or
+publisher portfolios (Bakker et al., 2024; Fraser et al., 2023; Jahn et
+al., 2022; Momeni et al., 2023; Pieper \& Broschinski, 2018; Schmal,
+2024), while large-scale studies relied on self-reported agreement data
+(Moskovkin et al., 2022), or used APC pricing lists (Shu \& Larivière,
+2023). In particular, data availability is a limiting factor when
+studying the impact of transformative agreements (Bakker et al., 2024),
+because bibliometric databases, even though many allow the retrieval of
+open access articles in hybrid journals, do not directly attribute them
+to specific transformative agreements. Likewise, article-level open
+access invoicing and cost data are only partly available (Jahn et al.,
+2022; Kramer, 2024).
 
@@
 Transformative Agreement Data dumps used in this study contained 729 of
-869 agreements listed in the ESAC registry by December 2023.
+869 agreements listed in the ESAC Transformative Agreement Registry by
+December 2023.
 
@@
 to the corresponding linking ISSN (ISSN-L), as provided by the ISSN
-International Centre, journals were associated with publishers using the
-ESAC ID, a unique identifier for transformative agreements in the ESAC
-Transformative Agreement Registry. Furthermore, journal subjects
-according to the All Science Journal Classification (ASJC) were added
-from the Scopus journal source list as of August 2023.
+International Centre, journals were associated with publishers according
+to the ESAC Transformative Agreement Registry. This reflects that some
+portfolios may include imprints. Furthermore, journal subjects according
+to the All Science Journal Classification (ASJC) were added from the
+Scopus journal source list as of August 2023.
 
@@
 (Borrego et al., 2021), only articles published in regular issues
-indicated by non-numeric pagination were included. Furthermore, paratext
+indicated by numeric pagination were included. Furthermore, paratext
 recognition was applied to exclude non-scholarly journal content such as
@@
 proportion above 95\% were excluded. This step allowed removing
-additional 241 fully open access journals.
+additional 134 fully open access journals. Together, these journals
+published 8,565 articles between 2018 and 2022.
 
@@
 90\% of the investigated articles had first author affiliation metadata
-in OpenAlex, which is a much larger proportion than previoulsy reported
+in OpenAlex, which is a much larger proportion than previously reported
 for the October 2022 snapshot (Zhang et al., 2024). First authors
@@
 analysis. It also includes code used to compile the data by connecting
-it to the cloud-based Google Big Query data warehouse, where scholarly
-big data from Crossref, OpenAlex and Unpaywall were imported, using
+it to a cloud-based Google BigQuery data warehouse, where scholarly big
+data from Crossref, OpenAlex and Unpaywall were made available, using
 bigrquery (Wickham \& Bryan, 2023). To increase computational
@@
 hybrid open access was made available through transformative agreements
-in 2021 and 2022, contributing 58\% of the total open access article
-volume in 2022. However, there was also a notable increase in open
-access provision through other means, presumably publication fees being
-not invoiced through transformative agreements, from 4.1\% (n = 62,625)
-in 2018 to 6.3\% (n = 105,896) in 2022.
+in 2021 and 2022. In 2022, 143,615 out of 249,511 open access articles
+were from eligible authors, representing 58\%. However, there was also a
+notable increase in open access provision through other means,
+presumably publication fees being not invoiced through transformative
+agreements, from 4.1\% (n = 62,625) in 2018 to 6.3\% (n = 105,896) in
+2022.
 
@@
 the latter made the majority of their open access articles available
-through such deals (Springer Nature 57\% resp. Wiley 55\%).
+through such deals (57\% Springer Nature resp. 55\% Wiley).
 
@@
 dominated hybrid open access publishing through transformative
-agreements. During this period, first authors affiliated with
-institutions from Organisation for Economic Co-operation and Development
-(OECD) member countries published 602,050 open access articles in hybrid
-journals, representing 81\% of the investigated open access articles.
-This disparity between OECD nations and other countries becomes even
-more evident when considering open access through transformative
-agreements, as 310,712 of 328,957, or 94\% of open access articles were
-associated with such agreements.
+agreements. To discern socio-economic differences, these countries were
+grouped according to their membership in the Organisation for Economic
+Co-operation and Development (OECD). During 2018 and 2022, first authors
+affiliated with institutions from OECD member countries published
+602,050 open access articles in hybrid journals, representing 81\% of
+the investigated open access articles. This disparity between OECD
+nations and other countries becomes even more evident when considering
+open access through transformative agreements, as 310,712 of 328,957, or
+94\% of open access articles were associated with such agreements.
 
@@
 of this journal business model. The results highlight a strong growth in
-open access between 2018 and 2022, driven by an increasing number of
-institutions that had transformative agreements in place. However, the
-majority of research literature published in hybrid journals in this
-five-year period remained behind publisher paywalls. Growth in the
-adoption of open access in hybrid journals, in particular through
-transformative agreements, can be largely attributed to three large
-commercial publishers -- Elsevier, Springer Nature, and Wiley -- but
-varies substantially across journals, publishers, disciplines, and
-country affiliations. Despite the limitations of the data, the findings
-indicate that the current level of implementation of transformative
-agreements is insufficient to bring about a large-scale transition to
-fully open access.
+open access between 2018 and 2022, driven by an increasing number
+transformative agreements. However, the majority of research literature
+published in hybrid journals in this five-year period remained behind
+publisher paywalls. Growth in the adoption of open access in hybrid
+journals, in particular through transformative agreements, can be
+largely attributed to three large commercial publishers -- Elsevier,
+Springer Nature, and Wiley -- but varies substantially across journals,
+publishers, disciplines, and country affiliations. Despite the
+limitations of the data, the findings indicate that the current level of
+implementation of transformative agreements is insufficient to bring
+about a large-scale transition to full open access.
 
@@
 previous research findings (Butler et al., 2023; Jahn et al., 2022;
-Wenaas, 2022), high uptake rates were observed across hybrid journals in
-Social Sciences and Humanities. However, it is important to emphasise
+Wenaas, 2022), high uptake rates were observed across Social Sciences
+and Humanities hybrid journals. However, it is important to emphasise
 that these hybrid journals do not encompass the entire field. For
 example, Khanna et al. (2022) found that approximately 60\% of journals
-utilizing the open-source publishing platform Open Journal Systems (OJS)
+utilising the open-source publishing platform Open Journal Systems (OJS)
 fall within the Social Sciences and Humanities. In these fields, the
@@
 influence of transformative agreements on the transition of hybrid
-journals to full open access. However, several limitations needs to be
+journals to full open access. However, several limitations need to be
 acknowledged. From a data perspective, estimations of open access
@@
 and Web of Science was beyond the scope of this analysis (Visser et al.,
-2021). Overall, the methodology is designed in such a way that the
-results underestimate rather than overestimate the adoption of open
-access through transformative agreements. In addition, the data used in
-this study are openly available, along with the code used for this
-analysis.
+2021). Overall, the methodology is designed to underestimate, rather
+than overestimate, the adoption of open access through transformative
+agreements. In addition, the data used in this study are openly
+available, along with the code used for this analysis.
 
@@
 libraries. Of primary concern should be the observed differences across
-countries, particularly the relationship between economic development
-and open access adoption. From a data perspective, the reporting of open
-access funding, including transformative agreements, is not harmonised,
-but often crowd-sourced from various sources. To improve the assessment
-of transformative agreements, libraries and publishers need to
-collaborate on standards and services to publicly share information
-about respective journal portfolios, participating institutions and open
-access invoicing.
+countries, particularly the relationship between socio-economic
+development and open access adoption. From a data perspective, the
+reporting of open access funding, including transformative agreements,
+is not harmonised, but often crowd-sourced from various sources. To
+improve the assessment of transformative agreements, libraries and
+publishers need to collaborate on standards and services to publicly
+share information about respective journal portfolios, participating
+institutions and open access invoicing.
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Brainard_2023}{}}%
-Brainard, J. (2023). {`Transformative'} journals get booted for
+Brainard, J. (2023). {``Transformative''} journals get booted for
 switching to open access too slowly. \emph{Science}.
 \url{https://doi.org/10.1126/science.adj3282}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Jisc_2024}{}}%
+Brayman, K., Devenney, A., Dobson, H., Marques, M., \& Vernon, A.
+(2024). \emph{A review of transitional agreements in the {UK}}. Zenodo.
+\url{https://doi.org/10.5281/zenodo.10787392}
 
@@
 \url{https://doi.org/10.1177/09610006221146771}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Kramer_2024}{}}%
+Kramer, B. (2024). \emph{Study on scientific publishing in {Europe} --
+{Development}, diversity, and transparency of costs}. Publications
+Office of the European Union. \url{https://doi.org/doi/10.2777/89349}
 

### 2024-09-11 analysis/manuscript.Rmd
Reviewer 1 suggestions to improve the introduction

@@
 
-As a consequence, libraries and their consortia began to develop licensing strategies aimed at avoiding such ‘double dipping’ scenarios, in which well-established commercial publishers gain twice from reading and open access publishing [@Mittermaier_2015], as well as to increase publisher-provided immediate open access [@BJ_RK_2014; @Schimmer_2015].
+As a consequence, libraries and their consortia began to develop licensing strategies aimed at avoiding such ‘double dipping’ scenarios, in which well-established commercial publishers gain twice from reading and open access publishing fees [@Mittermaier_2015], as well as to increase publisher-provided immediate open access [@BJ_RK_2014; @Schimmer_2015].
 These considerations resulted in transformative agreements[^agreements], which cover a broad range of contracts between library consortia and publishers from the mid-2010s onwards, where institutional spending for subscriptions and open access publishing are considered together [@Borrego_2020; @Hinchliffe_2019].
@@
 By the end of 2023, many transformative agreements had been implemented, but the interim outcomes were mixed.
-The ESAC Transformative Agreement Registry[^esacregistry], the largest resource for library consortia to publicly share their agreements, recorded more than 800 transformative agreements.
+The ESAC Transformative Agreement Registry[^esacregistry], the largest resource for library consortia to voluntarily and publicly share their agreements, recorded more than 800 transformative agreements.
 These agreements resulted in the publication of up to 900.000 open access articles published in both fully open access and hybrid journals, according to the accompanying ESAC Market Watch[^esacmarketwatch].

### 2024-09-11 analysis/manuscript.Rmd
Reviewer 1: Make description of CC and ROR-ID retrieval and journal inclusion clearer

@@
 
-Open access articles in hybrid journals were identified using the Creative Commons (CC) license information in Crossref metadata. 
+Open access articles in hybrid journals were identified using the Creative Commons (CC) license information in Crossref metadata, with consideration of any CC variant. 
 License information relative to the "accepted manuscript (AM)" version was not considered. 
@@
 Upon inspection, Transformative Agreement Data did not comprehensively cover associated institutions, such as university hospitals or institutes of large research organisations such as the Max Planck Society.
-To improve the matching, Transformative Agreement Data were automatically enriched with associated organisations using OpenAlex's institution entity data.
-
-The so compiled data set consists of `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l, cr_year, jn_all) |> pull(jn_all) |> sum()` articles published in `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l) |> nrow()` hybrid journals between 2018 and 2022 (see Figure \ref{fig:method_fig}B).
-Hybrid journals in transformative agreements represented `r cr_stats |> filter(src == "hoad") |> pull(n_articles) |> sum() / cr_stats |> pull(n_articles) |> sum() * 100`%  of the total global output over the same time period, according to Crossref, while fully open access journals recorded
+To improve the matching, Transformative Agreement Data were automatically enriched with ROR-IDs from associated organisations according to OpenAlex's institution entity data.
+
+The so compiled data set consists of `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l, cr_year, jn_all) |> pull(jn_all) |> sum()` articles published in `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l) |> nrow()` hybrid journals included in at least one transformative agreement between 2018 and 2022 (see Figure \ref{fig:method_fig}B).
+These hybrid journals in transformative agreements represented `r cr_stats |> filter(src == "hoad") |> pull(n_articles) |> sum() / cr_stats |> pull(n_articles) |> sum() * 100`%  of the total global output over the same time period, according to Crossref, while fully open access journals recorded
 `r cr_stats |> filter(is.na(src) & oa_status == "gold") |> pull(n_articles) |> sum() / cr_stats |> pull(n_articles) |> sum() * 100`%.

### 2024-09-11 analysis/manuscript.Rmd
Reviewer 1: Improve the presentation of the overall statistics and make the graph color- and printer-firendly (Thanks @gitti1)

@@
 esac_df <- readxl::read_xlsx(here::here("data", "esac.xlsx"), skip = 2)
+
+# TA statistics
+## TAs in ESAC as of 13 December
+tas_in_esac <- esac_df |> 
+  distinct(ID)
+tas_in_esac_and_hoad <- esac_df |>
+  filter(ID %in% hoaddata::jct_hybrid_jns$esac_id) |>
+  filter(Publisher != "National Academy of Sciences")
+tas_in_esac_and_hoad_active_in_2018_22 <- tas_in_esac_and_hoad |>
+  filter(`End date` >= as.Date("2018-01-01") & `Start date` < as.Date("2023-01-01"))
+
 #' OA journal check
@@
     values = c(
-      "hoad" = "#5789B6",
-      "other" = "#D8D4C9",
-      "gold" = "#DA6524"
+      "hoad" = "#998ec3",
+      "other" = "#f7f7f7",
+      "gold" = "#f1a340"
     ),
@@
 Matching also considered the duration of agreements according to the ESAC registry. 
+`r length(unique(tas_in_esac_and_hoad_active_in_2018_22$ID))` agreements were active between 2018 and 2022.
 Upon inspection, Transformative Agreement Data did not comprehensively cover associated institutions, such as university hospitals or institutes of large research organisations such as the Max Planck Society.
@@
 
-The so compiled data set consists of `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l, cr_year, jn_all) |> pull(jn_all) |> sum()` articles published in `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l) |> nrow()` hybrid journals included in at least one transformative agreement between 2018 and 2022 (see Figure \ref{fig:method_fig}B).
+The so compiled data set consists of `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l, cr_year, jn_all) |> pull(jn_all) |> sum()` articles published in `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l) |> nrow()` hybrid journals included in at least one transformative agreement between 2018 and 2022 (see purple bars in Figure \ref{fig:method_fig}B).
 These hybrid journals in transformative agreements represented `r cr_stats |> filter(src == "hoad") |> pull(n_articles) |> sum() / cr_stats |> pull(n_articles) |> sum() * 100`%  of the total global output over the same time period, according to Crossref, while fully open access journals recorded
@@
 (B) Number of articles in Crossref by journal types.
-The blue bars show the overall article volume of hybrid journals in transformative agreements, which were initially included in the study, in comparision with full open access journals according to OpenAlex.
-The remainder represents subscription-based journals not covered by transformative agreements."}
+The purple bars show the overall article volume of hybrid journals in transformative agreements, which were initially included in the study, in comparision with full open access journals according to OpenAlex.
+The remainder represents articles in subscription-based journals not covered by transformative agreements."}
 (p | cr_stats_plot) + plot_annotation(tag_levels = 'A') +  plot_layout(widths = c(0.6, 0.4))
@@
 
-Between 2018 and 2022, `r n_hybrid_jns` out of `r n_hybrid_jns_all` hybrid journals in transformative agreements published at least one open access article under a Creative Commons license.
-During this period, these hybrid journals provided open access to `r n_oa` out of `r n_total` articles, representing a five-year open access proportion of `r n_oa / n_total * 100`%.
-Authors who could make use of transformative agreements at the time of publication contributed `r n_ta_oa` open access articles.
-Overall, `r n_tas_with_oa` transformative agreements recorded at least one open access article from an eligible institution.
+Between 2018 and 2022, `r n_hybrid_jns` out of `r n_hybrid_jns_all` eligible hybrid journals in transformative agreements published at least one open access article under a Creative Commons license.
+These `r n_hybrid_jns` hybrid journals constituted the foundation for the subsequent analyses. 
+Collectively, they provided open access to `r n_oa` out of `r n_total` articles during the investigated period, representing a five-year open access proportion of `r n_oa / n_total * 100`%.
+Authors who could make use of transformative agreements at the time of publication contributed `r n_ta_oa` (`r n_ta_oa / n_oa * 100`%) open access articles.
+Overall, this investigation established a link between open access articles and eligible institutions for `r n_tas_with_oa` out of `r length(unique(tas_in_esac_and_hoad_active_in_2018_22$ID))` active transformative agreements between 2018 and 2022 through the analysis of publication metadata, including affiliations and agreement data.
 

### 2024-09-12 analysis/manuscript.Rmd
Reviewer 1: Introduce 50% grid lines, in case it was missing; fix wrong inline calculation for Big 3 publisher article volume

@@
 Authors who could make use of transformative agreements at the time of publication contributed `r n_ta_oa` (`r n_ta_oa / n_oa * 100`%) open access articles.
-Overall, this investigation established a link between open access articles and eligible institutions for `r n_tas_with_oa` out of `r length(unique(tas_in_esac_and_hoad_active_in_2018_22$ID))` active transformative agreements between 2018 and 2022 through the analysis of publication metadata, including affiliations and agreement data.
+Overall, this investigation was able to establish a link between open access articles and eligible institutions for `r n_tas_with_oa` out of `r length(unique(tas_in_esac_and_hoad_active_in_2018_22$ID))` (`r n_tas_with_oa / length(unique(tas_in_esac_and_hoad_active_in_2018_22$ID)) * 100`%) active transformative agreements between 2018 and 2022 through the analysis of publication metadata, including affiliations and agreement data.
 
@@
     legend.direction = "horizontal",
-    panel.grid.minor = element_blank(),
+ #   panel.grid.minor = element_blank(),
     panel.border = element_rect(color = "grey50", fill = NA),
@@
 Figure \ref{fig:results_overview}A shows a moderate growth in the proportion of open access articles in hybrid journals, comparing the overall open access uptake and the impact of transformative agreements on this trend.
-Over the five-years period from 2018 to 2022, open access increased from `r oa_by_year |> filter(cr_year == "2018") |> pull(prop)`% (n = `r oa_by_year |> filter(cr_year == "2018") |> pull(oa)`) to `r oa_by_year |> filter(cr_year == "2022") |> pull(prop)`% (n = `r oa_by_year |> filter(cr_year == "2022") |> pull(oa)`).
+Over the five-years period from 2018 to 2022, open access in hybrid journals increased from `r oa_by_year |> filter(cr_year == "2018") |> pull(prop)`% (n = `r oa_by_year |> filter(cr_year == "2018") |> pull(oa)`) to `r oa_by_year |> filter(cr_year == "2022") |> pull(prop)`% (n = `r oa_by_year |> filter(cr_year == "2022") |> pull(oa)`).
 Simultaneously, the total article volume of the investigated journals rose from `r oa_by_year |> filter(cr_year == "2018") |> pull(articles)` in 2018 to `r oa_by_year |> filter(cr_year == "2022") |> pull(articles)` in 2022.
@@
 In 2022, `r  oa_ta_plot_df |> filter(cr_year == "2022", type == "ta_oa") |> pull(value)` out of `r oa_by_year |> filter(cr_year == "2022") |> pull(oa)` open access articles were from eligible authors, representing `r oa_ta_plot_df |> filter(cr_year == "2022", type == "ta_oa") |> pull(value) / oa_by_year |> filter(cr_year == "2022") |> pull(oa) * 100`%.
-However, there was also a notable increase in open access provision through other means, presumably publication fees being not invoiced through transformative agreements, from `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> pull(value)`) in 2018 to  `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> pull(value)`) in 2022.
+However, there was also a notable increase in open access provision through other means, presumably publication fees being not invoiced through transformative agreements as indicated by the grey area in \ref{fig:results_overview}A, from `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2018", type == "no_ta") |> pull(value)`) in 2018 to  `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> mutate(prop = value / n * 100) |> pull(prop)`% (n = `r oa_ta_plot_df |> filter(cr_year == "2022", type == "no_ta") |> pull(value)`) in 2022.
 <!-- This suggests that publishers were able to gain equally from individual and institutional open access publishing options.-->
@@
 Analysing hybrid open access across publishers between 2018 and 2022 reveals a large market concentration. 
-Although `r length(unique(jn_ind_jct_all$esac_publisher))` publishers offered transformative agreements, the three commercial publishers Elsevier, Springer Nature, and Wiley, accounted for `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(jn_cum) * 100`% of the total article volume published (see Table \ref{tab:publisher_league_table}).
+Although `r length(unique(jn_ind_jct_all$esac_publisher))` publishers offered transformative agreements, the three commercial publishers Elsevier, Springer Nature, and Wiley accounted for `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(jn_cum) * 100`% of hybrid journals, representing `r publisher_league_tbl  |> filter(publisher != "Other") |> pull(publisher_total) |> sum()` or `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(cum) * 100`% of the total article volume (see Table \ref{tab:publisher_league_table}).
 Together, they published `r publisher_league_tbl  |> filter(publisher != "Other") |> pull(oa_total) |> sum()` or `r publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(oa_cum) * 100`% of the open access articles in hybrid journals.
@@
 Note that data on transformative agreements ending before June 2021 were not available for this study."}
-library(patchwork)
+
 design = "E
@@
     legend.direction = "horizontal",
-    panel.grid.minor = element_blank(),
+#    panel.grid.minor = element_blank(),
                 panel.border = element_rect(color = "grey50", fill = NA),

### 2024-09-12 analysis/manuscript.Rmd
Reviewer 1: Fix figure reference

@@
 However, the Springer Compact agreements were not included in the data as they concluded before the start of the transformative agreement data collection in June 2021. 
-Nonetheless, the results demonstrate the importance of agreements for Springer Nature's hybrid open access business over the past five years (Figure 2B).
+Nonetheless, the results demonstrate the importance of agreements for Springer Nature's hybrid open access business over the past five years (Figure \ref{fig:publisher_figure}).
 In 2022, `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature") |> pull(value) |> sum()*100`% (n = `r  publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature", type == "ta_oa") |> pull(value)`) of open access in Springer Nature hybrid journals was enabled through transformative agreements.

### 2024-09-12 analysis/manuscript.Rmd
Reviewer 1: Minor fix

@@
 In 2022, `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature") |> pull(value) |> sum()*100`% (n = `r  publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature", type == "ta_oa") |> pull(value)`) of open access in Springer Nature hybrid journals was enabled through transformative agreements.
-In the same year, `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley", type == "ta_oa") |> pull(value)`) of Wiley's open access articles could be linked to transformative agreements in 2022.
-By contrast, Elsevier published fewer than half of its open access articles through transformative agreements (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier", type == "ta_oa") |> pull(value) `; `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(value) |> sum()*100`%).
+In the same year, `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley", type == "ta_oa") |> pull(value)`) of Wiley's open access articles could be linked to transformative agreements.
+By contrast, Elsevier published a comparatively lower proportion of its open access articles through transformative agreements in 2022 (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier", type == "ta_oa") |> pull(value) `; `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(value) |> sum()*100`%).
 

### 2024-09-12 analysis/manuscript.Rmd
Reviewer 1: Fix bugs in country analysis

@@
   scale_fill_manual(NULL, values = c("OECD" = "#d01c8b", "BRICS" = "#4dac26", "Other" = "#f7f7f7")) +
-  geom_smooth(aes(all, oa / all), method = lm) +
+  geom_smooth(aes(all, oa / all), method = "lm", se = FALSE) +
   scale_size("Agreement", guide = FALSE) +
@@
 oecd_stats <- readr::read_csv(here::here("data", "oecd_stats.csv"))
+
+oecd_by_year <- oecd_stats |>
+  filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
+  filter(issn_l %in% hoaddata::cc_articles$issn_l) |>
+  group_by(cr_year) |>
+  summarise(yearly_volume = sum(articles, na.rm = TRUE))
+
 oecd_stats_df <- oecd_stats |>
@@
   pivot_longer(cols = c(ta_oa, oa_other, closed), names_to = "oa_type") |>
-  mutate(cat = ifelse(is.na(cat), "Other", cat))
-
-country_bar <- ggplot(oecd_stats_df, aes(gsub("^20", "'", as.character(cr_year)), value, group = oa_type, fill = oa_type)) +
-  geom_area(color = "#000000",
-            linewidth = 0.15) +
-  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels =  scales::number_format(big.mark = ",")) +
-  facet_wrap(~ cat) +
-  scale_fill_manual(NULL, values = c(oa_other = "#b3b3b3a0", ta_oa = "#56B4E9", closed = "#e1e1e1a0"),
-                    labels = c(oa_other = "Other", ta_oa = "Agreement", closed = "NA"),  guide = guide_legend(reverse = TRUE)) +
-  theme_minimal() +
-  labs(x = "Publication year", y = "Total articles")
+  mutate(cat = ifelse(is.na(cat), "Other", cat)) |>
+  left_join(oecd_by_year, by = "cr_year")
+
+country_bar <- ggplot(oecd_stats_df,
+                      aes(
+                        gsub("^20", "'", as.character(cr_year)),
+                        value / yearly_volume,
+                        group = oa_type,
+                        fill = oa_type
+                      )) +
+  geom_area(color = "#000000", linewidth = 0.15) +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels = scales::percent_format()) +
+  facet_wrap( ~ cat) +
+  scale_fill_manual(
+    NULL,
+    values = c(
+      oa_other = "#b3b3b3a0",
+      ta_oa = "#56B4E9",
+      closed = "#e1e1e1a0"
+    ),
+    labels = c(
+      oa_other = "Other",
+      ta_oa = "Agreement",
+      closed = "Closed"
+    ),
+    guide = guide_legend(reverse = TRUE)
+  ) +
+  theme_minimal() +
+  labs(x = "Publication year", y = "Hybrid OA Uptake")
 ```
@@
    theme_minimal() +
-  labs(x = "Publication year", y = "Total articles")
+  labs(x = "Publication year", y = "Hybrid OA Uptake")
 
@@
 
-Figure \ref{fig:country_patch}A shows the development of hybrid open access publishing by country, comparing the OECD area with the BRICS, an intergovernmental organisation, which comprised  Brazil, Russia, India, China, and South Africa as of 2022.
-The residual category "Other" includes the remaining countries. 
-Full counting was applied to account for multiple country affiliations [@Hottenrott_2021].
-From 2018 to 2022, the proportion of open access in hybrid journals increased from `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2018) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2018 to `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2022) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2022.
-In contrast, BRICS recorded a low uptake, moderately growing from `r country_bar$data |> filter(cat == "BRICS", oa_type != "closed", cr_year == 2018) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2018 to `r country_bar$data |> filter(cat == "BRICS", oa_type != "closed", cr_year == 2022) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2022.
-
-Despite the rise of open access across OECD countries, the overall publication output decreased sharply, dropping to `r country_bar$data |> filter(cat == "OECD", cr_year == "2022") |> distinct(all) |> pull(all)` in 2022 after peaking `r country_bar$data |> filter(cat == "OECD", cr_year == "2020") |> distinct(all) |> pull(all)` articles in 2020.
-In stark contrast, the number of articles published in hybrid journals by first authors affiliated with institutions from BRICS countries increased steadily over the years, more than doubling from `r country_bar$data |> filter(cat == "BRICS", cr_year == "2018") |> distinct(all) |> pull(all)` in 2018 to `r country_bar$data |> filter(cat == "OECD", cr_year == "2022") |> distinct(all) |> pull(all)` in 2022.
-Upon closer examination, this trend can be observed across all of the three largest publishers, although the shift towards BRICS is particularly evident in Elsevier's hybrid journal portfolio, particularly with regard to articles published in Physical Sciences journals.
-While OECD publication output in Elsevier's Physical Sciences journals declined from `r publ_subject_output_df |> filter(cr_year == 2018, top_level == "Physical Sciences", publisher == "Elsevier", cat == "OECD") |> pull(all)` articles in 2018 to `r publ_subject_output_df |> filter(cr_year == 2022, top_level == "Physical Sciences", publisher == "Elsevier", cat == "OECD") |> pull(all)` in 2022, BRICS output increased from `r publ_subject_output_df |> filter(cr_year == 2018, top_level == "Physical Sciences", publisher == "Elsevier", cat == "BRICS") |> pull(all)` to `r publ_subject_output_df |> filter(cr_year == 2022, top_level == "Physical Sciences", publisher == "Elsevier", cat == "BRICS") |> pull(all)` in the same five-year period.
-Furthermore, OECD publication output in Health Sciences and Life Sciences journals stagnated after peaking in 2020.
-
 ```{r country_patch, fig.asp=0.9, fig.cap="Development of hybrid open access publishing by country.
-(A) presents the number of articles published in hybrid journals included in transformative agreement, distinguishing between BRICS as of 20222, OECD and other countries.
-(B) Scatterplot contrasting total articles with open access article volume in 2022, by country.
-Point size represents the number of articles enabled through transformative agreements.
+(A) presents the relative number of articles published in hybrid journals included in transformative agreement by year, distinguishing between BRICS as of 2022, OECD and other countries.
+(B) Scatterplot contrasts total articles with open access article volume in 2022, by country and its OECD (purple-colored points) or BRICS (green) membership.
+Trend line obtained from linear regression, grey areas show pointwise symmetric 95% confidence bands.
 (C) Hybrid open access market share in 2022 by country.
 In (A) and (C), the blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established.
+The remainder shows closed access articles.
 Country names are represented as ISO two-letter country codes."}
-library(patchwork)
+
 wrap_plots(
@@
 
+Figure \ref{fig:country_patch}A shows the development of hybrid open access publishing by country, comparing the OECD area with the BRICS, an intergovernmental organisation, which comprised  Brazil, Russia, India, China, and South Africa as of 2022.
+The residual category "Other" includes the remaining countries. 
+Full counting was applied to account for multiple country affiliations [@Hottenrott_2021].
+Open access in hybrid journals increased in particular because of publications from authors affiliated with institutions from the OECD area, from `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2018) |> pull(value) |> sum()` in 2018 to `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2022) |> pull(value) |> sum()` in 2022.
+This growth was largely driven by transformative agreements.
+The share of open access by agreement rose from `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2018, oa_type == "ta_oa") |> pull(value) / country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2018, oa_type == "ta_oa") |> pull(oa) * 100`% in 2018 (n = `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2018, oa_type == "ta_oa") |> pull(value)`) to `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2022, oa_type == "ta_oa") |> pull(value) / country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2022, oa_type == "ta_oa") |> pull(oa) * 100`% (n = `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2022, oa_type == "ta_oa") |> pull(value)`) in 2022.
+In contrast, BRICS recorded a low uptake, moderately growing from `r country_bar$data |> filter(cat == "BRICS", oa_type != "closed", cr_year == 2018) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2018 to `r country_bar$data |> filter(cat == "BRICS", oa_type != "closed", cr_year == 2022) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2022.
+
+Despite the rise of open access across OECD countries, the overall publication output decreased sharply, dropping to `r country_bar$data |> filter(cat == "OECD", cr_year == "2022") |> distinct(all) |> pull(all)` in 2022 after peaking `r country_bar$data |> filter(cat == "OECD", cr_year == "2020") |> distinct(all) |> pull(all)` articles in 2020.
+In stark contrast, the number of articles published in hybrid journals by first authors affiliated with institutions from BRICS countries increased steadily over the years, from `r country_bar$data |> filter(cat == "BRICS", cr_year == "2018") |> distinct(all) |> pull(all)` in 2018 to `r country_bar$data |> filter(cat == "BRICS", cr_year == "2022") |> distinct(all) |> pull(all)` in 2022.
+This resulted in an increase in market share of the BRICS area from `r country_bar$data |> filter(cat == "BRICS", cr_year == "2018") |> distinct(all, yearly_volume) |> mutate(prop = all /yearly_volume) |> pull(prop) * 100`% to `r country_bar$data |> filter(cat == "BRICS", cr_year == "2022") |> distinct(all, yearly_volume) |> mutate(prop = all /yearly_volume) |> pull(prop) * 100`% between 2018 and 2022, whereas that of the OECD area decreased from `r country_bar$data |> filter(cat == "OECD", cr_year == "2018") |> distinct(all, yearly_volume) |> mutate(prop = all /yearly_volume) |> pull(prop) * 100`% to `r country_bar$data |> filter(cat == "OECD", cr_year == "2022") |> distinct(all, yearly_volume) |> mutate(prop = all /yearly_volume) |> pull(prop) * 100`% during the same period.
+Upon closer examination, this trend can be observed across all of the three largest publishers, although the shift towards BRICS is particularly evident in Elsevier's hybrid journal portfolio, particularly with regard to articles published in Physical Sciences journals.
+While OECD publication output in Elsevier's Physical Sciences journals declined from `r publ_subject_output_df |> filter(cr_year == 2018, top_level == "Physical Sciences", publisher == "Elsevier", cat == "OECD") |> pull(all)` articles in 2018 to `r publ_subject_output_df |> filter(cr_year == 2022, top_level == "Physical Sciences", publisher == "Elsevier", cat == "OECD") |> pull(all)` in 2022, BRICS output increased from `r publ_subject_output_df |> filter(cr_year == 2018, top_level == "Physical Sciences", publisher == "Elsevier", cat == "BRICS") |> pull(all)` to `r publ_subject_output_df |> filter(cr_year == 2022, top_level == "Physical Sciences", publisher == "Elsevier", cat == "BRICS") |> pull(all)` in the same five-year period.
+Furthermore, OECD publication output in Health Sciences and Life Sciences journals stagnated after peaking in 2020.
+
 To illustrate the situation in 2022, Figure \ref{fig:country_patch}B compares the total publication output with the number of open access articles. 

### 2024-09-13 analysis/manuscript.Rmd
Minor

@@
 For this study, country and institutional affiliations were retrieved from OpenAlex.
-However, because corresponding author affiliation was only available for 54% of the investigated articles in OpenAlex, this study focused on first authors and their affiliations instead; approximately 90% of the investigated articles had first author affiliation metadata in OpenAlex, which is a much larger proportion than previously reported for the October 2022 snapshot [@Zhang_2024].
-First authors typically contribute the most to a paper and are often considered being lead authors [@Larivi_re_2016]. 
-Related studies also assumed first authors as a proxy to examine open access payments and transformative agreements [@Zhang_2022; @Shu_2023; @Haucap_2021].
+Because the corresponding authorship field was not fully supported by OpenAlex at the time of analysis, and affiliations were only available for 54% of the investigated articles, this study focused on first authors and their affiliations instead; approximately 90% of the investigated articles had first author affiliation metadata in OpenAlex, which is a much larger proportion than previously reported for the October 2022 snapshot [@Zhang_2024].
+First authors typically contribute the most to a paper and are often considered lead authors [@Larivi_re_2016].
+Examining Web of Science and Scopus, @Chinchilla_Rodr_guez_2024 revealed that the corresponding authors were typically the first authors listed. 
+However, notable differences were observed between the two databases.
+Related studies also utilised first authors as a proxy to examine open access payments and transformative agreements [@Zhang_2022; @Shu_2023; @Haucap_2021].
 
@@
   theme_minimal() +
-  labs(x = "Publication year", y = "Hybrid OA Uptake")
+  labs(x = "Publication year", y = "% of total articles\nin hybrid journals")
 ```
@@
 However, several limitations need to be acknowledged.
-From a data perspective, estimations of open access through transformative agreements were established by linking first author affiliations with publicly available agreement data from cOAlition S, and not through invoicing data, which is usually not shared.
+From a data perspective, estimations of open access through transformative agreements were established by linking author affiliations with publicly available agreement data from cOAlition S, and not through invoicing data, which is usually not shared.
+Because data on corresponding authors, who are typically responsible for facilitating open access publication, were not fully supported by OpenAlex at the time of this study, first author affiliations were used instead.
+
 Moreover, this study is unable to account for the various types of transformative agreements, particularly with regard to article types and caps that limit the number of open access articles covered.
-Furthermore, assessing the quality of the OpenAlex and Crossref data used, particularly in terms of affiliations and article types, by combining them with established bibliometric databases such as Scopus and Web of Science was beyond the scope of this analysis [@Visser_2021].
+Furthermore, assessing the quality of the OpenAlex and Crossref data used, particularly in terms of affiliations, corresponding authorship, and article types, by combining them with established bibliometric databases such as Scopus and Web of Science was beyond the scope of this analysis [@Visser_2021; @Chinchilla_Rodr_guez_2024].
 Overall, the methodology is designed to underestimate, rather than overestimate, the adoption of open access through transformative agreements. 
@@
 To improve the assessment of transformative agreements, libraries and publishers need to collaborate on standards and services to publicly share information about respective journal portfolios, participating institutions and open access invoicing.
+Such collaboration would help overcome the limitations of current approaches that derive funding for open access from authorship data.
 

### 2024-09-13 analysis/manuscript.Rmd
Reviewer 1: Explain why first author affiliations were chosen and not corresponding author data provided by OpenAlex

@@
 For this study, country and institutional affiliations were retrieved from OpenAlex.
-Because the corresponding authorship field was not fully supported by OpenAlex at the time of analysis, and affiliations were only available for 54% of the investigated articles, this study focused on first authors and their affiliations instead; approximately 90% of the investigated articles had first author affiliation metadata in OpenAlex, which is a much larger proportion than previously reported for the October 2022 snapshot [@Zhang_2024].
+However, because the corresponding authorship field was not fully supported by OpenAlex at the time of analysis, and affiliations were only available for 54% of the investigated articles, this study focused on first authors and their affiliations instead; approximately 90% of the investigated articles had first author affiliation metadata in OpenAlex, which is a much larger proportion than previously reported for the October 2022 snapshot [@Zhang_2024].
 First authors typically contribute the most to a paper and are often considered lead authors [@Larivi_re_2016].

### 2024-09-13 analysis/manuscript.Rmd
Reviewer 2: Start adding helpful policy background information

--- 2024-09-13 analysis/manuscript.Rmd
@@
 The controversies surrounding hybrid open access and transformative agreements have led to varying policy conclusions. 
-For instance, the Association of Swedish Higher Education Institutions (Sveriges universitets- och
-högskoleförbund, SUHF) recommended only supporting agreements for publishing in fully open access journals by 2026 [^sweden].
-Likewise, most cOAlition S funders will discontinue financial support for transformative agreements by the end of 2024 [@Liverpool_2023].
+For instance, the UK's Joint Information Systems Committee (JISC) concluded from an evaluation of its transformative agreements [@Jisc_2024] that while these agreements had a significant impact on the growth of open access in the UK, they had limited effect in facilitating a global shift towards full open access.
+Consequently, the report argues for a reassessment of the UK's open access strategy, in which hybrid OA and transformative agreements have played an important role.
+After making similar observation, Norwegian and Swedish universities and their consortia also revised their policies [@norway; @Widding_2024]. 
+As a result, the Association of Swedish Higher Education Institutions (Sveriges universitets- och högskoleförbund, SUHF) recommended only supporting agreements for publishing in fully open access journals by 2026 [@Widding_2024].
+Likewise, cOAlition S will stop supporting transformative agreements at the end of 2024 [^coalitions].
 cOAlition S also removed the majority of hybrid journals from its Transformative Journal program in 2023 due to publishers' failure to meet self-defined open access growth targets [@Brainard_2023]. 
@@
 Despite these controversies around transformative agreements as a means of transitioning journal publishing to full open access, there is limited evidence available on the uptake of open access in hybrid journals, and the extent to which this can be attributed to transformative agreements.
-Previous studies have focused on specific countries [@P_l_nen_2020; @Huang_2020; @Haucap_2021; @Wenaas_2022; @Taubert_2023; @Jisc_2024] or publisher portfolios [@Pieper_2018; @Bakker_2024; @Fraser_2023; @Jahn_2021; @Momeni_2023; @Schmal_2024], while large-scale studies relied on self-reported agreement data [@Moskovkin_2022], or used APC pricing lists [@Shu_2023].
+Previous studies have focused on specific countries [@P_l_nen_2020; @Huang_2020; @Haucap_2021; @Wenaas_2022; @ludo_waltman_2022_7105355; @Taubert_2023; @Jisc_2024] or publisher portfolios [@Pieper_2018; @Bakker_2024; @Fraser_2023; @Jahn_2021; @Momeni_2023; @Schmal_2024], while large-scale studies relied on self-reported agreement data [@Moskovkin_2022], or used APC pricing lists [@Shu_2023].
 In particular, data availability is a limiting factor when studying the impact of transformative agreements [@Bakker_2024], because bibliometric databases, even though many allow the retrieval of open access articles in hybrid journals, do not directly attribute them to specific transformative agreements.
@@
 
-[^sweden]: <https://www.su.se/english/news/open-access-need-to-move-away-from-transformative-agreements-1.683787>
+[^coalitions]: <https://www.coalition-s.org/coalition-s-confirms-the-end-of-its-financial-support-for-open-access-publishing-under-transformative-arrangements-after-2024/>
 

### 2024-09-13 analysis/manuscript.Rmd

@@
 Because the corresponding authorship field was not fully supported by OpenAlex at the time of analysis, and affiliations were only available for 54% of the investigated articles, this study focused on first authors and their affiliations instead; approximately 90% of the investigated articles had first author affiliation metadata in OpenAlex, which is a much larger proportion than previously reported for the October 2022 snapshot [@Zhang_2024].
-First authors typically contribute the most to a paper and are often considered lead authors [@Larivi_re_2016].
-Examining Web of Science and Scopus, @Chinchilla_Rodr_guez_2024 revealed that the corresponding authors were typically the first authors listed. 
-However, notable differences were observed between the two databases.
-Related studies also utilised first authors as a proxy to examine open access payments and transformative agreements [@Zhang_2022; @Shu_2023; @Haucap_2021].
+First authors typically contribute the most to a paper and are often considered lead authors [@Larivi_re_2016], although notable differences in corresponding author metadata were observed across subjects, publication years and bibliometric databases [@Chinchilla_Rodr_guez_2024].
+In case of lacking data about corresponding authors in bibliometric databases, related studies also utilised first authors as a proxy to examine open access payments and transformative agreements [@Zhang_2022; @Shu_2023; @Haucap_2021].
 

### 2024-09-17 analysis/manuscript.Rmd
Reviever 2: Make limitations of data sources more transparent, provide a frank assessment about TAs and link to Barcelona Declaration.

@@
 The controversies surrounding hybrid open access and transformative agreements have led to varying policy conclusions. 
-For instance, the UK's Joint Information Systems Committee (JISC) concluded from an evaluation of its transformative agreements [@Jisc_2024] that while these agreements had a significant impact on the growth of open access in the UK, they had limited effect in facilitating a global shift towards full open access.
-Consequently, the report argues for a reassessment of the UK's open access strategy, in which hybrid OA and transformative agreements have played an important role.
-After making similar observation, Norwegian and Swedish universities and their consortia also revised their policies [@norway; @Widding_2024]. 
-As a result, the Association of Swedish Higher Education Institutions (Sveriges universitets- och högskoleförbund, SUHF) recommended only supporting agreements for publishing in fully open access journals by 2026 [@Widding_2024].
-Likewise, cOAlition S will stop supporting transformative agreements at the end of 2024 [^coalitions].
+For instance, the UK's Joint Information Systems Committee (JISC) evaluated its open access strategy, which included transformative agreements [@Jisc_2024]. 
+The evaluation revealed that while these agreements had a significant impact on the growth of open access in the UK, they had limited effects in facilitating a global shift towards full open access. 
+As a result, the report advised that the UK's open access strategy should be reassessed.
+After making similar observation, Norwegian and Swedish universities and their consortia also argued for policy changes [@norway; @Widding_2024]. 
+Notably, the Association of Swedish Higher Education Institutions (Sveriges universitets- och högskoleförbund, SUHF) recommended only supporting agreements for publishing in fully open access journals by 2026 [@Widding_2024].
+Furthermore, cOAlition S will conclude its financial support of transformative agreements at the end of 2024, but will continue to view open access resulting from such agreements as compliant.
+[^coalitions].
 cOAlition S also removed the majority of hybrid journals from its Transformative Journal program in 2023 due to publishers' failure to meet self-defined open access growth targets [@Brainard_2023]. 
@@
 In particular, data availability is a limiting factor when studying the impact of transformative agreements [@Bakker_2024], because bibliometric databases, even though many allow the retrieval of open access articles in hybrid journals, do not directly attribute them to specific transformative agreements.
-Likewise, article-level open access invoicing and cost data are only partly available [@Jahn_2021; @Kramer_2024].
+Likewise, article-level open access invoicing and cost data are only partly available, which would make it possible to establish a direct link between transformative agreements and open access publications [@Jahn_2021; @Kramer_2024].
 
@@
 After identifying hybrid journals by excluding fully open access journals, Crossref [@Hendricks_2020] served as the primary data source for article-level metadata including Creative Commons (CC) license information to indicate open access availability on publisher websites.
-To determine open access articles published through transformative agreements, first author affiliations from OpenAlex [@priem2022openalex] were subsequently linked to eligible institutions according to the transformative agreement data.
+Because of a lack of comprehensive publicly available invoicing data, open access articles published through transformative agreements were determined by first author affiliations from OpenAlex [@priem2022openalex], linked to eligible institutions according to the transformative agreement data.
 In the following, the steps are described in detail.
@@
 
-Data gathering started with obtaining journals included in transformative agreements from the publicly available Transformative Agreement Data dump[^tadump] used by the cOAlition S Journal Checker Tool.[^checkertool]
-The dump consists of multiple online Google spreadsheets where each data file represents one agreement listed in the ESAC Transformative Agreement Registry.[^esacregistry]
+Data gathering started with obtaining journals included in transformative agreements from the publicly available Transformative Agreement Data dump[^tadump] used by the cOAlition S Journal Checker Tool [^checkertool], a voluntary effort based on publicly disclosed contracts in the ESAC Transformative Agreement Registry.[^esacregistry].
+The dump consists of multiple online Google spreadsheets where each data file represents one agreement listed in the ESAC Transformative Agreement Registry.
 From the retrieved spreadsheet files, journals and institutions involved per agreement were obtained.
 
-A limitation of using the Journal Checker Tool and its underlying publicly available data dump to study the development of transformative agreements over time is that expired transformative agreements are constantly removed.
+It should be noted that although many library consortia see the need to register their agreements through the ESAC registry, many fail to disclose all details, including contracts [@Bakker_2024; @Kramer_2024].
+Furthermore, the extent to which ESAC is comprehensive remains uncertain, potentially limiting the coverage of transformative agreements in the Journal Checker Tool. 
+Another limitation of using the Journal Checker Tool and its underlying publicly available data dump to study the development of transformative agreements over time is that expired transformative agreements are constantly removed.
 To address this, four different snapshots were safeguarded and combined for this study: self-archived versions from July 2021, July 2022, and May 2023, as well as the most current dump downloaded on 11 December 2023.
@@
 For this study, country and institutional affiliations were retrieved from OpenAlex.
-Because the corresponding authorship field was not fully supported by OpenAlex at the time of analysis, and affiliations were only available for 54% of the investigated articles, this study focused on first authors and their affiliations instead; approximately 90% of the investigated articles had first author affiliation metadata in OpenAlex, which is a much larger proportion than previously reported for the October 2022 snapshot [@Zhang_2024].
+Because the corresponding authorship field was not fully supported by OpenAlex at the time of analysis, and respective affiliation data were only available for 54% of the investigated articles, this study focused on first authors and their affiliations instead; approximately 90% of the articles examined had first author affiliation metadata in OpenAlex, which is a much larger proportion than previously reported for the October 2022 snapshot [@Zhang_2024].
 First authors typically contribute the most to a paper and are often considered lead authors [@Larivi_re_2016], although notable differences in corresponding author metadata were observed across subjects, publication years and bibliometric databases [@Chinchilla_Rodr_guez_2024].
@@
   scale_fill_manual(NULL, values = c("OECD" = "#d01c8b", "BRICS" = "#4dac26", "Other" = "#f7f7f7")) +
-  geom_smooth(aes(all, oa / all), method = "lm", se = FALSE) +
+  geom_smooth(aes(all, oa / all), method = "lm", se = TRUE) +
   scale_size("Agreement", guide = FALSE) +
@@
 Because data on corresponding authors, who are typically responsible for facilitating open access publication, were not fully supported by OpenAlex at the time of this study, first author affiliations were used instead.
-
-Moreover, this study is unable to account for the various types of transformative agreements, particularly with regard to article types and caps that limit the number of open access articles covered.
+This study is also unable to account for the various types of transformative agreements due to a lack of data, especially regarding article types and caps limiting the number of open access articles covered.
 Furthermore, assessing the quality of the OpenAlex and Crossref data used, particularly in terms of affiliations, corresponding authorship, and article types, by combining them with established bibliometric databases such as Scopus and Web of Science was beyond the scope of this analysis [@Visser_2021; @Chinchilla_Rodr_guez_2024].
-Overall, the methodology is designed to underestimate, rather than overestimate, the adoption of open access through transformative agreements. 
-In addition, the data used in this study are openly available, along with the code used for this analysis.
+It must therefore be emphasised that, unlike related evaluations of national consortia based on invoice data [@Jisc_2024], this global overview can only provide an estimate of the support for open access through transformative agreements.
+Despite these limitations, the methodology was designed to underestimate, rather than overestimate, the adoption of open access through transformative agreements. 
+To promote transparency, the data used in this study, along with the code used for this analysis, are openly available.
 
@@
 Of primary concern should be the observed differences across countries, particularly the relationship between socio-economic development and open access adoption.
-From a data perspective, the reporting of open access funding, including transformative agreements, is not harmonised, but often crowd-sourced from various sources.
-To improve the assessment of transformative agreements, libraries and publishers need to collaborate on standards and services to publicly share information about respective journal portfolios, participating institutions and open access invoicing.
+In the last years, there was a notable increase in the number of closed access articles published in hybrid journals by authors from countries such as China and India, where transformative agreements and open access funding options were not widely available.
+In contrast, transformative agreements enabled a substantial increase in open access in OECD countries excluding the US, while their share of total articles in hybrid journals decreased. 
+This imbalance, where open access uptake depends largely on a few less productive high-income countries, makes it increasingly unlikely that collective action towards a large-scale transition of hybrid journals to full open access through transformation agreements will be successful.
+Rather, it highlights the importance of ongoing evaluation and adaptation of open access strategies in view of global publishing trends.
+
+From a data perspective, research institutions and funders need to push towards price transparency in scholarly publishing to enable such evaluations. 
+The reporting of open access funding, including transformative agreements, is also not harmonised, but often crowd-sourced from various sources.
+To improve the assessment of transformative agreements, libraries and publishers should collaborate on standards and services to publicly share information about respective journal portfolios, participating institutions and open access invoicing, for example in the international context of ESAC and the Barcelona Declaration on Open Research Information.
 Such collaboration would help overcome the limitations of current approaches that derive funding for open access from authorship data.

### 2024-09-17 analysis/manuscript.Rmd
Minor

@@
 
-From a data perspective, research institutions and funders need to push towards price transparency in scholarly publishing to enable such evaluations. 
+From a data perspective, research institutions and funders need to push towards price transparency in scholarly publishing to enable these evaluations. 
 The reporting of open access funding, including transformative agreements, is also not harmonised, but often crowd-sourced from various sources.

### 2024-09-18 analysis/manuscript.Rmd
Reviewer 2: Acknowledge impact of COVID-related literatue on OECD peak in 2020; and imprive discussion

@@
 
-Despite the rise of open access across OECD countries, the overall publication output decreased sharply, dropping to `r country_bar$data |> filter(cat == "OECD", cr_year == "2022") |> distinct(all) |> pull(all)` in 2022 after peaking `r country_bar$data |> filter(cat == "OECD", cr_year == "2020") |> distinct(all) |> pull(all)` articles in 2020.
+Despite the rise of open access across OECD countries, the overall publication output decreased sharply, dropping to `r country_bar$data |> filter(cat == "OECD", cr_year == "2022") |> distinct(all) |> pull(all)` in 2022 after peaking `r country_bar$data |> filter(cat == "OECD", cr_year == "2020") |> distinct(all) |> pull(all)` articles in 2020, which saw a massive growth in literature related to COVID-19, driven largely by researchers in the United States surpassing China and other countries [@Ioannidis_2021].
 In stark contrast, the number of articles published in hybrid journals by first authors affiliated with institutions from BRICS countries increased steadily over the years, from `r country_bar$data |> filter(cat == "BRICS", cr_year == "2018") |> distinct(all) |> pull(all)` in 2018 to `r country_bar$data |> filter(cat == "BRICS", cr_year == "2022") |> distinct(all) |> pull(all)` in 2022.
@@
 While OECD publication output in Elsevier's Physical Sciences journals declined from `r publ_subject_output_df |> filter(cr_year == 2018, top_level == "Physical Sciences", publisher == "Elsevier", cat == "OECD") |> pull(all)` articles in 2018 to `r publ_subject_output_df |> filter(cr_year == 2022, top_level == "Physical Sciences", publisher == "Elsevier", cat == "OECD") |> pull(all)` in 2022, BRICS output increased from `r publ_subject_output_df |> filter(cr_year == 2018, top_level == "Physical Sciences", publisher == "Elsevier", cat == "BRICS") |> pull(all)` to `r publ_subject_output_df |> filter(cr_year == 2022, top_level == "Physical Sciences", publisher == "Elsevier", cat == "BRICS") |> pull(all)` in the same five-year period.
-Furthermore, OECD publication output in Health Sciences and Life Sciences journals stagnated after peaking in 2020.
+Furthermore, OECD publication output in Health Sciences and Life Sciences journals stagnated after peaking in 2020, which saw an increase in publications due to the impact of the global COVID-19 pandemic, in particular from the United States [@Ioannidis_2021].
 
@@
 Of primary concern should be the observed differences across countries, particularly the relationship between socio-economic development and open access adoption.
-In the last years, there was a notable increase in the number of closed access articles published in hybrid journals by authors from countries such as China and India, where transformative agreements and open access funding options were not widely available.
-In contrast, transformative agreements enabled a substantial increase in open access in OECD countries excluding the US, while their share of total articles in hybrid journals decreased. 
-This imbalance, where open access uptake depends largely on a few less productive high-income countries, makes it increasingly unlikely that collective action towards a large-scale transition of hybrid journals to full open access through transformation agreements will be successful.
-Rather, it highlights the importance of ongoing evaluation and adaptation of open access strategies in view of global publishing trends.
+Between 2018 and 2022, there was a notable increase in the number of closed access articles published in hybrid journals by authors from countries such as China and India, where transformative agreements and open access funding options were not widely available.
+In contrast, the introduction of country-wide transformative agreements in numerous OECD countries led to a substantial increase in open access, while their share of total articles in hybrid journals decreased.
+This imbalance, whereby open access uptake depends largely on a few countries, makes it less likely that collective action towards a large-scale transition of hybrid journals to full open access through transformation agreements will succeed within the next years.
+Rather, and in accordance with recent policy recommendations from the UK and Nordic countries [@Jisc_2024; @Widding_2024], it emphasises the importance of ongoing evaluation and adaptation of open access strategies in view of global publishing trends.
 

### 2024-09-18 analysis/manuscript.Rmd
Minor fixes and recompilation of manuscript

@@
   theme_minimal() +
-  labs(x = "Publication year", y = "Total articles") +
+  labs(x = "Publication Year", y = "Total Articles") +
   guides(fill = guide_legend(reverse = TRUE)) +
@@
 
-The so compiled data set consists of `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l, cr_year, jn_all) |> pull(jn_all) |> sum()` articles published in `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l) |> nrow()` hybrid journals included in at least one transformative agreement between 2018 and 2022 (see purple bars in Figure \ref{fig:method_fig}B).
-These hybrid journals in transformative agreements represented `r cr_stats |> filter(src == "hoad") |> pull(n_articles) |> sum() / cr_stats |> pull(n_articles) |> sum() * 100`%  of the total global output over the same time period, according to Crossref, while fully open access journals recorded
-`r cr_stats |> filter(is.na(src) & oa_status == "gold") |> pull(n_articles) |> sum() / cr_stats |> pull(n_articles) |> sum() * 100`%.
-
 ```{r method_fig, fig.asp=0.55, fig.cap="Initial data characteristics.
@@
 ```
+
+The so compiled data set consists of `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l, cr_year, jn_all) |> pull(jn_all) |> sum()` articles published in `r hoaddata::jn_ind |> filter(issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |> distinct(issn_l) |> nrow()` hybrid journals included in at least one transformative agreement between 2018 and 2022 (see purple bars in Figure \ref{fig:method_fig}B).
+These hybrid journals in transformative agreements represented `r cr_stats |> filter(src == "hoad") |> pull(n_articles) |> sum() / cr_stats |> pull(n_articles) |> sum() * 100`%  of the total global output over the same time period, according to Crossref, while fully open access journals recorded
+`r cr_stats |> filter(is.na(src) & oa_status == "gold") |> pull(n_articles) |> sum() / cr_stats |> pull(n_articles) |> sum() * 100`%.
+
 <!-- To reflect the increasing trend towards multiple and cross-country affiliations (Hottenrott et al., 2021; Larivière et al., 2015), articles were counted once per country. -->
@@
     "Articles" = 2,
-    "OA articles" = 2,
-    "TA OA articles" = 2
+    "OA Articles" = 2,
+    "TA OA Articles" = 2
   ))
@@
 Figure \ref{fig:publisher_figure} takes a closer look into the growth of hybrid open access across publishers by year, with a focus on open access enabled by transformative agreements. 
-Although all publishers show a general long-term trend towards transformative agreements, Figure \ref{fig:publisher_figure}A and B indicate that Wiley experienced a substantial increase in its open access share from `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(value) |> sum()`) in 2018 to `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(prop) |> sum()*100`%  (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()`) in 2022, representing a 4.5-fold increase.
+Although all publishers show a general long-term trend towards transformative agreements, Figure \ref{fig:publisher_figure}A and B indicate that Wiley experienced a substantial increase in its open access share from `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Wiley") |> pull(value) |> sum()`) in 2018 to `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(prop) |> sum()*100`%  (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()`) in 2022.
 In contrast, Elsevier's hybrid journals demonstrated a more modest increase, from `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Elsevier") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2018", publisher == "Elsevier") |> pull(value) |> sum()`) in 2018 to `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(prop) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(value) |> sum()`) in 2022, which is a relatively low open access share compared to the general trend.
@@
 
-```{r publisher_figure, fig.asp=0.8, fig.cap="Developement of open access in hybrid journals included in transformative agreements between 2018 and 2022 by publishers.
+```{r publisher_figure, fig.asp=0.8, fig.cap="Development of open access in hybrid journals included in transformative agreements between 2018 and 2022 by publishers.
 The blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established.
@@
 
-Table \ref{tab:subject_summary_table} presents a high-level overview of hybrid open access by AJCS subject area using fractionalised counting to account for journals belonging to more than one category.
+Table \ref{tab:subject_summary_table} presents a high-level overview of hybrid open access by ASJC subject area using fractional counting to account for journals belonging to more than one category. 
 Between 2018 and 2022, most hybrid journals with at least one open access article could be attributed to the Social Sciences category, which also includes Arts and Humanities.
@@
 
-```{r subject_panel, fig.asp=0.8, fig.cap="Developement of open access in hybrid journals in transformative agreements between 2018 and 2022 by AJCS subject area.
+```{r subject_panel, fig.asp=0.8, fig.cap="Development of open access in hybrid journals in transformative agreements between 2018 and 2022 by ASJC subject area.
 The blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established.
-(A) Proportion of open access articles in hybrid journals per year and publisher.
-(B) Percentage of hybrid open access via agreements per year and publisher.
-Boxplots (C) show individual open access uptake rates by individual hybrid journals and open access funding per publication year and publisher.
-The individual outliers are not shown.
+(A) Proportion of open access articles in hybrid journals per year and subject area (B) Percentage of hybrid open access via agreements per year and subject area Boxplots (C) show individual open access uptake rates by individual hybrid journals and open access funding per publication year and subject area The individual outliers are not shown.
 Note that data on transformative agreements ending before June 2021 were not available for this study."}
@@
   theme_minimal() +
-  labs(x = "Total articles in hybrid journals in 2022\n(log-scale)", y = "OA Uptake")
+  labs(x = "Total Articles in Hybrid Journals in 2022\n(log-scale)", y = "OA Uptake")
 
@@
   theme_minimal() +
-  labs(x = "Publication year", y = "% of total articles\nin hybrid journals")
+  labs(x = "Publication Year", y = "% of Total Articles\nin Hybrid Journals")
 ```
@@
   theme_minimal() +
-  labs(x = "Publication year", y = "Total articles")
+  labs(x = "Publication Year", y = "Total Articles")
 
@@
    theme_minimal() +
-  labs(x = "Publication year", y = "Hybrid OA Uptake")
+  labs(x = "Publication Year", y = "Hybrid OA Uptake")
 
@@
    theme_minimal() +
-  labs(x = "Publication year", y = "Total articles")
+  labs(x = "Publication Year", y = "Total Articles")
 ```
@@
 
+```{r country_top_20_plot, fig.asp=0.8, fig.cap="Development of open access in hybrid journals in transformative agreements between 2018 and 2022, by the Top 20 most productive countries in terms of total articles published in the five-years period.
+Blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established.
+Country names are represented as ISO two-letter country codes.
+Facet subheadings show the total number of articles and corresponding market share."}
+country_plot
+```
+
 Figure \ref{fig:country_top_20_plot} illustrates the development of hybrid open access from 2018 to 2022, highlighting the top 20 most productive countries in terms of articles published in hybrid journals that were included in transformative agreements over the five-year period.
@@
 Prior to this, only a few organisations had agreements in place, for example, the Max Planck Society with Springer Compact.
-
-```{r country_top_20_plot, fig.asp=0.8, fig.cap="Development of open access in hybrid journals in transformative agreements between 2018 and 2022, by the Top 20 most productive countries in terms of total articles published in the five-years period.
-Blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established.
-Country names are represented as ISO two-letter country codes.
-Facet subheadings show the total number of articles and corresponding market share."}
-country_plot
-```
 
@@
 Although China supports Plan S [@Schiermeier_2018], this is not reflected in the data in terms of open access uptake in hybrid journals.
+Market observers also do not expect a broad implementation of transformative agreements in China in near future [@Owens_2024].
 As highlighted by @Koley_2023, India faces similar challenges in terms of the availablility of resources to pay for open access.

### 2024-09-18 analysis/manuscript.tex
Recompile tex
Minor fixes and recompilation of manuscript

@@
 % bibliography
+% definitions for citeproc citations
+\NewDocumentCommand\citeproctext{}{}
+\NewDocumentCommand\citeproc{mm}{%
+  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
+\makeatletter
+ % allow citations to break across lines
+ \let\@cite@ofmt\@firstofone
+ % avoid brackets around text for \cite:
+ \def\@biblabel#1{}
+ \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
+\makeatother
 \newlength{\cslhangindent}
 \setlength{\cslhangindent}{1.5em}
-\newenvironment{CSLReferences}%
-  {}%
-  {\par}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
+  {\begin{list}{}{%
+   \setlength{\itemindent}{0pt}
+   \setlength{\leftmargin}{0pt}
+   \setlength{\parsep}{0pt}
+   % turn on hanging indent if param 1 is 1
+   \ifodd #1
+    \setlength{\leftmargin}{\cslhangindent}
+    \setlength{\itemindent}{-1\cslhangindent}
+   \fi
+   % set entry spacing
+   \setlength{\itemsep}{#2\baselineskip}}}
+  {\end{list}}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 
-\abstract{The ongoing controversy surrounding transformative agreements, which aim to transition journal publishing to full open access, highlight the need for large-scale studies assessing the impact of these agreements on hybrid open access.
-By combining publicly available data from various sources, including cOAlition S Journal Checker, Crossref, and OpenAlex, this study presents a novel approach that analyses over 700 agreements.
-Results suggest a strong growth in open access between 2018 and 2022 from 4.3\% to 15\%.
-During this period, 11,189 hybrid journals provided open access to 742,369 out of 8,146,958 articles, representing a five-year open access proportion of 9.1\%.
-Authors who could make use of transformative agreements at the time of publication contributed 328,957 open access articles.
-In 2022, 143,615 out of 249,511 open access articles in hybrid journals (58\%) were enabled by transformative agreements.
-This trend was largely driven by the three commercial publishers Elsevier, Springer Nature, and Wiley, but the open access uptake varied substantially across journals, publishers, disciplines, and country affiliations.
-In particular, the OECD and BRICS areas revealed different publication trends.
+\abstract{The ongoing controversy surrounding transformative agreements, which aim to transition subscription-based journal publishing to full open access, highlights the need for large-scale studies assessing the impact of these agreements on hybrid open access.
+By combining multiple open data sources, including cOAlition S Journal Checker, Crossref, and OpenAlex, this study presents a novel approach that analyses over 700 agreements.
+Results suggest a strong growth in open access, from 4.3\% in 2018 to 15\% in 2022.
+Over five years, 11,189 hybrid journals provided open access to 742,369 out of 8,146,958 articles (9.1\%).
+Authors who could make use of transformative agreements contributed 328,957 open access articles (44\%) during this period, reaching a peak in 2022 with 143,615 out of 249,511 open access articles (58\%).
+While this trend was predominantly driven by the three commercial publishers Elsevier, Springer Nature, and Wiley, open access uptake varied substantially across journals, publishers, disciplines, and countries.
+Particularly, the OECD and BRICS areas revealed different publication trends.
 In conclusion, this study suggests that current levels of implementation of transformative agreements is insufficient to bring about a large-scale transition to full open access.}
@@
 
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
+\section{Introduction}\label{introduction}
 
@@
 in which well-established commercial publishers gain twice from reading
-and open access publishing (Mittermaier, 2015), as well as to increase
-publisher-provided immediate open access (Björk \& Solomon, 2014;
-Schimmer et al., 2015). These considerations resulted in transformative
-agreements\footnote{In this paper I use the term ``transformative
-  agreement'', addressing also offsetting, read-and-publish or
-  publish-and-read deals, and other variants (Borrego et al., 2021;
-  Hinchliffe, 2019). Although the term is critised as misleading and not
-  useful to describe the different types of open access agreements
-  between library consortia and commercial publishers (Babini et al.,
-  2022), it is widely used in policy discussions and in the research
-  literature.}, which cover a broad range of contracts between library
-consortia and publishers from the mid-2010s onwards, where institutional
-spending for subscriptions and open access publishing are considered
-together (Borrego et al., 2021; Hinchliffe, 2019). Transformative
-agreements seek to control costs while allowing a transitional phase for
-publishing more open access articles. Similar to big deals,
-transformative agreements mainly bundle hybrid and subscription-only
-journals from commercial publishers, but aim at a higher degree of
-transparency than previous big deals, where contracts including payments
-were confidential (Bergstrom et al., 2014).
+and open access publishing fees (Mittermaier, 2015), as well as to
+increase publisher-provided immediate open access (Björk \& Solomon,
+2014; Schimmer et al., 2015). These considerations resulted in
+transformative agreements\footnote{In this paper I use the term
+  ``transformative agreement'', addressing also offsetting,
+  read-and-publish or publish-and-read deals, and other variants
+  (Borrego et al., 2021; Hinchliffe, 2019). Although the term is
+  critised as misleading and not useful to describe the different types
+  of open access agreements between library consortia and commercial
+  publishers (Babini et al., 2022), it is widely used in policy
+  discussions and in the research literature.}, which cover a broad
+range of contracts between library consortia and publishers from the
+mid-2010s onwards, where institutional spending for subscriptions and
+open access publishing are considered together (Borrego et al., 2021;
+Hinchliffe, 2019). Transformative agreements seek to control costs while
+allowing a transitional phase for publishing more open access articles.
+Similar to big deals, transformative agreements mainly bundle hybrid and
+subscription-only journals from commercial publishers, but aim at a
+higher degree of transparency than previous big deals, where contracts
+including payments were confidential (Bergstrom et al., 2014).
 
@@
 Registry\footnote{\url{https://esac-initiative.org/about/transformative-agreements/agreement-registry/}},
-the largest resource for library consortia to publicly share their
-agreements, recorded more than 800 transformative agreements. These
-agreements resulted in the publication of up to 900.000 open access
-articles published in both fully open access and hybrid journals,
-according to the accompanying ESAC Market Watch\footnote{\url{https://esac-initiative.org/market-watch/}}.
-Library consortia reported increased open access volume, streamlined
-payment and monitoring procedures, as well as extensive utilisation of
-open access options by the researchers they serve (Marques \& Stone,
-2020; Parmhed \& Säll, 2023; Pinhasi et al., 2020). The ongoing
-standardisation of transformative agreements has contributed to improved
-transparency in terms of contracts and publisher-provided article
-metadata (Marques et al., 2019; Pinhasi et al., 2021). However, with the
-growing trend toward transformative agreements, continued reliance on
-big deals is perceived as problematic, because it perpetuates market
-concentration (Butler et al., 2023; Shu \& Larivière, 2023). Whether
-transformative agreements lead to reduced pricing remains uncertain
-(Borrego, 2023) and a substantial transition of hybrid journals towards
-full open access could not be observed (Matthias et al., 2019; Momeni et
-al., 2021). The focus on large commercial publishers might also increase
-inequality (Ross-Hellauer et al., 2022), because transformative
-agreements' focus on pay to publish mainly targets institutions from
-high-income countries, furthering a questionable journal prestige
-culture (Babini et al., 2022). Besides, an editorial-board resignation
-raised concerns that transformative agreements might encourage
-publishers to maximize journal publication volume ``without regard to
-quality'' (Rasmussen, 2023).
+the largest resource for library consortia to voluntarily and publicly
+share their agreements, recorded more than 800 transformative
+agreements. These agreements resulted in the publication of up to
+900.000 open access articles published in both fully open access and
+hybrid journals, according to the accompanying ESAC Market
+Watch\footnote{\url{https://esac-initiative.org/market-watch/}}. Library
+consortia reported increased open access volume, streamlined payment and
+monitoring procedures, as well as extensive utilisation of open access
+options by the researchers they serve (Marques \& Stone, 2020; Parmhed
+\& Säll, 2023; Pinhasi et al., 2020). The ongoing standardisation of
+transformative agreements has contributed to improved transparency in
+terms of contracts and publisher-provided article metadata (Marques et
+al., 2019; Pinhasi et al., 2021). However, with the growing trend toward
+transformative agreements, continued reliance on big deals is perceived
+as problematic, because it perpetuates market concentration (Butler et
+al., 2023; Shu \& Larivière, 2023). Whether transformative agreements
+lead to reduced pricing remains uncertain (Borrego, 2023) and a
+substantial transition of hybrid journals towards full open access could
+not be observed (Matthias et al., 2019; Momeni et al., 2021). The focus
+on large commercial publishers might also increase inequality
+(Ross-Hellauer et al., 2022), because transformative agreements' focus
+on pay to publish mainly targets institutions from high-income
+countries, furthering a questionable journal prestige culture (Babini et
+al., 2022). Besides, an editorial-board resignation raised concerns that
+transformative agreements might encourage publishers to maximize journal
+publication volume ``without regard to quality'' (Rasmussen, 2023).
 
@@
 agreements have led to varying policy conclusions. For instance, the
-Association of Swedish Higher Education Institutions (Sveriges
-universitets- och högskoleförbund, SUHF) recommended only supporting
-agreements for publishing in fully open access journals by 2026
-\footnote{\url{https://www.su.se/english/news/open-access-need-to-move-away-from-transformative-agreements-1.683787}}.
-Likewise, most cOAlition S funders will discontinue financial support
-for transformative agreements by the end of 2024 (Liverpool, 2023).
+UK's Joint Information Systems Committee (JISC) evaluated its open
+access strategy, which included transformative agreements (Brayman et
+al., 2024). The evaluation revealed that while these agreements had a
+significant impact on the growth of open access in the UK, they had
+limited effects in facilitating a global shift towards full open access.
+As a result, the report advised that the UK's open access strategy
+should be reassessed. After making similar observation, Norwegian and
+Swedish universities and their consortia also argued for policy changes
+(Holden et al., 2023; Widding, 2024). Notably, the Association of
+Swedish Higher Education Institutions (Sveriges universitets- och
+högskoleförbund, SUHF) recommended only supporting agreements for
+publishing in fully open access journals by 2026 (Widding, 2024).
+Furthermore, cOAlition S will conclude its financial support of
+transformative agreements at the end of 2024, but will continue to view
+open access resulting from such agreements as compliant.
+\footnote{\url{https://www.coalition-s.org/coalition-s-confirms-the-end-of-its-financial-support-for-open-access-publishing-under-transformative-arrangements-after-2024/}}.
 cOAlition S also removed the majority of hybrid journals from its
@@
 countries (Brayman et al., 2024; Haucap et al., 2021; Huang et al.,
-2020; Pölönen et al., 2020; Taubert et al., 2023; Wenaas, 2022) or
-publisher portfolios (Bakker et al., 2024; Fraser et al., 2023; Jahn et
-al., 2022; Momeni et al., 2023; Pieper \& Broschinski, 2018; Schmal,
-2024), while large-scale studies relied on self-reported agreement data
-(Moskovkin et al., 2022), or used APC pricing lists (Shu \& Larivière,
-2023). In particular, data availability is a limiting factor when
-studying the impact of transformative agreements (Bakker et al., 2024),
-because bibliometric databases, even though many allow the retrieval of
-open access articles in hybrid journals, do not directly attribute them
-to specific transformative agreements. Likewise, article-level open
-access invoicing and cost data are only partly available (Jahn et al.,
-2022; Kramer, 2024).
+2020; Pölönen et al., 2020; Taubert et al., 2023; Waltman \& Lamers,
+2022; Wenaas, 2022) or publisher portfolios (Bakker et al., 2024; Fraser
+et al., 2023; Jahn et al., 2022; Momeni et al., 2023; Pieper \&
+Broschinski, 2018; Schmal, 2024), while large-scale studies relied on
+self-reported agreement data (Moskovkin et al., 2022), or used APC
+pricing lists (Shu \& Larivière, 2023). In particular, data availability
+is a limiting factor when studying the impact of transformative
+agreements (Bakker et al., 2024), because bibliometric databases, even
+though many allow the retrieval of open access articles in hybrid
+journals, do not directly attribute them to specific transformative
+agreements. Likewise, article-level open access invoicing and cost data
+are only partly available, which would make it possible to establish a
+direct link between transformative agreements and open access
+publications (Jahn et al., 2022; Kramer, 2024).
 
@@
 
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
+\section{Methods}\label{methods}
 
@@
 license information to indicate open access availability on publisher
-websites. To determine open access articles published through
-transformative agreements, first author affiliations from OpenAlex
-(Priem et al., 2022) were subsequently linked to eligible institutions
-according to the transformative agreement data. In the following, the
-steps are described in detail.
+websites. Because of a lack of comprehensive publicly available
+invoicing data, open access articles published through transformative
+agreements were determined by first author affiliations from OpenAlex
+(Priem et al., 2022), linked to eligible institutions according to the
+transformative agreement data. In the following, the steps are described
+in detail.
 
@@
 
-\hypertarget{transformative-agreement-and-hybrid-journal-data}{%
 \subsection{Transformative agreement and hybrid journal
-data}\label{transformative-agreement-and-hybrid-journal-data}}
+data}\label{transformative-agreement-and-hybrid-journal-data}
 
@@
 Agreement Data dump\footnote{\url{https://journalcheckertool.org/transformative-agreements/}}
-used by the cOAlition S Journal Checker Tool.\footnote{\url{https://www.coalition-s.org/blog/enabling-accurate-results-within-the-journal-checker-tool/}}
+used by the cOAlition S Journal Checker Tool \footnote{\url{https://www.coalition-s.org/blog/enabling-accurate-results-within-the-journal-checker-tool/}},
+a voluntary effort based on publicly disclosed contracts in the ESAC
+Transformative Agreement Registry.\footnote{\url{https://esac-initiative.org/about/transformative-agreements/agreement-registry/}}.
 The dump consists of multiple online Google spreadsheets where each data
 file represents one agreement listed in the ESAC Transformative
-Agreement Registry.\footnote{\url{https://esac-initiative.org/about/transformative-agreements/agreement-registry/}}
-From the retrieved spreadsheet files, journals and institutions involved
-per agreement were obtained.
-
-A limitation of using the Journal Checker Tool and its underlying
-publicly available data dump to study the development of transformative
-agreements over time is that expired transformative agreements are
-constantly removed. To address this, four different snapshots were
-safeguarded and combined for this study: self-archived versions from
-July 2021, July 2022, and May 2023, as well as the most current dump
-downloaded on 11 December 2023. This ensured that transformative
-agreements, which ended from 2021 onwards, were included, representing
-the majority of transformative agreements. Overall, the four combined
-Transformative Agreement Data dumps used in this study contained 729 of
-869 agreements listed in the ESAC Transformative Agreement Registry by
-December 2023.
+Agreement Registry. From the retrieved spreadsheet files, journals and
+institutions involved per agreement were obtained.
+
+It should be noted that although many library consortia see the need to
+register their agreements through the ESAC registry, many fail to
+disclose all details, including contracts (Bakker et al., 2024; Kramer,
+2024). Furthermore, the extent to which ESAC is comprehensive remains
+uncertain, potentially limiting the coverage of transformative
+agreements in the Journal Checker Tool. Another limitation of using the
+Journal Checker Tool and its underlying publicly available data dump to
+study the development of transformative agreements over time is that
+expired transformative agreements are constantly removed. To address
+this, four different snapshots were safeguarded and combined for this
+study: self-archived versions from July 2021, July 2022, and May 2023,
+as well as the most current dump downloaded on 11 December 2023. This
+ensured that transformative agreements, which ended from 2021 onwards,
+were included, representing the majority of transformative agreements.
+Overall, the four combined Transformative Agreement Data dumps used in
+this study contained 729 of 869 agreements listed in the ESAC
+Transformative Agreement Registry by December 2023.
 
@@
 
-\hypertarget{article-and-author-metadata}{%
 \subsection{Article and author
-metadata}\label{article-and-author-metadata}}
+metadata}\label{article-and-author-metadata}
 
@@
 Open access articles in hybrid journals were identified using the
-Creative Commons (CC) license information in Crossref metadata. License
-information relative to the ``accepted manuscript (AM)'' version was not
-considered. Crossref was used for open access identification because
-transformative agreement workflows generally require publishers to
-deliver CC license information to this DOI registration agency (Geschuhn
-\& Stone, 2017).
+Creative Commons (CC) license information in Crossref metadata, with
+consideration of any CC variant. License information relative to the
+``accepted manuscript (AM)'' version was not considered. Crossref was
+used for open access identification because transformative agreement
+workflows generally require publishers to deliver CC license information
+to this DOI registration agency (Geschuhn \& Stone, 2017).
 
@@
 study, country and institutional affiliations were retrieved from
-OpenAlex. However, because corresponding author affiliation was only
-available for 54\% of the investigated articles in OpenAlex, this study
-focused on first authors and their affiliations instead; approximately
-90\% of the investigated articles had first author affiliation metadata
-in OpenAlex, which is a much larger proportion than previously reported
-for the October 2022 snapshot (Zhang et al., 2024). First authors
-typically contribute the most to a paper and are often considered being
-lead authors (Larivière et al., 2016). Related studies also assumed
-first authors as a proxy to examine open access payments and
+OpenAlex. Because the corresponding authorship field was not fully
+supported by OpenAlex at the time of analysis, and respective
+affiliation data were only available for 54\% of the investigated
+articles, this study focused on first authors and their affiliations
+instead; approximately 90\% of the articles examined had first author
+affiliation metadata in OpenAlex, which is a much larger proportion than
+previously reported for the October 2022 snapshot (Zhang et al., 2024).
+First authors typically contribute the most to a paper and are often
+considered lead authors (Larivière et al., 2016), although notable
+differences in corresponding author metadata were observed across
+subjects, publication years and bibliometric databases
+(Chinchilla-Rodríguez et al., 2024). In case of lacking data about
+corresponding authors in bibliometric databases, related studies also
+utilised first authors as a proxy to examine open access payments and
 transformative agreements (Haucap et al., 2021; Shu \& Larivière, 2023;
@@
 by OpenAlex using the ROR-ID. Matching also considered the duration of
-agreements according to the ESAC registry. Upon inspection,
-Transformative Agreement Data did not comprehensively cover associated
-institutions, such as university hospitals or institutes of large
-research organisations such as the Max Planck Society. To improve the
-matching, Transformative Agreement Data were automatically enriched with
-associated organisations using OpenAlex's institution entity data.
-
-The so compiled data set consists of 8,922,146 articles published in
-12,857 hybrid journals between 2018 and 2022 (see Figure
-\ref{fig:method_fig}B). Hybrid journals in transformative agreements
-represented 40\% of the total global output over the same time period,
-according to Crossref, while fully open access journals recorded 35\%.
+agreements according to the ESAC registry. 502 agreements were active
+between 2018 and 2022. Upon inspection, Transformative Agreement Data
+did not comprehensively cover associated institutions, such as
+university hospitals or institutes of large research organisations such
+as the Max Planck Society. To improve the matching, Transformative
+Agreement Data were automatically enriched with ROR-IDs from associated
+organisations according to OpenAlex's institution entity data.
 
@@
 (B) Number of articles in Crossref by journal types.
-The blue bars show the overall article volume of hybrid journals in transformative agreements, which were initially included in the study, in comparision with full open access journals according to OpenAlex.
-The remainder represents subscription-based journals not covered by transformative agreements.}\label{fig:method_fig}
+The purple bars show the overall article volume of hybrid journals in transformative agreements, which were initially included in the study, in comparision with full open access journals according to OpenAlex.
+The remainder represents articles in subscription-based journals not covered by transformative agreements.}\label{fig:method_fig}
 \end{figure}
 
-\hypertarget{data-analysis}{%
-\subsection{Data analysis}\label{data-analysis}}
+The so compiled data set consists of 8,922,146 articles published in
+12,857 hybrid journals included in at least one transformative agreement
+between 2018 and 2022 (see purple bars in Figure \ref{fig:method_fig}B).
+These hybrid journals in transformative agreements represented 40\% of
+the total global output over the same time period, according to
+Crossref, while fully open access journals recorded 35\%.
+
+\subsection{Data analysis}\label{data-analysis}
 
@@
 
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-\hypertarget{overview}{%
-\subsection{Overview}\label{overview}}
-
-Between 2018 and 2022, 11,189 out of 12,857 hybrid journals in
+\section{Results}\label{results}
+
+\subsection{Overview}\label{overview}
+
+Between 2018 and 2022, 11,189 out of 12,857 eligible hybrid journals in
 transformative agreements published at least one open access article
-under a Creative Commons license. During this period, these hybrid
-journals provided open access to 742,369 out of 8,146,958 articles,
-representing a five-year open access proportion of 9.1\%. Authors who
-could make use of transformative agreements at the time of publication
-contributed 328,957 open access articles. Overall, 394 transformative
-agreements recorded at least one open access article from an eligible
-institution.
+under a Creative Commons license. These 11,189 hybrid journals
+constituted the foundation for the subsequent analyses. Collectively,
+they provided open access to 742,369 out of 8,146,958 articles during
+the investigated period, representing a five-year open access proportion
+of 9.1\%. Authors who could make use of transformative agreements at the
+time of publication contributed 328,957 (44\%) open access articles.
+Overall, this investigation was able to establish a link between open
+access articles and eligible institutions for 394 out of 502 (78\%)
+active transformative agreements between 2018 and 2022 through the
+analysis of publication metadata, including affiliations and agreement
+data.
 
@@
 on this trend. Over the five-years period from 2018 to 2022, open access
-increased from 4.3\% (n = 65,486) to 15\% (n = 249,511). Simultaneously,
-the total article volume of the investigated journals rose from
-1,528,051 in 2018 to 1,676,928 in 2022.
+in hybrid journals increased from 4.3\% (n = 65,486) to 15\% (n =
+249,511). Simultaneously, the total article volume of the investigated
+journals rose from 1,528,051 in 2018 to 1,676,928 in 2022.
 
@@
 presumably publication fees being not invoiced through transformative
-agreements, from 4.1\% (n = 62,625) in 2018 to 6.3\% (n = 105,896) in
-2022.
+agreements as indicated by the grey area in \ref{fig:results_overview}A,
+from 4.1\% (n = 62,625) in 2018 to 6.3\% (n = 105,896) in 2022.
 
@@
 
-\hypertarget{publishing-market}{%
-\subsection{Publishing market}\label{publishing-market}}
+\subsection{Publishing market}\label{publishing-market}
 
@@
 transformative agreements, the three commercial publishers Elsevier,
-Springer Nature, and Wiley, accounted for 49\% of the total article
-volume published (see Table \ref{tab:publisher_league_table}). Together,
-they published 500,878 or 66\% of the open access articles in hybrid
-journals. Elsevier, Springer Nature, and Wiley made 243,891 articles
-open access in hybrid journals through transformative agreements,
-resulting in an even larger market share of 74\%.
+Springer Nature, and Wiley accounted for 49\% of hybrid journals,
+representing 5,144,308 or 63\% of the total article volume (see Table
+\ref{tab:publisher_league_table}). Together, they published 500,878 or
+66\% of the open access articles in hybrid journals. Elsevier, Springer
+Nature, and Wiley made 243,891 articles open access in hybrid journals
+through transformative agreements, resulting in an even larger market
+share of 74\%.
 
@@
 \toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Hybrid journals} & \multicolumn{2}{c}{Articles} & \multicolumn{2}{c}{OA articles} & \multicolumn{2}{c}{TA OA articles} \\
+\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Hybrid journals} & \multicolumn{2}{c}{Articles} & \multicolumn{2}{c}{OA Articles} & \multicolumn{2}{c}{TA OA Articles} \\
 \cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9}
@@
 substantial increase in its open access share from 5.9\% (n = 11,628) in
-2018 to 26\% (n = 53,503) in 2022, representing a 4.5-fold increase. In
-contrast, Elsevier's hybrid journals demonstrated a more modest
-increase, from 3.3\% (n = 16,872) in 2018 to 10\% (n = 60,821) in 2022,
-which is a relatively low open access share compared to the general
-trend. In 2018, Springer Nature had the largest open access proportion
-among the three publishers of 8.4\% (n = 19,701), but experienced a
-relatively slower growth, resulting in 18\% (n = 52,616) of articles
-being open access in Springer Nature hybrid journals in 2022.
+2018 to 26\% (n = 53,503) in 2022. In contrast, Elsevier's hybrid
+journals demonstrated a more modest increase, from 3.3\% (n = 16,872) in
+2018 to 10\% (n = 60,821) in 2022, which is a relatively low open access
+share compared to the general trend. In 2018, Springer Nature had the
+largest open access proportion among the three publishers of 8.4\% (n =
+19,701), but experienced a relatively slower growth, resulting in 18\%
+(n = 52,616) of articles being open access in Springer Nature hybrid
+journals in 2022.
 
@@
 
-\caption{Developement of open access in hybrid journals included in transformative agreements between 2018 and 2022 by publishers.
+\caption{Development of open access in hybrid journals included in transformative agreements between 2018 and 2022 by publishers.
 The blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established.
@@
 importance of agreements for Springer Nature's hybrid open access
-business over the past five years (Figure 2B). In 2022, 66\% (n =
-34,725) of open access in Springer Nature hybrid journals was enabled
-through transformative agreements. In the same year, 70\% (n = 37,316)
-of Wiley's open access articles could be linked to transformative
-agreements in 2022. By contrast, Elsevier published fewer than half of
-its open access articles through transformative agreements (n = 32,627;
-54\%).
+business over the past five years (Figure \ref{fig:publisher_figure}).
+In 2022, 66\% (n = 34,725) of open access in Springer Nature hybrid
+journals was enabled through transformative agreements. In the same
+year, 70\% (n = 37,316) of Wiley's open access articles could be linked
+to transformative agreements. By contrast, Elsevier published a
+comparatively lower proportion of its open access articles through
+transformative agreements in 2022 (n = 32,627; 54\%).
 
@@
 
-\hypertarget{journal-subjects}{%
-\subsection{Journal subjects}\label{journal-subjects}}
+\subsection{Journal subjects}\label{journal-subjects}
 
 Table \ref{tab:subject_summary_table} presents a high-level overview of
-hybrid open access by AJCS subject area using fractionalised counting to
+hybrid open access by ASJC subject area using fractional counting to
 account for journals belonging to more than one category. Between 2018
@@
 
-\caption{Developement of open access in hybrid journals in transformative agreements between 2018 and 2022 by AJCS subject area.
+\caption{Development of open access in hybrid journals in transformative agreements between 2018 and 2022 by ASJC subject area.
 The blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established.
-(A) Proportion of open access articles in hybrid journals per year and publisher.
-(B) Percentage of hybrid open access via agreements per year and publisher.
-Boxplots (C) show individual open access uptake rates by individual hybrid journals and open access funding per publication year and publisher.
-The individual outliers are not shown.
+(A) Proportion of open access articles in hybrid journals per year and subject area (B) Percentage of hybrid open access via agreements per year and subject area Boxplots (C) show individual open access uptake rates by individual hybrid journals and open access funding per publication year and subject area The individual outliers are not shown.
 Note that data on transformative agreements ending before June 2021 were not available for this study.}\label{fig:subject_panel}
@@
 
-\hypertarget{comparing-countries}{%
-\subsection{Comparing countries}\label{comparing-countries}}
+\subsection{Comparing countries}\label{comparing-countries}
 
@@
 
+\begin{figure}[ht!]
+
+{\centering \includegraphics[width=0.99\linewidth,]{fig/country_patch-1}
+
+}
+
+\caption{Development of hybrid open access publishing by country.
+(A) presents the relative number of articles published in hybrid journals included in transformative agreement by year, distinguishing between BRICS as of 2022, OECD and other countries.
+(B) Scatterplot contrasts total articles with open access article volume in 2022, by country and its OECD (purple-colored points) or BRICS (green) membership.
+Trend line obtained from linear regression, grey areas show pointwise symmetric 95\% confidence bands.
+(C) Hybrid open access market share in 2022 by country.
+In (A) and (C), the blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established.
+The remainder shows closed access articles.
+Country names are represented as ISO two-letter country codes.}\label{fig:country_patch}
+\end{figure}
+
 Figure \ref{fig:country_patch}A shows the development of hybrid open
@@
 includes the remaining countries. Full counting was applied to account
-for multiple country affiliations (Hottenrott et al., 2021). From 2018
-to 2022, the proportion of open access in hybrid journals increased from
-6.1\% in 2018 to 26\% in 2022. In contrast, BRICS recorded a low uptake,
-moderately growing from 1.6\% in 2018 to 3.7\% in 2022.
+for multiple country affiliations (Hottenrott et al., 2021). Open access
+in hybrid journals increased in particular because of publications from
+authors affiliated with institutions from the OECD area, from 52,154 in
+2018 to 202,787 in 2022. This growth was largely driven by
+transformative agreements. The share of open access by agreement rose
+from 5.5\% in 2018 (n = 2,859) to 68\% (n = 137,815) in 2022. In
+contrast, BRICS recorded a low uptake, moderately growing from 1.6\% in
+2018 to 3.7\% in 2022.
 
@@
 publication output decreased sharply, dropping to 786,903 in 2022 after
-peaking 892,197 articles in 2020. In stark contrast, the number of
-articles published in hybrid journals by first authors affiliated with
-institutions from BRICS countries increased steadily over the years,
-more than doubling from 356,632 in 2018 to 786,903 in 2022. Upon closer
-examination, this trend can be observed across all of the three largest
-publishers, although the shift towards BRICS is particularly evident in
-Elsevier's hybrid journal portfolio, particularly with regard to
-articles published in Physical Sciences journals. While OECD publication
-output in Elsevier's Physical Sciences journals declined from 112,822
-articles in 2018 to 103,766 in 2022, BRICS output increased from 104,654
-to 171,713 in the same five-year period. Furthermore, OECD publication
-output in Health Sciences and Life Sciences journals stagnated after
-peaking in 2020.
-
-\begin{figure}[ht!]
-
-{\centering \includegraphics[width=0.99\linewidth,]{fig/country_patch-1}
-
-}
-
-\caption{Development of hybrid open access publishing by country.
-(A) presents the number of articles published in hybrid journals included in transformative agreement, distinguishing between BRICS as of 20222, OECD and other countries.
-(B) Scatterplot contrasting total articles with open access article volume in 2022, by country.
-Point size represents the number of articles enabled through transformative agreements.
-(C) Hybrid open access market share in 2022 by country.
-In (A) and (C), the blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established.
-Country names are represented as ISO two-letter country codes.}\label{fig:country_patch}
-\end{figure}
+peaking 892,197 articles in 2020, which saw a massive growth in
+literature related to COVID-19, driven largely by researchers in the
+United States surpassing China and other countries (Ioannidis et al.,
+2021). In stark contrast, the number of articles published in hybrid
+journals by first authors affiliated with institutions from BRICS
+countries increased steadily over the years, from 356,632 in 2018 to
+535,828 in 2022. This resulted in an increase in market share of the
+BRICS area from 23\% to 31\% between 2018 and 2022, whereas that of the
+OECD area decreased from 55\% to 46\% during the same period. Upon
+closer examination, this trend can be observed across all of the three
+largest publishers, although the shift towards BRICS is particularly
+evident in Elsevier's hybrid journal portfolio, particularly with regard
+to articles published in Physical Sciences journals. While OECD
+publication output in Elsevier's Physical Sciences journals declined
+from 112,822 articles in 2018 to 103,766 in 2022, BRICS output increased
+from 104,654 to 171,713 in the same five-year period. Furthermore, OECD
+publication output in Health Sciences and Life Sciences journals
+stagnated after peaking in 2020, which saw an increase in publications
+due to the impact of the global COVID-19 pandemic, in particular from
+the United States (Ioannidis et al., 2021).
 
@@
 Netherlands, which was 7.1\%.
+
+\begin{figure}[ht!]
+
+{\centering \includegraphics[width=0.99\linewidth,]{fig/country_top_20_plot-1}
+
+}
+
+\caption{Development of open access in hybrid journals in transformative agreements between 2018 and 2022, by the Top 20 most productive countries in terms of total articles published in the five-years period.
+Blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established.
+Country names are represented as ISO two-letter country codes.
+Facet subheadings show the total number of articles and corresponding market share.}\label{fig:country_top_20_plot}
+\end{figure}
 
@@
 
-\begin{figure}[ht!]
-
-{\centering \includegraphics[width=0.99\linewidth,]{fig/country_top_20_plot-1}
-
-}
-
-\caption{Development of open access in hybrid journals in transformative agreements between 2018 and 2022, by the Top 20 most productive countries in terms of total articles published in the five-years period.
-Blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established.
-Country names are represented as ISO two-letter country codes.
-Facet subheadings show the total number of articles and corresponding market share.}\label{fig:country_top_20_plot}
-\end{figure}
-
 Since 2021, there has been a general trend towards hybrid open access
@@
 
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
+\section{Discussion}\label{discussion}
 
@@
 reflected in the data in terms of open access uptake in hybrid journals.
-As highlighted by Koley \& Lala (2023), India faces similar challenges
-in terms of the availablility of resources to pay for open access. At
-the same time, access to research literature is a pressing issue, which
-is addressed by the ``Indian one nation, one subscription'' policy
+Market observers also do not expect a broad implementation of
+transformative agreements in China in near future (Owens, 2024). As
+highlighted by Koley \& Lala (2023), India faces similar challenges in
+terms of the availablility of resources to pay for open access. At the
+same time, access to research literature is a pressing issue, which is
+addressed by the ``Indian one nation, one subscription'' policy
 proposal. However, this policy focuses on centrally negotiated
@@
 acknowledged. From a data perspective, estimations of open access
-through transformative agreements were established by linking first
-author affiliations with publicly available agreement data from
-cOAlition S, and not through invoicing data, which is usually not
-shared. Moreover, this study is unable to account for the various types
-of transformative agreements, particularly with regard to article types
-and caps that limit the number of open access articles covered.
-Furthermore, assessing the quality of the OpenAlex and Crossref data
-used, particularly in terms of affiliations and article types, by
-combining them with established bibliometric databases such as Scopus
-and Web of Science was beyond the scope of this analysis (Visser et al.,
-2021). Overall, the methodology is designed to underestimate, rather
-than overestimate, the adoption of open access through transformative
-agreements. In addition, the data used in this study are openly
-available, along with the code used for this analysis.
+through transformative agreements were established by linking author
+affiliations with publicly available agreement data from cOAlition S,
+and not through invoicing data, which is usually not shared. Because
+data on corresponding authors, who are typically responsible for
+facilitating open access publication, were not fully supported by
+OpenAlex at the time of this study, first author affiliations were used
+instead. This study is also unable to account for the various types of
+transformative agreements due to a lack of data, especially regarding
+article types and caps limiting the number of open access articles
+covered. Furthermore, assessing the quality of the OpenAlex and Crossref
+data used, particularly in terms of affiliations, corresponding
+authorship, and article types, by combining them with established
+bibliometric databases such as Scopus and Web of Science was beyond the
+scope of this analysis (Chinchilla-Rodríguez et al., 2024; Visser et
+al., 2021). It must therefore be emphasised that, unlike related
+evaluations of national consortia based on invoice data (Brayman et al.,
+2024), this global overview can only provide an estimate of the support
+for open access through transformative agreements. Despite these
+limitations, the methodology was designed to underestimate, rather than
+overestimate, the adoption of open access through transformative
+agreements. To promote transparency, the data used in this study, along
+with the code used for this analysis, are openly available.
 
@@
 countries, particularly the relationship between socio-economic
-development and open access adoption. From a data perspective, the
-reporting of open access funding, including transformative agreements,
-is not harmonised, but often crowd-sourced from various sources. To
-improve the assessment of transformative agreements, libraries and
-publishers need to collaborate on standards and services to publicly
-share information about respective journal portfolios, participating
-institutions and open access invoicing.
+development and open access adoption. Between 2018 and 2022, there was a
+notable increase in the number of closed access articles published in
+hybrid journals by authors from countries such as China and India, where
+transformative agreements and open access funding options were not
+widely available. In contrast, the introduction of country-wide
+transformative agreements in numerous OECD countries led to a
+substantial increase in open access, while their share of total articles
+in hybrid journals decreased. This imbalance, whereby open access uptake
+depends largely on a few countries, makes it less likely that collective
+action towards a large-scale transition of hybrid journals to full open
+access through transformation agreements will succeed within the next
+years. Rather, and in accordance with recent policy recommendations from
+the UK and Nordic countries (Brayman et al., 2024; Widding, 2024), it
+emphasises the importance of ongoing evaluation and adaptation of open
+access strategies in view of global publishing trends.
+
+From a data perspective, research institutions and funders need to push
+towards price transparency in scholarly publishing to enable these
+evaluations. The reporting of open access funding, including
+transformative agreements, is also not harmonised, but often
+crowd-sourced from various sources. To improve the assessment of
+transformative agreements, libraries and publishers should collaborate
+on standards and services to publicly share information about respective
+journal portfolios, participating institutions and open access
+invoicing, for example in the international context of ESAC and the
+Barcelona Declaration on Open Research Information. Such collaboration
+would help overcome the limitations of current approaches that derive
+funding for open access from authorship data.
 
@@
 
-\hypertarget{competing-interests}{%
-\section{Competing interests}\label{competing-interests}}
+\section{Competing interests}\label{competing-interests}
 
@@
 
-\hypertarget{funding-information}{%
-\section{Funding information}\label{funding-information}}
+\section{Funding information}\label{funding-information}
 
@@
 
-\hypertarget{data-and-code-availability}{%
-\section{Data and code availability}\label{data-and-code-availability}}
+\section{Data and code availability}\label{data-and-code-availability}
 
@@
 
-\hypertarget{references}{%
-\section*{References}\label{references}}
+\section*{References}\label{references}
 \addcontentsline{toc}{section}{References}
 
-\hypertarget{refs}{}
+\phantomsection\label{refs}
 \begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-Aviv_Reuven_2021}{}}%
+\bibitem[\citeproctext]{ref-Aviv_Reuven_2021}
 Aviv-Reuven, S., \& Rosenfeld, A. (2021). Publication patterns' changes
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-budapest}{}}%
+\bibitem[\citeproctext]{ref-budapest}
 Babini, D., Chan, L., Hagemann, M., Joseph, H., Kuchma, I., \& Suber, P.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Bakker_2024}{}}%
+\bibitem[\citeproctext]{ref-Bakker_2024}
 Bakker, C., Langham-Putrow, A., \& Riegelman, A. (2024). Impact of
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Bergstrom_2014}{}}%
+\bibitem[\citeproctext]{ref-Bergstrom_2014}
 Bergstrom, T. C., Courant, P. N., McAfee, R. P., \& Williams, M. A.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Bj_rk_2012}{}}%
+\bibitem[\citeproctext]{ref-Bj_rk_2012}
 Björk, B.-C. (2012). The hybrid model for open access publication of
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Bj_rk_2017}{}}%
+\bibitem[\citeproctext]{ref-Bj_rk_2017}
 Björk, B.-C. (2017). Growth of hybrid open access, 2009--2016.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-BJ_RK_2014}{}}%
+\bibitem[\citeproctext]{ref-BJ_RK_2014}
 Björk, B.-C., \& Solomon, D. (2014). How research funders can finance
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Borrego_2023}{}}%
+\bibitem[\citeproctext]{ref-Borrego_2023}
 Borrego, Á. (2023). Article processing charges for open access journal
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Borrego_2020}{}}%
+\bibitem[\citeproctext]{ref-Borrego_2020}
 Borrego, Á., Anglada, L., \& Abadal, E. (2021). Transformative
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-bosman_2021_4558704}{}}%
+\bibitem[\citeproctext]{ref-bosman_2021_4558704}
 Bosman, J., Frantsvåg, J. E., Kramer, B., Langlais, P.-C., \& Proudman,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Bosman_2021}{}}%
+\bibitem[\citeproctext]{ref-Bosman_2021}
 Bosman, J., Jonge, H. de, Kramer, B., \& Sondervan, J. (2021). Advancing
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Brainard_2021}{}}%
+\bibitem[\citeproctext]{ref-Brainard_2021}
 Brainard, J. (2021). California universities and {Elsevier} make up, ink
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Brainard_2023}{}}%
+\bibitem[\citeproctext]{ref-Brainard_2023}
 Brainard, J. (2023). {``Transformative''} journals get booted for
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Jisc_2024}{}}%
+\bibitem[\citeproctext]{ref-Jisc_2024}
 Brayman, K., Devenney, A., Dobson, H., Marques, M., \& Vernon, A.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-goldoa}{}}%
+\bibitem[\citeproctext]{ref-goldoa}
 Bruns, A., Cakir, Y., Kaya, S., \& Beidaghi, S. (2022).
-\emph{{ISSN-Matching of Gold OA Journals (ISSN-GOLD-OA) 5.0}}. Bielefeld
+\emph{{ISSN-matching of Gold OA journals (ISSN-GOLD-OA) 5.0}}. Bielefeld
 University. \url{https://doi.org/10.4119/unibi/2961544}
 
-\leavevmode\vadjust pre{\hypertarget{ref-Butler_2023}{}}%
+\bibitem[\citeproctext]{ref-Butler_2023}
 Butler, L.-A., Matthias, L., Simard, M.-A., Mongeon, P., \& Haustein, S.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-ifla}{}}%
+\bibitem[\citeproctext]{ref-ifla}
 Campbell, C., Dér, Á., Geschuhn, K., \& Valente, A. (2022). How are
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Dallmeier_2010}{}}%
+\bibitem[\citeproctext]{ref-Chinchilla_Rodr_guez_2024}
+Chinchilla-Rodríguez, Z., Costas, R., Robinson-García, N., \& Larivière,
+V. (2024). Examining the quality of the corresponding authorship field
+in {Web of Science} and {Scopus}. \emph{Quantitative Science Studies},
+\emph{5}(1), 76--97. \url{https://doi.org/10.1162/qss_a_00288}
+
+\bibitem[\citeproctext]{ref-Dallmeier_2010}
 Dallmeier-Tiessen, S., Goerner, B., Darby, R., Hyppoelae, J.,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Fraser_2021}{}}%
+\bibitem[\citeproctext]{ref-Fraser_2021}
 Fraser, N., Brierley, L., Dey, G., Polka, J. K., Pálfy, M., Nanni, F.,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Fraser_2023}{}}%
+\bibitem[\citeproctext]{ref-Fraser_2023}
 Fraser, N., Hobert, A., Jahn, N., Mayr, P., \& Peters, I. (2023). No
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Geschuhn_2017}{}}%
+\bibitem[\citeproctext]{ref-Geschuhn_2017}
 Geschuhn, K., \& Stone, G. (2017). It's the workflows, stupid! What is
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-hanson2023strain}{}}%
+\bibitem[\citeproctext]{ref-hanson2023strain}
 Hanson, M. A., Barreiro, P. G., Crosetto, P., \& Brockington, D. (2023).
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Haucap_2021}{}}%
+\bibitem[\citeproctext]{ref-Haucap_2021}
 Haucap, J., Moshgbar, N., \& Schmal, W. B. (2021). The impact of the
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Hendricks_2020}{}}%
+\bibitem[\citeproctext]{ref-Hendricks_2020}
 Hendricks, G., Tkaczyk, D., Lin, J., \& Feeney, P. (2020). Crossref: The
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Hinchliffe_2019}{}}%
+\bibitem[\citeproctext]{ref-Hinchliffe_2019}
 Hinchliffe, L. J. (2019). \emph{Transformative agreements: A primer}.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Hottenrott_2021}{}}%
+\bibitem[\citeproctext]{ref-norway}
+Holden, L., Skoie, M., Røeggen, V., Bjerde, K. W., Wenaas, L., Bakke,
+P., Løvhaug, J. W., Karlsen, E. S., \& Qvenild, M. (2023).
+\emph{Strategi for vitenskapelig publisering etter 2024}. Sikt.
+\url{https://doi.org/10.18711/2KZ1-BA97}
+
+\bibitem[\citeproctext]{ref-Hottenrott_2021}
 Hottenrott, H., Rose, M. E., \& Lawson, C. (2021). The rise of multiple
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Huang_2020}{}}%
+\bibitem[\citeproctext]{ref-Huang_2020}
 Huang, C.-K. (Karl), Neylon, C., Hosking, R., Montgomery, L., Wilson, K.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-hoaddata}{}}%
+\bibitem[\citeproctext]{ref-Ioannidis_2021}
+Ioannidis, J. P. A., Salholz-Hillel, M., Boyack, K. W., \& Baas, J.
+(2021). The rapid, massive growth of COVID-19 authors in the scientific
+literature. \emph{Royal Society Open Science}, \emph{8}(9).
+\url{https://doi.org/10.1098/rsos.210389}
+
+\bibitem[\citeproctext]{ref-hoaddata}
 Jahn, N. (2023). \emph{{hoaddata}: Data about hybrid open access journal
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-jahn2023}{}}%
+\bibitem[\citeproctext]{ref-jahn2023}
 Jahn, N., Haupka, N., \& Hobert, A. (2023). \emph{Analysing and
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Jahn_2021}{}}%
+\bibitem[\citeproctext]{ref-Jahn_2021}
 Jahn, N., Matthias, L., \& Laakso, M. (2022). Toward transparency of
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Jahn_2016}{}}%
+\bibitem[\citeproctext]{ref-Jahn_2016}
 Jahn, N., \& Tullney, M. (2016). A study of institutional spending on
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Jubb_2017}{}}%
+\bibitem[\citeproctext]{ref-Jubb_2017}
 Jubb, M., Plume, A., Oeben, S., Brammer, L., Johnson, R., Bütün, C., \&
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Khanna_2022}{}}%
+\bibitem[\citeproctext]{ref-Khanna_2022}
 Khanna, S., Ball, J., Alperin, J. P., \& Willinsky, J. (2022).
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Koley_2023}{}}%
+\bibitem[\citeproctext]{ref-Koley_2023}
 Koley, M., \& Lala, K. (2023). Limitations of the {{``Indian one nation,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Kramer_2024}{}}%
+\bibitem[\citeproctext]{ref-Kramer_2024}
 Kramer, B. (2024). \emph{Study on scientific publishing in {Europe} --
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Kwiek_2020}{}}%
+\bibitem[\citeproctext]{ref-Kwiek_2020}
 Kwiek, M. (2021). What large-scale publication and citation data tell us
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Laakso_2016}{}}%
+\bibitem[\citeproctext]{ref-Laakso_2016}
 Laakso, M., \& Björk, B.-C. (2016). Hybrid open access--a longitudinal
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Larivi_re_2016}{}}%
+\bibitem[\citeproctext]{ref-Larivi_re_2016}
 Larivière, V., Desrochers, N., Macaluso, B., Mongeon, P., Paul-Hus, A.,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Larivi_re_2015}{}}%
+\bibitem[\citeproctext]{ref-Larivi_re_2015}
 Larivière, V., Haustein, S., \& Mongeon, P. (2015). The oligopoly of
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Liverpool_2023}{}}%
-Liverpool, L. (2023). Open-access reformers launch next bold publishing
-plan. \emph{Nature}, \emph{623}(7986), 238--240.
-\url{https://doi.org/10.1038/d41586-023-03342-6}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Marques_2020}{}}%
+\bibitem[\citeproctext]{ref-Marques_2020}
 Marques, M., \& Stone, G. (2020). Transitioning to open access: An
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Marques_2019}{}}%
+\bibitem[\citeproctext]{ref-Marques_2019}
 Marques, M., Woutersen-Windhouwer, S., \& Tuuliniemi, A. (2019).
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Mart_n_Mart_n_2018}{}}%
+\bibitem[\citeproctext]{ref-Mart_n_Mart_n_2018}
 Martín-Martín, A., Costas, R., Leeuwen, T. van, \& López-Cózar, E. D.
-(2018). Evidence of open access of scientific publications in google
-scholar: A large-scale analysis. \emph{Journal of Informetrics},
+(2018). Evidence of open access of scientific publications in {Google
+Scholar}: A large-scale analysis. \emph{Journal of Informetrics},
 \emph{12}(3), 819--841. \url{https://doi.org/10.1016/j.joi.2018.06.012}
 
-\leavevmode\vadjust pre{\hypertarget{ref-Marwick_2018}{}}%
+\bibitem[\citeproctext]{ref-Marwick_2018}
 Marwick, B., Boettiger, C., \& Mullen, L. (2018). Packaging data
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Matthias_2019}{}}%
+\bibitem[\citeproctext]{ref-Matthias_2019}
 Matthias, L., Jahn, N., \& Laakso, M. (2019). The two-way street of open
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Mittermaier_2015}{}}%
+\bibitem[\citeproctext]{ref-Mittermaier_2015}
 Mittermaier, B. (2015). Double dipping in hybrid open access -- chimera
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Mittermaier_2021}{}}%
+\bibitem[\citeproctext]{ref-Mittermaier_2021}
 Mittermaier, B. (2021). {Rolle des Open Access Monitor Deutschland bei
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Momeni_2023}{}}%
+\bibitem[\citeproctext]{ref-Momeni_2023}
 Momeni, F., Dietze, S., Mayr, P., Biesenbender, K., \& Peters, I.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Momeni_2021}{}}%
+\bibitem[\citeproctext]{ref-Momeni_2021}
 Momeni, F., Mayr, P., Fraser, N., \& Peters, I. (2021). What happens
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Moskovkin_2022}{}}%
+\bibitem[\citeproctext]{ref-Moskovkin_2022}
 Moskovkin, V. M., Saprykina, T. V., \& Boichuk, I. V. (2022).
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Mu_oz_V_lez_2024}{}}%
+\bibitem[\citeproctext]{ref-Mu_oz_V_lez_2024}
 Muñoz-Vélez, H., Pallares, C., Echavarría, A. F., Contreras, J., Pavas,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Parmhed_2023}{}}%
+\bibitem[\citeproctext]{ref-Owens_2024}
+Owens, B. (2024). China's research clout leads to growth in homegrown
+science publishing. \emph{Nature}, \emph{630}(8015), S2--S4.
+\url{https://doi.org/10.1038/d41586-024-01596-2}
+
+\bibitem[\citeproctext]{ref-Parmhed_2023}
 Parmhed, S., \& Säll, J. (2023). Transformative agreements and their
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Pieper_2018}{}}%
+\bibitem[\citeproctext]{ref-Pieper_2018}
 Pieper, D., \& Broschinski, C. (2018). {OpenAPC}: A contribution to a
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Pinfield_2016}{}}%
+\bibitem[\citeproctext]{ref-Pinfield_2016}
 Pinfield, S., Salter, J., \& Bath, P. A. (2016). The "total cost of
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Pinhasi_2021}{}}%
+\bibitem[\citeproctext]{ref-Pinhasi_2021}
 Pinhasi, R., Hölbling, L., \& Kromp, B. (2021). Austrian transition to
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Pinhasi_2020}{}}%
+\bibitem[\citeproctext]{ref-Pinhasi_2020}
 Pinhasi, R., Kromp, B., Blechl, G., \& Hölbling, L. (2020). The impact
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Piwowar_2018}{}}%
+\bibitem[\citeproctext]{ref-Piwowar_2018}
 Piwowar, H., Priem, J., Larivière, V., Alperin, J. P., Matthias, L.,
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-P_l_nen_2020}{}}%
+\bibitem[\citeproctext]{ref-P_l_nen_2020}
 Pölönen, J., Laakso, M., Guns, R., Kulczycki, E., \& Sivertsen, G.
 (2020). Open access at the national level: A comprehensive analysis of
-publications by finnish researchers. \emph{Quantitative Science
+publications by {Finnish} researchers. \emph{Quantitative Science
 Studies}, \emph{1}(4), 1396--1428.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-priem2022openalex}{}}%
+\bibitem[\citeproctext]{ref-priem2022openalex}
 Priem, J., Piwowar, H., \& Orr, R. (2022). \emph{OpenAlex: A fully-open
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Prosser_2003}{}}%
+\bibitem[\citeproctext]{ref-Prosser_2003}
 Prosser, D. C. (2003). From here to there: A proposed mechanism for
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Goodin_2023}{}}%
+\bibitem[\citeproctext]{ref-Goodin_2023}
 Rasmussen, K. B. (2023). Interview with {Robert {``Bob''} E. Goodin}.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Robinson_Garcia_2020}{}}%
+\bibitem[\citeproctext]{ref-Robinson_Garcia_2020}
 Robinson-Garcia, N., Costas, R., \& Leeuwen, T. N. van. (2020). Open
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Ross_Hellauer_2022}{}}%
+\bibitem[\citeproctext]{ref-Ross_Hellauer_2022}
 Ross-Hellauer, T., Reichmann, S., Cole, N. L., Fessl, A., Klebel, T., \&
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Schares_2022}{}}%
+\bibitem[\citeproctext]{ref-Schares_2022}
 Schares, E. (2022). Unsub extender: A python-based web application for
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Schares_2023}{}}%
+\bibitem[\citeproctext]{ref-Schares_2023}
 Schares, E. (2023). Impact of the 2022 OSTP memo: A bibliometric
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Schiermeier_2018}{}}%
+\bibitem[\citeproctext]{ref-Schiermeier_2018}
 Schiermeier, Q. (2018). China backs bold plan to tear down journal
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Schiltz_2018}{}}%
+\bibitem[\citeproctext]{ref-Schiltz_2018}
 Schiltz, M. (2018). Science without publication paywalls: {cOAlition S}
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Schimmer_2015}{}}%
+\bibitem[\citeproctext]{ref-Schimmer_2015}
 Schimmer, R., Geschuhn, K., \& Vogler, A. (2015). \emph{{Disrupting the
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Schmal_2024}{}}%
+\bibitem[\citeproctext]{ref-Schmal_2024}
 Schmal, W. B. (2024). How transformative are transformative agreements?
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Schmidt_2012}{}}%
+\bibitem[\citeproctext]{ref-Schmidt_2012}
 Schmidt, B., \& Shearer, K. (2012). Licensing revisited: Open access
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Shu_2023}{}}%
+\bibitem[\citeproctext]{ref-Shu_2023}
 Shu, F., \& Larivière, V. (2023). The oligopoly of open access
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Taubert_2023}{}}%
+\bibitem[\citeproctext]{ref-Taubert_2023}
 Taubert, N., Hobert, A., Jahn, N., Bruns, A., \& Iravani, E. (2023).
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Visser_2021}{}}%
+\bibitem[\citeproctext]{ref-Visser_2021}
 Visser, M., Eck, N. J. van, \& Waltman, L. (2021). Large-scale
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Wenaas_2022}{}}%
+\bibitem[\citeproctext]{ref-ludo_waltman_2022_7105355}
+Waltman, L., \& Lamers, W. S. (2022). \emph{{Monitoring Open Access
+publishing of NWO-funded research (2015-2021)}} (Version 1). Zenodo.
+\url{https://doi.org/10.5281/zenodo.7105355}
+
+\bibitem[\citeproctext]{ref-Wenaas_2022}
 Wenaas, L. (2022). Choices of immediate open access and the relationship
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-bigrquery}{}}%
-Wickham, H., \& Bryan, J. (2023). \emph{Bigrquery: An interface to
-google's 'BigQuery' 'API'}.
+\bibitem[\citeproctext]{ref-bigrquery}
+Wickham, H., \& Bryan, J. (2023). \emph{{bigrquery: An interface to
+Google's 'BigQuery' 'API'}}.
 \url{https://CRAN.R-project.org/package=bigrquery}
 
-\leavevmode\vadjust pre{\hypertarget{ref-Zhang_2024}{}}%
+\bibitem[\citeproctext]{ref-Widding_2024}
+Widding, A. S. (2024). Beyond transformative agreements: Ways forward
+for universities. \emph{European Review}, 1--11.
+\url{https://doi.org/10.1017/s1062798724000036}
+
+\bibitem[\citeproctext]{ref-Zhang_2024}
 Zhang, L., Cao, Z., Shang, Y., Sivertsen, G., \& Huang, Y. (2024).
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Zhang_2022}{}}%
+\bibitem[\citeproctext]{ref-Zhang_2022}
 Zhang, L., Wei, Y., Huang, Y., \& Sivertsen, G. (2022). Should open
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Zhao_2023}{}}%
+\bibitem[\citeproctext]{ref-Zhao_2023}
 Zhao, X., Akbaritabar, A., Kashyap, R., \& Zagheni, E. (2023). A gender

### 2024-09-18 analysis/manuscript.Rmd
Language check

@@
 The controversies surrounding hybrid open access and transformative agreements have led to varying policy conclusions. 
-For instance, the UK's Joint Information Systems Committee (JISC) evaluated its open access strategy, which included transformative agreements [@Jisc_2024]. 
-The evaluation revealed that while these agreements had a significant impact on the growth of open access in the UK, they had limited effects in facilitating a global shift towards full open access. 
-As a result, the report advised that the UK's open access strategy should be reassessed.
+For instance, the British Joint Information Systems Committee (JISC) evaluated its open access strategy, which included transformative agreements [@Jisc_2024]. 
+The evaluation revealed that while these agreements had a significant impact on the growth of open access in the country, they had limited effects in facilitating a global shift towards full open access.
+As a result, the report advised that the British open access strategy should be reassessed.
 After making similar observation, Norwegian and Swedish universities and their consortia also argued for policy changes [@norway; @Widding_2024].
-Notably, the Association of Swedish Higher Education Institutions (Sveriges universitets- och högskoleförbund, SUHF) recommended only supporting agreements for publishing in fully open access journals by 2026 [@Widding_2024].
-Furthermore, cOAlition S will conclude its financial support of transformative agreements at the end of 2024, but will continue to view open access resulting from such agreements as compliant.
-[^coalitions].
+Furthermore, cOAlition S conclude its financial support of transformative agreements at the end of 2024, but continue to view open access resulting from such agreements as compliant. [^coalitions].
 cOAlition S also removed the majority of hybrid journals from its Transformative Journal program in 2023 due to publishers' failure to meet self-defined open access growth targets [@Brainard_2023]. 
@@
 In particular, data availability is a limiting factor when studying the impact of transformative agreements [@Bakker_2024], because bibliometric databases, even though many allow the retrieval of open access articles in hybrid journals, do not directly attribute them to specific transformative agreements.
-Likewise, article-level open access invoicing and cost data are only partly available, which would make it possible to establish a direct link between transformative agreements and open access publications [@Jahn_2021; @Kramer_2024].
+Likewise, article-level open access invoicing and cost data, which would make it possible to establish a direct link between transformative agreements and open access publications [@Jahn_2021; @Kramer_2024].
 
@@
 After identifying hybrid journals by excluding fully open access journals, Crossref [@Hendricks_2020] served as the primary data source for article-level metadata including Creative Commons (CC) license information to indicate open access availability on publisher websites.
-Because of a lack of comprehensive publicly available invoicing data, open access articles published through transformative agreements were determined by first author affiliations from OpenAlex [@priem2022openalex], linked to eligible institutions according to the transformative agreement data.
+Because of a lack of comprehensive publicly available invoicing data, open access articles published through transformative agreements were determined by linking first author affiliations from OpenAlex [@priem2022openalex] to eligible institutions according to the transformative agreement data.
 In the following, the steps are described in detail.
@@
 
-Data gathering started with obtaining journals included in transformative agreements from the publicly available Transformative Agreement Data dump[^tadump] used by the cOAlition S Journal Checker Tool [^checkertool], a voluntary effort based on publicly disclosed contracts in the ESAC Transformative Agreement Registry.[^esacregistry].
+Data gathering started with obtaining journals included in transformative agreements from the publicly available Transformative Agreement Data dump[^tadump] used by the cOAlition S Journal Checker Tool[^checkertool], a voluntary effort based on publicly disclosed contracts in the ESAC Transformative Agreement Registry[^esacregistry].
 The dump consists of multiple online Google spreadsheets where each data file represents one agreement listed in the ESAC Transformative Agreement Registry.
@@
 
-It should be noted that although many library consortia see the need to register their agreements through the ESAC registry, many fail to disclose all details, including contracts [@Bakker_2024; @Kramer_2024].
+It should be noted that although many library consortia see the need to register their agreements through the ESAC registry, some fail to disclose all details, including the full-text of contracts [@Bakker_2024; @Kramer_2024].
 Furthermore, the extent to which ESAC is comprehensive remains uncertain, potentially limiting the coverage of transformative agreements in the Journal Checker Tool. 
@@
 Because the corresponding authorship field was not fully supported by OpenAlex at the time of analysis, and respective affiliation data were only available for 54% of the investigated articles, this study focused on first authors and their affiliations instead; approximately 90% of the articles examined had first author affiliation metadata in OpenAlex, which is a much larger proportion than previously reported for the October 2022 snapshot [@Zhang_2024].
-First authors typically contribute the most to a paper and are often considered lead authors [@Larivi_re_2016], although notable differences in corresponding author metadata were observed across subjects, publication years and bibliometric databases [@Chinchilla_Rodr_guez_2024].
+First authors typically contribute the most to a paper and are often considered as lead authors [@Larivi_re_2016; @Chinchilla_Rodr_guez_2024].
 In case of lacking data about corresponding authors in bibliometric databases, related studies also utilised first authors as a proxy to examine open access payments and transformative agreements [@Zhang_2022; @Shu_2023; @Haucap_2021].
@@
 Matching also considered the duration of agreements according to the ESAC registry. 
-`r length(unique(tas_in_esac_and_hoad_active_in_2018_22$ID))` agreements were active between 2018 and 2022.
+In total, `r length(unique(tas_in_esac_and_hoad_active_in_2018_22$ID))` agreements were active between 2018 and 2022.
 Upon inspection, Transformative Agreement Data did not comprehensively cover associated institutions, such as university hospitals or institutes of large research organisations such as the Max Planck Society.
@@
 
-Between 2018 and 2022, `r n_hybrid_jns` out of `r n_hybrid_jns_all` eligible hybrid journals in transformative agreements published at least one open access article under a Creative Commons license.
-These `r n_hybrid_jns` hybrid journals constituted the foundation for the subsequent analyses. 
+Between 2018 and 2022, `r n_hybrid_jns` out of `r n_hybrid_jns_all` hybrid journals in transformative agreements published at least one open access article under a Creative Commons license.
+These eligible `r n_hybrid_jns` hybrid journals constituted the foundation for the subsequent analyses. 
 Collectively, they provided open access to `r n_oa` out of `r n_total` articles during the investigated period, representing a five-year open access proportion of `r n_oa / n_total * 100`%.
 Authors who could make use of transformative agreements at the time of publication contributed `r n_ta_oa` (`r n_ta_oa / n_oa * 100`%) open access articles.
-Overall, this investigation was able to establish a link between open access articles and eligible institutions for `r n_tas_with_oa` out of `r length(unique(tas_in_esac_and_hoad_active_in_2018_22$ID))` (`r n_tas_with_oa / length(unique(tas_in_esac_and_hoad_active_in_2018_22$ID)) * 100`%) active transformative agreements between 2018 and 2022 through the analysis of publication metadata, including affiliations and agreement data.
+Overall, this investigation was able to establish a link between open access articles and eligible institutions for `r n_tas_with_oa` out of `r length(unique(tas_in_esac_and_hoad_active_in_2018_22$ID))` (`r n_tas_with_oa / length(unique(tas_in_esac_and_hoad_active_in_2018_22$ID)) * 100`%) active transformative agreements between 2018 and 2022 through the linking of first author affiliations with agreement data.
 
@@
 Although the median generally follows the trend shown in Figure \ref{fig:results_overview}A, the farther stretch of upper quartiles and whiskers over the years illustrates that an increasing number of journals published an above-average proportion of open access articles.
-In 2022, 25% of hybrid journals (n = `r oa_jn_box_df |> filter(cr_year == "2022") |> mutate(top25 = quantile(prop, .75)) |> filter(prop >= top25) |> nrow()`) had an open access uptake of `r oa_jn_box_df |> filter(cr_year == "2022") |> pull(prop) |> quantile(0.75)*100`%, and `r (oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> nrow()) / length(unique(oa_jn_box_df$issn_l)) *100`% of journals (n = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> nrow()`) provided the majority of their articles under a Creative Commons license in the same year.
+In 2022, 25% of hybrid journals (n = `r oa_jn_box_df |> filter(cr_year == "2022") |> mutate(top25 = quantile(prop, .75)) |> filter(prop >= top25) |> nrow()`) had an open access uptake of `r oa_jn_box_df |> filter(cr_year == "2022") |> pull(prop) |> quantile(0.75)*100`%, and `r (oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> nrow()) / length(unique(oa_jn_box_df$issn_l)) *100`% of journals (n = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> nrow()`) provided the majority of their articles under a CC license in the same year.
 On average, these journals were smaller (M = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> pull(n) |> mean()`, SD = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> pull(n) |> sd()`) than those with an open access share below 50% (M = `r oa_jn_box_df |> filter(cr_year == "2022", prop <= 0.5) |> pull(n) |> mean()`, SD = `r oa_jn_box_df |> filter(cr_year == "2022", prop <= 0.5) |> pull(n) |> sd()`).
@@
 Although Elsevier published the largest volume of articles (n = `r publisher_league_tbl  |> filter(publisher == "Elsevier") |> pull(publisher_total)`, `r publisher_league_tbl  |> filter(publisher == "Elsevier") |> pull(cum)*100`%), it recorded a comparably low number of open access articles, including those that are associated with transformative agreements.
-In contrast, Springer Nature and Wiley provided open access to a larger proportion of their articles (`r (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_total)) / (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(publisher_total)) * 100`% of Springer Nature articles and `r (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(oa_total)) / (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(publisher_total)) * 100`% of Wiley articles were open access), leading to higher open access market share (`r publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_perc)*100`% Springer Nature resp.
+In contrast, Springer Nature and Wiley provided open access to a larger proportion of their articles (`r (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_total)) / (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(publisher_total)) * 100`% of Springer Nature articles and `r (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(oa_total)) / (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(publisher_total)) * 100`% of Wiley articles were open access), leading to higher open access market shares (`r publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_perc)*100`% Springer Nature resp.
 `r publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_perc)*100`% Wiley).
@@
 The varying degrees of adoption of open access across the three major publishers can be attributed to distinct approaches to transformative agreements. 
-Springer Nature, for example, started offering open access agreements for its hybrid journal portfolio to selected consortia such as the Max Planck Society, the Swedish Bibsam Consortium and the Finnish FinELib consortium in 2015 under the name Springer Compact[^springercompact].
+Springer Nature, for example, started offering open access agreements for its hybrid journal portfolio to selected consortia such as the Max Planck Society, the Swedish Bibsam consortium and the Finnish FinELib consortium in 2015 under the name Springer Compact[^springercompact].
 Prior to this, Springer had some pilot agreements with a small number of institutions, including the University of Göttingen [@Schmidt_2012].
 However, the Springer Compact agreements were not included in the data as they concluded before the start of the transformative agreement data collection in June 2021. 
-Nonetheless, the results demonstrate the importance of agreements for Springer Nature's hybrid open access business over the past five years (Figure \ref{fig:publisher_figure}).
+Nonetheless, the results demonstrate the importance of agreements for Springer Nature's hybrid open access business over the five-years period (Figure \ref{fig:publisher_figure}B).
 In 2022, `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature") |> pull(value) |> sum()*100`% (n = `r  publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature", type == "ta_oa") |> pull(value)`) of open access in Springer Nature hybrid journals was enabled through transformative agreements.
@@
 Figure \ref{fig:subject_panel} presents the relative growth of hybrid open access by subject area between 2018 and 2022.
-In particular, Social Sciences including Arts and Humanties journals accounted for the strongest growth in the five-years period from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Social Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Social Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(value) |> sum()`), followed by the Life Sciences from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Life Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Life Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Life Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Life Sciences") |> pull(value) |> sum()`) , Health Sciences from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Health Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Health Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Health Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Health Sciences") |> pull(value) |> sum()`) and Physical Sciences from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Physical Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Physical Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Physical Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Physical Sciences") |> pull(value) |> sum()`).
+In particular, Social Sciences including Arts and Humanities journals accounted for the strongest growth in the five-years period from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Social Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Social Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(value) |> sum()`), followed by the Life Sciences from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Life Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Life Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Life Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Life Sciences") |> pull(value) |> sum()`) , Health Sciences from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Health Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Health Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Health Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Health Sciences") |> pull(value) |> sum()`) and Physical Sciences from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Physical Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Physical Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Physical Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Physical Sciences") |> pull(value) |> sum()`).
 <!-- In particular, journals in the subject areas recorded large high rate (see supplement)-->
@@
 The blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established.
-(A) Proportion of open access articles in hybrid journals per year and subject area (B) Percentage of hybrid open access via agreements per year and subject area Boxplots (C) show individual open access uptake rates by individual hybrid journals and open access funding per publication year and subject area The individual outliers are not shown.
+(A) Proportion of open access articles in hybrid journals per year and subject area.
+(B) Percentage of hybrid open access via agreements per year and subject area Boxplots.
+(C) show individual open access uptake rates by individual hybrid journals and open access funding per publication year and subject area.
+The individual outliers are not shown.
 Note that data on transformative agreements ending before June 2021 were not available for this study."}
@@
 To discern socio-economic differences, these countries were grouped according to their membership in the Organisation for Economic Co-operation and Development (OECD).
-During 2018 and 2022, first authors affiliated with institutions from OECD member countries published `r oecd_stats_global |> filter(cat == "OECD") |> pull(oa)` open access articles in hybrid journals, representing `r oecd_stats_global |> filter(cat == "OECD") |> pull(oa) / n_oa * 100`% of the investigated open access articles.
+Overall, first authors affiliated with institutions from OECD member countries published `r oecd_stats_global |> filter(cat == "OECD") |> pull(oa)` open access articles in hybrid journals, representing `r oecd_stats_global |> filter(cat == "OECD") |> pull(oa) / n_oa * 100`% of the investigated open access articles.
 This disparity between OECD nations and other countries becomes even more evident when considering open access through transformative agreements, as  `r oecd_stats_global |> filter(cat == "OECD") |> pull(ta_oa)` of `r n_ta_oa`, or `r oecd_stats_global |> filter(cat == "OECD") |> pull(ta_oa) / n_ta_oa*100`% of open access articles were associated with such agreements.
@@
 (B) Scatterplot contrasts total articles with open access article volume in 2022, by country and its OECD (purple-colored points) or BRICS (green) membership.
-Trend line obtained from linear regression, grey areas show pointwise symmetric 95% confidence bands.
+Trend line obtained from linear regression, grey area show pointwise symmetric 95% confidence bands.
 (C) Hybrid open access market share in 2022 by country.
@@
 Full counting was applied to account for multiple country affiliations [@Hottenrott_2021].
-Open access in hybrid journals increased in particular because of publications from authors affiliated with institutions from the OECD area, from `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2018) |> pull(value) |> sum()` in 2018 to `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2022) |> pull(value) |> sum()` in 2022.
+Open access in hybrid journals increased in particular because of publications from first authors affiliated with institutions from the OECD area, from `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2018) |> pull(value) |> sum()` in 2018 to `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2022) |> pull(value) |> sum()` in 2022.
 This growth was largely driven by transformative agreements. 
-The share of open access by agreement rose from `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2018, oa_type == "ta_oa") |> pull(value) / country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2018, oa_type == "ta_oa") |> pull(oa) * 100`% in 2018 (n = `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2018, oa_type == "ta_oa") |> pull(value)`) to `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2022, oa_type == "ta_oa") |> pull(value) / country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2022, oa_type == "ta_oa") |> pull(oa) * 100`% (n = `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2022, oa_type == "ta_oa") |> pull(value)`) in 2022.
+Their share rose from `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2018, oa_type == "ta_oa") |> pull(value) / country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2018, oa_type == "ta_oa") |> pull(oa) * 100`% in 2018 (n = `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2018, oa_type == "ta_oa") |> pull(value)`) to `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2022, oa_type == "ta_oa") |> pull(value) / country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2022, oa_type == "ta_oa") |> pull(oa) * 100`% (n = `r country_bar$data |> filter(cat == "OECD", oa_type != "closed", cr_year == 2022, oa_type == "ta_oa") |> pull(value)`) in 2022.
 In contrast, BRICS recorded a low uptake, moderately growing from `r country_bar$data |> filter(cat == "BRICS", oa_type != "closed", cr_year == 2018) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2018 to `r country_bar$data |> filter(cat == "BRICS", oa_type != "closed", cr_year == 2022) |> mutate(prop = value / all) |> pull(prop) |> sum() * 100`% in 2022.
@@
 In contrast, Western and Northern European countries published a considerably high number of open access articles. 
-Particularly, Scandinavian countries, the Netherlands and Austria recorded above-average open access shares, as indicated by the linear trend line.
+Particularly, Nordic countries, the Netherlands and Austria recorded above-average open access shares, as indicated by the linear trend line.
 As shown in Figure \ref{fig:country_patch}C, transformative agreements contributed to these market positions. 
@@
 Specifically, the three largest commercial publishers Elsevier, Springer Nature, and Wiley dominate, particularly with regard to open access provided through transformative agreements.
-Together, the three publishers accounted for three-fourths of open access articles through transformative agreements, while recording less than half of the total publication volume published between 2018 and 2022.
+Together, the three publishers accounted for three-fourths of open access articles through transformative agreements, while recording about half of active hybrid journals included in transformative agreements between 2018 and 2022.
 This observation aligns with previous research on shifts in the publishing market following the introduction of funding opportunities for hybrid open access [@Jahn_2016; @Butler_2023; @Shu_2023]. 
@@
 However, it must be noted that the focus of transformative agreements on publishers with large hybrid journal portfolios is intentional [@ifla].
-Because of transformative agreements with a few large publishers, national consortia were able to substantially increase their country’s annual open access article volume [@Bosman_2021; @Pinhasi_2021; @Huang_2020; @Taubert_2023; @Wenaas_2022].
+Because of transformative agreements with a few large publishers, national consortia were able to substantially increase their country’s annual open access article volume [@Jisc_2024; @Bosman_2021; @Pinhasi_2021; @Huang_2020; @Taubert_2023; @Wenaas_2022; @Widding_2024].
 
@@
 Because the BRICS expansion can be exclusively attributed to closed access articles, this development has the potential to hinder the transition of academic journal publishing to full open access through transformative agreements and demands discussion.
-In China, the country with the highest volume of articles in hybrid journals, only limited research funding to pay for open access is available, with expenditures for open access publishing surpassing subscription costs [@Shu_2023].
+In China, the country with the highest volume of articles in hybrid journals, only limited research funding to pay for open access is available, with expenditures for open access publishing surpassing subscription costs at some universities [@Shu_2023].
 Furthermore, the focus of Chinese authors on established journals may contribute to this trend [@Zhang_2022]. 
@@
 Because data on corresponding authors, who are typically responsible for facilitating open access publication, were not fully supported by OpenAlex at the time of this study, first author affiliations were used instead.
-This study is also unable to account for the various types of transformative agreements due to a lack of data, especially regarding article types and caps limiting the number of open access articles covered.
+This study is also unable to account for the various types of transformative agreements due to a lack of data, especially regarding article types and caps that limit the number of open access articles covered.
 Furthermore, assessing the quality of the OpenAlex and Crossref data used, particularly in terms of affiliations, corresponding authorship, and article types, by combining them with established bibliometric databases such as Scopus and Web of Science was beyond the scope of this analysis [@Visser_2021; @Chinchilla_Rodr_guez_2024].
-It must therefore be emphasised that, unlike related evaluations of national consortia based on invoice data [@Jisc_2024], this global overview can only provide an estimate of the support for open access through transformative agreements.
+It must therefore be emphasised that, unlike evaluations from national consortia that could make use of invoices [@Jisc_2024], this global overview can only provide an estimate of the support for open access through transformative agreements.
 Despite these limitations, the methodology was designed to underestimate, rather than overestimate, the adoption of open access through transformative agreements. 
@@
 This study has practical implications for research funding and libraries.
-Of primary concern should be the observed differences across countries, particularly the relationship between socio-economic development and open access adoption.
+One concern should be the observed differences across countries, particularly the relationship between socio-economic development and open access adoption.
 Between 2018 and 2022, there was a notable increase in the number of closed access articles published in hybrid journals by authors from countries such as China and India, where transformative agreements and open access funding options were not widely available.
@@
 This imbalance, whereby open access uptake depends largely on a few countries, makes it less likely that collective action towards a large-scale transition of hybrid journals to full open access through transformation agreements will succeed within the next years.
-Rather, and in accordance with recent policy recommendations from the UK and Nordic countries [@Jisc_2024; @Widding_2024], it emphasises the importance of ongoing evaluation and adaptation of open access strategies in view of global publishing trends.
+Rather, and in accordance with recent British and Scandinavian policy recommendations [@Jisc_2024; @Widding_2024], it emphasises the importance of ongoing evaluation and adaptation of open access strategies in view of global publishing trends.
 

### 2024-09-25 analysis/manuscript.Rmd
Minor language fixes

@@
 However, there are notable differences between the three large publishers.
-Although Elsevier published the largest volume of articles (n = `r publisher_league_tbl  |> filter(publisher == "Elsevier") |> pull(publisher_total)`, `r publisher_league_tbl  |> filter(publisher == "Elsevier") |> pull(cum)*100`%), it recorded a comparably low number of open access articles, including those that are associated with transformative agreements.
+Although Elsevier published the largest volume of articles (n = `r publisher_league_tbl  |> filter(publisher == "Elsevier") |> pull(publisher_total)`; `r publisher_league_tbl  |> filter(publisher == "Elsevier") |> pull(cum)*100`%), it recorded a comparably low number of open access articles, including those that are associated with transformative agreements.
 In contrast, Springer Nature and Wiley provided open access to a larger proportion of their articles (`r (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_total)) / (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(publisher_total)) * 100`% of Springer Nature articles and `r (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(oa_total)) / (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(publisher_total)) * 100`% of Wiley articles were open access), leading to higher open access market shares (`r publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_perc)*100`% Springer Nature resp.
@@
 (A) Proportion of open access articles in hybrid journals per year and subject area.
-(B) Percentage of hybrid open access via agreements per year and subject area Boxplots.
-(C) show individual open access uptake rates by individual hybrid journals and open access funding per publication year and subject area.
+(B) Percentage of hybrid open access via agreements per year and subject area.
+(C) Boxplots show individual open access uptake rates by individual hybrid journals and open access funding per publication year and subject area.
 The individual outliers are not shown.
@@
 
-Surprisingly, the total publication output of high-income countries belonging to the OECD in hybrid journals declined substantially after peaking in 2020, whereas that of BRICS countries doubled from 2018 to 2022.
+Surprisingly, the total publication output of high-income countries belonging to the OECD in hybrid journals declined substantially after peaking in 2020, whereas that of BRICS countries continuously grew from 2018 to 2022.
 Because the BRICS expansion can be exclusively attributed to closed access articles, this development has the potential to hinder the transition of academic journal publishing to full open access through transformative agreements and demands discussion.
@@
 Market observers also do not expect a broad implementation of transformative agreements in China in near future [@Owens_2024].
-As highlighted by @Koley_2023, India faces similar challenges in terms of the availablility of resources to pay for open access.
+As highlighted by @Koley_2023, India faces similar challenges in terms of the availability of resources to pay for open access.
 At the same time, access to research literature is a pressing issue, which is addressed by the "Indian one nation, one subscription" policy proposal. 
@@
 This imbalance, whereby open access uptake depends largely on a few countries, makes it less likely that collective action towards a large-scale transition of hybrid journals to full open access through transformation agreements will succeed within the next years.
-Rather, and in accordance with recent British and Scandinavian policy recommendations [@Jisc_2024; @Widding_2024], it emphasises the importance of ongoing evaluation and adaptation of open access strategies in view of global publishing trends.
+Rather, and in accordance with recent British and Scandinavian policy recommendations [@Jisc_2024; @Widding_2024], it emphasises the importance of ongoing evaluation and adaptation of open access strategies in view of global publishing trends and related challenges to equity and journal quality [@Ross_Hellauer_2022; @Goodin_2023].
 

### 2024-09-25 analysis/manuscript.tex
Minor language fixes

@@
 agreements have led to varying policy conclusions. For instance, the
-UK's Joint Information Systems Committee (JISC) evaluated its open
+British Joint Information Systems Committee (JISC) evaluated its open
 access strategy, which included transformative agreements (Brayman et
 al., 2024). The evaluation revealed that while these agreements had a
-significant impact on the growth of open access in the UK, they had
+significant impact on the growth of open access in the country, they had
 limited effects in facilitating a global shift towards full open access.
-As a result, the report advised that the UK's open access strategy
+As a result, the report advised that the British open access strategy
 should be reassessed. After making similar observation, Norwegian and
 Swedish universities and their consortia also argued for policy changes
-(Holden et al., 2023; Widding, 2024). Notably, the Association of
-Swedish Higher Education Institutions (Sveriges universitets- och
-högskoleförbund, SUHF) recommended only supporting agreements for
-publishing in fully open access journals by 2026 (Widding, 2024).
-Furthermore, cOAlition S will conclude its financial support of
-transformative agreements at the end of 2024, but will continue to view
-open access resulting from such agreements as compliant.
-\footnote{\url{https://www.coalition-s.org/coalition-s-confirms-the-end-of-its-financial-support-for-open-access-publishing-under-transformative-arrangements-after-2024/}}.
+(Holden et al., 2023; Widding, 2024). Furthermore, cOAlition S conclude
+its financial support of transformative agreements at the end of 2024,
+but continue to view open access resulting from such agreements as
+compliant. \footnote{\url{https://www.coalition-s.org/coalition-s-confirms-the-end-of-its-financial-support-for-open-access-publishing-under-transformative-arrangements-after-2024/}}.
 cOAlition S also removed the majority of hybrid journals from its
@@
 journals, do not directly attribute them to specific transformative
-agreements. Likewise, article-level open access invoicing and cost data
-are only partly available, which would make it possible to establish a
-direct link between transformative agreements and open access
-publications (Jahn et al., 2022; Kramer, 2024).
+agreements. Likewise, article-level open access invoicing and cost data,
+which would make it possible to establish a direct link between
+transformative agreements and open access publications (Jahn et al.,
+2022; Kramer, 2024).
 
@@
 invoicing data, open access articles published through transformative
-agreements were determined by first author affiliations from OpenAlex
-(Priem et al., 2022), linked to eligible institutions according to the
+agreements were determined by linking first author affiliations from
+OpenAlex (Priem et al., 2022) to eligible institutions according to the
 transformative agreement data. In the following, the steps are described
@@
 Agreement Data dump\footnote{\url{https://journalcheckertool.org/transformative-agreements/}}
-used by the cOAlition S Journal Checker Tool \footnote{\url{https://www.coalition-s.org/blog/enabling-accurate-results-within-the-journal-checker-tool/}},
+used by the cOAlition S Journal Checker Tool\footnote{\url{https://www.coalition-s.org/blog/enabling-accurate-results-within-the-journal-checker-tool/}},
 a voluntary effort based on publicly disclosed contracts in the ESAC
-Transformative Agreement Registry.\footnote{\url{https://esac-initiative.org/about/transformative-agreements/agreement-registry/}}.
+Transformative Agreement Registry\footnote{\url{https://esac-initiative.org/about/transformative-agreements/agreement-registry/}}.
 The dump consists of multiple online Google spreadsheets where each data
@@
 It should be noted that although many library consortia see the need to
-register their agreements through the ESAC registry, many fail to
-disclose all details, including contracts (Bakker et al., 2024; Kramer,
-2024). Furthermore, the extent to which ESAC is comprehensive remains
-uncertain, potentially limiting the coverage of transformative
-agreements in the Journal Checker Tool. Another limitation of using the
-Journal Checker Tool and its underlying publicly available data dump to
-study the development of transformative agreements over time is that
-expired transformative agreements are constantly removed. To address
-this, four different snapshots were safeguarded and combined for this
-study: self-archived versions from July 2021, July 2022, and May 2023,
-as well as the most current dump downloaded on 11 December 2023. This
-ensured that transformative agreements, which ended from 2021 onwards,
-were included, representing the majority of transformative agreements.
-Overall, the four combined Transformative Agreement Data dumps used in
-this study contained 729 of 869 agreements listed in the ESAC
-Transformative Agreement Registry by December 2023.
+register their agreements through the ESAC registry, some fail to
+disclose all details, including the full-text of contracts (Bakker et
+al., 2024; Kramer, 2024). Furthermore, the extent to which ESAC is
+comprehensive remains uncertain, potentially limiting the coverage of
+transformative agreements in the Journal Checker Tool. Another
+limitation of using the Journal Checker Tool and its underlying publicly
+available data dump to study the development of transformative
+agreements over time is that expired transformative agreements are
+constantly removed. To address this, four different snapshots were
+safeguarded and combined for this study: self-archived versions from
+July 2021, July 2022, and May 2023, as well as the most current dump
+downloaded on 11 December 2023. This ensured that transformative
+agreements, which ended from 2021 onwards, were included, representing
+the majority of transformative agreements. Overall, the four combined
+Transformative Agreement Data dumps used in this study contained 729 of
+869 agreements listed in the ESAC Transformative Agreement Registry by
+December 2023.
 
@@
 First authors typically contribute the most to a paper and are often
-considered lead authors (Larivière et al., 2016), although notable
-differences in corresponding author metadata were observed across
-subjects, publication years and bibliometric databases
-(Chinchilla-Rodríguez et al., 2024). In case of lacking data about
-corresponding authors in bibliometric databases, related studies also
-utilised first authors as a proxy to examine open access payments and
-transformative agreements (Haucap et al., 2021; Shu \& Larivière, 2023;
-Zhang et al., 2022).
+considered as lead authors (Chinchilla-Rodríguez et al., 2024; Larivière
+et al., 2016). In case of lacking data about corresponding authors in
+bibliometric databases, related studies also utilised first authors as a
+proxy to examine open access payments and transformative agreements
+(Haucap et al., 2021; Shu \& Larivière, 2023; Zhang et al., 2022).
 
@@
 by OpenAlex using the ROR-ID. Matching also considered the duration of
-agreements according to the ESAC registry. 502 agreements were active
-between 2018 and 2022. Upon inspection, Transformative Agreement Data
-did not comprehensively cover associated institutions, such as
+agreements according to the ESAC registry. In total, 502 agreements were
+active between 2018 and 2022. Upon inspection, Transformative Agreement
+Data did not comprehensively cover associated institutions, such as
 university hospitals or institutes of large research organisations such
@@
 
-Between 2018 and 2022, 11,189 out of 12,857 eligible hybrid journals in
+Between 2018 and 2022, 11,189 out of 12,857 hybrid journals in
 transformative agreements published at least one open access article
-under a Creative Commons license. These 11,189 hybrid journals
+under a Creative Commons license. These eligible 11,189 hybrid journals
 constituted the foundation for the subsequent analyses. Collectively,
@@
 active transformative agreements between 2018 and 2022 through the
-analysis of publication metadata, including affiliations and agreement
-data.
+linking of first author affiliations with agreement data.
 
@@
 an open access uptake of 29\%, and 6.6\% of journals (n = 744) provided
-the majority of their articles under a Creative Commons license in the
-same year. On average, these journals were smaller (M = 75, SD = 186)
-than those with an open access share below 50\% (M = 164, SD = 347).
+the majority of their articles under a CC license in the same year. On
+average, these journals were smaller (M = 75, SD = 186) than those with
+an open access share below 50\% (M = 164, SD = 347).
 
@@
 publishers. Although Elsevier published the largest volume of articles
-(n = 2,770,826, 34\%), it recorded a comparably low number of open
+(n = 2,770,826; 34\%), it recorded a comparably low number of open
 access articles, including those that are associated with transformative
@@
 articles and 15\% of Wiley articles were open access), leading to higher
-open access market share (23\% Springer Nature resp. 23\% Wiley). This
+open access market shares (23\% Springer Nature resp. 23\% Wiley). This
 difference between Elsevier on the one hand and Springer Nature and
@@
 agreements for its hybrid journal portfolio to selected consortia such
-as the Max Planck Society, the Swedish Bibsam Consortium and the Finnish
+as the Max Planck Society, the Swedish Bibsam consortium and the Finnish
 FinELib consortium in 2015 under the name Springer Compact\footnote{\url{https://web.archive.org/web/20180414062853id_/http://www.liber2015.org.uk/wp-content/uploads/2015/03/Springer-Compact.pdf}}.
@@
 importance of agreements for Springer Nature's hybrid open access
-business over the past five years (Figure \ref{fig:publisher_figure}).
-In 2022, 66\% (n = 34,725) of open access in Springer Nature hybrid
-journals was enabled through transformative agreements. In the same
-year, 70\% (n = 37,316) of Wiley's open access articles could be linked
-to transformative agreements. By contrast, Elsevier published a
-comparatively lower proportion of its open access articles through
-transformative agreements in 2022 (n = 32,627; 54\%).
+business over the five-years period (Figure
+\ref{fig:publisher_figure}B). In 2022, 66\% (n = 34,725) of open access
+in Springer Nature hybrid journals was enabled through transformative
+agreements. In the same year, 70\% (n = 37,316) of Wiley's open access
+articles could be linked to transformative agreements. By contrast,
+Elsevier published a comparatively lower proportion of its open access
+articles through transformative agreements in 2022 (n = 32,627; 54\%).
 
@@
 open access by subject area between 2018 and 2022. In particular, Social
-Sciences including Arts and Humanties journals accounted for the
+Sciences including Arts and Humanities journals accounted for the
 strongest growth in the five-years period from 6.4\% (n = 8,361) to 23\%
@@
 The blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established.
-(A) Proportion of open access articles in hybrid journals per year and subject area (B) Percentage of hybrid open access via agreements per year and subject area Boxplots (C) show individual open access uptake rates by individual hybrid journals and open access funding per publication year and subject area The individual outliers are not shown.
+(A) Proportion of open access articles in hybrid journals per year and subject area.
+(B) Percentage of hybrid open access via agreements per year and subject area.
+(C) Boxplots show individual open access uptake rates by individual hybrid journals and open access funding per publication year and subject area.
+The individual outliers are not shown.
 Note that data on transformative agreements ending before June 2021 were not available for this study.}\label{fig:subject_panel}
@@
 grouped according to their membership in the Organisation for Economic
-Co-operation and Development (OECD). During 2018 and 2022, first authors
-affiliated with institutions from OECD member countries published
-602,050 open access articles in hybrid journals, representing 81\% of
-the investigated open access articles. This disparity between OECD
-nations and other countries becomes even more evident when considering
-open access through transformative agreements, as 310,712 of 328,957, or
-94\% of open access articles were associated with such agreements.
+Co-operation and Development (OECD). Overall, first authors affiliated
+with institutions from OECD member countries published 602,050 open
+access articles in hybrid journals, representing 81\% of the
+investigated open access articles. This disparity between OECD nations
+and other countries becomes even more evident when considering open
+access through transformative agreements, as 310,712 of 328,957, or 94\%
+of open access articles were associated with such agreements.
 
@@
 (B) Scatterplot contrasts total articles with open access article volume in 2022, by country and its OECD (purple-colored points) or BRICS (green) membership.
-Trend line obtained from linear regression, grey areas show pointwise symmetric 95\% confidence bands.
+Trend line obtained from linear regression, grey area show pointwise symmetric 95\% confidence bands.
 (C) Hybrid open access market share in 2022 by country.
@@
 in hybrid journals increased in particular because of publications from
-authors affiliated with institutions from the OECD area, from 52,154 in
-2018 to 202,787 in 2022. This growth was largely driven by
-transformative agreements. The share of open access by agreement rose
-from 5.5\% in 2018 (n = 2,859) to 68\% (n = 137,815) in 2022. In
-contrast, BRICS recorded a low uptake, moderately growing from 1.6\% in
-2018 to 3.7\% in 2022.
+first authors affiliated with institutions from the OECD area, from
+52,154 in 2018 to 202,787 in 2022. This growth was largely driven by
+transformative agreements. Their share rose from 5.5\% in 2018 (n =
+2,859) to 68\% (n = 137,815) in 2022. In contrast, BRICS recorded a low
+uptake, moderately growing from 1.6\% in 2018 to 3.7\% in 2022.
 
@@
 contrast, Western and Northern European countries published a
-considerably high number of open access articles. Particularly,
-Scandinavian countries, the Netherlands and Austria recorded
-above-average open access shares, as indicated by the linear trend line.
-As shown in Figure \ref{fig:country_patch}C, transformative agreements
-contributed to these market positions. Interestingly, the United States
-had a notable open access market share of 15\%, although transformative
-agreements contributed to a lesser extent. Similarly, China's open
-access market share of 7.2\% in 2022 was comparable to that of the
-Netherlands, which was 7.1\%.
+considerably high number of open access articles. Particularly, Nordic
+countries, the Netherlands and Austria recorded above-average open
+access shares, as indicated by the linear trend line. As shown in Figure
+\ref{fig:country_patch}C, transformative agreements contributed to these
+market positions. Interestingly, the United States had a notable open
+access market share of 15\%, although transformative agreements
+contributed to a lesser extent. Similarly, China's open access market
+share of 7.2\% in 2022 was comparable to that of the Netherlands, which
+was 7.1\%.
 
@@
 of open access articles through transformative agreements, while
-recording less than half of the total publication volume published
-between 2018 and 2022. This observation aligns with previous research on
-shifts in the publishing market following the introduction of funding
-opportunities for hybrid open access (Butler et al., 2023; Jahn \&
-Tullney, 2016; Shu \& Larivière, 2023). Additionally, the results
-confirm previously observed variations by publisher, with Elsevier
-exhibiting a different development than Springer Nature and Wiley
-(Butler et al., 2023; Jahn et al., 2022). However, it must be noted that
-the focus of transformative agreements on publishers with large hybrid
-journal portfolios is intentional (Campbell et al., 2022). Because of
-transformative agreements with a few large publishers, national
-consortia were able to substantially increase their country's annual
-open access article volume (Bosman, Jonge, et al., 2021; Huang et al.,
-2020; Pinhasi et al., 2021; Taubert et al., 2023; Wenaas, 2022).
+recording about half of active hybrid journals included in
+transformative agreements between 2018 and 2022. This observation aligns
+with previous research on shifts in the publishing market following the
+introduction of funding opportunities for hybrid open access (Butler et
+al., 2023; Jahn \& Tullney, 2016; Shu \& Larivière, 2023). Additionally,
+the results confirm previously observed variations by publisher, with
+Elsevier exhibiting a different development than Springer Nature and
+Wiley (Butler et al., 2023; Jahn et al., 2022). However, it must be
+noted that the focus of transformative agreements on publishers with
+large hybrid journal portfolios is intentional (Campbell et al., 2022).
+Because of transformative agreements with a few large publishers,
+national consortia were able to substantially increase their country's
+annual open access article volume (Bosman, Jonge, et al., 2021; Brayman
+et al., 2024; Huang et al., 2020; Pinhasi et al., 2021; Taubert et al.,
+2023; Wenaas, 2022; Widding, 2024).
 
@@
 belonging to the OECD in hybrid journals declined substantially after
-peaking in 2020, whereas that of BRICS countries doubled from 2018 to
-2022. Because the BRICS expansion can be exclusively attributed to
-closed access articles, this development has the potential to hinder the
-transition of academic journal publishing to full open access through
-transformative agreements and demands discussion. In China, the country
-with the highest volume of articles in hybrid journals, only limited
-research funding to pay for open access is available, with expenditures
-for open access publishing surpassing subscription costs (Shu \&
-Larivière, 2023). Furthermore, the focus of Chinese authors on
-established journals may contribute to this trend (Zhang et al., 2022).
-Although China supports Plan S (Schiermeier, 2018), this is not
-reflected in the data in terms of open access uptake in hybrid journals.
-Market observers also do not expect a broad implementation of
-transformative agreements in China in near future (Owens, 2024). As
-highlighted by Koley \& Lala (2023), India faces similar challenges in
-terms of the availablility of resources to pay for open access. At the
-same time, access to research literature is a pressing issue, which is
-addressed by the ``Indian one nation, one subscription'' policy
-proposal. However, this policy focuses on centrally negotiated
-subscriptions and does not entail open access.
+peaking in 2020, whereas that of BRICS countries continuously grew from
+2018 to 2022. Because the BRICS expansion can be exclusively attributed
+to closed access articles, this development has the potential to hinder
+the transition of academic journal publishing to full open access
+through transformative agreements and demands discussion. In China, the
+country with the highest volume of articles in hybrid journals, only
+limited research funding to pay for open access is available, with
+expenditures for open access publishing surpassing subscription costs at
+some universities (Shu \& Larivière, 2023). Furthermore, the focus of
+Chinese authors on established journals may contribute to this trend
+(Zhang et al., 2022). Although China supports Plan S (Schiermeier,
+2018), this is not reflected in the data in terms of open access uptake
+in hybrid journals. Market observers also do not expect a broad
+implementation of transformative agreements in China in near future
+(Owens, 2024). As highlighted by Koley \& Lala (2023), India faces
+similar challenges in terms of the availability of resources to pay for
+open access. At the same time, access to research literature is a
+pressing issue, which is addressed by the ``Indian one nation, one
+subscription'' policy proposal. However, this policy focuses on
+centrally negotiated subscriptions and does not entail open access.
 
@@
 transformative agreements due to a lack of data, especially regarding
-article types and caps limiting the number of open access articles
+article types and caps that limit the number of open access articles
 covered. Furthermore, assessing the quality of the OpenAlex and Crossref
@@
 scope of this analysis (Chinchilla-Rodríguez et al., 2024; Visser et
-al., 2021). It must therefore be emphasised that, unlike related
-evaluations of national consortia based on invoice data (Brayman et al.,
+al., 2021). It must therefore be emphasised that, unlike evaluations
+from national consortia that could make use of invoices (Brayman et al.,
 2024), this global overview can only provide an estimate of the support
@@
 This study has practical implications for research funding and
-libraries. Of primary concern should be the observed differences across
+libraries. One concern should be the observed differences across
 countries, particularly the relationship between socio-economic
@@
 access through transformation agreements will succeed within the next
-years. Rather, and in accordance with recent policy recommendations from
-the UK and Nordic countries (Brayman et al., 2024; Widding, 2024), it
+years. Rather, and in accordance with recent British and Scandinavian
+policy recommendations (Brayman et al., 2024; Widding, 2024), it
 emphasises the importance of ongoing evaluation and adaptation of open
-access strategies in view of global publishing trends.
+access strategies in view of global publishing trends and related
+challenges to equity and journal quality (Rasmussen, 2023; Ross-Hellauer
+et al., 2022).
 
@@
 Schmal, W. B. (2024). How transformative are transformative agreements?
-Evidence from germany across disciplines. \emph{Scientometrics}.
+Evidence from {Germany} across disciplines. \emph{Scientometrics},
+\emph{129}(3), 1863--1889.
 \url{https://doi.org/10.1007/s11192-024-04955-y}

### 2024-09-25 analysis/manuscript.Rmd
Add figure data

@@
   mutate(prop = n / sum(n)) |>
+  ungroup() |>
   ggplot(aes(gsub("^20", "'", as.character(cr_year)), n, fill = cat)) +
@@
 (p | cr_stats_plot) + plot_annotation(tag_levels = 'A') +  plot_layout(widths = c(0.6, 0.4))
+
+# Data backup
+
+my_venn_df <- stack(my_venn)
+write_csv(my_venn_df, here::here("analysis/fig_data", "fig_2A.csv"))
+write_csv(cr_stats_plot$data, here::here("analysis/fig_data", "fig_2B.csv"))
 ```
@@
 oa_jn_box <- oa_jn_box_df |>
+  ungroup() |>
   ggplot(aes(gsub("^20", "'", as.character(cr_year)), prop)) +
@@
     plot_annotation(tag_levels = 'A')
+
+# Data export
+write_csv(oa_ta_plot$data, here::here("analysis/fig_data", "fig_3A.csv"))
+write_csv(ta_share_plot$data, here::here("analysis/fig_data", "fig_3B.csv"))
+write_csv(oa_jn_box$data, here::here("analysis/fig_data", "fig_3C.csv"))
+write_csv(box_ta_prop$data, here::here("analysis/fig_data", "fig_3D.csv"))
 ```
@@
 publisher_summary_df <- inner_join(oa_publisher, all_publisher) |>
-  mutate(prop = value / jn_all)
+  mutate(prop = value / jn_all) |>
+  ungroup()
 
@@
 publisher_ta_prop_plot <- publisher_ta_prop |>
+  ungroup() |>
   ggplot(aes(gsub("^20", "'", as.character(cr_year)), prop)) +
@@
     plot_annotation(tag_levels = 'A')
+
+# Data backup
+write_csv(publisher_prop$data, here::here("analysis/fig_data", "fig_4A.csv"))
+write_csv(publisher_ta_prop_plot$data, here::here("analysis/fig_data", "fig_4B.csv"))
+write_csv(publisher_boxplot$data, here::here("analysis/fig_data", "fig_4C.csv"))
 ```
@@
   mutate(prop = value / articles_frac) |>
+  ungroup() |>
    ggplot(aes(
@@
   mutate(prop = ta_oa_frac / oa_frac) |>
+  ungroup() |>
   ggplot(aes(gsub("^20", "'", as.character(cr_year)), prop)) +
@@
   filter(top_level != "Multidisciplinary", !cr_year %in% c("2017", "2023")) |>
+  ungroup() |>
   ggplot(aes(gsub("^20", "'", as.character(cr_year)), value / jn_all, fill = type)) +
@@
     plot_annotation(tag_levels = 'A')
+
+# Data backup
+
+write_csv(subject_trend_plot$data, here::here("analysis/fig_data", "fig_5A.csv"))
+write_csv(subject_ta_trend_plot$data, here::here("analysis/fig_data", "fig_5B.csv"))
+write_csv(subject_box_trend$data, here::here("analysis/fig_data", "fig_5C.csv"))
 ```
@@
   values_to = "value",
-  names_to = "type") 
+  names_to = "type") |>
+  ungroup()
 ```
@@
             ta_oa = sum(ta_oa), .groups = "drop") |>
-  head(20) |>
-  mutate(cat_labels = case_when(
-   cat == "OECD" ~ glue::glue("<span style = 'color: #3fc1c9'>{country_code}</span>"),
-   cat == "BRICS" ~ glue::glue("<span style = 'color: #fc5185'>{country_code}</span>"),
-   cat == "Other" ~ glue::glue("<span style = 'color: #e1e1e1'>{country_code}</span>")
-  )) |>
-  mutate(cat_labels = fct_inorder(cat_labels))
+  head(20)
 
@@
   names_to = "type") |>
+  ungroup() |>
   ggplot(aes(reorder(as.character(country_code),  value), value / sum(value), fill = type)) +
@@
   filter(cr_year == "2022", !is.na(country_code), all > 5000) |>
+  ungroup() |>
   ggplot(aes(all, oa / all)) +
@@
   mutate(cat = ifelse(is.na(cat), "Other", cat)) |>
-  left_join(oecd_by_year, by = "cr_year")
+  left_join(oecd_by_year, by = "cr_year") |>
+  ungroup()
 
@@
     plot_annotation(tag_levels = 'A')
+
+# Data backup
+write_csv(country_bar$data, here::here("analysis/fig_data", "fig_6A.csv"))
+write_csv(country_scatter_22$data, here::here("analysis/fig_data", "fig_6B.csv"))
+write_csv(top_20_plot$data, here::here("analysis/fig_data", "fig_6C.csv"))
 ```
@@
 country_plot
+
+# Data backup 
+write_csv(country_plot$data, here::here("analysis/fig_data", "fig_7.csv"))
 ```

### 2024-09-26 analysis/manuscript.Rmd
Submitted revision

@@
 After making similar observation, Norwegian and Swedish universities and their consortia also argued for policy changes [@norway; @Widding_2024].
-Furthermore, cOAlition S conclude its financial support of transformative agreements at the end of 2024, but continue to view open access resulting from such agreements as compliant. [^coalitions].
+Furthermore, cOAlition S conclude its financial support of transformative agreements at the end of 2024, but continue to view open access resulting from such agreements as compliant.[^coalitions]
 cOAlition S also removed the majority of hybrid journals from its Transformative Journal program in 2023 due to publishers' failure to meet self-defined open access growth targets [@Brainard_2023]. 

### 2024-09-26 analysis/manuscript.tex
Submitted revision

@@
 but continue to view open access resulting from such agreements as
-compliant. \footnote{\url{https://www.coalition-s.org/coalition-s-confirms-the-end-of-its-financial-support-for-open-access-publishing-under-transformative-arrangements-after-2024/}}.
+compliant.\footnote{\url{https://www.coalition-s.org/coalition-s-confirms-the-end-of-its-financial-support-for-open-access-publishing-under-transformative-arrangements-after-2024/}}
 cOAlition S also removed the majority of hybrid journals from its
